### 2017-09-25 doc/aboutme.Rmd
Adding some documentation for perspective on runs

---
title: "Adapting toolmark based chumbley score to bullets"
author: "Ganesh, Heike"
date: "September 25, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Toolmarks:

Used an optimization window size of n = 500 and a validation window size of m = 50. It is visually clear that when the two tool marks match, the U-statistic tends to be relatively large, and when they do not match, the value tends to be relatively small.

The p-Values associated with each U-statistic, are calculated under the assumption that the nonmatch distribution is standard normal. As should be the case, the p-values for nonmatching pairs are approximately uniformly distributed between zero and one, while the p-values for matching pairs are concentrated on small values (suggesting that they would be declared to be “matching”).

The classification rates corresponding to the p-values where the rows correspond to the known status of the tool marks and columns represent the model conclusion. Of 50 pairs of known matching marks, we misclassified three pairs as nonmatches for false-negative error rate of 6% and 0 pairs of known nonmatching tool marks were identified as being matches
```{r cars, echo=FALSE}
a<- matrix(NA, ncol = 3, nrow = 2)
toolmark.falsepositive<- data.frame(a)
colnames(toolmark.falsepositive)<- c("Classification", "Match", "Non-Match")
toolmark.falsepositive$Classification<- c("Match", "Non-Match")
toolmark.falsepositive$Match<- c(47,0)
toolmark.falsepositive$`Non-Match`<- c(3,50)
toolmark.falsepositive

```

Following on similar lines, the first step here is to first identify what difference does different window sizes of optimization and the validation step have, when adapting the toolmark method to bullets.

The marking made on bullets are smaller than toolmarks and is also less wider. The idea is to find out possible areas of error while adapting the score based method proposed for toolmarks.

Bullet signatures being compared at this time are from the Hamby 44 and Hamby 252 data, which have the best set of known-matched and known non-matches.

Bullet signatures are chosen by
1. Filtering out Land_id for Profiles from the Hamby 44 and Hamby 252 data and removing 
all NA values
2. run_id = 3 is chosen, Seems like the signatures generated from this run_id give the closest
match. Different run_id's have some different settings for generating the signatures
(The level of smoothing does not seem to be one of them)
3. The bullet signatures when generated already included the LOESS smoothing
## Including Plots

For a significance level of 0.05

```{r echo=FALSE}
#library(tidyverse)
#install.packages("tidyverse")
falspostive<- function(sig_alpha){
  library(dplyr)
  if(is.null(all)){
    files <- dir("././data", pattern="csv")
    all <- data.frame()
    for (file in files) {
      tmp <- read.csv(file=file.path(path="././data", file)) 
      if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% select(-x)
      all <- rbind(all, tmp)
    }
  }
    summ <- all %>% filter(!is.na(p_value)) %>%
      mutate(signif = p_value < sig_alpha) %>%
      group_by(wv, wo, match, signif) %>% tally()
    summ$error <- with(summ, match != signif)
    rates <- summ %>% group_by(wv, wo, match) %>% summarize(
      rate = n[error==TRUE]/sum(n)
    )
    combo<- list(summ, rates)
    names(combo)<- c("summ", "rates")
    return(combo)
}

a<- falspostive(0.05)

a$summ %>% ggplot(aes( x  = factor(wv), weight = n, fill=error)) + geom_bar(position="fill") +
      facet_grid(match~wo, labeller="label_both")
a$rates %>% ggplot(aes(x = wv, y = rate)) + geom_point() + facet_grid(match~wo, labeller="label_both") +  geom_line()
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2017-09-25 doc/eval_windows.R

@@ -1,0 +1,56 @@
+
+falspostive<- function(sig_alpha){
+  if(is.null(all)){
+    library(tidyverse)
+    files <- dir("././data", pattern="csv")
+    all <- data.frame()
+    for (file in files) {
+      tmp <- read.csv(file=file.path(path="././data", file)) 
+      if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% select(-x)
+      all <- rbind(all, tmp)
+    }
+  }
+    summ <- all %>% filter(!is.na(p_value)) %>%
+      mutate(signif = p_value < sig_alpha) %>%
+      group_by(wv, wo, match, signif) %>% tally()
+    summ$error <- with(summ, match != signif)
+    rates <- summ %>% group_by(wv, wo, match) %>% summarize(
+      rate = n[error==TRUE]/sum(n)
+    )
+    combo<- list(summ, rates)
+    names(combo)<- c("summ", "rates")
+    return(combo)
+}
+
+    summ %>% ggplot(aes( x  = factor(wv), weight = n, fill=error)) + geom_bar(position="fill") +
+      facet_grid(match~wo, labeller="label_both")
+
+    rates %>% ggplot(aes(x = wv, y = rate)) + geom_point() +
+      facet_grid(match~wo, labeller="label_both") +
+      geom_line()
+
+    # library(tidyverse)
+    # files <- dir("./data", pattern="csv")
+    # all <- data.frame()
+    # for (file in files) {
+    #   tmp <- read.csv(file=file.path(path="./data", file)) 
+    #   if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% select(-x)
+    #   all <- rbind(all, tmp)
+    # }
+    # 
+    # summ <- all %>% filter(!is.na(p_value)) %>%
+    #   mutate(signif = p_value < 0.01) %>%
+    #   group_by(wv, wo, match, signif) %>% tally()
+    # summ$error <- with(summ, match != signif)
+    # summ %>% ggplot(aes( x  = factor(wv), weight = n, fill=error)) + geom_bar(position="fill") +
+    #   facet_grid(match~wo, labeller="label_both")
+    # 
+    # rates <- summ %>% group_by(wv, wo, match) %>% summarize(
+    #   rate = n[error==TRUE]/sum(n)
+    # )
+    # rates %>% ggplot(aes(x = wv, y = rate)) + geom_point() +
+    #   facet_grid(match~wo, labeller="label_both") + 
+    #   geom_line()
+    
+
+    

### 2017-09-25 doc/aboutme.Rmd
update

@@
 The classification rates corresponding to the p-values where the rows correspond to the known status of the tool marks and columns represent the model conclusion.
-Of 50 pairs of known matching marks, we misclassified three pairs as nonmatches for false-negative error rate of 6% and 0 pairs of known nonmatching tool marks were identified as being matches
+Of 50 pairs of known matching marks, we misclassified three pairs as nonmatches for false-negative error rate of 6% and 0 pairs of known nonmatching tool marks were identified as being matches.
+
+Null Hypothesis: The toolmarks were not made by the same tool.
+Alternate: Toolmarks made by the same tool.
 ```{r cars, echo=FALSE}
@@
 (The level of smoothing does not seem to be one of them)
-3. The bullet signatures when generated already included the LOESS smoothing
+
+The bullet signatures when generated already included the LOWESS smoothing.
+Therefore, the coarseness factor is set to 1 while running the chumbley_non_random()
+which generates the same_shift, different_shift, U-Stat and P_value parameters.
+
 ## Including Plots
@@
 ```{r echo=FALSE}
-#library(tidyverse)
-#install.packages("tidyverse")
-falspostive<- function(sig_alpha){
-  library(dplyr)
-  if(is.null(all)){
-    files <- dir("././data", pattern="csv")
-    all <- data.frame()
-    for (file in files) {
-      tmp <- read.csv(file=file.path(path="././data", file)) 
-      if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% select(-x)
-      all <- rbind(all, tmp)
-    }
-  }
-    summ <- all %>% filter(!is.na(p_value)) %>%
-      mutate(signif = p_value < sig_alpha) %>%
-      group_by(wv, wo, match, signif) %>% tally()
-    summ$error <- with(summ, match != signif)
-    rates <- summ %>% group_by(wv, wo, match) %>% summarize(
-      rate = n[error==TRUE]/sum(n)
-    )
-    combo<- list(summ, rates)
-    names(combo)<- c("summ", "rates")
-    return(combo)
-}
 
-a<- falspostive(0.05)
-
-a$summ %>% ggplot(aes( x  = factor(wv), weight = n, fill=error)) + geom_bar(position="fill") +
-      facet_grid(match~wo, labeller="label_both")
-a$rates %>% ggplot(aes(x = wv, y = rate)) + geom_point() + facet_grid(match~wo, labeller="label_both") +  geom_line()
 ```

### 2017-09-25 doc/eval_windows.R

@@ -22,12 +22,19 @@ falspostive<- function(sig_alpha){
     return(combo)
 }
 
-    summ %>% ggplot(aes( x  = factor(wv), weight = n, fill=error)) + geom_bar(position="fill") +
-      facet_grid(match~wo, labeller="label_both")
+a<- falspostive(0.05)
 
-    rates %>% ggplot(aes(x = wv, y = rate)) + geom_point() +
-      facet_grid(match~wo, labeller="label_both") +
-      geom_line()
+a$summ %>% ggplot(aes( x  = factor(wv), weight = n, fill=error)) + geom_bar(position="fill") +
+  facet_grid(match~wo, labeller="label_both")
+a$rates %>% ggplot(aes(x = wv, y = rate)) + geom_point() + facet_grid(match~wo, labeller="label_both") +
+  geom_line()
+    
+# summ %>% ggplot(aes( x  = factor(wv), weight = n, fill=error)) + geom_bar(position="fill") +
+#       facet_grid(match~wo, labeller="label_both")
+# 
+#     rates %>% ggplot(aes(x = wv, y = rate)) + geom_point() +
+#       facet_grid(match~wo, labeller="label_both") +
+#       geom_line()
 
     # library(tidyverse)
     # files <- dir("./data", pattern="csv")

### 2017-09-25 doc/aboutme.Rmd
checking some things

@@
 title: "Adapting toolmark based chumbley score to bullets"
-author: "Ganesh, Heike"
+author: "Ganesh Krishnan, Heike Hofmann"
 date: "September 25, 2017"
@@
 
-```{r echo=FALSE}
+![](../doc/plots/error_pt05.png)
+For a significance level of 0.01
 
-```
+![](../doc/plots/error_pt01.png)
+
+For a significance level of 0.001
+
+![](../doc/plots/error_pt001.png)
 

### 2017-09-25 doc/aboutme.Rmd
.

@@
 
-## R Markdown
-
-Toolmarks:
+## Toolmarks
 
@@
 ```
+## Bullets
 
@@
 
+WO 320 wv 50
+```{r echo=FALSE}
+a<- matrix(NA, ncol = 3, nrow = 2)
+bullet.falsepositive<- data.frame(a)
+colnames(bullet.falsepositive)<- c("Classification", "FALSe", "TRUe")
+bullet.falsepositive$Classification<- c("FALSe", "TRUe")
+bullet.falsepositive$FALSe<- c(77564,412)
+bullet.falsepositive$TRUe<- c(4022,674)
+bullet.falsepositive
+
+```
+Wo 200 WV 50
+```{r echo=FALSE}
+bullet.falsepositive$FALSe<- c(78817,383)
+bullet.falsepositive$TRUe<- c(4004,798)
+bullet.falsepositive
+
+```
+Wo 80 wv 50
+```{r echo=FALSE}
+bullet.falsepositive$FALSe<- c(78817,450)
+bullet.falsepositive$TRUe<- c(4680,775)
+bullet.falsepositive
+
+```
+
 ## Including Plots
@@
 
+![](../doc/plots/rate_pt001.png)
+
 Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2017-09-25 doc/eval_windows.R

@@ -59,5 +59,14 @@ a$rates %>% ggplot(aes(x = wv, y = rate)) + geom_point() + facet_grid(match~wo, 
     #   facet_grid(match~wo, labeller="label_both") + 
     #   geom_line()
     
+aa1<- read.csv2("./data/chumbley-out-wo-80-wv-50.csv",header = TRUE, sep = ",")
+xtabs(~signif+match, data =aa1)
 
+
+aa2<- read.csv2("./data/chumbley_wo-200_wv-50.csv",header = TRUE, sep = ",")
+xtabs(~signif+match, data =aa2)
+
+aa3<- read.csv2("./data/chumbley_wo-320_wv-50.csv",header = TRUE, sep = ",")
+xtabs(~signif+match, data =aa3)
     
+

### 2017-09-26 code/evaluate.R
mini changes

@@ -8,7 +8,7 @@ for (file in files) {
 }
 
 summ <- all %>% filter(!is.na(p_value)) %>%
-  mutate(signif = p_value < 0.001) %>%
+  mutate(signif = p_value < 0.05) %>%
   group_by(wv, wo, match, signif) %>% tally()
 summ$error <- with(summ, match != signif)
 summ %>% ggplot(aes( x  = factor(wv), weight = n, fill=error)) + geom_bar(position="fill") +

### 2017-09-26 code/evaluate.R
couple of summary plots

@@ -7,16 +7,63 @@ for (file in files) {
   all <- rbind(all, tmp)
 }
 
-summ <- all %>% filter(!is.na(p_value)) %>%
-  mutate(signif = p_value < 0.05) %>%
+
+errorrate <- function(data, alpha) {
+summ <- data %>% filter(!is.na(p_value)) %>%
+  mutate(signif = p_value < alpha) %>%
   group_by(wv, wo, match, signif) %>% tally()
 summ$error <- with(summ, match != signif)
-summ %>% ggplot(aes( x  = factor(wv), weight = n, fill=error)) + geom_bar(position="fill") +
-  facet_grid(match~wo, labeller="label_both")
+summ$alpha = alpha
+#summ %>% ggplot(aes( x  = factor(wv), weight = n, fill=error)) + geom_bar(position="fill") +
+#  facet_grid(match~wo, labeller="label_both")
 
-rates <- summ %>% group_by(wv, wo, match) %>% summarize(
+rates <- summ %>% group_by(wv, wo, match, alpha) %>% summarize(
   rate = n[error==TRUE]/sum(n)
 )
-rates %>% ggplot(aes(x = wv, y = rate)) + geom_point() +
-  facet_grid(match~wo, labeller="label_both") + 
-  geom_line()
+rates %>% spread(match,rate) %>% rename(
+  actual = `FALSE`,
+  beta = `TRUE`
+)
+}
+
+errors <- rbind(errorrate(all, 0.001), 
+                errorrate(all, 0.005), 
+                errorrate(all, 0.01), 
+                errorrate(all, 0.05))
+errors %>% ggplot(aes(x = wv, y = beta, colour=factor(alpha))) + geom_point() +
+  facet_grid(.~wo, labeller="label_both") + 
+  geom_line(aes(group=alpha))
+
+errors %>% ggplot(aes(x = wv, y = actual, colour=factor(alpha))) + geom_point() +
+  facet_grid(.~wo, labeller="label_both") + 
+  geom_line(aes(group=alpha))
+
+
+errors %>% ggplot(aes(x = wo, y = beta, colour=factor(alpha))) + geom_point() +
+#  facet_grid(.~wo, labeller="label_both") + 
+  geom_smooth(aes(group=alpha), se=FALSE, method="loess") +
+  theme_bw() +
+  scale_colour_brewer("Nominal type I error", palette="Greens") +
+  xlab("Window size for optimization") +
+  ylab("Type II error rate")
[…]

### 2017-09-29 code/evaluate.R
change to pics

@@ -38,12 +38,12 @@ errors %>% ggplot(aes(x = wv, y = actual, colour=factor(alpha))) + geom_point() 
   facet_grid(.~wo, labeller="label_both") + 
   geom_line(aes(group=alpha))
 
-
+greens <- RColorBrewer::brewer.pal(name="Greens", n = 6)
 errors %>% ggplot(aes(x = wo, y = beta, colour=factor(alpha))) + geom_point() +
 #  facet_grid(.~wo, labeller="label_both") + 
   geom_smooth(aes(group=alpha), se=FALSE, method="loess") +
   theme_bw() +
-  scale_colour_brewer("Nominal type I error", palette="Greens") +
+  scale_colour_manual("Nominal type I error", values=greens[-(1:2)]) +
   xlab("Window size for optimization") +
   ylab("Type II error rate")
 

### 2017-09-29 code/evaluate.R
change to pics

@@ -30,25 +30,29 @@ errors <- rbind(errorrate(all, 0.001),
                 errorrate(all, 0.005), 
                 errorrate(all, 0.01), 
                 errorrate(all, 0.05))
-errors %>% ggplot(aes(x = wv, y = beta, colour=factor(alpha))) + geom_point() +
+errors %>% filter(wv %in% c(30,50)) %>% ggplot(aes(x = wv, y = beta, colour=factor(alpha))) + geom_point() +
   facet_grid(.~wo, labeller="label_both") + 
   geom_line(aes(group=alpha))
 
-errors %>% ggplot(aes(x = wv, y = actual, colour=factor(alpha))) + geom_point() +
+errors %>% filter(wv %in% c(30,50)) %>% ggplot(aes(x = wv, y = actual, colour=factor(alpha))) + geom_point() +
   facet_grid(.~wo, labeller="label_both") + 
   geom_line(aes(group=alpha))
 
 greens <- RColorBrewer::brewer.pal(name="Greens", n = 6)
-errors %>% ggplot(aes(x = wo, y = beta, colour=factor(alpha))) + geom_point() +
+errors %>% filter(wv %in% c(30,50)) %>%
+  ggplot(aes(x = wo, y = beta, colour=factor(alpha))) + 
+  geom_point(aes(shape=factor(wv))) +
 #  facet_grid(.~wo, labeller="label_both") + 
   geom_smooth(aes(group=alpha), se=FALSE, method="loess") +
   theme_bw() +
   scale_colour_manual("Nominal type I error", values=greens[-(1:2)]) +
+  scale_shape_discrete("Size of \nvalidation window") +
   xlab("Window size for optimization") +
   ylab("Type II error rate")
 
 # Relative error for type I
-errors %>% ggplot(aes(x = wo, y = actual/alpha, colour=factor(alpha))) +
+errors %>% filter(wv %in% c(30,50)) %>%
+  ggplot(aes(x = wo, y = actual/alpha, colour=factor(alpha))) +
   geom_hline(yintercept = 1, colour="grey30", size=0.5) +
   geom_point() +
   #  facet_grid(.~wo, labeller="label_both") + 
@@ -59,11 +63,14 @@ errors %>% ggplot(aes(x = wo, y = actual/alpha, colour=factor(alpha))) +
   xlab("Window size for optimization") +
   scale_colour_brewer("Nominal type I error", palette="Set2")
 
[…]

### 2017-09-29 code/evaluate.R
another data combo

@@ -74,3 +74,15 @@ errors %>% filter(wv %in% c(30,50)) %>%
   ylab("Observed type I error rate") +
   xlab("Window size for optimization") +
   scale_colour_brewer("Nominal type I error", palette="Set2")
+
+# focus on wo = 120
+
+errors %>% filter(wo == 120) %>% ggplot(aes(x = wv, y = actual, colour=factor(alpha))) + geom_point() +
+  facet_grid(.~wo, labeller="label_both") + 
+  geom_line(aes(group=alpha))
+# type I error rate increases 
+
+errors %>% filter(wo == 120) %>% ggplot(aes(x = wv, y = beta, colour=factor(alpha))) + geom_point() +
+  facet_grid(.~wo, labeller="label_both") + 
+  geom_line(aes(group=alpha))
+

### 2017-09-29 code/evaluate.R
adjust paths to new structure

@@ -1,8 +1,8 @@
 library(tidyverse)
-files <- dir("../data", pattern="csv")
+files <- dir("../data/signatures", pattern="csv")
 all <- data.frame()
 for (file in files) {
-  tmp <- read.csv(file=file.path(path="../data", file)) 
+  tmp <- read.csv(file=file.path(path="../data/signatures", file)) 
   if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% select(-x)
   all <- rbind(all, tmp)
 }

### 2017-09-29 doc/eval_windows.R

@@ -2,10 +2,10 @@
 falspostive<- function(sig_alpha){
   if(is.null(all)){
     library(tidyverse)
-    files <- dir("././data", pattern="csv")
+    files <- dir("./data/signatures", pattern="csv")
     all <- data.frame()
     for (file in files) {
-      tmp <- read.csv(file=file.path(path="././data", file)) 
+      tmp <- read.csv(file=file.path(path="./data/signatures", file)) 
       if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% select(-x)
       all <- rbind(all, tmp)
     }
@@ -59,14 +59,14 @@ a$rates %>% ggplot(aes(x = wv, y = rate)) + geom_point() + facet_grid(match~wo, 
     #   facet_grid(match~wo, labeller="label_both") + 
     #   geom_line()
     
-aa1<- read.csv2("./data/chumbley-out-wo-80-wv-50.csv",header = TRUE, sep = ",")
+aa1<- read.csv2("./data/signatures/chumbley-out-wo-80-wv-50.csv",header = TRUE, sep = ",")
 xtabs(~signif+match, data =aa1)
 
 
-aa2<- read.csv2("./data/chumbley_wo-200_wv-50.csv",header = TRUE, sep = ",")
+aa2<- read.csv2("./data/signatures/chumbley_wo-200_wv-50.csv",header = TRUE, sep = ",")
 xtabs(~signif+match, data =aa2)
 
-aa3<- read.csv2("./data/chumbley_wo-320_wv-50.csv",header = TRUE, sep = ",")
+aa3<- read.csv2("./data/signatures/chumbley_wo-320_wv-50.csv",header = TRUE, sep = ",")
 xtabs(~signif+match, data =aa3)
     
 

### 2017-09-29 doc/paper.Rmd
change of paper structure

--- 2017-09-25 doc/aboutme.Rmd
@@
 ---
-title: "Adapting toolmark based chumbley score to bullets"
-author: "Ganesh Krishnan, Heike Hofmann"
-date: "September 25, 2017"
-output: html_document
+title: Adaption of the Chumbley Score to matching of bullet striation marks
+
+# to produce blinded version set to 1
+blinded: 0
+
+authors: 
+- name: Ganesh Krishnan
+  thanks: The authors gratefully acknowledge ...
+  affiliation: Department of YYY, University of XXX
+
+- name: Heike Hofmann
+  affiliation: Department of ZZZ, University of WWW
+
+keywords:
+- 3 to 6 keywords
+- that do not appear in the title
+
+abstract: |
+
+bibliography: bibliography
+biblio-style: apsr
+
+output: 
+  pdf_document:
+    citation_package: natbib
+    keep_tex: true
+    fig_caption: true
+    latex_engine: pdflatex
+    template: template.tex
+
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+```{r, echo = FALSE, message = FALSE, warning = FALSE}
+library(knitr)
+knitr::opts_chunk$set(
+  message = FALSE,
+  warning = FALSE,
+  error = FALSE, 
+  collapse = TRUE,
+  comment = "",
+  fig.height = 6,
+  fig.width = 6,
+  fig.align = "center",
+  out.width= '\\textwidth',
+  cache = FALSE,
+  fig.path='figures/',
+  echo=FALSE
+)
+options(knitr.table.format = "latex")
 ```
+
+## Introduction
+
+What is the chumbley score?
 

### 2017-09-30 doc/paper.Rmd
type 2 error rate figure

@@
 
-```{r, echo = FALSE, message = FALSE, warning = FALSE}
+```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
 library(knitr)
@@
 options(knitr.table.format = "latex")
+library(tidyverse)
 ```
 
-## Introduction
+```{r functions}
+errorrate <- function(data, alpha) {
+  summ <- data %>% filter(!is.na(p_value)) %>%
+    mutate(signif = p_value < alpha) %>%
+    group_by(wv, wo, match, signif) %>% tally()
+  summ$error <- with(summ, match != signif)
+  summ$alpha <- alpha
+
+  rates <- summ %>% group_by(wv, wo, match, alpha) %>% summarize(
+    rate = n[error==TRUE]/sum(n)
+  )
+  rates %>% spread(match,rate) %>% rename(
+    actual = `FALSE`,
+    beta = `TRUE`
+  )
+}
+```
+
+```{r data}
+if (!file.exists("../data/all-sigs.rds")) {
+  files <- dir("../data/signatures", pattern="csv")
+  all <- data.frame()
+  for (file in files) {
+    tmp <- read.csv(file=file.path(path="../data/signatures", file)) 
+    if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% select(-x)
+    all <- rbind(all, tmp)
+  }
+  saveRDS(all, file="../data/all-sigs.rds")
+} else {
+  all <- readRDS("../data/all-sigs.rds")
+}
+
+errors <- rbind(errorrate(all, 0.001), 
+                errorrate(all, 0.005), 
+                errorrate(all, 0.01), 
+                errorrate(all, 0.05))
+```
+
+# Introduction
 
@@
 
-## Toolmarks
+# Toolmarks
 
@@
 ```
-## Bullets
+# Bullets
 
@@
 
-## Including Plots
+# Results
 
-For a significance level of 0.05
+## Signatures
 
-![](../doc/plots/error_pt05.png)
-For a significance level of 0.01
+Signatures of lands for all Hamby-44 and Hamby-252 scans made available through the NIST ballistics database \citep{nist} were considered.
 
-![](../doc/plots/error_pt01.png)
+Figure \ref{fig:type2} gives an overview of type II error rates observed when varying the window size in the optimization step.
 
-For a significance level of 0.001
-
-![](../doc/plots/error_pt001.png)
-
-![](../doc/plots/rate_pt001.png)
-
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
+```{r type2, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap="Type II error rates observed across a range of window sizes for optimization $wo$.
+For a window size of $wo = 120$ we see a drop in type II error rate across all type I rates considered.
+Smaller validation sizes $wv$ are typically associated with a smaller type II error."}
+greens <- RColorBrewer::brewer.pal(name="Greens", n = 6)
+errors %>% filter(wv %in% c(30,50)) %>%
+  ggplot(aes(x = wo, y = beta, colour=factor(alpha))) + 
+  geom_point(aes(shape=factor(wv))) +
+#  facet_grid(.~wo, labeller="label_both") + 
+  geom_smooth(aes(group=alpha), se=FALSE, method="loess") +
+  theme_bw() +
+  scale_colour_manual("Nominal type I error", values=greens[-(1:2)]) +
+  scale_shape_discrete("Size of \nvalidation window") +
+  xlab("Window size for optimization") +
+  ylab("Type II error rate")
+```

### 2017-09-30 doc/paper.Rmd
type I error rate

@@
 ---
+
+\newcommand{\hh}[1]{{\textcolor{orange}{#1}}}
+\newcommand{\cited}[1]{{\textcolor{red}{#1}}}
 
@@
 
-What is the chumbley score?
+\hh{Ganesh, could you start to flesh out the intro?}
+
+\begin{enumerate}
+\item Problem statement: same source matching based on striations
+\item Existing solution for toolmarks: Chumbley Score 
+\item How is the chumbley score computed? 
+\item Potential limitations in the application to striation marks on bullet lands: smaller in width and curved - need to adjust parameter settings
+\item Using cross-validation setup to identify appropriate parameter settings for (a) signatures and (b) profiles directly
+\end{enumerate}
 
@@
 
-Used an optimization window size of n = 500 and a validation window size of m = 50.
-It is visually clear that when the two tool marks match, the U-statistic tends to be relatively large, and when they do not match, the value tends to be relatively small.
+\hh{Ganesh, it's never a good idea to copy directly from another paper, even if you want to cite it.
+Always make sure to make it absolutely clear that it is not your work, otherwise you might be accused of plagiarism.}
 
-The p-Values associated with each U-statistic, are calculated under the assumption that the nonmatch distribution is standard normal.
-As should be the case, the p-values for nonmatching pairs are approximately uniformly distributed between zero and one, while the p-values for matching pairs are concentrated on small values (suggesting that they would be declared to be “matching”).
+Citation from \citet{chumbley}:
 
-The classification rates corresponding to the p-values where the rows correspond to the known status of the tool marks and columns represent the model conclusion.
-Of 50 pairs of known matching marks, we misclassified three pairs as nonmatches for false-negative error rate of 6% and 0 pairs of known nonmatching tool marks were identified as being matches.
+\cited{Used an optimization window size of n = 500 and a validation window size of m = 50.
+It is visually clear that when the two tool marks match, the U-statistic tends to be relatively large, and when they do not match, the value tends to be relatively small.}
 
-Null Hypothesis: The toolmarks were not made by the same tool.
-Alternate: Toolmarks made by the same tool.
+\cited{The p-Values associated with each U-statistic, are calculated under the assumption that the nonmatch distribution is standard normal.
+As should be the case, the p-values for nonmatching pairs are approximately uniformly distributed between zero and one, while the p-values for matching pairs are concentrated on small values (suggesting that they would be declared to be “matching”).}
+
+\cited{The classification rates corresponding to the p-values where the rows correspond to the known status of the tool marks and columns represent the model conclusion.
+Of 50 pairs of known matching marks, we misclassified three pairs as nonmatches for false-negative error rate of 6\% and 0 pairs of known nonmatching tool marks were identified as being matches.}
+
+\cited{Null Hypothesis: The toolmarks were not made by the same tool.
+Alternate: Toolmarks made by the same tool.}
+
 ```{r cars, echo=FALSE}
@@
 ```
+
 # Bullets
@@
 ```
+
+```{r type1, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap="Comparison of observed and nominal type I error rates  across a range of window sizes for optimization $wo$.
+The horizontal line in each facet indicates the nominal type I error rate.
+As the optimization window increase the observed type I error rate gets smaller.
+A smaller validation window tends to be associated with a higher type I error rate."}
+errors %>% filter(wv %in% c(30,50)) %>%
+  ggplot(aes(x = wo, y = actual, colour=factor(alpha))) + 
+  geom_hline(aes(yintercept=alpha), colour="grey30") +
+  geom_point(aes(shape=factor(wv))) +
+  #  facet_grid(.~wo, labeller="label_both") + 
+  geom_smooth(aes(group=alpha), se=FALSE, method="lm") +
+  theme_bw() +
+#  scale_y_log10(breaks=c(0.001,.005, 0.01, 0.05)) +
+  scale_shape_discrete("Size of \nvalidation window") +
+  ylab("Observed type I error rate") +
+  xlab("Window size for optimization") +
+  scale_colour_brewer("Nominal type I error alpha", palette="Set2") +
+  facet_wrap(~alpha, labeller="label_both", scales="free") +
+  theme(legend.position="bottom")
+
+```

### 2017-09-30 code/chumbley-test-signatures.R
code for setting up chumbley-non-random from the DB

@@ -1,0 +1,74 @@
+library(toolmaRk)
+library(RMySQL)
+library(dplyr)
+library(dbplyr)
+#library(dbConnect)
+library(ggplot2)
+library(purrr)
+###
+### DB Connection
+###
+
+dbname <- "bullets"
+user <- "buser"
+password <- readLines("buser_pass.txt")
+host <- "127.0.0.1"
+con <- src_mysql(dbname, host, user=user, password=password)
+conn<- dbConnect(MySQL(), user= user, password= password , dbname=dbname, host=host)
+
+result.metadata_derived<- metadata_derived %>% collect()
+result.metadata<- metadata %>% collect()
+
+matches <- dbReadTable(conn, "matches")
+profiles <- dbReadTable(conn, "profiles")
+sigs <- dbReadTable(conn, "signatures")
+ccf <- dbReadTable(conn, "ccf")
+
+data<- tbl(conn, "data")
+metadata_derived <- tbl(conn, "metadata_derived")
+ccf<- tbl(conn, "ccf")
+metadata<- tbl(conn, "metadata")
+runs<- tbl(conn, "runs")
+
+meta_sub <- metadata %>% filter(study %in% c("Hamby44", "Hamby252")) %>% collect()
+profiles_sub <- profiles %>% filter(land_id %in% unique(meta_sub$land_id))
+sigs <- sigs %>% left_join(profiles_sub %>% select(land_id, profile_id), by="profile_id")
+sigs <- sigs %>% filter(!is.na(land_id))
+sigs <- sigs %>% filter(run_id ==3)
+
+sigs_nest <- sigs %>% group_by(land_id, profile_id, run_id) %>% tidyr::nest()
+
+wo <- 110
+wv <- 30
+
+k <- 1
+n <- nrow(sigs_nest)
+dframe <- data.frame(matrix(rep(NA, n*(n-1)*5), ncol=5))
+names(dframe) <- c("land1_id", "land2_id", "profile1_id", "profile2_id", "chumbley")
+for (i in 1:(n-1)) {
+  for (j in (i+1):n) {
+    dframe$chumbley[k] <- I(list(chumbley_non_random(matrix(sigs_nest$data[[i]]$l30), matrix(sigs_nest$data[[j]]$l30),  window_opt=wo, window_val=wv, coarse=1)))
+    dframe$land1_id[k] <- sigs_nest$land_id[i]
[…]

### 2017-10-01 doc/paper.Rmd
bit more fleshed out

@@
   }
+  all <- all %>% filter(land1_id < land2_id)
   saveRDS(all, file="../data/all-sigs.rds")
@@
 Signatures of lands for all Hamby-44 and Hamby-252 scans made available through the NIST ballistics database \citep{nist} were considered.
+Both of these sets of scans are part of the larger Hamby study \citep{hamby} and each consist of twenty known bullets (two each from ten consecutively rifled Ruger P85 barrels) and fifteen questioned bullets (each matching one of the ten barrels).
+Ground truth for both of these Hamby sets is known and was used to assess correctness of the tests results.
+
+We used the adjusted Chumbley method as proposed in \citet{hadler} and implemented in the R package `toolmaRk` \citep{toolmark} on all pairwise land-to-land comparisons of the Hamby scans (a total of 85,491 comparisons) in the following manner:
+
+\hh{ XXX maybe we should include a table with an overview of the different parameter settings}
+
+### Signatures
 
@@
 ```
+
+Figure \ref{fig:type1} compares nominal (fixed) type I error and actually observed type I errors for the parameter settings in table XXX.
 
@@
 ```
+
+\hh{Some of the test results are missing - we should look into the numbers and parameter settings for those.}
+
+```{r}
+ns <- all %>% group_by(wv, wo, match) %>% summarize(
+  n_ss = mean(same_shift_n),
+  n_ds = mean(diff_shift_n),
+  miss = sum(is.na(p_value))
+)
+ns %>% filter(wv %in% c(30, 50)) %>% 
+  ggplot(aes(x = match, y = n_ds, shape=match, colour = factor(wv))) + 
+  geom_point() + geom_line(aes(group=wv)) + facet_wrap(~wo)
+```

### 2017-10-01 doc/paper.Rmd
failed tests included

@@
 
-```{r}
+```{r, fig.width=8, fig.height=5, fig.cap="Number of failed tests."}
 ns <- all %>% group_by(wv, wo, match) %>% summarize(
@@
   n_ds = mean(diff_shift_n),
+  zero_ss = sum(same_shift_n==0),
+  zero_ds = sum(diff_shift_n==0),
   miss = sum(is.na(p_value))
@@
 ns %>% filter(wv %in% c(30, 50)) %>% 
-  ggplot(aes(x = match, y = n_ds, shape=match, colour = factor(wv))) + 
-  geom_point() + geom_line(aes(group=wv)) + facet_wrap(~wo)
+  ggplot(aes(x = wo, y = miss, shape=factor(wv), colour = factor(match))) + 
+  geom_point() + 
+  theme_bw() +
+  ylab("Number of failed tests.")
 ```

### 2017-10-01 code/evaluate.R

@@ -1,8 +1,9 @@
 library(tidyverse)
-files <- dir("../data/signatures", pattern="csv")
+path <- "data/signatures"
+files <- dir(path, pattern="csv")
 all <- data.frame()
 for (file in files) {
-  tmp <- read.csv(file=file.path(path="../data/signatures", file)) 
+  tmp <- read.csv(file=file.path(path=path, file)) 
   if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% select(-x)
   all <- rbind(all, tmp)
 }

### 2017-10-01 doc/paper.Rmd
fiddling

--- 2017-10-01 doc/paper.Rmd
@@
 Figure \ref{fig:type1} compares nominal (fixed) type I error and actually observed type I errors for the parameter settings in table XXX.
+With an increasing size of the window used in the optimization step the observed type I error rate decreases (slighty).
+This might be related to the increasing number of tests that fail for larger window sizes, in particular for non-matching striae (see fig \ref{fig:missings}).
 
@@
 
-```{r}
+```{r missings, fig.width=8, fig.height=5, fig.cap="Number of failed tests."}
 ns <- all %>% group_by(wv, wo, match) %>% summarize(
@@
 ns %>% filter(wv %in% c(30, 50)) %>% 
-  ggplot(aes(x = match, y = n_ds, shape=match, colour = factor(wv))) + 
-  geom_point() + geom_line(aes(group=wv)) + facet_wrap(~wo)
+  ggplot(aes(x = wo, y = miss/85491*100, shape=factor(wv), colour = factor(match))) + 
+  geom_point() + 
+  theme_bw() +
+  ylab("Percent of failed tests.")
 ```

### 2017-10-04 doc/paper.Rmd
reference to joint-slip pliers

@@
 \item How is the chumbley score computed? 
-\item Potential limitations in the application to striation marks on bullet lands: smaller in width and curved - need to adjust parameter settings
+\item Potential limitations in the application to striation marks on bullet lands: smaller in width and curved - need to adjust parameter settings; something similar is done in \citet{afte-chumbley} for toolmark comparisons of slip-joint pliers.
 \item Using cross-validation setup to identify appropriate parameter settings for (a) signatures and (b) profiles directly
@@
 ```{r echo=FALSE}
-a<- matrix(NA, ncol = 3, nrow = 2)
-bullet.falsepositive<- data.frame(a)
-colnames(bullet.falsepositive)<- c("Classification", "FALSe", "TRUe")
-bullet.falsepositive$Classification<- c("FALSe", "TRUe")
-bullet.falsepositive$FALSe<- c(77564,412)
-bullet.falsepositive$TRUe<- c(4022,674)
-bullet.falsepositive
+# a<- matrix(NA, ncol = 3, nrow = 2)
+# bullet.falsepositive<- data.frame(a)
+# colnames(bullet.falsepositive)<- c("Classification", "FALSe", "TRUe")
+# bullet.falsepositive$Classification<- c("FALSe", "TRUe")
+# bullet.falsepositive$FALSe<- c(77564,412)
+# bullet.falsepositive$TRUe<- c(4022,674)
+# bullet.falsepositive
 
+all %>% filter(wo==320, wv==50) %>% xtabs(data=., ~signif+match)
 ```
@@
 ```{r echo=FALSE}
-bullet.falsepositive$FALSe<- c(78817,383)
-bullet.falsepositive$TRUe<- c(4004,798)
-bullet.falsepositive
+# bullet.falsepositive$FALSe<- c(78817,383)
+# bullet.falsepositive$TRUe<- c(4004,798)
+# bullet.falsepositive
 
+all %>% filter(wo==200, wv==50) %>% xtabs(data=., ~signif+match)
 ```
@@
 ```{r echo=FALSE}
-bullet.falsepositive$FALSe<- c(78817,450)
-bullet.falsepositive$TRUe<- c(4680,775)
-bullet.falsepositive
+# bullet.falsepositive$FALSe<- c(78817,450)
+# bullet.falsepositive$TRUe<- c(4680,775)
+# bullet.falsepositive
+
+all %>% filter(wo==80, wv==50) %>% xtabs(data=., ~signif+match)
 
@@
 ```
+
+```{r wo120, fig.width = 8, fig.height=4.5}
+wo120 <- errors %>% filter(wo==120)
+wo120 %>% gather(type, error, actual:beta) %>%
+  ggplot(aes(x = wv, y = error, colour = factor(alpha)))  + 
+  geom_line() + geom_point() + facet_wrap(~type, scales="free") +
+  theme_bw()+ 
+  theme(legend.position="bottom")
+```

### 2017-10-05 doc/paper.Rmd
small changes

@@
   ggplot(aes(x = wo, y = miss/85491*100, shape=factor(wv), colour = factor(match))) + 
+  geom_smooth(method="lm", size=.5, se=FALSE) +
   geom_point() + 
   theme_bw() +
-  ylab("Percent of failed tests.")
+  ylab("Percent of failed tests.") +
+  xlab("Size of optimization window") +
+  scale_colour_brewer("Match", palette="Set1") +
+  scale_shape_discrete("Size of \nvalidation\nwindow")
+
 ```

### 2017-10-10 doc/paper.Rmd
small stuff

--- 2017-10-01 doc/paper.Rmd
@@
 ---
-title: Adaption of the Chumbley Score to matching of bullet striation marks
-
-# to produce blinded version set to 1
+title: "Adaption of the Chumbley Score to matching of bullet striation marks"
+authors:
+- affiliation: Department of YYY, University of XXX
+  name: Ganesh Krishnan
+  thanks: The authors gratefully acknowledge ...
+- affiliation: Department of ZZZ, University of WWW
+  name: Heike Hofmann
+biblio-style: apsr
+output:
+  pdf_document:
+    citation_package: natbib
+    fig_caption: yes
+    keep_tex: yes
+    latex_engine: pdflatex
+    template: template.tex
+  html_document: default
 blinded: 0
-
-authors: 
-- name: Ganesh Krishnan
-  thanks: The authors gratefully acknowledge ...
-  affiliation: Department of YYY, University of XXX
-
-- name: Heike Hofmann
-  affiliation: Department of ZZZ, University of WWW
-
 keywords:
@@
 - that do not appear in the title
-
-abstract: |
-
 bibliography: bibliography
-biblio-style: apsr
-
-output: 
-  pdf_document:
-    citation_package: natbib
-    keep_tex: true
-    fig_caption: true
-    latex_engine: pdflatex
-    template: template.tex
-
+abstract: null
 ---
@@
 }
+
+# Multiple plot function
+#
+# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
+# - cols:   Number of columns in layout
+# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
+#
+# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
+# then plot 1 will go in the upper left, 2 will go in the upper right, and
+# 3 will go all the way across the bottom.
+#
+multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
+  library(grid)
+
+  # Make a list from the ... arguments and plotlist
+  plots <- c(list(...), plotlist)
+
+  numPlots = length(plots)
+
+  # If layout is NULL, then use 'cols' to determine layout
+  if (is.null(layout)) {
+    # Make the panel
+    # ncol: Number of columns of plots
+    # nrow: Number of rows needed, calculated from # of cols
+    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
+                    ncol = cols, nrow = ceiling(numPlots/cols))
+  }
+
+ if (numPlots==1) {
+    print(plots[[1]])
+
+  } else {
+    # Set up the page
+    grid.newpage()
+    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
+
+    # Make each plot, in the correct location
+    for (i in 1:numPlots) {
+      # Get the i,j matrix positions of the regions that contain this subplot
+      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
+
+      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
+                                      layout.pos.col = matchidx$col))
+    }
+  }
+}
+  #  ref: http://www.cookbook-r.com/Graphs/Multiple_graphs_on_one_page_(ggplot2)/
 ```
@@
 ```{r data}
-if (!file.exists("../data/all-sigs.rds")) {
+if (!file.exists("../all-sigs.rds")) {
   files <- dir("../data/signatures", pattern="csv")
@@
 } else {
-  all <- readRDS("../data/all-sigs.rds")
+  all <- readRDS("../all-sigs.rds")
 }
@@
 
-\hh{Ganesh, could you start to flesh out the intro?}
+## Problem statement: same source matching based on striations
 
-\begin{enumerate}
-\item Problem statement: same source matching based on striations
-\item Existing solution for toolmarks: Chumbley Score 
-\item How is the chumbley score computed? 
-\item Potential limitations in the application to striation marks on bullet lands: smaller in width and curved - need to adjust parameter settings
-\item Using cross-validation setup to identify appropriate parameter settings for (a) signatures and (b) profiles directly
-\end{enumerate}
+Striation marks created on Bullets by the Barrel of a firearm are considered to be unique to the barrel of the firearm.
+This was first stated by the AFTE Criteria fo Identification Committee in 1992 (AFTE 1992 article) \citet{afte-article1992}.
+The maximum number of consecutively matching striae (CMS) as first seen by Biasotti 1959 \citet{biasotti} and later mentioned in \citet{chu2013} are the striated markings that line up exactly without any dissimilarity.
+This definition involves counting a single peak as a striae.
+A modified meaning of the terminology as explained by (Hare, Hofmann and Carriquiry) includes both peaks and valleys as means of identifying striae.
+Such identification methods more often than not include subject bia and are error prone and in case of CMS as seen in the work of \citet{miller} the number of CMS may turn to out to be high even when the bullet is not fires by the same firearm.
 
-# Toolmarks
+Thus, an important aspect of same source matching in firearms is identification of bullet microtopographies that are unique to be chosen for a match, Bullet profiles and signatures prove to be such features.
+The choice of a bullet signature to uniquely define a bullet depends on finding first, a stable region on a Bullet land which minimal noise and many pronounced striation marks.
+\citet{aoas}
 
-\hh{Ganesh, it's never a good idea to copy directly from another paper, even if you want to cite it.
-Always make sure to make it absolutely clear that it is not your work, otherwise you might be accused of plagiarism.}
+The method employed by \citet{aoas} uses the cross correlation factor as a means to identify the stable region.
+The markings or striae in this region are supposed to have a high CCF with each other.
+Therefore a Bullet profile is chosen from this region by taking the cross section at a given height.
+A loess fit on the bullet profile produces  residuay which are termed as signatures and can be used as a unique identifier that can be used while matching two bullets to a firearm.
 
-Citation from \citet{chumbley}:
+## Existing solution for toolmarks: Chumbley Score
 
-\cited{Used an optimization window size of n = 500 and a validation window size of m = 50.
-It is visually clear that when the two tool marks match, the U-statistic tends to be relatively large, and when they do not match, the value tends to be relatively small.}
+Compairing pairs of toolmarks with the intention of matching it to a tool has been studied in the past, and \citet{chumbley} have described in their paper an algorithm and analytic method that compares two toolmarks and comes to the conclusion if they are from the same tool.
+The motivation for this lies in the idea of developing analytical methods for which error rates can be found.
+This moves reduces subject bias and empirical based studies to designate two toolmarks as matches or non-matches, although \citet{chumbley} use this kind of a setup to validate their proposed analytic and quantitative algorithm.
 
-\cited{The p-Values associated with each U-statistic, are calculated under the assumption that the nonmatch distribution is standard normal.
-As should be the case, the p-values for nonmatching pairs are approximately uniformly distributed between zero and one, while the p-values for matching pairs are concentrated on small values (suggesting that they would be declared to be “matching”).}
+The data that is used by \citet{chumbley} is generated by a surface profileometer that gives the height in terms of distance along a linear trace.
+This is taken perpendicular to the striations present in the toolmark.
+Two such trace are then compared using the algorithm.
 
-\cited{The classification rates corresponding to the p-values where the rows correspond to the known status of the tool marks and columns represent the model conclusion.
-Of 50 pairs of known matching marks, we misclassified three pairs as nonmatches for false-negative error rate of 6\% and 0 pairs of known nonmatching tool marks were identified as being matches.}
+Th algoritm works in two phases, namely, an optimization step and a validation step, at the end of which a Mann Whitney U statistic is calculated.
+\citet{hadler} in their paper proposed an improvement to this algorithm by trying to remove mutual dependence of parameters (due to serial correlation in surface depth values of a toolmark and because of a random sampling sub step in the validation phase which makes a group of pixels to be chosen more than once and hence introduces lack of independence) in certain steps, especially because the Mann-Whitney U statistic that is later calculated in the algorithm and used as a measure to differentiate between matches and non-matches works under the assumption of independence of parameters.
 
-\cited{Null Hypothesis: The toolmarks were not made by the same tool.
-Alternate: Toolmarks made by the same tool.}
+Since we are only interested in a non-parameteric U statistic, Hadler et al.
+proposed a normalization procedure in the Validation step that goes to some extent to address the issue of mutual dependence.
+Also the same shift and different shift substep was modified to use a deterministic rule for sampling sam-shhift and dofferent shift-samples as opposed to the originally proposed random samples.
 
-```{r cars, echo=FALSE}
-a<- matrix(NA, ncol = 3, nrow = 2)
-toolmark.falsepositive<- data.frame(a)
-colnames(toolmark.falsepositive)<- c("Classification", "Match", "Non-Match")
-toolmark.falsepositive$Classification<- c("Match", "Non-Match")
-toolmark.falsepositive$Match<- c(47,0)
-toolmark.falsepositive$`Non-Match`<- c(3,50)
-toolmark.falsepositive
+## Chumbley Algorithm and Chumbley score
 
-```
+###Optimization step
+The idea behind this step is to first identify the area of best agreement in the two toolmark data.
 
-# Bullets
+Comparison window size is predefined by the user, which in case of screwdriver toolmarks was chosen by Chumbley et al and Hadler et al as around 10 percent of the length of the toolmarks, which was around 500.
+This window is henceforth referred to as Window of optimization.
 
-Following on similar lines, the first step here is to first identify what difference does different window sizes of optimization and the validation step have, when adapting the toolmark method to bullets.
+A maximum correlation statistic is used to identify the region of best agreement, with the maximum usually seen being near 1 for both cases of matches ( which is intuitively what we expect) and non matches (which is not what we intuitively expect)
+
+First, there are a very large number of correlations calculated, for eg if the window of optimization was defined as 200 and toolmark pixel length is around 1000 then we have 1200-(200-1) = 1001.
+Can be thought of ordered arrangements without replacement where n = 1001 and r = 1.
+
+This is the number of windows we have for one toolmark and each window is compared with all windows of the second toolmark (the number of windows is again similar), and the window with the maximum correlation is identified to be the region where the toolmarks are in maximum agreement with each other
+
+###Aside (Bullets):
+Window of optimization will be shorter as the signatures are smaller.
+Idea is to keep the number of windows of optimization sufficiently large.
+This means shorter Trace segments (partition of signature or toolmark with length = size of window of optimization) and lets us compare smaller segments of one signature to another.
+
+This introduces a problem as if we go too small in the window of optimization the unique feature of the trace segments are lost and seem similar, and too large sizes vastly reduces the weightage of small features that would otherwise uniquely classify a signature and hence identify the region of agreement.
+
+This means it has a direct influence on whether we are
+Falsely rejecting the null when it is true (Type I) or Falsely accepting the null when it is false (Type II)
+
+Hence identification of the optimum size of window of optimization is very important.
+
+###Validation Step
+
+####Same-shift: 
+In this step a series of windows are chosen at random (originally by chumbley et al) and deterministically (by hadler et al), but at a common distance (rigid-shift) from the window identified as the region of best agreement in the vaildation step.
+The correlation of these windows would be as intuitively assumed i.e.
+lower than the maximum correlation window (Optimization step), but the significance is that these same shift windows will still have large enough correlation values for two toolmarks or signatures that are in reality a match.
+
+This also validates that if in the optimization step the maximum correlation window chosen (with correlation value near 1) was by accident (like in case of signatures that in reality are a non-match), all these same- shift correlations (A fixed number of these trace segments are identified) would not be anywhere large enough for all same-shift windows.
+
+####Different Shift:
+Primary reason for this substep is to give perspective to the correlation values of the same shift window correlation values.
+
+This time there are no rigid-shifts but different shifts (distance from Window of Opt with max correlation) chosen randomly by \citet{chumbley} and deterministically by Hadler et al.
+such that there is an equal possibilty of comparing a trace segment from one signature or toolmark to any one in the second signature or toolamrk.
+
+Neither of above sets of correlation are allowed to include the maximum correlation window as identified earlier.
+
+Therefore the assumption is that if two toolmarks or signatures match each other the same-shift correlations would be larger than the different-shift windows,
+and if they are not a match the correlations in the two sets will be very similar.
+
+###U Statistic:
+
+This is computed from the joint rank of all correlations of both the same and different shift samples. As given by \citet{hadler}
+
+Null Hypothesis: If the toolmarks were not match i.e not made by the same tool.
+
+Let ns and nd be the number of same shift and different shift windows
+$N = n_{s} + n_{d}$
+
+The mann whitney U statistic is given by
+$U =\sum^{ns}_{i=1}R_{s}\left( i\right)$
+
+with the standardized version which includes provision for rank ties
+
+$\overline{U}= \dfrac{U-M}{\sqrt{V}}$
+
+where prior to normalization the U-statistic has the mean as
+
+$M = n_{s}\left(\dfrac{N+1}{2}\right)$
+
+and variance
+
+$V = \dfrac{n_{s}n_{d}}{N\left(N-1\right)}\left[\Sigma^{n_{s}}R_{s}\left(i\right)^{2}+\Sigma^{n_{d}}R_{d}\left(j\right)^{2}\right] -\dfrac{n_{s} n_{d}\left(N+1\right)^{2}}{4\left(N-1\right)}$
+
+## Potential limitations in the application to striation marks on bullet lands: smaller in width and curved - need to adjust parameter settings; something similar is done in \citet{afte-chumbley} for toolmark comparisons of slip-joint pliers.
+
+Bullets are much smaller in length, width, are not flat and curved in the cross-sectional topography as opposed to tools like screw driver tips which produced longer and pronounced markings.
+This means the makings made by barrels in comparison to toolmarks may have a problem in distinctiveness.
+The majority of Bullet profiles and signatures extracted by procedures mentioned by Hare, Hofmann and Carriquiry \citet{aoas} are almost 1/4 th the size of toolmarks as used by \citet{chumbley} or even smaller.
+Striations on Bullets are made on their curved surfaces, whereas the algorithm developed by \citet{chumbley} and \citet{hadler} has only been tested for flatter and wider surfaces which have negligible curvature.
+Therefore, using methods proposed for toolmarks may need adaptation in order to give tangible results for bullets.
+Moreoverr, in order to to get flat bullet signatures and remove the curvatures some kind of smoothing needs to be applied as a pre-step which needs further investigation as to whether the level of smoothing does effect the working of the algorithm on Bullets.
+
+## Using cross-validation setup to identify appropriate parameter settings for (a) signatures and (b) profiles directly
+
+Following on similar lines to the setup of toolamrks, the first step here is to first identify what difference does different window sizes of optimization and the validation step have, when adapting the toolmark method to bullets.
 
@@
 
-WO 320 wv 50
+```{r win-comparison, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap=" The two plots on the left show how the same shift behaves in case of a matching pair and the two plots on the right show how the different shift behaves in case of a matching pair"}
+a<- matrix(NA, ncol = 3, nrow = 2)
+toolmark.falsepositive<- data.frame(a)
+colnames(toolmark.falsepositive)<- c("Classification", "Match", "Non-Match")
+toolmark.falsepositive$Classification<- c("Match", "Non-Match")
+toolmark.falsepositive$Match<- c(47,0)
+toolmark.falsepositive$`Non-Match`<- c(3,50)
+toolmark.falsepositive
+
+asd<- all %>% filter(wo==120, wv==50, match == TRUE, land1_id == 1)
+
+sigs_graphics<- readRDS("../data/sigs_generate_window.rds")
+
+sigs_graphics.match<- as.data.frame(sigs_graphics[1])
+sigs_graphics.nonmatch<- as.data.frame(sigs_graphics[2])
+colnames(sigs_graphics.match)<- gsub("^.*\\.","", colnames(sigs_graphics.match))
+colnames(sigs_graphics.nonmatch)<- gsub("^.*\\.","", colnames(sigs_graphics.nonmatch))
+
+# Matching signature
+#gi<-
+#sigs_graphics.match<- sigs_graphics.match %>% filter(land_id == c(1,9))  
+#sigs_graphics.match<- sigs_graphics.match %>% select(y, l30, land_id)
+d1<- sigs_graphics.match[which(sigs_graphics.match$land_id==1,arr.ind = TRUE),]
+d2<- sigs_graphics.match[which(sigs_graphics.match$land_id==19,arr.ind = TRUE),]
+data1<-d1 %>%select(l30)
+#data1<- data1$y
+data2<- d2 %>%select(l30)
+#data2<- data2$y
+window_opt = 120 
+window_val = 50
+coarse = 1
+data1<- matrix(unlist(data1))
+data2<- matrix(unlist(data2))
+
+  unity <- function(x) {x / sqrt(sum(x^2))} ## normalize columns of a matrix to make correlation computation faster
+
+  ####################################################
+  ##Clean the marks and compute the smooth residuals##
+  ####################################################
+
+  data1 <- matrix(data1[round((0.01*nrow(data1))):round(0.99*nrow(data1)),], ncol = 1)
+  data2 <- matrix(data2[round((0.01*nrow(data2))):round(0.99*nrow(data2)),], ncol = 1)
+
+  ##Normalize the tool marks
+  y1 <- data1 - lowess(y = data1,  x = 1:nrow(data1), f= coarse)$y
+  y2 <- data2 - lowess(y = data2,  x = 1:nrow(data2), f= coarse)$y
+
+  ############################################
+  ##Compute the observed maximum correlation##
+  ############################################
+
+  #####################
+  ##Optimization step##
+  #####################
+  ##Each column in these matrices corresponds to a window in the respective tool mark
+  y1_mat_opt <- matrix(NA, ncol = length(1:(length(y1) - (window_opt - 1))), nrow = window_opt)
+  for(l in 1:(length(y1) - (window_opt - 1))){
+    y1_mat_opt[,l] <- y1[l:(l+(window_opt - 1))]
+  }
+  y2_mat_opt <- matrix(NA, ncol = length(1:(length(y2) - (window_opt - 1))), nrow = window_opt)
+  for(l in 1:(length(y2) - (window_opt - 1))){
+    y2_mat_opt[,l] <- y2[l:(l+(window_opt - 1))]
+  }
+
+  ##Compute the correlation between all pairs of windows for the two marks
+  ##Rows in the following matrix are mark 2, columns are mark 1
+  y2_mat_opt <- apply(scale(y2_mat_opt), 2, unity)
+  y1_mat_opt <- apply(scale(y1_mat_opt), 2, unity)
+  corr_mat_opt <- t(y2_mat_opt) %*% y1_mat_opt ##correlation matrix
+  max_corr_opt_loc <- which(corr_mat_opt == max(corr_mat_opt), arr.ind = TRUE) ##pair of windows maximizing the correlation
+
+  s1.rect.opt <- data.frame(xmin=max_corr_opt_loc[1,2], xmax= max_corr_opt_loc[1,2]+window_opt, ymin=-Inf, ymax=Inf)
+    s1.rect.sameshift1 <- data.frame(xmin=max_corr_opt_loc[1,2]-window_val, xmax= max_corr_opt_loc[1,2]- 2*window_val, ymin=-Inf, ymax=Inf)
+       s1.rect.sameshift2 <- data.frame(xmin=max_corr_opt_loc[1,2]+window_opt+ window_val, xmax= max_corr_opt_loc[1,2]+ window_opt+2*window_val, ymin=-Inf, ymax=Inf)
+
+  s2.rect.opt <- data.frame(xmin=max_corr_opt_loc[1,1], xmax= max_corr_opt_loc[1,1]-window_opt, ymin=-Inf, ymax=Inf)
+    s2.rect.sameshift1 <- data.frame(xmin=max_corr_opt_loc[1,1]-window_opt-window_val, xmax= max_corr_opt_loc[1,1]-window_opt- 2*window_val, ymin=-Inf, ymax=Inf)
+       s2.rect.sameshift2 <- data.frame(xmin=max_corr_opt_loc[1,1]+ window_val, xmax= max_corr_opt_loc[1,1]+ 2*window_val, ymin=-Inf, ymax=Inf)
+
+p1<-  ggplot(data = d1, aes(x = y, y = l30)) + geom_line() + 
+  geom_rect(data=s1.rect.opt, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "black",alpha=0.5, inherit.aes = FALSE)+
+    geom_rect(data=s1.rect.sameshift1, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "red",alpha=0.1, inherit.aes = FALSE)+
+    geom_rect(data=s1.rect.sameshift2, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "blue",alpha=0.1, inherit.aes = FALSE)+
+  theme_bw() + labs(y = "Signature Land Id = 1")
+
+ p2<-   ggplot(data = d2, aes(x = y, y = l30)) + geom_line() + 
+   geom_rect(data=s2.rect.opt, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "black",alpha=0.5, inherit.aes = FALSE)+
+    geom_rect(data=s2.rect.sameshift1, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "red",alpha=0.1, inherit.aes = FALSE)+
+    geom_rect(data=s2.rect.sameshift2, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "blue",alpha=0.1, inherit.aes = FALSE)+
+  theme_bw() + labs(y = "Signature Land Id = 9")
+ ##########
+
+ d1.rect.opt <- data.frame(xmin=max_corr_opt_loc[1,2], xmax= max_corr_opt_loc[1,2]+window_opt, ymin=-Inf, ymax=Inf)
+    d1.rect.sameshift1 <- data.frame(xmin=max_corr_opt_loc[1,2]-2*window_opt, xmax= max_corr_opt_loc[1,2]-2*window_opt-window_val, ymin=-Inf, ymax=Inf)
+       d1.rect.sameshift2 <- data.frame(xmin=max_corr_opt_loc[1,2]+window_opt+ window_val, xmax= max_corr_opt_loc[1,2]+ window_opt+2*window_val, ymin=-Inf, ymax=Inf)
+
+  d2.rect.opt <- data.frame(xmin=max_corr_opt_loc[1,1], xmax= max_corr_opt_loc[1,1]-window_opt, ymin=-Inf, ymax=Inf)
+    d2.rect.sameshift1 <- data.frame(xmin=max_corr_opt_loc[1,1]-window_opt-window_val, xmax= max_corr_opt_loc[1,1]-window_opt- 2*window_val, ymin=-Inf, ymax=Inf)
+       d2.rect.sameshift2 <- data.frame(xmin=max_corr_opt_loc[1,1]+ 4*window_val, xmax= max_corr_opt_loc[1,1]+ 5*window_val, ymin=-Inf, ymax=Inf)
+
+p3<-  ggplot(data = d1, aes(x = y, y = l30)) + geom_line() + 
+  geom_rect(data=d1.rect.opt, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "black",alpha=0.5, inherit.aes = FALSE)+
+    geom_rect(data=d1.rect.sameshift1, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "red",alpha=0.1, inherit.aes = FALSE)+
+    geom_rect(data=d1.rect.sameshift2, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "blue",alpha=0.1, inherit.aes = FALSE)+
+  theme_bw() + labs(y = "Signature Land Id = 1")
+
+ p4<-   ggplot(data = d2, aes(x = y, y = l30)) + geom_line() + 
+   geom_rect(data=d2.rect.opt, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "black",alpha=0.5, inherit.aes = FALSE)+
+    geom_rect(data=d2.rect.sameshift1, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "red",alpha=0.1, inherit.aes = FALSE)+
+    geom_rect(data=d2.rect.sameshift2, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "blue",alpha=0.1, inherit.aes = FALSE)+
+  theme_bw() + labs(y = "Signature Land Id = 9")
+
+multiplot(p1, p2, p3, p4, cols=2)
+```
+
+window Optimization 320 Window Validation 50
 ```{r echo=FALSE}
-a<- matrix(NA, ncol = 3, nrow = 2)
-bullet.falsepositive<- data.frame(a)
-colnames(bullet.falsepositive)<- c("Classification", "FALSe", "TRUe")
-bullet.falsepositive$Classification<- c("FALSe", "TRUe")
-bullet.falsepositive$FALSe<- c(77564,412)
-bullet.falsepositive$TRUe<- c(4022,674)
-bullet.falsepositive
-
+all %>% filter(wo==320, wv==50) %>% xtabs(data=., ~signif+match)
 ```
-Wo 200 WV 50
+window Optimization 120 Window Validation 50
 ```{r echo=FALSE}
-bullet.falsepositive$FALSe<- c(78817,383)
-bullet.falsepositive$TRUe<- c(4004,798)
-bullet.falsepositive
-
+all %>% filter(wo==120, wv==50) %>% xtabs(data=., ~signif+match)
 ```
-Wo 80 wv 50
+window Optimization 80 Window Validation 50
 ```{r echo=FALSE}
-bullet.falsepositive$FALSe<- c(78817,450)
-bullet.falsepositive$TRUe<- c(4680,775)
-bullet.falsepositive
+all %>% filter(wo==80, wv==50) %>% xtabs(data=., ~signif+match)
 
@@
 
-\hh{ XXX maybe we should include a table with an overview of the different parameter settings}
+\hh{table with an overview of the different parameter settings}
+```{r echo=FALSE}
+param_setting<- all %>% select(wo, wv) %>% group_by(wo, wv) %>% tally() %>% select(wo,wv)
+#param_setting
+param_setting %>% tally() 
+```
 
@@
 
-\hh{Some of the test results are missing - we should look into the numbers and parameter settings for those.}
-
 ```{r missings, fig.width=8, fig.height=5, fig.cap="Number of failed tests."}
@@
   zero_ds = sum(diff_shift_n==0),
-  miss = sum(is.na(p_value))
+  miss = sum(is.na(p_value)),
+  missperc = miss/85491*100
 )
 ns %>% filter(wv %in% c(30, 50)) %>% 
-  ggplot(aes(x = wo, y = miss/85491*100, shape=factor(wv), colour = factor(match))) + 
+  ggplot(aes(x = wo, y = missperc, shape=factor(wv), colour = factor(match))) + 
+  geom_smooth(method="lm", size=.5, se=FALSE) +
   geom_point() + 
   theme_bw() +
-  ylab("Percent of failed tests.")
+  ylab("Percent of failed tests.") +
+  xlab("Size of optimization window") +
+  scale_colour_brewer("Match", palette="Set1") +
+  scale_shape_discrete("Size of \nvalidation\nwindow")
+
 ```
+Figure \ref{fig:missings} gives an overview of the number of failed tests, i.e.
+tests in which a particular parameter setting did not return a valid result.
+This happens, when the shift to align two signatures is so large, that the remaining overlap is too small to accommodate  windows for validation.
+The problem is therefore exacerbated by a larger validation window.
+Figure \ref{fig:missings} also shows that the number of failed tests is approximately linear in the size of the optimization window.
+
+```{r lms, results='asis'}
+nsnest <- ns %>% filter(wv %in% c(30,50)) %>% group_by(match, wv) %>% nest()
+nsnest <- nsnest %>% mutate(
+  model = data %>% purrr::map(.f = function(d) lm(I(missperc*100)~wo, data=d)),
+  coefs = model %>% purrr::map(.f = broom::tidy)
+)
+tab <- nsnest %>% unnest(coefs) %>% filter(term=="wo") %>% arrange(match) %>%
+  select(-term, -statistic, -p.value)
+knitr::kable(tab, digits=3, caption = "Estimates of the increase in percent of failed tests corresponding to a 100 point increase in the optimization window.", booktabs=TRUE)
+```
+
+```{r wo120, fig.width = 8, fig.height=4.5}
+wo120 <- errors %>% filter(wo==120)
+wo120 %>% gather(type, error, actual:beta) %>%
+  ggplot(aes(x = wv, y = error, colour = factor(alpha)))  + 
+  geom_line() + geom_point() + facet_wrap(~type, scales="free") +
+  theme_bw()+ 
+  theme(legend.position="bottom")
+```

### 2017-10-10 doc/paper.Rmd
still working on the bib

@@
 ```{r data}
-if (!file.exists("../all-sigs.rds")) {
+if (!file.exists("./all-sigs.rds")) {
   files <- dir("../data/signatures", pattern="csv")
@@
 } else {
-  all <- readRDS("../all-sigs.rds")
+  all <- readRDS("./all-sigs.rds")
 }
@@
 Striation marks created on Bullets by the Barrel of a firearm are considered to be unique to the barrel of the firearm.
-This was first stated by the AFTE Criteria fo Identification Committee in 1992 (AFTE 1992 article) \citet{afte-article1992}.
-The maximum number of consecutively matching striae (CMS) as first seen by Biasotti 1959 \citet{biasotti} and later mentioned in \citet{chu2013} are the striated markings that line up exactly without any dissimilarity.
+This was first stated by the AFTE Criteria fo Identification Committee in 1992 (AFTE 1992 article).
+The maximum number of consecutively matching striae (CMS) as first seen by Biasotti 1959 and later mentioned in Chu et al.
+2013 are the striated markings that line up exactly without any dissimilarity.
 This definition involves counting a single peak as a striae.
 A modified meaning of the terminology as explained by (Hare, Hofmann and Carriquiry) includes both peaks and valleys as means of identifying striae.
-Such identification methods more often than not include subject bia and are error prone and in case of CMS as seen in the work of \citet{miller} the number of CMS may turn to out to be high even when the bullet is not fires by the same firearm.
+Such identification methods more often than not include subject bia and are error prone and in case of CMS as seen in the work of Miller 1998 the number of CMS may turn to out to be high even when the bullet is not fires by the same firearm.
 
@@
 The choice of a bullet signature to uniquely define a bullet depends on finding first, a stable region on a Bullet land which minimal noise and many pronounced striation marks.
-\citet{aoas}
+(Hare, Hofmann and Carriquiry)
 
-The method employed by \citet{aoas} uses the cross correlation factor as a means to identify the stable region.
+The method employed by Hare, Hofmann and Carriquiry uses the cross correlation factor as a means to identify the stable region.
 The markings or striae in this region are supposed to have a high CCF with each other.
@@
 
-Compairing pairs of toolmarks with the intention of matching it to a tool has been studied in the past, and \citet{chumbley} have described in their paper an algorithm and analytic method that compares two toolmarks and comes to the conclusion if they are from the same tool.
+Compairing pairs of toolmarks with the intention of matching it to a tool has been studied in the past, and Chumbley et al.
+have described in their paper an algorithm and analytic method that compares two toolmarks and comes to the conclusion if they are from the same tool.
 The motivation for this lies in the idea of developing analytical methods for which error rates can be found.
-This moves reduces subject bias and empirical based studies to designate two toolmarks as matches or non-matches, although \citet{chumbley} use this kind of a setup to validate their proposed analytic and quantitative algorithm.
+This moves reduces subject bias and empirical based studies to designate two toolmarks as matches or non-matches, although Chumbley et.
+al do use this kind of a setup to validate their proposed analytic and quantitative algorithm.
 
-The data that is used by \citet{chumbley} is generated by a surface profileometer that gives the height in terms of distance along a linear trace.
+The data that is used by Chumbley et al.
+is generated by a surface profileometer that gives the height in terms of distance along a linear trace.
 This is taken perpendicular to the striations present in the toolmark.
@@
 Th algoritm works in two phases, namely, an optimization step and a validation step, at the end of which a Mann Whitney U statistic is calculated.
-\citet{hadler} in their paper proposed an improvement to this algorithm by trying to remove mutual dependence of parameters (due to serial correlation in surface depth values of a toolmark and because of a random sampling sub step in the validation phase which makes a group of pixels to be chosen more than once and hence introduces lack of independence) in certain steps, especially because the Mann-Whitney U statistic that is later calculated in the algorithm and used as a measure to differentiate between matches and non-matches works under the assumption of independence of parameters.
+Hadler et al.
+in their paper proposed an improvement to this algorithm by trying to remove mutual dependence of parameters (due to serial correlation in surface depth values of a toolmark and because of a random sampling sub step in the validation phase which makes a group of pixels to be chosen more than once and hence introduces lack of independence) in certain steps, especially because the Mann-Whitney U statistic that is later calculated in the algorithm and used as a measure to differentiate between matches and non-matches works under the assumption of independence of parameters.
 
@@
 
-## Chumbley Algorithm and Chumbley score
+## How is the chumbley score computed?
+
+Chumbley Algorithm and Chumbley score
 
@@
 ###Aside (Bullets):
-Window of optimization will be shorter as the signatures are smaller.
+Window of optimization will be shorterr as the signatures are smaller.
 Idea is to keep the number of windows of optimization sufficiently large.
@@
 ####Different Shift:
-Primary reason for this substep is to give perspective to the correlation values of the same shift window correlation values.
+Primary reason for this substep is to give perpective to the correlation values of the same shift window correlation values.
 
-This time there are no rigid-shifts but different shifts (distance from Window of Opt with max correlation) chosen randomly by \citet{chumbley} and deterministically by Hadler et al.
+This time there are no rigid-shifts but different shifts (distance from Window of Opt with max correlation) chosen randomly by Chumbley et al.
+and deterministically by Hadler et al.
 such that there is an equal possibilty of comparing a trace segment from one signature or toolmark to any one in the second signature or toolamrk.
@@
 
-This is computed from the joint rank of all correlations of both the same and different shift samples. As given by \citet{hadler}
+This is computed from the joint rank of all correlations of both the same and different shift samples.
+As given by Hadler et al.\citet{hadler}
 
-Null Hypothesis: If the toolmarks were not match i.e not made by the same tool.
+Null Hypothesis: The toolmarks were not match i.e not made by the same tool.
 
 Let ns and nd be the number of same shift and different shift windows
-$N = n_{s} + n_{d}$
+N = ns + nd
 
@@
 with the standardized version which includes provision for rank ties
-
 $\overline{U}= \dfrac{U-M}{\sqrt{V}}$
 
-where prior to normalization the U-statistic has the mean as
-
+where the mean 
 $M = n_{s}\left(\dfrac{N+1}{2}\right)$
@@
 
-$V = \dfrac{n_{s}n_{d}}{N\left(N-1\right)}\left[\Sigma^{n_{s}}R_{s}\left(i\right)^{2}+\Sigma^{n_{d}}R_{d}\left(j\right)^{2}\right] -\dfrac{n_{s} n_{d}\left(N+1\right)^{2}}{4\left(N-1\right)}$
+$V = \dfrac{n_{s}n_{d}}{N\left(N-1\right)}\left[\Sigma^{ns}R_{s}\left(i\right)^{2}+\Sigma^{nd}R_{d}\left(j\right)^{2}\right] -\dfrac{n_{s} n_{d}\left(N+1\right)^{2}}{4\left(N-1\right)}$
+ of the unnormalized U-statistic
 
@@
 This means the makings made by barrels in comparison to toolmarks may have a problem in distinctiveness.
-The majority of Bullet profiles and signatures extracted by procedures mentioned by Hare, Hofmann and Carriquiry \citet{aoas} are almost 1/4 th the size of toolmarks as used by \citet{chumbley} or even smaller.
+The majority of Bullet profiles and signatures extracted by procedures mentioned by Hare, Hofmann and Carriquiry \citet{hare1} are almost 1/4 th the size of toolmarks as used by \citet{chumbley} or even smaller.
 Striations on Bullets are made on their curved surfaces, whereas the algorithm developed by \citet{chumbley} and \citet{hadler} has only been tested for flatter and wider surfaces which have negligible curvature.
@@
 
-```{r win-comparison, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap=" The two plots on the left show how the same shift behaves in case of a matching pair and the two plots on the right show how the different shift behaves in case of a matching pair"}
+```{r cars, echo=FALSE}
 a<- matrix(NA, ncol = 3, nrow = 2)
@@
 
-sigs_graphics<- readRDS("../data/sigs_generate_window.rds")
+sigs_graphics<- readRDS("./sigs_generate_window.rds")
 
@@
 param_setting<- all %>% select(wo, wv) %>% group_by(wo, wv) %>% tally() %>% select(wo,wv)
-#param_setting
+param_setting
 param_setting %>% tally() 
@@
 
+\hh{Some of the test results are missing - we should look into the numbers and parameter settings for those.}
+
 ```{r missings, fig.width=8, fig.height=5, fig.cap="Number of failed tests."}
@@
   zero_ds = sum(diff_shift_n==0),
-  miss = sum(is.na(p_value)),
-  missperc = miss/85491*100
+  miss = sum(is.na(p_value))
 )
 ns %>% filter(wv %in% c(30, 50)) %>% 
-  ggplot(aes(x = wo, y = missperc, shape=factor(wv), colour = factor(match))) + 
+  ggplot(aes(x = wo, y = miss/85491*100, shape=factor(wv), colour = factor(match))) + 
   geom_smooth(method="lm", size=.5, se=FALSE) +
@@
 
-```
-Figure \ref{fig:missings} gives an overview of the number of failed tests, i.e.
-tests in which a particular parameter setting did not return a valid result.
-This happens, when the shift to align two signatures is so large, that the remaining overlap is too small to accommodate  windows for validation.
-The problem is therefore exacerbated by a larger validation window.
-Figure \ref{fig:missings} also shows that the number of failed tests is approximately linear in the size of the optimization window.
-
-```{r lms, results='asis'}
-nsnest <- ns %>% filter(wv %in% c(30,50)) %>% group_by(match, wv) %>% nest()
-nsnest <- nsnest %>% mutate(
-  model = data %>% purrr::map(.f = function(d) lm(I(missperc*100)~wo, data=d)),
-  coefs = model %>% purrr::map(.f = broom::tidy)
-)
-tab <- nsnest %>% unnest(coefs) %>% filter(term=="wo") %>% arrange(match) %>%
-  select(-term, -statistic, -p.value)
-knitr::kable(tab, digits=3, caption = "Estimates of the increase in percent of failed tests corresponding to a 100 point increase in the optimization window.", booktabs=TRUE)
 ```

### 2017-10-10 doc/paper.Rmd
correction to the file location

@@
 ```{r data}
-if (!file.exists("./all-sigs.rds")) {
+if (!file.exists("../all-sigs.rds")) {
   files <- dir("../data/signatures", pattern="csv")

### 2017-10-11 doc/paper.Rmd
couple of comments

@@
 ## Problem statement: same source matching based on striations
+
+\hh{'considered to be unique' by AFTE - that is right, but we should not use this argument, and I don't think we have to.
+It's better to focus on the problem at hand: striation marks are used to determine same source for bullets.
+XXX we could include some figures here XXX Current standards ask for a visual inspection of the marks under a comparison microscope by a firearms examiner.
+One of the problems raised by the PCAST report in 2009 is the subjectivity involved in this comparison.
+There are different approaches used to achieve more objectivity and assess error rates.
+XXX here we need some examples of approaches XXX
+For toolmarks, the Chumbley score has been suggested to determine same-source. XXX summary of results from original paper and the adjustment XXX
+In this paper we are interested in how this method has to be adjusted to best be applied to the smaller marks on bullets. }
 
@@
   zero_ds = sum(diff_shift_n==0),
-  miss = sum(is.na(p_value))
+  miss = sum(is.na(p_value)),
+  missperc = miss/85491*100
 )
 ns %>% filter(wv %in% c(30, 50)) %>% 
-  ggplot(aes(x = wo, y = miss/85491*100, shape=factor(wv), colour = factor(match))) + 
+  ggplot(aes(x = wo, y = missperc, shape=factor(wv), colour = factor(match))) + 
   geom_smooth(method="lm", size=.5, se=FALSE) +
@@
 
+```
+Figure \ref{fig:missings} gives an overview of the number of failed tests, i.e.
+tests in which a particular parameter setting did not return a valid result.
+This happens, when the shift to align two signatures is so large, that the remaining overlap is too small to accommodate  windows for validation.
+The problem is therefore exacerbated by a larger validation window.
+Figure \ref{fig:missings} also shows that the number of failed tests is approximately linear in the size of the optimization window.
+
+```{r lms, results='asis'}
+nsnest <- ns %>% filter(wv %in% c(30,50)) %>% group_by(match, wv) %>% nest()
+nsnest <- nsnest %>% mutate(
+  model = data %>% purrr::map(.f = function(d) lm(I(missperc*100)~wo, data=d)),
+  coefs = model %>% purrr::map(.f = broom::tidy)
+)
+tab <- nsnest %>% unnest(coefs) %>% filter(term=="wo") %>% arrange(match) %>%
+  select(-term, -statistic, -p.value)
+knitr::kable(tab, digits=3, caption = "Estimates of the increase in percent of failed tests corresponding to a 100 point increase in the optimization window.", booktabs=TRUE)
 ```

### 2017-10-11 doc/paper.Rmd
small stuff

@@
 \hh{table with an overview of the different parameter settings}
-```{r echo=FALSE}
+```{r, echo=FALSE, results='asis'}
 param_setting<- all %>% select(wo, wv) %>% group_by(wo, wv) %>% tally() %>% select(wo,wv)
-#param_setting
-param_setting %>% tally() 
+#t(param_setting)
+#param_setting %>% tally()
+
+knitr::kable(t(param_setting), digits=0, caption = "Overview of parameter settings used for optimization and validation windows for bullet land signatures.", booktabs=TRUE)
 ```
@@
 Smaller validation sizes $wv$ are typically associated with a smaller type II error."}
-greens <- RColorBrewer::brewer.pal(name="Greens", n = 6)
 errors %>% filter(wv %in% c(30,50)) %>%
@@
   theme_bw() +
-  scale_colour_manual("Nominal type I error", values=greens[-(1:2)]) +
-  scale_shape_discrete("Size of \nvalidation window") +
-  xlab("Window size for optimization") +
+  scale_colour_brewer(expression("Nominal type I error"~alpha), palette="Set2") +
+  scale_shape_discrete("Size of validation\nwindow, wv") +
+  xlab(expression("Window size for optimization, "~wo)) +
   ylab("Type II error rate")
+
+labels = expression(P[M1](tilde(z)>0),P[M0](tilde(z)>0))
 ```
@@
 #  scale_y_log10(breaks=c(0.001,.005, 0.01, 0.05)) +
-  scale_shape_discrete("Size of \nvalidation window") +
+  scale_shape_discrete("Size of validation window, wv") +
   ylab("Observed type I error rate") +
-  xlab("Window size for optimization") +
-  scale_colour_brewer("Nominal type I error alpha", palette="Set2") +
+  xlab(expression("Window size for optimization, "~wo)) +
+  scale_colour_brewer(expression("Nominal type I error "~alpha), palette="Set2") +
   facet_wrap(~alpha, labeller="label_both", scales="free") +
@@
 
-```{r missings, fig.width=8, fig.height=5, fig.cap="Number of failed tests."}
+```{r missings, fig.width=8, fig.height=5, fig.cap="Number of failed tests by the window optimization size, wo, and ground truth.
+Test results from different sources have a much higher chance to fail, raising the question, whether failed tests should be treated as rejections of the null hypothesis of same source."}
 ns <- all %>% group_by(wv, wo, match) %>% summarize(
@@
   ylab("Percent of failed tests.") +
-  xlab("Size of optimization window") +
+  xlab("Size of optimization window, wo") +
   scale_colour_brewer("Match", palette="Set1") +
-  scale_shape_discrete("Size of \nvalidation\nwindow")
-
+  scale_shape_discrete("Size of \nvalidation\nwindow, wv")
 ```

### 2017-10-18 doc/paper.Rmd
references added
comments and some updates to bib info

@@
 authors:
-- affiliation: Department of YYY, University of XXX
+- affiliation: Department of Statistics, Iowa State University 
   name: Ganesh Krishnan
   thanks: The authors gratefully acknowledge ...
-- affiliation: Department of ZZZ, University of WWW
+- affiliation: Department of Statistics and csafe, Iowa State University 
   name: Heike Hofmann
@@
   fig.path='figures/',
-  echo=FALSE
+  echo=FALSE,
+  cache=TRUE
 )
@@
 ```{r data}
-if (!file.exists("../all-sigs.rds")) {
+if (!file.exists("../data/all-sigs.rds")) {
   files <- dir("../data/signatures", pattern="csv")
@@
 } else {
-  all <- readRDS("../all-sigs.rds")
+  all <- readRDS("../data/all-sigs.rds")
 }
@@
 
-## Problem statement: same source matching based on striations
+## Problem statement and motivation
 
@@
 XXX we could include some figures here XXX Current standards ask for a visual inspection of the marks under a comparison microscope by a firearms examiner.
-One of the problems raised by the PCAST report in 2009 is the subjectivity involved in this comparison.
+One of the problems raised by the NAS report in 2009 is the subjectivity involved in this comparison and the need for determining error rates \citep{NAS:2009}.
 There are different approaches used to achieve more objectivity and assess error rates.
@@
 In this paper we are interested in how this method has to be adjusted to best be applied to the smaller marks on bullets. }
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{images/B6-B2-L6.png}
+```{r, echo=FALSE, fig.width=8, fig.height=3, out.width='\\textwidth'}
+land <- read.csv("../data/b6b2l6.csv")
+land %>% ggplot(aes(x = y, y=resid)) +geom_line() +theme_bw()
+```
+
+\caption{\label{fig:rgl} Image of a bullet land from a confocal light microscope at 20 fold magnification.}
+
+\end{figure}
 
@@
 Compairing pairs of toolmarks with the intention of matching it to a tool has been studied in the past, and \citet{chumbley} have described in their paper an algorithm and analytic method that compares two toolmarks and comes to the conclusion if they are from the same tool.
+\hh{XXX anything to do with motivation has to come first - move this to the first paragraph of the intro XXX}
 The motivation for this lies in the idea of developing analytical methods for which error rates can be found.
@@
 
-The data that is used by \citet{chumbley} is generated by a surface profileometer that gives the height in terms of distance along a linear trace.
+\hh{XXX data and results come after the intro to chumbley.}
+The data that is used by \citet{chumbley} is generated by a surface profilometer that gives the height in terms of distance along a linear trace.
 This is taken perpendicular to the striations present in the toolmark.
@@
 
+\hh{XXX start with defining the data, i.e.
+we are working with random processes $x(t)$ and $y(t)$ ...
+you should also go into the coarseness parameter here - that is a pre-processing step before the optimization.
+}
 Th algoritm works in two phases, namely, an optimization step and a validation step, at the end of which a Mann Whitney U statistic is calculated.
@@
 
-A maximum correlation statistic is used to identify the region of best agreement, with the maximum usually seen being near 1 for both cases of matches ( which is intuitively what we expect) and non matches (which is not what we intuitively expect)
+A maximum correlation statistic is used to identify the region of best agreement, with the maximum usually seen being near 1 for both cases of matches ( which is intuitively what we expect) and non matches (which is not what we intuitively expect) \hh{XXX get rid of the parentheses - put the content into a sentence after your first statement.}
 
+\hh{XXX the process you describe below is the cross-correlation of two series $x(t)$ and $y(t)$. }
 First, there are a very large number of correlations calculated, for eg if the window of optimization was defined as 200 and toolmark pixel length is around 1000 then we have 1200-(200-1) = 1001.
@@
 
-This introduces a problem as if we go too small in the window of optimization the unique feature of the trace segments are lost and seem similar, and too large sizes vastly reduces the weightage of small features that would otherwise uniquely classify a signature and hence identify the region of agreement.
+\hh{XXX discussion of limitations should come after introducing the method. }
+This introduces a problem as if we go too small in the window of optimization the unique feature of the trace segments are lost and seem similar, and too large sizes vastly reduces the weight of small features that would otherwise uniquely classify a signature and hence identify the region of agreement.
 
@@
 
-####Different Shift:
-Primary reason for this substep is to give perspective to the correlation values of the same shift window correlation values.
-
-This time there are no rigid-shifts but different shifts (distance from Window of Opt with max correlation) chosen randomly by \citet{chumbley} and deterministically by Hadler et al.
-such that there is an equal possibilty of comparing a trace segment from one signature or toolmark to any one in the second signature or toolamrk.
-
-Neither of above sets of correlation are allowed to include the maximum correlation window as identified earlier.
-
-Therefore the assumption is that if two toolmarks or signatures match each other the same-shift correlations would be larger than the different-shift windows,
-and if they are not a match the correlations in the two sets will be very similar.
-
-###U Statistic:
-
-This is computed from the joint rank of all correlations of both the same and different shift samples. As given by \citet{hadler}
-
-Null Hypothesis: If the toolmarks were not match i.e not made by the same tool.
-
-Let ns and nd be the number of same shift and different shift windows
-$N = n_{s} + n_{d}$
-
-The mann whitney U statistic is given by
-$U =\sum^{ns}_{i=1}R_{s}\left( i\right)$
-
-with the standardized version which includes provision for rank ties
-
-$\overline{U}= \dfrac{U-M}{\sqrt{V}}$
-
-where prior to normalization the U-statistic has the mean as
-
-$M = n_{s}\left(\dfrac{N+1}{2}\right)$
-
-and variance
-
-$V = \dfrac{n_{s}n_{d}}{N\left(N-1\right)}\left[\Sigma^{n_{s}}R_{s}\left(i\right)^{2}+\Sigma^{n_{d}}R_{d}\left(j\right)^{2}\right] -\dfrac{n_{s} n_{d}\left(N+1\right)^{2}}{4\left(N-1\right)}$
-
-## Potential limitations in the application to striation marks on bullet lands: smaller in width and curved - need to adjust parameter settings; something similar is done in \citet{afte-chumbley} for toolmark comparisons of slip-joint pliers.
-
-Bullets are much smaller in length, width, are not flat and curved in the cross-sectional topography as opposed to tools like screw driver tips which produced longer and pronounced markings.
-This means the makings made by barrels in comparison to toolmarks may have a problem in distinctiveness.
-The majority of Bullet profiles and signatures extracted by procedures mentioned by Hare, Hofmann and Carriquiry \citet{aoas} are almost 1/4 th the size of toolmarks as used by \citet{chumbley} or even smaller.
-Striations on Bullets are made on their curved surfaces, whereas the algorithm developed by \citet{chumbley} and \citet{hadler} has only been tested for flatter and wider surfaces which have negligible curvature.
-Therefore, using methods proposed for toolmarks may need adaptation in order to give tangible results for bullets.
-Moreoverr, in order to to get flat bullet signatures and remove the curvatures some kind of smoothing needs to be applied as a pre-step which needs further investigation as to whether the level of smoothing does effect the working of the algorithm on Bullets.
-
-## Using cross-validation setup to identify appropriate parameter settings for (a) signatures and (b) profiles directly
-
-Following on similar lines to the setup of toolamrks, the first step here is to first identify what difference does different window sizes of optimization and the validation step have, when adapting the toolmark method to bullets.
-
-The marking made on bullets are smaller than toolmarks and is also less wider. The idea is to find out possible areas of error while adapting the score based method proposed for toolmarks.
-
-Bullet signatures being compared at this time are from the Hamby 44 and Hamby 252 data, which have the best set of known-matched and known non-matches.
-
-Bullet signatures are chosen by
-1. Filtering out Land_id for Profiles from the Hamby 44 and Hamby 252 data and removing 
-all NA values
-2. run_id = 3 is chosen, Seems like the signatures generated from this run_id give the closest
-match. Different run_id's have some different settings for generating the signatures
-(The level of smoothing does not seem to be one of them)
-
-The bullet signatures when generated already included the LOWESS smoothing.
-Therefore, the coarseness factor is set to 1 while running the chumbley_non_random()
-which generates the same_shift, different_shift, U-Stat and P_value parameters.
-
-```{r win-comparison, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap=" The two plots on the left show how the same shift behaves in case of a matching pair and the two plots on the right show how the different shift behaves in case of a matching pair"}
+```{r win-comparison, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap=" The two plots on the left show how the same shift behaves in case of a matching pair and the two plots on the right show how the different shift behaves in case of a matching pair."}
 a<- matrix(NA, ncol = 3, nrow = 2)
@@
 
+####Different Shift:
+Primary reason for this substep is to give perspective to the correlation values of the same shift window correlation values.
+
+This time there are no rigid-shifts but different shifts (distance from Window of Opt with max correlation) chosen randomly by \citet{chumbley} and deterministically by Hadler et al.
+such that there is an equal possibilty of comparing a trace segment from one signature or toolmark to any one in the second signature or toolamrk.
+
+Neither of above sets of correlation are allowed to include the maximum correlation window as identified earlier.
+
+Therefore the assumption is that if two toolmarks or signatures match each other the same-shift correlations would be larger than the different-shift windows,
+and if they are not a match the correlations in the two sets will be very similar.
+
+###U Statistic:
+
+This is computed from the joint rank of all correlations of both the same and different shift samples. As given by \citet{hadler}
+
+Null Hypothesis: If the toolmarks were not match i.e not made by the same tool.
+
+Let ns and nd be the number of same shift and different shift windows
+$N = n_{s} + n_{d}$
+
+The mann whitney U statistic is given by
+$U =\sum^{ns}_{i=1}R_{s}\left( i\right)$
+
+with the standardized version which includes provision for rank ties
+
+$\overline{U}= \dfrac{U-M}{\sqrt{V}}$
+
+where prior to normalization the U-statistic has the mean as
+
+$M = n_{s}\left(\dfrac{N+1}{2}\right)$
+
+and variance
+
+$V = \dfrac{n_{s}n_{d}}{N\left(N-1\right)}\left[\Sigma^{n_{s}}R_{s}\left(i\right)^{2}+\Sigma^{n_{d}}R_{d}\left(j\right)^{2}\right] -\dfrac{n_{s} n_{d}\left(N+1\right)^{2}}{4\left(N-1\right)}$
+
+## Potential limitations in the application to striation marks on bullet lands: smaller in width and curved - need to adjust parameter settings; something similar is done in \citet{afte-chumbley} for toolmark comparisons of slip-joint pliers.
+
+Bullets are much smaller in length, width, are not flat and curved in the cross-sectional topography as opposed to tools like screw driver tips which produced longer and pronounced markings.
+This means the makings made by barrels in comparison to toolmarks may have a problem in distinctiveness.
+The majority of Bullet profiles and signatures extracted by procedures mentioned by Hare, Hofmann and Carriquiry \citet{aoas} are almost 1/4 th the size of toolmarks as used by \citet{chumbley} or even smaller.
+Striations on Bullets are made on their curved surfaces, whereas the algorithm developed by \citet{chumbley} and \citet{hadler} has only been tested for flatter and wider surfaces which have negligible curvature.
+Therefore, using methods proposed for toolmarks may need adaptation in order to give tangible results for bullets.
+Moreoverr, in order to to get flat bullet signatures and remove the curvatures some kind of smoothing needs to be applied as a pre-step which needs further investigation as to whether the level of smoothing does effect the working of the algorithm on Bullets.
+
+## Using cross-validation setup to identify appropriate parameter settings for (a) signatures and (b) profiles directly
+
+Following on similar lines to the setup of toolamrks, the first step here is to first identify what difference does different window sizes of optimization and the validation step have, when adapting the toolmark method to bullets.
+
+The marking made on bullets are smaller than toolmarks and is also less wider. The idea is to find out possible areas of error while adapting the score based method proposed for toolmarks.
+
+Bullet signatures being compared at this time are from the Hamby 44 and Hamby 252 data, which have the best set of known-matched and known non-matches.
+
+Bullet signatures are chosen by
+1. Filtering out Land_id for Profiles from the Hamby 44 and Hamby 252 data and removing 
+all NA values
+2. run_id = 3 is chosen, Seems like the signatures generated from this run_id give the closest
+match. Different run_id's have some different settings for generating the signatures
+(The level of smoothing does not seem to be one of them)
+
+The bullet signatures when generated already included the LOWESS smoothing.
+Therefore, the coarseness factor is set to 1 while running the chumbley_non_random()
+which generates the same_shift, different_shift, U-Stat and P_value parameters.
+
 window Optimization 320 Window Validation 50
@@
   theme_bw() +
-  scale_colour_brewer(expression("Nominal type I error"~alpha), palette="Set2") +
+  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") +
   scale_shape_discrete("Size of validation\nwindow, wv") +

### 2017-10-20 doc/paper.Rmd
intro revised- and some other changes made

@@
 \newcommand{\hh}[1]{{\textcolor{orange}{#1}}}
+\newcommand{\gk}[1]{{\textcolor{green}{#1}}}
 \newcommand{\cited}[1]{{\textcolor{red}{#1}}}
@@
 In this paper we are interested in how this method has to be adjusted to best be applied to the smaller marks on bullets. }
+\gk{done}
+
+Striation marks are used to determine same source for bullets.
+Current standards ask for a visual inspection of the marks under a comparison microscope by a firearms examiner.
+One of the problems raised by the NAS report in 2009 is the subjectivity involved in this comparison and the need for determining error rates \citep{NAS:2009}.
+The issue of subjectivity of a firearms examiner has been reviewed by many authors where things like scientific principle testability and error rates are considered to be defining aspects for an objective analysis such that without identifying the uncertainities associated with a method of comparison, it is inconclusive to regards an analysis complete.
+This means that with the usual method of visual comparison that is adopted by firearms examiners, lies the problem of lack of any systemized quantification and as such there seems no certain way of associating it with any kind of mathematical probability.
+Since determining error rates has been duly noted as a fundamental problem in forensic science \citep{NAS:2009}, there is a need of methods that address this problem.
+Some of the methods that have been used to estimate error rates in firearm examinations include Automatic cartridge case comparison where same source identification of cartridge cases and associated error rates have been provided by \citep{riva}, although as noted by \citep{aoas}, in this case alignments of striae involves roation of planes, which cannot be generalized for bullets.
+In case of methods that use machine learning algorithms some methods which are bootsrap based methods often tend to give over-estimated error rates \citep{efron}, but there are alternate error estimation techniques as described by \citep{aoas}, \citep{efron} and  \citep{vorburger2016} which give better results.
+
+In case of methods that do not use machine learning algorithm, matching two striation marks with each other in order to identify if it is from the same source, can also be done using a well defined comparative statistical algorithm.
+The statistical algorithm is therefore expected  to go through a step by step procedure of identifying the class characteristics i.e striation marks for bullets, choosing the striae, and then follow a systematic procedure of comparing two striae with each other on the basis of criteria like cross-correlation (or others).
+Such a methodology would make it possible to determine error rates too and give definite results regarding what proportion of cases would the analysis be successful in identifying a match correctly and in how many cases would it end up being a false positive.
+
+Identification criteria like consecutively matching striae (CMS) as first seen by Biasotti 1959 \citet{biasotti} and later mentioned in \citet{chu2013} more often than not include subject bias and are error prone and as seen in the work of \citet{miller} the number of CMS may turn to out to be high even when the bullet is not fires by the same firearm.
+Parameters like cross-correlation factor as seen in the extensive comparisons made by \citep{aoas} seem to perform much better matching purposes.
+
+An important aspect of same source matching in firearms is identification of bullet microtopographies that are unique to be chosen for a match, Bullet profiles and signatures prove to be such features.
+The choice of a bullet signature to uniquely define a bullet depends on finding first, a stable region on a Bullet land which minimal noise and many pronounced striation marks.
+\citet{aoas}
+
+The method employed by \citet{aoas} uses the cross correlation factor as a means to identify the stable region.
+The markings or striae in this region are supposed to have a high CCF with each other.
+Therefore a Bullet profile is chosen from this region by taking the cross section at a given height.
+A loess fit on the bullet profile produces  residuay which are termed as signatures and can be used as a unique identifier that can be used while matching two bullets to a firearm.
+
+Compairing pairs of toolmarks on the otherhand, with the intention of matching it to a tool has been studied relatively more in the past, and \citet{chumbley} have described in their paper an algorithm and analytic method that compares two toolmarks and come to the conclusion if they are from the same tool or not.
+The method also determines the error rates, reduces subject bias and designate the two toolmarks as matches or non-matches with respect to a source.
+\citet{chumbley} used an empirical based setup to validate their proposed analytic and quantitative algorithm.
+
+\gk{XXX summary of results from original paper and the adjustment XXX}
 
@@
 
-Striation marks created on Bullets by the Barrel of a firearm are considered to be unique to the barrel of the firearm.
-This was first stated by the AFTE Criteria fo Identification Committee in 1992 (AFTE 1992 article) \citet{afte-article1992}.
-The maximum number of consecutively matching striae (CMS) as first seen by Biasotti 1959 \citet{biasotti} and later mentioned in \citet{chu2013} are the striated markings that line up exactly without any dissimilarity.
-This definition involves counting a single peak as a striae.
-A modified meaning of the terminology as explained by (Hare, Hofmann and Carriquiry) includes both peaks and valleys as means of identifying striae.
-Such identification methods more often than not include subject bia and are error prone and in case of CMS as seen in the work of \citet{miller} the number of CMS may turn to out to be high even when the bullet is not fires by the same firearm.
-
-Thus, an important aspect of same source matching in firearms is identification of bullet microtopographies that are unique to be chosen for a match, Bullet profiles and signatures prove to be such features.
-The choice of a bullet signature to uniquely define a bullet depends on finding first, a stable region on a Bullet land which minimal noise and many pronounced striation marks.
-\citet{aoas}
-
-The method employed by \citet{aoas} uses the cross correlation factor as a means to identify the stable region.
-The markings or striae in this region are supposed to have a high CCF with each other.
-Therefore a Bullet profile is chosen from this region by taking the cross section at a given height.
-A loess fit on the bullet profile produces  residuay which are termed as signatures and can be used as a unique identifier that can be used while matching two bullets to a firearm.
-
 ## Existing solution for toolmarks: Chumbley Score
 
-Compairing pairs of toolmarks with the intention of matching it to a tool has been studied in the past, and \citet{chumbley} have described in their paper an algorithm and analytic method that compares two toolmarks and comes to the conclusion if they are from the same tool.
 \hh{XXX anything to do with motivation has to come first - move this to the first paragraph of the intro XXX}
-The motivation for this lies in the idea of developing analytical methods for which error rates can be found.
-This moves reduces subject bias and empirical based studies to designate two toolmarks as matches or non-matches, although \citet{chumbley} use this kind of a setup to validate their proposed analytic and quantitative algorithm.
-
 \hh{XXX data and results come after the intro to chumbley.}
-The data that is used by \citet{chumbley} is generated by a surface profilometer that gives the height in terms of distance along a linear trace.
-This is taken perpendicular to the striations present in the toolmark.
-Two such trace are then compared using the algorithm.
-
 \hh{XXX start with defining the data, i.e.
@@
 }
-Th algoritm works in two phases, namely, an optimization step and a validation step, at the end of which a Mann Whitney U statistic is calculated.
+\gk{done}
+
+The Chumbley score algorithm takes input as two random processes $x(t_1)$ and $y(t_2)$ which denote two sets of marks or striae.
+These marks or striae are potentially from two different bullets or two different toolmarks whose source needs to be identified as being same or different.
+The marks or striae are indexed by the pixel location where $t_1$ is for the first striae referred to as $x$ and $t_2$ is for the second striae which is referred as $y$.
+$t_1$ and $t_2$ need not be the same but are usually of similar lenghtsThe similarity is then judged by the algorithm on the basis of cross-correlation of a fixed and constant number consecutive pixels say $k$ taken from the two indexed marks $x(t_1)$ and $y(t_2)$ such that in theory $k$ remains smaller that length of the two striae or marks which is the same as $t_1$ and $t_2$.
+Depending on the what stage of the algorithm we are in, matching of different pixel lengths and locations is done, which at the end effectively compares all possible windows that would guarantee in quantifying the two marks or striae as coming from the same source or not, which also lets us assess the error rates by checking for a large number of cases.
+
+The algorithm works in two phases, namely, an optimization step and a validation step, at the end of which a Mann Whitney U statistic is calculated.
+A pre-processing step to the algorithm is to choose a coarseness value which  is used as a parameter to the LOWESS smoothing function.
+The coarseness essentially gives the proportion of points which influence the smooth at each value, which means larger values lead to more smoothness.
+The LOWESS smoothing is applied to each of two sets of vectorized striae or marks $x(t_1)$ and $y(t_2)$, before proceeding to the algorithm.
+
 \citet{hadler} in their paper proposed an improvement to this algorithm by trying to remove mutual dependence of parameters (due to serial correlation in surface depth values of a toolmark and because of a random sampling sub step in the validation phase which makes a group of pixels to be chosen more than once and hence introduces lack of independence) in certain steps, especially because the Mann-Whitney U statistic that is later calculated in the algorithm and used as a measure to differentiate between matches and non-matches works under the assumption of independence of parameters.
@@
 
+The data that is used by \citet{chumbley} is generated by a surface profilometer that gives the height in terms of distance along a linear trace.
+This is taken perpendicular to the striations present in the toolmark.
+Two such trace are then compared using the algorithm.
+
 ## Chumbley Algorithm and Chumbley score
@@
 The idea behind this step is to first identify the area of best agreement in the two toolmark data.
-
 Comparison window size is predefined by the user, which in case of screwdriver toolmarks was chosen by Chumbley et al and Hadler et al as around 10 percent of the length of the toolmarks, which was around 500.
@@
 
-A maximum correlation statistic is used to identify the region of best agreement, with the maximum usually seen being near 1 for both cases of matches ( which is intuitively what we expect) and non matches (which is not what we intuitively expect) \hh{XXX get rid of the parentheses - put the content into a sentence after your first statement.}
+A maximum correlation statistic is used to identify the region of best agreement, with the maximum usually seen being near 1 for both cases which is what we intuitively expect for matches, and something which we do not intuitively expect for non matches \gk{done}\hh{XXX get rid of the parentheses - put the content into a sentence after your first statement.}
 
 \hh{XXX the process you describe below is the cross-correlation of two series $x(t)$ and $y(t)$. }
-First, there are a very large number of correlations calculated, for eg if the window of optimization was defined as 200 and toolmark pixel length is around 1000 then we have 1200-(200-1) = 1001.
+
+First, there are a very large number of cross-correlations calculated for the two series of striae or marks $x(t_1)$ and $y(t_2)$, for eg if the window of optimization was defined as 200 and toolmark pixel length is around 1000 then we have 1200-(200-1) = 1001.
 Can be thought of ordered arrangements without replacement where n = 1001 and r = 1.
@@
 This is the number of windows we have for one toolmark and each window is compared with all windows of the second toolmark (the number of windows is again similar), and the window with the maximum correlation is identified to be the region where the toolmarks are in maximum agreement with each other
-
-###Aside (Bullets):
-Window of optimization will be shorter as the signatures are smaller.
-Idea is to keep the number of windows of optimization sufficiently large.
-This means shorter Trace segments (partition of signature or toolmark with length = size of window of optimization) and lets us compare smaller segments of one signature to another.
-
-\hh{XXX discussion of limitations should come after introducing the method. }
-This introduces a problem as if we go too small in the window of optimization the unique feature of the trace segments are lost and seem similar, and too large sizes vastly reduces the weight of small features that would otherwise uniquely classify a signature and hence identify the region of agreement.
-
-This means it has a direct influence on whether we are
-Falsely rejecting the null when it is true (Type I) or Falsely accepting the null when it is false (Type II)
-
-Hence identification of the optimum size of window of optimization is very important.
 
@@
 This means the makings made by barrels in comparison to toolmarks may have a problem in distinctiveness.
-The majority of Bullet profiles and signatures extracted by procedures mentioned by Hare, Hofmann and Carriquiry \citet{aoas} are almost 1/4 th the size of toolmarks as used by \citet{chumbley} or even smaller.
+The majority of Bullet profiles and signatures extracted by procedures mentioned by  \citet{aoas} are almost 1/4 th the size of toolmarks as used by \citet{chumbley} or even smaller.
 Striations on Bullets are made on their curved surfaces, whereas the algorithm developed by \citet{chumbley} and \citet{hadler} has only been tested for flatter and wider surfaces which have negligible curvature.
 Therefore, using methods proposed for toolmarks may need adaptation in order to give tangible results for bullets.
-Moreoverr, in order to to get flat bullet signatures and remove the curvatures some kind of smoothing needs to be applied as a pre-step which needs further investigation as to whether the level of smoothing does effect the working of the algorithm on Bullets.
+Moreover, in order to to get flat bullet signatures and remove the curvatures some kind of smoothing needs to be applied as a pre-step which needs further investigation as to whether the level of smoothing does effect the working of the algorithm on Bullets.
+
+Also when in the optimization step, the Window of optimization for bullets will be shorter as the signatures are smaller.
+The idea is to keep the number of windows of optimization sufficiently large, which means shorter Trace segments (partition of signature or toolmark with length = size of window of optimization) that lets us compare smaller segments of one signature to another.
+
+This introduces a problem as, if we go too small in the window of optimization, the unique features of the trace segments are lost and seem similar, while too large sizes vastly reduces the weight of small features that would otherwise uniquely classify a signature and hence identify the region of agreement.
+
+Thus the Window of Optimization has a direct influence on whether we are
+Falsely rejecting the null when it is true (Type I) or Falsely accepting the null when it is false (Type II) making identification of the optimum size of window of optimization very important.
 
@@
 
-\hh{table with an overview of the different parameter settings}
 ```{r, echo=FALSE, results='asis'}
 param_setting<- all %>% select(wo, wv) %>% group_by(wo, wv) %>% tally() %>% select(wo,wv)
+#param_setting<- data.frame(matrix(param_setting, nrow = 16))
 #t(param_setting)
 #param_setting %>% tally() 
-
-knitr::kable(t(param_setting), digits=0, caption = "Overview of parameter settings used for optimization and validation windows for bullet land signatures.", booktabs=TRUE)
+library(kableExtra)
+knitr::kable(t(param_setting), digits=0, caption = "Overview of parameter settings used for optimization and validation windows for bullet land signatures.", format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("scale_down"))
 ```

### 2017-10-20 doc/paper.Rmd
results from chumbley and hadler

@@
 Therefore a Bullet profile is chosen from this region by taking the cross section at a given height.
-A loess fit on the bullet profile produces  residuay which are termed as signatures and can be used as a unique identifier that can be used while matching two bullets to a firearm.
+A loess fit on the bullet profile produces  residuals which are termed as signatures and can be used as a unique identifier that can be used while matching two bullets to a firearm.
 
-Compairing pairs of toolmarks on the otherhand, with the intention of matching it to a tool has been studied relatively more in the past, and \citet{chumbley} have described in their paper an algorithm and analytic method that compares two toolmarks and come to the conclusion if they are from the same tool or not.
+Compairing pairs of toolmarks on the otherhand, with the intention of matching it to a tool has been studied relatively more in the past as compared to bullets, and \citet{chumbley} have described in their paper an algorithm and analytic method that compares two toolmarks and come to the conclusion if they are from the same tool or not.
 The method also determines the error rates, reduces subject bias and designate the two toolmarks as matches or non-matches with respect to a source.
@@
 
-\gk{XXX summary of results from original paper and the adjustment XXX}
-
+\citep{chumbley} found that the algorithm gives false-positives in slightly over 2 cases (or 2%) of the time and about 9 false-negatives (or 9%) for every 100 comparisons.
+\citep{hadler} on the other hand, in the improved version of the algorithm first described by \citep{chumbley} found a false-positives for 0 cases and 3 false-negatives (or 6%) for 50 comparisons of knwon matches.
 \begin{figure}
@@
 
-The Chumbley score algorithm takes input as two random processes $x(t_1)$ and $y(t_2)$ which denote two sets of marks or striae.
+The Chumbley score algorithm takes input as two vectorized processes $x(t_1)$ and $y(t_2)$ which denote two sets of marks or striae.
 These marks or striae are potentially from two different bullets or two different toolmarks whose source needs to be identified as being same or different.
@@
 Two such trace are then compared using the algorithm.
+
+In case of
 

### 2017-10-20 doc/paper.Rmd
,,

@@
 \citep{chumbley} found that the algorithm gives false-positives in slightly over 2 cases (or 2%) of the time and about 9 false-negatives (or 9%) for every 100 comparisons.
-\citep{hadler} on the other hand, in the improved version of the algorithm first described by \citep{chumbley} found a false-positives for 0 cases and 3 false-negatives (or 6%) for 50 comparisons of knwon matches.
+\citep{hadler} on the other hand, in the improved version of the algorithm first described by \citep{chumbley} found false-positives for 0 cases and 3 false-negatives (or 6%) for 50 comparisons of knwon matches.
 \begin{figure}
@@
 Two such trace are then compared using the algorithm.
-
-In case of
 

### 2017-10-29 doc/paper.Rmd
introduction and other changes, results from chumbley and hadler included.
some information on profiles and how they are being used in the chumbley algorithm _.

@@
 
-\hh{XXX the process you describe below is the cross-correlation of two series $x(t)$ and $y(t)$. }
+\hh{XXX the process you describe below is the cross-correlation of two series $x(t)$ and $y(t)$.}
 
@@
 ```
+
+### Profiles
+
+The profiles are cross-sectional values of the the bullet striation mark which are chosen at an optimum height (x as used by \citet{aoas}).
+This x or height is not a randomly chosen level.
+The rationale behind the choice has been explained by \citet{aoas}.
+A region is first chosen where the cross-correlation seems to change very less and in this region an optimum height is chosen.
+The profiles generally resemble a curve which is more or less similar to a quadratic curve (a quadratic fit to the raw data values of the profile is not an exact fit but it does show a similar trend).
+Profiles are the set of raw values representing the striation marks, and signatures are generated from these by removal of the inherent curvature and applying some smoothing (the signatures generated by \citet{aoas} use a loess function for smoothing).
+
+Coming back to the database of Hamby 44 and Hamby 252 datasets, the run_id = 3 was used when applying the chumbley algorithm on the signatures.
+The run_id not only defines the level of smoothing but also signifies the chosen height at which the profiles were selected initially.
+Another important aspect is the range of horizontal values (which is referred to as the y values in \citet{aoas}) in the signatures.
+Thes have already been pre-processed in the database to not include any grooves.
+
+Therefore for the sake of comparison the run_id = 3 is still chosen so as to ensure that the horizontal values remain the same as that of the signatures.
+This also gives us profiles with the grooves removed.
+
+The idea therefore is to first use these raw values of the profile directly in the chumbley algorithm, and see how the algorithm performs for different coarseness values (smoothing parameter as referred in the function LOWESS used in the chumbley algorithm).

### 2017-10-29 doc/paper.Rmd
some information on profiles and how they are being used in the chumbley algorithm _.

@@
 
-Coming back to the database of Hamby 44 and Hamby 252 datasets, the run_id = 3 was used when applying the chumbley algorithm on the signatures.
+Coming back to the database of Hamby-44 and Hamby-252 datasets, the run_id = 3 was used when applying the chumbley algorithm on the signatures.
 The run_id not only defines the level of smoothing but also signifies the chosen height at which the profiles were selected initially.
@@
 
-The idea therefore is to first use these raw values of the profile directly in the chumbley algorithm, and see how the algorithm performs for different coarseness values (smoothing parameter as referred in the function LOWESS used in the chumbley algorithm).
+The idea therefore is to first use these raw values of the profile directly in the chumbley algorithm, and see how the algorithm performs for different coarseness values (smoothing parameter as referred in the function LOWESS used in the chumbley algorithm)

### 2017-11-11 doc/paper.Rmd
couple of comments

@@
 \newcommand{\cited}[1]{{\textcolor{red}{#1}}}
+
+\tableofcontents
 
@@
 
-# Introduction
+# Introduction and Background
 
@@
 
-## Existing solution for toolmarks: Chumbley Score
+## The Chumbley Score Test
 
-\hh{XXX anything to do with motivation has to come first - move this to the first paragraph of the intro XXX}
-\hh{XXX data and results come after the intro to chumbley.}
-\hh{XXX start with defining the data, i.e.
-we are working with random processes $x(t)$ and $y(t)$ ...
-you should also go into the coarseness parameter here - that is a pre-processing step before the optimization.
-}
-\gk{done}
+\hh{You need to distinguish between $x(t)$ and $x(t_1)$ - $x(t)$ is a spatial process for some $t$.
+$x(t_1)$ is the realization of the process in $t_1$.
+- Alternatively you could also follow the notation in Hadler et al on 'digitized toolmarks at a pixel level' - where our 'pixel' is the resolution of the confocal light microscope, i.e.
+a pixel corresponds to 0.645 microns.}
 
@@
 
-## Chumbley Algorithm and Chumbley score
+## Detailed algorithm
 
@@
 
-## Potential limitations in the application to striation marks on bullet lands: smaller in width and curved - need to adjust parameter settings; something similar is done in \citet{afte-chumbley} for toolmark comparisons of slip-joint pliers.
+## Potential limitations of the Chumbley Score
+
+\hh{in the application to striation marks on bullet lands: smaller in width and curved - need to adjust parameter settings; something similar is done in \citet{afte-chumbley} for toolmark comparisons of slip-joint pliers.
+XXX work in }
 
@@
 
-## Using cross-validation setup to identify appropriate parameter settings for (a) signatures and (b) profiles directly
+## Testing setup:
+\hh{Using cross-validation setup to identify appropriate parameter settings for (a) signatures and (b) profiles directly}
 
-Following on similar lines to the setup of toolamrks, the first step here is to first identify what difference does different window sizes of optimization and the validation step have, when adapting the toolmark method to bullets.
+Following on similar lines to the setup of toolmarks, the first step here is to first identify what difference does different window sizes of optimization and the validation step have, when adapting the toolmark method to bullets.
 
@@
 The idea therefore is to first use these raw values of the profile directly in the chumbley algorithm, and see how the algorithm performs for different coarseness values (smoothing parameter as referred in the function LOWESS used in the chumbley algorithm)
+
+## Discussion

### 2017-11-11 doc/paper.Rmd
results for profiles

@@
 
+errorratecoarse <- function(data, alpha) {
+  summ <- data %>% filter(!is.na(p_value)) %>%
+    mutate(signif = p_value < alpha) %>%
+    group_by(wv, wo, coarse, match, signif) %>% tally()
+  summ$error <- with(summ, match != signif)
+  summ$alpha <- alpha
+
+  rates <- summ %>% group_by(wv, wo, coarse, match, alpha) %>% summarize(
+    rate = n[error==TRUE]/sum(n)
+  )
+  rates %>% spread(match,rate) %>% rename(
+    actual = `FALSE`,
+    beta = `TRUE`
+  )
+}
 # Multiple plot function
@@
   all <- readRDS("../data/all-sigs.rds")
+}
+
+errors2 <- rbind(errorratecoarse(all, 0.001), 
+                errorratecoarse(all, 0.005), 
+                errorratecoarse(all, 0.01), 
+                errorratecoarse(all, 0.05))
+```
+
+```{r data-profiles}
+if (!file.exists("../data/all-profiles.rds")) {
+  files <- dir("../data/profiles/", pattern="csv")
+  all <- data.frame()
+  for (file in files) {
+    tmp <- read.csv(file=file.path(path="../data/profiles", file)) 
+    #print(names(tmp))
+    if (length(grep("coarse",names(tmp))) == 0)
+      tmp$coarse <- gsub(".*coar-(.*).csv", "\\1", file)
+    if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% select(-x)
+    all <- rbind(all, tmp)
+  }
+  all <- all %>% filter(land1_id < land2_id)
+  all <- all %>% mutate(coarse = as.numeric(gsub("pt", ".", coarse)))
+  saveRDS(all, file="../data/all-profiles.rds")
+} else {
+  all <- readRDS("../data/all-profiles.rds")
 }
@@
 
+\hh{XXX ideas should go to the previous section on setting up the study; in this section we focus on results only XXX}
 The idea therefore is to first use these raw values of the profile directly in the chumbley algorithm, and see how the algorithm performs for different coarseness values (smoothing parameter as referred in the function LOWESS used in the chumbley algorithm)
 
+```{r, fig.cap="Type II errors for different levels of coarseness. XXX We need more values around .25"}
+errors2 %>% ggplot(aes(x = coarse, y=beta, colour=factor(alpha))) +
+  geom_point() + geom_line() +
+  theme_bw() +
+  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2")
+```
+
 ## Discussion

### 2017-11-12 doc/paper.Rmd
attempt at summarizing overal total errors

@@
     mutate(signif = p_value < alpha) %>%
-    group_by(wv, wo, match, signif) %>% tally()
-  summ$error <- with(summ, match != signif)
-  summ$alpha <- alpha
-
-  rates <- summ %>% group_by(wv, wo, match, alpha) %>% summarize(
-    rate = n[error==TRUE]/sum(n)
-  )
-  rates %>% spread(match,rate) %>% rename(
-    actual = `FALSE`,
-    beta = `TRUE`
-  )
-}
-
-errorratecoarse <- function(data, alpha) {
-  summ <- data %>% filter(!is.na(p_value)) %>%
-    mutate(signif = p_value < alpha) %>%
-    group_by(wv, wo, coarse, match, signif) %>% tally()
+    group_by(wv, wo, match, coarse, signif) %>% tally()
   summ$error <- with(summ, match != signif)
@@
   )
-  rates %>% spread(match,rate) %>% rename(
+  totals <- summ %>% group_by(wv, wo, coarse, alpha) %>% summarize(
+    total= sum(n[error==TRUE])/sum(n)
+  )
+  rates <- rates %>% spread(match,rate) %>% rename(
     actual = `FALSE`,
@@
   )
+  left_join(rates, totals)
 }
+
 # Multiple plot function
@@
 
-errors2 <- rbind(errorratecoarse(all, 0.001), 
-                errorratecoarse(all, 0.005), 
-                errorratecoarse(all, 0.01), 
-                errorratecoarse(all, 0.05))
+all$coarse <- 1
+errors <- rbind(errorrate(all, 0.001), 
+                errorrate(all, 0.005), 
+                errorrate(all, 0.01), 
+                errorrate(all, 0.05))
 ```
@@
 
-errors <- rbind(errorrate(all, 0.001), 
+errors2 <- rbind(errorrate(all, 0.001), 
                 errorrate(all, 0.005), 
@@
 
+### comparison of total error with signatures and profiles
+
+```{r}
+errors %>% ggplot(aes(x = wv, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + facet_grid(~wo) +
+  geom_point(data=errors2, aes(shape="profiles"))
+```
 ## Discussion

### 2017-12-26 doc/paper.tex
modification to algorithm in the same shift step..search vicinity

--- 2017-11-11 doc/paper.Rmd
@@
----
-title: "Adaption of the Chumbley Score to matching of bullet striation marks"
-authors:
-- affiliation: Department of Statistics, Iowa State University 
-  name: Ganesh Krishnan
-  thanks: The authors gratefully acknowledge ...
-- affiliation: Department of Statistics and csafe, Iowa State University 
-  name: Heike Hofmann
-biblio-style: apsr
-output:
-  pdf_document:
-    citation_package: natbib
-    fig_caption: yes
-    keep_tex: yes
-    latex_engine: pdflatex
-    template: template.tex
-  html_document: default
-blinded: 0
-keywords:
-- 3 to 6 keywords
-- that do not appear in the title
-bibliography: bibliography
-abstract: null
----
+% !TeX program = pdfLaTeX
+\documentclass[12pt]{article}
+\usepackage{amsmath}
+\usepackage{amsthm}
+\usepackage{graphicx,psfrag,epsf}
+\usepackage{enumerate}
+\usepackage{natbib}
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage[table]{xcolor}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{hyperref}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+
+\usepackage{url} % not crucial - just used below for the URL
+
+%\pdfminorversion=4
+% NOTE: To produce blinded version, replace "0" with "1" below.
+\newcommand{\blind}{0}
+
+% DON'T change margins - should be 1 inch all around.
+\addtolength{\oddsidemargin}{-.5in}%
+\addtolength{\evensidemargin}{-.5in}%
+\addtolength{\textwidth}{1in}%
+\addtolength{\textheight}{1.3in}%
+\addtolength{\topmargin}{-.8in}%
+
+\newenvironment{definition}[1]% environment name 
+{% begin code 
+  \par\vspace{.75\baselineskip}\noindent 
+  \textbf{Definition (#1)}\begin{itshape}% 
+  \par\vspace{.5\baselineskip}\noindent\ignorespaces 
+}% 
+{% end code 
+  \end{itshape}\ignorespacesafterend 
+}
+
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
+\begin{document}
+
+\def\spacingset#1{\renewcommand{\baselinestretch}%
+{#1}\small\normalsize} \spacingset{1}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\if0\blind
+{
+  \title{\bf Adaption of the Chumbley Score to matching of bullet striation marks}
+
+  \author{
+        Ganesh Krishnan \thanks{The authors gratefully acknowledge \ldots{}} \\
+    Department of Statistics, Iowa State University\\
+     and \\     Heike Hofmann \\
+    Department of Statistics and csafe, Iowa State University\\
+      }
+  \maketitle
+} \fi
+
+\if1\blind
+{
+  \bigskip
+  \bigskip
+  \bigskip
+  \begin{center}
+    {\LARGE\bf Adaption of the Chumbley Score to matching of bullet striation marks}
+  \end{center}
+  \medskip
+} \fi
+
+\bigskip
+\begin{abstract}
+
+\end{abstract}
+
+\noindent%
+{\it Keywords:} 3 to 6 keywords, that do not appear in the title
+\vfill
+
+\newpage
+\spacingset{1.45} % DON'T change the spacing!
 
@@
 
-\tableofcontents
+\section{Introduction}\label{introduction}
 
-```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
-library(knitr)
-knitr::opts_chunk$set(
-  message = FALSE,
-  warning = FALSE,
-  error = FALSE, 
-  collapse = TRUE,
-  comment = "",
-  fig.height = 6,
-  fig.width = 6,
-  fig.align = "center",
-  out.width= '\\textwidth',
-  cache = FALSE,
-  fig.path='figures/',
-  echo=FALSE,
-  cache=TRUE
-)
-options(knitr.table.format = "latex")
-library(tidyverse)
-```
-
-```{r functions}
-errorrate <- function(data, alpha) {
-  summ <- data %>% filter(!is.na(p_value)) %>%
-    mutate(signif = p_value < alpha) %>%
-    group_by(wv, wo, match, signif) %>% tally()
-  summ$error <- with(summ, match != signif)
-  summ$alpha <- alpha
-
-  rates <- summ %>% group_by(wv, wo, match, alpha) %>% summarize(
-    rate = n[error==TRUE]/sum(n)
-  )
-  rates %>% spread(match,rate) %>% rename(
-    actual = `FALSE`,
-    beta = `TRUE`
-  )
-}
-
-# Multiple plot function
-#
-# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
-# - cols:   Number of columns in layout
-# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
-#
-# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
-# then plot 1 will go in the upper left, 2 will go in the upper right, and
-# 3 will go all the way across the bottom.
-#
-multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
-  library(grid)
-
-  # Make a list from the ... arguments and plotlist
-  plots <- c(list(...), plotlist)
-
-  numPlots = length(plots)
-
-  # If layout is NULL, then use 'cols' to determine layout
-  if (is.null(layout)) {
-    # Make the panel
-    # ncol: Number of columns of plots
-    # nrow: Number of rows needed, calculated from # of cols
-    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
-                    ncol = cols, nrow = ceiling(numPlots/cols))
-  }
-
- if (numPlots==1) {
-    print(plots[[1]])
-
-  } else {
-    # Set up the page
-    grid.newpage()
-    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
-
-    # Make each plot, in the correct location
-    for (i in 1:numPlots) {
-      # Get the i,j matrix positions of the regions that contain this subplot
-      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
-
-      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
-                                      layout.pos.col = matchidx$col))
-    }
-  }
-}
-  #  ref: http://www.cookbook-r.com/Graphs/Multiple_graphs_on_one_page_(ggplot2)/
-```
-
-```{r data}
-if (!file.exists("../data/all-sigs.rds")) {
-  files <- dir("../data/signatures", pattern="csv")
-  all <- data.frame()
-  for (file in files) {
-    tmp <- read.csv(file=file.path(path="../data/signatures", file)) 
-    if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% select(-x)
-    all <- rbind(all, tmp)
-  }
-  all <- all %>% filter(land1_id < land2_id)
-  saveRDS(all, file="../data/all-sigs.rds")
-} else {
-  all <- readRDS("../data/all-sigs.rds")
-}
-
-errors <- rbind(errorrate(all, 0.001), 
-                errorrate(all, 0.005), 
-                errorrate(all, 0.01), 
-                errorrate(all, 0.05))
-```
-
-# Introduction and Background
-
-## Problem statement and motivation
+\subsection{Problem statement and
+motivation}\label{problem-statement-and-motivation}
 
@@
 
-Striation marks are used to determine same source for bullets.
-Current standards ask for a visual inspection of the marks under a comparison microscope by a firearms examiner.
-One of the problems raised by the NAS report in 2009 is the subjectivity involved in this comparison and the need for determining error rates \citep{NAS:2009}.
-The issue of subjectivity of a firearms examiner has been reviewed by many authors where things like scientific principle testability and error rates are considered to be defining aspects for an objective analysis such that without identifying the uncertainities associated with a method of comparison, it is inconclusive to regards an analysis complete.
-This means that with the usual method of visual comparison that is adopted by firearms examiners, lies the problem of lack of any systemized quantification and as such there seems no certain way of associating it with any kind of mathematical probability.
-Since determining error rates has been duly noted as a fundamental problem in forensic science \citep{NAS:2009}, there is a need of methods that address this problem.
-Some of the methods that have been used to estimate error rates in firearm examinations include Automatic cartridge case comparison where same source identification of cartridge cases and associated error rates have been provided by \citep{riva}, although as noted by \citep{aoas}, in this case alignments of striae involves roation of planes, which cannot be generalized for bullets.
-In case of methods that use machine learning algorithms some methods which are bootsrap based methods often tend to give over-estimated error rates \citep{efron}, but there are alternate error estimation techniques as described by \citep{aoas}, \citep{efron} and  \citep{vorburger2016} which give better results.
+Striation marks are used to determine same source for bullets. Current
+standards ask for a visual inspection of the marks under a comparison
+microscope by a firearms examiner. One of the problems raised by the NAS
+report in 2009 is the subjectivity involved in this comparison and the
+need for determining error rates \citep{NAS:2009}. The issue of
+subjectivity of a firearms examiner has been reviewed by many authors
+where things like scientific principle testability and error rates are
+considered to be defining aspects for an objective analysis such that
+without identifying the uncertainities associated with a method of
+comparison, it is inconclusive to regards an analysis complete. This
+means that with the usual method of visual comparison that is adopted by
+firearms examiners, lies the problem of lack of any systemized
+quantification and as such there seems no certain way of associating it
+with any kind of mathematical probability. Since determining error rates
+has been duly noted as a fundamental problem in forensic science
+\citep{NAS:2009}, there is a need of methods that address this problem.
+Some of the methods that have been used to estimate error rates in
+firearm examinations include Automatic cartridge case comparison where
+same source identification of cartridge cases and associated error rates
+have been provided by \citep{riva}, although as noted by \citep{aoas},
+in this case alignments of striae involves roation of planes, which
+cannot be generalized for bullets. In case of methods that use machine
+learning algorithms some methods which are bootsrap based methods often
+tend to give over-estimated error rates \citep{efron}, but there are
+alternate error estimation techniques as described by \citep{aoas},
+\citep{efron} and \citep{vorburger2016} which give better results.
 
-In case of methods that do not use machine learning algorithm, matching two striation marks with each other in order to identify if it is from the same source, can also be done using a well defined comparative statistical algorithm.
-The statistical algorithm is therefore expected  to go through a step by step procedure of identifying the class characteristics i.e striation marks for bullets, choosing the striae, and then follow a systematic procedure of comparing two striae with each other on the basis of criteria like cross-correlation (or others).
-Such a methodology would make it possible to determine error rates too and give definite results regarding what proportion of cases would the analysis be successful in identifying a match correctly and in how many cases would it end up being a false positive.
+In case of methods that do not use machine learning algorithm, matching
+two striation marks with each other in order to identify if it is from
+the same source, can also be done using a well defined comparative
+statistical algorithm. The statistical algorithm is therefore expected
+to go through a step by step procedure of identifying the class
+characteristics i.e striation marks for bullets, choosing the striae,
+and then follow a systematic procedure of comparing two striae with each
+other on the basis of criteria like cross-correlation (or others). Such
+a methodology would make it possible to determine error rates too and
+give definite results regarding what proportion of cases would the
+analysis be successful in identifying a match correctly and in how many
+cases would it end up being a false positive.
 
-Identification criteria like consecutively matching striae (CMS) as first seen by Biasotti 1959 \citet{biasotti} and later mentioned in \citet{chu2013} more often than not include subject bias and are error prone and as seen in the work of \citet{miller} the number of CMS may turn to out to be high even when the bullet is not fires by the same firearm.
-Parameters like cross-correlation factor as seen in the extensive comparisons made by \citep{aoas} seem to perform much better matching purposes.
+Identification criteria like consecutively matching striae (CMS) as
+first seen by Biasotti 1959 \citet{biasotti} and later mentioned in
+\citet{chu2013} more often than not include subject bias and are error
+prone and as seen in the work of \citet{miller} the number of CMS may
+turn to out to be high even when the bullet is not fires by the same
+firearm. Parameters like cross-correlation factor as seen in the
+extensive comparisons made by \citep{aoas} seem to perform much better
+matching purposes.
 
-An important aspect of same source matching in firearms is identification of bullet microtopographies that are unique to be chosen for a match, Bullet profiles and signatures prove to be such features.
-The choice of a bullet signature to uniquely define a bullet depends on finding first, a stable region on a Bullet land which minimal noise and many pronounced striation marks.
-\citet{aoas}
+An important aspect of same source matching in firearms is
+identification of bullet microtopographies that are unique to be chosen
+for a match, Bullet profiles and signatures prove to be such features.
+The choice of a bullet signature to uniquely define a bullet depends on
+finding first, a stable region on a Bullet land which minimal noise and
+many pronounced striation marks. \citet{aoas}
 
-The method employed by \citet{aoas} uses the cross correlation factor as a means to identify the stable region.
-The markings or striae in this region are supposed to have a high CCF with each other.
-Therefore a Bullet profile is chosen from this region by taking the cross section at a given height.
-A loess fit on the bullet profile produces  residuals which are termed as signatures and can be used as a unique identifier that can be used while matching two bullets to a firearm.
+The method employed by \citet{aoas} uses the cross correlation factor as
+a means to identify the stable region. The markings or striae in this
+region are supposed to have a high CCF with each other. Therefore a
+Bullet profile is chosen from this region by taking the cross section at
+a given height. A loess fit on the bullet profile produces residuals
+which are termed as signatures and can be used as a unique identifier
+that can be used while matching two bullets to a firearm.
 
-Compairing pairs of toolmarks on the otherhand, with the intention of matching it to a tool has been studied relatively more in the past as compared to bullets, and \citet{chumbley} have described in their paper an algorithm and analytic method that compares two toolmarks and come to the conclusion if they are from the same tool or not.
-The method also determines the error rates, reduces subject bias and designate the two toolmarks as matches or non-matches with respect to a source.
-\citet{chumbley} used an empirical based setup to validate their proposed analytic and quantitative algorithm.
+Compairing pairs of toolmarks on the otherhand, with the intention of
+matching it to a tool has been studied relatively more in the past as
+compared to bullets, and \citet{chumbley} have described in their paper
+an algorithm and analytic method that compares two toolmarks and come to
+the conclusion if they are from the same tool or not. The method also
+determines the error rates, reduces subject bias and designate the two
+toolmarks as matches or non-matches with respect to a source.
+\citet{chumbley} used an empirical based setup to validate their
+proposed analytic and quantitative algorithm.
 
-\citep{chumbley} found that the algorithm gives false-positives in slightly over 2 cases (or 2%) of the time and about 9 false-negatives (or 9%) for every 100 comparisons.
-\citep{hadler} on the other hand, in the improved version of the algorithm first described by \citep{chumbley} found false-positives for 0 cases and 3 false-negatives (or 6%) for 50 comparisons of knwon matches.
+\citep{chumbley} found that the algorithm gives false-positives in
+slightly over 2 cases (or 2\%) of the time and about 9 false-negatives
+(or 9\%) for every 100 comparisons. \citep{hadler} on the other hand, in
+the improved version of the algorithm first described by
+\citep{chumbley} found false-positives for 0 cases and 3 false-negatives
+(or 6\%) for 50 comparisons of knwon matches.
+
 \begin{figure}
@@
 \includegraphics[width=\textwidth]{images/B6-B2-L6.png}
-```{r, echo=FALSE, fig.width=8, fig.height=3, out.width='\\textwidth'}
-land <- read.csv("../data/b6b2l6.csv")
-land %>% ggplot(aes(x = y, y=resid)) +geom_line() +theme_bw()
-```
+
+\begin{center}\includegraphics[width=\textwidth]{figures/unnamed-chunk-1-1} \end{center}
 
@@
 
-## The Chumbley Score Test
+\subsection{Existing solution for toolmarks: Chumbley
+Score}\label{existing-solution-for-toolmarks-chumbley-score}
 
-\hh{You need to distinguish between $x(t)$ and $x(t_1)$ - $x(t)$ is a spatial process for some $t$.
-$x(t_1)$ is the realization of the process in $t_1$.
-- Alternatively you could also follow the notation in Hadler et al on 'digitized toolmarks at a pixel level' - where our 'pixel' is the resolution of the confocal light microscope, i.e.
-a pixel corresponds to 0.645 microns.}
+\hh{XXX anything to do with motivation has to come first - move this to the first paragraph of the intro XXX}
+\hh{XXX data and results come after the intro to chumbley.}
+\hh{XXX start with defining the data, i.e.
+we are working with random processes $x(t)$ and $y(t)$ ...
+you should also go into the coarseness parameter here - that is a pre-processing step before the optimization.
+}
+\gk{done}
 
-The Chumbley score algorithm takes input as two vectorized processes $x(t_1)$ and $y(t_2)$ which denote two sets of marks or striae.
-These marks or striae are potentially from two different bullets or two different toolmarks whose source needs to be identified as being same or different.
-The marks or striae are indexed by the pixel location where $t_1$ is for the first striae referred to as $x$ and $t_2$ is for the second striae which is referred as $y$.
-$t_1$ and $t_2$ need not be the same but are usually of similar lenghtsThe similarity is then judged by the algorithm on the basis of cross-correlation of a fixed and constant number consecutive pixels say $k$ taken from the two indexed marks $x(t_1)$ and $y(t_2)$ such that in theory $k$ remains smaller that length of the two striae or marks which is the same as $t_1$ and $t_2$.
-Depending on the what stage of the algorithm we are in, matching of different pixel lengths and locations is done, which at the end effectively compares all possible windows that would guarantee in quantifying the two marks or striae as coming from the same source or not, which also lets us assess the error rates by checking for a large number of cases.
+The Chumbley score algorithm takes input as two vectorized processes
+\(x(t_1)\) and \(y(t_2)\) which denote two sets of marks or striae.
+These marks or striae are potentially from two different bullets or two
+different toolmarks whose source needs to be identified as being same or
+different. The marks or striae are indexed by the pixel location where
+\(t_1\) is for the first striae referred to as \(x\) and \(t_2\) is for
+the second striae which is referred as \(y\). \(t_1\) and \(t_2\) need
+not be the same but are usually of similar lenghtsThe similarity is then
+judged by the algorithm on the basis of cross-correlation of a fixed and
+constant number consecutive pixels say \(k\) taken from the two indexed
+marks \(x(t_1)\) and \(y(t_2)\) such that in theory \(k\) remains
+smaller that length of the two striae or marks which is the same as
+\(t_1\) and \(t_2\). Depending on the what stage of the algorithm we are
+in, matching of different pixel lengths and locations is done, which at
+the end effectively compares all possible windows that would guarantee
+in quantifying the two marks or striae as coming from the same source or
+not, which also lets us assess the error rates by checking for a large
+number of cases.
 
-The algorithm works in two phases, namely, an optimization step and a validation step, at the end of which a Mann Whitney U statistic is calculated.
-A pre-processing step to the algorithm is to choose a coarseness value which  is used as a parameter to the LOWESS smoothing function.
-The coarseness essentially gives the proportion of points which influence the smooth at each value, which means larger values lead to more smoothness.
-The LOWESS smoothing is applied to each of two sets of vectorized striae or marks $x(t_1)$ and $y(t_2)$, before proceeding to the algorithm.
+The algorithm works in two phases, namely, an optimization step and a
+validation step, at the end of which a Mann Whitney U statistic is
+calculated. A pre-processing step to the algorithm is to choose a
+coarseness value which is used as a parameter to the LOWESS smoothing
+function. The coarseness essentially gives the proportion of points
+which influence the smooth at each value, which means larger values lead
+to more smoothness. The LOWESS smoothing is applied to each of two sets
+of vectorized striae or marks \(x(t_1)\) and \(y(t_2)\), before
+proceeding to the algorithm.
 
-\citet{hadler} in their paper proposed an improvement to this algorithm by trying to remove mutual dependence of parameters (due to serial correlation in surface depth values of a toolmark and because of a random sampling sub step in the validation phase which makes a group of pixels to be chosen more than once and hence introduces lack of independence) in certain steps, especially because the Mann-Whitney U statistic that is later calculated in the algorithm and used as a measure to differentiate between matches and non-matches works under the assumption of independence of parameters.
+\citet{hadler} in their paper proposed an improvement to this algorithm
+by trying to remove mutual dependence of parameters (due to serial
+correlation in surface depth values of a toolmark and because of a
+random sampling sub step in the validation phase which makes a group of
+pixels to be chosen more than once and hence introduces lack of
+independence) in certain steps, especially because the Mann-Whitney U
+statistic that is later calculated in the algorithm and used as a
+measure to differentiate between matches and non-matches works under the
+assumption of independence of parameters.
 
-Since we are only interested in a non-parameteric U statistic, Hadler et al.
-proposed a normalization procedure in the Validation step that goes to some extent to address the issue of mutual dependence.
-Also the same shift and different shift substep was modified to use a deterministic rule for sampling sam-shhift and dofferent shift-samples as opposed to the originally proposed random samples.
+Since we are only interested in a non-parameteric U statistic, Hadler et
+al. proposed a normalization procedure in the Validation step that goes
+to some extent to address the issue of mutual dependence. Also the same
+shift and different shift substep was modified to use a deterministic
+rule for sampling sam-shhift and dofferent shift-samples as opposed to
+the originally proposed random samples.
 
-The data that is used by \citet{chumbley} is generated by a surface profilometer that gives the height in terms of distance along a linear trace.
-This is taken perpendicular to the striations present in the toolmark.
-Two such trace are then compared using the algorithm.
+The data that is used by \citet{chumbley} is generated by a surface
+profilometer that gives the height in terms of distance along a linear
+trace. This is taken perpendicular to the striations present in the
+toolmark. Two such trace are then compared using the algorithm.
 
-## Detailed algorithm
+\subsection{Chumbley Algorithm and Chumbley
+score}\label{chumbley-algorithm-and-chumbley-score}
 
-###Optimization step
-The idea behind this step is to first identify the area of best agreement in the two toolmark data.
-Comparison window size is predefined by the user, which in case of screwdriver toolmarks was chosen by Chumbley et al and Hadler et al as around 10 percent of the length of the toolmarks, which was around 500.
-This window is henceforth referred to as Window of optimization.
+\subsubsection{Optimization step}\label{optimization-step}
 
-A maximum correlation statistic is used to identify the region of best agreement, with the maximum usually seen being near 1 for both cases which is what we intuitively expect for matches, and something which we do not intuitively expect for non matches \gk{done}\hh{XXX get rid of the parentheses - put the content into a sentence after your first statement.}
+The idea behind this step is to first identify the area of best
+agreement in the two toolmark data. Comparison window size is predefined
+by the user, which in case of screwdriver toolmarks was chosen by
+Chumbley et al and Hadler et al as around 10 percent of the length of
+the toolmarks, which was around 500. This window is henceforth referred
+to as Window of optimization.
+
+A maximum correlation statistic is used to identify the region of best
+agreement, with the maximum usually seen being near 1 for both cases
+which is what we intuitively expect for matches, and something which we
+do not intuitively expect for non matches
+\gk{done}\hh{XXX get rid of the parentheses - put the content into a sentence after your first statement.}
 
@@
 
-First, there are a very large number of cross-correlations calculated for the two series of striae or marks $x(t_1)$ and $y(t_2)$, for eg if the window of optimization was defined as 200 and toolmark pixel length is around 1000 then we have 1200-(200-1) = 1001.
-Can be thought of ordered arrangements without replacement where n = 1001 and r = 1.
+First, there are a very large number of cross-correlations calculated
+for the two series of striae or marks \(x(t_1)\) and \(y(t_2)\), for eg
+if the window of optimization was defined as 200 and toolmark pixel
+length is around 1000 then we have 1200-(200-1) = 1001. Can be thought
+of ordered arrangements without replacement where n = 1001 and r = 1.
 
-This is the number of windows we have for one toolmark and each window is compared with all windows of the second toolmark (the number of windows is again similar), and the window with the maximum correlation is identified to be the region where the toolmarks are in maximum agreement with each other
+This is the number of windows we have for one toolmark and each window
+is compared with all windows of the second toolmark (the number of
+windows is again similar), and the window with the maximum correlation
+is identified to be the region where the toolmarks are in maximum
+agreement with each other
 
-###Validation Step
+\subsubsection{Validation Step}\label{validation-step}
 
-####Same-shift: 
-In this step a series of windows are chosen at random (originally by chumbley et al) and deterministically (by hadler et al), but at a common distance (rigid-shift) from the window identified as the region of best agreement in the vaildation step.
-The correlation of these windows would be as intuitively assumed i.e.
-lower than the maximum correlation window (Optimization step), but the significance is that these same shift windows will still have large enough correlation values for two toolmarks or signatures that are in reality a match.
+\paragraph{Same-shift:}\label{same-shift}
 
-This also validates that if in the optimization step the maximum correlation window chosen (with correlation value near 1) was by accident (like in case of signatures that in reality are a non-match), all these same- shift correlations (A fixed number of these trace segments are identified) would not be anywhere large enough for all same-shift windows.
+In this step a series of windows are chosen at random (originally by
+chumbley et al) and deterministically (by hadler et al), but at a common
+distance (rigid-shift) from the window identified as the region of best
+agreement in the vaildation step. The correlation of these windows would
+be as intuitively assumed i.e.~lower than the maximum correlation window
+(Optimization step), but the significance is that these same shift
+windows will still have large enough correlation values for two
+toolmarks or signatures that are in reality a match.
 
-```{r win-comparison, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap=" The two plots on the left show how the same shift behaves in case of a matching pair and the two plots on the right show how the different shift behaves in case of a matching pair."}
-a<- matrix(NA, ncol = 3, nrow = 2)
-toolmark.falsepositive<- data.frame(a)
-colnames(toolmark.falsepositive)<- c("Classification", "Match", "Non-Match")
-toolmark.falsepositive$Classification<- c("Match", "Non-Match")
-toolmark.falsepositive$Match<- c(47,0)
-toolmark.falsepositive$`Non-Match`<- c(3,50)
-toolmark.falsepositive
+This also validates that if in the optimization step the maximum
+correlation window chosen (with correlation value near 1) was by
+accident (like in case of signatures that in reality are a non-match),
+all these same- shift correlations (A fixed number of these trace
+segments are identified) would not be anywhere large enough for all
+same-shift windows.
 
-asd<- all %>% filter(wo==120, wv==50, match == TRUE, land1_id == 1)
+\begin{verbatim}
+  Classification Match Non-Match
+1          Match    47         3
+2      Non-Match     0        50
+\end{verbatim}
 
-sigs_graphics<- readRDS("../data/sigs_generate_window.rds")
+\begin{figure}
 
-sigs_graphics.match<- as.data.frame(sigs_graphics[1])
-sigs_graphics.nonmatch<- as.data.frame(sigs_graphics[2])
-colnames(sigs_graphics.match)<- gsub("^.*\\.","", colnames(sigs_graphics.match))
-colnames(sigs_graphics.nonmatch)<- gsub("^.*\\.","", colnames(sigs_graphics.nonmatch))
+{\centering \includegraphics[width=\textwidth]{figures/win-comparison-1}
 
-# Matching signature
-#gi<-
-#sigs_graphics.match<- sigs_graphics.match %>% filter(land_id == c(1,9))  
-#sigs_graphics.match<- sigs_graphics.match %>% select(y, l30, land_id)
-d1<- sigs_graphics.match[which(sigs_graphics.match$land_id==1,arr.ind = TRUE),]
-d2<- sigs_graphics.match[which(sigs_graphics.match$land_id==19,arr.ind = TRUE),]
-data1<-d1 %>%select(l30)
-#data1<- data1$y
-data2<- d2 %>%select(l30)
-#data2<- data2$y
-window_opt = 120 
-window_val = 50
-coarse = 1
-data1<- matrix(unlist(data1))
-data2<- matrix(unlist(data2))
-
-  unity <- function(x) {x / sqrt(sum(x^2))} ## normalize columns of a matrix to make correlation computation faster
-
-  ####################################################
-  ##Clean the marks and compute the smooth residuals##
-  ####################################################
-
-  data1 <- matrix(data1[round((0.01*nrow(data1))):round(0.99*nrow(data1)),], ncol = 1)
-  data2 <- matrix(data2[round((0.01*nrow(data2))):round(0.99*nrow(data2)),], ncol = 1)
-
-  ##Normalize the tool marks
-  y1 <- data1 - lowess(y = data1,  x = 1:nrow(data1), f= coarse)$y
-  y2 <- data2 - lowess(y = data2,  x = 1:nrow(data2), f= coarse)$y
-
-  ############################################
-  ##Compute the observed maximum correlation##
-  ############################################
-
-  #####################
-  ##Optimization step##
-  #####################
-  ##Each column in these matrices corresponds to a window in the respective tool mark
-  y1_mat_opt <- matrix(NA, ncol = length(1:(length(y1) - (window_opt - 1))), nrow = window_opt)
-  for(l in 1:(length(y1) - (window_opt - 1))){
-    y1_mat_opt[,l] <- y1[l:(l+(window_opt - 1))]
-  }
-  y2_mat_opt <- matrix(NA, ncol = length(1:(length(y2) - (window_opt - 1))), nrow = window_opt)
-  for(l in 1:(length(y2) - (window_opt - 1))){
-    y2_mat_opt[,l] <- y2[l:(l+(window_opt - 1))]
 }
 
-  ##Compute the correlation between all pairs of windows for the two marks
-  ##Rows in the following matrix are mark 2, columns are mark 1
-  y2_mat_opt <- apply(scale(y2_mat_opt), 2, unity)
-  y1_mat_opt <- apply(scale(y1_mat_opt), 2, unity)
-  corr_mat_opt <- t(y2_mat_opt) %*% y1_mat_opt ##correlation matrix
-  max_corr_opt_loc <- which(corr_mat_opt == max(corr_mat_opt), arr.ind = TRUE) ##pair of windows maximizing the correlation
+\caption{ The two plots on the left show how the same shift behaves in case of a matching pair and the two plots on the right show how the different shift behaves in case of a matching pair.}\label{fig:win-comparison}
+\end{figure}
 
-  s1.rect.opt <- data.frame(xmin=max_corr_opt_loc[1,2], xmax= max_corr_opt_loc[1,2]+window_opt, ymin=-Inf, ymax=Inf)
-    s1.rect.sameshift1 <- data.frame(xmin=max_corr_opt_loc[1,2]-window_val, xmax= max_corr_opt_loc[1,2]- 2*window_val, ymin=-Inf, ymax=Inf)
-       s1.rect.sameshift2 <- data.frame(xmin=max_corr_opt_loc[1,2]+window_opt+ window_val, xmax= max_corr_opt_loc[1,2]+ window_opt+2*window_val, ymin=-Inf, ymax=Inf)
+\paragraph{Different Shift:}\label{different-shift}
 
-  s2.rect.opt <- data.frame(xmin=max_corr_opt_loc[1,1], xmax= max_corr_opt_loc[1,1]-window_opt, ymin=-Inf, ymax=Inf)
-    s2.rect.sameshift1 <- data.frame(xmin=max_corr_opt_loc[1,1]-window_opt-window_val, xmax= max_corr_opt_loc[1,1]-window_opt- 2*window_val, ymin=-Inf, ymax=Inf)
-       s2.rect.sameshift2 <- data.frame(xmin=max_corr_opt_loc[1,1]+ window_val, xmax= max_corr_opt_loc[1,1]+ 2*window_val, ymin=-Inf, ymax=Inf)
+Primary reason for this substep is to give perspective to the
+correlation values of the same shift window correlation values.
 
-p1<-  ggplot(data = d1, aes(x = y, y = l30)) + geom_line() + 
-  geom_rect(data=s1.rect.opt, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "black",alpha=0.5, inherit.aes = FALSE)+
-    geom_rect(data=s1.rect.sameshift1, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "red",alpha=0.1, inherit.aes = FALSE)+
-    geom_rect(data=s1.rect.sameshift2, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "blue",alpha=0.1, inherit.aes = FALSE)+
-  theme_bw() + labs(y = "Signature Land Id = 1")
+This time there are no rigid-shifts but different shifts (distance from
+Window of Opt with max correlation) chosen randomly by \citet{chumbley}
+and deterministically by Hadler et al. such that there is an equal
+possibilty of comparing a trace segment from one signature or toolmark
+to any one in the second signature or toolamrk.
 
- p2<-   ggplot(data = d2, aes(x = y, y = l30)) + geom_line() + 
-   geom_rect(data=s2.rect.opt, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "black",alpha=0.5, inherit.aes = FALSE)+
-    geom_rect(data=s2.rect.sameshift1, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "red",alpha=0.1, inherit.aes = FALSE)+
-    geom_rect(data=s2.rect.sameshift2, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "blue",alpha=0.1, inherit.aes = FALSE)+
-  theme_bw() + labs(y = "Signature Land Id = 9")
- ##########
+Neither of above sets of correlation are allowed to include the maximum
+correlation window as identified earlier.
 
- d1.rect.opt <- data.frame(xmin=max_corr_opt_loc[1,2], xmax= max_corr_opt_loc[1,2]+window_opt, ymin=-Inf, ymax=Inf)
-    d1.rect.sameshift1 <- data.frame(xmin=max_corr_opt_loc[1,2]-2*window_opt, xmax= max_corr_opt_loc[1,2]-2*window_opt-window_val, ymin=-Inf, ymax=Inf)
-       d1.rect.sameshift2 <- data.frame(xmin=max_corr_opt_loc[1,2]+window_opt+ window_val, xmax= max_corr_opt_loc[1,2]+ window_opt+2*window_val, ymin=-Inf, ymax=Inf)
+Therefore the assumption is that if two toolmarks or signatures match
+each other the same-shift correlations would be larger than the
+different-shift windows, and if they are not a match the correlations in
+the two sets will be very similar.
 
-  d2.rect.opt <- data.frame(xmin=max_corr_opt_loc[1,1], xmax= max_corr_opt_loc[1,1]-window_opt, ymin=-Inf, ymax=Inf)
-    d2.rect.sameshift1 <- data.frame(xmin=max_corr_opt_loc[1,1]-window_opt-window_val, xmax= max_corr_opt_loc[1,1]-window_opt- 2*window_val, ymin=-Inf, ymax=Inf)
-       d2.rect.sameshift2 <- data.frame(xmin=max_corr_opt_loc[1,1]+ 4*window_val, xmax= max_corr_opt_loc[1,1]+ 5*window_val, ymin=-Inf, ymax=Inf)
+\subsubsection{U Statistic:}\label{u-statistic}
 
-p3<-  ggplot(data = d1, aes(x = y, y = l30)) + geom_line() + 
-  geom_rect(data=d1.rect.opt, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "black",alpha=0.5, inherit.aes = FALSE)+
-    geom_rect(data=d1.rect.sameshift1, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "red",alpha=0.1, inherit.aes = FALSE)+
-    geom_rect(data=d1.rect.sameshift2, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "blue",alpha=0.1, inherit.aes = FALSE)+
-  theme_bw() + labs(y = "Signature Land Id = 1")
+This is computed from the joint rank of all correlations of both the
+same and different shift samples. As given by \citet{hadler}
 
- p4<-   ggplot(data = d2, aes(x = y, y = l30)) + geom_line() + 
-   geom_rect(data=d2.rect.opt, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "black",alpha=0.5, inherit.aes = FALSE)+
-    geom_rect(data=d2.rect.sameshift1, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "red",alpha=0.1, inherit.aes = FALSE)+
-    geom_rect(data=d2.rect.sameshift2, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "blue",alpha=0.1, inherit.aes = FALSE)+
-  theme_bw() + labs(y = "Signature Land Id = 9")
-
-multiplot(p1, p2, p3, p4, cols=2)
-```
-
-####Different Shift:
-Primary reason for this substep is to give perspective to the correlation values of the same shift window correlation values.
-
-This time there are no rigid-shifts but different shifts (distance from Window of Opt with max correlation) chosen randomly by \citet{chumbley} and deterministically by Hadler et al.
-such that there is an equal possibilty of comparing a trace segment from one signature or toolmark to any one in the second signature or toolamrk.
-
-Neither of above sets of correlation are allowed to include the maximum correlation window as identified earlier.
-
-Therefore the assumption is that if two toolmarks or signatures match each other the same-shift correlations would be larger than the different-shift windows,
-and if they are not a match the correlations in the two sets will be very similar.
-
-###U Statistic:
-
-This is computed from the joint rank of all correlations of both the same and different shift samples. As given by \citet{hadler}
-
-Null Hypothesis: If the toolmarks were not match i.e not made by the same tool.
+Null Hypothesis: If the toolmarks were not match i.e not made by the
+same tool.
 
 Let ns and nd be the number of same shift and different shift windows
-$N = n_{s} + n_{d}$
+\(N = n_{s} + n_{d}\)
 
 The mann whitney U statistic is given by
-$U =\sum^{ns}_{i=1}R_{s}\left( i\right)$
+\(U =\sum^{ns}_{i=1}R_{s}\left( i\right)\)
 
@@
 
-$\overline{U}= \dfrac{U-M}{\sqrt{V}}$
+\(\overline{U}= \dfrac{U-M}{\sqrt{V}}\)
 
@@
 
-$M = n_{s}\left(\dfrac{N+1}{2}\right)$
+\(M = n_{s}\left(\dfrac{N+1}{2}\right)\)
 
@@
 
-$V = \dfrac{n_{s}n_{d}}{N\left(N-1\right)}\left[\Sigma^{n_{s}}R_{s}\left(i\right)^{2}+\Sigma^{n_{d}}R_{d}\left(j\right)^{2}\right] -\dfrac{n_{s} n_{d}\left(N+1\right)^{2}}{4\left(N-1\right)}$
+\(V = \dfrac{n_{s}n_{d}}{N\left(N-1\right)}\left[\Sigma^{n_{s}}R_{s}\left(i\right)^{2}+\Sigma^{n_{d}}R_{d}\left(j\right)^{2}\right] -\dfrac{n_{s} n_{d}\left(N+1\right)^{2}}{4\left(N-1\right)}\)
 
-## Potential limitations of the Chumbley Score
+\subsection{\texorpdfstring{Potential limitations in the application to
+striation marks on bullet lands: smaller in width and curved - need to
+adjust parameter settings; something similar is done in
+\citet{afte-chumbley} for toolmark comparisons of slip-joint
+pliers.}{Potential limitations in the application to striation marks on bullet lands: smaller in width and curved - need to adjust parameter settings; something similar is done in  for toolmark comparisons of slip-joint pliers.}}\label{potential-limitations-in-the-application-to-striation-marks-on-bullet-lands-smaller-in-width-and-curved---need-to-adjust-parameter-settings-something-similar-is-done-in-for-toolmark-comparisons-of-slip-joint-pliers.}
 
-\hh{in the application to striation marks on bullet lands: smaller in width and curved - need to adjust parameter settings; something similar is done in \citet{afte-chumbley} for toolmark comparisons of slip-joint pliers.
-XXX work in }
+Bullets are much smaller in length, width, are not flat and curved in
+the cross-sectional topography as opposed to tools like screw driver
+tips which produced longer and pronounced markings. This means the
+makings made by barrels in comparison to toolmarks may have a problem in
+distinctiveness. The majority of Bullet profiles and signatures
+extracted by procedures mentioned by \citet{aoas} are almost 1/4 th the
+size of toolmarks as used by \citet{chumbley} or even smaller.
+Striations on Bullets are made on their curved surfaces, whereas the
+algorithm developed by \citet{chumbley} and \citet{hadler} has only been
+tested for flatter and wider surfaces which have negligible curvature.
+Therefore, using methods proposed for toolmarks may need adaptation in
+order to give tangible results for bullets. Moreover, in order to to get
+flat bullet signatures and remove the curvatures some kind of smoothing
+needs to be applied as a pre-step which needs further investigation as
+to whether the level of smoothing does effect the working of the
+algorithm on Bullets.
 
-Bullets are much smaller in length, width, are not flat and curved in the cross-sectional topography as opposed to tools like screw driver tips which produced longer and pronounced markings.
-This means the makings made by barrels in comparison to toolmarks may have a problem in distinctiveness.
-The majority of Bullet profiles and signatures extracted by procedures mentioned by  \citet{aoas} are almost 1/4 th the size of toolmarks as used by \citet{chumbley} or even smaller.
-Striations on Bullets are made on their curved surfaces, whereas the algorithm developed by \citet{chumbley} and \citet{hadler} has only been tested for flatter and wider surfaces which have negligible curvature.
-Therefore, using methods proposed for toolmarks may need adaptation in order to give tangible results for bullets.
-Moreover, in order to to get flat bullet signatures and remove the curvatures some kind of smoothing needs to be applied as a pre-step which needs further investigation as to whether the level of smoothing does effect the working of the algorithm on Bullets.
+Also when in the optimization step, the Window of optimization for
+bullets will be shorter as the signatures are smaller. The idea is to
+keep the number of windows of optimization sufficiently large, which
+means shorter Trace segments (partition of signature or toolmark with
+length = size of window of optimization) that lets us compare smaller
+segments of one signature to another.
 
-Also when in the optimization step, the Window of optimization for bullets will be shorter as the signatures are smaller.
-The idea is to keep the number of windows of optimization sufficiently large, which means shorter Trace segments (partition of signature or toolmark with length = size of window of optimization) that lets us compare smaller segments of one signature to another.
-
-This introduces a problem as, if we go too small in the window of optimization, the unique features of the trace segments are lost and seem similar, while too large sizes vastly reduces the weight of small features that would otherwise uniquely classify a signature and hence identify the region of agreement.
+This introduces a problem as, if we go too small in the window of
+optimization, the unique features of the trace segments are lost and
+seem similar, while too large sizes vastly reduces the weight of small
+features that would otherwise uniquely classify a signature and hence
+identify the region of agreement.
 
 Thus the Window of Optimization has a direct influence on whether we are
-Falsely rejecting the null when it is true (Type I) or Falsely accepting the null when it is false (Type II) making identification of the optimum size of window of optimization very important.
+Falsely rejecting the null when it is true (Type I) or Falsely accepting
+the null when it is false (Type II) making identification of the optimum
+size of window of optimization very important.
 
-## Testing setup:
-\hh{Using cross-validation setup to identify appropriate parameter settings for (a) signatures and (b) profiles directly}
+\subsection{Using cross-validation setup to identify appropriate
+parameter settings for (a) signatures and (b) profiles
+directly}\label{using-cross-validation-setup-to-identify-appropriate-parameter-settings-for-a-signatures-and-b-profiles-directly}
 
-Following on similar lines to the setup of toolmarks, the first step here is to first identify what difference does different window sizes of optimization and the validation step have, when adapting the toolmark method to bullets.
+Following on similar lines to the setup of toolamrks, the first step
+here is to first identify what difference does different window sizes of
+optimization and the validation step have, when adapting the toolmark
+method to bullets.
 
-The marking made on bullets are smaller than toolmarks and is also less wider. The idea is to find out possible areas of error while adapting the score based method proposed for toolmarks.
+The marking made on bullets are smaller than toolmarks and is also less
+wider. The idea is to find out possible areas of error while adapting
+the score based method proposed for toolmarks.
 
-Bullet signatures being compared at this time are from the Hamby 44 and Hamby 252 data, which have the best set of known-matched and known non-matches.
+Bullet signatures being compared at this time are from the Hamby 44 and
+Hamby 252 data, which have the best set of known-matched and known
+non-matches.
 
-Bullet signatures are chosen by
-1. Filtering out Land_id for Profiles from the Hamby 44 and Hamby 252 data and removing 
-all NA values
-2. run_id = 3 is chosen, Seems like the signatures generated from this run_id give the closest
-match. Different run_id's have some different settings for generating the signatures
-(The level of smoothing does not seem to be one of them)
+Bullet signatures are chosen by 1. Filtering out Land\_id for Profiles
+from the Hamby 44 and Hamby 252 data and removing all NA values 2.
+run\_id = 3 is chosen, Seems like the signatures generated from this
+run\_id give the closest match. Different run\_id's have some different
+settings for generating the signatures (The level of smoothing does not
+seem to be one of them)
 
-The bullet signatures when generated already included the LOWESS smoothing.
-Therefore, the coarseness factor is set to 1 while running the chumbley_non_random()
-which generates the same_shift, different_shift, U-Stat and P_value parameters.
+The bullet signatures when generated already included the LOWESS
+smoothing. Therefore, the coarseness factor is set to 1 while running
+the chumbley\_non\_random() which generates the same\_shift,
+different\_shift, U-Stat and P\_value parameters.
 
 window Optimization 320 Window Validation 50
-```{r echo=FALSE}
-all %>% filter(wo==320, wv==50) %>% xtabs(data=., ~signif+match)
-```
+
+\begin{verbatim}
+       match
+signif  FALSE  TRUE
+  FALSE 77564   412
+  TRUE   4022   674
+\end{verbatim}
+
 window Optimization 120 Window Validation 50
-```{r echo=FALSE}
-all %>% filter(wo==120, wv==50) %>% xtabs(data=., ~signif+match)
-```
+
+\begin{verbatim}
+       match
+signif  FALSE  TRUE
+  FALSE 79249   386
+  TRUE   4523   854
+\end{verbatim}
+
 window Optimization 80 Window Validation 50
-```{r echo=FALSE}
-all %>% filter(wo==80, wv==50) %>% xtabs(data=., ~signif+match)
 
-```
+\begin{verbatim}
+       match
+signif  FALSE  TRUE
+  FALSE 78817   450
+  TRUE   4680   775
+\end{verbatim}
 
-# Results
+\section{Results}\label{results}
 
-## Signatures
+\subsection{Signatures}\label{signatures}
 
-Signatures of lands for all Hamby-44 and Hamby-252 scans made available through the NIST ballistics database \citep{nist} were considered.
-Both of these sets of scans are part of the larger Hamby study \citep{hamby} and each consist of twenty known bullets (two each from ten consecutively rifled Ruger P85 barrels) and fifteen questioned bullets (each matching one of the ten barrels).
-Ground truth for both of these Hamby sets is known and was used to assess correctness of the tests results.
+Signatures of lands for all Hamby-44 and Hamby-252 scans made available
+through the NIST ballistics database \citep{nist} were considered. Both
+of these sets of scans are part of the larger Hamby study \citep{hamby}
+and each consist of twenty known bullets (two each from ten
+consecutively rifled Ruger P85 barrels) and fifteen questioned bullets
+(each matching one of the ten barrels). Ground truth for both of these
+Hamby sets is known and was used to assess correctness of the tests
+results.
 
-We used the adjusted Chumbley method as proposed in \citet{hadler} and implemented in the R package `toolmaRk` \citep{toolmark} on all pairwise land-to-land comparisons of the Hamby scans (a total of 85,491 comparisons) in the following manner:
+We used the adjusted Chumbley method as proposed in \citet{hadler} and
+implemented in the R package \texttt{toolmaRk} \citep{toolmark} on all
+pairwise land-to-land comparisons of the Hamby scans (a total of 85,491
+comparisons) in the following manner:
 
-```{r, echo=FALSE, results='asis'}
-param_setting<- all %>% select(wo, wv) %>% group_by(wo, wv) %>% tally() %>% select(wo,wv)
-#param_setting<- data.frame(matrix(param_setting, nrow = 16))
-#t(param_setting)
-#param_setting %>% tally() 
-library(kableExtra)
-knitr::kable(t(param_setting), digits=0, caption = "Overview of parameter settings used for optimization and validation windows for bullet land signatures.", format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("scale_down"))
-```
+\begin{table}
 
-### Signatures
+\caption{\label{tab:unnamed-chunk-5}Overview of parameter settings used for optimization and validation windows for bullet land signatures.}
+\centering
+\resizebox{\linewidth}{!}{\begin{tabular}[t]{lrrrrrrrrrrrrrrrrrrrrrrrrrrrrrrrrr}
+\toprule
+wo & 50 & 50 & 60 & 60 & 80 & 80 & 90 & 90 & 100 & 100 & 110 & 110 & 120 & 120 & 120 & 120 & 120 & 130 & 130 & 140 & 140 & 150 & 150 & 160 & 160 & 200 & 200 & 200 & 240 & 240 & 280 & 280 & 320\\
+wv & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 10 & 20 & 30 & 50 & 60 & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 20 & 30 & 50 & 30 & 50 & 30 & 50 & 50\\
+\bottomrule
+\end{tabular}}
+\end{table}
 
-Figure \ref{fig:type2} gives an overview of type II error rates observed when varying the window size in the optimization step.
+\subsubsection{Signatures}\label{signatures-1}
 
-```{r type2, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap="Type II error rates observed across a range of window sizes for optimization $wo$.
+Figure \ref{fig:type2} gives an overview of type II error rates observed
+when varying the window size in the optimization step.
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figures/type2-1}
+
+}
+
+\caption{Type II error rates observed across a range of window sizes for optimization $wo$.
 For a window size of $wo = 120$ we see a drop in type II error rate across all type I rates considered.
-Smaller validation sizes $wv$ are typically associated with a smaller type II error."}
-errors %>% filter(wv %in% c(30,50)) %>%
-  ggplot(aes(x = wo, y = beta, colour=factor(alpha))) + 
-  geom_point(aes(shape=factor(wv))) +
-#  facet_grid(.~wo, labeller="label_both") + 
-  geom_smooth(aes(group=alpha), se=FALSE, method="loess") +
-  theme_bw() +
-  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") +
-  scale_shape_discrete("Size of validation\nwindow, wv") +
-  xlab(expression("Window size for optimization, "~wo)) +
-  ylab("Type II error rate")
+Smaller validation sizes $wv$ are typically associated with a smaller type II error.}\label{fig:type2}
+\end{figure}
 
-labels = expression(P[M1](tilde(z)>0),P[M0](tilde(z)>0))
-```
+Figure \ref{fig:type1} compares nominal (fixed) type I error and
+actually observed type I errors for the parameter settings in table XXX.
+With an increasing size of the window used in the optimization step the
+observed type I error rate decreases (slighty). This might be related to
+the increasing number of tests that fail for larger window sizes, in
+particular for non-matching striae (see fig \ref{fig:missings}).
 
-Figure \ref{fig:type1} compares nominal (fixed) type I error and actually observed type I errors for the parameter settings in table XXX.
-With an increasing size of the window used in the optimization step the observed type I error rate decreases (slighty).
-This might be related to the increasing number of tests that fail for larger window sizes, in particular for non-matching striae (see fig \ref{fig:missings}).
+\begin{figure}
 
-```{r type1, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap="Comparison of observed and nominal type I error rates  across a range of window sizes for optimization $wo$.
+{\centering \includegraphics[width=\textwidth]{figures/type1-1}
+
+}
+
+\caption{Comparison of observed and nominal type I error rates  across a range of window sizes for optimization $wo$.
 The horizontal line in each facet indicates the nominal type I error rate.
 As the optimization window increase the observed type I error rate gets smaller.
-A smaller validation window tends to be associated with a higher type I error rate."}
-errors %>% filter(wv %in% c(30,50)) %>%
-  ggplot(aes(x = wo, y = actual, colour=factor(alpha))) + 
-  geom_hline(aes(yintercept=alpha), colour="grey30") +
-  geom_point(aes(shape=factor(wv))) +
-  #  facet_grid(.~wo, labeller="label_both") + 
-  geom_smooth(aes(group=alpha), se=FALSE, method="lm") +
-  theme_bw() +
-#  scale_y_log10(breaks=c(0.001,.005, 0.01, 0.05)) +
-  scale_shape_discrete("Size of validation window, wv") +
-  ylab("Observed type I error rate") +
-  xlab(expression("Window size for optimization, "~wo)) +
-  scale_colour_brewer(expression("Nominal type I error "~alpha), palette="Set2") +
-  facet_wrap(~alpha, labeller="label_both", scales="free") +
-  theme(legend.position="bottom")
+A smaller validation window tends to be associated with a higher type I error rate.}\label{fig:type1}
+\end{figure}
 
-```
+\begin{figure}
 
-```{r missings, fig.width=8, fig.height=5, fig.cap="Number of failed tests by the window optimization size, wo, and ground truth.
-Test results from different sources have a much higher chance to fail, raising the question, whether failed tests should be treated as rejections of the null hypothesis of same source."}
-ns <- all %>% group_by(wv, wo, match) %>% summarize(
-  n_ss = mean(same_shift_n),
-  n_ds = mean(diff_shift_n),
-  zero_ss = sum(same_shift_n==0),
-  zero_ds = sum(diff_shift_n==0),
-  miss = sum(is.na(p_value)),
-  missperc = miss/85491*100
-)
-ns %>% filter(wv %in% c(30, 50)) %>% 
-  ggplot(aes(x = wo, y = missperc, shape=factor(wv), colour = factor(match))) + 
-  geom_smooth(method="lm", size=.5, se=FALSE) +
-  geom_point() + 
-  theme_bw() +
-  ylab("Percent of failed tests.") +
-  xlab("Size of optimization window, wo") +
-  scale_colour_brewer("Match", palette="Set1") +
-  scale_shape_discrete("Size of \nvalidation\nwindow, wv")
-```
-Figure \ref{fig:missings} gives an overview of the number of failed tests, i.e.
-tests in which a particular parameter setting did not return a valid result.
-This happens, when the shift to align two signatures is so large, that the remaining overlap is too small to accommodate  windows for validation.
-The problem is therefore exacerbated by a larger validation window.
-Figure \ref{fig:missings} also shows that the number of failed tests is approximately linear in the size of the optimization window.
+{\centering \includegraphics[width=\textwidth]{figures/missings-1}
 
-```{r lms, results='asis'}
-nsnest <- ns %>% filter(wv %in% c(30,50)) %>% group_by(match, wv) %>% nest()
-nsnest <- nsnest %>% mutate(
-  model = data %>% purrr::map(.f = function(d) lm(I(missperc*100)~wo, data=d)),
-  coefs = model %>% purrr::map(.f = broom::tidy)
-)
-tab <- nsnest %>% unnest(coefs) %>% filter(term=="wo") %>% arrange(match) %>%
-  select(-term, -statistic, -p.value)
-knitr::kable(tab, digits=3, caption = "Estimates of the increase in percent of failed tests corresponding to a 100 point increase in the optimization window.", booktabs=TRUE)
-```
+}
 
-```{r wo120, fig.width = 8, fig.height=4.5}
-wo120 <- errors %>% filter(wo==120)
-wo120 %>% gather(type, error, actual:beta) %>%
-  ggplot(aes(x = wv, y = error, colour = factor(alpha)))  + 
-  geom_line() + geom_point() + facet_wrap(~type, scales="free") +
-  theme_bw()+ 
-  theme(legend.position="bottom")
-```
+\caption{Number of failed tests by the window optimization size, wo, and ground truth.
+Test results from different sources have a much higher chance to fail, raising the question, whether failed tests should be treated as rejections of the null hypothesis of same source.}\label{fig:missings}
+\end{figure}
 
-### Profiles
+Figure \ref{fig:missings} gives an overview of the number of failed
+tests, i.e.~tests in which a particular parameter setting did not return
+a valid result. This happens, when the shift to align two signatures is
+so large, that the remaining overlap is too small to accommodate windows
+for validation. The problem is therefore exacerbated by a larger
+validation window. Figure \ref{fig:missings} also shows that the number
+of failed tests is approximately linear in the size of the optimization
+window.
 
-The profiles are cross-sectional values of the the bullet striation mark which are chosen at an optimum height (x as used by \citet{aoas}).
-This x or height is not a randomly chosen level.
-The rationale behind the choice has been explained by \citet{aoas}.
-A region is first chosen where the cross-correlation seems to change very less and in this region an optimum height is chosen.
-The profiles generally resemble a curve which is more or less similar to a quadratic curve (a quadratic fit to the raw data values of the profile is not an exact fit but it does show a similar trend).
-Profiles are the set of raw values representing the striation marks, and signatures are generated from these by removal of the inherent curvature and applying some smoothing (the signatures generated by \citet{aoas} use a loess function for smoothing).
+\begin{table}
 
-Coming back to the database of Hamby-44 and Hamby-252 datasets, the run_id = 3 was used when applying the chumbley algorithm on the signatures.
-The run_id not only defines the level of smoothing but also signifies the chosen height at which the profiles were selected initially.
-Another important aspect is the range of horizontal values (which is referred to as the y values in \citet{aoas}) in the signatures.
-Thes have already been pre-processed in the database to not include any grooves.
+\caption{\label{tab:lms}Estimates of the increase in percent of failed tests corresponding to a 100 point increase in the optimization window.}
+\centering
+\begin{tabular}[t]{lrrr}
+\toprule
+match & wv & estimate & std.error\\
+\midrule
+FALSE & 30 & 0.452 & 0.027\\
+FALSE & 50 & 0.922 & 0.035\\
+TRUE & 30 & 0.037 & 0.004\\
+TRUE & 50 & 0.056 & 0.005\\
+\bottomrule
+\end{tabular}
+\end{table}
 
-Therefore for the sake of comparison the run_id = 3 is still chosen so as to ensure that the horizontal values remain the same as that of the signatures.
-This also gives us profiles with the grooves removed.
+\begin{center}\includegraphics[width=\textwidth]{figures/wo120-1} \end{center}
 
-The idea therefore is to first use these raw values of the profile directly in the chumbley algorithm, and see how the algorithm performs for different coarseness values (smoothing parameter as referred in the function LOWESS used in the chumbley algorithm)
+\subsubsection{Profiles}\label{profiles}
 
-## Discussion
+The profiles are cross-sectional values of the the bullet striation mark
+which are chosen at an optimum height (x as used by \citet{aoas}). This
+x or height is not a randomly chosen level. The rationale behind the
+choice has been explained by \citet{aoas}. A region is first chosen
+where the cross-correlation seems to change very less and in this region
+an optimum height is chosen. The profiles generally resemble a curve
+which is more or less similar to a quadratic curve (a quadratic fit to
+the raw data values of the profile is not an exact fit but it does show
+a similar trend). Profiles are the set of raw values representing the
+striation marks, and signatures are generated from these by removal of
+the inherent curvature and applying some smoothing (the signatures
+generated by \citet{aoas} use a loess function for smoothing).
+
+Coming back to the database of Hamby-44 and Hamby-252 datasets, the
+run\_id = 3 was used when applying the chumbley algorithm on the
+signatures. The run\_id not only defines the level of smoothing but also
+signifies the chosen height at which the profiles were selected
+initially. Another important aspect is the range of horizontal values
+(which is referred to as the y values in \citet{aoas}) in the
+signatures. Thes have already been pre-processed in the database to not
+include any grooves.
+
+Therefore for the sake of comparison the run\_id = 3 is still chosen so
+as to ensure that the horizontal values remain the same as that of the
+signatures. This also gives us profiles with the grooves removed.
+
+The idea therefore is to first use these raw values of the profile
+directly in the chumbley algorithm, and see how the algorithm performs
+for different coarseness values (smoothing parameter as referred in the
+function LOWESS used in the chumbley algorithm)
+
+\bibliographystyle{agsm}
+\bibliography{bibliography}
+
+\end{document}

### 2017-12-26 code/algo_modification.R

@@ -1,0 +1,239 @@
+#' Modified Chumbley Non-Random
+#' 
+#' The modification is made in the same-shift, where for the 2nd signature 
+#' the window is moved a bit towards the left and right to find if the 
+#' correlation of the moved windows are higher than the original same shift (OSS) window
+#' If its higher the new correlation values are stored
+#' 
+#' This function computes the Chumbley U-Statistic on systemically chosen pairs of windows rather 
+#' than the original method which selects randomly chosen pairs of windows
+#' @param data1 The first tool mark as a 1-column matrix
+#' @param data2 The second tool mark as a 1-column matrix
+#' @param window_opt size of the window to be used in the optimization step
+#' @param window_val Size of the window to be used in the validation step
+#' @param coarse smoothing parameter for the normalization smooth
+#' @importFrom stats pnorm
+#' @importFrom ggplot2 ggplot
+#' @export
+#' @return list with
+#' \itemize{
+#' \item {same_shift_n} Number of same shift offsets used
+#' \item {diff_shift_n} Number of different shift offsets used
+#' \item {U} observed U statistic
+#' \item {p_value} Corresponding p-value
+#' }
+chumbley_non_random_modified <- function(data1, data2, window_opt = 500, window_val = 50, coarse = .25){
+  
+  unity <- function(x) {x / sqrt(sum(x^2))} ## normalize columns of a matrix to make correlation computation faster
+  
+  ####################################################
+  ##Clean the marks and compute the smooth residuals##
+  ####################################################
+  
+  data1 <- matrix(data1[round((0.01*nrow(data1))):round(0.99*nrow(data1)),], ncol = 1)
+  data2 <- matrix(data2[round((0.01*nrow(data2))):round(0.99*nrow(data2)),], ncol = 1)
+  
+  ##Normalize the tool marks
+  y1 <- data1 - lowess(y = data1,  x = 1:nrow(data1), f= coarse)$y
+  y2 <- data2 - lowess(y = data2,  x = 1:nrow(data2), f= coarse)$y
+  
+  
+  ############################################
+  ##Compute the observed maximum correlation##
+  ############################################
+  
[…]

### 2017-12-26 doc/paper.Rmd
small edits

@@
   files <- dir("../data/profiles/", pattern="csv")
-  all <- data.frame()
+  allprofiles <- data.frame()
   for (file in files) {
@@
     if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% select(-x)
-    all <- rbind(all, tmp)
+    allprofiles <- rbind(allprofiles, tmp)
   }
-  all <- all %>% filter(land1_id < land2_id)
-  all <- all %>% mutate(coarse = as.numeric(gsub("pt", ".", coarse)))
-  saveRDS(all, file="../data/all-profiles.rds")
+  allprofiles <- allprofiles %>% filter(land1_id < land2_id)
+  allprofiles <- allprofiles %>% mutate(coarse = as.numeric(gsub("pt", ".", coarse)))
+  saveRDS(allprofiles, file="../data/all-profiles.rds")
 } else {
-  all <- readRDS("../data/all-profiles.rds")
+  allprofiles <- readRDS("../data/all-profiles.rds")
 }
 
-errors2 <- rbind(errorrate(all, 0.001), 
-                errorrate(all, 0.005), 
-                errorrate(all, 0.01), 
-                errorrate(all, 0.05))
+errors2 <- rbind(errorrate(allprofiles, 0.001), 
+                errorrate(allprofiles, 0.005), 
+                errorrate(allprofiles, 0.01), 
+                errorrate(allprofiles, 0.05))
 ```
@@
 Since determining error rates has been duly noted as a fundamental problem in forensic science \citep{NAS:2009}, there is a need of methods that address this problem.
-Some of the methods that have been used to estimate error rates in firearm examinations include Automatic cartridge case comparison where same source identification of cartridge cases and associated error rates have been provided by \citep{riva}, although as noted by \citep{aoas}, in this case alignments of striae involves roation of planes, which cannot be generalized for bullets.
+Some of the methods that have been used to estimate error rates in firearm examinations include Automatic cartridge case comparison where same source identification of cartridge cases and associated error rates have been provided by \citep{riva}, although as noted by \citep{aoas}, in this case alignments of striae involves a rotation of planes, which cannot be generalized for bullets.
 In case of methods that use machine learning algorithms some methods which are bootsrap based methods often tend to give over-estimated error rates \citep{efron}, but there are alternate error estimation techniques as described by \citep{aoas}, \citep{efron} and  \citep{vorburger2016} which give better results.
+\hh{XXX give examples for these techniques}
 
@@
 An important aspect of same source matching in firearms is identification of bullet microtopographies that are unique to be chosen for a match, Bullet profiles and signatures prove to be such features.
-The choice of a bullet signature to uniquely define a bullet depends on finding first, a stable region on a Bullet land which minimal noise and many pronounced striation marks.
-\citet{aoas}
+The choice of a bullet signature to uniquely define a bullet depends on finding first, a stable region on a Bullet land with minimal noise and pronounced striation marks \citep{aoas}.
 
@@
 \includegraphics[width=\textwidth]{images/B6-B2-L6.png}
-```{r, echo=FALSE, fig.width=8, fig.height=3, out.width='\\textwidth'}
+```{r, echo=FALSE, fig.width=8, fig.height=3, fig.align="center", out.width='.9\\textwidth'}
 land <- read.csv("../data/b6b2l6.csv")
-land %>% ggplot(aes(x = y, y=resid)) +geom_line() +theme_bw()
+land %>% ggplot(aes(x = y, y=resid)) +geom_line() +theme_bw() +
+  theme(axis.text.y = element_blank(), axis.title.y = element_blank())
 ```
@@
 
-###Optimization step
+### Optimization step
 The idea behind this step is to first identify the area of best agreement in the two toolmark data.
-Comparison window size is predefined by the user, which in case of screwdriver toolmarks was chosen by Chumbley et al and Hadler et al as around 10 percent of the length of the toolmarks, which was around 500.
-This window is henceforth referred to as Window of optimization.
+Comparison window size is predefined by the user, which in case of screwdriver toolmarks was chosen by \citet{chumbley} and \citet{hadler} as around 10 percent of the length of the toolmarks, which was around 500.
+\hh{XXX it's not clear that the 500 is referring to the 10 percent.} This window is henceforth referred to as Window of optimization.
 
-A maximum correlation statistic is used to identify the region of best agreement, with the maximum usually seen being near 1 for both cases which is what we intuitively expect for matches, and something which we do not intuitively expect for non matches \gk{done}\hh{XXX get rid of the parentheses - put the content into a sentence after your first statement.}
+A maximum correlation statistic is used to identify the region of best agreement, with the maximum usually seen being near 1 for both cases which is what we intuitively expect for matches, and something which we do not intuitively expect for non matches \gk{done} \hh{XXX get rid of the parentheses - put the content into a sentence after your first statement.}
 
@@
 
-```{r win-comparison, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap=" The two plots on the left show how the same shift behaves in case of a matching pair and the two plots on the right show how the different shift behaves in case of a matching pair."}
+\hh{XXX Figure~\ref{win-comparison} is an orphan - cite it in the correct place.}
+\hh{XXX let's include two or three more windows.
+We won't be able to use color - so we need to think of  a different way to distinguish between optimization and validation and link corresponding validation windows.
+Also include the code in the folder on the repo that generates sigs\_generate\_windows.rds }
+
+```{r win-comparison, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap=" The two plots on the left show how the same shift behaves in case of a matching pair and the two plots on the right show how the different shift behaves in case of a matching pair.
+"}
 a<- matrix(NA, ncol = 3, nrow = 2)
@@
 
-## Testing setup:
+# Testing setup:
 \hh{Using cross-validation setup to identify appropriate parameter settings for (a) signatures and (b) profiles directly}
@@
 
-# Results
-
-## Signatures
+## Study setup
 
@@
 library(kableExtra)
-knitr::kable(t(param_setting), digits=0, caption = "Overview of parameter settings used for optimization and validation windows for bullet land signatures.", format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("scale_down"))
+knitr::kable(t(param_setting), digits=10, caption = "Overview of parameter settings used for optimization and validation windows for bullet land signatures.", format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("scale_down"))
 ```
+
+# Results
 
@@
 ## Discussion
+
+## Results of Chumbley score
+- Nominal alpha value shows dependency on the size of the window of optimization
+- Test Fail depends on whether known-match or known non-matches has predictive value
+- Type II error least bad for window of validation 30 and window of optimization 120
+
+# Algorithm Modification
+
+The proposed algorithm modification is at the same shift step which comes after the shift distance is identified by finding two windows in the two markings that have the highest correlation.
+
+The modification allows for a "wiggle" room in the second marking of the two sets of markings being compared.
+This means that each set of same shift windows that are being compared, the second marking will window that is under consideration is allowd to move a little towards the left and a little towards the right.
+
+This gives us a new set of 4 windows, 2 to the left and 2 to the right of the original comparison window.
+Then the correlations for each one of these 5 windows with the window under consideration of the 1st marking is retrieved (from the validation step correlation matrix) or computed.
+
+Then the window that has the maximum correlation (from the set of 5 windows in the 2nd marking) with the window of the 1st marking is chosen.
+
+This new maximum correlation is then used to compute the U-statistics using the same method as before.
+
+## Expected advantage of modification
+
+We are expecting that this modification would improve the type I and II errors for the good. Less number of false negatives and false positives.
+The reason for this expectation is many a times the bullet markings are not made at the same distances for two or more bullets because of the way it comes out of the barrel.
+
+Therefore rigid same shifts might not necessarily compare the right set of windows.
+In the same shift step for one set of windows, allowing the window in the 2nd marking to wiggle left and right and finding the best match to the window in the 1st marking, lets us adjust for situations where markings are compressed or elongated.
+
+## Dependence on the delta of new windows from the original same shift window
+
+From initial tests the amount of movement to get new windows left and right of the original same shift window seems to directly influence the
+the pvalue and U statistic that we look at.
+
+### other questions
+How much movement is reasonable and should be used?
+Is there a way to understand if the 2nd marking in comparison to the 1st marking is compressed or elongated or neither?
+Should the amount of movement selected depend on this compression or elongation?
+or using a static delta movement value justifieable.

### 2017-12-27 doc/paper.Rmd
information on algorithm

--- 2017-11-12 doc/paper.Rmd
@@
 ## Discussion
+
+## Results of Chumbley score
+- Nominal alpha value shows dependency on the size of the window of optimization
+- Test Fail depends on whether known-match or known non-matches has predictive value
+- Type II error least bad for window of validation 30 and window of optimization 120
+
+# Algorithm Modification
+
+The proposed algorithm modification is at the same shift step which comes after the shift distance is identified by finding two windows in the two markings that have the highest correlation.
+
+The modification allows for a "wiggle" room in the second marking of the two sets of markings being compared.
+This means that each set of same shift windows that are being compared, the second marking will window that is under consideration is allowd to move a little towards the left and a little towards the right.
+
+This gives us a new set of 4 windows, 2 to the left and 2 to the right of the original comparison window.
+Then the correlations for each one of these 5 windows with the window under consideration of the 1st marking is retrieved (from the validation step correlation matrix) or computed.
+
+Then the window that has the maximum correlation (from the set of 5 windows in the 2nd marking) with the window of the 1st marking is chosen.
+
+This new maximum correlation is then used to compute the U-statistics using the same method as before.
+
+## Expected advantage of modification
+
+We are expecting that this modification would improve the type I and II errors for the good. Less number of false negatives and false positives.
+The reason for this expectation is many a times the bullet markings are not made at the same distances for two or more bullets because of the way it comes out of the barrel.
+
+Therefore rigid same shifts might not necessarily compare the right set of windows.
+In the same shift step for one set of windows, allowing the window in the 2nd marking to wiggle left and right and finding the best match to the window in the 1st marking, lets us adjust for situations where markings are compressed or elongated.
+
+## Dependence on the delta of new windows from the original same shift window
+
+From initial tests the amount of movement to get new windows left and right of the original same shift window seems to directly influence the
+penultimate pvalue and U statistic that we look at.
+
+So the question that arises is how much of a movement is reasonable and should be used?
+Is there a way to understand if the 2nd marking in comparison to the 1st marking is compressed or elongated?
+Should the amount of movement selected depend on this compression or elongation?
+or using a static delta movement value justifieable.

### 2017-12-27 doc/paper.Rmd
.

@@
 From initial tests the amount of movement to get new windows left and right of the original same shift window seems to directly influence the
-penultimate pvalue and U statistic that we look at.
+the pvalue and U statistic that we look at.
 
-So the question that arises is how much of a movement is reasonable and should be used?
-Is there a way to understand if the 2nd marking in comparison to the 1st marking is compressed or elongated?
+### how much of a movement is reasonable and should be used?
+Is there a way to understand if the 2nd marking in comparison to the 1st marking is compressed or elongated or neither?
 Should the amount of movement selected depend on this compression or elongation?

### 2018-01-02 p2/p2.Rmd
.

--- 2017-12-27 doc/paper.Rmd
@@
   files <- dir("../data/profiles/", pattern="csv")
-  all <- data.frame()
+  allp <- data.frame()
   for (file in files) {
@@
     if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% select(-x)
-    all <- rbind(all, tmp)
+    allp <- rbind(allp, tmp)
   }
-  all <- all %>% filter(land1_id < land2_id)
-  all <- all %>% mutate(coarse = as.numeric(gsub("pt", ".", coarse)))
-  saveRDS(all, file="../data/all-profiles.rds")
+  allp <- allp %>% filter(land1_id < land2_id)
+  allp <- allp %>% mutate(coarse = as.numeric(gsub("pt", ".", coarse)))
+  saveRDS(allp, file="../data/all-profiles.rds")
 } else {
-  all <- readRDS("../data/all-profiles.rds")
+  allp <- readRDS("../data/all-profiles.rds")
 }
 
-errors2 <- rbind(errorrate(all, 0.001), 
-                errorrate(all, 0.005), 
-                errorrate(all, 0.01), 
-                errorrate(all, 0.05))
+errors2 <- rbind(errorrate(allp, 0.001), 
+                errorrate(allp, 0.005), 
+                errorrate(allp, 0.01), 
+                errorrate(allp, 0.05))
+
+```
+
+```{r data-sig-coars}
+if (!file.exists("../data/all-sig-c.rds")) {
+  files <- dir("../data/sig_diff_coarseness/", pattern="csv")
+  all_sig_c <- data.frame()
+  for (file in files) {
+    tmp <- read.csv(file=file.path(path="../data/sig_diff_coarseness/", file)) 
+    #print(names(tmp))
+    if (length(grep("coarse",names(tmp))) == 0)
+      tmp$coarse <- gsub(".*coarse-(.*).csv", "\\1", file)
+    if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% select(-x)
+    all_sig_c <- rbind(all_sig_c, tmp)
+  }
+  all_sig_c <- all_sig_c %>% filter(land1_id < land2_id)
+  all_sig_c <- all_sig_c %>% mutate(coarse = as.numeric(gsub("pt", ".", coarse)))
+  saveRDS(all_sig_c, file="../data/all-sig-c.rds")
+} else {
+  all_sig_c <- readRDS("../data/all-sig-c.rds")
+}
+
+errors_sig_c <- rbind(errorrate(all_sig_c, 0.001), 
+                errorrate(all_sig_c, 0.005), 
+                errorrate(all_sig_c, 0.01), 
+                errorrate(all_sig_c, 0.05))
+
 ```
@@
 
-window Optimization 320 Window Validation 50
-```{r echo=FALSE}
-all %>% filter(wo==320, wv==50) %>% xtabs(data=., ~signif+match)
+<!-- \begin{minipage}{0.3\textwidth} -->
+```{r }
+library(kableExtra)
+a27<- all %>% filter(wo==320, wv==50) %>% xtabs(data= ., ~signif+match)
+knitr::kable(a27, digits=0, caption = "Window of Optimization 320 and Window of Validation 50", format = "latex" ,booktabs=TRUE)%>% kableExtra::kable_styling(latex_options = c("hold"))
+
 ```
-window Optimization 120 Window Validation 50
-```{r echo=FALSE}
-all %>% filter(wo==120, wv==50) %>% xtabs(data=., ~signif+match)
+<!-- \end{minipage} -->
+<!-- \begin{minipage}{0.3\textwidth} -->
+```{r}
+a27<- all %>% filter(wo==120, wv==50) %>% xtabs(data= ., ~signif+match)
+knitr::kable(a27, digits=0, caption = "Window of Optimization 120 and Window of Validation 50", format = "latex" ,booktabs=TRUE)%>% kableExtra::kable_styling(latex_options = c("hold"))
 ```
-window Optimization 80 Window Validation 50
-```{r echo=FALSE}
-all %>% filter(wo==80, wv==50) %>% xtabs(data=., ~signif+match)
-
+<!-- \end{minipage} -->
+```{r}
+a27<- all %>% filter(wo==80, wv==50) %>% xtabs(data= ., ~signif+match)
+knitr::kable(a27, digits=0, caption = "Window of Optimization 80 and Window of Validation 50", format = "latex" ,booktabs=TRUE)%>% kableExtra::kable_styling(latex_options = c("hold"))
 ```
@@
 library(kableExtra)
-knitr::kable(t(param_setting), digits=0, caption = "Overview of parameter settings used for optimization and validation windows for bullet land signatures.", format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("scale_down"))
+knitr::kable(t(param_setting)[,1:17], digits=0, caption = "Overview of parameter settings used for optimization and validation windows for bullet land signatures.", format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("scale_down", "hold"))
+
+knitr::kable(t(param_setting)[,17:33], digits=0, format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("scale_down","hold"))
+
+# library(xtable)
+# strCaption <- "Overview of parameter settings used for optimization and validation windows for bullet land signatures."
+#   
+#   # set up xtable output
+# print(xtable(t(param_setting)[,1:17], 
+#              caption = strCaption, label = ""),
+#       include.rownames = TRUE,
+#       include.colnames = FALSE,
+#       caption.placement = "top",
+#       comment=FALSE,
+#       floating=FALSE
+#       )
+# print(xtable(t(param_setting)[,17:33],label = ""),
+#       include.rownames = FALSE,
+#       include.colnames = FALSE,
+#       caption.placement = "bottom",
+#       comment=FALSE,
+#       floating=TRUE
+#       #table.placement = "!ht"
+#       )
 ```
@@
 )
+
+#ns1<- ns[which(ns$wv == c(30,50), arr.ind = T),]
 ns %>% filter(wv %in% c(30, 50)) %>%
+#ns %>% filter(wv == c(30, 50)) %>%
   ggplot(aes(x = wo, y = missperc, shape=factor(wv), colour = factor(match))) + 
@@
 ```{r}
-errors %>% ggplot(aes(x = wv, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + facet_grid(~wo) +
-  geom_point(data=errors2, aes(shape="profiles"))
+#errors %>% ggplot(aes(x = wv, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + facet_grid(~wo) +geom_point(data=errors2, aes(shape="profiles"))
+
+errors_sig_c %>% ggplot(aes(x = coarse, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + geom_line() + geom_point(data=errors2, aes(shape="profiles")) +geom_line(data=errors2, aes(shape="profiles"))
 ```
@@
 
-### how much of a movement is reasonable and should be used?
+### other questions
+How much movement is reasonable and should be used?
 Is there a way to understand if the 2nd marking in comparison to the 1st marking is compressed or elongated or neither?

### 2018-01-02 p2/p2.tex
.

--- 2017-12-26 doc/paper.tex
@@
 
-\section{Introduction}\label{introduction}
+\tableofcontents
+
+\section{Introduction and Background}\label{introduction-and-background}
 
@@
 
-\subsection{Existing solution for toolmarks: Chumbley
-Score}\label{existing-solution-for-toolmarks-chumbley-score}
+\subsection{The Chumbley Score Test}\label{the-chumbley-score-test}
 
-\hh{XXX anything to do with motivation has to come first - move this to the first paragraph of the intro XXX}
-\hh{XXX data and results come after the intro to chumbley.}
-\hh{XXX start with defining the data, i.e.
-we are working with random processes $x(t)$ and $y(t)$ ...
-you should also go into the coarseness parameter here - that is a pre-processing step before the optimization.
-}
-\gk{done}
+\hh{You need to distinguish between $x(t)$ and $x(t_1)$ - $x(t)$ is a spatial process for some $t$.
+$x(t_1)$ is the realization of the process in $t_1$.
+- Alternatively you could also follow the notation in Hadler et al on 'digitized toolmarks at a pixel level' - where our 'pixel' is the resolution of the confocal light microscope, i.e.
+a pixel corresponds to 0.645 microns.}
 
@@
 
-\subsection{Chumbley Algorithm and Chumbley
-score}\label{chumbley-algorithm-and-chumbley-score}
+\subsection{Detailed algorithm}\label{detailed-algorithm}
 
@@
 
-\subsection{\texorpdfstring{Potential limitations in the application to
-striation marks on bullet lands: smaller in width and curved - need to
-adjust parameter settings; something similar is done in
-\citet{afte-chumbley} for toolmark comparisons of slip-joint
-pliers.}{Potential limitations in the application to striation marks on bullet lands: smaller in width and curved - need to adjust parameter settings; something similar is done in  for toolmark comparisons of slip-joint pliers.}}\label{potential-limitations-in-the-application-to-striation-marks-on-bullet-lands-smaller-in-width-and-curved---need-to-adjust-parameter-settings-something-similar-is-done-in-for-toolmark-comparisons-of-slip-joint-pliers.}
+\subsection{Potential limitations of the Chumbley
+Score}\label{potential-limitations-of-the-chumbley-score}
+
+\hh{in the application to striation marks on bullet lands: smaller in width and curved - need to adjust parameter settings; something similar is done in \citet{afte-chumbley} for toolmark comparisons of slip-joint pliers.
+XXX work in }
 
@@
 
-\subsection{Using cross-validation setup to identify appropriate
-parameter settings for (a) signatures and (b) profiles
-directly}\label{using-cross-validation-setup-to-identify-appropriate-parameter-settings-for-a-signatures-and-b-profiles-directly}
+\subsection{Testing setup:}\label{testing-setup}
 
-Following on similar lines to the setup of toolamrks, the first step
+\hh{Using cross-validation setup to identify appropriate parameter settings for (a) signatures and (b) profiles directly}
+
+Following on similar lines to the setup of toolmarks, the first step
 here is to first identify what difference does different window sizes of
@@
 
-window Optimization 320 Window Validation 50
+\begin{table}[!h]
 
-\begin{verbatim}
-       match
-signif  FALSE  TRUE
-  FALSE 77564   412
-  TRUE   4022   674
-\end{verbatim}
+\caption{\label{tab:unnamed-chunk-2}Window of Optimization 320 and Window of Validation 50}
+\centering
+\begin{tabular}[t]{lrr}
+\toprule
+  & FALSE & TRUE\\
+\midrule
+FALSE & 77564 & 412\\
+TRUE & 4022 & 674\\
+\bottomrule
+\end{tabular}
+\end{table}
 
-window Optimization 120 Window Validation 50
+\begin{table}[!h]
 
-\begin{verbatim}
-       match
-signif  FALSE  TRUE
-  FALSE 79249   386
-  TRUE   4523   854
-\end{verbatim}
+\caption{\label{tab:unnamed-chunk-3}Window of Optimization 120 and Window of Validation 50}
+\centering
+\begin{tabular}[t]{lrr}
+\toprule
+  & FALSE & TRUE\\
+\midrule
+FALSE & 78938 & 371\\
+TRUE & 4281 & 845\\
+\bottomrule
+\end{tabular}
+\end{table}
 
-window Optimization 80 Window Validation 50
+\begin{table}[!h]
 
-\begin{verbatim}
-       match
-signif  FALSE  TRUE
-  FALSE 78817   450
-  TRUE   4680   775
-\end{verbatim}
+\caption{\label{tab:unnamed-chunk-4}Window of Optimization 80 and Window of Validation 50}
+\centering
+\begin{tabular}[t]{lrr}
+\toprule
+  & FALSE & TRUE\\
+\midrule
+FALSE & 78817 & 450\\
+TRUE & 4680 & 775\\
+\bottomrule
+\end{tabular}
+\end{table}
 
@@
 
-\begin{table}
+\begin{table}[!h]
 
@@
 \centering
-\resizebox{\linewidth}{!}{\begin{tabular}[t]{lrrrrrrrrrrrrrrrrrrrrrrrrrrrrrrrrr}
+\resizebox{\linewidth}{!}{\begin{tabular}[t]{lrrrrrrrrrrrrrrrrr}
 \toprule
-wo & 50 & 50 & 60 & 60 & 80 & 80 & 90 & 90 & 100 & 100 & 110 & 110 & 120 & 120 & 120 & 120 & 120 & 130 & 130 & 140 & 140 & 150 & 150 & 160 & 160 & 200 & 200 & 200 & 240 & 240 & 280 & 280 & 320\\
-wv & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 10 & 20 & 30 & 50 & 60 & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 20 & 30 & 50 & 30 & 50 & 30 & 50 & 50\\
+wo & 50 & 50 & 60 & 60 & 80 & 80 & 90 & 90 & 100 & 100 & 110 & 110 & 120 & 120 & 120 & 120 & 120\\
+wv & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 10 & 20 & 30 & 50 & 60\\
+\bottomrule
+\end{tabular}}
+\end{table}
+
+\begin{table}[!h]
+\centering
+\resizebox{\linewidth}{!}{\begin{tabular}{lrrrrrrrrrrrrrrrrr}
+\toprule
+wo & 120 & 130 & 130 & 140 & 140 & 150 & 150 & 160 & 160 & 200 & 200 & 200 & 240 & 240 & 280 & 280 & 320\\
+wv & 60 & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 20 & 30 & 50 & 30 & 50 & 30 & 50 & 50\\
 \bottomrule
@@
 
+\hh{XXX ideas should go to the previous section on setting up the study; in this section we focus on results only XXX}
 The idea therefore is to first use these raw values of the profile
@@
 
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figures/unnamed-chunk-6-1}
+
+}
+
+\caption{Type II errors for different levels of coarseness. XXX We need more values around .25}\label{fig:unnamed-chunk-6}
+\end{figure}
+
+\subsubsection{comparison of total error with signatures and
+profiles}\label{comparison-of-total-error-with-signatures-and-profiles}
+
+\begin{center}\includegraphics[width=\textwidth]{figures/unnamed-chunk-7-1} \end{center}
+
+\subsection{Discussion}\label{discussion}
+
+\subsection{Results of Chumbley score}\label{results-of-chumbley-score}
+
+\begin{itemize}
+\tightlist
+\item
+  Nominal alpha value shows dependency on the size of the window of
+  optimization
+\item
+  Test Fail depends on whether known-match or known non-matches has
+  predictive value
+\item
+  Type II error least bad for window of validation 30 and window of
+  optimization 120
+\end{itemize}
+
+\section{Algorithm Modification}\label{algorithm-modification}
+
+The proposed algorithm modification is at the same shift step which
+comes after the shift distance is identified by finding two windows in
+the two markings that have the highest correlation.
+
+The modification allows for a ``wiggle'' room in the second marking of
+the two sets of markings being compared. This means that each set of
+same shift windows that are being compared, the second marking will
+window that is under consideration is allowd to move a little towards
+the left and a little towards the right.
+
+This gives us a new set of 4 windows, 2 to the left and 2 to the right
+of the original comparison window. Then the correlations for each one of
+these 5 windows with the window under consideration of the 1st marking
+is retrieved (from the validation step correlation matrix) or computed.
+
+Then the window that has the maximum correlation (from the set of 5
+windows in the 2nd marking) with the window of the 1st marking is
+chosen.
+
+This new maximum correlation is then used to compute the U-statistics
+using the same method as before.
+
+\subsection{Expected advantage of
+modification}\label{expected-advantage-of-modification}
+
+We are expecting that this modification would improve the type I and II
+errors for the good. Less number of false negatives and false positives.
+The reason for this expectation is many a times the bullet markings are
+not made at the same distances for two or more bullets because of the
+way it comes out of the barrel.
+
+Therefore rigid same shifts might not necessarily compare the right set
+of windows. In the same shift step for one set of windows, allowing the
+window in the 2nd marking to wiggle left and right and finding the best
+match to the window in the 1st marking, lets us adjust for situations
+where markings are compressed or elongated.
+
+\subsection{Dependence on the delta of new windows from the original
+same shift
+window}\label{dependence-on-the-delta-of-new-windows-from-the-original-same-shift-window}
+
+From initial tests the amount of movement to get new windows left and
+right of the original same shift window seems to directly influence the
+the pvalue and U statistic that we look at.
+
+\subsubsection{other questions}\label{other-questions}
+
+How much movement is reasonable and should be used? Is there a way to
+understand if the 2nd marking in comparison to the 1st marking is
+compressed or elongated or neither? Should the amount of movement
+selected depend on this compression or elongation? or using a static
+delta movement value justifieable.
+
 \bibliographystyle{agsm}

### 2018-01-02 p2/eval_windows.R

@@ -1,0 +1,72 @@
+
+falspostive<- function(sig_alpha){
+  if(is.null(all)){
+    library(tidyverse)
+    files <- dir("./data/signatures", pattern="csv")
+    all <- data.frame()
+    for (file in files) {
+      tmp <- read.csv(file=file.path(path="./data/signatures", file)) 
+      if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% select(-x)
+      all <- rbind(all, tmp)
+    }
+  }
+    summ <- all %>% filter(!is.na(p_value)) %>%
+      mutate(signif = p_value < sig_alpha) %>%
+      group_by(wv, wo, match, signif) %>% tally()
+    summ$error <- with(summ, match != signif)
+    rates <- summ %>% group_by(wv, wo, match) %>% summarize(
+      rate = n[error==TRUE]/sum(n)
+    )
+    combo<- list(summ, rates)
+    names(combo)<- c("summ", "rates")
+    return(combo)
+}
+
+a<- falspostive(0.05)
+
+a$summ %>% ggplot(aes( x  = factor(wv), weight = n, fill=error)) + geom_bar(position="fill") +
+  facet_grid(match~wo, labeller="label_both")
+a$rates %>% ggplot(aes(x = wv, y = rate)) + geom_point() + facet_grid(match~wo, labeller="label_both") +
+  geom_line()
+    
+# summ %>% ggplot(aes( x  = factor(wv), weight = n, fill=error)) + geom_bar(position="fill") +
+#       facet_grid(match~wo, labeller="label_both")
+# 
+#     rates %>% ggplot(aes(x = wv, y = rate)) + geom_point() +
+#       facet_grid(match~wo, labeller="label_both") +
+#       geom_line()
+
+    # library(tidyverse)
+    # files <- dir("./data", pattern="csv")
+    # all <- data.frame()
+    # for (file in files) {
+    #   tmp <- read.csv(file=file.path(path="./data", file)) 
+    #   if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% select(-x)
+    #   all <- rbind(all, tmp)
+    # }
+    # 
+    # summ <- all %>% filter(!is.na(p_value)) %>%
+    #   mutate(signif = p_value < 0.01) %>%
+    #   group_by(wv, wo, match, signif) %>% tally()
[…]

### 2018-01-02 doc/paper.Rmd
.
rearranging files

@@
 library(kableExtra)
-knitr::kable(t(param_setting), digits=0, caption = "Overview of parameter settings used for optimization and validation windows for bullet land signatures.", format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("scale_down"))
+knitr::kable(t(param_setting), digits=10, caption = "Overview of parameter settings used for optimization and validation windows for bullet land signatures.", format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("scale_down"))
 ```

### 2018-01-02 p2/p2.Rmd
rearranging files

@@
 
+# Multiple plot function
+#
+# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
+# - cols:   Number of columns in layout
+# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
+#
+# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
+# then plot 1 will go in the upper left, 2 will go in the upper right, and
+# 3 will go all the way across the bottom.
+#
 # Multiple plot function
@@
 
-\hh{'considered to be unique' by AFTE - that is right, but we should not use this argument, and I don't think we have to.
-It's better to focus on the problem at hand: striation marks are used to determine same source for bullets.
-XXX we could include some figures here XXX Current standards ask for a visual inspection of the marks under a comparison microscope by a firearms examiner.
-One of the problems raised by the NAS report in 2009 is the subjectivity involved in this comparison and the need for determining error rates \citep{NAS:2009}.
-There are different approaches used to achieve more objectivity and assess error rates.
-XXX here we need some examples of approaches XXX
-For toolmarks, the Chumbley score has been suggested to determine same-source. XXX summary of results from original paper and the adjustment XXX
-In this paper we are interested in how this method has to be adjusted to best be applied to the smaller marks on bullets. }
-\gk{done}
-
 Striation marks are used to determine same source for bullets.
@@
 
-A maximum correlation statistic is used to identify the region of best agreement, with the maximum usually seen being near 1 for both cases which is what we intuitively expect for matches, and something which we do not intuitively expect for non matches \gk{done}\hh{XXX get rid of the parentheses - put the content into a sentence after your first statement.}
-
-\hh{XXX the process you describe below is the cross-correlation of two series $x(t)$ and $y(t)$.}
+A maximum correlation statistic is used to identify the region of best agreement, with the maximum usually seen being near 1 for both cases which is what we intuitively expect for matches, and something which we do not intuitively expect for non matches
 
 First, there are a very large number of cross-correlations calculated for the two series of striae or marks $x(t_1)$ and $y(t_2)$, for eg if the window of optimization was defined as 200 and toolmark pixel length is around 1000 then we have 1200-(200-1) = 1001.
-Can be thought of ordered arrangements without replacement where n = 1001 and r = 1.
 
@@
 toolmark.falsepositive$`Non-Match`<- c(3,50)
-toolmark.falsepositive
+
+knitr::kable(toolmark.falsepositive, digits=0, format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("hold"))
 
@@
 Let ns and nd be the number of same shift and different shift windows
-$N = n_{s} + n_{d}$
+$$N = n_{s} + n_{d}$$
 
 The mann whitney U statistic is given by
-$U =\sum^{ns}_{i=1}R_{s}\left( i\right)$
+$$U =\sum^{ns}_{i=1}R_{s}\left( i\right)$$
 
@@
 
-$\overline{U}= \dfrac{U-M}{\sqrt{V}}$
+$$\overline{U}= \dfrac{U-M}{\sqrt{V}}$$
 
@@
 
-$M = n_{s}\left(\dfrac{N+1}{2}\right)$
+$$M = n_{s}\left(\dfrac{N+1}{2}\right)$$
 
@@
 
-$V = \dfrac{n_{s}n_{d}}{N\left(N-1\right)}\left[\Sigma^{n_{s}}R_{s}\left(i\right)^{2}+\Sigma^{n_{d}}R_{d}\left(j\right)^{2}\right] -\dfrac{n_{s} n_{d}\left(N+1\right)^{2}}{4\left(N-1\right)}$
+$$V = \dfrac{n_{s}n_{d}}{N\left(N-1\right)}\left[\Sigma^{n_{s}}R_{s}\left(i\right)^{2}+\Sigma^{n_{d}}R_{d}\left(j\right)^{2}\right] -\dfrac{n_{s} n_{d}\left(N+1\right)^{2}}{4\left(N-1\right)}$$
 
@@
 
-<!-- \begin{minipage}{0.3\textwidth} -->
 ```{r }
 library(kableExtra)
-a27<- all %>% filter(wo==320, wv==50) %>% xtabs(data= ., ~signif+match)
+a27<- all %>% filter(wo==200, wv==30) %>% xtabs(data= ., ~signif+match)
 knitr::kable(a27, digits=0, caption = "Window of Optimization 320 and Window of Validation 50", format = "latex" ,booktabs=TRUE)%>% kableExtra::kable_styling(latex_options = c("hold"))
@@
 ```
-<!-- \end{minipage} -->
-<!-- \begin{minipage}{0.3\textwidth} -->
+
 ```{r}
-a27<- all %>% filter(wo==120, wv==50) %>% xtabs(data= ., ~signif+match)
+a27<- all %>% filter(wo==120, wv==30) %>% xtabs(data= ., ~signif+match)
 knitr::kable(a27, digits=0, caption = "Window of Optimization 120 and Window of Validation 50", format = "latex" ,booktabs=TRUE)%>% kableExtra::kable_styling(latex_options = c("hold"))
 ```
-<!-- \end{minipage} -->
+
 ```{r}
-a27<- all %>% filter(wo==80, wv==50) %>% xtabs(data= ., ~signif+match)
+a27<- all %>% filter(wo==80, wv==30) %>% xtabs(data= ., ~signif+match)
 knitr::kable(a27, digits=0, caption = "Window of Optimization 80 and Window of Validation 50", format = "latex" ,booktabs=TRUE)%>% kableExtra::kable_styling(latex_options = c("hold"))
@@
 
-knitr::kable(t(param_setting)[,17:33], digits=0, format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("scale_down","hold"))
+knitr::kable(t(param_setting)[,18:33], digits=0, format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("scale_down","hold"))
 
@@
 ```{r, fig.cap="Type II errors for different levels of coarseness. XXX We need more values around .25"}
-errors2 %>% ggplot(aes(x = coarse, y=beta, colour=factor(alpha))) +
+errors2 %>% filter(wv==30)%>% ggplot(aes(x = coarse, y=beta, colour=factor(alpha))) +
   geom_point() + geom_line() +
@@
 
-### comparison of total error with signatures and profiles
+### Comparison of total error with signatures and profiles
 
@@
 
-errors_sig_c %>% ggplot(aes(x = coarse, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + geom_line() + geom_point(data=errors2, aes(shape="profiles")) +geom_line(data=errors2, aes(shape="profiles"))
+total1<- errors_sig_c %>% ggplot(aes(x = coarse, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + geom_line()  + geom_point(data= errors2 %>% filter(wv==30), aes(shape="profiles")) + geom_line(data=errors2 %>% filter(wv==30), aes(shape="profiles")) + theme_bw() +
+  scale_colour_brewer(expression("type I error"~alpha), palette="Set2")
+
+total2<- wo120 %>% filter(wv>10) %>% ggplot(aes(x = wv, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + geom_line() +geom_point(data=errors2 %>% filter(wv>10) %>% filter(coarse==0.25), aes(shape="profiles")) + geom_line(data=errors2 %>% filter(wv>10)%>% filter(coarse==0.25), aes(shape="profiles")) + theme_bw()
+
+multiplot(total1, total2)
 ```

### 2018-01-02 p2/p2.Rmd
.

@@
   thanks: The authors gratefully acknowledge ...
-- affiliation: Department of Statistics and csafe, Iowa State University 
+- affiliation: Department of Statistics and CSAFE, Iowa State University 
   name: Heike Hofmann
@@
 
-```{r, fig.cap="Type II errors for different levels of coarseness. XXX We need more values around .25"}
+```{r, fig.cap="Type II errors for different levels of coarseness."}
 errors2 %>% filter(wv==30)%>% ggplot(aes(x = coarse, y=beta, colour=factor(alpha))) +

### 2018-01-02 p2/p2.Rmd
.
fp fn definitions

@@
 
-## The Chumbley Score Test
+# The Chumbley Score Test
 
@@
 
-## Testing setup:
+## Testing setup
 \hh{Using cross-validation setup to identify appropriate parameter settings for (a) signatures and (b) profiles directly}
@@
 a27<- all %>% filter(wo==200, wv==30) %>% xtabs(data= ., ~signif+match)
+a27<- as.data.frame(a27)
+a27$Type<- c("True Negative","False Positive (Type I)","False Negative (Type II)","True Positive")
 knitr::kable(a27, digits=0, caption = "Window of Optimization 320 and Window of Validation 50", format = "latex" ,booktabs=TRUE)%>% kableExtra::kable_styling(latex_options = c("hold"))
@@
 a27<- all %>% filter(wo==120, wv==30) %>% xtabs(data= ., ~signif+match)
+a27<- as.data.frame(a27)
+a27$Type<- c("True Negative","False Positive (Type I)","False Negative (Type II)","True Positive")
 knitr::kable(a27, digits=0, caption = "Window of Optimization 120 and Window of Validation 50", format = "latex" ,booktabs=TRUE)%>% kableExtra::kable_styling(latex_options = c("hold"))
@@
 a27<- all %>% filter(wo==80, wv==30) %>% xtabs(data= ., ~signif+match)
+a27<- as.data.frame(a27)
+a27$Type<- c("True Negative","False Positive (Type I)","False Negative (Type II)","True Positive")
 knitr::kable(a27, digits=0, caption = "Window of Optimization 80 and Window of Validation 50", format = "latex" ,booktabs=TRUE)%>% kableExtra::kable_styling(latex_options = c("hold"))

### 2018-01-02 p2/p2.Rmd
.

--- 2018-01-02 p2/p2.Rmd
@@
 a27<- all %>% filter(wo==200, wv==30) %>% xtabs(data= ., ~signif+match)
-knitr::kable(a27, digits=0, caption = "Window of Optimization 320 and Window of Validation 50", format = "latex" ,booktabs=TRUE)%>% kableExtra::kable_styling(latex_options = c("hold"))
+a27<- as.data.frame(a27)
+a27$Type<- c("True Negative","False Positive (Type I)","False Negative (Type II)","True Positive")
+#knitr::kable(a27, digits=0, caption = "Window of Optimization 320 and Window of Validation 50", format = "latex" ,booktabs=TRUE)%>% kableExtra::kable_styling(latex_options = c("hold"))
 
@@
 ```{r}
-a27<- all %>% filter(wo==120, wv==30) %>% xtabs(data= ., ~signif+match)
-knitr::kable(a27, digits=0, caption = "Window of Optimization 120 and Window of Validation 50", format = "latex" ,booktabs=TRUE)%>% kableExtra::kable_styling(latex_options = c("hold"))
+a28<- all %>% filter(wo==120, wv==30) %>% xtabs(data= ., ~signif+match)
+a28<- as.data.frame(a28)
+a28$Type<- c("True Negative","False Positive (Type I)","False Negative (Type II)","True Positive")
+#knitr::kable(a27, digits=0, caption = "Window of Optimization 120 and Window of Validation 50", format = "latex" ,booktabs=TRUE)%>% kableExtra::kable_styling(latex_options = c("hold"))
 ```
@@
 ```{r}
-a27<- all %>% filter(wo==80, wv==30) %>% xtabs(data= ., ~signif+match)
-knitr::kable(a27, digits=0, caption = "Window of Optimization 80 and Window of Validation 50", format = "latex" ,booktabs=TRUE)%>% kableExtra::kable_styling(latex_options = c("hold"))
+a29<- all %>% filter(wo==80, wv==30) %>% xtabs(data= ., ~signif+match)
+a29<- as.data.frame(a29)
+a29$Type<- c("True Negative","False Positive (Type I)","False Negative (Type II)","True Positive")
+
+knitr::kable(list(a27,a28,a29), digits=0, caption = "Confusion Table for different optimization window sizes", format = "latex" ,booktabs=TRUE)%>% kableExtra::kable_styling(latex_options = c("hold"), "striped", full_width = F) %>% group_rows("Window of Optimization 200 and Window of Validation 30",1,5) %>% group_rows("Window of Optimization 120 and Window of Validation 30", 6, 10) %>%  group_rows("Window of Optimization 80 and Window of Validation 30", 11,14)
 ```

### 2018-01-03 p2/p2.Rmd
.

--- 2018-01-02 p2/p2.Rmd
@@
 
-Identification criteria like consecutively matching striae (CMS) as first seen by Biasotti 1959 \citet{biasotti} and later mentioned in \citet{chu2013} more often than not include subject bias and are error prone and as seen in the work of \citet{miller} the number of CMS may turn to out to be high even when the bullet is not fires by the same firearm.
-Parameters like cross-correlation factor as seen in the extensive comparisons made by \citep{aoas} seem to perform much better matching purposes.
+Identification criteria like consecutively matching striae (CMS) as first seen by Biasotti 1959 \citet{biasotti} and later mentioned in \citet{chu2013} more often than not include subject bias and are error prone and as seen in the work of \citet{miller} the number of CMS may turn out to be high even when the bullet is not fired by the same firearm.
+Parameters like cross-correlation factor as seen in the extensive comparisons made by \citep{aoas} seem to perform much better for matching purposes.
 
@@
 
-## The Chumbley Score Test
+# The Chumbley Score Test
 
-\hh{You need to distinguish between $x(t)$ and $x(t_1)$ - $x(t)$ is a spatial process for some $t$.
-$x(t_1)$ is the realization of the process in $t_1$.
-- Alternatively you could also follow the notation in Hadler et al on 'digitized toolmarks at a pixel level' - where our 'pixel' is the resolution of the confocal light microscope, i.e.
-a pixel corresponds to 0.645 microns.}
-
-The Chumbley score algorithm takes input as two vectorized processes $x(t_1)$ and $y(t_2)$ which denote two sets of marks or striae.
+The Chumbley score algorithm takes input as two vectorized processes $x(t_1)$, $t_1 = 1,2,...T_1$ and $y(t_2)$, $t_2 = 1,2...T_2$ which denote two sets of marks or striae.
+Here the processes are of the form $z(s)$ which is a spatial process for some $t$, while $z(s_1)$ is the realization of the process in $s_1$.
+The $t_1$ and $t_2$ can be better understood as equally spaced pixel locations for the two marks under consideration, where a 'pixel' is the resolution of the confocal light microscope.
+In the case for bullet signatures a pixel corresponds to 0.645 microns.
 These marks or striae are potentially from two different bullets or two different toolmarks whose source needs to be identified as being same or different.
 The marks or striae are indexed by the pixel location where $t_1$ is for the first striae referred to as $x$ and $t_2$ is for the second striae which is referred as $y$.
-$t_1$ and $t_2$ need not be the same but are usually of similar lenghtsThe similarity is then judged by the algorithm on the basis of cross-correlation of a fixed and constant number consecutive pixels say $k$ taken from the two indexed marks $x(t_1)$ and $y(t_2)$ such that in theory $k$ remains smaller that length of the two striae or marks which is the same as $t_1$ and $t_2$.
+$T_1$ and $T_2$ being the respective lengths of the markings, need not be the same but are usually of similar lenghtsThe similarity is then judged by the algorithm on the basis of cross-correlation of a fixed and constant number consecutive pixels say $k$ taken from the two indexed marks $x(t_1)$ and $y(t_2)$ such that in theory $k$ remains smaller that length of the two striae or marks which is the same as $t_1$ and $t_2$.
 Depending on the what stage of the algorithm we are in, matching of different pixel lengths and locations is done, which at the end effectively compares all possible windows that would guarantee in quantifying the two marks or striae as coming from the same source or not, which also lets us assess the error rates by checking for a large number of cases.
@@
 
-## Testing setup:
-\hh{Using cross-validation setup to identify appropriate parameter settings for (a) signatures and (b) profiles directly}
+## Testing setup
 
@@
 
-The marking made on bullets are smaller than toolmarks and is also less wider. The idea is to find out possible areas of error while adapting the score based method proposed for toolmarks.
+The marking made on bullets are smaller than toolmarks and is also less wider.
+The idea is to find out possible areas of error while adapting the score based method proposed for toolmarks, using cross-validation setup to identify appropriate parameter settings for (a) signatures and (b) profiles directly
 
-Bullet signatures being compared at this time are from the Hamby 44 and Hamby 252 data, which have the best set of known-matched and known non-matches.
-
-Bullet signatures are chosen by
-1. Filtering out Land_id for Profiles from the Hamby 44 and Hamby 252 data and removing 
-all NA values
-2. run_id = 3 is chosen, Seems like the signatures generated from this run_id give the closest
-match. Different run_id's have some different settings for generating the signatures
-(The level of smoothing does not seem to be one of them)
-
-The bullet signatures when generated already included the LOWESS smoothing.
-Therefore, the coarseness factor is set to 1 while running the chumbley_non_random()
-which generates the same_shift, different_shift, U-Stat and P_value parameters.
-
-```{r }
-library(kableExtra)
-a27<- all %>% filter(wo==200, wv==30) %>% xtabs(data= ., ~signif+match)
-knitr::kable(a27, digits=0, caption = "Window of Optimization 320 and Window of Validation 50", format = "latex" ,booktabs=TRUE)%>% kableExtra::kable_styling(latex_options = c("hold"))
-
-```
-
-```{r}
-a27<- all %>% filter(wo==120, wv==30) %>% xtabs(data= ., ~signif+match)
-knitr::kable(a27, digits=0, caption = "Window of Optimization 120 and Window of Validation 50", format = "latex" ,booktabs=TRUE)%>% kableExtra::kable_styling(latex_options = c("hold"))
-```
-
-```{r}
-a27<- all %>% filter(wo==80, wv==30) %>% xtabs(data= ., ~signif+match)
-knitr::kable(a27, digits=0, caption = "Window of Optimization 80 and Window of Validation 50", format = "latex" ,booktabs=TRUE)%>% kableExtra::kable_styling(latex_options = c("hold"))
-```
-
-# Results
-
-## Signatures
+### Signatures
 
@@
 
-We used the adjusted Chumbley method as proposed in \citet{hadler} and implemented in the R package `toolmaRk` \citep{toolmark} on all pairwise land-to-land comparisons of the Hamby scans (a total of 85,491 comparisons) in the following manner:
+Bullet signatures being compared at this time are therefore from the Hamby 44 and Hamby 252 data.
+The database setup and pre-processing system used for choosing the Bullet signatures are as described by \citet{aoas}.
+In order to choose the bullet signatures we first filter out Land_id for Profiles from the Hamby 44 and Hamby 252 data and remove all NA values.
+Then run_id = 3 is chosen as the signatures generated from this run_id give the closest match.
+Different run_id's have some different settings for generating the signatures.(The level of smoothing does not seem to be one of them)
 
-```{r, echo=FALSE, results='asis'}
+The bullet signatures when generated by this process already includes a loess smoothing.
+Therefore, the coarseness factor is set to 1 while running the chumbley non random algorithm for comparing different optimization windows.The algorithm generates the same_shift, different_shift, U-Stat and P_value parameters which are then used to calculate the errors associated with different sets of window sizes.
+
+We used the adjusted Chumbley method as proposed in \citet{hadler} and implemented in the R package `toolmaRk` \citep{toolmark} on all pairwise land-to-land comparisons of the Hamby scans (a total of 85,491 comparisons) with the pairwise sets for the comparisons given in the following table.
+
+```{r param, echo=FALSE, results='asis'}
 param_setting<- all %>% select(wo, wv) %>% group_by(wo, wv) %>% tally() %>% select(wo,wv)
@@
 
+### Profiles
+
+The profiles are cross-sectional values of the the bullet striation mark which are chosen at an optimum height (x as used by \citet{aoas}).
+This x or height is not a randomly chosen level.
+The rationale behind the choice has been explained by \citet{aoas}.
+A region is first chosen where the cross-correlation seems to change very less and in this region an optimum height is chosen.
+The profiles generally resemble a curve which is more or less similar to a quadratic curve (a quadratic fit to the raw data values of the profile is not an exact fit but it does show a similar trend).
+Profiles are the set of raw values representing the striation marks, and signatures are generated from these by removal of the inherent curvature and applying some smoothing (the signatures generated by \citet{aoas} use a loess function for smoothing).
+
+Similar to signatures the run_id = 3 was used when applying the chumbley algorithm using the database setup given by \citet{aoas} of Hamby-44 and Hamby-252 datasets, on the profiles.
+The run_id not only defines the level of smoothing but also signifies the chosen height at which the profiles were selected initially.
+Another important aspect is the range of horizontal values (which is referred to as the y values in \citet{aoas}) in the signatures.
+These have already been pre-processed in the database to not include any grooves.
+
+Therefore for the sake of comparison the run_id = 3 is still chosen so as to ensure that the horizontal values remain the same as that of the signatures.
+This also gives us profiles with the grooves removed.
+
+The idea therefore is to first use these raw values of the profile directly in the chumbley algorithm, and see how the algorithm performs for different coarseness values (smoothing parameter as referred in the function lowess used in the chumbley algorithm)
+
+```{r }
+
+#knitr::kable(a27, digits=0, caption = "Window of Optimization 320 and Window of Validation 50", format = "latex" ,booktabs=TRUE)%>% kableExtra::kable_styling(latex_options = c("hold"))
+
+```
+
+```{r}
+
+#knitr::kable(a27, digits=0, caption = "Window of Optimization 120 and Window of Validation 50", format = "latex" ,booktabs=TRUE)%>% kableExtra::kable_styling(latex_options = c("hold"))
+```
+
+```{r confusion}
+library(kableExtra)
+a27<- all %>% filter(wo==200, wv==30) %>% xtabs(data= ., ~signif+match)
+a27<- as.data.frame(a27)
+a27$Type<- c("True Negative","False Positive (Type I)","False Negative (Type II)","True Positive")
+
+a28<- all %>% filter(wo==120, wv==30) %>% xtabs(data= ., ~signif+match)
+a28<- as.data.frame(a28)
+a28$Type<- c("True Negative","False Positive (Type I)","False Negative (Type II)","True Positive")
+
+a29<- all %>% filter(wo==80, wv==30) %>% xtabs(data= ., ~signif+match)
+a29<- as.data.frame(a29)
+a29$Type<- c("True Negative","False Positive (Type I)","False Negative (Type II)","True Positive")
+
+knitr::kable(list(a27,a28,a29), digits=0, caption = "Confusion Table for different optimization window sizes with validation window size as 30", format = "latex" ,booktabs=TRUE)%>% kableExtra::kable_styling(latex_options = c("hold"), "striped", full_width = F) %>% group_rows("Size of Optimization Window = 200",1,5) %>% group_rows("Size of Optimization Window = 120", 6, 10) %>%  group_rows("Size of Optimization Window = 80", 11,14)
+```
+
+# Results
+
 ### Signatures
@@
 
-Figure \ref{fig:type1} compares nominal (fixed) type I error and actually observed type I errors for the parameter settings in table XXX.
+Figure \ref{fig:type1} compares nominal (fixed) type I error and actually observed type I errors for the parameter settings in table \ref{tab:param}.
 With an increasing size of the window used in the optimization step the observed type I error rate decreases (slighty).
@@
 
-```{r wo120, fig.width = 8, fig.height=4.5}
+```{r wo120, fig.width = 8, fig.height=4.5, fig.cap="For optimization window size 120 and different validation window size how Actual and Type II (beta) errors vary for various chosen nominal alpha levels"}
 wo120 <- errors %>% filter(wo==120)
@@
 ### Profiles
-
-The profiles are cross-sectional values of the the bullet striation mark which are chosen at an optimum height (x as used by \citet{aoas}).
-This x or height is not a randomly chosen level.
-The rationale behind the choice has been explained by \citet{aoas}.
-A region is first chosen where the cross-correlation seems to change very less and in this region an optimum height is chosen.
-The profiles generally resemble a curve which is more or less similar to a quadratic curve (a quadratic fit to the raw data values of the profile is not an exact fit but it does show a similar trend).
-Profiles are the set of raw values representing the striation marks, and signatures are generated from these by removal of the inherent curvature and applying some smoothing (the signatures generated by \citet{aoas} use a loess function for smoothing).
-
-Coming back to the database of Hamby-44 and Hamby-252 datasets, the run_id = 3 was used when applying the chumbley algorithm on the signatures.
-The run_id not only defines the level of smoothing but also signifies the chosen height at which the profiles were selected initially.
-Another important aspect is the range of horizontal values (which is referred to as the y values in \citet{aoas}) in the signatures.
-Thes have already been pre-processed in the database to not include any grooves.
-
-Therefore for the sake of comparison the run_id = 3 is still chosen so as to ensure that the horizontal values remain the same as that of the signatures.
-This also gives us profiles with the grooves removed.
-
-\hh{XXX ideas should go to the previous section on setting up the study; in this section we focus on results only XXX}
-The idea therefore is to first use these raw values of the profile directly in the chumbley algorithm, and see how the algorithm performs for different coarseness values (smoothing parameter as referred in the function LOWESS used in the chumbley algorithm)
 
@@
 total1<- errors_sig_c %>% ggplot(aes(x = coarse, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + geom_line()  + geom_point(data= errors2 %>% filter(wv==30), aes(shape="profiles")) + geom_line(data=errors2 %>% filter(wv==30), aes(shape="profiles")) + theme_bw() +
-  scale_colour_brewer(expression("type I error"~alpha), palette="Set2")
+  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2")
 
-total2<- wo120 %>% filter(wv>10) %>% ggplot(aes(x = wv, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + geom_line() +geom_point(data=errors2 %>% filter(wv>10) %>% filter(coarse==0.25), aes(shape="profiles")) + geom_line(data=errors2 %>% filter(wv>10)%>% filter(coarse==0.25), aes(shape="profiles")) + theme_bw()
+total2<- wo120 %>% filter(wv>10) %>% ggplot(aes(x = wv, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + geom_line() +geom_point(data=errors2 %>% filter(wv>10) %>% filter(coarse==0.25), aes(shape="profiles")) + geom_line(data=errors2 %>% filter(wv>10)%>% filter(coarse==0.25), aes(shape="profiles")) + theme_bw() +
+  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2")
 
@@
 - Type II error least bad for window of validation 30 and window of optimization 120
+- Type II error for profiles increases with the amount of smoothing and least for LOWESS smoothing coarseness value about 0.25 or 0.3.
+- Profiles in comparison with Signatures seem to perform worse with total error either equal or more for all sizes of validation window.
+-
 

### 2018-01-03 p2/p2.Rmd
failed test info for comparison between profiles and signatures

@@
 
-```{r, fig.cap="Type II errors for different levels of coarseness."}
-errors2 %>% filter(wv==30)%>% ggplot(aes(x = coarse, y=beta, colour=factor(alpha))) +
+```{r, fig.cap="Type II errors for different levels of coarseness of profiles."}
+e1<-errors2 %>% filter(wv==30)%>% ggplot(aes(x = coarse, y=beta, colour=factor(alpha))) +
   geom_point() + geom_line() +
   theme_bw() +
-  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2")
+  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + xlab("Profile Coarseness")+
+  ylab("Type II error")
+
+e2<- errors_sig_c %>% filter(wv==30)%>% ggplot(aes(x = coarse, y=beta, colour=factor(alpha))) +
+  geom_point() + geom_line() +
+  theme_bw() +
+  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2")+ xlab("Signature Coarseness") + ylab("Type II error")
+
+multiplot(e1,e2, cols = 1)
+
 ```
@@
 total1<- errors_sig_c %>% ggplot(aes(x = coarse, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + geom_line()  + geom_point(data= errors2 %>% filter(wv==30), aes(shape="profiles")) + geom_line(data=errors2 %>% filter(wv==30), aes(shape="profiles")) + theme_bw() +
-  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2")
+  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + ylab("Total error")
 
 total2<- wo120 %>% filter(wv>10) %>% ggplot(aes(x = wv, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + geom_line() +geom_point(data=errors2 %>% filter(wv>10) %>% filter(coarse==0.25), aes(shape="profiles")) + geom_line(data=errors2 %>% filter(wv>10)%>% filter(coarse==0.25), aes(shape="profiles")) + theme_bw() +
-  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2")
+  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + ylab("Total error")
 
-multiplot(total1, total2)
 ```
+
+```{r prof_missings, fig.width=8, fig.height=5, fig.cap="Number of failed tests by the window optimization size, wo, and ground truth.
+Test results from different sources have a much higher chance to fail, raising the question, whether failed tests should be treated as rejections of the null hypothesis of same source."}
+nsp <- allp %>% group_by(wv, wo, match, coarse) %>% summarize(
+  n_ss = mean(same_shift_n),
+  n_ds = mean(diff_shift_n),
+  zero_ss = sum(same_shift_n==0),
+  zero_ds = sum(diff_shift_n==0),
+  miss = sum(is.na(p_value)),
+  missperc = miss/85491*100
+)
+
+ns_sig_c <- all_sig_c %>% group_by(wv, wo, match, coarse) %>% summarize(
+  n_ss = mean(same_shift_n),
+  n_ds = mean(diff_shift_n),
+  zero_ss = sum(same_shift_n==0),
+  zero_ds = sum(diff_shift_n==0),
+  miss = sum(is.na(p_value)),
+  missperc = miss/85491*100
+)
+
+#ns1<- ns[which(ns$wv == c(30,50), arr.ind = T),]
+miss_1<- nsp %>% filter(wv == 30) %>%
+#ns %>% filter(wv == c(30, 50)) %>%
+  ggplot(aes(x = coarse, y = missperc, shape= "profile", colour = factor(match))) + 
+  #geom_smooth(method="auto", size=.5, se=FALSE) +
+  geom_point(size = 2) + 
+  theme_bw() +
+  ylab("Percent of failed tests.") +
+  xlab("Coarseness") +
+  scale_colour_brewer("Match", palette="Set1") + geom_point(data=ns  %>% filter(wv %in% c(30,50)) , aes(shape="signature"), size = 2)# + scale_colour_brewer("Match", palette="Set2") #+geom_smooth(method="auto", size=.5, se=FALSE)
+  #scale_shape_discrete("Size of \nvalidation\nwindow, #wv")
+
+miss_2<- nsp %>% filter(coarse == 0.25) %>%
+#ns %>% filter(wv == c(30, 50)) %>%
+  ggplot(aes(x = wv, y = missperc, shape= "profile", colour = factor(match))) + 
+  #geom_smooth(method="auto", size=.5, se=FALSE) +
+  geom_point(size = 2) + 
+  theme_bw() +
+  ylab("Percent of failed tests.") +
+  xlab("wv") +
+  scale_colour_brewer("Match", palette="Set1") + geom_point(data=ns %>% filter(wo == 120)  %>% filter(coarse == 1) , aes(shape="signature"), size = 2) #+ scale_colour_brewer("Match", palette="Set2")
+
+multiplot(plotlist = list(total1, total2, miss_1, miss_2), cols = 2)#, matrix(c(1,2,3,3), nrow=2, byrow=TRUE))
+```
+
 ## Discussion

### 2018-01-03 p2/p2.Rmd
.

@@
 
-```{r, fig.cap="Type II errors for different levels of coarseness of profiles."}
+```{r, fig.cap="Type II errors for different levels of coarseness."}
 e1<-errors2 %>% filter(wv==30)%>% ggplot(aes(x = coarse, y=beta, colour=factor(alpha))) +
@@
 total1<- errors_sig_c %>% ggplot(aes(x = coarse, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + geom_line()  + geom_point(data= errors2 %>% filter(wv==30), aes(shape="profiles")) + geom_line(data=errors2 %>% filter(wv==30), aes(shape="profiles")) + theme_bw() +
-  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + ylab("Total error")
+  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + xlab("Coarseness") +ylab("Total error")
 
@@
 
-```{r prof_missings, fig.width=8, fig.height=5, fig.cap="Number of failed tests by the window optimization size, wo, and ground truth.
-Test results from different sources have a much higher chance to fail, raising the question, whether failed tests should be treated as rejections of the null hypothesis of same source."}
+```{r prof_missings, fig.width=8, fig.height=5, fig.cap="Bottom row shows Total error and Number of failed tests by the window validation size, wv, and ground truth, while the top row shows Total error and Number of failed tests with Coarseness for both profiles and signatures"}
 nsp <- allp %>% group_by(wv, wo, match, coarse) %>% summarize(
@@
   xlab("Coarseness") +
-  scale_colour_brewer("Match", palette="Set1") + geom_point(data=ns  %>% filter(wv %in% c(30,50)) , aes(shape="signature"), size = 2)# + scale_colour_brewer("Match", palette="Set2") #+geom_smooth(method="auto", size=.5, se=FALSE)
+  scale_colour_brewer("Match", palette="Set1") + geom_point(data=ns_sig_c  %>% filter(wv %in% c(30,50)) , aes(shape="signature"), size = 2)# + scale_colour_brewer("Match", palette="Set2") #+geom_smooth(method="auto", size=.5, se=FALSE)
   #scale_shape_discrete("Size of \nvalidation\nwindow, #wv")

### 2018-01-03 p2/p2.Rmd
abst edit

@@
 ## Results of Chumbley score
-- Nominal alpha value shows dependency on the size of the window of optimization
+
+Compairing pairs of toolmarks with the intention of matching it to a tool has been studied relatively more in the past as compared to bullets, and \citet{chumbley} have described in their paper an algorithm and a deterministic method that compares
+two toolmarks and come to the conclusion if they are from the same tool or not.
+The method also determines the error rates, reduces subject bias and designate the two toolmarks as matches or non-matches with respect to a source.
+This project tries to adapt the Chumbley algorithm as modified by \citet{hadler}, to bullets which are much smaller in length, width, are not flat and curved in the cross-sectional topography as opposed to tools like screw driver tips which produced longer and pronounced markings.
+The majority of Bullet profiles and signatures extracted by procedures mentioned by  \citet{aoas} are almost 1/4 th the size of toolmarks as used by \citet{chumbley} or even smaller.
+Striations on Bullets are made on their curved surfaces, whereas the algorithm developed by \citet{chumbley} and \citet{hadler} has only been tested for flatter and wider surfaces which have negligible curvature.
+Therefore, using methods proposed for toolmarks may need adaptation in order to give tangible results for bullets.
+Moreover, in order to to get flat bullet signatures and remove the curvatures some kind of smoothing needs to be applied as a pre-step which needs further investigation as to whether the level of smoothing does effect the working of the algorithm on Bullets.
+Another important aspect of adapting the algorithm is to find the sizes of the two comparison windows  \citet{hadler} that minimizes the associated errors.
+This identification is not obvious as,  if we go too small in the comparison windows, the unique features of the trace segments are lost and seem similar, while too large sizes vastly reduces the weight of small features that would otherwise uniquely classify a signature and hence identify the region of agreement.
+
+- Nominal alpha value shows dependency on the size of the window of optimization and the window of validation.
+For a given window of optimization the actual Type I error is near the nominal significance level only for validation window sizes around 30 to 50
 - Test Fail depends on whether known-match or known non-matches has predictive value
@@
 - Type II error for profiles increases with the amount of smoothing and least for LOWESS smoothing coarseness value about 0.25 or 0.3.
-- Profiles in comparison with Signatures seem to perform worse with total error either equal or more for all sizes of validation window.
--
+- On comparing Profiles and Signatures with respect to validation window size,  have a total error greater than or equal to the total error of signatures for all sizes of validation window.
+- Profiles also fail (i.e.
+incorrect classification of a same source match) more number of times than signatures, which lets us conclude that the behaviour of the algorithm for the raw profiles instead of pre-processed signatures is not better.
 

### 2018-01-03 p2/p2.Rmd
short version added

@@
 The method also determines the error rates, reduces subject bias and designate the two toolmarks as matches or non-matches with respect to a source.
-This project tries to adapt the Chumbley algorithm as modified by \citet{hadler}, to bullets which are much smaller in length, width, are not flat and curved in the cross-sectional topography as opposed to tools like screw driver tips which produced longer and pronounced markings.
+This project tries to adapt the Chumbley algorithm as modified by \citet{hadler}, to bullets which are much smaller in length, width, are not flat and curved in the cross-sectional topography as opposed to tools like screw driver tips which produces longer and pronounced markings.
 The majority of Bullet profiles and signatures extracted by procedures mentioned by  \citet{aoas} are almost 1/4 th the size of toolmarks as used by \citet{chumbley} or even smaller.
@@
 Therefore, using methods proposed for toolmarks may need adaptation in order to give tangible results for bullets.
-Moreover, in order to to get flat bullet signatures and remove the curvatures some kind of smoothing needs to be applied as a pre-step which needs further investigation as to whether the level of smoothing does effect the working of the algorithm on Bullets.
+Moreover, in order to to get flat bullet signatures and remove the curvatures some kind of smoothing needs to be applied as a pre-step.
+This needs further investigation as to whether the level of smoothing does effect the working of the algorithm on Bullets.
 Another important aspect of adapting the algorithm is to find the sizes of the two comparison windows  \citet{hadler} that minimizes the associated errors.
@@
 
-- Nominal alpha value shows dependency on the size of the window of optimization and the window of validation.
-For a given window of optimization the actual Type I error is near the nominal significance level only for validation window sizes around 30 to 50
-- Test Fail depends on whether known-match or known non-matches has predictive value
-- Type II error least bad for window of validation 30 and window of optimization 120
-- Type II error for profiles increases with the amount of smoothing and least for LOWESS smoothing coarseness value about 0.25 or 0.3.
-- On comparing Profiles and Signatures with respect to validation window size,  have a total error greater than or equal to the total error of signatures for all sizes of validation window.
-- Profiles also fail (i.e.
-incorrect classification of a same source match) more number of times than signatures, which lets us conclude that the behaviour of the algorithm for the raw profiles instead of pre-processed signatures is not better.
+An objective analysis of signatures (pre-processed markings) and profiles (raw markings) of bullet lands for all Hamby-44 and Hamby-252 scans pairwise land to land comparisons (a total of 85,491 comparisons)\citep{hamby} made available through the NIST ballistics database \citep{nist}(Zheng 2016) was done to identify the effects of the two comparison windows and coarseness parameter on the error rates in as proposed in the adjusted chumbley algorithm for toolmarks by \citet{hadler}.
+The results suggested that the Nominal type I error $\alpha$ value shows dependency on the size of the window of optimization and the window of validation.
+For a given window of optimization the actual Type I error is comparable the nominal level only for only a select few validation window sizes.
+A Test Fail, which is the percentage of incorrect classifications, depends on whether known-match or known non-matches has predictive value, with test results from different sources having a much higher chance to fail.
+On conducting an analysis of all known bullet lands using the adjusted chumbley algorithm, Type II error was identified to be least bad for window of validation 30 and window of optimization 120.
+In case of unsmoothed raw marks (profiles), Type II error increases with the amount of smoothing and least for LOWESS smoothing coarseness value about 0.25 or 0.3.
+In an effort to identify the level of adaptiveness of the algorithm, comparisons were made between signatures and profiles.
+Their comparison with respect to validation window size for a fixed optimization window side suggeted that, profiles have a total error greater than or equal to the total error of signatures for all sizes of validation window.
+Profiles also fail (i.e.
+incorrectly classify known-matches and known non-matches) more number of times than signatures, which lets us conclude that the behaviour of the algorithm for the profiles instead of pre-processed signatures is not better.
+Finally it needs to be noted that the current version of the adjusted chumbley algorithm seems to falls short when compared to other machine-learning based methods \citet{aoas}, and some level of modification to the deterministic algorithm needs to be identified and tested that would reduces the number of incorrect classifications.
 

### 2018-01-03 short/short.Rmd
short version added

--- 2018-01-03 p2/p2.Rmd
@@
 keywords:
-- 3 to 6 keywords
-- that do not appear in the title
+- Bullets, Signatures, Lands, Errors, Known match, Known non-match, Comparisons
 bibliography: bibliography
-abstract: null
+<!-- abstract: null -->
 ---
@@
 
-\tableofcontents
+<!-- \tableofcontents -->
 
@@
 ```
-
-```{r functions}
-errorrate <- function(data, alpha) {
-  summ <- data %>% filter(!is.na(p_value)) %>%
-    mutate(signif = p_value < alpha) %>%
-    group_by(wv, wo, match, coarse, signif) %>% tally()
-  summ$error <- with(summ, match != signif)
-  summ$alpha <- alpha
-
-  rates <- summ %>% group_by(wv, wo, coarse, match, alpha) %>% summarize(
-    rate = n[error==TRUE]/sum(n)
-  )
-  totals <- summ %>% group_by(wv, wo, coarse, alpha) %>% summarize(
-    total= sum(n[error==TRUE])/sum(n)
-  )
-  rates <- rates %>% spread(match,rate) %>% rename(
-    actual = `FALSE`,
-    beta = `TRUE`
-  )
-  left_join(rates, totals)
-}
-
-# Multiple plot function
-#
-# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
-# - cols:   Number of columns in layout
-# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
-#
-# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
-# then plot 1 will go in the upper left, 2 will go in the upper right, and
-# 3 will go all the way across the bottom.
-#
-# Multiple plot function
-#
-# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
-# - cols:   Number of columns in layout
-# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
-#
-# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
-# then plot 1 will go in the upper left, 2 will go in the upper right, and
-# 3 will go all the way across the bottom.
-#
-multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
-  library(grid)
-
-  # Make a list from the ... arguments and plotlist
-  plots <- c(list(...), plotlist)
-
-  numPlots = length(plots)
-
-  # If layout is NULL, then use 'cols' to determine layout
-  if (is.null(layout)) {
-    # Make the panel
-    # ncol: Number of columns of plots
-    # nrow: Number of rows needed, calculated from # of cols
-    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
-                    ncol = cols, nrow = ceiling(numPlots/cols))
-  }
-
- if (numPlots==1) {
-    print(plots[[1]])
-
-  } else {
-    # Set up the page
-    grid.newpage()
-    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
-
-    # Make each plot, in the correct location
-    for (i in 1:numPlots) {
-      # Get the i,j matrix positions of the regions that contain this subplot
-      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
-
-      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
-                                      layout.pos.col = matchidx$col))
-    }
-  }
-}
-#  ref: http://www.cookbook-r.com/Graphs/Multiple_graphs_on_one_page_(ggplot2)/
-```
-
-```{r data}
-if (!file.exists("../data/all-sigs.rds")) {
-  files <- dir("../data/signatures", pattern="csv")
-  all <- data.frame()
-  for (file in files) {
-    tmp <- read.csv(file=file.path(path="../data/signatures", file)) 
-    if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% select(-x)
-    all <- rbind(all, tmp)
-  }
-  all <- all %>% filter(land1_id < land2_id)
-  saveRDS(all, file="../data/all-sigs.rds")
-} else {
-  all <- readRDS("../data/all-sigs.rds")
-}
-
-all$coarse <- 1
-errors <- rbind(errorrate(all, 0.001), 
-                errorrate(all, 0.005), 
-                errorrate(all, 0.01), 
-                errorrate(all, 0.05))
-```
-
-```{r data-profiles}
-if (!file.exists("../data/all-profiles.rds")) {
-  files <- dir("../data/profiles/", pattern="csv")
-  allp <- data.frame()
-  for (file in files) {
-    tmp <- read.csv(file=file.path(path="../data/profiles", file)) 
-    #print(names(tmp))
-    if (length(grep("coarse",names(tmp))) == 0)
-      tmp$coarse <- gsub(".*coar-(.*).csv", "\\1", file)
-    if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% select(-x)
-    allp <- rbind(allp, tmp)
-  }
-  allp <- allp %>% filter(land1_id < land2_id)
-  allp <- allp %>% mutate(coarse = as.numeric(gsub("pt", ".", coarse)))
-  saveRDS(allp, file="../data/all-profiles.rds")
-} else {
-  allp <- readRDS("../data/all-profiles.rds")
-}
-
-errors2 <- rbind(errorrate(allp, 0.001), 
-                errorrate(allp, 0.005), 
-                errorrate(allp, 0.01), 
-                errorrate(allp, 0.05))
-
-```
-
-```{r data-sig-coars}
-if (!file.exists("../data/all-sig-c.rds")) {
-  files <- dir("../data/sig_diff_coarseness/", pattern="csv")
-  all_sig_c <- data.frame()
-  for (file in files) {
-    tmp <- read.csv(file=file.path(path="../data/sig_diff_coarseness/", file)) 
-    #print(names(tmp))
-    if (length(grep("coarse",names(tmp))) == 0)
-      tmp$coarse <- gsub(".*coarse-(.*).csv", "\\1", file)
-    if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% select(-x)
-    all_sig_c <- rbind(all_sig_c, tmp)
-  }
-  all_sig_c <- all_sig_c %>% filter(land1_id < land2_id)
-  all_sig_c <- all_sig_c %>% mutate(coarse = as.numeric(gsub("pt", ".", coarse)))
-  saveRDS(all_sig_c, file="../data/all-sig-c.rds")
-} else {
-  all_sig_c <- readRDS("../data/all-sig-c.rds")
-}
-
-errors_sig_c <- rbind(errorrate(all_sig_c, 0.001), 
-                errorrate(all_sig_c, 0.005), 
-                errorrate(all_sig_c, 0.01), 
-                errorrate(all_sig_c, 0.05))
-
-```
-
-# Introduction and Background
-
-## Problem statement and motivation
-
-Striation marks are used to determine same source for bullets.
-Current standards ask for a visual inspection of the marks under a comparison microscope by a firearms examiner.
-One of the problems raised by the NAS report in 2009 is the subjectivity involved in this comparison and the need for determining error rates \citep{NAS:2009}.
-The issue of subjectivity of a firearms examiner has been reviewed by many authors where things like scientific principle testability and error rates are considered to be defining aspects for an objective analysis such that without identifying the uncertainities associated with a method of comparison, it is inconclusive to regards an analysis complete.
-This means that with the usual method of visual comparison that is adopted by firearms examiners, lies the problem of lack of any systemized quantification and as such there seems no certain way of associating it with any kind of mathematical probability.
-Since determining error rates has been duly noted as a fundamental problem in forensic science \citep{NAS:2009}, there is a need of methods that address this problem.
-Some of the methods that have been used to estimate error rates in firearm examinations include Automatic cartridge case comparison where same source identification of cartridge cases and associated error rates have been provided by \citep{riva}, although as noted by \citep{aoas}, in this case alignments of striae involves roation of planes, which cannot be generalized for bullets.
-In case of methods that use machine learning algorithms some methods which are bootsrap based methods often tend to give over-estimated error rates \citep{efron}, but there are alternate error estimation techniques as described by \citep{aoas}, \citep{efron} and  \citep{vorburger2016} which give better results.
-
-In case of methods that do not use machine learning algorithm, matching two striation marks with each other in order to identify if it is from the same source, can also be done using a well defined comparative statistical algorithm.
-The statistical algorithm is therefore expected  to go through a step by step procedure of identifying the class characteristics i.e striation marks for bullets, choosing the striae, and then follow a systematic procedure of comparing two striae with each other on the basis of criteria like cross-correlation (or others).
-Such a methodology would make it possible to determine error rates too and give definite results regarding what proportion of cases would the analysis be successful in identifying a match correctly and in how many cases would it end up being a false positive.
-
-Identification criteria like consecutively matching striae (CMS) as first seen by Biasotti 1959 \citet{biasotti} and later mentioned in \citet{chu2013} more often than not include subject bias and are error prone and as seen in the work of \citet{miller} the number of CMS may turn out to be high even when the bullet is not fired by the same firearm.
-Parameters like cross-correlation factor as seen in the extensive comparisons made by \citep{aoas} seem to perform much better for matching purposes.
-
-An important aspect of same source matching in firearms is identification of bullet microtopographies that are unique to be chosen for a match, Bullet profiles and signatures prove to be such features.
-The choice of a bullet signature to uniquely define a bullet depends on finding first, a stable region on a Bullet land which minimal noise and many pronounced striation marks.
-\citet{aoas}
-
-The method employed by \citet{aoas} uses the cross correlation factor as a means to identify the stable region.
-The markings or striae in this region are supposed to have a high CCF with each other.
-Therefore a Bullet profile is chosen from this region by taking the cross section at a given height.
-A loess fit on the bullet profile produces  residuals which are termed as signatures and can be used as a unique identifier that can be used while matching two bullets to a firearm.
-
-Compairing pairs of toolmarks on the otherhand, with the intention of matching it to a tool has been studied relatively more in the past as compared to bullets, and \citet{chumbley} have described in their paper an algorithm and analytic method that compares two toolmarks and come to the conclusion if they are from the same tool or not.
-The method also determines the error rates, reduces subject bias and designate the two toolmarks as matches or non-matches with respect to a source.
-\citet{chumbley} used an empirical based setup to validate their proposed analytic and quantitative algorithm.
-
-\citep{chumbley} found that the algorithm gives false-positives in slightly over 2 cases (or 2%) of the time and about 9 false-negatives (or 9%) for every 100 comparisons.
-\citep{hadler} on the other hand, in the improved version of the algorithm first described by \citep{chumbley} found false-positives for 0 cases and 3 false-negatives (or 6%) for 50 comparisons of knwon matches.
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{images/B6-B2-L6.png}
-```{r, echo=FALSE, fig.width=8, fig.height=3, out.width='\\textwidth'}
-land <- read.csv("../data/b6b2l6.csv")
-land %>% ggplot(aes(x = y, y=resid)) +geom_line() +theme_bw()
-```
-
-\caption{\label{fig:rgl} Image of a bullet land from a confocal light microscope at 20 fold magnification.}
-
-\end{figure}
-
-# The Chumbley Score Test
-
-The Chumbley score algorithm takes input as two vectorized processes $x(t_1)$, $t_1 = 1,2,...T_1$ and $y(t_2)$, $t_2 = 1,2...T_2$ which denote two sets of marks or striae.
-Here the processes are of the form $z(s)$ which is a spatial process for some $t$, while $z(s_1)$ is the realization of the process in $s_1$.
-The $t_1$ and $t_2$ can be better understood as equally spaced pixel locations for the two marks under consideration, where a 'pixel' is the resolution of the confocal light microscope.
-In the case for bullet signatures a pixel corresponds to 0.645 microns.
-These marks or striae are potentially from two different bullets or two different toolmarks whose source needs to be identified as being same or different.
-The marks or striae are indexed by the pixel location where $t_1$ is for the first striae referred to as $x$ and $t_2$ is for the second striae which is referred as $y$.
-$T_1$ and $T_2$ being the respective lengths of the markings, need not be the same but are usually of similar lenghtsThe similarity is then judged by the algorithm on the basis of cross-correlation of a fixed and constant number consecutive pixels say $k$ taken from the two indexed marks $x(t_1)$ and $y(t_2)$ such that in theory $k$ remains smaller that length of the two striae or marks which is the same as $t_1$ and $t_2$.
-Depending on the what stage of the algorithm we are in, matching of different pixel lengths and locations is done, which at the end effectively compares all possible windows that would guarantee in quantifying the two marks or striae as coming from the same source or not, which also lets us assess the error rates by checking for a large number of cases.
-
-The algorithm works in two phases, namely, an optimization step and a validation step, at the end of which a Mann Whitney U statistic is calculated.
-A pre-processing step to the algorithm is to choose a coarseness value which  is used as a parameter to the LOWESS smoothing function.
-The coarseness essentially gives the proportion of points which influence the smooth at each value, which means larger values lead to more smoothness.
-The LOWESS smoothing is applied to each of two sets of vectorized striae or marks $x(t_1)$ and $y(t_2)$, before proceeding to the algorithm.
-
-\citet{hadler} in their paper proposed an improvement to this algorithm by trying to remove mutual dependence of parameters (due to serial correlation in surface depth values of a toolmark and because of a random sampling sub step in the validation phase which makes a group of pixels to be chosen more than once and hence introduces lack of independence) in certain steps, especially because the Mann-Whitney U statistic that is later calculated in the algorithm and used as a measure to differentiate between matches and non-matches works under the assumption of independence of parameters.
-
-Since we are only interested in a non-parameteric U statistic, Hadler et al.
-proposed a normalization procedure in the Validation step that goes to some extent to address the issue of mutual dependence.
-Also the same shift and different shift substep was modified to use a deterministic rule for sampling sam-shhift and dofferent shift-samples as opposed to the originally proposed random samples.
-
-The data that is used by \citet{chumbley} is generated by a surface profilometer that gives the height in terms of distance along a linear trace.
-This is taken perpendicular to the striations present in the toolmark.
-Two such trace are then compared using the algorithm.
-
-## Detailed algorithm
-
-###Optimization step
-The idea behind this step is to first identify the area of best agreement in the two toolmark data.
-Comparison window size is predefined by the user, which in case of screwdriver toolmarks was chosen by Chumbley et al and Hadler et al as around 10 percent of the length of the toolmarks, which was around 500.
-This window is henceforth referred to as Window of optimization.
-
-A maximum correlation statistic is used to identify the region of best agreement, with the maximum usually seen being near 1 for both cases which is what we intuitively expect for matches, and something which we do not intuitively expect for non matches
-
-First, there are a very large number of cross-correlations calculated for the two series of striae or marks $x(t_1)$ and $y(t_2)$, for eg if the window of optimization was defined as 200 and toolmark pixel length is around 1000 then we have 1200-(200-1) = 1001.
-
-This is the number of windows we have for one toolmark and each window is compared with all windows of the second toolmark (the number of windows is again similar), and the window with the maximum correlation is identified to be the region where the toolmarks are in maximum agreement with each other
-
-###Validation Step
-
-####Same-shift: 
-In this step a series of windows are chosen at random (originally by chumbley et al) and deterministically (by hadler et al), but at a common distance (rigid-shift) from the window identified as the region of best agreement in the vaildation step.
-The correlation of these windows would be as intuitively assumed i.e.
-lower than the maximum correlation window (Optimization step), but the significance is that these same shift windows will still have large enough correlation values for two toolmarks or signatures that are in reality a match.
-
-This also validates that if in the optimization step the maximum correlation window chosen (with correlation value near 1) was by accident (like in case of signatures that in reality are a non-match), all these same- shift correlations (A fixed number of these trace segments are identified) would not be anywhere large enough for all same-shift windows.
-
-```{r win-comparison, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap=" The two plots on the left show how the same shift behaves in case of a matching pair and the two plots on the right show how the different shift behaves in case of a matching pair."}
-a<- matrix(NA, ncol = 3, nrow = 2)
-toolmark.falsepositive<- data.frame(a)
-colnames(toolmark.falsepositive)<- c("Classification", "Match", "Non-Match")
-toolmark.falsepositive$Classification<- c("Match", "Non-Match")
-toolmark.falsepositive$Match<- c(47,0)
-toolmark.falsepositive$`Non-Match`<- c(3,50)
-
-knitr::kable(toolmark.falsepositive, digits=0, format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("hold"))
-
-asd<- all %>% filter(wo==120, wv==50, match == TRUE, land1_id == 1)
-
-sigs_graphics<- readRDS("../data/sigs_generate_window.rds")
-
-sigs_graphics.match<- as.data.frame(sigs_graphics[1])
-sigs_graphics.nonmatch<- as.data.frame(sigs_graphics[2])
-colnames(sigs_graphics.match)<- gsub("^.*\\.","", colnames(sigs_graphics.match))
-colnames(sigs_graphics.nonmatch)<- gsub("^.*\\.","", colnames(sigs_graphics.nonmatch))
-
-# Matching signature
-#gi<-
-#sigs_graphics.match<- sigs_graphics.match %>% filter(land_id == c(1,9))  
-#sigs_graphics.match<- sigs_graphics.match %>% select(y, l30, land_id)
-d1<- sigs_graphics.match[which(sigs_graphics.match$land_id==1,arr.ind = TRUE),]
-d2<- sigs_graphics.match[which(sigs_graphics.match$land_id==19,arr.ind = TRUE),]
-data1<-d1 %>%select(l30)
-#data1<- data1$y
-data2<- d2 %>%select(l30)
-#data2<- data2$y
-window_opt = 120 
-window_val = 50
-coarse = 1
-data1<- matrix(unlist(data1))
-data2<- matrix(unlist(data2))
-
-  unity <- function(x) {x / sqrt(sum(x^2))} ## normalize columns of a matrix to make correlation computation faster
-
-  ####################################################
-  ##Clean the marks and compute the smooth residuals##
-  ####################################################
-
-  data1 <- matrix(data1[round((0.01*nrow(data1))):round(0.99*nrow(data1)),], ncol = 1)
-  data2 <- matrix(data2[round((0.01*nrow(data2))):round(0.99*nrow(data2)),], ncol = 1)
-
-  ##Normalize the tool marks
-  y1 <- data1 - lowess(y = data1,  x = 1:nrow(data1), f= coarse)$y
-  y2 <- data2 - lowess(y = data2,  x = 1:nrow(data2), f= coarse)$y
-
-  ############################################
-  ##Compute the observed maximum correlation##
-  ############################################
-
-  #####################
-  ##Optimization step##
-  #####################
-  ##Each column in these matrices corresponds to a window in the respective tool mark
-  y1_mat_opt <- matrix(NA, ncol = length(1:(length(y1) - (window_opt - 1))), nrow = window_opt)
-  for(l in 1:(length(y1) - (window_opt - 1))){
-    y1_mat_opt[,l] <- y1[l:(l+(window_opt - 1))]
-  }
-  y2_mat_opt <- matrix(NA, ncol = length(1:(length(y2) - (window_opt - 1))), nrow = window_opt)
-  for(l in 1:(length(y2) - (window_opt - 1))){
-    y2_mat_opt[,l] <- y2[l:(l+(window_opt - 1))]
-  }
-
-  ##Compute the correlation between all pairs of windows for the two marks
-  ##Rows in the following matrix are mark 2, columns are mark 1
-  y2_mat_opt <- apply(scale(y2_mat_opt), 2, unity)
-  y1_mat_opt <- apply(scale(y1_mat_opt), 2, unity)
-  corr_mat_opt <- t(y2_mat_opt) %*% y1_mat_opt ##correlation matrix
-  max_corr_opt_loc <- which(corr_mat_opt == max(corr_mat_opt), arr.ind = TRUE) ##pair of windows maximizing the correlation
-
-  s1.rect.opt <- data.frame(xmin=max_corr_opt_loc[1,2], xmax= max_corr_opt_loc[1,2]+window_opt, ymin=-Inf, ymax=Inf)
-    s1.rect.sameshift1 <- data.frame(xmin=max_corr_opt_loc[1,2]-window_val, xmax= max_corr_opt_loc[1,2]- 2*window_val, ymin=-Inf, ymax=Inf)
-       s1.rect.sameshift2 <- data.frame(xmin=max_corr_opt_loc[1,2]+window_opt+ window_val, xmax= max_corr_opt_loc[1,2]+ window_opt+2*window_val, ymin=-Inf, ymax=Inf)
-
-  s2.rect.opt <- data.frame(xmin=max_corr_opt_loc[1,1], xmax= max_corr_opt_loc[1,1]-window_opt, ymin=-Inf, ymax=Inf)
-    s2.rect.sameshift1 <- data.frame(xmin=max_corr_opt_loc[1,1]-window_opt-window_val, xmax= max_corr_opt_loc[1,1]-window_opt- 2*window_val, ymin=-Inf, ymax=Inf)
-       s2.rect.sameshift2 <- data.frame(xmin=max_corr_opt_loc[1,1]+ window_val, xmax= max_corr_opt_loc[1,1]+ 2*window_val, ymin=-Inf, ymax=Inf)
-
-p1<-  ggplot(data = d1, aes(x = y, y = l30)) + geom_line() + 
-  geom_rect(data=s1.rect.opt, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "black",alpha=0.5, inherit.aes = FALSE)+
-    geom_rect(data=s1.rect.sameshift1, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "red",alpha=0.1, inherit.aes = FALSE)+
-    geom_rect(data=s1.rect.sameshift2, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "blue",alpha=0.1, inherit.aes = FALSE)+
-  theme_bw() + labs(y = "Signature Land Id = 1")
-
- p2<-   ggplot(data = d2, aes(x = y, y = l30)) + geom_line() + 
-   geom_rect(data=s2.rect.opt, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "black",alpha=0.5, inherit.aes = FALSE)+
-    geom_rect(data=s2.rect.sameshift1, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "red",alpha=0.1, inherit.aes = FALSE)+
-    geom_rect(data=s2.rect.sameshift2, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "blue",alpha=0.1, inherit.aes = FALSE)+
-  theme_bw() + labs(y = "Signature Land Id = 9")
- ##########
-
- d1.rect.opt <- data.frame(xmin=max_corr_opt_loc[1,2], xmax= max_corr_opt_loc[1,2]+window_opt, ymin=-Inf, ymax=Inf)
-    d1.rect.sameshift1 <- data.frame(xmin=max_corr_opt_loc[1,2]-2*window_opt, xmax= max_corr_opt_loc[1,2]-2*window_opt-window_val, ymin=-Inf, ymax=Inf)
-       d1.rect.sameshift2 <- data.frame(xmin=max_corr_opt_loc[1,2]+window_opt+ window_val, xmax= max_corr_opt_loc[1,2]+ window_opt+2*window_val, ymin=-Inf, ymax=Inf)
-
-  d2.rect.opt <- data.frame(xmin=max_corr_opt_loc[1,1], xmax= max_corr_opt_loc[1,1]-window_opt, ymin=-Inf, ymax=Inf)
-    d2.rect.sameshift1 <- data.frame(xmin=max_corr_opt_loc[1,1]-window_opt-window_val, xmax= max_corr_opt_loc[1,1]-window_opt- 2*window_val, ymin=-Inf, ymax=Inf)
-       d2.rect.sameshift2 <- data.frame(xmin=max_corr_opt_loc[1,1]+ 4*window_val, xmax= max_corr_opt_loc[1,1]+ 5*window_val, ymin=-Inf, ymax=Inf)
-
-p3<-  ggplot(data = d1, aes(x = y, y = l30)) + geom_line() + 
-  geom_rect(data=d1.rect.opt, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "black",alpha=0.5, inherit.aes = FALSE)+
-    geom_rect(data=d1.rect.sameshift1, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "red",alpha=0.1, inherit.aes = FALSE)+
-    geom_rect(data=d1.rect.sameshift2, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "blue",alpha=0.1, inherit.aes = FALSE)+
-  theme_bw() + labs(y = "Signature Land Id = 1")
-
- p4<-   ggplot(data = d2, aes(x = y, y = l30)) + geom_line() + 
-   geom_rect(data=d2.rect.opt, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "black",alpha=0.5, inherit.aes = FALSE)+
-    geom_rect(data=d2.rect.sameshift1, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "red",alpha=0.1, inherit.aes = FALSE)+
-    geom_rect(data=d2.rect.sameshift2, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "blue",alpha=0.1, inherit.aes = FALSE)+
-  theme_bw() + labs(y = "Signature Land Id = 9")
-
-multiplot(p1, p2, p3, p4, cols=2)
-```
-
-####Different Shift:
-Primary reason for this substep is to give perspective to the correlation values of the same shift window correlation values.
-
-This time there are no rigid-shifts but different shifts (distance from Window of Opt with max correlation) chosen randomly by \citet{chumbley} and deterministically by Hadler et al.
-such that there is an equal possibilty of comparing a trace segment from one signature or toolmark to any one in the second signature or toolamrk.
-
-Neither of above sets of correlation are allowed to include the maximum correlation window as identified earlier.
-
-Therefore the assumption is that if two toolmarks or signatures match each other the same-shift correlations would be larger than the different-shift windows,
-and if they are not a match the correlations in the two sets will be very similar.
-
-###U Statistic:
-
-This is computed from the joint rank of all correlations of both the same and different shift samples. As given by \citet{hadler}
-
-Null Hypothesis: If the toolmarks were not match i.e not made by the same tool.
-
-Let ns and nd be the number of same shift and different shift windows
-$$N = n_{s} + n_{d}$$
-
-The mann whitney U statistic is given by
-$$U =\sum^{ns}_{i=1}R_{s}\left( i\right)$$
-
-with the standardized version which includes provision for rank ties
-
-$$\overline{U}= \dfrac{U-M}{\sqrt{V}}$$
-
-where prior to normalization the U-statistic has the mean as
-
-$$M = n_{s}\left(\dfrac{N+1}{2}\right)$$
-
-and variance
-
-$$V = \dfrac{n_{s}n_{d}}{N\left(N-1\right)}\left[\Sigma^{n_{s}}R_{s}\left(i\right)^{2}+\Sigma^{n_{d}}R_{d}\left(j\right)^{2}\right] -\dfrac{n_{s} n_{d}\left(N+1\right)^{2}}{4\left(N-1\right)}$$
-
-## Potential limitations of the Chumbley Score
-
-\hh{in the application to striation marks on bullet lands: smaller in width and curved - need to adjust parameter settings; something similar is done in \citet{afte-chumbley} for toolmark comparisons of slip-joint pliers.
-XXX work in }
-
-Bullets are much smaller in length, width, are not flat and curved in the cross-sectional topography as opposed to tools like screw driver tips which produced longer and pronounced markings.
-This means the makings made by barrels in comparison to toolmarks may have a problem in distinctiveness.
-The majority of Bullet profiles and signatures extracted by procedures mentioned by  \citet{aoas} are almost 1/4 th the size of toolmarks as used by \citet{chumbley} or even smaller.
-Striations on Bullets are made on their curved surfaces, whereas the algorithm developed by \citet{chumbley} and \citet{hadler} has only been tested for flatter and wider surfaces which have negligible curvature.
-Therefore, using methods proposed for toolmarks may need adaptation in order to give tangible results for bullets.
-Moreover, in order to to get flat bullet signatures and remove the curvatures some kind of smoothing needs to be applied as a pre-step which needs further investigation as to whether the level of smoothing does effect the working of the algorithm on Bullets.
-
-Also when in the optimization step, the Window of optimization for bullets will be shorter as the signatures are smaller.
-The idea is to keep the number of windows of optimization sufficiently large, which means shorter Trace segments (partition of signature or toolmark with length = size of window of optimization) that lets us compare smaller segments of one signature to another.
-
-This introduces a problem as, if we go too small in the window of optimization, the unique features of the trace segments are lost and seem similar, while too large sizes vastly reduces the weight of small features that would otherwise uniquely classify a signature and hence identify the region of agreement.
-
-Thus the Window of Optimization has a direct influence on whether we are
-Falsely rejecting the null when it is true (Type I) or Falsely accepting the null when it is false (Type II) making identification of the optimum size of window of optimization very important.
-
-## Testing setup
-
-Following on similar lines to the setup of toolmarks, the first step here is to first identify what difference does different window sizes of optimization and the validation step have, when adapting the toolmark method to bullets.
-
-The marking made on bullets are smaller than toolmarks and is also less wider.
-The idea is to find out possible areas of error while adapting the score based method proposed for toolmarks, using cross-validation setup to identify appropriate parameter settings for (a) signatures and (b) profiles directly
-
-### Signatures
-
-Signatures of lands for all Hamby-44 and Hamby-252 scans made available through the NIST ballistics database \citep{nist} were considered.
-Both of these sets of scans are part of the larger Hamby study \citep{hamby} and each consist of twenty known bullets (two each from ten consecutively rifled Ruger P85 barrels) and fifteen questioned bullets (each matching one of the ten barrels).
-Ground truth for both of these Hamby sets is known and was used to assess correctness of the tests results.
-
-Bullet signatures being compared at this time are therefore from the Hamby 44 and Hamby 252 data.
-The database setup and pre-processing system used for choosing the Bullet signatures are as described by \citet{aoas}.
-In order to choose the bullet signatures we first filter out Land_id for Profiles from the Hamby 44 and Hamby 252 data and remove all NA values.
-Then run_id = 3 is chosen as the signatures generated from this run_id give the closest match.
-Different run_id's have some different settings for generating the signatures.(The level of smoothing does not seem to be one of them)
-
-The bullet signatures when generated by this process already includes a loess smoothing.
-Therefore, the coarseness factor is set to 1 while running the chumbley non random algorithm for comparing different optimization windows.The algorithm generates the same_shift, different_shift, U-Stat and P_value parameters which are then used to calculate the errors associated with different sets of window sizes.
-
-We used the adjusted Chumbley method as proposed in \citet{hadler} and implemented in the R package `toolmaRk` \citep{toolmark} on all pairwise land-to-land comparisons of the Hamby scans (a total of 85,491 comparisons) with the pairwise sets for the comparisons given in the following table.
-
-```{r param, echo=FALSE, results='asis'}
-param_setting<- all %>% select(wo, wv) %>% group_by(wo, wv) %>% tally() %>% select(wo,wv)
-#param_setting<- data.frame(matrix(param_setting, nrow = 16))
-#t(param_setting)
-#param_setting %>% tally() 
-library(kableExtra)
-knitr::kable(t(param_setting)[,1:17], digits=0, caption = "Overview of parameter settings used for optimization and validation windows for bullet land signatures.", format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("scale_down", "hold"))
-
-knitr::kable(t(param_setting)[,18:33], digits=0, format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("scale_down","hold"))
-
-# library(xtable)
-# strCaption <- "Overview of parameter settings used for optimization and validation windows for bullet land signatures."
-#   
-#   # set up xtable output
-# print(xtable(t(param_setting)[,1:17], 
-#              caption = strCaption, label = ""),
-#       include.rownames = TRUE,
-#       include.colnames = FALSE,
-#       caption.placement = "top",
-#       comment=FALSE,
-#       floating=FALSE
-#       )
-# print(xtable(t(param_setting)[,17:33],label = ""),
-#       include.rownames = FALSE,
-#       include.colnames = FALSE,
-#       caption.placement = "bottom",
-#       comment=FALSE,
-#       floating=TRUE
-#       #table.placement = "!ht"
-#       )
-```
-
-### Profiles
-
-The profiles are cross-sectional values of the the bullet striation mark which are chosen at an optimum height (x as used by \citet{aoas}).
-This x or height is not a randomly chosen level.
-The rationale behind the choice has been explained by \citet{aoas}.
-A region is first chosen where the cross-correlation seems to change very less and in this region an optimum height is chosen.
-The profiles generally resemble a curve which is more or less similar to a quadratic curve (a quadratic fit to the raw data values of the profile is not an exact fit but it does show a similar trend).
-Profiles are the set of raw values representing the striation marks, and signatures are generated from these by removal of the inherent curvature and applying some smoothing (the signatures generated by \citet{aoas} use a loess function for smoothing).
-
-Similar to signatures the run_id = 3 was used when applying the chumbley algorithm using the database setup given by \citet{aoas} of Hamby-44 and Hamby-252 datasets, on the profiles.
-The run_id not only defines the level of smoothing but also signifies the chosen height at which the profiles were selected initially.
-Another important aspect is the range of horizontal values (which is referred to as the y values in \citet{aoas}) in the signatures.
-These have already been pre-processed in the database to not include any grooves.
-
-Therefore for the sake of comparison the run_id = 3 is still chosen so as to ensure that the horizontal values remain the same as that of the signatures.
-This also gives us profiles with the grooves removed.
-
-The idea therefore is to first use these raw values of the profile directly in the chumbley algorithm, and see how the algorithm performs for different coarseness values (smoothing parameter as referred in the function lowess used in the chumbley algorithm)
-
-```{r }
-
-#knitr::kable(a27, digits=0, caption = "Window of Optimization 320 and Window of Validation 50", format = "latex" ,booktabs=TRUE)%>% kableExtra::kable_styling(latex_options = c("hold"))
-
-```
-
-```{r}
-
-#knitr::kable(a27, digits=0, caption = "Window of Optimization 120 and Window of Validation 50", format = "latex" ,booktabs=TRUE)%>% kableExtra::kable_styling(latex_options = c("hold"))
-```
-
-```{r confusion}
-library(kableExtra)
-a27<- all %>% filter(wo==200, wv==30) %>% xtabs(data= ., ~signif+match)
-a27<- as.data.frame(a27)
-a27$Type<- c("True Negative","False Positive (Type I)","False Negative (Type II)","True Positive")
-
-a28<- all %>% filter(wo==120, wv==30) %>% xtabs(data= ., ~signif+match)
-a28<- as.data.frame(a28)
-a28$Type<- c("True Negative","False Positive (Type I)","False Negative (Type II)","True Positive")
-
-a29<- all %>% filter(wo==80, wv==30) %>% xtabs(data= ., ~signif+match)
-a29<- as.data.frame(a29)
-a29$Type<- c("True Negative","False Positive (Type I)","False Negative (Type II)","True Positive")
-
-knitr::kable(list(a27,a28,a29), digits=0, caption = "Confusion Table for different optimization window sizes with validation window size as 30", format = "latex" ,booktabs=TRUE)%>% kableExtra::kable_styling(latex_options = c("hold"), "striped", full_width = F) %>% group_rows("Size of Optimization Window = 200",1,5) %>% group_rows("Size of Optimization Window = 120", 6, 10) %>%  group_rows("Size of Optimization Window = 80", 11,14)
-```
-
-# Results
-
-### Signatures
-
-Figure \ref{fig:type2} gives an overview of type II error rates observed when varying the window size in the optimization step.
-
-```{r type2, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap="Type II error rates observed across a range of window sizes for optimization $wo$.
-For a window size of $wo = 120$ we see a drop in type II error rate across all type I rates considered.
-Smaller validation sizes $wv$ are typically associated with a smaller type II error."}
-errors %>% filter(wv %in% c(30,50)) %>%
-  ggplot(aes(x = wo, y = beta, colour=factor(alpha))) + 
-  geom_point(aes(shape=factor(wv))) +
-#  facet_grid(.~wo, labeller="label_both") + 
-  geom_smooth(aes(group=alpha), se=FALSE, method="loess") +
-  theme_bw() +
-  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") +
-  scale_shape_discrete("Size of validation\nwindow, wv") +
-  xlab(expression("Window size for optimization, "~wo)) +
-  ylab("Type II error rate")
-
-labels = expression(P[M1](tilde(z)>0),P[M0](tilde(z)>0))
-```
-
-Figure \ref{fig:type1} compares nominal (fixed) type I error and actually observed type I errors for the parameter settings in table \ref{tab:param}.
-With an increasing size of the window used in the optimization step the observed type I error rate decreases (slighty).
-This might be related to the increasing number of tests that fail for larger window sizes, in particular for non-matching striae (see fig \ref{fig:missings}).
-
-```{r type1, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap="Comparison of observed and nominal type I error rates  across a range of window sizes for optimization $wo$.
-The horizontal line in each facet indicates the nominal type I error rate.
-As the optimization window increase the observed type I error rate gets smaller.
-A smaller validation window tends to be associated with a higher type I error rate."}
-errors %>% filter(wv %in% c(30,50)) %>%
-  ggplot(aes(x = wo, y = actual, colour=factor(alpha))) + 
-  geom_hline(aes(yintercept=alpha), colour="grey30") +
-  geom_point(aes(shape=factor(wv))) +
-  #  facet_grid(.~wo, labeller="label_both") + 
-  geom_smooth(aes(group=alpha), se=FALSE, method="lm") +
-  theme_bw() +
-#  scale_y_log10(breaks=c(0.001,.005, 0.01, 0.05)) +
-  scale_shape_discrete("Size of validation window, wv") +
-  ylab("Observed type I error rate") +
-  xlab(expression("Window size for optimization, "~wo)) +
-  scale_colour_brewer(expression("Nominal type I error "~alpha), palette="Set2") +
-  facet_wrap(~alpha, labeller="label_both", scales="free") +
-  theme(legend.position="bottom")
-
-```
-
-```{r missings, fig.width=8, fig.height=5, fig.cap="Number of failed tests by the window optimization size, wo, and ground truth.
-Test results from different sources have a much higher chance to fail, raising the question, whether failed tests should be treated as rejections of the null hypothesis of same source."}
-ns <- all %>% group_by(wv, wo, match) %>% summarize(
-  n_ss = mean(same_shift_n),
-  n_ds = mean(diff_shift_n),
-  zero_ss = sum(same_shift_n==0),
-  zero_ds = sum(diff_shift_n==0),
-  miss = sum(is.na(p_value)),
-  missperc = miss/85491*100
-)
-
-#ns1<- ns[which(ns$wv == c(30,50), arr.ind = T),]
-ns %>% filter(wv %in% c(30, 50)) %>%
-#ns %>% filter(wv == c(30, 50)) %>%
-  ggplot(aes(x = wo, y = missperc, shape=factor(wv), colour = factor(match))) + 
-  geom_smooth(method="lm", size=.5, se=FALSE) +
-  geom_point() + 
-  theme_bw() +
-  ylab("Percent of failed tests.") +
-  xlab("Size of optimization window, wo") +
-  scale_colour_brewer("Match", palette="Set1") +
-  scale_shape_discrete("Size of \nvalidation\nwindow, wv")
-```
-Figure \ref{fig:missings} gives an overview of the number of failed tests, i.e.
-tests in which a particular parameter setting did not return a valid result.
-This happens, when the shift to align two signatures is so large, that the remaining overlap is too small to accommodate  windows for validation.
-The problem is therefore exacerbated by a larger validation window.
-Figure \ref{fig:missings} also shows that the number of failed tests is approximately linear in the size of the optimization window.
-
-```{r lms, results='asis'}
-nsnest <- ns %>% filter(wv %in% c(30,50)) %>% group_by(match, wv) %>% nest()
-nsnest <- nsnest %>% mutate(
-  model = data %>% purrr::map(.f = function(d) lm(I(missperc*100)~wo, data=d)),
-  coefs = model %>% purrr::map(.f = broom::tidy)
-)
-tab <- nsnest %>% unnest(coefs) %>% filter(term=="wo") %>% arrange(match) %>%
-  select(-term, -statistic, -p.value)
-knitr::kable(tab, digits=3, caption = "Estimates of the increase in percent of failed tests corresponding to a 100 point increase in the optimization window.", booktabs=TRUE)
-```
-
-```{r wo120, fig.width = 8, fig.height=4.5, fig.cap="For optimization window size 120 and different validation window size how Actual and Type II (beta) errors vary for various chosen nominal alpha levels"}
-wo120 <- errors %>% filter(wo==120)
-wo120 %>% gather(type, error, actual:beta) %>%
-  ggplot(aes(x = wv, y = error, colour = factor(alpha)))  + 
-  geom_line() + geom_point() + facet_wrap(~type, scales="free") +
-  theme_bw()+ 
-  theme(legend.position="bottom")
-```
-
-### Profiles
-
-```{r, fig.cap="Type II errors for different levels of coarseness."}
-e1<-errors2 %>% filter(wv==30)%>% ggplot(aes(x = coarse, y=beta, colour=factor(alpha))) +
-  geom_point() + geom_line() +
-  theme_bw() +
-  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + xlab("Profile Coarseness")+
-  ylab("Type II error")
-
-e2<- errors_sig_c %>% filter(wv==30)%>% ggplot(aes(x = coarse, y=beta, colour=factor(alpha))) +
-  geom_point() + geom_line() +
-  theme_bw() +
-  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2")+ xlab("Signature Coarseness") + ylab("Type II error")
-
-multiplot(e1,e2, cols = 1)
-
-```
-
-### Comparison of total error with signatures and profiles
-
-```{r}
-#errors %>% ggplot(aes(x = wv, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + facet_grid(~wo) +geom_point(data=errors2, aes(shape="profiles"))
-
-total1<- errors_sig_c %>% ggplot(aes(x = coarse, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + geom_line()  + geom_point(data= errors2 %>% filter(wv==30), aes(shape="profiles")) + geom_line(data=errors2 %>% filter(wv==30), aes(shape="profiles")) + theme_bw() +
-  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + xlab("Coarseness") +ylab("Total error")
-
-total2<- wo120 %>% filter(wv>10) %>% ggplot(aes(x = wv, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + geom_line() +geom_point(data=errors2 %>% filter(wv>10) %>% filter(coarse==0.25), aes(shape="profiles")) + geom_line(data=errors2 %>% filter(wv>10)%>% filter(coarse==0.25), aes(shape="profiles")) + theme_bw() +
-  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + ylab("Total error")
-
-```
-
-```{r prof_missings, fig.width=8, fig.height=5, fig.cap="Bottom row shows Total error and Number of failed tests by the window validation size, wv, and ground truth, while the top row shows Total error and Number of failed tests with Coarseness for both profiles and signatures"}
-nsp <- allp %>% group_by(wv, wo, match, coarse) %>% summarize(
-  n_ss = mean(same_shift_n),
-  n_ds = mean(diff_shift_n),
-  zero_ss = sum(same_shift_n==0),
-  zero_ds = sum(diff_shift_n==0),
-  miss = sum(is.na(p_value)),
-  missperc = miss/85491*100
-)
-
-ns_sig_c <- all_sig_c %>% group_by(wv, wo, match, coarse) %>% summarize(
-  n_ss = mean(same_shift_n),
-  n_ds = mean(diff_shift_n),
-  zero_ss = sum(same_shift_n==0),
-  zero_ds = sum(diff_shift_n==0),
-  miss = sum(is.na(p_value)),
-  missperc = miss/85491*100
-)
-
-#ns1<- ns[which(ns$wv == c(30,50), arr.ind = T),]
-miss_1<- nsp %>% filter(wv == 30) %>%
-#ns %>% filter(wv == c(30, 50)) %>%
-  ggplot(aes(x = coarse, y = missperc, shape= "profile", colour = factor(match))) + 
-  #geom_smooth(method="auto", size=.5, se=FALSE) +
-  geom_point(size = 2) + 
-  theme_bw() +
-  ylab("Percent of failed tests.") +
-  xlab("Coarseness") +
-  scale_colour_brewer("Match", palette="Set1") + geom_point(data=ns_sig_c  %>% filter(wv %in% c(30,50)) , aes(shape="signature"), size = 2)# + scale_colour_brewer("Match", palette="Set2") #+geom_smooth(method="auto", size=.5, se=FALSE)
-  #scale_shape_discrete("Size of \nvalidation\nwindow, #wv")
-
-miss_2<- nsp %>% filter(coarse == 0.25) %>%
-#ns %>% filter(wv == c(30, 50)) %>%
-  ggplot(aes(x = wv, y = missperc, shape= "profile", colour = factor(match))) + 
-  #geom_smooth(method="auto", size=.5, se=FALSE) +
-  geom_point(size = 2) + 
-  theme_bw() +
-  ylab("Percent of failed tests.") +
-  xlab("wv") +
-  scale_colour_brewer("Match", palette="Set1") + geom_point(data=ns %>% filter(wo == 120)  %>% filter(coarse == 1) , aes(shape="signature"), size = 2) #+ scale_colour_brewer("Match", palette="Set2")
-
-multiplot(plotlist = list(total1, total2, miss_1, miss_2), cols = 2)#, matrix(c(1,2,3,3), nrow=2, byrow=TRUE))
-```
-
-## Discussion
-
-## Results of Chumbley score
+# Introduction and Results
 
@@
 
-An objective analysis of signatures (pre-processed markings) and profiles (raw markings) of bullet lands for all Hamby-44 and Hamby-252 scans pairwise land to land comparisons (a total of 85,491 comparisons)\citep{hamby} made available through the NIST ballistics database \citep{nist}(Zheng 2016) was done to identify the effects of the two comparison windows and coarseness parameter on the error rates in as proposed in the adjusted chumbley algorithm for toolmarks by \citet{hadler}.
+An objective analysis of signatures (pre-processed markings) and profiles (raw markings) of bullet lands for all Hamby-44 and Hamby-252 scans pairwise land to land comparisons (a total of 85,491 comparisons)\citep{hamby} made available through the NIST ballistics database \citep{nist} was done to identify the effects of the two comparison windows and coarseness parameter on the error rates in as proposed in the adjusted chumbley algorithm for toolmarks by \citet{hadler}.
 The results suggested that the Nominal type I error $\alpha$ value shows dependency on the size of the window of optimization and the window of validation.
@@
 Finally it needs to be noted that the current version of the adjusted chumbley algorithm seems to falls short when compared to other machine-learning based methods \citet{aoas}, and some level of modification to the deterministic algorithm needs to be identified and tested that would reduces the number of incorrect classifications.
-
-# Algorithm Modification
-
-The proposed algorithm modification is at the same shift step which comes after the shift distance is identified by finding two windows in the two markings that have the highest correlation.
-
-The modification allows for a "wiggle" room in the second marking of the two sets of markings being compared.
-This means that each set of same shift windows that are being compared, the second marking will window that is under consideration is allowd to move a little towards the left and a little towards the right.
-
-This gives us a new set of 4 windows, 2 to the left and 2 to the right of the original comparison window.
-Then the correlations for each one of these 5 windows with the window under consideration of the 1st marking is retrieved (from the validation step correlation matrix) or computed.
-
-Then the window that has the maximum correlation (from the set of 5 windows in the 2nd marking) with the window of the 1st marking is chosen.
-
-This new maximum correlation is then used to compute the U-statistics using the same method as before.
-
-## Expected advantage of modification
-
-We are expecting that this modification would improve the type I and II errors for the good. Less number of false negatives and false positives.
-The reason for this expectation is many a times the bullet markings are not made at the same distances for two or more bullets because of the way it comes out of the barrel.
-
-Therefore rigid same shifts might not necessarily compare the right set of windows.
-In the same shift step for one set of windows, allowing the window in the 2nd marking to wiggle left and right and finding the best match to the window in the 1st marking, lets us adjust for situations where markings are compressed or elongated.
-
-## Dependence on the delta of new windows from the original same shift window
-
-From initial tests the amount of movement to get new windows left and right of the original same shift window seems to directly influence the
-the pvalue and U statistic that we look at.
-
-### other questions
-How much movement is reasonable and should be used?
-Is there a way to understand if the 2nd marking in comparison to the 1st marking is compressed or elongated or neither?
-Should the amount of movement selected depend on this compression or elongation?
-or using a static delta movement value justifieable.

### 2018-01-03 short/short.tex
short version added

--- 2018-01-03 short/short.Rmd
@@
----
-title: "Adaption of the Chumbley Score to matching of bullet striation marks"
-authors:
-- affiliation: Department of Statistics, Iowa State University 
-  name: Ganesh Krishnan
-  thanks: The authors gratefully acknowledge ...
-- affiliation: Department of Statistics and CSAFE, Iowa State University 
-  name: Heike Hofmann
-biblio-style: apsr
-output:
-  pdf_document:
-    citation_package: natbib
-    fig_caption: yes
-    keep_tex: yes
-    latex_engine: pdflatex
-    template: template.tex
-  html_document: default
-blinded: 0
-keywords:
-- Bullets, Signatures, Lands, Errors, Known match, Known non-match, Comparisons
-bibliography: bibliography
-<!-- abstract: null -->
----
+% !TeX program = pdfLaTeX
+\documentclass[12pt]{article}
+\usepackage{amsmath}
+\usepackage{amsthm}
+\usepackage{graphicx,psfrag,epsf}
+\usepackage{enumerate}
+\usepackage{natbib}
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage[table]{xcolor}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{hyperref}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+
+\usepackage{url} % not crucial - just used below for the URL
+
+%\pdfminorversion=4
+% NOTE: To produce blinded version, replace "0" with "1" below.
+\newcommand{\blind}{0}
+
+% DON'T change margins - should be 1 inch all around.
+\addtolength{\oddsidemargin}{-.5in}%
+\addtolength{\evensidemargin}{-.5in}%
+\addtolength{\textwidth}{1in}%
+\addtolength{\textheight}{1.3in}%
+\addtolength{\topmargin}{-.8in}%
+
+\newenvironment{definition}[1]% environment name 
+{% begin code 
+  \par\vspace{.75\baselineskip}\noindent 
+  \textbf{Definition (#1)}\begin{itshape}% 
+  \par\vspace{.5\baselineskip}\noindent\ignorespaces 
+}% 
+{% end code 
+  \end{itshape}\ignorespacesafterend 
+}
+
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
+\begin{document}
+
+\def\spacingset#1{\renewcommand{\baselinestretch}%
+{#1}\small\normalsize} \spacingset{1}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\if0\blind
+{
+  \title{\bf Adaption of the Chumbley Score to matching of bullet striation marks}
+
+  \author{
+        Ganesh Krishnan \thanks{The authors gratefully acknowledge \ldots{}} \\
+    Department of Statistics, Iowa State University\\
+     and \\     Heike Hofmann \\
+    Department of Statistics and CSAFE, Iowa State University\\
+      }
+  \maketitle
+} \fi
+
+\if1\blind
+{
+  \bigskip
+  \bigskip
+  \bigskip
+  \begin{center}
+    {\LARGE\bf Adaption of the Chumbley Score to matching of bullet striation marks}
+  \end{center}
+  \medskip
+} \fi
+
+\bigskip
+% \begin{abstract}
+% 
+% \end{abstract}
+
+\noindent%
+{\it Keywords:} Bullets, Signatures, Lands, Errors, Known match, Known non-match,
+Comparisons
+\vfill
+
+\newpage
+\spacingset{1.45} % DON'T change the spacing!
 
@@
 
-<!-- \tableofcontents -->
+\section{Introduction and Results}\label{introduction-and-results}
 
-```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
-library(knitr)
-knitr::opts_chunk$set(
-  message = FALSE,
-  warning = FALSE,
-  error = FALSE, 
-  collapse = TRUE,
-  comment = "",
-  fig.height = 6,
-  fig.width = 6,
-  fig.align = "center",
-  out.width= '\\textwidth',
-  cache = FALSE,
-  fig.path='figures/',
-  echo=FALSE,
-  cache=TRUE
-)
-options(knitr.table.format = "latex")
-library(tidyverse)
-```
-# Introduction and Results
+Compairing pairs of toolmarks with the intention of matching it to a
+tool has been studied relatively more in the past as compared to
+bullets, and \citet{chumbley} have described in their paper an algorithm
+and a deterministic method that compares two toolmarks and come to the
+conclusion if they are from the same tool or not. The method also
+determines the error rates, reduces subject bias and designate the two
+toolmarks as matches or non-matches with respect to a source. This
+project tries to adapt the Chumbley algorithm as modified by
+\citet{hadler}, to bullets which are much smaller in length, width, are
+not flat and curved in the cross-sectional topography as opposed to
+tools like screw driver tips which produces longer and pronounced
+markings. The majority of Bullet profiles and signatures extracted by
+procedures mentioned by \citet{aoas} are almost 1/4 th the size of
+toolmarks as used by \citet{chumbley} or even smaller. Striations on
+Bullets are made on their curved surfaces, whereas the algorithm
+developed by \citet{chumbley} and \citet{hadler} has only been tested
+for flatter and wider surfaces which have negligible curvature.
+Therefore, using methods proposed for toolmarks may need adaptation in
+order to give tangible results for bullets. Moreover, in order to to get
+flat bullet signatures and remove the curvatures some kind of smoothing
+needs to be applied as a pre-step. This needs further investigation as
+to whether the level of smoothing does effect the working of the
+algorithm on Bullets. Another important aspect of adapting the algorithm
+is to find the sizes of the two comparison windows \citet{hadler} that
+minimizes the associated errors. This identification is not obvious as,
+if we go too small in the comparison windows, the unique features of the
+trace segments are lost and seem similar, while too large sizes vastly
+reduces the weight of small features that would otherwise uniquely
+classify a signature and hence identify the region of agreement.
 
-Compairing pairs of toolmarks with the intention of matching it to a tool has been studied relatively more in the past as compared to bullets, and \citet{chumbley} have described in their paper an algorithm and a deterministic method that compares
-two toolmarks and come to the conclusion if they are from the same tool or not.
-The method also determines the error rates, reduces subject bias and designate the two toolmarks as matches or non-matches with respect to a source.
-This project tries to adapt the Chumbley algorithm as modified by \citet{hadler}, to bullets which are much smaller in length, width, are not flat and curved in the cross-sectional topography as opposed to tools like screw driver tips which produces longer and pronounced markings.
-The majority of Bullet profiles and signatures extracted by procedures mentioned by  \citet{aoas} are almost 1/4 th the size of toolmarks as used by \citet{chumbley} or even smaller.
-Striations on Bullets are made on their curved surfaces, whereas the algorithm developed by \citet{chumbley} and \citet{hadler} has only been tested for flatter and wider surfaces which have negligible curvature.
-Therefore, using methods proposed for toolmarks may need adaptation in order to give tangible results for bullets.
-Moreover, in order to to get flat bullet signatures and remove the curvatures some kind of smoothing needs to be applied as a pre-step.
-This needs further investigation as to whether the level of smoothing does effect the working of the algorithm on Bullets.
-Another important aspect of adapting the algorithm is to find the sizes of the two comparison windows  \citet{hadler} that minimizes the associated errors.
-This identification is not obvious as,  if we go too small in the comparison windows, the unique features of the trace segments are lost and seem similar, while too large sizes vastly reduces the weight of small features that would otherwise uniquely classify a signature and hence identify the region of agreement.
+An objective analysis of signatures (pre-processed markings) and
+profiles (raw markings) of bullet lands for all Hamby-44 and Hamby-252
+scans pairwise land to land comparisons (a total of 85,491
+comparisons)\citep{hamby} made available through the NIST ballistics
+database \citep{nist} was done to identify the effects of the two
+comparison windows and coarseness parameter on the error rates in as
+proposed in the adjusted chumbley algorithm for toolmarks by
+\citet{hadler}. The results suggested that the Nominal type I error
+\(\alpha\) value shows dependency on the size of the window of
+optimization and the window of validation. For a given window of
+optimization the actual Type I error is comparable the nominal level
+only for only a select few validation window sizes. A Test Fail, which
+is the percentage of incorrect classifications, depends on whether
+known-match or known non-matches has predictive value, with test results
+from different sources having a much higher chance to fail. On
+conducting an analysis of all known bullet lands using the adjusted
+chumbley algorithm, Type II error was identified to be least bad for
+window of validation 30 and window of optimization 120. In case of
+unsmoothed raw marks (profiles), Type II error increases with the amount
+of smoothing and least for LOWESS smoothing coarseness value about 0.25
+or 0.3. In an effort to identify the level of adaptiveness of the
+algorithm, comparisons were made between signatures and profiles. Their
+comparison with respect to validation window size for a fixed
+optimization window side suggeted that, profiles have a total error
+greater than or equal to the total error of signatures for all sizes of
+validation window. Profiles also fail (i.e.~incorrectly classify
+known-matches and known non-matches) more number of times than
+signatures, which lets us conclude that the behaviour of the algorithm
+for the profiles instead of pre-processed signatures is not better.
+Finally it needs to be noted that the current version of the adjusted
+chumbley algorithm seems to falls short when compared to other
+machine-learning based methods \citet{aoas}, and some level of
+modification to the deterministic algorithm needs to be identified and
+tested that would reduces the number of incorrect classifications.
 
-An objective analysis of signatures (pre-processed markings) and profiles (raw markings) of bullet lands for all Hamby-44 and Hamby-252 scans pairwise land to land comparisons (a total of 85,491 comparisons)\citep{hamby} made available through the NIST ballistics database \citep{nist} was done to identify the effects of the two comparison windows and coarseness parameter on the error rates in as proposed in the adjusted chumbley algorithm for toolmarks by \citet{hadler}.
-The results suggested that the Nominal type I error $\alpha$ value shows dependency on the size of the window of optimization and the window of validation.
-For a given window of optimization the actual Type I error is comparable the nominal level only for only a select few validation window sizes.
-A Test Fail, which is the percentage of incorrect classifications, depends on whether known-match or known non-matches has predictive value, with test results from different sources having a much higher chance to fail.
-On conducting an analysis of all known bullet lands using the adjusted chumbley algorithm, Type II error was identified to be least bad for window of validation 30 and window of optimization 120.
-In case of unsmoothed raw marks (profiles), Type II error increases with the amount of smoothing and least for LOWESS smoothing coarseness value about 0.25 or 0.3.
-In an effort to identify the level of adaptiveness of the algorithm, comparisons were made between signatures and profiles.
-Their comparison with respect to validation window size for a fixed optimization window side suggeted that, profiles have a total error greater than or equal to the total error of signatures for all sizes of validation window.
-Profiles also fail (i.e.
-incorrectly classify known-matches and known non-matches) more number of times than signatures, which lets us conclude that the behaviour of the algorithm for the profiles instead of pre-processed signatures is not better.
-Finally it needs to be noted that the current version of the adjusted chumbley algorithm seems to falls short when compared to other machine-learning based methods \citet{aoas}, and some level of modification to the deterministic algorithm needs to be identified and tested that would reduces the number of incorrect classifications.
+\bibliographystyle{agsm}
+\bibliography{bibliography}
+
+\end{document}

### 2018-01-03 short/short.Rmd
minor corrections

@@
 Another important aspect of adapting the algorithm is to find the sizes of the two comparison windows  \citet{hadler} that minimizes the associated errors.
-This identification is not obvious as,  if we go too small in the comparison windows, the unique features of the trace segments are lost and seem similar, while too large sizes vastly reduces the weight of small features that would otherwise uniquely classify a signature and hence identify the region of agreement.
+This identification is not obvious as,  if we go too small in the comparison windows, the unique features of the trace segments are lost and seem similar, whereas very large sizes vastly reduce the weight of small features that would otherwise uniquely classify a signature and hence identify the region of agreement.
 
-An objective analysis of signatures (pre-processed markings) and profiles (raw markings) of bullet lands for all Hamby-44 and Hamby-252 scans pairwise land to land comparisons (a total of 85,491 comparisons)\citep{hamby} made available through the NIST ballistics database \citep{nist} was done to identify the effects of the two comparison windows and coarseness parameter on the error rates in as proposed in the adjusted chumbley algorithm for toolmarks by \citet{hadler}.
-The results suggested that the Nominal type I error $\alpha$ value shows dependency on the size of the window of optimization and the window of validation.
-For a given window of optimization the actual Type I error is comparable the nominal level only for only a select few validation window sizes.
+An objective analysis of signatures (pre-processed markings) and profiles (raw markings) of bullet lands for all Hamby-44 and Hamby-252 scans pairwise land to land comparisons (a total of 85,491 comparisons)\citep{hamby} made available through the NIST ballistics database \citep{nist} was done to identify the effects of the two comparison windows and coarseness parameter on the error rates as proposed in the adjusted chumbley algorithm for toolmarks by \citet{hadler}.
+The results suggest that the Nominal type I error $\alpha$ value shows dependence on the size of the window of optimization and the window of validation.
+For a given window of optimization the actual Type I error is comparable to the nominal level for only a select few validation window sizes.
 A Test Fail, which is the percentage of incorrect classifications, depends on whether known-match or known non-matches has predictive value, with test results from different sources having a much higher chance to fail.
@@
 In an effort to identify the level of adaptiveness of the algorithm, comparisons were made between signatures and profiles.
-Their comparison with respect to validation window size for a fixed optimization window side suggeted that, profiles have a total error greater than or equal to the total error of signatures for all sizes of validation window.
+Their comparison with respect to validation window size for a fixed optimization window size suggested that, profiles have a total error greater than or equal to the total error of signatures for all sizes of validation window.
 Profiles also fail (i.e.
 incorrectly classify known-matches and known non-matches) more number of times than signatures, which lets us conclude that the behaviour of the algorithm for the profiles instead of pre-processed signatures is not better.
-Finally it needs to be noted that the current version of the adjusted chumbley algorithm seems to falls short when compared to other machine-learning based methods \citet{aoas}, and some level of modification to the deterministic algorithm needs to be identified and tested that would reduces the number of incorrect classifications.
+Finally it should be noted that the current version of the adjusted chumbley algorithm seems to fall short when compared to other machine-learning based methods \citet{aoas}, and some level of modification to the deterministic algorithm needs to be identified and tested that would reduce the number of incorrect classifications.

### 2018-01-03 doc_v2/doc_v2.Rmd
directory change

--- 2018-01-03 p2/p2.Rmd
@@
 
-## Results of Chumbley score
-- Nominal alpha value shows dependency on the size of the window of optimization
-- Test Fail depends on whether known-match or known non-matches has predictive value
-- Type II error least bad for window of validation 30 and window of optimization 120
-- Type II error for profiles increases with the amount of smoothing and least for LOWESS smoothing coarseness value about 0.25 or 0.3.
-- Profiles in comparison with Signatures seem to perform worse with total error either equal or more for all sizes of validation window.
--
-
 # Algorithm Modification

### 2018-01-03 short/short.Rmd
altered incorrect line about test

@@
 An objective analysis of signatures (pre-processed markings) and profiles (raw markings) of bullet lands for all Hamby-44 and Hamby-252 scans pairwise land to land comparisons (a total of 85,491 comparisons)\citep{hamby} made available through the NIST ballistics database \citep{nist} was done to identify the effects of the two comparison windows and coarseness parameter on the error rates as proposed in the adjusted chumbley algorithm for toolmarks by \citet{hadler}.
-The results suggest that the Nominal type I error $\alpha$ value shows dependence on the size of the window of optimization and the window of validation.
-For a given window of optimization the actual Type I error is comparable to the nominal level for only a select few validation window sizes.
-A Test Fail, which is the percentage of incorrect classifications, depends on whether known-match or known non-matches has predictive value, with test results from different sources having a much higher chance to fail.
+The results suggest that the Nominal type I error $\alpha$ value shows dependence on the size of the window of optimization.
+For a given window of optimization the actual Type I error is comparable to the nominal level for only a select few validation window sizes and for comparable validation window sizes of 30 and 50 as done here, the actual type I error does not seem to vary as much as it varies with the optimization window sizes .
+A Test Fail, i.e.
+tests in which a particular parameter setting did not return a valid result, happens, when the shift to align two signatures is so large, that the remaining overlap is too small to accommodate  windows for validation, depends on whether known-match or known non-matches has predictive value, with test results from different sources having a much higher chance to fail.
 On conducting an analysis of all known bullet lands using the adjusted chumbley algorithm, Type II error was identified to be least bad for window of validation 30 and window of optimization 120.
@@
 In an effort to identify the level of adaptiveness of the algorithm, comparisons were made between signatures and profiles.
-Their comparison with respect to validation window size for a fixed optimization window size suggested that, profiles have a total error greater than or equal to the total error of signatures for all sizes of validation window.
-Profiles also fail (i.e.
-incorrectly classify known-matches and known non-matches) more number of times than signatures, which lets us conclude that the behaviour of the algorithm for the profiles instead of pre-processed signatures is not better.
+Their comparison with respect to validation window size for a fixed optimization window size suggested that, profiles have a total error (i.e all incorrect classification of known-matches and known non-matches) greater than or equal to the total error of signatures for all sizes of validation window.
+Profiles also fail more number of times than signatures in a test fail (for different coarseness keeping windows fixed and also for different validation windows keeping coarseness fixed) which lets us conclude that the behaviour of the algorithm for the profiles instead of pre-processed signatures is not better.
 Finally it should be noted that the current version of the adjusted chumbley algorithm seems to fall short when compared to other machine-learning based methods \citet{aoas}, and some level of modification to the deterministic algorithm needs to be identified and tested that would reduce the number of incorrect classifications.

### 2018-01-03 doc_v2/doc_v2.Rmd
formatting

@@
 \tableofcontents
-
+\newpage
 ```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
@@
 library(tidyverse)
+library(kableExtra)
 ```
@@
 
-## Problem statement and motivation
+## Motivation
 
@@
 \citep{hadler} on the other hand, in the improved version of the algorithm first described by \citep{chumbley} found false-positives for 0 cases and 3 false-negatives (or 6%) for 50 comparisons of knwon matches.
+```{r}
+a<- matrix(NA, ncol = 3, nrow = 2)
+toolmark.falsepositive<- data.frame(a)
+colnames(toolmark.falsepositive)<- c("Classification", "Match", "Non-Match")
+toolmark.falsepositive$Classification<- c("Match", "Non-Match")
+toolmark.falsepositive$Match<- c(47,0)
+toolmark.falsepositive$`Non-Match`<- c(3,50)
+
+knitr::kable(toolmark.falsepositive, digits=0, format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("hold"))
+```
+
+## Potential limitations of the Chumbley Score adaptation to bullets 
+Bullets are much smaller in length, width, are not flat and curved in the cross-sectional topography as opposed to tools like screw driver tips which produced longer and pronounced markings.
+This means the makings made by barrels in comparison to toolmarks may have a problem in distinctiveness.
+The majority of Bullet profiles and signatures extracted by procedures mentioned by  \citet{aoas} are almost 1/4 th the size of toolmarks as used by \citet{chumbley} or even smaller.
+Striations on Bullets are made on their curved surfaces, whereas the algorithm developed by \citet{chumbley} and \citet{hadler} has only been tested for flatter and wider surfaces which have negligible curvature.
+Therefore, using methods proposed for toolmarks may need adaptation in order to give tangible results for bullets.
+Moreover, in order to to get flat bullet signatures and remove the curvatures some kind of smoothing needs to be applied as a pre-step which needs further investigation as to whether the level of smoothing does effect the working of the algorithm on Bullets.
+
+Also when in the optimization step, the Window of optimization for bullets will be shorter as the signatures are smaller.
+The idea is to keep the number of windows of optimization sufficiently large, which means shorter Trace segments (partition of signature or toolmark with length = size of window of optimization) that lets us compare smaller segments of one signature to another.
+This introduces a problem as, if we go too small in the window of optimization, the unique features of the trace segments are lost and seem similar, while too large sizes vastly reduces the weight of small features that would otherwise uniquely classify a signature and hence identify the region of agreement.
+
+Thus the Window of Optimization has a direct influence on whether we are Falsely rejecting the null when it is true (Type I) or Falsely accepting the null when it is false (Type II) making identification of the optimum size of window of optimization very important.
+This raises important questions about what parameter settings need to be chosen for bullet land comparison.
+Something similar is done in \citet{afte-chumbley} for toolmark comparisons of slip-joint pliers where optimum window sizes are determined.
+There is a need to figure out the best parameter settings which minimizes the errors for unsmoothed markings and pre-processed signatures as determined by \citet{aoas}.
+An analysis of these error rates and comparison with other methods will help us understand the adaptibility of the chumbley score to bullets.
+
 \begin{figure}
@@
 ```{r win-comparison, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap=" The two plots on the left show how the same shift behaves in case of a matching pair and the two plots on the right show how the different shift behaves in case of a matching pair."}
-a<- matrix(NA, ncol = 3, nrow = 2)
-toolmark.falsepositive<- data.frame(a)
-colnames(toolmark.falsepositive)<- c("Classification", "Match", "Non-Match")
-toolmark.falsepositive$Classification<- c("Match", "Non-Match")
-toolmark.falsepositive$Match<- c(47,0)
-toolmark.falsepositive$`Non-Match`<- c(3,50)
-
-knitr::kable(toolmark.falsepositive, digits=0, format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("hold"))
 
@@
 
-## Potential limitations of the Chumbley Score
-
-\hh{in the application to striation marks on bullet lands: smaller in width and curved - need to adjust parameter settings; something similar is done in \citet{afte-chumbley} for toolmark comparisons of slip-joint pliers.
-XXX work in }
-
-Bullets are much smaller in length, width, are not flat and curved in the cross-sectional topography as opposed to tools like screw driver tips which produced longer and pronounced markings.
-This means the makings made by barrels in comparison to toolmarks may have a problem in distinctiveness.
-The majority of Bullet profiles and signatures extracted by procedures mentioned by  \citet{aoas} are almost 1/4 th the size of toolmarks as used by \citet{chumbley} or even smaller.
-Striations on Bullets are made on their curved surfaces, whereas the algorithm developed by \citet{chumbley} and \citet{hadler} has only been tested for flatter and wider surfaces which have negligible curvature.
-Therefore, using methods proposed for toolmarks may need adaptation in order to give tangible results for bullets.
-Moreover, in order to to get flat bullet signatures and remove the curvatures some kind of smoothing needs to be applied as a pre-step which needs further investigation as to whether the level of smoothing does effect the working of the algorithm on Bullets.
-
-Also when in the optimization step, the Window of optimization for bullets will be shorter as the signatures are smaller.
-The idea is to keep the number of windows of optimization sufficiently large, which means shorter Trace segments (partition of signature or toolmark with length = size of window of optimization) that lets us compare smaller segments of one signature to another.
-
-This introduces a problem as, if we go too small in the window of optimization, the unique features of the trace segments are lost and seem similar, while too large sizes vastly reduces the weight of small features that would otherwise uniquely classify a signature and hence identify the region of agreement.
-
-Thus the Window of Optimization has a direct influence on whether we are
-Falsely rejecting the null when it is true (Type I) or Falsely accepting the null when it is false (Type II) making identification of the optimum size of window of optimization very important.
-
 ## Testing setup
@@
 
-We used the adjusted Chumbley method as proposed in \citet{hadler} and implemented in the R package `toolmaRk` \citep{toolmark} on all pairwise land-to-land comparisons of the Hamby scans (a total of 85,491 comparisons) with the pairwise sets for the comparisons given in the following table.
-
-```{r param, echo=FALSE, results='asis'}
-param_setting<- all %>% select(wo, wv) %>% group_by(wo, wv) %>% tally() %>% select(wo,wv)
-#param_setting<- data.frame(matrix(param_setting, nrow = 16))
-#t(param_setting)
-#param_setting %>% tally() 
-library(kableExtra)
-knitr::kable(t(param_setting)[,1:17], digits=0, caption = "Overview of parameter settings used for optimization and validation windows for bullet land signatures.", format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("scale_down", "hold"))
-
-knitr::kable(t(param_setting)[,18:33], digits=0, format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("scale_down","hold"))
-
-# library(xtable)
-# strCaption <- "Overview of parameter settings used for optimization and validation windows for bullet land signatures."
-#   
-#   # set up xtable output
-# print(xtable(t(param_setting)[,1:17], 
-#              caption = strCaption, label = ""),
-#       include.rownames = TRUE,
-#       include.colnames = FALSE,
-#       caption.placement = "top",
-#       comment=FALSE,
-#       floating=FALSE
-#       )
-# print(xtable(t(param_setting)[,17:33],label = ""),
-#       include.rownames = FALSE,
-#       include.colnames = FALSE,
-#       caption.placement = "bottom",
-#       comment=FALSE,
-#       floating=TRUE
-#       #table.placement = "!ht"
-#       )
-```
-
 ### Profiles
@@
 
-The idea therefore is to first use these raw values of the profile directly in the chumbley algorithm, and see how the algorithm performs for different coarseness values (smoothing parameter as referred in the function lowess used in the chumbley algorithm)
+The idea therefore is to first use these raw values of the profile directly in the chumbley algorithm, and see how the algorithm performs for different coarseness values (smoothing parameter as referred in the function lowess used in the chumbley algorithm).
+
+\pagebreak
 
@@
 #knitr::kable(a27, digits=0, caption = "Window of Optimization 120 and Window of Validation 50", format = "latex" ,booktabs=TRUE)%>% kableExtra::kable_styling(latex_options = c("hold"))
+```
+
+```{r param, echo=FALSE, results='asis'}
+param_setting<- all %>% select(wo, wv) %>% group_by(wo, wv) %>% tally() %>% select(wo,wv)
+#param_setting<- data.frame(matrix(param_setting, nrow = 16))
+#t(param_setting)
+#param_setting %>% tally() 
+library(kableExtra)
+#knitr::kable(t(param_setting)[,1:17], digits=0, caption = "Overview of parameter settings used for optimization and validation windows for bullet land signatures.", format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("scale_down", "hold"))
+
+par_tab1<-t(param_setting)[,1:18]
+par_tab2<-t(param_setting)[,19:33]
+knitr::kable(list(par_tab1,par_tab2), digits=0,caption = "Overview of parameter settings used for optimization and validation windows for bullet land signatures.", format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("scale_down","hold"))%>% kableExtra::kable_styling(latex_options = c("scale_down","hold")) #%>% group_rows("",1,2) %>% group_rows("",3,4)
+
 ```
@@
 
+We used the adjusted Chumbley method as proposed in \citet{hadler} and implemented in the R package `toolmaRk` \citep{toolmark} on all pairwise land-to-land comparisons of the Hamby scans (a total of 85,491 comparisons) with the pairwise sets for the comparisons given in the table \ref{tab:param}.
+
 ### Signatures
@@
 Figure \ref{fig:type2} gives an overview of type II error rates observed when varying the window size in the optimization step.
+Two levels of validation window size 30 and 50 were chosen as to compare the error rates for different nominal type I errors.
+We notice that the trends for these nominal type I errors are similar and in most cases a validation window of 50 has higher type II error than for 30.
+A change in this trend is seen for a 0.05 $\alpha$ level, although the difference between the two windows is very small for this case.
+We can also notice an obvious trend of increase in the Type II error as the window of optimization increases and see a minimum around the optimization window size of 120 pixels.
+Hence we are inclined to choose a smaller validation window size and optimization window as 120.
 
@@
 Figure \ref{fig:type1} compares nominal (fixed) type I error and actually observed type I errors for the parameter settings in table \ref{tab:param}.
-With an increasing size of the window used in the optimization step the observed type I error rate decreases (slighty).
+With an increasing size of the window used in the optimization step the observed type I error rate decreases (slighty)or As the optimization window increase the observed type I error rate gets smaller, while, A smaller validation window tends to be associated with a higher type I error rate.
+This can be better imagined for a given window of optimization, when the actual Type I error is comparable to the nominal level for only a select few validation window sizes and for comparable validation window sizes of 30 and 50 as done here, the actual type I error although increases negligibly, it does not seem to vary as much as it varies with the optimization window sizes.
 This might be related to the increasing number of tests that fail for larger window sizes, in particular for non-matching striae (see fig \ref{fig:missings}).
@@
 ```{r type1, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap="Comparison of observed and nominal type I error rates  across a range of window sizes for optimization $wo$.
-The horizontal line in each facet indicates the nominal type I error rate.
-As the optimization window increase the observed type I error rate gets smaller.
-A smaller validation window tends to be associated with a higher type I error rate."}
+The horizontal line in each facet indicates the nominal type I error rate."}
 errors %>% filter(wv %in% c(30,50)) %>%

### 2018-01-03 doc_v2/doc_v2.Rmd
.

@@
 
-```{r }
-
-#knitr::kable(a27, digits=0, caption = "Window of Optimization 320 and Window of Validation 50", format = "latex" ,booktabs=TRUE)%>% kableExtra::kable_styling(latex_options = c("hold"))
-
-```
-
-```{r}
-
-#knitr::kable(a27, digits=0, caption = "Window of Optimization 120 and Window of Validation 50", format = "latex" ,booktabs=TRUE)%>% kableExtra::kable_styling(latex_options = c("hold"))
-```
-
-```{r param, echo=FALSE, results='asis'}
+# Results
+We used the adjusted Chumbley method as proposed in \citet{hadler} and implemented in the R package `toolmaRk` \citep{toolmark} on all pairwise land-to-land comparisons of the Hamby scans (a total of 85,491 comparisons) with the pairwise sets for the comparisons given in the table \ref{tab:param}.
+```{r param}
 param_setting<- all %>% select(wo, wv) %>% group_by(wo, wv) %>% tally() %>% select(wo,wv)
@@
 ```
-
-# Results
-
-We used the adjusted Chumbley method as proposed in \citet{hadler} and implemented in the R package `toolmaRk` \citep{toolmark} on all pairwise land-to-land comparisons of the Hamby scans (a total of 85,491 comparisons) with the pairwise sets for the comparisons given in the table \ref{tab:param}.
 

### 2018-01-03 doc_v2/doc_v2.Rmd
.

@@
 
-```{r missings, fig.width=8, fig.height=5, fig.cap="Number of failed tests by the window optimization size, wo, and ground truth.
-Test results from different sources have a much higher chance to fail, raising the question, whether failed tests should be treated as rejections of the null hypothesis of same source."}
+```{r missings, fig.width=8, fig.height=5, fig.cap="Number of failed tests by the window optimization size, wo, and ground truth."}
 ns <- all %>% group_by(wv, wo, match) %>% summarize(
@@
 Figure \ref{fig:missings} also shows that the number of failed tests is approximately linear in the size of the optimization window.
+Test results from different sources have a much higher chance to fail, raising the question, whether failed tests should be treated as rejections of the null hypothesis of same source.
+For known non-matches there is a higher possibility that in the optimization pair of windows where cross-correlations are maximum are too far apart, and same shifts of this order hit the end of the signature.
+
+Another reason for failed tests can be incorrect identification of maximum correlation windows in the optimization step as seen in \ref{fig:prof_missings}(c) because of the level of smoothing, as too much smoothing would subdue intricate features that might otherwise help in the correlation calculations and correct identification of maximum correlation windows irrespective of the size.
+This would again cause a simiar effect as explained for \ref{fig:missings} with validation windows, irrespective of size, during the shifts end up at the ends of the markings resulting in an invalid calculation and failed comparison attempt.
+
+In \ref{fig:prof_missings}(c) and (d), we compare profiles and signatures on the basis of number of failed tests.
+The profiles chosen for \ref{fig:prof_missings}(d) have a constant coarseness of 0.25 and window of optimization as 120.
+The signatures in this case are not smoothed using the chumbley algorithm step of LOWESS smoothing.
+Instead signatures are used as calculated by \citet{aoas}.
+The smoothing in these signatures were determined and fixed on the basis of their performance in the random forest based algorithm proposed by \citet{aoas}.
+The comparison of profiles and signatures with variation of validation window size therefore is made on even footing.
+The trends are similar to \ref{fig:missings} in the sense that for known non-matches the number of failed tests are more for both signatures and profiles and increasing linearly with the validation window size.
+The problem is however, worse for profiles which has higher number of failed tests than signatures for all validation windows.
 
@@
 total1<- errors_sig_c %>% ggplot(aes(x = coarse, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + geom_line()  + geom_point(data= errors2 %>% filter(wv==30), aes(shape="profiles")) + geom_line(data=errors2 %>% filter(wv==30), aes(shape="profiles")) + theme_bw() +
-  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + xlab("Coarseness") +ylab("Total error")
+  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + xlab("Coarseness") +ylab("Total error") + ggtitle(label = "(a)")
 
 total2<- wo120 %>% filter(wv>10) %>% ggplot(aes(x = wv, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + geom_line() +geom_point(data=errors2 %>% filter(wv>10) %>% filter(coarse==0.25), aes(shape="profiles")) + geom_line(data=errors2 %>% filter(wv>10)%>% filter(coarse==0.25), aes(shape="profiles")) + theme_bw() +
-  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + ylab("Total error")
+  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + ylab("Total error") + ggtitle(label = "(b)")
 
@@
   xlab("Coarseness") +
-  scale_colour_brewer("Match", palette="Set1") + geom_point(data=ns_sig_c  %>% filter(wv %in% c(30,50)) , aes(shape="signature"), size = 2)# + scale_colour_brewer("Match", palette="Set2") #+geom_smooth(method="auto", size=.5, se=FALSE)
+  scale_colour_brewer("Match", palette="Set1") + geom_point(data=ns_sig_c  %>% filter(wv %in% c(30,50)) , aes(shape="signature"), size = 2) + ggtitle(label = "(c)")
+# + scale_colour_brewer("Match", palette="Set2") #+geom_smooth(method="auto", size=.5, se=FALSE) 
   #scale_shape_discrete("Size of \nvalidation\nwindow, #wv")
@@
   xlab("wv") +
-  scale_colour_brewer("Match", palette="Set1") + geom_point(data=ns %>% filter(wo == 120)  %>% filter(coarse == 1) , aes(shape="signature"), size = 2) #+ scale_colour_brewer("Match", palette="Set2")
+  scale_colour_brewer("Match", palette="Set1") + geom_point(data=ns %>% filter(wo == 120)  %>% filter(coarse == 1) , aes(shape="signature"), size = 2) + ggtitle(label = "(d)")#+ scale_colour_brewer("Match", palette="Set2")
 

### 2018-01-03 doc_v2/doc_v2.Rmd
.

@@
 
-Another reason for failed tests can be incorrect identification of maximum correlation windows in the optimization step as seen in \ref{fig:prof_missings}(c) because of the level of smoothing, as too much smoothing would subdue intricate features that might otherwise help in the correlation calculations and correct identification of maximum correlation windows irrespective of the size.
-This would again cause a simiar effect as explained for \ref{fig:missings} with validation windows, irrespective of size, during the shifts end up at the ends of the markings resulting in an invalid calculation and failed comparison attempt.
+Another reason for failed tests can be incorrect identification of maximum correlation windows in the optimization step as seen in figure \ref{fig:prof_missings}(c) because of the level of smoothing, as too much smoothing would subdue intricate features that might otherwise help in the correlation calculations and correct identification of maximum correlation windows irrespective of the size.
+This would again cause a simiar effect as explained for figure \ref{fig:missings} with validation windows, irrespective of size, during the shifts end up at the ends of the markings resulting in an invalid calculation and failed comparison attempt.
 
-In \ref{fig:prof_missings}(c) and (d), we compare profiles and signatures on the basis of number of failed tests.
-The profiles chosen for \ref{fig:prof_missings}(d) have a constant coarseness of 0.25 and window of optimization as 120.
+In figure \ref{fig:prof_missings}(c) and (d), we compare profiles and signatures on the basis of number of failed tests.
+The profiles chosen for figure \ref{fig:prof_missings}(d) have a constant coarseness of 0.25 and window of optimization as 120.
 The signatures in this case are not smoothed using the chumbley algorithm step of LOWESS smoothing.
@@
 The comparison of profiles and signatures with variation of validation window size therefore is made on even footing.
-The trends are similar to \ref{fig:missings} in the sense that for known non-matches the number of failed tests are more for both signatures and profiles and increasing linearly with the validation window size.
+The trends are similar to figure \ref{fig:missings} in the sense that for known non-matches the number of failed tests are more for both signatures and profiles and increasing linearly with the validation window size.
 The problem is however, worse for profiles which has higher number of failed tests than signatures for all validation windows.

### 2018-01-04 doc_v2/doc_v2.Rmd
alignment

@@
 
-```{r missings, fig.width=8, fig.height=5, fig.cap="Number of failed tests by the window optimization size, wo, and ground truth."}
+```{r wo120, fig.width = 8, fig.height=4.5, fig.cap="For optimization window size 120 and different validation window size how Actual and Type II (beta) errors vary for various chosen nominal alpha levels"}
+wo120 <- errors %>% filter(wo==120)
+wo120 %>% gather(type, error, actual:beta) %>%
+  ggplot(aes(x = wv, y = error, colour = factor(alpha)))  + 
+  geom_line() + geom_point() + facet_wrap(~type, scales="free") +
+  theme_bw()+ 
+  theme(legend.position="bottom")
+```
+
+```{r }
 ns <- all %>% group_by(wv, wo, match) %>% summarize(
@@
 
-#ns1<- ns[which(ns$wv == c(30,50), arr.ind = T),]
-ns %>% filter(wv %in% c(30, 50)) %>%
-#ns %>% filter(wv == c(30, 50)) %>%
-  ggplot(aes(x = wo, y = missperc, shape=factor(wv), colour = factor(match))) + 
-  geom_smooth(method="lm", size=.5, se=FALSE) +
-  geom_point() + 
-  theme_bw() +
-  ylab("Percent of failed tests.") +
-  xlab("Size of optimization window, wo") +
-  scale_colour_brewer("Match", palette="Set1") +
-  scale_shape_discrete("Size of \nvalidation\nwindow, wv")
-```
-Figure \ref{fig:missings} gives an overview of the number of failed tests, i.e.
-tests in which a particular parameter setting did not return a valid result.
-This happens, when the shift to align two signatures is so large, that the remaining overlap is too small to accommodate  windows for validation.
-The problem is therefore exacerbated by a larger validation window.
-Figure \ref{fig:missings} also shows that the number of failed tests is approximately linear in the size of the optimization window.
-Test results from different sources have a much higher chance to fail, raising the question, whether failed tests should be treated as rejections of the null hypothesis of same source.
-For known non-matches there is a higher possibility that in the optimization pair of windows where cross-correlations are maximum are too far apart, and same shifts of this order hit the end of the signature.
-
-Another reason for failed tests can be incorrect identification of maximum correlation windows in the optimization step as seen in figure \ref{fig:prof_missings}(c) because of the level of smoothing, as too much smoothing would subdue intricate features that might otherwise help in the correlation calculations and correct identification of maximum correlation windows irrespective of the size.
-This would again cause a simiar effect as explained for figure \ref{fig:missings} with validation windows, irrespective of size, during the shifts end up at the ends of the markings resulting in an invalid calculation and failed comparison attempt.
-
-In figure \ref{fig:prof_missings}(c) and (d), we compare profiles and signatures on the basis of number of failed tests.
-The profiles chosen for figure \ref{fig:prof_missings}(d) have a constant coarseness of 0.25 and window of optimization as 120.
-The signatures in this case are not smoothed using the chumbley algorithm step of LOWESS smoothing.
-Instead signatures are used as calculated by \citet{aoas}.
-The smoothing in these signatures were determined and fixed on the basis of their performance in the random forest based algorithm proposed by \citet{aoas}.
-The comparison of profiles and signatures with variation of validation window size therefore is made on even footing.
-The trends are similar to figure \ref{fig:missings} in the sense that for known non-matches the number of failed tests are more for both signatures and profiles and increasing linearly with the validation window size.
-The problem is however, worse for profiles which has higher number of failed tests than signatures for all validation windows.
-
-```{r lms, results='asis'}
-nsnest <- ns %>% filter(wv %in% c(30,50)) %>% group_by(match, wv) %>% nest()
-nsnest <- nsnest %>% mutate(
-  model = data %>% purrr::map(.f = function(d) lm(I(missperc*100)~wo, data=d)),
-  coefs = model %>% purrr::map(.f = broom::tidy)
-)
-tab <- nsnest %>% unnest(coefs) %>% filter(term=="wo") %>% arrange(match) %>%
-  select(-term, -statistic, -p.value)
-knitr::kable(tab, digits=3, caption = "Estimates of the increase in percent of failed tests corresponding to a 100 point increase in the optimization window.", booktabs=TRUE)
-```
-
-```{r wo120, fig.width = 8, fig.height=4.5, fig.cap="For optimization window size 120 and different validation window size how Actual and Type II (beta) errors vary for various chosen nominal alpha levels"}
-wo120 <- errors %>% filter(wo==120)
-wo120 %>% gather(type, error, actual:beta) %>%
-  ggplot(aes(x = wv, y = error, colour = factor(alpha)))  + 
-  geom_line() + geom_point() + facet_wrap(~type, scales="free") +
-  theme_bw()+ 
-  theme(legend.position="bottom")
-```
-
-### Profiles
-
-```{r, fig.cap="Type II errors for different levels of coarseness."}
-e1<-errors2 %>% filter(wv==30)%>% ggplot(aes(x = coarse, y=beta, colour=factor(alpha))) +
-  geom_point() + geom_line() +
-  theme_bw() +
-  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + xlab("Profile Coarseness")+
-  ylab("Type II error")
-
-e2<- errors_sig_c %>% filter(wv==30)%>% ggplot(aes(x = coarse, y=beta, colour=factor(alpha))) +
-  geom_point() + geom_line() +
-  theme_bw() +
-  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2")+ xlab("Signature Coarseness") + ylab("Type II error")
-
-multiplot(e1,e2, cols = 1)
-
-```
-
-### Comparison of total error with signatures and profiles
-
-```{r}
-#errors %>% ggplot(aes(x = wv, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + facet_grid(~wo) +geom_point(data=errors2, aes(shape="profiles"))
-
-total1<- errors_sig_c %>% ggplot(aes(x = coarse, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + geom_line()  + geom_point(data= errors2 %>% filter(wv==30), aes(shape="profiles")) + geom_line(data=errors2 %>% filter(wv==30), aes(shape="profiles")) + theme_bw() +
-  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + xlab("Coarseness") +ylab("Total error") + ggtitle(label = "(a)")
-
-total2<- wo120 %>% filter(wv>10) %>% ggplot(aes(x = wv, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + geom_line() +geom_point(data=errors2 %>% filter(wv>10) %>% filter(coarse==0.25), aes(shape="profiles")) + geom_line(data=errors2 %>% filter(wv>10)%>% filter(coarse==0.25), aes(shape="profiles")) + theme_bw() +
-  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + ylab("Total error") + ggtitle(label = "(b)")
-
-```
-
-```{r prof_missings, fig.width=8, fig.height=5, fig.cap="Bottom row shows Total error and Number of failed tests by the window validation size, wv, and ground truth, while the top row shows Total error and Number of failed tests with Coarseness for both profiles and signatures"}
 nsp <- allp %>% group_by(wv, wo, match, coarse) %>% summarize(
@@
 
+```
+
+```{r missings, fig.width=5, fig.height=2, fig.cap="Number of failed tests by the window optimization size, wo, and ground truth.",fig.align = "center"}
+#ns1<- ns[which(ns$wv == c(30,50), arr.ind = T),]
+ns %>% filter(wv %in% c(30, 50)) %>%
+#ns %>% filter(wv == c(30, 50)) %>%
+  ggplot(aes(x = wo, y = missperc, shape=factor(wv), colour = factor(match)), size=.3) + 
+  geom_smooth(method="lm", size=.2, se=FALSE) +
+  geom_point() + 
+  theme_bw() +
+  ylab("Percent of failed tests.") +
+  xlab("Size of optimization window, wo") +
+  scale_colour_brewer("Match", palette="Set1") +
+  scale_shape_discrete("Size of \nvalidation\nwindow, wv")
+```
+
+```{r lms, results='asis'}
+nsnest <- ns %>% filter(wv %in% c(30,50)) %>% group_by(match, wv) %>% nest()
+nsnest <- nsnest %>% mutate(
+  model = data %>% purrr::map(.f = function(d) lm(I(missperc*100)~wo, data=d)),
+  coefs = model %>% purrr::map(.f = broom::tidy)
+)
+tab <- nsnest %>% unnest(coefs) %>% filter(term=="wo") %>% arrange(match) %>%
+  select(-term, -statistic, -p.value)
+knitr::kable(tab, digits=3, caption = "Estimates of the increase in percent of failed tests corresponding to a 100 point increase in the optimization window.", booktabs=TRUE)
+```
+Figure \ref{fig:missings} gives an overview of the number of failed tests, i.e.
+tests in which a particular parameter setting did not return a valid result.
+This happens, when the shift to align two signatures is so large, that the remaining overlap is too small to accommodate  windows for validation.
+The problem is therefore exacerbated by a larger validation window.
+Figure \ref{fig:missings} also shows that the number of failed tests is approximately linear in the size of the optimization window.
+Test results from different sources have a much higher chance to fail, raising the question, whether failed tests should be treated as rejections of the null hypothesis of same source.
+For known non-matches there is a higher possibility that in the optimization pair of windows where cross-correlations are maximum are too far apart, and same shifts of this order hit the end of the signature.
+### Profiles
+
+### Comparison of signatures and profiles
+
+Another reason for failed tests can be incorrect identification of maximum correlation windows in the optimization step as seen in figure \ref{fig:prof_missings}(d) because of the level of smoothing, as too much smoothing would subdue intricate features that might otherwise help in the correlation calculations and correct identification of maximum correlation windows irrespective of the size.
+This would again cause a simiar effect as explained for figure \ref{fig:missings} with validation windows, irrespective of size, during the shifts end up at the ends of the markings resulting in an invalid calculation and failed comparison attempt.
+
+In figure \ref{fig:prof_missings}(d) and (f), we compare profiles and signatures on the basis of number of failed tests.
+The profiles chosen for figure \ref{fig:prof_missings}(f) have a constant coarseness of 0.25 and window of optimization as 120.
+The signatures in this case are not smoothed using the chumbley algorithm step of LOWESS smoothing.
+Instead signatures are used as calculated by \citet{aoas}.
+The smoothing in these signatures were determined and fixed on the basis of their performance in the random forest based algorithm proposed by \citet{aoas}.
+The comparison of profiles and signatures with variation of validation window size therefore is made on even footing.
+The trends are similar to figure \ref{fig:missings} in the sense that for known non-matches the number of failed tests are more for both signatures and profiles and increasing linearly with the validation window size.
+The problem is however, worse for profiles which has higher number of failed tests than signatures for all validation windows.
+
+The total error for different validation window sizes for signatures and profiles can be seen in figure \ref{fig:prof_missings} (e).The optimization window size is 120 and profiles are calculated at a default 0.25 coarseness level while signatures as before are not smoothed again in the modified chumbley algorithm.
+We can see that the total error is always higher for profiles as compared to signatures for all sizes of validation window.
+
+```{r, fig.cap="Type II errors for different levels of coarseness."}
+e1<-errors2 %>% filter(wv==30)%>% ggplot(aes(x = coarse, y=beta, colour=factor(alpha))) +
+  geom_point() + geom_line() +
+  theme_bw() +
+  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + xlab("Profile Coarseness")+
+  ylab("Type II error") + ggtitle(label = "(a)")
+
+e2<- errors_sig_c %>% filter(wv==30)%>% ggplot(aes(x = coarse, y=beta, colour=factor(alpha))) +
+  geom_point() + geom_line() +
+  theme_bw() +
+  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2")+ xlab("Signature Coarseness") + ylab("Type II error") + ggtitle(label = "(b)")
+
+#multiplot(e1,e2, cols = 1)
+
+```
+
+```{r}
+#errors %>% ggplot(aes(x = wv, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + facet_grid(~wo) +geom_point(data=errors2, aes(shape="profiles"))
+
+total1<- errors_sig_c %>% ggplot(aes(x = coarse, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + geom_line()  + geom_point(data= errors2 %>% filter(wv==30), aes(shape="profiles")) + geom_line(data=errors2 %>% filter(wv==30), aes(shape="profiles")) + theme_bw() +
+  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + xlab("Coarseness") +ylab("Total error") + ggtitle(label = "(c)")
+
+total2<- wo120 %>% filter(wv>10) %>% ggplot(aes(x = wv, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + geom_line() +geom_point(data=errors2 %>% filter(wv>10) %>% filter(coarse==0.25), aes(shape="profiles")) + geom_line(data=errors2 %>% filter(wv>10)%>% filter(coarse==0.25), aes(shape="profiles")) + theme_bw() +
+  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + ylab("Total error") + ggtitle(label = "(e)")
+
+```
+
+```{r prof_missings, fig.width=10, fig.height=10, fig.cap="Bottom row shows Total error and Number of failed tests by the window validation size, wv, and ground truth, while the top row shows Total error and Number of failed tests with Coarseness for both profiles and signatures"}
+
 #ns1<- ns[which(ns$wv == c(30,50), arr.ind = T),]
@@
   xlab("Coarseness") +
-  scale_colour_brewer("Match", palette="Set1") + geom_point(data=ns_sig_c  %>% filter(wv %in% c(30,50)) , aes(shape="signature"), size = 2) + ggtitle(label = "(c)")
+  scale_colour_brewer("Match", palette="Set1") + geom_point(data=ns_sig_c  %>% filter(wv %in% c(30,50)) , aes(shape="signature"), size = 2) + ggtitle(label = "(d)")
 # + scale_colour_brewer("Match", palette="Set2") #+geom_smooth(method="auto", size=.5, se=FALSE) 
@@
   xlab("wv") +
-  scale_colour_brewer("Match", palette="Set1") + geom_point(data=ns %>% filter(wo == 120)  %>% filter(coarse == 1) , aes(shape="signature"), size = 2) + ggtitle(label = "(d)")#+ scale_colour_brewer("Match", palette="Set2")
+  scale_colour_brewer("Match", palette="Set1") + geom_point(data=ns %>% filter(wo == 120)  %>% filter(coarse == 1) , aes(shape="signature"), size = 2) + ggtitle(label = "(f)")#+ scale_colour_brewer("Match", palette="Set2")
 
-multiplot(plotlist = list(total1, total2, miss_1, miss_2), cols = 2)#, matrix(c(1,2,3,3), nrow=2, byrow=TRUE))
+#multiplot(plotlist = list(total1, total2, miss_1, miss_2, e1, e2), cols = 2)#, matrix(c(1,2,3,3), nrow=2, byrow=TRUE))
+multiplot(plotlist = list(e1, total1, total2,e2, miss_1, miss_2), cols = 2)
 ```
-
-## Discussion
-
-# Algorithm Modification
-
-The proposed algorithm modification is at the same shift step which comes after the shift distance is identified by finding two windows in the two markings that have the highest correlation.
-
-The modification allows for a "wiggle" room in the second marking of the two sets of markings being compared.
-This means that each set of same shift windows that are being compared, the second marking will window that is under consideration is allowd to move a little towards the left and a little towards the right.
-
-This gives us a new set of 4 windows, 2 to the left and 2 to the right of the original comparison window.
-Then the correlations for each one of these 5 windows with the window under consideration of the 1st marking is retrieved (from the validation step correlation matrix) or computed.
-
-Then the window that has the maximum correlation (from the set of 5 windows in the 2nd marking) with the window of the 1st marking is chosen.
-
-This new maximum correlation is then used to compute the U-statistics using the same method as before.
-
-## Expected advantage of modification
-
-We are expecting that this modification would improve the type I and II errors for the good. Less number of false negatives and false positives.
-The reason for this expectation is many a times the bullet markings are not made at the same distances for two or more bullets because of the way it comes out of the barrel.
-
-Therefore rigid same shifts might not necessarily compare the right set of windows.
-In the same shift step for one set of windows, allowing the window in the 2nd marking to wiggle left and right and finding the best match to the window in the 1st marking, lets us adjust for situations where markings are compressed or elongated.
-
-## Dependence on the delta of new windows from the original same shift window
-
-From initial tests the amount of movement to get new windows left and right of the original same shift window seems to directly influence the
-the pvalue and U statistic that we look at.
-
-### other questions
-How much movement is reasonable and should be used?
-Is there a way to understand if the 2nd marking in comparison to the 1st marking is compressed or elongated or neither?
-Should the amount of movement selected depend on this compression or elongation?
-or using a static delta movement value justifieable.
+ \pagebreak

### 2018-01-04 doc_v2/doc_v2.Rmd
.

@@
 library(kableExtra)
-a27<- all %>% filter(wo==200, wv==30) %>% xtabs(data= ., ~signif+match)
+a27<- all %>% filter(wo==280, wv==30) %>% xtabs(data= ., ~signif+match)
 a27<- as.data.frame(a27)
@@
 
-knitr::kable(list(a27,a28,a29), digits=0, caption = "Confusion Table for different optimization window sizes with validation window size as 30", format = "latex" ,booktabs=TRUE)%>% kableExtra::kable_styling(latex_options = c("hold"), "striped", full_width = F) %>% group_rows("Size of Optimization Window = 200",1,5) %>% group_rows("Size of Optimization Window = 120", 6, 10) %>%  group_rows("Size of Optimization Window = 80", 11,14)
+knitr::kable(list(a27,a28,a29), digits=0, caption = "Confusion Table for different optimization window sizes with validation window size as 30.", format = "latex" ,booktabs=TRUE)%>% kableExtra::kable_styling(latex_options = c("hold"), "striped", full_width = F) %>% group_rows("Size of Optimization Window = 280",1,5) %>% group_rows("Size of Optimization Window = 120", 6, 10) %>%  group_rows("Size of Optimization Window = 80", 11,14)
 ```
 
-### Signatures
+## Sigatures
 
@@
 Hence we are inclined to choose a smaller validation window size and optimization window as 120.
+
+Table \ref{tab:confusion} shows the confusion tables with the classification of type I and type II errors and how the numbers change with a change in the optimization window.
+The windows represent areas to the left of the window with minimum type II, near the minimum type II window and to the far right of the minimum type II error
 
@@
 
+The actual type I error and type II error for signatures were also compared for different validation window sizes.
+Figure \ref{fig:wo120} shows the actual rates for different nominal $\alpha$ levels.
+We can see that the actual $\alpha$ levels are well below the nominal rates for all validation windows, while the type II error rises with higher validation windows for the smaller nominal $\alpha$ levels while for the nominal $\alpha$ = 0.05 its almost constant.
 ```{r }
@@
 For known non-matches there is a higher possibility that in the optimization pair of windows where cross-correlations are maximum are too far apart, and same shifts of this order hit the end of the signature.
-### Profiles
 
-### Comparison of signatures and profiles
+## Effect of coarseness
+
+Figure \ref{fig:prof_missings} (a) shows the type II error rates for profiles for the optimization window 120 and validation window 30 with varying level of coarseness.
+We can see that the type II error for all the nominal $\alpha$ levels is lowest in the range of 0.20 to 0.35.
+Therefore, a value of 0.25 can be used keeping in mind it keeps the type II error lowest while runnning simulations.
+Thus for comparisons of different window sizes etc as seen in the differnt parts of Figure \ref{fig:prof_missings} this coarseness value is used.
+
+On the other hand Figure \ref{fig:prof_missings} (b) shows if the coarseness level set in the chumbley agorithm has any effect on the signatures, which are pre-processed and already smoothed to a certain extent.
+From Figure \ref{fig:prof_missings} (b) we can notice that for different nominal $\alpha$ levels, the type II error fluctuates slightly but does not change much, thereby helping us conclude that the coarseness levels set in the LOWESS smoothing in the chumbley alggorithm does effect the type II error much for signatures.
+
+### Comparison of profiles and signatures
 

### 2018-01-04 doc_v2/doc_v2.Rmd
conclusion added

@@
 
-## Sigatures
+## Signatures
 
@@
 
-## Effect of coarseness
+## Profiles
 
@@
  \pagebreak
+
+## Conclusion
+
+The results suggest that the Nominal type I error $\alpha$ value shows dependence on the size of the window of optimization.
+For a given window of optimization the actual Type I error is comparable to the nominal level for only a select few validation window sizes and for comparable validation window sizes of 30 and 50 as done here, the actual type I error does not seem to vary as much as it varies with the optimization window sizes .
+A Test Fail, i.e.
+tests in which a particular parameter setting did not return a valid result, happens, when the shift to align two signatures is so large, that the remaining overlap is too small to accommodate  windows for validation, depends on whether known-match or known non-matches has predictive value, with test results from different sources having a much higher chance to fail.
+On conducting an analysis of all known bullet lands using the adjusted chumbley algorithm, Type II error was identified to be least bad for window of validation 30 and window of optimization 120.
+In case of unsmoothed raw marks (profiles), Type II error increases with the amount of smoothing and least for LOWESS smoothing coarseness value about 0.25 or 0.3.
+In an effort to identify the level of adaptiveness of the algorithm, comparisons were made between signatures and profiles.
+Their comparison with respect to validation window size for a fixed optimization window size suggested that, profiles have a total error (i.e all incorrect classification of known-matches and known non-matches) greater than or equal to the total error of signatures for all sizes of validation window.
+Profiles also fail more number of times than signatures in a test fail (for different coarseness keeping windows fixed and also for different validation windows keeping coarseness fixed) which lets us conclude that the behaviour of the algorithm for the profiles instead of pre-processed signatures is not better.
+Finally it should be noted that the current version of the adjusted chumbley algorithm seems to fall short when compared to other machine-learning based methods \citet{aoas}, and some level of modification to the deterministic algorithm needs to be identified and tested that would reduce the number of incorrect classifications.

### 2018-01-04 doc_v2/doc_v2.Rmd
rephrasing sentences

@@
 Figure \ref{fig:type1} compares nominal (fixed) type I error and actually observed type I errors for the parameter settings in table \ref{tab:param}.
-With an increasing size of the window used in the optimization step the observed type I error rate decreases (slighty)or As the optimization window increase the observed type I error rate gets smaller, while, A smaller validation window tends to be associated with a higher type I error rate.
-This can be better imagined for a given window of optimization, when the actual Type I error is comparable to the nominal level for only a select few validation window sizes and for comparable validation window sizes of 30 and 50 as done here, the actual type I error although increases negligibly, it does not seem to vary as much as it varies with the optimization window sizes.
-This might be related to the increasing number of tests that fail for larger window sizes, in particular for non-matching striae (see fig \ref{fig:missings}).
+With an increasing size of the window used in the optimization step the observed type I error rate decreases (slighty).
+This means as the optimization window increase the observed type I error rate gets smaller.
+A smaller validation window on the other hand, tends to be associated with a higher type I error rate.
+This can be better imagined for a given window of optimization, where the actual Type I error is comparable to the nominal level for only a select few validation window sizes.
+For these comparable validation window sizes of 30 and 50 as done here, the actual type I error increases very slightly and can be seen in Figure \ref{fig:type1}.
+This increase is not as much when compared to the variation seen with the optimization window sizes.
+This effect might be related to the increasing number of tests that fail for larger optimization window sizes, in particular for non-matching striae (see fig \ref{fig:missings}).
 

### 2018-01-04 doc_v2/doc_v2.Rmd
.

@@
 
-```{r wo120, fig.width = 8, fig.height=4.5, fig.cap="For optimization window size 120 and different validation window size how Actual and Type II (beta) errors vary for various chosen nominal alpha levels"}
+```{r wo120, fig.width = 8, fig.height=4.5, fig.cap="The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120"}
 wo120 <- errors %>% filter(wo==120)
@@
 
-```{r prof_missings, fig.width=10, fig.height=10, fig.cap="Bottom row shows Total error and Number of failed tests by the window validation size, wv, and ground truth, while the top row shows Total error and Number of failed tests with Coarseness for both profiles and signatures"}
+```{r prof_missings, fig.width=10, fig.height=10, fig.cap="Row 3:  Total error and Number of failed tests by the window validation size, wv, and ground truth, Row 2: Total error and Number of failed tests with Coarseness for both profiles and signatures, Row 1: Type II error for different coarseness levels as used in the modified chumbley algorithm for profiles and signatures"}
 

### 2018-01-11 doc_v2/doc_v2.Rmd
rescaled image

@@
 
+\hh{figure \ref{fig:rgl} needs to be mentioned somewhere in the write-up}
+
 \begin{figure}
 \centering
-\includegraphics[width=\textwidth]{images/B6-B2-L6.png}
-```{r, echo=FALSE, fig.width=8, fig.height=3, out.width='\\textwidth'}
+\includegraphics[width=\textwidth]{images/B6-B2-L6-rescaled.png}
+```{r, eval=FALSE, echo=FALSE, fig.width=8, fig.height=3, out.width='\\textwidth'}
 land <- read.csv("../data/b6b2l6.csv")
-land %>% ggplot(aes(x = y, y=resid)) +geom_line() +theme_bw()
+land %>% ggplot(aes(x = y, y=resid)) +geom_line() +theme_bw() +
+  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
+  xlab(expression(paste("Relative Location (in ",mu,m,")", sep="")))
 ```
 
-\caption{\label{fig:rgl} Image of a bullet land from a confocal light microscope at 20 fold magnification.}
+\caption{\label{fig:rgl} Image of a bullet land from a confocal light microscope at 20 fold magnification (top) and a chart of the corresponding signature of the same land (bottom).
+The dotted lines connect some peaks visible in both visualizations.}
 

### 2018-01-22 doc_v2/doc_v2.Rmd
pdf

@@
   theme_bw()+ 
-  theme(legend.position="bottom")
+  theme(legend.position="bottom") +  ggtitle("Signatures" , subtitle = "Varying validation window sizes, Optimization window = 120")
 ```
@@
 Finally it should be noted that the current version of the adjusted chumbley algorithm seems to fall short when compared to other machine-learning based methods \citet{aoas}, and some level of modification to the deterministic algorithm needs to be identified and tested that would reduce the number of incorrect classifications.
+
+```{r wo120-profile, fig.width = 8, fig.height=4.5, fig.cap="The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120"}
+wo120_profile <- errors2 %>% filter(wo==120, coarse == 0.25) %>% 
+  gather(type, error, actual:beta) %>% 
+  ggplot(aes(x = wv, y = error, colour = factor(alpha)))  + 
+  geom_line() + geom_point() + facet_wrap(~type, scales="free") +
+    theme_bw()+ 
+  theme(legend.position="bottom") + ggtitle("Profile" , subtitle = "Varying validation window sizes, Optimization window = 120, Coarseness = 0.25")
+
+wo120_1<- wo120 %>% gather(type, error, actual:beta) %>%  
+  ggplot(aes(x = wv, y = error, colour = factor(alpha)))+
+  geom_line(linetype=5)+  geom_point(shape= 24) + facet_wrap(~type, scales="free") +
+  theme_bw()+ 
+  theme(legend.position="bottom") + ggtitle("Signatures" , subtitle = "Varying validation window sizes, Optimization window = 120")
+
+wo120_profile
+wo120_1
+#multiplot(wo120_1, wo120_profile1)
+```

### 2018-02-05 short/Feb_rewrite/Introduction to adaptation of chumbley.Rmd
intro

---
title: "Adaption of the Chumbley Score to matching of bullet striation marks"
authors:
- affiliation: Department of Statistics, Iowa State University 
  name: Ganesh Krishnan
  thanks: The authors gratefully acknowledge ...
- affiliation: Department of Statistics and CSAFE, Iowa State University 
  name: Heike Hofmann
biblio-style: apsr
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    keep_tex: yes
    latex_engine: pdflatex
    template: template.tex
  html_document: default
blinded: 0
keywords:
- Bullets, Signatures, Lands, Errors, Known match, Known non-match, Comparisons
bibliography: bibliography
<!-- abstract: null -->
---

\newcommand{\hh}[1]{{\textcolor{orange}{#1}}}
\newcommand{\gk}[1]{{\textcolor{green}{#1}}}
\newcommand{\cited}[1]{{\textcolor{red}{#1}}}

<!-- \tableofcontents -->

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  error = FALSE, 
  collapse = TRUE,
  comment = "",
  fig.height = 6,
  fig.width = 6,
  fig.align = "center",
  out.width= '\\textwidth',
  cache = FALSE,
  fig.path='figures/',
  echo=FALSE,
  cache=TRUE
)
options(knitr.table.format = "latex")
library(tidyverse)
```
# Introduction

Compairing pairs of toolmarks with the intention of matching it to a tool has been studied many times in the past. Extensive examples in literature for tools and toolmark research ranging from screwdrivers to groove pliers to slip-joint pliers can be found in the work of  \citet{manytoolmarks1}, \citet{manytoolmarks2}, \citet{miller}, \cite{afte-chumbley} and many more. In comparison to this, same source matching of bullets to firearms has not been examined as prominently as that of toolmarks, with even less information available on validity of methods and error rates associated with firearms examination. The National Academy of Sciences in its report in 2009 \citep{NAS:2009} discussed the need for determining error rates in methods proposed for firearms examination. In the context of same source matching and error rates determination, as seen in the case of most forensic applications, the first step involves identification of unique features that are characteristic of the object at hand. For the case of bullets and firearms, striation marks on the surface of the bullet are considered to be such markings that can be used in methods for same source matching. These marks are often a product of rifling and impurities and defects due to manufacturing in the barrel of the gun, which leads to engravings on the bullet surface \citep{afte-article1992}. In current practice, firearm examiners invariably make visual comparisons of bullet striae and use visual assesment tools to dignify bullets as being matches and non-matches. One way of accomplishing anykind of comparison between bullets is to do a comparison between surface marking of two or more bullet lands. Bullet Lands are considered to be areas between grooves made by the rifling action of the barrel. These marking are considered to be unique. The land engraved markings or sometimes termed as Bullet profiles \citep{aoas} are striation marks made on Bullet lands and often used for these land to land comparisons. Bullet Signatures is another word used in literature as seen in the work of \citet{chu2013} and \citet{aoas}. In our context bullet signatures refer to a processed version of the raw land engraved markings or profiles. The generation of bullet signatures involves first extraction of a bullet profile by taking the cross-sectional of the surface at a given height and then using loess fits to model the structure. The residuals of this fit are called signatures, which are considered to be noise free and a good reflection of the class charecteristics and unique features of a bullet. A more detailed version of the extraction technique of signatures is discussed by \citet{aoas}, where comprehensive details about the height at which profile is to be selected, removing curvature, smoothing, identifying groove locations are explained.

In the study conducted by \citet{aoas} a machine learning based algorithm was developed for same source matching of bullets and error rates were discussed using the database from the Hamby Study \citep{hamby}. In this paper, we first try to adapt a deterministic algorithm and method developed for toolmarks by \citet{chumbley} and improved by \citet{hadler}, to bullets. Then we consequently discuss about the efforts in doing so along with the associated error rates. The data used in this paper also belongs to the Hamby Study \citep{hamby}. This gives us a common platform for comparing the performance of the chumbley method on bullets with an already existing method proposed by \citet{aoas} for bullets. The proposed algorithm and method of \citet{chumbley}, in their paper, compares two toolmarks with the intention of determining if it comes from the same source (same tool). The method also provides a means to determine error rates and claims to reduce subject bias. As mentioned earlier, subject bias and error rate determination have been a long standing issue in firearm examination \citep{NAS:2009}. This remains one of the motivations to explore the adaptibility of the Chumbley score methodology to bullets. \citet{chumbley} used an empirical based setup to validate their proposed algorithm and quantitative method which calculates a U-statisic for the purpose of classification of toolmarks as matching or non-matching. The data for their study was obtained from 50 sequentially manufactured screwdriver tips, and preselected comparison window sizes were given as inputs to the algorithm. The algorithm then compares the two toolmarks and comes up with a U-statistic and an associated p-value to designate them as matches or non-matches. The performance for every 100 comparisons, of the algorithm proposed by \citet{chumbley} and the improvement proposed by \citet{hadler} are listed in the table below.

```{r, results='asis'}
library(knitr)
library(kableExtra)
library(xtable)

a1<- matrix(NA, ncol = 3, nrow = 2)
toolmark1.confusion<- data.frame(a1)
colnames(toolmark1.confusion)<- c("Classification", "Match", "Non-Match")
toolmark1.confusion$Classification<- c("Match", "Non-Match")
toolmark1.confusion$Match<- c(41,2)
toolmark1.confusion$`Non-Match`<- c(9,48)

a2<- matrix(NA, ncol = 3, nrow = 2)
toolmark2.confusion<- data.frame(a2)
colnames(toolmark2.confusion)<- c("Classification", "Match", "Non-Match")
toolmark2.confusion$Classification<- c("Match", "Non-Match")
toolmark2.confusion$Match<- c(47,0)
toolmark2.confusion$`Non-Match`<- c(3,50)

chumbley.confusion<-knitr::kable(toolmark1.confusion, digits=0, format = "latex" ,booktabs=TRUE) #%>% kableExtra::kable_styling(latex_options = c("hold"))
hadler.confusion<-knitr::kable(toolmark2.confusion, digits=0, format = "latex" ,booktabs=TRUE) #%>% kableExtra::kable_styling(latex_options = c("hold"))

cat(c("\\begin{table}[!htb]
    \\begin{minipage}{.5\\linewidth}
      \\caption{Chumbley et al. 2010}
      \\centering",
        chumbley.confusion,
    "\\end{minipage}%
    \\begin{minipage}{.5\\linewidth}
      \\centering
        \\caption{Hadler et. al. (2017)}",
       hadler.confusion,
    "\\end{minipage} 
\\end{table}"
))

```
\newpage

### 2018-02-05 short/Feb_rewrite/Introduction_to_adaptation_of_chumbley.tex
intro

--- 2018-02-05 short/Feb_rewrite/Introduction to adaptation of chumbley.Rmd
@@
 
-<!-- \tableofcontents -->
+\section{Introduction}\label{introduction}
 
-```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
-library(knitr)
-knitr::opts_chunk$set(
-  message = FALSE,
-  warning = FALSE,
-  error = FALSE, 
-  collapse = TRUE,
-  comment = "",
-  fig.height = 6,
-  fig.width = 6,
-  fig.align = "center",
-  out.width= '\\textwidth',
-  cache = FALSE,
-  fig.path='figures/',
-  echo=FALSE,
-  cache=TRUE
-)
-options(knitr.table.format = "latex")
-library(tidyverse)
-```
-# Introduction
+Compairing pairs of toolmarks with the intention of matching it to a
+tool has been studied many times in the past. Extensive examples in
+literature for tools and toolmark research ranging from screwdrivers to
+groove pliers to slip-joint pliers can be found in the work of
+\citet{manytoolmarks1}, \citet{manytoolmarks2}, \citet{miller},
+\cite{afte-chumbley} and many more. In comparison to this, same source
+matching of bullets to firearms has not been examined as prominently as
+that of toolmarks, with even less information available on validity of
+methods and error rates associated with firearms examination. The
+National Academy of Sciences in its report in 2009 \citep{NAS:2009}
+discussed the need for determining error rates in methods proposed for
+firearms examination. In the context of same source matching and error
+rates determination, as seen in the case of most forensic applications,
+the first step involves identification of unique features that are
+characteristic of the object at hand. For the case of bullets and
+firearms, striation marks on the surface of the bullet are considered to
+be such markings that can be used in methods for same source matching.
+These marks are often a product of rifling and impurities and defects
+due to manufacturing in the barrel of the gun, which leads to engravings
+on the bullet surface \citep{afte-article1992}. In current practice,
+firearm examiners invariably make visual comparisons of bullet striae
+and use visual assesment tools to dignify bullets as being matches and
+non-matches. One way of accomplishing anykind of comparison between
+bullets is to do a comparison between surface marking of two or more
+bullet lands. Bullet Lands are considered to be areas between grooves
+made by the rifling action of the barrel. These marking are considered
+to be unique. The land engraved markings or sometimes termed as Bullet
+profiles \citep{aoas} are striation marks made on Bullet lands and often
+used for these land to land comparisons. Bullet Signatures is another
+word used in literature as seen in the work of \citet{chu2013} and
+\citet{aoas}. In our context bullet signatures refer to a processed
+version of the raw land engraved markings or profiles. The generation of
+bullet signatures involves first extraction of a bullet profile by
+taking the cross-sectional of the surface at a given height and then
+using loess fits to model the structure. The residuals of this fit are
+called signatures, which are considered to be noise free and a good
+reflection of the class charecteristics and unique features of a bullet.
+A more detailed version of the extraction technique of signatures is
+discussed by \citet{aoas}, where comprehensive details about the height
+at which profile is to be selected, removing curvature, smoothing,
+identifying groove locations are explained.
 
-Compairing pairs of toolmarks with the intention of matching it to a tool has been studied many times in the past.
-Extensive examples in literature for tools and toolmark research ranging from screwdrivers to groove pliers to slip-joint pliers can be found in the work of  \citet{manytoolmarks1}, \citet{manytoolmarks2}, \citet{miller}, \cite{afte-chumbley} and many more.
-In comparison to this, same source matching of bullets to firearms has not been examined as prominently as that of toolmarks, with even less information available on validity of methods and error rates associated with firearms examination.
-The National Academy of Sciences in its report in 2009 \citep{NAS:2009} discussed the need for determining error rates in methods proposed for firearms examination.
-In the context of same source matching and error rates determination, as seen in the case of most forensic applications, the first step involves identification of unique features that are characteristic of the object at hand.
-For the case of bullets and firearms, striation marks on the surface of the bullet are considered to be such markings that can be used in methods for same source matching.
-These marks are often a product of rifling and impurities and defects due to manufacturing in the barrel of the gun, which leads to engravings on the bullet surface \citep{afte-article1992}.
-In current practice, firearm examiners invariably make visual comparisons of bullet striae and use visual assesment tools to dignify bullets as being matches and non-matches.
-One way of accomplishing anykind of comparison between bullets is to do a comparison between surface marking of two or more bullet lands.
-Bullet Lands are considered to be areas between grooves made by the rifling action of the barrel.
-These marking are considered to be unique.
-The land engraved markings or sometimes termed as Bullet profiles \citep{aoas} are striation marks made on Bullet lands and often used for these land to land comparisons.
-Bullet Signatures is another word used in literature as seen in the work of \citet{chu2013} and \citet{aoas}.
-In our context bullet signatures refer to a processed version of the raw land engraved markings or profiles.
-The generation of bullet signatures involves first extraction of a bullet profile by taking the cross-sectional of the surface at a given height and then using loess fits to model the structure.
-The residuals of this fit are called signatures, which are considered to be noise free and a good reflection of the class charecteristics and unique features of a bullet.
-A more detailed version of the extraction technique of signatures is discussed by \citet{aoas}, where comprehensive details about the height at which profile is to be selected, removing curvature, smoothing, identifying groove locations are explained.
+In the study conducted by \citet{aoas} a machine learning based
+algorithm was developed for same source matching of bullets and error
+rates were discussed using the database from the Hamby Study
+\citep{hamby}. In this paper, we first try to adapt a deterministic
+algorithm and method developed for toolmarks by \citet{chumbley} and
+improved by \citet{hadler}, to bullets. Then we consequently discuss
+about the efforts in doing so along with the associated error rates. The
+data used in this paper also belongs to the Hamby Study \citep{hamby}.
+This gives us a common platform for comparing the performance of the
+chumbley method on bullets with an already existing method proposed by
+\citet{aoas} for bullets. The proposed algorithm and method of
+\citet{chumbley}, in their paper, compares two toolmarks with the
+intention of determining if it comes from the same source (same tool).
+The method also provides a means to determine error rates and claims to
+reduce subject bias. As mentioned earlier, subject bias and error rate
+determination have been a long standing issue in firearm examination
+\citep{NAS:2009}. This remains one of the motivations to explore the
+adaptibility of the Chumbley score methodology to bullets.
+\citet{chumbley} used an empirical based setup to validate their
+proposed algorithm and quantitative method which calculates a U-statisic
+for the purpose of classification of toolmarks as matching or
+non-matching. The data for their study was obtained from 50 sequentially
+manufactured screwdriver tips, and preselected comparison window sizes
+were given as inputs to the algorithm. The algorithm then compares the
+two toolmarks and comes up with a U-statistic and an associated p-value
+to designate them as matches or non-matches. The performance for every
+100 comparisons, of the algorithm proposed by \citet{chumbley} and the
+improvement proposed by \citet{hadler} are listed in the table below.
 
-In the study conducted by \citet{aoas} a machine learning based algorithm was developed for same source matching of bullets and error rates were discussed using the database from the Hamby Study \citep{hamby}.
-In this paper, we first try to adapt a deterministic algorithm and method developed for toolmarks by \citet{chumbley} and improved by \citet{hadler}, to bullets.
-Then we consequently discuss about the efforts in doing so along with the associated error rates.
-The data used in this paper also belongs to the Hamby Study \citep{hamby}.
-This gives us a common platform for comparing the performance of the chumbley method on bullets with an already existing method proposed by \citet{aoas} for bullets.
-The proposed algorithm and method of \citet{chumbley}, in their paper, compares two toolmarks with the intention of determining if it comes from the same source (same tool).
-The method also provides a means to determine error rates and claims to reduce subject bias.
-As mentioned earlier, subject bias and error rate determination have been a long standing issue in firearm examination \citep{NAS:2009}.
-This remains one of the motivations to explore the adaptibility of the Chumbley score methodology to bullets.
-\citet{chumbley} used an empirical based setup to validate their proposed algorithm and quantitative method which calculates a U-statisic for the purpose of classification of toolmarks as matching or non-matching.
-The data for their study was obtained from 50 sequentially manufactured screwdriver tips, and preselected comparison window sizes were given as inputs to the algorithm.
-The algorithm then compares the two toolmarks and comes up with a U-statistic and an associated p-value to designate them as matches or non-matches.
-The performance for every 100 comparisons, of the algorithm proposed by \citet{chumbley} and the improvement proposed by \citet{hadler} are listed in the table below.
+\begin{table}[!htb]
+    \begin{minipage}{.5\linewidth}
+      \caption{Chumbley et al. 2010}
+      \centering 
+\begin{tabular}{lrr}
+\toprule
+Classification & Match & Non-Match\\
+\midrule
+Match & 41 & 9\\
+Non-Match & 2 & 48\\
+\bottomrule
+\end{tabular} \end{minipage}%
+    \begin{minipage}{.5\linewidth}
+      \centering
+        \caption{Hadler et. al. (2017)} 
+\begin{tabular}{lrr}
+\toprule
+Classification & Match & Non-Match\\
+\midrule
+Match & 47 & 3\\
+Non-Match & 0 & 50\\
+\bottomrule
+\end{tabular} \end{minipage} 
+\end{table}\newpage
 
-```{r, results='asis'}
-library(knitr)
-library(kableExtra)
-library(xtable)
+\bibliographystyle{agsm}
+\bibliography{bibliography}
 
-a1<- matrix(NA, ncol = 3, nrow = 2)
-toolmark1.confusion<- data.frame(a1)
-colnames(toolmark1.confusion)<- c("Classification", "Match", "Non-Match")
-toolmark1.confusion$Classification<- c("Match", "Non-Match")
-toolmark1.confusion$Match<- c(41,2)
-toolmark1.confusion$`Non-Match`<- c(9,48)
-
-a2<- matrix(NA, ncol = 3, nrow = 2)
-toolmark2.confusion<- data.frame(a2)
-colnames(toolmark2.confusion)<- c("Classification", "Match", "Non-Match")
-toolmark2.confusion$Classification<- c("Match", "Non-Match")
-toolmark2.confusion$Match<- c(47,0)
-toolmark2.confusion$`Non-Match`<- c(3,50)
-
-chumbley.confusion<-knitr::kable(toolmark1.confusion, digits=0, format = "latex" ,booktabs=TRUE) #%>% kableExtra::kable_styling(latex_options = c("hold"))
-hadler.confusion<-knitr::kable(toolmark2.confusion, digits=0, format = "latex" ,booktabs=TRUE) #%>% kableExtra::kable_styling(latex_options = c("hold"))
-
-cat(c("\\begin{table}[!htb]
-    \\begin{minipage}{.5\\linewidth}
-      \\caption{Chumbley et al. 2010}
-      \\centering",
-        chumbley.confusion,
-    "\\end{minipage}%
-    \\begin{minipage}{.5\\linewidth}
-      \\centering
-        \\caption{Hadler et. al. (2017)}",
-       hadler.confusion,
-    "\\end{minipage} 
-\\end{table}"
-))
-
-```
-\newpage
+\end{document}

### 2018-02-05 code/algo_modification_v2.R
algo-modification

@@ -396,7 +396,7 @@ sigs_nest <- sigs %>% group_by(land_id, profile_id, run_id) %>% tidyr::nest()
 wo <- 120
 wv <- 30
 coarse <- 1
-eps<- 1
+eps<- 4
 
 k <- 1
 n <- nrow(sigs_nest)
@@ -429,10 +429,10 @@ xtabs(~signif+match, data =dframe)
 
 #saveRDS(object = dframe, file = "chumbley_wo-280_wv-30")
 
-write.table(dframe, file = "algo-mod_v2_sig_chumbley_wo-120_wv-30_coarse1_eps1.csv", sep = ",", append = FALSE, col.names = TRUE)
+write.table(dframe, file = "algo-mod_v2_sig_chumbley_wo-120_wv-30_coarse1_eps4.csv", sep = ",", append = FALSE, col.names = TRUE)
 
 }
-aa<- read.csv2("./data/Algo_modification/Signatures/algo-mod2_v2_sig_chumbley_wo-120_wv-30_coarse1_eps1.csv",header = TRUE, sep = ",")
+aa<- read.csv2("./data/Algo_modification/Signatures/algo-mod2_v2_sig_chumbley_wo-120_wv-30_coarse1_eps4.csv",header = TRUE, sep = ",")
 
 #  dframe<- read.csv2("./toolmark-method-assessment/data/chumbley_wo-320_wv-50.csv",header = TRUE, sep = ",")
 # dframe$U<- as.numeric(as.character(dframe$U))

### 2018-02-05 code/algo_modification_v3_larger_psuedo_wv.R

@@ -1,0 +1,439 @@
+#' Modified Chumbley Non-Random
+#' 
+#' The modification is made in the same-shift, where for the 2nd signature 
+#' the window is moved a bit towards the left and right to find if the 
+#' correlation of the moved windows are higher than the original same shift (OSS) window
+#' If its higher the new correlation values are stored
+#' 
+#' This function computes the Chumbley U-Statistic on systemically chosen pairs of windows rather 
+#' than the original method which selects randomly chosen pairs of windows
+#' @param data1 The first tool mark as a 1-column matrix
+#' @param data2 The second tool mark as a 1-column matrix
+#' @param window_opt size of the window to be used in the optimization step
+#' @param window_val Size of the window to be used in the validation step
+#' @param coarse smoothing parameter for the normalization smooth
+#' @importFrom stats pnorm
+#' @importFrom ggplot2 ggplot
+#' @export
+#' @return list with
+#' \itemize{
+#' \item {same_shift_n} Number of same shift offsets used
+#' \item {diff_shift_n} Number of different shift offsets used
+#' \item {U} observed U statistic
+#' \item {p_value} Corresponding p-value
+#' }
+{
+  chumbley_non_random_modified <- function(data1, data2, window_opt = 500, window_val = 50, coarse = .25, eps = 1){
+    
+    unity <- function(x) {x / sqrt(sum(x^2))} ## normalize columns of a matrix to make correlation computation faster
+    
+    ####################################################
+    ##Clean the marks and compute the smooth residuals##
+    ####################################################
+    
+    data1 <- matrix(data1[round((0.01*nrow(data1))):round(0.99*nrow(data1)),], ncol = 1)
+    data2 <- matrix(data2[round((0.01*nrow(data2))):round(0.99*nrow(data2)),], ncol = 1)
+    
+    ##Normalize the tool marks
+    y1 <- data1 - lowess(y = data1,  x = 1:nrow(data1), f= coarse)$y
+    y2 <- data2 - lowess(y = data2,  x = 1:nrow(data2), f= coarse)$y
+    
+    
+    ############################################
+    ##Compute the observed maximum correlation##
+    ############################################
+    
[…]

### 2018-02-05 doc/paper.Rmd
updating from doc_v2

--- 2017-12-27 doc/paper.Rmd
@@
 
-## Problem statement and motivation
-
-\hh{'considered to be unique' by AFTE - that is right, but we should not use this argument, and I don't think we have to.
-It's better to focus on the problem at hand: striation marks are used to determine same source for bullets.
-XXX we could include some figures here XXX Current standards ask for a visual inspection of the marks under a comparison microscope by a firearms examiner.
-One of the problems raised by the NAS report in 2009 is the subjectivity involved in this comparison and the need for determining error rates \citep{NAS:2009}.
-There are different approaches used to achieve more objectivity and assess error rates.
-XXX here we need some examples of approaches XXX
-For toolmarks, the Chumbley score has been suggested to determine same-source. XXX summary of results from original paper and the adjustment XXX
-In this paper we are interested in how this method has to be adjusted to best be applied to the smaller marks on bullets. }
-\gk{done}
+## Motivation
 
@@
 Current standards ask for a visual inspection of the marks under a comparison microscope by a firearms examiner.
-One of the problems raised by the NAS report in 2009 is the subjectivity involved in this comparison and the need for determining error rates \citep{NAS:2009}.
+One of the problems raised by the NAS report in 2009 is the subjectivity involved in this comparison and the need for determining error rates \citet{NAS:2009}.
 The issue of subjectivity of a firearms examiner has been reviewed by many authors where things like scientific principle testability and error rates are considered to be defining aspects for an objective analysis such that without identifying the uncertainities associated with a method of comparison, it is inconclusive to regards an analysis complete.
 This means that with the usual method of visual comparison that is adopted by firearms examiners, lies the problem of lack of any systemized quantification and as such there seems no certain way of associating it with any kind of mathematical probability.
-Since determining error rates has been duly noted as a fundamental problem in forensic science \citep{NAS:2009}, there is a need of methods that address this problem.
-Some of the methods that have been used to estimate error rates in firearm examinations include Automatic cartridge case comparison where same source identification of cartridge cases and associated error rates have been provided by \citep{riva}, although as noted by \citep{aoas}, in this case alignments of striae involves roation of planes, which cannot be generalized for bullets.
-In case of methods that use machine learning algorithms some methods which are bootsrap based methods often tend to give over-estimated error rates \citep{efron}, but there are alternate error estimation techniques as described by \citep{aoas}, \citep{efron} and  \citep{vorburger2016} which give better results.
+Since determining error rates has been duly noted as a fundamental problem in forensic science \citet{NAS:2009}, there is a need of methods that address this problem.
+Some of the methods that have been used to estimate error rates in firearm examinations include Automatic cartridge case comparison where same source identification of cartridge cases and associated error rates have been provided by \citet{riva}, although as noted by \citet{aoas}, in this case alignments of striae involves roation of planes, which cannot be generalized for bullets.
+In case of methods that use machine learning algorithms some methods which are bootsrap based methods often tend to give over-estimated error rates \citet{efron}, but there are alternate error estimation techniques as described by \citet{aoas}, \citet{efron} and  \citet{vorburger2016} which give better results.
 
@@
 
-\citep{chumbley} found that the algorithm gives false-positives in slightly over 2 cases (or 2%) of the time and about 9 false-negatives (or 9%) for every 100 comparisons.
-\citep{hadler} on the other hand, in the improved version of the algorithm first described by \citep{chumbley} found false-positives for 0 cases and 3 false-negatives (or 6%) for 50 comparisons of knwon matches.
+\citet{chumbley} found that the algorithm gives false-positives in slightly over 2 cases (or 2%) of the time and about 9 false-negatives (or 9%) for every 100 comparisons.
+\citet{hadler} on the other hand, in the improved version of the algorithm first described by \citet{chumbley} found false-positives for 0 cases and 3 false-negatives (or 6%) for 50 comparisons of knwon matches.
+```{r}
+a<- matrix(NA, ncol = 3, nrow = 2)
+toolmark.falsepositive<- data.frame(a)
+colnames(toolmark.falsepositive)<- c("Classification", "Match", "Non-Match")
+toolmark.falsepositive$Classification<- c("Match", "Non-Match")
+toolmark.falsepositive$Match<- c(47,0)
+toolmark.falsepositive$`Non-Match`<- c(3,50)
+
+knitr::kable(toolmark.falsepositive, digits=0, format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("hold"))
+```
+
+## Potential limitations of the Chumbley Score adaptation to bullets 
+Bullets are much smaller in length, width, are not flat and curved in the cross-sectional topography as opposed to tools like screw driver tips which produced longer and pronounced markings.
+This means the makings made by barrels in comparison to toolmarks may have a problem in distinctiveness.
+The majority of Bullet profiles and signatures extracted by procedures mentioned by  \citet{aoas} are almost 1/4 th the size of toolmarks as used by \citet{chumbley} or even smaller.
+Striations on Bullets are made on their curved surfaces, whereas the algorithm developed by \citet{chumbley} and \citet{hadler} has only been tested for flatter and wider surfaces which have negligible curvature.
+Therefore, using methods proposed for toolmarks may need adaptation in order to give tangible results for bullets.
+Moreover, in order to to get flat bullet signatures and remove the curvatures some kind of smoothing needs to be applied as a pre-step which needs further investigation as to whether the level of smoothing does effect the working of the algorithm on Bullets.
+
+Also when in the optimization step, the Window of optimization for bullets will be shorter as the signatures are smaller.
+The idea is to keep the number of windows of optimization sufficiently large, which means shorter Trace segments (partition of signature or toolmark with length = size of window of optimization) that lets us compare smaller segments of one signature to another.
+This introduces a problem as, if we go too small in the window of optimization, the unique features of the trace segments are lost and seem similar, while too large sizes vastly reduces the weight of small features that would otherwise uniquely classify a signature and hence identify the region of agreement.
+
+Thus the Window of Optimization has a direct influence on whether we are Falsely rejecting the null when it is true (Type I) or Falsely accepting the null when it is false (Type II) making identification of the optimum size of window of optimization very important.
+This raises important questions about what parameter settings need to be chosen for bullet land comparison.
+Something similar is done in \citet{afte-chumbley} for toolmark comparisons of slip-joint pliers where optimum window sizes are determined.
+There is a need to figure out the best parameter settings which minimizes the errors for unsmoothed markings and pre-processed signatures as determined by \citet{aoas}.
+An analysis of these error rates and comparison with other methods will help us understand the adaptibility of the chumbley score to bullets.
+
+\hh{figure \ref{fig:rgl} needs to be mentioned somewhere in the write-up}
+
 \begin{figure}
 \centering
-\includegraphics[width=\textwidth]{images/B6-B2-L6.png}
-```{r, echo=FALSE, fig.width=8, fig.height=3, out.width='\\textwidth'}
+\includegraphics[width=\textwidth]{images/B6-B2-L6-rescaled.png}
+```{r, eval=FALSE, echo=FALSE, fig.width=8, fig.height=3, out.width='\\textwidth'}
 land <- read.csv("../data/b6b2l6.csv")
-land %>% ggplot(aes(x = y, y=resid)) +geom_line() +theme_bw()
+land %>% ggplot(aes(x = y, y=resid)) +geom_line() +theme_bw() +
+  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
+  xlab(expression(paste("Relative Location (in ",mu,m,")", sep="")))
 ```
 
-\caption{\label{fig:rgl} Image of a bullet land from a confocal light microscope at 20 fold magnification.}
+\caption{\label{fig:rgl} Image of a bullet land from a confocal light microscope at 20 fold magnification (top) and a chart of the corresponding signature of the same land (bottom).
+The dotted lines connect some peaks visible in both visualizations.}
 
@@
 ```{r win-comparison, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap=" The two plots on the left show how the same shift behaves in case of a matching pair and the two plots on the right show how the different shift behaves in case of a matching pair."}
-a<- matrix(NA, ncol = 3, nrow = 2)
-toolmark.falsepositive<- data.frame(a)
-colnames(toolmark.falsepositive)<- c("Classification", "Match", "Non-Match")
-toolmark.falsepositive$Classification<- c("Match", "Non-Match")
-toolmark.falsepositive$Match<- c(47,0)
-toolmark.falsepositive$`Non-Match`<- c(3,50)
-toolmark.falsepositive
 
@@
 
-$V = \dfrac{n_{s}n_{d}}{N\left(N-1\right)}\left[\Sigma^{n_{s}}R_{s}\left(i\right)^{2}+\Sigma^{n_{d}}R_{d}\left(j\right)^{2}\right] -\dfrac{n_{s} n_{d}\left(N+1\right)^{2}}{4\left(N-1\right)}$
+$$V = \dfrac{n_{s}n_{d}}{N\left(N-1\right)}\left[\Sigma^{n_{s}}R_{s}\left(i\right)^{2}+\Sigma^{n_{d}}R_{d}\left(j\right)^{2}\right] -\dfrac{n_{s} n_{d}\left(N+1\right)^{2}}{4\left(N-1\right)}$$
 
-## Potential limitations of the Chumbley Score
-
-\hh{in the application to striation marks on bullet lands: smaller in width and curved - need to adjust parameter settings; something similar is done in \citet{afte-chumbley} for toolmark comparisons of slip-joint pliers.
-XXX work in }
-
-Bullets are much smaller in length, width, are not flat and curved in the cross-sectional topography as opposed to tools like screw driver tips which produced longer and pronounced markings.
-This means the makings made by barrels in comparison to toolmarks may have a problem in distinctiveness.
-The majority of Bullet profiles and signatures extracted by procedures mentioned by  \citet{aoas} are almost 1/4 th the size of toolmarks as used by \citet{chumbley} or even smaller.
-Striations on Bullets are made on their curved surfaces, whereas the algorithm developed by \citet{chumbley} and \citet{hadler} has only been tested for flatter and wider surfaces which have negligible curvature.
-Therefore, using methods proposed for toolmarks may need adaptation in order to give tangible results for bullets.
-Moreover, in order to to get flat bullet signatures and remove the curvatures some kind of smoothing needs to be applied as a pre-step which needs further investigation as to whether the level of smoothing does effect the working of the algorithm on Bullets.
-
-Also when in the optimization step, the Window of optimization for bullets will be shorter as the signatures are smaller.
-The idea is to keep the number of windows of optimization sufficiently large, which means shorter Trace segments (partition of signature or toolmark with length = size of window of optimization) that lets us compare smaller segments of one signature to another.
-
-This introduces a problem as, if we go too small in the window of optimization, the unique features of the trace segments are lost and seem similar, while too large sizes vastly reduces the weight of small features that would otherwise uniquely classify a signature and hence identify the region of agreement.
-
-Thus the Window of Optimization has a direct influence on whether we are
-Falsely rejecting the null when it is true (Type I) or Falsely accepting the null when it is false (Type II) making identification of the optimum size of window of optimization very important.
-
-## Testing setup:
-\hh{Using cross-validation setup to identify appropriate parameter settings for (a) signatures and (b) profiles directly}
+## Testing setup
 
@@
 
-The marking made on bullets are smaller than toolmarks and is also less wider. The idea is to find out possible areas of error while adapting the score based method proposed for toolmarks.
+The marking made on bullets are smaller than toolmarks and is also less wider.
+The idea is to find out possible areas of error while adapting the score based method proposed for toolmarks, using cross-validation setup to identify appropriate parameter settings for (a) signatures and (b) profiles directly
 
-Bullet signatures being compared at this time are from the Hamby 44 and Hamby 252 data, which have the best set of known-matched and known non-matches.
-
-Bullet signatures are chosen by
-1. Filtering out Land_id for Profiles from the Hamby 44 and Hamby 252 data and removing 
-all NA values
-2. run_id = 3 is chosen, Seems like the signatures generated from this run_id give the closest
-match. Different run_id's have some different settings for generating the signatures
-(The level of smoothing does not seem to be one of them)
-
-The bullet signatures when generated already included the LOWESS smoothing.
-Therefore, the coarseness factor is set to 1 while running the chumbley_non_random()
-which generates the same_shift, different_shift, U-Stat and P_value parameters.
-
-window Optimization 320 Window Validation 50
-```{r echo=FALSE}
-all %>% filter(wo==320, wv==50) %>% xtabs(data=., ~signif+match)
-```
-window Optimization 120 Window Validation 50
-```{r echo=FALSE}
-all %>% filter(wo==120, wv==50) %>% xtabs(data=., ~signif+match)
-```
-window Optimization 80 Window Validation 50
-```{r echo=FALSE}
-all %>% filter(wo==80, wv==50) %>% xtabs(data=., ~signif+match)
-
-```
-
-# Results
-
-## Signatures
+### Signatures
 
@@
 
-We used the adjusted Chumbley method as proposed in \citet{hadler} and implemented in the R package `toolmaRk` \citep{toolmark} on all pairwise land-to-land comparisons of the Hamby scans (a total of 85,491 comparisons) in the following manner:
+Bullet signatures being compared at this time are therefore from the Hamby 44 and Hamby 252 data.
+The database setup and pre-processing system used for choosing the Bullet signatures are as described by \citet{aoas}.
+In order to choose the bullet signatures we first filter out Land_id for Profiles from the Hamby 44 and Hamby 252 data and remove all NA values.
+Then run_id = 3 is chosen as the signatures generated from this run_id give the closest match.
+Different run_id's have some different settings for generating the signatures.(The level of smoothing does not seem to be one of them)
 
-```{r, echo=FALSE, results='asis'}
+The bullet signatures when generated by this process already includes a loess smoothing.
+Therefore, the coarseness factor is set to 1 while running the chumbley non random algorithm for comparing different optimization windows.The algorithm generates the same_shift, different_shift, U-Stat and P_value parameters which are then used to calculate the errors associated with different sets of window sizes.
+
+### Profiles
+
+The profiles are cross-sectional values of the the bullet striation mark which are chosen at an optimum height (x as used by \citet{aoas}).
+This x or height is not a randomly chosen level.
+The rationale behind the choice has been explained by \citet{aoas}.
+A region is first chosen where the cross-correlation seems to change very less and in this region an optimum height is chosen.
+The profiles generally resemble a curve which is more or less similar to a quadratic curve (a quadratic fit to the raw data values of the profile is not an exact fit but it does show a similar trend).
+Profiles are the set of raw values representing the striation marks, and signatures are generated from these by removal of the inherent curvature and applying some smoothing (the signatures generated by \citet{aoas} use a loess function for smoothing).
+
+Similar to signatures the run_id = 3 was used when applying the chumbley algorithm using the database setup given by \citet{aoas} of Hamby-44 and Hamby-252 datasets, on the profiles.
+The run_id not only defines the level of smoothing but also signifies the chosen height at which the profiles were selected initially.
+Another important aspect is the range of horizontal values (which is referred to as the y values in \citet{aoas}) in the signatures.
+These have already been pre-processed in the database to not include any grooves.
+
+Therefore for the sake of comparison the run_id = 3 is still chosen so as to ensure that the horizontal values remain the same as that of the signatures.
+This also gives us profiles with the grooves removed.
+
+The idea therefore is to first use these raw values of the profile directly in the chumbley algorithm, and see how the algorithm performs for different coarseness values (smoothing parameter as referred in the function lowess used in the chumbley algorithm).
+
+\pagebreak
+
+# Results
+We used the adjusted Chumbley method as proposed in \citet{hadler} and implemented in the R package `toolmaRk` \citep{toolmark} on all pairwise land-to-land comparisons of the Hamby scans (a total of 85,491 comparisons) with the pairwise sets for the comparisons given in the table \ref{tab:param}.
+```{r param}
 param_setting<- all %>% select(wo, wv) %>% group_by(wo, wv) %>% tally() %>% select(wo,wv)
@@
 library(kableExtra)
-knitr::kable(t(param_setting), digits=0, caption = "Overview of parameter settings used for optimization and validation windows for bullet land signatures.", format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("scale_down"))
+#knitr::kable(t(param_setting)[,1:17], digits=0, caption = "Overview of parameter settings used for optimization and validation windows for bullet land signatures.", format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("scale_down", "hold"))
+
+par_tab1<-t(param_setting)[,1:18]
+par_tab2<-t(param_setting)[,19:33]
+knitr::kable(list(par_tab1,par_tab2), digits=0,caption = "Overview of parameter settings used for optimization and validation windows for bullet land signatures.", format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("scale_down","hold"))%>% kableExtra::kable_styling(latex_options = c("scale_down","hold")) #%>% group_rows("",1,2) %>% group_rows("",3,4)
+
 ```
 
-### Signatures
+```{r confusion}
+library(kableExtra)
+a27<- all %>% filter(wo==280, wv==30) %>% xtabs(data= ., ~signif+match)
+a27<- as.data.frame(a27)
+a27$Type<- c("True Negative","False Positive (Type I)","False Negative (Type II)","True Positive")
+
+a28<- all %>% filter(wo==120, wv==30) %>% xtabs(data= ., ~signif+match)
+a28<- as.data.frame(a28)
+a28$Type<- c("True Negative","False Positive (Type I)","False Negative (Type II)","True Positive")
+
+a29<- all %>% filter(wo==80, wv==30) %>% xtabs(data= ., ~signif+match)
+a29<- as.data.frame(a29)
+a29$Type<- c("True Negative","False Positive (Type I)","False Negative (Type II)","True Positive")
+
+knitr::kable(list(a27,a28,a29), digits=0, caption = "Confusion Table for different optimization window sizes with validation window size as 30.", format = "latex" ,booktabs=TRUE)%>% kableExtra::kable_styling(latex_options = c("hold"), "striped", full_width = F) %>% group_rows("Size of Optimization Window = 280",1,5) %>% group_rows("Size of Optimization Window = 120", 6, 10) %>%  group_rows("Size of Optimization Window = 80", 11,14)
+```
+
+## Signatures
 
 Figure \ref{fig:type2} gives an overview of type II error rates observed when varying the window size in the optimization step.
+Two levels of validation window size 30 and 50 were chosen as to compare the error rates for different nominal type I errors.
+We notice that the trends for these nominal type I errors are similar and in most cases a validation window of 50 has higher type II error than for 30.
+A change in this trend is seen for a 0.05 $\alpha$ level, although the difference between the two windows is very small for this case.
+We can also notice an obvious trend of increase in the Type II error as the window of optimization increases and see a minimum around the optimization window size of 120 pixels.
+Hence we are inclined to choose a smaller validation window size and optimization window as 120.
+
+Table \ref{tab:confusion} shows the confusion tables with the classification of type I and type II errors and how the numbers change with a change in the optimization window.
+The windows represent areas to the left of the window with minimum type II, near the minimum type II window and to the far right of the minimum type II error
 
@@
 
-Figure \ref{fig:type1} compares nominal (fixed) type I error and actually observed type I errors for the parameter settings in table XXX.
+Figure \ref{fig:type1} compares nominal (fixed) type I error and actually observed type I errors for the parameter settings in table \ref{tab:param}.
 With an increasing size of the window used in the optimization step the observed type I error rate decreases (slighty).
-This might be related to the increasing number of tests that fail for larger window sizes, in particular for non-matching striae (see fig \ref{fig:missings}).
+This means as the optimization window increase the observed type I error rate gets smaller.
+A smaller validation window on the other hand, tends to be associated with a higher type I error rate.
+This can be better imagined for a given window of optimization, where the actual Type I error is comparable to the nominal level for only a select few validation window sizes.
+For these comparable validation window sizes of 30 and 50 as done here, the actual type I error increases very slightly and can be seen in Figure \ref{fig:type1}.
+This increase is not as much when compared to the variation seen with the optimization window sizes.
+This effect might be related to the increasing number of tests that fail for larger optimization window sizes, in particular for non-matching striae (see fig \ref{fig:missings}).
 
 ```{r type1, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap="Comparison of observed and nominal type I error rates  across a range of window sizes for optimization $wo$.
-The horizontal line in each facet indicates the nominal type I error rate.
-As the optimization window increase the observed type I error rate gets smaller.
-A smaller validation window tends to be associated with a higher type I error rate."}
+The horizontal line in each facet indicates the nominal type I error rate."}
 errors %>% filter(wv %in% c(30,50)) %>%
@@
 
-```{r missings, fig.width=8, fig.height=5, fig.cap="Number of failed tests by the window optimization size, wo, and ground truth.
-Test results from different sources have a much higher chance to fail, raising the question, whether failed tests should be treated as rejections of the null hypothesis of same source."}
+```{r wo120, fig.width = 4, fig.height=4, fig.cap="The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120"}
+wo120 <- errors %>% filter(wo==120)
+wo120 %>% gather(type, error, beta) %>%#actual:beta) %>%
+  ggplot(aes(x = wv, y = error, colour = factor(alpha)))  +
+  geom_line() + geom_point() + facet_wrap(~type, scales="free") +
+  theme_bw()+
+  theme(legend.position="bottom") +  ggtitle("Signatures" , subtitle = "Varying validation window sizes, Optimization window = 120")
+```
+
+The actual type I error and type II error for signatures were also compared for different validation window sizes.
+Figure \ref{fig:wo120} shows the actual rates for different nominal $\alpha$ levels.
+We can see that the type II error rises with higher validation windows for the smaller nominal $\alpha$ levels while for the nominal $\alpha$ = 0.05 its almost constant.
+```{r }
 ns <- all %>% group_by(wv, wo, match) %>% summarize(
@@
 )
-ns %>% filter(wv %in% c(30, 50)) %>% 
-  ggplot(aes(x = wo, y = missperc, shape=factor(wv), colour = factor(match))) + 
-  geom_smooth(method="lm", size=.5, se=FALSE) +
+
+nsp <- allp %>% group_by(wv, wo, match, coarse) %>% summarize(
+  n_ss = mean(same_shift_n),
+  n_ds = mean(diff_shift_n),
+  zero_ss = sum(same_shift_n==0),
+  zero_ds = sum(diff_shift_n==0),
+  miss = sum(is.na(p_value)),
+  missperc = miss/85491*100
+)
+
+ns_sig_c <- all_sig_c %>% group_by(wv, wo, match, coarse) %>% summarize(
+  n_ss = mean(same_shift_n),
+  n_ds = mean(diff_shift_n),
+  zero_ss = sum(same_shift_n==0),
+  zero_ds = sum(diff_shift_n==0),
+  miss = sum(is.na(p_value)),
+  missperc = miss/85491*100
+)
+
+```
+
+```{r missings, fig.width=6, fig.height=4, fig.cap="Number of failed tests by the window optimization size, wo, and ground truth.",fig.align = "center"}
+#ns1<- ns[which(ns$wv == c(30,50), arr.ind = T),]
+missings<- ns %>% filter(wv %in% c(30, 50)) %>%
+#ns %>% filter(wv == c(30, 50)) %>%
+  ggplot(aes(x = wo, y = missperc, shape=factor(wv), colour = factor(match)), size=.3) + 
+  geom_smooth(method="lm", size=.2, se=FALSE) +
   geom_point() + 
@@
   scale_shape_discrete("Size of \nvalidation\nwindow, wv")
+missings
 ```
-Figure \ref{fig:missings} gives an overview of the number of failed tests, i.e.
-tests in which a particular parameter setting did not return a valid result.
-This happens, when the shift to align two signatures is so large, that the remaining overlap is too small to accommodate  windows for validation.
-The problem is therefore exacerbated by a larger validation window.
-Figure \ref{fig:missings} also shows that the number of failed tests is approximately linear in the size of the optimization window.
 
@@
 ```
+Figure \ref{fig:missings} gives an overview of the number of failed tests, i.e.
+tests in which a particular parameter setting did not return a valid result.
+This happens, when the shift to align two signatures is so large, that the remaining overlap is too small to accommodate  windows for validation.
+The problem is therefore exacerbated by a larger validation window.
+Figure \ref{fig:missings} also shows that the number of failed tests is approximately linear in the size of the optimization window.
+Test results from different sources have a much higher chance to fail, raising the question, whether failed tests should be treated as rejections of the null hypothesis of same source.
+For known non-matches there is a higher possibility that in the optimization pair of windows where cross-correlations are maximum are too far apart, and same shifts of this order hit the end of the signature.
 
-```{r wo120, fig.width = 8, fig.height=4.5}
-wo120 <- errors %>% filter(wo==120)
-wo120 %>% gather(type, error, actual:beta) %>%
+## Profiles
+
+Figure \ref{fig:prof_missings} (a) shows the type II error rates for profiles for the optimization window 120 and validation window 30 with varying level of coarseness.
+We can see that the type II error for all the nominal $\alpha$ levels is lowest in the range of 0.20 to 0.35.
+Therefore, a value of 0.25 can be used keeping in mind it keeps the type II error lowest while runnning simulations.
+Thus for comparisons of different window sizes etc as seen in the differnt parts of Figure \ref{fig:prof_missings} this coarseness value is used.
+
+On the other hand Figure \ref{fig:prof_missings} (b) shows if the coarseness level set in the chumbley agorithm has any effect on the signatures, which are pre-processed and already smoothed to a certain extent.
+From Figure \ref{fig:prof_missings} (b) we can notice that for different nominal $\alpha$ levels, the type II error fluctuates slightly but does not change much, thereby helping us conclude that the coarseness levels set in the LOWESS smoothing in the chumbley alggorithm does effect the type II error much for signatures.
+
+### Comparison of profiles and signatures
+
+Another reason for failed tests can be incorrect identification of maximum correlation windows in the optimization step as seen in figure \ref{fig:prof_missings}(d) because of the level of smoothing, as too much smoothing would subdue intricate features that might otherwise help in the correlation calculations and correct identification of maximum correlation windows irrespective of the size.
+This would again cause a simiar effect as explained for figure \ref{fig:missings} with validation windows, irrespective of size, during the shifts end up at the ends of the markings resulting in an invalid calculation and failed comparison attempt.
+
+In figure \ref{fig:prof_missings}(d) and (f), we compare profiles and signatures on the basis of number of failed tests.
+The profiles chosen for figure \ref{fig:prof_missings}(f) have a constant coarseness of 0.25 and window of optimization as 120.
+The signatures in this case are not smoothed using the chumbley algorithm step of LOWESS smoothing.
+Instead signatures are used as calculated by \citet{aoas}.
+The smoothing in these signatures were determined and fixed on the basis of their performance in the random forest based algorithm proposed by \citet{aoas}.
+The comparison of profiles and signatures with variation of validation window size therefore is made on even footing.
+The trends are similar to figure \ref{fig:missings} in the sense that for known non-matches the number of failed tests are more for both signatures and profiles and increasing linearly with the validation window size.
+The problem is however, worse for profiles which has higher number of failed tests than signatures for all validation windows.
+
+The total error for different validation window sizes for signatures and profiles can be seen in figure \ref{fig:prof_missings} (e).The optimization window size is 120 and profiles are calculated at a default 0.25 coarseness level while signatures as before are not smoothed again in the modified chumbley algorithm.
+We can see that the total error is always higher for profiles as compared to signatures for all sizes of validation window.
+
+```{r, fig.cap="Type II errors for different levels of coarseness."}
+e1<-errors2 %>% filter(wv==30)%>% ggplot(aes(x = coarse, y=beta, colour=factor(alpha))) +
+  geom_point() + geom_line() +
+  theme_bw() +
+  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + xlab("Profile Coarseness")+
+  ylab("Type II error") + ggtitle(label = "(a)")
+
+e2<- errors_sig_c %>% filter(wv==30)%>% ggplot(aes(x = coarse, y=beta, colour=factor(alpha))) +
+  geom_point() + geom_line() +
+  theme_bw() +
+  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2")+ xlab("Signature Coarseness") + ylab("Type II error") + ggtitle(label = "(b)")
+
+#multiplot(e1,e2, cols = 1)
+
+```
+
+```{r}
+#errors %>% ggplot(aes(x = wv, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + facet_grid(~wo) +geom_point(data=errors2, aes(shape="profiles"))
+
+total1<- errors_sig_c %>% ggplot(aes(x = coarse, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + geom_line()  + geom_point(data= errors2 %>% filter(wv==30), aes(shape="profiles")) + geom_line(data=errors2 %>% filter(wv==30), aes(shape="profiles")) + theme_bw() +
+  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + xlab("Coarseness") +ylab("Total error") + ggtitle(label = "(c)")
+
+total2<- wo120 %>% filter(wv>10) %>% ggplot(aes(x = wv, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + geom_line() +geom_point(data=errors2 %>% filter(wv>10) %>% filter(coarse==0.25), aes(shape="profiles")) + geom_line(data=errors2 %>% filter(wv>10)%>% filter(coarse==0.25), aes(shape="profiles")) + theme_bw() +
+  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + ylab("Total error") + ggtitle(label = "(e)")
+
+```
+
+```{r prof_missings, fig.width=10, fig.height=10, fig.cap="Row 3:  Total error and Number of failed tests by the window validation size, wv, and ground truth, Row 2: Total error and Number of failed tests with Coarseness for both profiles and signatures, Row 1: Type II error for different coarseness levels as used in the modified chumbley algorithm for profiles and signatures"}
+
+#ns1<- ns[which(ns$wv == c(30,50), arr.ind = T),]
+miss_1<- nsp %>% filter(wv == 30) %>%
+#ns %>% filter(wv == c(30, 50)) %>%
+  ggplot(aes(x = coarse, y = missperc, shape= "profile", colour = factor(match))) + 
+  #geom_smooth(method="auto", size=.5, se=FALSE) +
+  geom_point(size = 2) + 
+  theme_bw() +
+  ylab("Percent of failed tests.") +
+  xlab("Coarseness") +
+  scale_colour_brewer("Match", palette="Set1") + geom_point(data=ns_sig_c  %>% filter(wv %in% c(30,50)) , aes(shape="signature"), size = 2) + ggtitle(label = "(d)")
+# + scale_colour_brewer("Match", palette="Set2") #+geom_smooth(method="auto", size=.5, se=FALSE) 
+  #scale_shape_discrete("Size of \nvalidation\nwindow, #wv")
+
+miss_2<- nsp %>% filter(coarse == 0.25) %>%
+#ns %>% filter(wv == c(30, 50)) %>%
+  ggplot(aes(x = wv, y = missperc, shape= "profile", colour = factor(match))) + 
+  #geom_smooth(method="auto", size=.5, se=FALSE) +
+  geom_point(size = 2) + 
+  theme_bw() +
+  ylab("Percent of failed tests.") +
+  xlab("wv") +
+  scale_colour_brewer("Match", palette="Set1") + geom_point(data=ns %>% filter(wo == 120)  %>% filter(coarse == 1) , aes(shape="signature"), size = 2) + ggtitle(label = "(f)")#+ scale_colour_brewer("Match", palette="Set2")
+
+#multiplot(plotlist = list(total1, total2, miss_1, miss_2, e1, e2), cols = 2)#, matrix(c(1,2,3,3), nrow=2, byrow=TRUE))
+multiplot(plotlist = list(e1, total1, total2,e2, miss_1, miss_2), cols = 2)
+```
+ \pagebreak
+
+## Conclusion
+
+The results suggest that the Nominal type I error $\alpha$ value shows dependence on the size of the window of optimization.
+For a given window of optimization the actual Type I error is comparable to the nominal level for only a select few validation window sizes and for comparable validation window sizes of 30 and 50 as done here, the actual type I error does not seem to vary as much as it varies with the optimization window sizes .
+A Test Fail, i.e.
+tests in which a particular parameter setting did not return a valid result, happens, when the shift to align two signatures is so large, that the remaining overlap is too small to accommodate  windows for validation, depends on whether known-match or known non-matches has predictive value, with test results from different sources having a much higher chance to fail.
+On conducting an analysis of all known bullet lands using the adjusted chumbley algorithm, Type II error was identified to be least bad for window of validation 30 and window of optimization 120.
+In case of unsmoothed raw marks (profiles), Type II error increases with the amount of smoothing and least for LOWESS smoothing coarseness value about 0.25 or 0.3.
+In an effort to identify the level of adaptiveness of the algorithm, comparisons were made between signatures and profiles.
+Their comparison with respect to validation window size for a fixed optimization window size suggested that, profiles have a total error (i.e all incorrect classification of known-matches and known non-matches) greater than or equal to the total error of signatures for all sizes of validation window.
+Profiles also fail more number of times than signatures in a test fail (for different coarseness keeping windows fixed and also for different validation windows keeping coarseness fixed) which lets us conclude that the behaviour of the algorithm for the profiles instead of pre-processed signatures is not better.
+Finally it should be noted that the current version of the adjusted chumbley algorithm seems to fall short when compared to other machine-learning based methods \citet{aoas}, and some level of modification to the deterministic algorithm needs to be identified and tested that would reduce the number of incorrect classifications.
+
+```{r wo120-profile, fig.width = 6, fig.height=4, fig.cap="The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120"}
+wo120_profile <- errors2 %>% filter(wo==120, coarse == 0.25) %>% 
+  #gather(type, error, actual:beta) %>% 
+  gather(type, error, beta) %>% 
   ggplot(aes(x = wv, y = error, colour = factor(alpha)))  + 
@@
     theme_bw()+ 
-  theme(legend.position="bottom")
+  theme(legend.position="bottom") + ggtitle("Profile" , subtitle = "Varying validation window sizes, Optimization window = 120, Coarseness = 0.25")
+
+#combined120 <- full_join(errors2 %>% filter(wo==120, coarse == 0.25), wo120 %>%
+
+wo120_1<- wo120 %>% gather(type, error, beta) %>%  
+  ggplot(aes(x = wv, y = error, colour = factor(alpha)))+
+  geom_line(linetype=5)+  geom_point(shape= 24) + facet_wrap(~type, scales="free") +
+  theme_bw()+ 
+  theme(legend.position="bottom") + ggtitle("Signatures" , subtitle = "Varying validation window sizes, Optimization window = 120")
+
+combined120<- 
+  ggplot(data = wo120 %>% gather(type, error, beta), aes(x = wv, y = error, colour = factor(alpha)))+
+  geom_line(aes(linetype="Signatures"))+  geom_point(shape= 24) + 
+  geom_line(data =  errors2 %>% filter(wo==120, coarse == 0.25) %>% 
+  gather(type, error, beta), aes(x = wv, y = error, colour = factor(alpha), linetype = "Profiles")) +       geom_point(data =  errors2 %>% filter(wo==120, coarse == 0.25) %>% 
+  gather(type, error, beta)) +
+  facet_wrap(~type, scales="free") +
+  scale_linetype_manual("Lines",values=c("Signatures"=2,"Profiles"=1))+
+  guides(fill = guide_legend(keywidth = 1, keyheight = 1),
+    linetype=guide_legend(keywidth = 3, keyheight = 1),
+    colour=guide_legend(keywidth = 3, keyheight = 1))+
+  theme_bw()+ 
+  theme(legend.position="right") + ggtitle("Type II error with Validation window" , subtitle = "Profiles at coarseness 0.25") + xlab("Type II error")
+
+combined120
+#wo120_profile
+#wo120_1
+#multiplot(wo120_1, wo120_profile1)
+```
+```{r failtest-coarseness, fig.width = 6, fig.height=4, fig.cap="The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120"}
+
+miss_1
+```
+```{r failtest-wv, fig.width = 6, fig.height=4, fig.cap="The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120"}
+
+miss_2
+```
+```{r failtest-wo, fig.width = 6, fig.height=4, fig.cap="The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120"}
+
+miss_2
 ```
 
-### Profiles
+<!-- # Algorithm Modification -->
 
-The profiles are cross-sectional values of the the bullet striation mark which are chosen at an optimum height (x as used by \citet{aoas}).
-This x or height is not a randomly chosen level.
-The rationale behind the choice has been explained by \citet{aoas}.
-A region is first chosen where the cross-correlation seems to change very less and in this region an optimum height is chosen.
-The profiles generally resemble a curve which is more or less similar to a quadratic curve (a quadratic fit to the raw data values of the profile is not an exact fit but it does show a similar trend).
-Profiles are the set of raw values representing the striation marks, and signatures are generated from these by removal of the inherent curvature and applying some smoothing (the signatures generated by \citet{aoas} use a loess function for smoothing).
+<!-- The proposed algorithm modification is at the same shift step which comes after the shift distance is identified by finding two windows in the two markings that have the highest correlation. -->
 
-Coming back to the database of Hamby-44 and Hamby-252 datasets, the run_id = 3 was used when applying the chumbley algorithm on the signatures.
-The run_id not only defines the level of smoothing but also signifies the chosen height at which the profiles were selected initially.
-Another important aspect is the range of horizontal values (which is referred to as the y values in \citet{aoas}) in the signatures.
-Thes have already been pre-processed in the database to not include any grooves.
+<!-- The modification allows for a "wiggle" room in the second marking of the two sets of markings being compared.
+This means that each set of same shift windows that are being compared, the second marking will window that is under consideration is allowd to move a little towards the left and a little towards the right.
+-->
 
-Therefore for the sake of comparison the run_id = 3 is still chosen so as to ensure that the horizontal values remain the same as that of the signatures.
-This also gives us profiles with the grooves removed.
+<!-- This gives us a new set of 4 windows, 2 to the left and 2 to the right of the original comparison window.
+Then the correlations for each one of these 5 windows with the window under consideration of the 1st marking is retrieved (from the validation step correlation matrix) or computed.
+-->
 
-\hh{XXX ideas should go to the previous section on setting up the study; in this section we focus on results only XXX}
-The idea therefore is to first use these raw values of the profile directly in the chumbley algorithm, and see how the algorithm performs for different coarseness values (smoothing parameter as referred in the function LOWESS used in the chumbley algorithm)
+<!-- Then the window that has the maximum correlation (from the set of 5 windows in the 2nd marking) with the window of the 1st marking is chosen. -->
 
-```{r, fig.cap="Type II errors for different levels of coarseness. XXX We need more values around .25"}
-errors2 %>% ggplot(aes(x = coarse, y=beta, colour=factor(alpha))) +
-  geom_point() + geom_line() +
-  theme_bw() +
-  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2")
-```
+<!-- This new maximum correlation is then used to compute the U-statistics using the same method as before. -->
 
-### comparison of total error with signatures and profiles
+<!-- ## Expected advantage of modification -->
 
-```{r}
-errors %>% ggplot(aes(x = wv, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + facet_grid(~wo) +
-  geom_point(data=errors2, aes(shape="profiles"))
-```
-## Discussion
+<!-- We are expecting that this modification would improve the type I and II errors for the good. Less number of false negatives and false positives. -->
+<!-- The reason for this expectation is many a times the bullet markings are not made at the same distances for two or more bullets because of the way it comes out of the barrel. -->
 
-## Results of Chumbley score
-- Nominal alpha value shows dependency on the size of the window of optimization
-- Test Fail depends on whether known-match or known non-matches has predictive value
-- Type II error least bad for window of validation 30 and window of optimization 120
+<!-- Therefore rigid same shifts might not necessarily compare the right set of windows.
+In the same shift step for one set of windows, allowing the window in the 2nd marking to wiggle left and right and finding the best match to the window in the 1st marking, lets us adjust for situations where markings are compressed or elongated.
+-->
 
-# Algorithm Modification
+<!-- ## Dependence on the delta of new windows from the original same shift window -->
 
-The proposed algorithm modification is at the same shift step which comes after the shift distance is identified by finding two windows in the two markings that have the highest correlation.
+<!-- From initial tests the amount of movement to get new windows left and right of the original same shift window seems to directly influence the -->
+<!-- the pvalue and U statistic that we look at.  -->
 
-The modification allows for a "wiggle" room in the second marking of the two sets of markings being compared.
-This means that each set of same shift windows that are being compared, the second marking will window that is under consideration is allowd to move a little towards the left and a little towards the right.
-
-This gives us a new set of 4 windows, 2 to the left and 2 to the right of the original comparison window.
-Then the correlations for each one of these 5 windows with the window under consideration of the 1st marking is retrieved (from the validation step correlation matrix) or computed.
-
-Then the window that has the maximum correlation (from the set of 5 windows in the 2nd marking) with the window of the 1st marking is chosen.
-
-This new maximum correlation is then used to compute the U-statistics using the same method as before.
-
-## Expected advantage of modification
-
-We are expecting that this modification would improve the type I and II errors for the good. Less number of false negatives and false positives.
-The reason for this expectation is many a times the bullet markings are not made at the same distances for two or more bullets because of the way it comes out of the barrel.
-
-Therefore rigid same shifts might not necessarily compare the right set of windows.
-In the same shift step for one set of windows, allowing the window in the 2nd marking to wiggle left and right and finding the best match to the window in the 1st marking, lets us adjust for situations where markings are compressed or elongated.
-
-## Dependence on the delta of new windows from the original same shift window
-
-From initial tests the amount of movement to get new windows left and right of the original same shift window seems to directly influence the
-the pvalue and U statistic that we look at.
-
-### how much of a movement is reasonable and should be used?
+<!-- ### other questions -->
+<!-- How much movement is reasonable and should be used?
 Is there a way to understand if the 2nd marking in comparison to the 1st marking is compressed or elongated or neither?
@@
 or using a static delta movement value justifieable.
+-->

### 2018-02-05 doc/paper.tex
updating from doc_v2

@@
      and \\     Heike Hofmann \\
-    Department of Statistics and csafe, Iowa State University\\
+    Department of Statistics and CSAFE, Iowa State University\\
       }
@@
 
-\section{Introduction}\label{introduction}
+\tableofcontents
+\newpage
 
-\subsection{Problem statement and
-motivation}\label{problem-statement-and-motivation}
+\section{Introduction and Background}\label{introduction-and-background}
 
-\hh{'considered to be unique' by AFTE - that is right, but we should not use this argument, and I don't think we have to.
-It's better to focus on the problem at hand: striation marks are used to determine same source for bullets.
-XXX we could include some figures here XXX Current standards ask for a visual inspection of the marks under a comparison microscope by a firearms examiner.
-One of the problems raised by the NAS report in 2009 is the subjectivity involved in this comparison and the need for determining error rates \citep{NAS:2009}.
-There are different approaches used to achieve more objectivity and assess error rates.
-XXX here we need some examples of approaches XXX
-For toolmarks, the Chumbley score has been suggested to determine same-source. XXX summary of results from original paper and the adjustment XXX
-In this paper we are interested in how this method has to be adjusted to best be applied to the smaller marks on bullets. }
-\gk{done}
+\subsection{Motivation}\label{motivation}
 
@@
 report in 2009 is the subjectivity involved in this comparison and the
-need for determining error rates \citep{NAS:2009}. The issue of
+need for determining error rates \citet{NAS:2009}. The issue of
 subjectivity of a firearms examiner has been reviewed by many authors
@@
 has been duly noted as a fundamental problem in forensic science
-\citep{NAS:2009}, there is a need of methods that address this problem.
+\citet{NAS:2009}, there is a need of methods that address this problem.
 Some of the methods that have been used to estimate error rates in
@@
 same source identification of cartridge cases and associated error rates
-have been provided by \citep{riva}, although as noted by \citep{aoas},
+have been provided by \citet{riva}, although as noted by \citet{aoas},
 in this case alignments of striae involves roation of planes, which
@@
 learning algorithms some methods which are bootsrap based methods often
-tend to give over-estimated error rates \citep{efron}, but there are
-alternate error estimation techniques as described by \citep{aoas},
-\citep{efron} and \citep{vorburger2016} which give better results.
+tend to give over-estimated error rates \citet{efron}, but there are
+alternate error estimation techniques as described by \citet{aoas},
+\citet{efron} and \citet{vorburger2016} which give better results.
 
@@
 prone and as seen in the work of \citet{miller} the number of CMS may
-turn to out to be high even when the bullet is not fires by the same
+turn out to be high even when the bullet is not fired by the same
 firearm. Parameters like cross-correlation factor as seen in the
 extensive comparisons made by \citep{aoas} seem to perform much better
-matching purposes.
+for matching purposes.
 
@@
 
-\citep{chumbley} found that the algorithm gives false-positives in
+\citet{chumbley} found that the algorithm gives false-positives in
 slightly over 2 cases (or 2\%) of the time and about 9 false-negatives
-(or 9\%) for every 100 comparisons. \citep{hadler} on the other hand, in
+(or 9\%) for every 100 comparisons. \citet{hadler} on the other hand, in
 the improved version of the algorithm first described by
-\citep{chumbley} found false-positives for 0 cases and 3 false-negatives
+\citet{chumbley} found false-positives for 0 cases and 3 false-negatives
 (or 6\%) for 50 comparisons of knwon matches.
+
+\begin{table}[!h]
+\centering
+\begin{tabular}{lrr}
+\toprule
+Classification & Match & Non-Match\\
+\midrule
+Match & 47 & 3\\
+Non-Match & 0 & 50\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\subsection{Potential limitations of the Chumbley Score adaptation to
+bullets}\label{potential-limitations-of-the-chumbley-score-adaptation-to-bullets}
+
+Bullets are much smaller in length, width, are not flat and curved in
+the cross-sectional topography as opposed to tools like screw driver
+tips which produced longer and pronounced markings. This means the
+makings made by barrels in comparison to toolmarks may have a problem in
+distinctiveness. The majority of Bullet profiles and signatures
+extracted by procedures mentioned by \citet{aoas} are almost 1/4 th the
+size of toolmarks as used by \citet{chumbley} or even smaller.
+Striations on Bullets are made on their curved surfaces, whereas the
+algorithm developed by \citet{chumbley} and \citet{hadler} has only been
+tested for flatter and wider surfaces which have negligible curvature.
+Therefore, using methods proposed for toolmarks may need adaptation in
+order to give tangible results for bullets. Moreover, in order to to get
+flat bullet signatures and remove the curvatures some kind of smoothing
+needs to be applied as a pre-step which needs further investigation as
+to whether the level of smoothing does effect the working of the
+algorithm on Bullets.
+
+Also when in the optimization step, the Window of optimization for
+bullets will be shorter as the signatures are smaller. The idea is to
+keep the number of windows of optimization sufficiently large, which
+means shorter Trace segments (partition of signature or toolmark with
+length = size of window of optimization) that lets us compare smaller
+segments of one signature to another. This introduces a problem as, if
+we go too small in the window of optimization, the unique features of
+the trace segments are lost and seem similar, while too large sizes
+vastly reduces the weight of small features that would otherwise
+uniquely classify a signature and hence identify the region of
+agreement.
+
+Thus the Window of Optimization has a direct influence on whether we are
+Falsely rejecting the null when it is true (Type I) or Falsely accepting
+the null when it is false (Type II) making identification of the optimum
+size of window of optimization very important. This raises important
+questions about what parameter settings need to be chosen for bullet
+land comparison. Something similar is done in \citet{afte-chumbley} for
+toolmark comparisons of slip-joint pliers where optimum window sizes are
+determined. There is a need to figure out the best parameter settings
+which minimizes the errors for unsmoothed markings and pre-processed
+signatures as determined by \citet{aoas}. An analysis of these error
+rates and comparison with other methods will help us understand the
+adaptibility of the chumbley score to bullets.
+
+\hh{figure \ref{fig:rgl} needs to be mentioned somewhere in the write-up}
 
@@
 \centering
-\includegraphics[width=\textwidth]{images/B6-B2-L6.png}
+\includegraphics[width=\textwidth]{images/B6-B2-L6-rescaled.png}
 
-\begin{center}\includegraphics[width=\textwidth]{figures/unnamed-chunk-1-1} \end{center}
-
-\caption{\label{fig:rgl} Image of a bullet land from a confocal light microscope at 20 fold magnification.}
+\caption{\label{fig:rgl} Image of a bullet land from a confocal light microscope at 20 fold magnification (top) and a chart of the corresponding signature of the same land (bottom).
+The dotted lines connect some peaks visible in both visualizations.}
 
@@
 
-\subsection{Existing solution for toolmarks: Chumbley
-Score}\label{existing-solution-for-toolmarks-chumbley-score}
-
-\hh{XXX anything to do with motivation has to come first - move this to the first paragraph of the intro XXX}
-\hh{XXX data and results come after the intro to chumbley.}
-\hh{XXX start with defining the data, i.e.
-we are working with random processes $x(t)$ and $y(t)$ ...
-you should also go into the coarseness parameter here - that is a pre-processing step before the optimization.
-}
-\gk{done}
+\section{The Chumbley Score Test}\label{the-chumbley-score-test}
 
 The Chumbley score algorithm takes input as two vectorized processes
-\(x(t_1)\) and \(y(t_2)\) which denote two sets of marks or striae.
-These marks or striae are potentially from two different bullets or two
-different toolmarks whose source needs to be identified as being same or
-different. The marks or striae are indexed by the pixel location where
-\(t_1\) is for the first striae referred to as \(x\) and \(t_2\) is for
-the second striae which is referred as \(y\). \(t_1\) and \(t_2\) need
-not be the same but are usually of similar lenghtsThe similarity is then
-judged by the algorithm on the basis of cross-correlation of a fixed and
-constant number consecutive pixels say \(k\) taken from the two indexed
-marks \(x(t_1)\) and \(y(t_2)\) such that in theory \(k\) remains
-smaller that length of the two striae or marks which is the same as
-\(t_1\) and \(t_2\). Depending on the what stage of the algorithm we are
-in, matching of different pixel lengths and locations is done, which at
-the end effectively compares all possible windows that would guarantee
-in quantifying the two marks or striae as coming from the same source or
-not, which also lets us assess the error rates by checking for a large
-number of cases.
+\(x(t_1)\), \(t_1 = 1,2,...T_1\) and \(y(t_2)\), \(t_2 = 1,2...T_2\)
+which denote two sets of marks or striae. Here the processes are of the
+form \(z(s)\) which is a spatial process for some \(t\), while
+\(z(s_1)\) is the realization of the process in \(s_1\). The \(t_1\) and
+\(t_2\) can be better understood as equally spaced pixel locations for
+the two marks under consideration, where a `pixel' is the resolution of
+the confocal light microscope. In the case for bullet signatures a pixel
+corresponds to 0.645 microns. These marks or striae are potentially from
+two different bullets or two different toolmarks whose source needs to
+be identified as being same or different. The marks or striae are
+indexed by the pixel location where \(t_1\) is for the first striae
+referred to as \(x\) and \(t_2\) is for the second striae which is
+referred as \(y\). \(T_1\) and \(T_2\) being the respective lengths of
+the markings, need not be the same but are usually of similar lenghtsThe
+similarity is then judged by the algorithm on the basis of
+cross-correlation of a fixed and constant number consecutive pixels say
+\(k\) taken from the two indexed marks \(x(t_1)\) and \(y(t_2)\) such
+that in theory \(k\) remains smaller that length of the two striae or
+marks which is the same as \(t_1\) and \(t_2\). Depending on the what
+stage of the algorithm we are in, matching of different pixel lengths
+and locations is done, which at the end effectively compares all
+possible windows that would guarantee in quantifying the two marks or
+striae as coming from the same source or not, which also lets us assess
+the error rates by checking for a large number of cases.
 
@@
 do not intuitively expect for non matches
-\gk{done}\hh{XXX get rid of the parentheses - put the content into a sentence after your first statement.}
-
-\hh{XXX the process you describe below is the cross-correlation of two series $x(t)$ and $y(t)$.}
 
@@
 if the window of optimization was defined as 200 and toolmark pixel
-length is around 1000 then we have 1200-(200-1) = 1001. Can be thought
-of ordered arrangements without replacement where n = 1001 and r = 1.
+length is around 1000 then we have 1200-(200-1) = 1001.
 
@@
 same-shift windows.
-
-\begin{verbatim}
-  Classification Match Non-Match
-1          Match    47         3
-2      Non-Match     0        50
-\end{verbatim}
 
@@
 Let ns and nd be the number of same shift and different shift windows
-\(N = n_{s} + n_{d}\)
+\[N = n_{s} + n_{d}\]
 
 The mann whitney U statistic is given by
-\(U =\sum^{ns}_{i=1}R_{s}\left( i\right)\)
+\[U =\sum^{ns}_{i=1}R_{s}\left( i\right)\]
 
@@
 
-\(\overline{U}= \dfrac{U-M}{\sqrt{V}}\)
+\[\overline{U}= \dfrac{U-M}{\sqrt{V}}\]
 
@@
 
-\(M = n_{s}\left(\dfrac{N+1}{2}\right)\)
+\[M = n_{s}\left(\dfrac{N+1}{2}\right)\]
 
@@
 
-\(V = \dfrac{n_{s}n_{d}}{N\left(N-1\right)}\left[\Sigma^{n_{s}}R_{s}\left(i\right)^{2}+\Sigma^{n_{d}}R_{d}\left(j\right)^{2}\right] -\dfrac{n_{s} n_{d}\left(N+1\right)^{2}}{4\left(N-1\right)}\)
+\[V = \dfrac{n_{s}n_{d}}{N\left(N-1\right)}\left[\Sigma^{n_{s}}R_{s}\left(i\right)^{2}+\Sigma^{n_{d}}R_{d}\left(j\right)^{2}\right] -\dfrac{n_{s} n_{d}\left(N+1\right)^{2}}{4\left(N-1\right)}\]
 
-\subsection{\texorpdfstring{Potential limitations in the application to
-striation marks on bullet lands: smaller in width and curved - need to
-adjust parameter settings; something similar is done in
-\citet{afte-chumbley} for toolmark comparisons of slip-joint
-pliers.}{Potential limitations in the application to striation marks on bullet lands: smaller in width and curved - need to adjust parameter settings; something similar is done in  for toolmark comparisons of slip-joint pliers.}}\label{potential-limitations-in-the-application-to-striation-marks-on-bullet-lands-smaller-in-width-and-curved---need-to-adjust-parameter-settings-something-similar-is-done-in-for-toolmark-comparisons-of-slip-joint-pliers.}
+\subsection{Testing setup}\label{testing-setup}
 
-Bullets are much smaller in length, width, are not flat and curved in
-the cross-sectional topography as opposed to tools like screw driver
-tips which produced longer and pronounced markings. This means the
-makings made by barrels in comparison to toolmarks may have a problem in
-distinctiveness. The majority of Bullet profiles and signatures
-extracted by procedures mentioned by \citet{aoas} are almost 1/4 th the
-size of toolmarks as used by \citet{chumbley} or even smaller.
-Striations on Bullets are made on their curved surfaces, whereas the
-algorithm developed by \citet{chumbley} and \citet{hadler} has only been
-tested for flatter and wider surfaces which have negligible curvature.
-Therefore, using methods proposed for toolmarks may need adaptation in
-order to give tangible results for bullets. Moreover, in order to to get
-flat bullet signatures and remove the curvatures some kind of smoothing
-needs to be applied as a pre-step which needs further investigation as
-to whether the level of smoothing does effect the working of the
-algorithm on Bullets.
-
-Also when in the optimization step, the Window of optimization for
-bullets will be shorter as the signatures are smaller. The idea is to
-keep the number of windows of optimization sufficiently large, which
-means shorter Trace segments (partition of signature or toolmark with
-length = size of window of optimization) that lets us compare smaller
-segments of one signature to another.
-
-This introduces a problem as, if we go too small in the window of
-optimization, the unique features of the trace segments are lost and
-seem similar, while too large sizes vastly reduces the weight of small
-features that would otherwise uniquely classify a signature and hence
-identify the region of agreement.
-
-Thus the Window of Optimization has a direct influence on whether we are
-Falsely rejecting the null when it is true (Type I) or Falsely accepting
-the null when it is false (Type II) making identification of the optimum
-size of window of optimization very important.
-
-\subsection{Using cross-validation setup to identify appropriate
-parameter settings for (a) signatures and (b) profiles
-directly}\label{using-cross-validation-setup-to-identify-appropriate-parameter-settings-for-a-signatures-and-b-profiles-directly}
-
-Following on similar lines to the setup of toolamrks, the first step
+Following on similar lines to the setup of toolmarks, the first step
 here is to first identify what difference does different window sizes of
@@
 wider. The idea is to find out possible areas of error while adapting
-the score based method proposed for toolmarks.
+the score based method proposed for toolmarks, using cross-validation
+setup to identify appropriate parameter settings for (a) signatures and
+(b) profiles directly
 
-Bullet signatures being compared at this time are from the Hamby 44 and
-Hamby 252 data, which have the best set of known-matched and known
-non-matches.
-
-Bullet signatures are chosen by 1. Filtering out Land\_id for Profiles
-from the Hamby 44 and Hamby 252 data and removing all NA values 2.
-run\_id = 3 is chosen, Seems like the signatures generated from this
-run\_id give the closest match. Different run\_id's have some different
-settings for generating the signatures (The level of smoothing does not
-seem to be one of them)
-
-The bullet signatures when generated already included the LOWESS
-smoothing. Therefore, the coarseness factor is set to 1 while running
-the chumbley\_non\_random() which generates the same\_shift,
-different\_shift, U-Stat and P\_value parameters.
-
-window Optimization 320 Window Validation 50
-
-\begin{verbatim}
-       match
-signif  FALSE  TRUE
-  FALSE 77564   412
-  TRUE   4022   674
-\end{verbatim}
-
-window Optimization 120 Window Validation 50
-
-\begin{verbatim}
-       match
-signif  FALSE  TRUE
-  FALSE 79249   386
-  TRUE   4523   854
-\end{verbatim}
-
-window Optimization 80 Window Validation 50
-
-\begin{verbatim}
-       match
-signif  FALSE  TRUE
-  FALSE 78817   450
-  TRUE   4680   775
-\end{verbatim}
-
-\section{Results}\label{results}
-
-\subsection{Signatures}\label{signatures}
+\subsubsection{Signatures}\label{signatures}
 
@@
 
+Bullet signatures being compared at this time are therefore from the
+Hamby 44 and Hamby 252 data. The database setup and pre-processing
+system used for choosing the Bullet signatures are as described by
+\citet{aoas}. In order to choose the bullet signatures we first filter
+out Land\_id for Profiles from the Hamby 44 and Hamby 252 data and
+remove all NA values. Then run\_id = 3 is chosen as the signatures
+generated from this run\_id give the closest match. Different run\_id's
+have some different settings for generating the signatures.(The level of
+smoothing does not seem to be one of them)
+
+The bullet signatures when generated by this process already includes a
+loess smoothing. Therefore, the coarseness factor is set to 1 while
+running the chumbley non random algorithm for comparing different
+optimization windows.The algorithm generates the same\_shift,
+different\_shift, U-Stat and P\_value parameters which are then used to
+calculate the errors associated with different sets of window sizes.
+
+\subsubsection{Profiles}\label{profiles}
+
+The profiles are cross-sectional values of the the bullet striation mark
+which are chosen at an optimum height (x as used by \citet{aoas}). This
+x or height is not a randomly chosen level. The rationale behind the
+choice has been explained by \citet{aoas}. A region is first chosen
+where the cross-correlation seems to change very less and in this region
+an optimum height is chosen. The profiles generally resemble a curve
+which is more or less similar to a quadratic curve (a quadratic fit to
+the raw data values of the profile is not an exact fit but it does show
+a similar trend). Profiles are the set of raw values representing the
+striation marks, and signatures are generated from these by removal of
+the inherent curvature and applying some smoothing (the signatures
+generated by \citet{aoas} use a loess function for smoothing).
+
+Similar to signatures the run\_id = 3 was used when applying the
+chumbley algorithm using the database setup given by \citet{aoas} of
+Hamby-44 and Hamby-252 datasets, on the profiles. The run\_id not only
+defines the level of smoothing but also signifies the chosen height at
+which the profiles were selected initially. Another important aspect is
+the range of horizontal values (which is referred to as the y values in
+\citet{aoas}) in the signatures. These have already been pre-processed
+in the database to not include any grooves.
+
+Therefore for the sake of comparison the run\_id = 3 is still chosen so
+as to ensure that the horizontal values remain the same as that of the
+signatures. This also gives us profiles with the grooves removed.
+
+The idea therefore is to first use these raw values of the profile
+directly in the chumbley algorithm, and see how the algorithm performs
+for different coarseness values (smoothing parameter as referred in the
+function lowess used in the chumbley algorithm).
+
+\pagebreak
+
+\section{Results}\label{results}
+
 We used the adjusted Chumbley method as proposed in \citet{hadler} and
@@
 pairwise land-to-land comparisons of the Hamby scans (a total of 85,491
-comparisons) in the following manner:
+comparisons) with the pairwise sets for the comparisons given in the
+table \ref{tab:param}.
 
-\begin{table}
+\begin{table}[!h][!h]
+\caption{\label{tab:param}Overview of parameter settings used for optimization and validation windows for bullet land signatures.}
 
-\caption{\label{tab:unnamed-chunk-5}Overview of parameter settings used for optimization and validation windows for bullet land signatures.}
 \centering
-\resizebox{\linewidth}{!}{\begin{tabular}[t]{lrrrrrrrrrrrrrrrrrrrrrrrrrrrrrrrrr}
+\resizebox{\linewidth}{!}{\resizebox{\linewidth}{!}{\begin{tabular}[t]{lrrrrrrrrrrrrrrrrrr}
 \toprule
-wo & 50 & 50 & 60 & 60 & 80 & 80 & 90 & 90 & 100 & 100 & 110 & 110 & 120 & 120 & 120 & 120 & 120 & 130 & 130 & 140 & 140 & 150 & 150 & 160 & 160 & 200 & 200 & 200 & 240 & 240 & 280 & 280 & 320\\
-wv & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 10 & 20 & 30 & 50 & 60 & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 20 & 30 & 50 & 30 & 50 & 30 & 50 & 50\\
+wo & 50 & 50 & 60 & 60 & 80 & 80 & 90 & 90 & 100 & 100 & 110 & 110 & 120 & 120 & 120 & 120 & 120 & 120\\
+wv & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 10 & 20 & 30 & 40 & 50 & 60\\
 \bottomrule
-\end{tabular}}
+\end{tabular}}}
+\centering
+\begin{tabular}[t]{lrrrrrrrrrrrrrrr}
+\toprule
+wo & 130 & 130 & 140 & 140 & 150 & 150 & 160 & 160 & 200 & 200 & 200 & 240 & 240 & 280 & 280\\
+wv & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 20 & 30 & 50 & 30 & 50 & 30 & 50\\
+\bottomrule
+\end{tabular}
 \end{table}
 
-\subsubsection{Signatures}\label{signatures-1}
+\begin{table}[!h]
+\caption{\label{tab:confusion}Confusion Table for different optimization window sizes with validation window size as 30.}
+
+\centering
+\begin{tabular}[t]{llrl}
+\toprule
+\hspace{1em}\hspace{1em}signif & match & Freq & Type\\
+\midrule
+\addlinespace[0.3em]
+\multicolumn{4}{l}{\textbf{Size of Optimization Window = 280}}\\
+\hspace{1em}FALSE & FALSE & 78909 & True Negative\\
+\hspace{1em}TRUE & FALSE & 4192 & False Positive (Type I)\\
+\hspace{1em}FALSE & TRUE & 446 & False Negative (Type II)\\
+\hspace{1em}TRUE & TRUE & 731 & True Positive\\
+\bottomrule
+\end{tabular}
+\centering
+\begin{tabular}[t]{llrl}
+\toprule
+signif & match & Freq & Type\\
+\midrule
+\addlinespace[0.3em]
+\multicolumn{4}{l}{\textbf{Size of Optimization Window = 120}}\\
+\hspace{1em}FALSE & FALSE & 79249 & True Negative\\
+\hspace{1em}TRUE & FALSE & 4523 & False Positive (Type I)\\
+\hspace{1em}FALSE & TRUE & 386 & False Negative (Type II)\\
+\hspace{1em}TRUE & TRUE & 854 & True Positive\\
+\bottomrule
+\end{tabular}
+\centering
+\begin{tabular}[t]{llrl}
+\toprule
+signif & match & Freq & Type\\
+\midrule
+\addlinespace[0.3em]
+\multicolumn{4}{l}{\textbf{Size of Optimization Window = 80}}\\
+\hspace{1em}FALSE & FALSE & 79477 & True Negative\\
+\hspace{1em}TRUE & FALSE & 4503 & False Positive (Type I)\\
+\hspace{1em}FALSE & TRUE & 463 & False Negative (Type II)\\
+\hspace{1em}TRUE & TRUE & 781 & True Positive\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\subsection{Signatures}\label{signatures-1}
 
 Figure \ref{fig:type2} gives an overview of type II error rates observed
-when varying the window size in the optimization step.
+when varying the window size in the optimization step. Two levels of
+validation window size 30 and 50 were chosen as to compare the error
+rates for different nominal type I errors. We notice that the trends for
+these nominal type I errors are similar and in most cases a validation
+window of 50 has higher type II error than for 30. A change in this
+trend is seen for a 0.05 \(\alpha\) level, although the difference
+between the two windows is very small for this case. We can also notice
+an obvious trend of increase in the Type II error as the window of
+optimization increases and see a minimum around the optimization window
+size of 120 pixels. Hence we are inclined to choose a smaller validation
+window size and optimization window as 120.
+
+Table \ref{tab:confusion} shows the confusion tables with the
+classification of type I and type II errors and how the numbers change
+with a change in the optimization window. The windows represent areas to
+the left of the window with minimum type II, near the minimum type II
+window and to the far right of the minimum type II error
 
@@
 Figure \ref{fig:type1} compares nominal (fixed) type I error and
-actually observed type I errors for the parameter settings in table XXX.
-With an increasing size of the window used in the optimization step the
-observed type I error rate decreases (slighty). This might be related to
-the increasing number of tests that fail for larger window sizes, in
-particular for non-matching striae (see fig \ref{fig:missings}).
+actually observed type I errors for the parameter settings in table
+\ref{tab:param}. With an increasing size of the window used in the
+optimization step the observed type I error rate decreases (slighty).
+This means as the optimization window increase the observed type I error
+rate gets smaller. A smaller validation window on the other hand, tends
+to be associated with a higher type I error rate. This can be better
+imagined for a given window of optimization, where the actual Type I
+error is comparable to the nominal level for only a select few
+validation window sizes. For these comparable validation window sizes of
+30 and 50 as done here, the actual type I error increases very slightly
+and can be seen in Figure \ref{fig:type1}. This increase is not as much
+when compared to the variation seen with the optimization window sizes.
+This effect might be related to the increasing number of tests that fail
+for larger optimization window sizes, in particular for non-matching
+striae (see fig \ref{fig:missings}).
 
@@
 \caption{Comparison of observed and nominal type I error rates  across a range of window sizes for optimization $wo$.
-The horizontal line in each facet indicates the nominal type I error rate.
-As the optimization window increase the observed type I error rate gets smaller.
-A smaller validation window tends to be associated with a higher type I error rate.}\label{fig:type1}
+The horizontal line in each facet indicates the nominal type I error rate.}\label{fig:type1}
 \end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figures/wo120-1}
+
+}
+
+\caption{The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120}\label{fig:wo120}
+\end{figure}
+
+The actual type I error and type II error for signatures were also
+compared for different validation window sizes. Figure \ref{fig:wo120}
+shows the actual rates for different nominal \(\alpha\) levels. We can
+see that the type II error rises with higher validation windows for the
+smaller nominal \(\alpha\) levels while for the nominal \(\alpha\) =
+0.05 its almost constant.
 
@@
 
-\caption{Number of failed tests by the window optimization size, wo, and ground truth.
-Test results from different sources have a much higher chance to fail, raising the question, whether failed tests should be treated as rejections of the null hypothesis of same source.}\label{fig:missings}
+\caption{Number of failed tests by the window optimization size, wo, and ground truth.}\label{fig:missings}
 \end{figure}
-
-Figure \ref{fig:missings} gives an overview of the number of failed
-tests, i.e.~tests in which a particular parameter setting did not return
-a valid result. This happens, when the shift to align two signatures is
-so large, that the remaining overlap is too small to accommodate windows
-for validation. The problem is therefore exacerbated by a larger
-validation window. Figure \ref{fig:missings} also shows that the number
-of failed tests is approximately linear in the size of the optimization
-window.
 
@@
 
-\begin{center}\includegraphics[width=\textwidth]{figures/wo120-1} \end{center}
+Figure \ref{fig:missings} gives an overview of the number of failed
+tests, i.e.~tests in which a particular parameter setting did not return
+a valid result. This happens, when the shift to align two signatures is
+so large, that the remaining overlap is too small to accommodate windows
+for validation. The problem is therefore exacerbated by a larger
+validation window. Figure \ref{fig:missings} also shows that the number
+of failed tests is approximately linear in the size of the optimization
+window. Test results from different sources have a much higher chance to
+fail, raising the question, whether failed tests should be treated as
+rejections of the null hypothesis of same source. For known non-matches
+there is a higher possibility that in the optimization pair of windows
+where cross-correlations are maximum are too far apart, and same shifts
+of this order hit the end of the signature.
 
-\subsubsection{Profiles}\label{profiles}
+\subsection{Profiles}\label{profiles-1}
 
-The profiles are cross-sectional values of the the bullet striation mark
-which are chosen at an optimum height (x as used by \citet{aoas}). This
-x or height is not a randomly chosen level. The rationale behind the
-choice has been explained by \citet{aoas}. A region is first chosen
-where the cross-correlation seems to change very less and in this region
-an optimum height is chosen. The profiles generally resemble a curve
-which is more or less similar to a quadratic curve (a quadratic fit to
-the raw data values of the profile is not an exact fit but it does show
-a similar trend). Profiles are the set of raw values representing the
-striation marks, and signatures are generated from these by removal of
-the inherent curvature and applying some smoothing (the signatures
-generated by \citet{aoas} use a loess function for smoothing).
+Figure \ref{fig:prof_missings} (a) shows the type II error rates for
+profiles for the optimization window 120 and validation window 30 with
+varying level of coarseness. We can see that the type II error for all
+the nominal \(\alpha\) levels is lowest in the range of 0.20 to 0.35.
+Therefore, a value of 0.25 can be used keeping in mind it keeps the type
+II error lowest while runnning simulations. Thus for comparisons of
+different window sizes etc as seen in the differnt parts of Figure
+\ref{fig:prof_missings} this coarseness value is used.
 
-Coming back to the database of Hamby-44 and Hamby-252 datasets, the
-run\_id = 3 was used when applying the chumbley algorithm on the
-signatures. The run\_id not only defines the level of smoothing but also
-signifies the chosen height at which the profiles were selected
-initially. Another important aspect is the range of horizontal values
-(which is referred to as the y values in \citet{aoas}) in the
-signatures. Thes have already been pre-processed in the database to not
-include any grooves.
+On the other hand Figure \ref{fig:prof_missings} (b) shows if the
+coarseness level set in the chumbley agorithm has any effect on the
+signatures, which are pre-processed and already smoothed to a certain
+extent. From Figure \ref{fig:prof_missings} (b) we can notice that for
+different nominal \(\alpha\) levels, the type II error fluctuates
+slightly but does not change much, thereby helping us conclude that the
+coarseness levels set in the LOWESS smoothing in the chumbley alggorithm
+does effect the type II error much for signatures.
 
-Therefore for the sake of comparison the run\_id = 3 is still chosen so
-as to ensure that the horizontal values remain the same as that of the
-signatures. This also gives us profiles with the grooves removed.
+\subsubsection{Comparison of profiles and
+signatures}\label{comparison-of-profiles-and-signatures}
 
-The idea therefore is to first use these raw values of the profile
-directly in the chumbley algorithm, and see how the algorithm performs
-for different coarseness values (smoothing parameter as referred in the
-function LOWESS used in the chumbley algorithm)
+Another reason for failed tests can be incorrect identification of
+maximum correlation windows in the optimization step as seen in figure
+\ref{fig:prof_missings}(d) because of the level of smoothing, as too
+much smoothing would subdue intricate features that might otherwise help
+in the correlation calculations and correct identification of maximum
+correlation windows irrespective of the size. This would again cause a
+simiar effect as explained for figure \ref{fig:missings} with validation
+windows, irrespective of size, during the shifts end up at the ends of
+the markings resulting in an invalid calculation and failed comparison
+attempt.
+
+In figure \ref{fig:prof_missings}(d) and (f), we compare profiles and
+signatures on the basis of number of failed tests. The profiles chosen
+for figure \ref{fig:prof_missings}(f) have a constant coarseness of 0.25
+and window of optimization as 120. The signatures in this case are not
+smoothed using the chumbley algorithm step of LOWESS smoothing. Instead
+signatures are used as calculated by \citet{aoas}. The smoothing in
+these signatures were determined and fixed on the basis of their
+performance in the random forest based algorithm proposed by
+\citet{aoas}. The comparison of profiles and signatures with variation
+of validation window size therefore is made on even footing. The trends
+are similar to figure \ref{fig:missings} in the sense that for known
+non-matches the number of failed tests are more for both signatures and
+profiles and increasing linearly with the validation window size. The
+problem is however, worse for profiles which has higher number of failed
+tests than signatures for all validation windows.
+
+The total error for different validation window sizes for signatures and
+profiles can be seen in figure \ref{fig:prof_missings} (e).The
+optimization window size is 120 and profiles are calculated at a default
+0.25 coarseness level while signatures as before are not smoothed again
+in the modified chumbley algorithm. We can see that the total error is
+always higher for profiles as compared to signatures for all sizes of
+validation window.
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figures/prof_missings-1}
+
+}
+
+\caption{Row 3:  Total error and Number of failed tests by the window validation size, wv, and ground truth, Row 2: Total error and Number of failed tests with Coarseness for both profiles and signatures, Row 1: Type II error for different coarseness levels as used in the modified chumbley algorithm for profiles and signatures}\label{fig:prof_missings}
+\end{figure}
+
+\pagebreak
+
+\subsection{Conclusion}\label{conclusion}
+
+The results suggest that the Nominal type I error \(\alpha\) value shows
+dependence on the size of the window of optimization. For a given window
+of optimization the actual Type I error is comparable to the nominal
+level for only a select few validation window sizes and for comparable
+validation window sizes of 30 and 50 as done here, the actual type I
+error does not seem to vary as much as it varies with the optimization
+window sizes . A Test Fail, i.e.~tests in which a particular parameter
+setting did not return a valid result, happens, when the shift to align
+two signatures is so large, that the remaining overlap is too small to
+accommodate windows for validation, depends on whether known-match or
+known non-matches has predictive value, with test results from different
+sources having a much higher chance to fail. On conducting an analysis
+of all known bullet lands using the adjusted chumbley algorithm, Type II
+error was identified to be least bad for window of validation 30 and
+window of optimization 120. In case of unsmoothed raw marks (profiles),
+Type II error increases with the amount of smoothing and least for
+LOWESS smoothing coarseness value about 0.25 or 0.3. In an effort to
+identify the level of adaptiveness of the algorithm, comparisons were
+made between signatures and profiles. Their comparison with respect to
+validation window size for a fixed optimization window size suggested
+that, profiles have a total error (i.e all incorrect classification of
+known-matches and known non-matches) greater than or equal to the total
+error of signatures for all sizes of validation window. Profiles also
+fail more number of times than signatures in a test fail (for different
+coarseness keeping windows fixed and also for different validation
+windows keeping coarseness fixed) which lets us conclude that the
+behaviour of the algorithm for the profiles instead of pre-processed
+signatures is not better. Finally it should be noted that the current
+version of the adjusted chumbley algorithm seems to fall short when
+compared to other machine-learning based methods \citet{aoas}, and some
+level of modification to the deterministic algorithm needs to be
+identified and tested that would reduce the number of incorrect
+classifications.
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figures/wo120-profile-1}
+
+}
+
+\caption{The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120}\label{fig:wo120-profile}
+\end{figure}\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figures/failtest-coarseness-1}
+
+}
+
+\caption{The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120}\label{fig:failtest-coarseness}
+\end{figure}\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figures/failtest-wv-1}
+
+}
+
+\caption{The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120}\label{fig:failtest-wv}
+\end{figure}\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figures/failtest-wo-1}
+
+}
+
+\caption{The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120}\label{fig:failtest-wo}
+\end{figure}
 

### 2018-02-05 doc_v2/eval_windows.R
removing doc_v2

@@ -1,72 +1,0 @@
-
-falspostive<- function(sig_alpha){
-  if(is.null(all)){
-    library(tidyverse)
-    files <- dir("./data/signatures", pattern="csv")
-    all <- data.frame()
-    for (file in files) {
-      tmp <- read.csv(file=file.path(path="./data/signatures", file)) 
-      if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% select(-x)
-      all <- rbind(all, tmp)
-    }
-  }
-    summ <- all %>% filter(!is.na(p_value)) %>%
-      mutate(signif = p_value < sig_alpha) %>%
-      group_by(wv, wo, match, signif) %>% tally()
-    summ$error <- with(summ, match != signif)
-    rates <- summ %>% group_by(wv, wo, match) %>% summarize(
-      rate = n[error==TRUE]/sum(n)
-    )
-    combo<- list(summ, rates)
-    names(combo)<- c("summ", "rates")
-    return(combo)
-}
-
-a<- falspostive(0.05)
-
-a$summ %>% ggplot(aes( x  = factor(wv), weight = n, fill=error)) + geom_bar(position="fill") +
-  facet_grid(match~wo, labeller="label_both")
-a$rates %>% ggplot(aes(x = wv, y = rate)) + geom_point() + facet_grid(match~wo, labeller="label_both") +
-  geom_line()
-    
-# summ %>% ggplot(aes( x  = factor(wv), weight = n, fill=error)) + geom_bar(position="fill") +
-#       facet_grid(match~wo, labeller="label_both")
-# 
-#     rates %>% ggplot(aes(x = wv, y = rate)) + geom_point() +
-#       facet_grid(match~wo, labeller="label_both") +
-#       geom_line()
-
-    # library(tidyverse)
-    # files <- dir("./data", pattern="csv")
-    # all <- data.frame()
-    # for (file in files) {
-    #   tmp <- read.csv(file=file.path(path="./data", file)) 
-    #   if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% select(-x)
-    #   all <- rbind(all, tmp)
-    # }
-    # 
-    # summ <- all %>% filter(!is.na(p_value)) %>%
-    #   mutate(signif = p_value < 0.01) %>%
-    #   group_by(wv, wo, match, signif) %>% tally()
[…]

### 2018-02-05 doc/paper.Rmd
adding modified introduction

@@
 ## Motivation
+Compairing pairs of toolmarks with the intention of matching it to a tool has been studied many times in the past.
+Extensive examples in literature for tools and toolmark research ranging from screwdrivers to groove pliers to slip-joint pliers can be found in the work of  \citet{manytoolmarks1}, \citet{manytoolmarks2}, \citet{miller}, \cite{afte-chumbley} and many more.
+In comparison to this, same source matching of bullets to firearms has not been examined as prominently as that of toolmarks, with even less information available on validity of methods and error rates associated with firearms examination.
+The National Academy of Sciences in its report in 2009 \citep{NAS:2009} discussed the need for determining error rates in methods proposed for firearms examination.
+In the context of same source matching and error rates determination, as seen in the case of most forensic applications, the first step involves identification of unique features that are characteristic of the object at hand.
+For the case of bullets and firearms, striation marks on the surface of the bullet are considered to be such markings that can be used in methods for same source matching.
+These marks are often a product of rifling and impurities and defects due to manufacturing in the barrel of the gun, which leads to engravings on the bullet surface \citep{afte-article1992}.
+In current practice, firearm examiners invariably make visual comparisons of bullet striae and use visual assesment tools to dignify bullets as being matches and non-matches.
+One way of accomplishing anykind of comparison between bullets is to do a comparison between surface marking of two or more bullet lands.
+Bullet Lands are considered to be areas between grooves made by the rifling action of the barrel.
+These marking are considered to be unique.
+The land engraved markings or sometimes termed as Bullet profiles \citep{aoas} are striation marks made on Bullet lands and often used for these land to land comparisons.
+Bullet Signatures is another word used in literature as seen in the work of \citet{chu2013} and \citet{aoas}.
+In our context bullet signatures refer to a processed version of the raw land engraved markings or profiles.
+The generation of bullet signatures involves first extraction of a bullet profile by taking the cross-sectional of the surface at a given height and then using loess fits to model the structure.
+The residuals of this fit are called signatures, which are considered to be noise free and a good reflection of the class charecteristics and unique features of a bullet.
+A more detailed version of the extraction technique of signatures is discussed by \citet{aoas}, where comprehensive details about the height at which profile is to be selected, removing curvature, smoothing, identifying groove locations are explained.
 
-Striation marks are used to determine same source for bullets.
+In the study conducted by \citet{aoas} a machine learning based algorithm was developed for same source matching of bullets and error rates were discussed using the database from the Hamby Study \citep{hamby}.
+In this paper, we first try to adapt a deterministic algorithm and method developed for toolmarks by \citet{chumbley} and improved by \citet{hadler}, to bullets.
+Then we consequently discuss about the efforts in doing so along with the associated error rates.
+The data used in this paper also belongs to the Hamby Study \citep{hamby}.
+This gives us a common platform for comparing the performance of the chumbley method on bullets with an already existing method proposed by \citet{aoas} for bullets.
+The proposed algorithm and method of \citet{chumbley}, in their paper, compares two toolmarks with the intention of determining if it comes from the same source (same tool).
+The method also provides a means to determine error rates and claims to reduce subject bias.
+As mentioned earlier, subject bias and error rate determination have been a long standing issue in firearm examination \citep{NAS:2009}.
+This remains one of the motivations to explore the adaptibility of the Chumbley score methodology to bullets.
+\citet{chumbley} used an empirical based setup to validate their proposed algorithm and quantitative method which calculates a U-statisic for the purpose of classification of toolmarks as matching or non-matching.
+The data for their study was obtained from 50 sequentially manufactured screwdriver tips, and preselected comparison window sizes were given as inputs to the algorithm.
+The algorithm then compares the two toolmarks and comes up with a U-statistic and an associated p-value to designate them as matches or non-matches.
+The performance for every 100 comparisons, of the algorithm proposed by \citet{chumbley} and the improvement proposed by \citet{hadler} are listed in the table below.
+
+```{r, results='asis'}
+library(knitr)
+library(kableExtra)
+library(xtable)
+
+a1<- matrix(NA, ncol = 3, nrow = 2)
+toolmark1.confusion<- data.frame(a1)
+colnames(toolmark1.confusion)<- c("Classification", "Match", "Non-Match")
+toolmark1.confusion$Classification<- c("Match", "Non-Match")
+toolmark1.confusion$Match<- c(41,2)
+toolmark1.confusion$`Non-Match`<- c(9,48)
+
+a2<- matrix(NA, ncol = 3, nrow = 2)
+toolmark2.confusion<- data.frame(a2)
+colnames(toolmark2.confusion)<- c("Classification", "Match", "Non-Match")
+toolmark2.confusion$Classification<- c("Match", "Non-Match")
+toolmark2.confusion$Match<- c(47,0)
+toolmark2.confusion$`Non-Match`<- c(3,50)
+
+chumbley.confusion<-knitr::kable(toolmark1.confusion, digits=0, format = "latex" ,booktabs=TRUE) #%>% kableExtra::kable_styling(latex_options = c("hold"))
+hadler.confusion<-knitr::kable(toolmark2.confusion, digits=0, format = "latex" ,booktabs=TRUE) #%>% kableExtra::kable_styling(latex_options = c("hold"))
+
+cat(c("\\begin{table}[!htb]
+    \\begin{minipage}{.5\\linewidth}
+      \\caption{Chumbley et al. 2010}
+      \\centering",
+        chumbley.confusion,
+    "\\end{minipage}%
+    \\begin{minipage}{.5\\linewidth}
+      \\centering
+        \\caption{Hadler et. al. (2017)}",
+       hadler.confusion,
+    "\\end{minipage} 
+\\end{table}"
+))
+
+```
+<!-- Striation marks are used to determine same source for bullets.
 Current standards ask for a visual inspection of the marks under a comparison microscope by a firearms examiner.
@@
 In case of methods that use machine learning algorithms some methods which are bootsrap based methods often tend to give over-estimated error rates \citet{efron}, but there are alternate error estimation techniques as described by \citet{aoas}, \citet{efron} and  \citet{vorburger2016} which give better results.
+-->
 
-In case of methods that do not use machine learning algorithm, matching two striation marks with each other in order to identify if it is from the same source, can also be done using a well defined comparative statistical algorithm.
+<!-- In case of methods that do not use machine learning algorithm, matching two striation marks with each other in order to identify if it is from the same source, can also be done using a well defined comparative statistical algorithm.
 The statistical algorithm is therefore expected  to go through a step by step procedure of identifying the class characteristics i.e striation marks for bullets, choosing the striae, and then follow a systematic procedure of comparing two striae with each other on the basis of criteria like cross-correlation (or others).
 Such a methodology would make it possible to determine error rates too and give definite results regarding what proportion of cases would the analysis be successful in identifying a match correctly and in how many cases would it end up being a false positive.
+-->
 
-Identification criteria like consecutively matching striae (CMS) as first seen by Biasotti 1959 \citet{biasotti} and later mentioned in \citet{chu2013} more often than not include subject bias and are error prone and as seen in the work of \citet{miller} the number of CMS may turn out to be high even when the bullet is not fired by the same firearm.
+<!-- Identification criteria like consecutively matching striae (CMS) as first seen by Biasotti 1959 \citet{biasotti} and later mentioned in \citet{chu2013} more often than not include subject bias and are error prone and as seen in the work of \citet{miller} the number of CMS may turn out to be high even when the bullet is not fired by the same firearm.
 Parameters like cross-correlation factor as seen in the extensive comparisons made by \citep{aoas} seem to perform much better for matching purposes.
+-->
 
-An important aspect of same source matching in firearms is identification of bullet microtopographies that are unique to be chosen for a match, Bullet profiles and signatures prove to be such features.
+<!-- An important aspect of same source matching in firearms is identification of bullet microtopographies that are unique to be chosen for a match, Bullet profiles and signatures prove to be such features.
 The choice of a bullet signature to uniquely define a bullet depends on finding first, a stable region on a Bullet land which minimal noise and many pronounced striation marks.
-\citet{aoas}
+\citet{aoas} -->
 
-The method employed by \citet{aoas} uses the cross correlation factor as a means to identify the stable region.
+<!-- The method employed by \citet{aoas} uses the cross correlation factor as a means to identify the stable region.
 The markings or striae in this region are supposed to have a high CCF with each other.
@@
 A loess fit on the bullet profile produces  residuals which are termed as signatures and can be used as a unique identifier that can be used while matching two bullets to a firearm.
+-->
 
-Compairing pairs of toolmarks on the otherhand, with the intention of matching it to a tool has been studied relatively more in the past as compared to bullets, and \citet{chumbley} have described in their paper an algorithm and analytic method that compares two toolmarks and come to the conclusion if they are from the same tool or not.
+<!-- Compairing pairs of toolmarks on the otherhand, with the intention of matching it to a tool has been studied relatively more in the past as compared to bullets, and \citet{chumbley} have described in their paper an algorithm and analytic method that compares two toolmarks and come to the conclusion if they are from the same tool or not.
 The method also determines the error rates, reduces subject bias and designate the two toolmarks as matches or non-matches with respect to a source.
 \citet{chumbley} used an empirical based setup to validate their proposed analytic and quantitative algorithm.
+-->
 
-\citet{chumbley} found that the algorithm gives false-positives in slightly over 2 cases (or 2%) of the time and about 9 false-negatives (or 9%) for every 100 comparisons.
+<!-- \citet{chumbley} found that the algorithm gives false-positives in slightly over 2 cases (or 2%) of the time and about 9 false-negatives (or 9%) for every 100 comparisons.
 \citet{hadler} on the other hand, in the improved version of the algorithm first described by \citet{chumbley} found false-positives for 0 cases and 3 false-negatives (or 6%) for 50 comparisons of knwon matches.
-```{r}
-a<- matrix(NA, ncol = 3, nrow = 2)
-toolmark.falsepositive<- data.frame(a)
-colnames(toolmark.falsepositive)<- c("Classification", "Match", "Non-Match")
-toolmark.falsepositive$Classification<- c("Match", "Non-Match")
-toolmark.falsepositive$Match<- c(47,0)
-toolmark.falsepositive$`Non-Match`<- c(3,50)
+-->
+<!-- ```{r} -->
+<!-- a<- matrix(NA, ncol = 3, nrow = 2) -->
+<!-- toolmark.falsepositive<- data.frame(a) -->
+<!-- colnames(toolmark.falsepositive)<- c("Classification", "Match", "Non-Match") -->
+<!-- toolmark.falsepositive$Classification<- c("Match", "Non-Match") -->
+<!-- toolmark.falsepositive$Match<- c(47,0) -->
+<!-- toolmark.falsepositive$`Non-Match`<- c(3,50) -->
 
-knitr::kable(toolmark.falsepositive, digits=0, format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("hold"))
-```
+<!-- knitr::kable(toolmark.falsepositive, digits=0, format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("hold")) -->
+<!-- ``` -->
 
@@
 
-```{r wo120-profile, fig.width = 6, fig.height=4, fig.cap="The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120"}
+```{r wo120-profile, fig.width = 6, fig.height=4, fig.cap="The figure shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120"}
 wo120_profile <- errors2 %>% filter(wo==120, coarse == 0.25) %>% 
@@
 ```
-```{r failtest-coarseness, fig.width = 6, fig.height=4, fig.cap="The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120"}
+```{r failtest-coarseness, fig.width = 6, fig.height=4, fig.cap="The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120", eval= FALSE}
 
@@
 ```
-```{r failtest-wv, fig.width = 6, fig.height=4, fig.cap="The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120"}
+```{r failtest-wv, fig.width = 6, fig.height=4, fig.cap="The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120", eval= FALSE}
 
@@
 ```
-```{r failtest-wo, fig.width = 6, fig.height=4, fig.cap="The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120"}
+```{r failtest-wo, fig.width = 6, fig.height=4, fig.cap="The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120", eval= FALSE}
 

### 2018-02-05 doc/paper.tex
adding modified introduction

@@
 
-Striation marks are used to determine same source for bullets. Current
-standards ask for a visual inspection of the marks under a comparison
-microscope by a firearms examiner. One of the problems raised by the NAS
-report in 2009 is the subjectivity involved in this comparison and the
-need for determining error rates \citet{NAS:2009}. The issue of
-subjectivity of a firearms examiner has been reviewed by many authors
-where things like scientific principle testability and error rates are
-considered to be defining aspects for an objective analysis such that
-without identifying the uncertainities associated with a method of
-comparison, it is inconclusive to regards an analysis complete. This
-means that with the usual method of visual comparison that is adopted by
-firearms examiners, lies the problem of lack of any systemized
-quantification and as such there seems no certain way of associating it
-with any kind of mathematical probability. Since determining error rates
-has been duly noted as a fundamental problem in forensic science
-\citet{NAS:2009}, there is a need of methods that address this problem.
-Some of the methods that have been used to estimate error rates in
-firearm examinations include Automatic cartridge case comparison where
-same source identification of cartridge cases and associated error rates
-have been provided by \citet{riva}, although as noted by \citet{aoas},
-in this case alignments of striae involves roation of planes, which
-cannot be generalized for bullets. In case of methods that use machine
-learning algorithms some methods which are bootsrap based methods often
-tend to give over-estimated error rates \citet{efron}, but there are
-alternate error estimation techniques as described by \citet{aoas},
-\citet{efron} and \citet{vorburger2016} which give better results.
+Compairing pairs of toolmarks with the intention of matching it to a
+tool has been studied many times in the past. Extensive examples in
+literature for tools and toolmark research ranging from screwdrivers to
+groove pliers to slip-joint pliers can be found in the work of
+\citet{manytoolmarks1}, \citet{manytoolmarks2}, \citet{miller},
+\cite{afte-chumbley} and many more. In comparison to this, same source
+matching of bullets to firearms has not been examined as prominently as
+that of toolmarks, with even less information available on validity of
+methods and error rates associated with firearms examination. The
+National Academy of Sciences in its report in 2009 \citep{NAS:2009}
+discussed the need for determining error rates in methods proposed for
+firearms examination. In the context of same source matching and error
+rates determination, as seen in the case of most forensic applications,
+the first step involves identification of unique features that are
+characteristic of the object at hand. For the case of bullets and
+firearms, striation marks on the surface of the bullet are considered to
+be such markings that can be used in methods for same source matching.
+These marks are often a product of rifling and impurities and defects
+due to manufacturing in the barrel of the gun, which leads to engravings
+on the bullet surface \citep{afte-article1992}. In current practice,
+firearm examiners invariably make visual comparisons of bullet striae
+and use visual assesment tools to dignify bullets as being matches and
+non-matches. One way of accomplishing anykind of comparison between
+bullets is to do a comparison between surface marking of two or more
+bullet lands. Bullet Lands are considered to be areas between grooves
+made by the rifling action of the barrel. These marking are considered
+to be unique. The land engraved markings or sometimes termed as Bullet
+profiles \citep{aoas} are striation marks made on Bullet lands and often
+used for these land to land comparisons. Bullet Signatures is another
+word used in literature as seen in the work of \citet{chu2013} and
+\citet{aoas}. In our context bullet signatures refer to a processed
+version of the raw land engraved markings or profiles. The generation of
+bullet signatures involves first extraction of a bullet profile by
+taking the cross-sectional of the surface at a given height and then
+using loess fits to model the structure. The residuals of this fit are
+called signatures, which are considered to be noise free and a good
+reflection of the class charecteristics and unique features of a bullet.
+A more detailed version of the extraction technique of signatures is
+discussed by \citet{aoas}, where comprehensive details about the height
+at which profile is to be selected, removing curvature, smoothing,
+identifying groove locations are explained.
 
-In case of methods that do not use machine learning algorithm, matching
-two striation marks with each other in order to identify if it is from
-the same source, can also be done using a well defined comparative
-statistical algorithm. The statistical algorithm is therefore expected
-to go through a step by step procedure of identifying the class
-characteristics i.e striation marks for bullets, choosing the striae,
-and then follow a systematic procedure of comparing two striae with each
-other on the basis of criteria like cross-correlation (or others). Such
-a methodology would make it possible to determine error rates too and
-give definite results regarding what proportion of cases would the
-analysis be successful in identifying a match correctly and in how many
-cases would it end up being a false positive.
+In the study conducted by \citet{aoas} a machine learning based
+algorithm was developed for same source matching of bullets and error
+rates were discussed using the database from the Hamby Study
+\citep{hamby}. In this paper, we first try to adapt a deterministic
+algorithm and method developed for toolmarks by \citet{chumbley} and
+improved by \citet{hadler}, to bullets. Then we consequently discuss
+about the efforts in doing so along with the associated error rates. The
+data used in this paper also belongs to the Hamby Study \citep{hamby}.
+This gives us a common platform for comparing the performance of the
+chumbley method on bullets with an already existing method proposed by
+\citet{aoas} for bullets. The proposed algorithm and method of
+\citet{chumbley}, in their paper, compares two toolmarks with the
+intention of determining if it comes from the same source (same tool).
+The method also provides a means to determine error rates and claims to
+reduce subject bias. As mentioned earlier, subject bias and error rate
+determination have been a long standing issue in firearm examination
+\citep{NAS:2009}. This remains one of the motivations to explore the
+adaptibility of the Chumbley score methodology to bullets.
+\citet{chumbley} used an empirical based setup to validate their
+proposed algorithm and quantitative method which calculates a U-statisic
+for the purpose of classification of toolmarks as matching or
+non-matching. The data for their study was obtained from 50 sequentially
+manufactured screwdriver tips, and preselected comparison window sizes
+were given as inputs to the algorithm. The algorithm then compares the
+two toolmarks and comes up with a U-statistic and an associated p-value
+to designate them as matches or non-matches. The performance for every
+100 comparisons, of the algorithm proposed by \citet{chumbley} and the
+improvement proposed by \citet{hadler} are listed in the table below.
 
-Identification criteria like consecutively matching striae (CMS) as
-first seen by Biasotti 1959 \citet{biasotti} and later mentioned in
-\citet{chu2013} more often than not include subject bias and are error
-prone and as seen in the work of \citet{miller} the number of CMS may
-turn out to be high even when the bullet is not fired by the same
-firearm. Parameters like cross-correlation factor as seen in the
-extensive comparisons made by \citep{aoas} seem to perform much better
-for matching purposes.
-
-An important aspect of same source matching in firearms is
-identification of bullet microtopographies that are unique to be chosen
-for a match, Bullet profiles and signatures prove to be such features.
-The choice of a bullet signature to uniquely define a bullet depends on
-finding first, a stable region on a Bullet land which minimal noise and
-many pronounced striation marks. \citet{aoas}
-
-The method employed by \citet{aoas} uses the cross correlation factor as
-a means to identify the stable region. The markings or striae in this
-region are supposed to have a high CCF with each other. Therefore a
-Bullet profile is chosen from this region by taking the cross section at
-a given height. A loess fit on the bullet profile produces residuals
-which are termed as signatures and can be used as a unique identifier
-that can be used while matching two bullets to a firearm.
-
-Compairing pairs of toolmarks on the otherhand, with the intention of
-matching it to a tool has been studied relatively more in the past as
-compared to bullets, and \citet{chumbley} have described in their paper
-an algorithm and analytic method that compares two toolmarks and come to
-the conclusion if they are from the same tool or not. The method also
-determines the error rates, reduces subject bias and designate the two
-toolmarks as matches or non-matches with respect to a source.
-\citet{chumbley} used an empirical based setup to validate their
-proposed analytic and quantitative algorithm.
-
-\citet{chumbley} found that the algorithm gives false-positives in
-slightly over 2 cases (or 2\%) of the time and about 9 false-negatives
-(or 9\%) for every 100 comparisons. \citet{hadler} on the other hand, in
-the improved version of the algorithm first described by
-\citet{chumbley} found false-positives for 0 cases and 3 false-negatives
-(or 6\%) for 50 comparisons of knwon matches.
-
-\begin{table}[!h]
+\begin{table}[!htb]
+    \begin{minipage}{.5\linewidth}
+      \caption{Chumbley et al. 2010}
       \centering 
+\begin{tabular}{lrr}
+\toprule
+Classification & Match & Non-Match\\
+\midrule
+Match & 41 & 9\\
+Non-Match & 2 & 48\\
+\bottomrule
+\end{tabular} \end{minipage}%
+    \begin{minipage}{.5\linewidth}
+      \centering
+        \caption{Hadler et. al. (2017)} 
 \begin{tabular}{lrr}
@@
 \bottomrule
-\end{tabular}
+\end{tabular} \end{minipage} 
 \end{table}
@@
 
-\caption{The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120}\label{fig:wo120-profile}
-\end{figure}\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figures/failtest-coarseness-1}
-
-}
-
-\caption{The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120}\label{fig:failtest-coarseness}
-\end{figure}\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figures/failtest-wv-1}
-
-}
-
-\caption{The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120}\label{fig:failtest-wv}
-\end{figure}\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figures/failtest-wo-1}
-
-}
-
-\caption{The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120}\label{fig:failtest-wo}
+\caption{The figure shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120}\label{fig:wo120-profile}
 \end{figure}

### 2018-02-05 doc/paper.Rmd
cleaning files

@@
 
-```{r wo120-profile, fig.width = 6, fig.height=4, fig.cap="The figure shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120"}
+```{r wo120-profile, fig.width = 6, fig.height=4, fig.cap="The figure shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120", eval = FALSE}
 wo120_profile <- errors2 %>% filter(wo==120, coarse == 0.25) %>% 

### 2018-02-08 doc/paper.Rmd
minor changes to language

@@
 One way of accomplishing anykind of comparison between bullets is to do a comparison between surface marking of two or more bullet lands.
-Bullet Lands are considered to be areas between grooves made by the rifling action of the barrel.
-These marking are considered to be unique.
+Bullet Lands are considered to be areas between grooves made by the rifling action of the barrel and the markings on them are considered to be unique.
 The land engraved markings or sometimes termed as Bullet profiles \citep{aoas} are striation marks made on Bullet lands and often used for these land to land comparisons.
@@
 In the study conducted by \citet{aoas} a machine learning based algorithm was developed for same source matching of bullets and error rates were discussed using the database from the Hamby Study \citep{hamby}.
-In this paper, we first try to adapt a deterministic algorithm and method developed for toolmarks by \citet{chumbley} and improved by \citet{hadler}, to bullets.
-Then we consequently discuss about the efforts in doing so along with the associated error rates.
+In this paper, we first try to adapt a deterministic algorithm and method to bullets, which was developed for toolmarks by \citet{chumbley}.
+The method was later improved by \citet{hadler}.
+Then we consequently discuss about the efforts in doing so, along with the associated error rates.
 The data used in this paper also belongs to the Hamby Study \citep{hamby}.
 This gives us a common platform for comparing the performance of the chumbley method on bullets with an already existing method proposed by \citet{aoas} for bullets.
-The proposed algorithm and method of \citet{chumbley}, in their paper, compares two toolmarks with the intention of determining if it comes from the same source (same tool).
-The method also provides a means to determine error rates and claims to reduce subject bias.
+\citet{chumbley}, in their paper, compare two toolmarks with the intention of determining if it comes from the same source (same tool).
 As mentioned earlier, subject bias and error rate determination have been a long standing issue in firearm examination \citep{NAS:2009}.
-This remains one of the motivations to explore the adaptibility of the Chumbley score methodology to bullets.
+For an objective analysis, scientific principle testability and error rate determination are considered to be defining aspects.
+Identification of the uncertainities associated with a method of comparison and coming up with a systemized method of quantification, is therefore important.
+Associating mathematical probabilites in a systemized manner to these methods are one such way of quantification.
+The method proposed by \citet{chumbley} provides a means to determine error rates and claims to reduce subject bias.
+Therefore, this remains a strong motivation to explore the adaptibility of the Chumbley score methodology to bullets.
 \citet{chumbley} used an empirical based setup to validate their proposed algorithm and quantitative method which calculates a U-statisic for the purpose of classification of toolmarks as matching or non-matching.

### 2018-02-08 doc/paper.Rmd
some more changes

@@
 Extensive examples in literature for tools and toolmark research ranging from screwdrivers to groove pliers to slip-joint pliers can be found in the work of  \citet{manytoolmarks1}, \citet{manytoolmarks2}, \citet{miller}, \cite{afte-chumbley} and many more.
-In comparison to this, same source matching of bullets to firearms has not been examined as prominently as that of toolmarks, with even less information available on validity of methods and error rates associated with firearms examination.
+In comparison to this, same source matching of bullets to firearms has not been examined as prominently as that of toolmarks.
+Even less information is available on validity of methods and error rates associated with firearms examination.
 The National Academy of Sciences in its report in 2009 \citep{NAS:2009} discussed the need for determining error rates in methods proposed for firearms examination.
-In the context of same source matching and error rates determination, as seen in the case of most forensic applications, the first step involves identification of unique features that are characteristic of the object at hand.
+As seen in the case of most forensic applications, the first step for same source matching and error rate determination involves identification of unique features that are characteristic of the object at hand.
 For the case of bullets and firearms, striation marks on the surface of the bullet are considered to be such markings that can be used in methods for same source matching.
@@
 
-In the study conducted by \citet{aoas} a machine learning based algorithm was developed for same source matching of bullets and error rates were discussed using the database from the Hamby Study \citep{hamby}.
-In this paper, we first try to adapt a deterministic algorithm and method to bullets, which was developed for toolmarks by \citet{chumbley}.
+As mentioned earlier, subject bias and error rate determination have been a long standing issue in firearm examination \citep{NAS:2009}.
+For an objective analysis, scientific principle testability and error rate determination are considered to be defining aspects.
+Identification of the uncertainities associated with a method of comparison and coming up with a systemized method of quantification, is therefore important.
+Associating mathematical probabilites in a systematic manner to these methods are one such way of quantification.
+In a previous study conducted by \citet{aoas} a machine learning based algorithm was developed for same source matching of bullets and error rates were discussed using the database from the Hamby Study \citep{hamby}.
+The method proposed by \citet{chumbley} also provides a means to determine error rates and claims to reduce subject bias.
+Therefore, giving us a strong motivation to explore the adaptibility of the Chumbley score methodology to bullets.
+In this paper, we first try to adapt the deterministic algorithm and method developed for toolmarks by \citet{chumbley} to bullets.
 The method was later improved by \citet{hadler}.
-Then we consequently discuss about the efforts in doing so, along with the associated error rates.
+After this, we consequently discuss about the efforts in doing so, along with the associated error rates.
 The data used in this paper also belongs to the Hamby Study \citep{hamby}.
@@
 \citet{chumbley}, in their paper, compare two toolmarks with the intention of determining if it comes from the same source (same tool).
-As mentioned earlier, subject bias and error rate determination have been a long standing issue in firearm examination \citep{NAS:2009}.
-For an objective analysis, scientific principle testability and error rate determination are considered to be defining aspects.
-Identification of the uncertainities associated with a method of comparison and coming up with a systemized method of quantification, is therefore important.
-Associating mathematical probabilites in a systemized manner to these methods are one such way of quantification.
-The method proposed by \citet{chumbley} provides a means to determine error rates and claims to reduce subject bias.
-Therefore, this remains a strong motivation to explore the adaptibility of the Chumbley score methodology to bullets.
 \citet{chumbley} used an empirical based setup to validate their proposed algorithm and quantitative method which calculates a U-statisic for the purpose of classification of toolmarks as matching or non-matching.

### 2018-02-08 doc/paper.Rmd
additions to intro

@@
 Therefore, giving us a strong motivation to explore the adaptibility of the Chumbley score methodology to bullets.
-In this paper, we first try to adapt the deterministic algorithm and method developed for toolmarks by \citet{chumbley} to bullets.
+In this paper, we try to adapt this deterministic algorithm and method developed for toolmarks, to bullets.
 The method was later improved by \citet{hadler}.
@@
 The data used in this paper also belongs to the Hamby Study \citep{hamby}.
-This gives us a common platform for comparing the performance of the chumbley method on bullets with an already existing method proposed by \citet{aoas} for bullets.
-\citet{chumbley}, in their paper, compare two toolmarks with the intention of determining if it comes from the same source (same tool).
+This gives us a common platform for comparing the performance of the chumbley method on bullets with the already existing method proposed by \citet{aoas} for bullets.
+As a brief overview, \citet{chumbley}, in their paper, compare two toolmarks with the intention of determining if it comes from the same source (same tool).
 \citet{chumbley} used an empirical based setup to validate their proposed algorithm and quantitative method which calculates a U-statisic for the purpose of classification of toolmarks as matching or non-matching.

### 2018-02-08 doc/paper.Rmd
rewording .
rewording

@@
 Compairing pairs of toolmarks with the intention of matching it to a tool has been studied many times in the past.
-Extensive examples in literature for tools and toolmark research ranging from screwdrivers to groove pliers to slip-joint pliers can be found in the work of  \citet{manytoolmarks1}, \citet{manytoolmarks2}, \citet{miller}, \cite{afte-chumbley} and many more.
+Extensive examples can be found in literature for tools and toolmark research ranging from screwdrivers \citep{manytoolmarks1, manytoolmarks2, chumbley} to groove pliers \citep{manytoolmarks2} to slip-joint pliers \citep{afte-chumbley} and many more.
 In comparison to this, same source matching of bullets to firearms has not been examined as prominently as that of toolmarks.
@@
 As mentioned earlier, subject bias and error rate determination have been a long standing issue in firearm examination \citep{NAS:2009}.
+This issue of removing subjectivity and making firearm analysis objective  was also raised in the \citet{pcast2016} report.
 For an objective analysis, scientific principle testability and error rate determination are considered to be defining aspects.
@@
 In a previous study conducted by \citet{aoas} a machine learning based algorithm was developed for same source matching of bullets and error rates were discussed using the database from the Hamby Study \citep{hamby}.
-The method proposed by \citet{chumbley} also provides a means to determine error rates and claims to reduce subject bias.
+The method proposed by \citet{chumbley} (and later improved by \citet{hadler}) also provides a means to determine error rates and claims to reduce subject bias.
 Therefore, giving us a strong motivation to explore the adaptibility of the Chumbley score methodology to bullets.
 In this paper, we try to adapt this deterministic algorithm and method developed for toolmarks, to bullets.
-The method was later improved by \citet{hadler}.
 After this, we consequently discuss about the efforts in doing so, along with the associated error rates.
@@
 This gives us a common platform for comparing the performance of the chumbley method on bullets with the already existing method proposed by \citet{aoas} for bullets.
-As a brief overview, \citet{chumbley}, in their paper, compare two toolmarks with the intention of determining if it comes from the same source (same tool).
-\citet{chumbley} used an empirical based setup to validate their proposed algorithm and quantitative method which calculates a U-statisic for the purpose of classification of toolmarks as matching or non-matching.
+As a brief overview, \citet{chumbley}, in their paper, compare two toolmarks with the intention of determining if it comes from the same tool.
+They use an empirical based setup to validate their proposed algorithm and quantitative method which calculates a U-statisic for the purpose of classification of toolmarks as matching or non-matching.
 The data for their study was obtained from 50 sequentially manufactured screwdriver tips, and preselected comparison window sizes were given as inputs to the algorithm.
-The algorithm then compares the two toolmarks and comes up with a U-statistic and an associated p-value to designate them as matches or non-matches.
+The algorithm then compares the two toolmarks and comes up with a Mann-Whitney U-statistic and an associated p-value to designate them as matches or non-matches.
 The performance for every 100 comparisons, of the algorithm proposed by \citet{chumbley} and the improvement proposed by \citet{hadler} are listed in the table below.

### 2018-02-11 doc/paper.Rmd
minor changes to limitations of chumbley score

@@
 Therefore, giving us a strong motivation to explore the adaptibility of the Chumbley score methodology to bullets.
-In this paper, we try to adapt this deterministic algorithm and method developed for toolmarks, to bullets.
+In this paper, we take this toolmark based methodology and try to apply and adapt it to bullets.
 After this, we consequently discuss about the efforts in doing so, along with the associated error rates.
@@
 ## Potential limitations of the Chumbley Score adaptation to bullets
-Bullets are much smaller in length, width, are not flat and curved in the cross-sectional topography as opposed to tools like screw driver tips which produced longer and pronounced markings.
-This means the makings made by barrels in comparison to toolmarks may have a problem in distinctiveness.
-The majority of Bullet profiles and signatures extracted by procedures mentioned by  \citet{aoas} are almost 1/4 th the size of toolmarks as used by \citet{chumbley} or even smaller.
-Striations on Bullets are made on their curved surfaces, whereas the algorithm developed by \citet{chumbley} and \citet{hadler} has only been tested for flatter and wider surfaces which have negligible curvature.
+Bullet striations are different from toolmarks.
+Bullet markings are typically produced by the barrel on the bullets, while toolmarks can be made by tool tips on different surfaces.
+Tools like screw driver tips as used by \citet{chumbley}, produce longer and pronounced markings.
+Bullets on the other hand are much smaller in length, width, and curved in the cross-sectional topography.
+This means the markings made by barrels as opposed to markings made by tools, may have a problem in distinctiveness.
+The majority of Bullet profiles and signatures of the Hamby study \citep{hamby} used in this paper are extracted by procedures mentioned by  \citet{aoas}.
+The Ruger pistol barrels used in the  Hamby study have a 9mm caliber.
+This means that bullets used in these pistols have land markings of almost $(1/6)^{th}$ the size of screwdriver toolmarks used by \citet{chumbley}.
+Also, striations on bullets are made on their curved surfaces, whereas the algorithm developed by \citet{chumbley} and \citet{hadler} has only been tested for flatter and wider surfaces which have negligible curvature.
 Therefore, using methods proposed for toolmarks may need adaptation in order to give tangible results for bullets.
-Moreover, in order to to get flat bullet signatures and remove the curvatures some kind of smoothing needs to be applied as a pre-step which needs further investigation as to whether the level of smoothing does effect the working of the algorithm on Bullets.
+Moreover, in order to get flat bullet signatures and remove the curvatures some kind of smoothing needs to be applied as a pre-step wihch needs further investigation as to whether the level of smoothing does effect the working of the algorithm on Bullets.
 
-Also when in the optimization step, the Window of optimization for bullets will be shorter as the signatures are smaller.
-The idea is to keep the number of windows of optimization sufficiently large, which means shorter Trace segments (partition of signature or toolmark with length = size of window of optimization) that lets us compare smaller segments of one signature to another.
+The Chumbley method and algorithm works on the premise of comparison windows which are supposed to represent small segments of pixels of equal sizes in the two markings that are to be compared in a predetermined fashion.
+They term them as windows of optimization and validation.
+The optimization window is selected in the first step called the optimization step where areas/windows of best agreement between the the two markings are found.
+The window of optimization for bullet striations are shorter as bullet signatures are smaller compared to toolmarks .
+The idea is to keep the number of windows of optimization sufficiently large.
+This means we have shorter trace segments that lets us compare smaller segments of one signature to another than those for toolmarks.
+Here, trace segments are the partitions of marking with the length of the partition equal to the size of window of optimization.
 This introduces a problem as, if we go too small in the window of optimization, the unique features of the trace segments are lost and seem similar, while too large sizes vastly reduces the weight of small features that would otherwise uniquely classify a signature and hence identify the region of agreement.
+The problem of distinctiveness in the structure of bullet striations may also lead to potential failure of the algorithm to identify correct windows of agreements or not be able to identify these windows at all.
+This may lead to incorrect classification of bullets as matches or non-matches.
 
-Thus the Window of Optimization has a direct influence on whether we are Falsely rejecting the null when it is true (Type I) or Falsely accepting the null when it is false (Type II) making identification of the optimum size of window of optimization very important.
-This raises important questions about what parameter settings need to be chosen for bullet land comparison.
+Thus, the comaparison windows have a direct influence on the false positives and false negatives, making identification of the optimum sizes of these windows very important.
+This raises important questions about how and what parameter settings need to be chosen for bullet land comparison.
 Something similar is done in \citet{afte-chumbley} for toolmark comparisons of slip-joint pliers where optimum window sizes are determined.
-There is a need to figure out the best parameter settings which minimizes the errors for unsmoothed markings and pre-processed signatures as determined by \citet{aoas}.
+There is a need to figure out the best parameter settings which minimizes the errors for unsmoothed markings or profiles and pre-processed signatures.
 An analysis of these error rates and comparison with other methods will help us understand the adaptibility of the chumbley score to bullets.
 
-\hh{figure \ref{fig:rgl} needs to be mentioned somewhere in the write-up}
+Figure \ref{fig:rgl} shows us how the extracted signature from a bullet land lines up with the image of the land from which it was extracted.
+We can see also see in the figure how the depth and relative position of the striation markings seen in the image are interpreted as the signature.
 
@@
 The $t_1$ and $t_2$ can be better understood as equally spaced pixel locations for the two marks under consideration, where a 'pixel' is the resolution of the confocal light microscope.
-In the case for bullet signatures a pixel corresponds to 0.645 microns.
+In the case for bullet signatures extracted from NIST scans of the Hamby study, a pixel corresponds to approximately 1.56 microns.
 These marks or striae are potentially from two different bullets or two different toolmarks whose source needs to be identified as being same or different.
 The marks or striae are indexed by the pixel location where $t_1$ is for the first striae referred to as $x$ and $t_2$ is for the second striae which is referred as $y$.
-$T_1$ and $T_2$ being the respective lengths of the markings, need not be the same but are usually of similar lenghtsThe similarity is then judged by the algorithm on the basis of cross-correlation of a fixed and constant number consecutive pixels say $k$ taken from the two indexed marks $x(t_1)$ and $y(t_2)$ such that in theory $k$ remains smaller that length of the two striae or marks which is the same as $t_1$ and $t_2$.
+$T_1$ and $T_2$ being the respective lengths of the markings, need not be the same but are usually of similar lenghts.
+The similarity is then judged by the algorithm on the basis of cross-correlation of a fixed and constant number consecutive pixels say $k$ taken from the two indexed marks $x(t_1)$ and $y(t_2)$ such that in theory $k$ remains smaller that length of the two striae or marks which is the same as $t_1$ and $t_2$.
 Depending on the what stage of the algorithm we are in, matching of different pixel lengths and locations is done, which at the end effectively compares all possible windows that would guarantee in quantifying the two marks or striae as coming from the same source or not, which also lets us assess the error rates by checking for a large number of cases.
@@
 The algorithm works in two phases, namely, an optimization step and a validation step, at the end of which a Mann Whitney U statistic is calculated.
-A pre-processing step to the algorithm is to choose a coarseness value which  is used as a parameter to the LOWESS smoothing function.
+A pre-processing step to the algorithm is to choose a coarseness value which  is used as a parameter to the lowess smoothing function.
 The coarseness essentially gives the proportion of points which influence the smooth at each value, which means larger values lead to more smoothness.
-The LOWESS smoothing is applied to each of two sets of vectorized striae or marks $x(t_1)$ and $y(t_2)$, before proceeding to the algorithm.
+The lowess smoothing is applied to each of two sets of vectorized striae or marks $x(t_1)$ and $y(t_2)$, before proceeding to the algorithm.
 
@@
 
-Since we are only interested in a non-parameteric U statistic, Hadler et al.
-proposed a normalization procedure in the Validation step that goes to some extent to address the issue of mutual dependence.
-Also the same shift and different shift substep was modified to use a deterministic rule for sampling sam-shhift and dofferent shift-samples as opposed to the originally proposed random samples.
+Since we are only interested in a non-parameteric U statistic, \citet{hadler} proposed a normalization procedure in the Validation step that goes to some extent to address the issue of mutual dependence.
+Also the same shift and different shift substep was modified by \citet{hadler} to use a deterministic rule for sampling same-shift and different shift-samples as opposed to the originally proposed random samples.
 
-The data that is used by \citet{chumbley} is generated by a surface profilometer that gives the height in terms of distance along a linear trace.
+<!-- The data that is used by \citet{chumbley} is generated by a surface profilometer that gives the height in terms of distance along a linear trace.
 This is taken perpendicular to the striations present in the toolmark.
 Two such trace are then compared using the algorithm.
+-->
 

### 2018-02-16 doc/paper.Rmd
section chumbley score test reworded reorganized

@@
 
-# The Chumbley Score Test
+## The Chumbley Score Test
 
-The Chumbley score algorithm takes input as two vectorized processes $x(t_1)$, $t_1 = 1,2,...T_1$ and $y(t_2)$, $t_2 = 1,2...T_2$ which denote two sets of marks or striae.
-Here the processes are of the form $z(s)$ which is a spatial process for some $t$, while $z(s_1)$ is the realization of the process in $s_1$.
-The $t_1$ and $t_2$ can be better understood as equally spaced pixel locations for the two marks under consideration, where a 'pixel' is the resolution of the confocal light microscope.
+The Chumbley score algorithm takes input as two vectorized processes.
+Here the processes are of the form $z(t)$ which is a spatial process for some location indexed with $t$.
+This means that while $z(t)$ represents any striae, $z_1(t_1)$ is a realization of the spatial process.
+$t$ on the other hand is a vector of equally spaced pixel locations for the striation marks under consideration.
+The word 'pixel' refers to the resolution of the confocal light microscope.
 In the case for bullet signatures extracted from NIST scans of the Hamby study, a pixel corresponds to approximately 1.56 microns.
-These marks or striae are potentially from two different bullets or two different toolmarks whose source needs to be identified as being same or different.
-The marks or striae are indexed by the pixel location where $t_1$ is for the first striae referred to as $x$ and $t_2$ is for the second striae which is referred as $y$.
-$T_1$ and $T_2$ being the respective lengths of the markings, need not be the same but are usually of similar lenghts.
-The similarity is then judged by the algorithm on the basis of cross-correlation of a fixed and constant number consecutive pixels say $k$ taken from the two indexed marks $x(t_1)$ and $y(t_2)$ such that in theory $k$ remains smaller that length of the two striae or marks which is the same as $t_1$ and $t_2$.
-Depending on the what stage of the algorithm we are in, matching of different pixel lengths and locations is done, which at the end effectively compares all possible windows that would guarantee in quantifying the two marks or striae as coming from the same source or not, which also lets us assess the error rates by checking for a large number of cases.
+The vectorized processes representing two sets of striaes are therefore shown as $x(t_1)$, $t_1 = 1,2,...T_1$ and $y(t_2)$, $t_2 = 1,2...T_2$.
+Here $x$ and $y$ denote two realizations of the process $z(*)$ which means two striation marks, while the indexing $t$ or pixel range is shown as $t_1$ and $t_2$ for $x$ and $y$ respectively.
+This representation of $t$ as $t_1$ and $t_2$ lets us identify them as either of same or different lengths.
+The striation marks under consideration are potentially from two different bullets whose source needs to be identified as being same or different.
+$T_1$ and $T_2$, as represented above, are the final pixel indexes of each marking and therefore give the respective lengths of the markings.
+The similarity in the two markings is then judged by the algorithm on the basis of cross-correlation of a fixed and constant number consecutive pixels (say $k$) taken from the two markings.
+This can for example be $k$ taken from the two indexed marking $x(t_1)$ and $y(t_2)$ such that in theory $k$ remains smaller than length of the two striae or marks.
+Depending on what stage of the algorithm we are in, matching of different pixel lengths and locations is done, which at the end effectively compares all possible windows that would guarantee in quantifying the two marks or striae as coming from the same source or not, which also lets us assess the error rates by checking for a large number of cases.
 
-The algorithm works in two phases, namely, an optimization step and a validation step, at the end of which a Mann Whitney U statistic is calculated.
+The algorithm works in two phases, namely, an optimization step and a validation step, at the end of which a Mann Whitney U statistic is calculated as a measure to differentiate between matches and non-matches.
 A pre-processing step to the algorithm is to choose a coarseness value which  is used as a parameter to the lowess smoothing function.
@@
 The lowess smoothing is applied to each of two sets of vectorized striae or marks $x(t_1)$ and $y(t_2)$, before proceeding to the algorithm.
+The algorithm starts with the optimization step where the area of best agreement in the two markings being compared is identified.
+Comparison window size is predefined.
+Each window of the first marking is compared with all windows of the second marking.
+A maximum correlation statistic is used to identify the region of best agreement, with the maximum usually seen being near 1 for both cases which is what is intuitively expected for matches, and not expected intuitively for non matches.
+\citet{hadler} in their paper proposed an improvement to this algorithm by trying to remove mutual dependence of parameters.
+The mututal dependence was due to the serial correlation in surface depth values of a marking.
+Also a random sampling sub step in the validation phase makes a group of pixels to be chosen more than once and hence introduces lack of independence in certain steps.
+The reason for removing the mutual dependence was the Mann-Whitney U statistic works under the assumption of independence of parameters.
+The validation step builds on this with two sub-steps namely, same shift and different shift.
+As the aim is to come up with a non-parameteric U statistic, \citet{hadler} proposed a normalization procedure in the validation step which goes to some extent to address the issue of mutual dependence.
+A series of windows are chosen in both the same and different shift sub-steps for the purpose of comparison.
+Both substeps were modified by \citet{hadler} who introduced a deterministic rule for sampling of windows in the same-shift and different shift as opposed to the originally proposed random sampling by \citet{chumbley}.
+In th same shift substep the chosen series of windows are at a common distance (rigid-shifts) from the window earlier identified as the region of best agreement in the optimization step.
+The correlation of these windows generally turn out to be lower than the maximum correlation window.
+The significance is that these same shift windows still have large enough correlation values for the two markings being compared that are in reality a match.
+Any choice of maximum correlation windows (where correlation values are often near 1) for non-matches in the optimization step are also validated by this sub-step.
+This is because the same-shift correlations would not be anywhere as large for same-shift windows when all windows in the sub-step are compared.
+The different shift substep on the other hand gives perspective to the correlation values of the same shift window correlation values.
+There are no rigid-shifts but different shifts where distance from the maximum correlation window are chosen randomly by \citet{chumbley} and deterministically as per \citet{hadler}.
+This means that there is an equal possibilty of comparing a trace segment from one marking to any trace segment or window in the second marking.
 
-\citet{hadler} in their paper proposed an improvement to this algorithm by trying to remove mutual dependence of parameters (due to serial correlation in surface depth values of a toolmark and because of a random sampling sub step in the validation phase which makes a group of pixels to be chosen more than once and hence introduces lack of independence) in certain steps, especially because the Mann-Whitney U statistic that is later calculated in the algorithm and used as a measure to differentiate between matches and non-matches works under the assumption of independence of parameters.
-
-Since we are only interested in a non-parameteric U statistic, \citet{hadler} proposed a normalization procedure in the Validation step that goes to some extent to address the issue of mutual dependence.
-Also the same shift and different shift substep was modified by \citet{hadler} to use a deterministic rule for sampling same-shift and different shift-samples as opposed to the originally proposed random samples.
+Neither of above sets of correlation in the two sub-steps are allowed to include the maximum correlation window as identified earlier.
+Therefore the assumption is that if two markings match each other, the same-shift correlations would be larger than the different-shift windows.
+And if they are not a match the correlations in the two sets will be very similar.
+The U-statistic tests for the null hypothesis that the two markings are not a match and are therefore not made by the same source.
+As given by \citet{hadler}, it is computed from the joint rank of all correlations of both the same and different shift samples.
 
@@
 
-## Detailed algorithm
+<!-- ## Detailed algorithm -->
 
-###Optimization step
-The idea behind this step is to first identify the area of best agreement in the two toolmark data.
+<!-- ###Optimization step -->
+<!-- The idea behind this step is to first identify the area of best agreement in the two toolmark data.
 Comparison window size is predefined by the user, which in case of screwdriver toolmarks was chosen by Chumbley et al and Hadler et al as around 10 percent of the length of the toolmarks, which was around 500.
 This window is henceforth referred to as Window of optimization.
+-->
 
-A maximum correlation statistic is used to identify the region of best agreement, with the maximum usually seen being near 1 for both cases which is what we intuitively expect for matches, and something which we do not intuitively expect for non matches
+<!-- A maximum correlation statistic is used to identify the region of best agreement, with the maximum usually seen being near 1 for both cases which is what we intuitively expect for matches, and something which we do not intuitively expect for non matches  -->
 
-First, there are a very large number of cross-correlations calculated for the two series of striae or marks $x(t_1)$ and $y(t_2)$, for eg if the window of optimization was defined as 200 and toolmark pixel length is around 1000 then we have 1200-(200-1) = 1001.
+<!-- First, there are a very large number of cross-correlations calculated for the two series of striae or marks $x(t_1)$ and $y(t_2)$, for eg if the window of optimization was defined as 200 and toolmark pixel length is around 1000 then we have 1200-(200-1) = 1001.
+-->
 
-This is the number of windows we have for one toolmark and each window is compared with all windows of the second toolmark (the number of windows is again similar), and the window with the maximum correlation is identified to be the region where the toolmarks are in maximum agreement with each other
+<!-- This is the number of windows we have for one toolmark and each window is compared with all windows of the second toolmark (the number of windows is again similar), and the window with the maximum correlation is identified to be the region where the toolmarks are in maximum agreement with each other -->
 
-###Validation Step
+<!-- ###Validation Step -->
 
-####Same-shift: 
-In this step a series of windows are chosen at random (originally by chumbley et al) and deterministically (by hadler et al), but at a common distance (rigid-shift) from the window identified as the region of best agreement in the vaildation step.
-The correlation of these windows would be as intuitively assumed i.e.
+<!-- ####Same-shift:  -->
+<!-- In this step a series of windows are chosen at random (originally by \citet{chumbley}) and deterministically (by \citet{hadler}), but at a common distance (rigid-shift) from the window identified as the region of best agreement in the vaildation step.
+-->
+<!-- The correlation of these windows would be as intuitively assumed i.e.
 lower than the maximum correlation window (Optimization step), but the significance is that these same shift windows will still have large enough correlation values for two toolmarks or signatures that are in reality a match.
+-->
 
-This also validates that if in the optimization step the maximum correlation window chosen (with correlation value near 1) was by accident (like in case of signatures that in reality are a non-match), all these same- shift correlations (A fixed number of these trace segments are identified) would not be anywhere large enough for all same-shift windows.
+<!-- This also validates that if in the optimization step the maximum correlation window chosen (with correlation value near 1) was by accident (like in case of signatures that in reality are a non-match), all these same- shift correlations (A fixed number of these trace segments are identified) would not be anywhere large enough for all same-shift windows.
+-->
 
@@
 
-####Different Shift:
-Primary reason for this substep is to give perspective to the correlation values of the same shift window correlation values.
+<!-- ####Different Shift: -->
+<!-- Primary reason for this substep is to give perspective to the correlation values of the same shift window correlation values. -->
 
-This time there are no rigid-shifts but different shifts (distance from Window of Opt with max correlation) chosen randomly by \citet{chumbley} and deterministically by Hadler et al.
+<!-- This time there are no rigid-shifts but different shifts (distance from Window of Opt with max correlation) chosen randomly by \citet{chumbley} and deterministically by Hadler et al.
 such that there is an equal possibilty of comparing a trace segment from one signature or toolmark to any one in the second signature or toolamrk.
+-->
 
-Neither of above sets of correlation are allowed to include the maximum correlation window as identified earlier.
+<!-- Neither of above sets of correlation are allowed to include the maximum correlation window as identified earlier. -->
 
-Therefore the assumption is that if two toolmarks or signatures match each other the same-shift correlations would be larger than the different-shift windows,
-and if they are not a match the correlations in the two sets will be very similar.
+<!-- Therefore the assumption is that if two toolmarks or signatures match each other the same-shift correlations would be larger than the different-shift windows, -->
+<!-- and if they are not a match the correlations in the two sets will be very similar. -->
 
-###U Statistic:
+<!-- ###U Statistic: -->
 
-This is computed from the joint rank of all correlations of both the same and different shift samples. As given by \citet{hadler}
+<!-- This is computed from the joint rank of all correlations of both the same and different shift samples. As given by \citet{hadler} -->
 
-Null Hypothesis: If the toolmarks were not match i.e not made by the same tool.
+<!-- Null Hypothesis: If the toolmarks were not match i.e not made by the same tool. -->
 
-Let ns and nd be the number of same shift and different shift windows
-$$N = n_{s} + n_{d}$$
+<!-- Let ns and nd be the number of same shift and different shift windows -->
+<!-- $$N = n_{s} + n_{d}$$ -->
 
-The mann whitney U statistic is given by
-$$U =\sum^{ns}_{i=1}R_{s}\left( i\right)$$
+<!-- The mann whitney U statistic is given by -->
+<!-- $$U =\sum^{ns}_{i=1}R_{s}\left( i\right)$$ -->
 
-with the standardized version which includes provision for rank ties
+<!-- with the standardized version which includes provision for rank ties -->
 
-$$\overline{U}= \dfrac{U-M}{\sqrt{V}}$$
+<!-- $$\overline{U}= \dfrac{U-M}{\sqrt{V}}$$ -->
 
-where prior to normalization the U-statistic has the mean as
+<!-- where prior to normalization the U-statistic has the mean as -->
 
-$$M = n_{s}\left(\dfrac{N+1}{2}\right)$$
+<!-- $$M = n_{s}\left(\dfrac{N+1}{2}\right)$$ -->
 
-and variance
+<!-- and variance  -->
 
-$$V = \dfrac{n_{s}n_{d}}{N\left(N-1\right)}\left[\Sigma^{n_{s}}R_{s}\left(i\right)^{2}+\Sigma^{n_{d}}R_{d}\left(j\right)^{2}\right] -\dfrac{n_{s} n_{d}\left(N+1\right)^{2}}{4\left(N-1\right)}$$
+<!-- $$V = \dfrac{n_{s}n_{d}}{N\left(N-1\right)}\left[\Sigma^{n_{s}}R_{s}\left(i\right)^{2}+\Sigma^{n_{d}}R_{d}\left(j\right)^{2}\right] -\dfrac{n_{s} n_{d}\left(N+1\right)^{2}}{4\left(N-1\right)}$$ -->
 

### 2018-02-16 doc/paper.Rmd
minor changes cited nist database
minor change

@@
 The word 'pixel' refers to the resolution of the confocal light microscope.
-In the case for bullet signatures extracted from NIST scans of the Hamby study, a pixel corresponds to approximately 1.56 microns.
+In the case for bullet signatures extracted from NIST ballistics database \citep{nist} of the Hamby study, a pixel corresponds to approximately 1.56 microns.
 The vectorized processes representing two sets of striaes are therefore shown as $x(t_1)$, $t_1 = 1,2,...T_1$ and $y(t_2)$, $t_2 = 1,2...T_2$.
-Here $x$ and $y$ denote two realizations of the process $z(*)$ which means two striation marks, while the indexing $t$ or pixel range is shown as $t_1$ and $t_2$ for $x$ and $y$ respectively.
+Here $x$ and $y$ denote two instances of the process $z(*)$ which means two striation marks, while the indexing $t$ or pixel range is shown as $t_1$ and $t_2$ for $x$ and $y$ respectively.
 This representation of $t$ as $t_1$ and $t_2$ lets us identify them as either of same or different lengths.

### 2018-02-16 doc/paper.Rmd
minor chng

@@
 Here the processes are of the form $z(t)$ which is a spatial process for some location indexed with $t$.
-This means that while $z(t)$ represents any striae, $z_1(t_1)$ is a realization of the spatial process.
-$t$ on the other hand is a vector of equally spaced pixel locations for the striation marks under consideration.
+This means that while $z(t)$ represents any striae, $z(t_1)$ is a realization of the spatial process.
+$t$ here denotes a vector of equally spaced pixel locations for the striation marks under consideration.
 The word 'pixel' refers to the resolution of the confocal light microscope.

### 2018-02-16 doc/paper.Rmd
figure info place rearranged

@@
 
+Figure \ref{fig:rgl} shows us how the extracted signature from a bullet land lines up with the image of the land from which it was extracted.
+We can see also see in the figure how the depth and relative position of the striation markings seen in the image are interpreted as the signature.
+
 The Chumbley method and algorithm works on the premise of comparison windows which are supposed to represent small segments of pixels of equal sizes in the two markings that are to be compared in a predetermined fashion.
@@
 An analysis of these error rates and comparison with other methods will help us understand the adaptibility of the chumbley score to bullets.
-
-Figure \ref{fig:rgl} shows us how the extracted signature from a bullet land lines up with the image of the land from which it was extracted.
-We can see also see in the figure how the depth and relative position of the striation markings seen in the image are interpreted as the signature.
 
@@
 This can for example be $k$ taken from the two indexed marking $x(t_1)$ and $y(t_2)$ such that in theory $k$ remains smaller than length of the two striae or marks.
-Depending on what stage of the algorithm we are in, matching of different pixel lengths and locations is done, which at the end effectively compares all possible windows that would guarantee in quantifying the two marks or striae as coming from the same source or not, which also lets us assess the error rates by checking for a large number of cases.
+Depending on what stage of the algorithm we are in, matching of different pixel lengths and locations is done.
+This in the end, effectively compares all possible windows that would guarantee in quantifying the two marks or striae as coming from the same source or not.
+It also lets us assess the error rates by checking for a large number of cases.
 

### 2018-02-19 doc/paper.Rmd
rename section.
figure 1 info
removed a sentence in intro
meeting monday morning

@@
 ## Motivation
+\hh{XXX first sentence: Same source analyses are a major part of an Forensic Toolmark examiners job.
+XXX second sentence: describe standard practice: In current practice firearm examiners  make these comparisons visually under a comparison microscope and report XXX what is being reported: AFTE.
+XXX find: AFTE theory of identification as it relates to toolmarks  }
+\hh{One of the main criticisms in this toolmark examination first raised by the National Research Council in 2009 [@NAS:2009] and later emphasized further by the President's Council of Advisors on Science and Technology [@pcast2016] are the amount of subjectivity in the assessment and the lack of error rates.
+}
+\hh{XXX we have digitized images, e.g. profilometers or Confocal microscopy allow us to measure the surface [@vorburger2016]}
+
+\hh{Using a statistical approach removes both subjectivity from the assessment and allows us to provide error rates for both false positive and false negative identifications. }
+\hh{Multiple toolmarks have been studied in the literature:  \citep{manytoolmarks1, manytoolmarks2, chumbley} have studied screwdriver marks digitized using a profilometer,  \citep{manytoolmarks2} investigate groove pliers XXX data?
+XXX  to slip-joint pliers \citep{afte-chumbley} XXX.
+In all of these cases, the statistical test used, is a variation of the Chumbley scoring method, first introduced by XXX and later improved by XXX.
+}
+
+\hh{XXX In this paper, we are investigating the applicability of the Chumbley scoring method by @hadler to assessing striation marks on bullets for same-source identification. }
+
+\hh{Striation marks on bullets are made by impurities in the barrel.
+As the bullet travels through the barrel, these imperfections leave "scratches" on the bullet surface.
+Typically, only striation marks in the land engraved areas (LEAs) are considered (XXX citation) [@afte-article1992].
+XXX explain lands and grooves XXX.
+Compared to toolmarks made by screwdrivers striation marks on bullets are typically much smaller, both in length and in width.
+Bullets also have a curved cross-sectional topography.
+}
+
+\hh{XXX How are bullets matched now? \citet{chu2013}: CMS, automatically - error rate?, using ccf XXX Vorburger paper XXX  and \citet{aoas}: multiple features, error rates }
+
+\hh{The Chumbley score  provides us with another approach in the same-source assessment of bullet striation marks. }
+
+\hh{end of intro: remaining paper is structured as follows: introduce to the data we get from confocal microscopy, introduce to profiles and signatures.
+Introduce to chumbley score method, apply chumbley and discuss results ....
+}
+
+## Scans for land engraved areas
+
+- scans available: NIST database (citation), Hamby 44 and Hamby 252 (Hamby citation)
+- move figure up
+- discuss cross section, profiles and signatures
+
 Compairing pairs of toolmarks with the intention of matching it to a tool has been studied many times in the past.
@@
 A more detailed version of the extraction technique of signatures is discussed by \citet{aoas}, where comprehensive details about the height at which profile is to be selected, removing curvature, smoothing, identifying groove locations are explained.
+Figure \ref{fig:rgl} shows us how the extracted signature from a bullet land lines up with the image of the land from which it was extracted.
+We can see also see in the figure how the depth and relative position of the striation markings seen in the image are interpreted as the signature.
 
@@
 
-Figure \ref{fig:rgl} shows us how the extracted signature from a bullet land lines up with the image of the land from which it was extracted.
-We can see also see in the figure how the depth and relative position of the striation markings seen in the image are interpreted as the signature.
-
 The Chumbley method and algorithm works on the premise of comparison windows which are supposed to represent small segments of pixels of equal sizes in the two markings that are to be compared in a predetermined fashion.
@@
 This in the end, effectively compares all possible windows that would guarantee in quantifying the two marks or striae as coming from the same source or not.
-It also lets us assess the error rates by checking for a large number of cases.
 
@@
 
-## Testing setup
+# Simulation setup
 

### 2018-02-19 doc/paper.tex
pdf recompiled
meeting monday morning

@@
 
+\hh{XXX first sentence: Same source analyses are a major part of an Forensic Toolmark examiners job.
+XXX second sentence: describe standard practice: In current practice firearm examiners  make these comparisons visually under a comparison microscope and report XXX what is being reported: AFTE.
+XXX find: AFTE theory of identification as it relates to toolmarks  }
+\hh{One of the main criticisms in this toolmark examination first raised by the National Research Council in 2009 [@NAS:2009] and later emphasized further by the President's Council of Advisors on Science and Technology [@pcast2016] are the amount of subjectivity in the assessment and the lack of error rates.
+}
+\hh{XXX we have digitized images, e.g. profilometers or Confocal microscopy allow us to measure the surface}
+
+\hh{Using a statistical approach removes both subjectivity from the assessment and allows us to provide error rates for both false positive and false negative identifications. }
+\hh{Multiple toolmarks have been studied in the literature:  \citep{manytoolmarks1, manytoolmarks2, chumbley} have studied screwdriver marks digitized using a profilometer,  \citep{manytoolmarks2} investigate groove pliers XXX data?
+XXX  to slip-joint pliers \citep{afte-chumbley} XXX.
+In all of these cases, the statistical test used, is a variation of the Chumbley scoring method, first introduced by XXX and later improved by XXX.
+}
+
+\hh{XXX In this pper, we are investigating the applicability of the Chumbley scoring method by @hadler to assessing striation marks on bullets for same-source identification. }
+
+\hh{Striation marks on bullets are made by impurities in the barrel.
+As the bullet travels through the barrel, these imperfections leave "scratches' on the bullet surface.
+Typically, only striation marks in the land engraved areas (LEAs) are considered (XXX citation) [@afte-article1992].
+Compared to toolmarks made by screwdrivers striation marks on bullets are typically much smaller, both in length and in width.
+Bullets also have a curved cross-sectional topography.
+}
+
 Compairing pairs of toolmarks with the intention of matching it to a
-tool has been studied many times in the past. Extensive examples in
-literature for tools and toolmark research ranging from screwdrivers to
-groove pliers to slip-joint pliers can be found in the work of
-\citet{manytoolmarks1}, \citet{manytoolmarks2}, \citet{miller},
-\cite{afte-chumbley} and many more. In comparison to this, same source
-matching of bullets to firearms has not been examined as prominently as
-that of toolmarks, with even less information available on validity of
-methods and error rates associated with firearms examination. The
-National Academy of Sciences in its report in 2009 \citep{NAS:2009}
-discussed the need for determining error rates in methods proposed for
-firearms examination. In the context of same source matching and error
-rates determination, as seen in the case of most forensic applications,
-the first step involves identification of unique features that are
-characteristic of the object at hand. For the case of bullets and
-firearms, striation marks on the surface of the bullet are considered to
-be such markings that can be used in methods for same source matching.
-These marks are often a product of rifling and impurities and defects
-due to manufacturing in the barrel of the gun, which leads to engravings
-on the bullet surface \citep{afte-article1992}. In current practice,
-firearm examiners invariably make visual comparisons of bullet striae
-and use visual assesment tools to dignify bullets as being matches and
-non-matches. One way of accomplishing anykind of comparison between
-bullets is to do a comparison between surface marking of two or more
-bullet lands. Bullet Lands are considered to be areas between grooves
-made by the rifling action of the barrel. These marking are considered
-to be unique. The land engraved markings or sometimes termed as Bullet
+tool has been studied many times in the past. Extensive examples can be
+found in literature for tools and toolmark research ranging from
+screwdrivers \citep{manytoolmarks1, manytoolmarks2, chumbley} to groove
+pliers \citep{manytoolmarks2} to slip-joint pliers \citep{afte-chumbley}
+and many more. In comparison to this, same source matching of bullets to
+firearms has not been examined as prominently as that of toolmarks. Even
+less information is available on validity of methods and error rates
+associated with firearms examination. The National Academy of Sciences
+in its report in 2009 \citep{NAS:2009} discussed the need for
+determining error rates in methods proposed for firearms examination. As
+seen in the case of most forensic applications, the first step for same
+source matching and error rate determination involves identification of
+unique features that are characteristic of the object at hand. For the
+case of bullets and firearms, striation marks on the surface of the
+bullet are considered to be such markings that can be used in methods
+for same source matching. These marks are often a product of rifling and
+impurities and defects due to manufacturing in the barrel of the gun,
+which leads to engravings on the bullet surface
+\citep{afte-article1992}. In current practice, firearm examiners
+invariably make visual comparisons of bullet striae and use visual
+assesment tools to dignify bullets as being matches and non-matches. One
+way of accomplishing anykind of comparison between bullets is to do a
+comparison between surface marking of two or more bullet lands. Bullet
+Lands are considered to be areas between grooves made by the rifling
+action of the barrel and the markings on them are considered to be
+unique. The land engraved markings or sometimes termed as Bullet
 profiles \citep{aoas} are striation marks made on Bullet lands and often
@@
 at which profile is to be selected, removing curvature, smoothing,
-identifying groove locations are explained.
+identifying groove locations are explained. Figure \ref{fig:rgl} shows
+us how the extracted signature from a bullet land lines up with the
+image of the land from which it was extracted. We can see also see in
+the figure how the depth and relative position of the striation markings
+seen in the image are interpreted as the signature.
 
-In the study conducted by \citet{aoas} a machine learning based
-algorithm was developed for same source matching of bullets and error
-rates were discussed using the database from the Hamby Study
-\citep{hamby}. In this paper, we first try to adapt a deterministic
-algorithm and method developed for toolmarks by \citet{chumbley} and
-improved by \citet{hadler}, to bullets. Then we consequently discuss
-about the efforts in doing so along with the associated error rates. The
-data used in this paper also belongs to the Hamby Study \citep{hamby}.
-This gives us a common platform for comparing the performance of the
-chumbley method on bullets with an already existing method proposed by
-\citet{aoas} for bullets. The proposed algorithm and method of
-\citet{chumbley}, in their paper, compares two toolmarks with the
-intention of determining if it comes from the same source (same tool).
-The method also provides a means to determine error rates and claims to
-reduce subject bias. As mentioned earlier, subject bias and error rate
-determination have been a long standing issue in firearm examination
-\citep{NAS:2009}. This remains one of the motivations to explore the
-adaptibility of the Chumbley score methodology to bullets.
-\citet{chumbley} used an empirical based setup to validate their
-proposed algorithm and quantitative method which calculates a U-statisic
-for the purpose of classification of toolmarks as matching or
-non-matching. The data for their study was obtained from 50 sequentially
-manufactured screwdriver tips, and preselected comparison window sizes
-were given as inputs to the algorithm. The algorithm then compares the
-two toolmarks and comes up with a U-statistic and an associated p-value
+As mentioned earlier, subject bias and error rate determination have
+been a long standing issue in firearm examination \citep{NAS:2009}. This
+issue of removing subjectivity and making firearm analysis objective was
+also raised in the \citet{pcast2016} report. For an objective analysis,
+scientific principle testability and error rate determination are
+considered to be defining aspects. Identification of the uncertainities
+associated with a method of comparison and coming up with a systemized
+method of quantification, is therefore important. Associating
+mathematical probabilites in a systematic manner to these methods are
+one such way of quantification. In a previous study conducted by
+\citet{aoas} a machine learning based algorithm was developed for same
+source matching of bullets and error rates were discussed using the
+database from the Hamby Study \citep{hamby}. The method proposed by
+\citet{chumbley} (and later improved by \citet{hadler}) also provides a
+means to determine error rates and claims to reduce subject bias.
+Therefore, giving us a strong motivation to explore the adaptibility of
+the Chumbley score methodology to bullets. In this paper, we take this
+toolmark based methodology and try to apply and adapt it to bullets.
+After this, we consequently discuss about the efforts in doing so, along
+with the associated error rates. The data used in this paper also
+belongs to the Hamby Study \citep{hamby}. This gives us a common
+platform for comparing the performance of the chumbley method on bullets
+with the already existing method proposed by \citet{aoas} for bullets.
+As a brief overview, \citet{chumbley}, in their paper, compare two
+toolmarks with the intention of determining if it comes from the same
+tool. They use an empirical based setup to validate their proposed
+algorithm and quantitative method which calculates a U-statisic for the
+purpose of classification of toolmarks as matching or non-matching. The
+data for their study was obtained from 50 sequentially manufactured
+screwdriver tips, and preselected comparison window sizes were given as
+inputs to the algorithm. The algorithm then compares the two toolmarks
+and comes up with a Mann-Whitney U-statistic and an associated p-value
 to designate them as matches or non-matches. The performance for every
@@
 
-Bullets are much smaller in length, width, are not flat and curved in
-the cross-sectional topography as opposed to tools like screw driver
-tips which produced longer and pronounced markings. This means the
-makings made by barrels in comparison to toolmarks may have a problem in
-distinctiveness. The majority of Bullet profiles and signatures
-extracted by procedures mentioned by \citet{aoas} are almost 1/4 th the
-size of toolmarks as used by \citet{chumbley} or even smaller.
-Striations on Bullets are made on their curved surfaces, whereas the
-algorithm developed by \citet{chumbley} and \citet{hadler} has only been
-tested for flatter and wider surfaces which have negligible curvature.
-Therefore, using methods proposed for toolmarks may need adaptation in
-order to give tangible results for bullets. Moreover, in order to to get
-flat bullet signatures and remove the curvatures some kind of smoothing
-needs to be applied as a pre-step which needs further investigation as
-to whether the level of smoothing does effect the working of the
-algorithm on Bullets.
+Bullet striations are different from toolmarks. Bullet markings are
+typically produced by the barrel on the bullets, while toolmarks can be
+made by tool tips on different surfaces. Tools like screw driver tips as
+used by \citet{chumbley}, produce longer and pronounced markings.
+Bullets on the other hand are much smaller in length, width, and curved
+in the cross-sectional topography. This means the markings made by
+barrels as opposed to markings made by tools, may have a problem in
+distinctiveness. The majority of Bullet profiles and signatures of the
+Hamby study \citep{hamby} used in this paper are extracted by procedures
+mentioned by \citet{aoas}. The Ruger pistol barrels used in the Hamby
+study have a 9mm caliber. This means that bullets used in these pistols
+have land markings of almost \((1/6)^{th}\) the size of screwdriver
+toolmarks used by \citet{chumbley}. Also, striations on bullets are made
+on their curved surfaces, whereas the algorithm developed by
+\citet{chumbley} and \citet{hadler} has only been tested for flatter and
+wider surfaces which have negligible curvature. Therefore, using methods
+proposed for toolmarks may need adaptation in order to give tangible
+results for bullets. Moreover, in order to get flat bullet signatures
+and remove the curvatures some kind of smoothing needs to be applied as
+a pre-step wihch needs further investigation as to whether the level of
+smoothing does effect the working of the algorithm on Bullets.
 
-Also when in the optimization step, the Window of optimization for
-bullets will be shorter as the signatures are smaller. The idea is to
-keep the number of windows of optimization sufficiently large, which
-means shorter Trace segments (partition of signature or toolmark with
-length = size of window of optimization) that lets us compare smaller
-segments of one signature to another. This introduces a problem as, if
-we go too small in the window of optimization, the unique features of
-the trace segments are lost and seem similar, while too large sizes
-vastly reduces the weight of small features that would otherwise
-uniquely classify a signature and hence identify the region of
-agreement.
+The Chumbley method and algorithm works on the premise of comparison
+windows which are supposed to represent small segments of pixels of
+equal sizes in the two markings that are to be compared in a
+predetermined fashion. They term them as windows of optimization and
+validation. The optimization window is selected in the first step called
+the optimization step where areas/windows of best agreement between the
+the two markings are found. The window of optimization for bullet
+striations are shorter as bullet signatures are smaller compared to
+toolmarks . The idea is to keep the number of windows of optimization
+sufficiently large. This means we have shorter trace segments that lets
+us compare smaller segments of one signature to another than those for
+toolmarks. Here, trace segments are the partitions of marking with the
+length of the partition equal to the size of window of optimization.
+This introduces a problem as, if we go too small in the window of
+optimization, the unique features of the trace segments are lost and
+seem similar, while too large sizes vastly reduces the weight of small
+features that would otherwise uniquely classify a signature and hence
+identify the region of agreement. The problem of distinctiveness in the
+structure of bullet striations may also lead to potential failure of the
+algorithm to identify correct windows of agreements or not be able to
+identify these windows at all. This may lead to incorrect classification
+of bullets as matches or non-matches.
 
-Thus the Window of Optimization has a direct influence on whether we are
-Falsely rejecting the null when it is true (Type I) or Falsely accepting
-the null when it is false (Type II) making identification of the optimum
-size of window of optimization very important. This raises important
-questions about what parameter settings need to be chosen for bullet
-land comparison. Something similar is done in \citet{afte-chumbley} for
+Thus, the comaparison windows have a direct influence on the false
+positives and false negatives, making identification of the optimum
+sizes of these windows very important. This raises important questions
+about how and what parameter settings need to be chosen for bullet land
+comparison. Something similar is done in \citet{afte-chumbley} for
 toolmark comparisons of slip-joint pliers where optimum window sizes are
 determined. There is a need to figure out the best parameter settings
-which minimizes the errors for unsmoothed markings and pre-processed
-signatures as determined by \citet{aoas}. An analysis of these error
-rates and comparison with other methods will help us understand the
-adaptibility of the chumbley score to bullets.
-
-\hh{figure \ref{fig:rgl} needs to be mentioned somewhere in the write-up}
+which minimizes the errors for unsmoothed markings or profiles and
+pre-processed signatures. An analysis of these error rates and
+comparison with other methods will help us understand the adaptibility
+of the chumbley score to bullets.
 
@@
 
-\section{The Chumbley Score Test}\label{the-chumbley-score-test}
+\subsection{The Chumbley Score Test}\label{the-chumbley-score-test}
 
-The Chumbley score algorithm takes input as two vectorized processes
-\(x(t_1)\), \(t_1 = 1,2,...T_1\) and \(y(t_2)\), \(t_2 = 1,2...T_2\)
-which denote two sets of marks or striae. Here the processes are of the
-form \(z(s)\) which is a spatial process for some \(t\), while
-\(z(s_1)\) is the realization of the process in \(s_1\). The \(t_1\) and
-\(t_2\) can be better understood as equally spaced pixel locations for
-the two marks under consideration, where a `pixel' is the resolution of
-the confocal light microscope. In the case for bullet signatures a pixel
-corresponds to 0.645 microns. These marks or striae are potentially from
-two different bullets or two different toolmarks whose source needs to
-be identified as being same or different. The marks or striae are
-indexed by the pixel location where \(t_1\) is for the first striae
-referred to as \(x\) and \(t_2\) is for the second striae which is
-referred as \(y\). \(T_1\) and \(T_2\) being the respective lengths of
-the markings, need not be the same but are usually of similar lenghtsThe
-similarity is then judged by the algorithm on the basis of
-cross-correlation of a fixed and constant number consecutive pixels say
-\(k\) taken from the two indexed marks \(x(t_1)\) and \(y(t_2)\) such
-that in theory \(k\) remains smaller that length of the two striae or
-marks which is the same as \(t_1\) and \(t_2\). Depending on the what
-stage of the algorithm we are in, matching of different pixel lengths
-and locations is done, which at the end effectively compares all
-possible windows that would guarantee in quantifying the two marks or
-striae as coming from the same source or not, which also lets us assess
-the error rates by checking for a large number of cases.
+The Chumbley score algorithm takes input as two vectorized processes.
+Here the processes are of the form \(z(t)\) which is a spatial process
+for some location indexed with \(t\). This means that while \(z(t)\)
+represents any striae, \(z(t_1)\) is a realization of the spatial
+process. \(t\) here denotes a vector of equally spaced pixel locations
+for the striation marks under consideration. The word `pixel' refers to
+the resolution of the confocal light microscope. In the case for bullet
+signatures extracted from NIST ballistics database \citep{nist} of the
+Hamby study, a pixel corresponds to approximately 1.56 microns. The
+vectorized processes representing two sets of striaes are therefore
+shown as \(x(t_1)\), \(t_1 = 1,2,...T_1\) and \(y(t_2)\),
+\(t_2 = 1,2...T_2\). Here \(x\) and \(y\) denote two instances of the
+process \(z(*)\) which means two striation marks, while the indexing
+\(t\) or pixel range is shown as \(t_1\) and \(t_2\) for \(x\) and \(y\)
+respectively. This representation of \(t\) as \(t_1\) and \(t_2\) lets
+us identify them as either of same or different lengths. The striation
+marks under consideration are potentially from two different bullets
+whose source needs to be identified as being same or different. \(T_1\)
+and \(T_2\), as represented above, are the final pixel indexes of each
+marking and therefore give the respective lengths of the markings. The
+similarity in the two markings is then judged by the algorithm on the
+basis of cross-correlation of a fixed and constant number consecutive
+pixels (say \(k\)) taken from the two markings. This can for example be
+\(k\) taken from the two indexed marking \(x(t_1)\) and \(y(t_2)\) such
+that in theory \(k\) remains smaller than length of the two striae or
+marks. Depending on what stage of the algorithm we are in, matching of
+different pixel lengths and locations is done. This in the end,
+effectively compares all possible windows that would guarantee in
+quantifying the two marks or striae as coming from the same source or
+not.
 
@@
 validation step, at the end of which a Mann Whitney U statistic is
-calculated. A pre-processing step to the algorithm is to choose a
-coarseness value which is used as a parameter to the LOWESS smoothing
+calculated as a measure to differentiate between matches and
+non-matches. A pre-processing step to the algorithm is to choose a
+coarseness value which is used as a parameter to the lowess smoothing
 function. The coarseness essentially gives the proportion of points
 which influence the smooth at each value, which means larger values lead
-to more smoothness. The LOWESS smoothing is applied to each of two sets
+to more smoothness. The lowess smoothing is applied to each of two sets
 of vectorized striae or marks \(x(t_1)\) and \(y(t_2)\), before
-proceeding to the algorithm.
+proceeding to the algorithm. The algorithm starts with the optimization
+step where the area of best agreement in the two markings being compared
+is identified. Comparison window size is predefined. Each window of the
+first marking is compared with all windows of the second marking. A
+maximum correlation statistic is used to identify the region of best
+agreement, with the maximum usually seen being near 1 for both cases
+which is what is intuitively expected for matches, and not expected
+intuitively for non matches. \citet{hadler} in their paper proposed an
+improvement to this algorithm by trying to remove mutual dependence of
+parameters. The mututal dependence was due to the serial correlation in
+surface depth values of a marking. Also a random sampling sub step in
+the validation phase makes a group of pixels to be chosen more than once
+and hence introduces lack of independence in certain steps. The reason
+for removing the mutual dependence was the Mann-Whitney U statistic
+works under the assumption of independence of parameters. The validation
+step builds on this with two sub-steps namely, same shift and different
+shift. As the aim is to come up with a non-parameteric U statistic,
+\citet{hadler} proposed a normalization procedure in the validation step
+which goes to some extent to address the issue of mutual dependence. A
+series of windows are chosen in both the same and different shift
+sub-steps for the purpose of comparison. Both substeps were modified by
+\citet{hadler} who introduced a deterministic rule for sampling of
+windows in the same-shift and different shift as opposed to the
+originally proposed random sampling by \citet{chumbley}. In th same
+shift substep the chosen series of windows are at a common distance
+(rigid-shifts) from the window earlier identified as the region of best
+agreement in the optimization step. The correlation of these windows
+generally turn out to be lower than the maximum correlation window. The
+significance is that these same shift windows still have large enough
+correlation values for the two markings being compared that are in
+reality a match. Any choice of maximum correlation windows (where
+correlation values are often near 1) for non-matches in the optimization
+step are also validated by this sub-step. This is because the same-shift
+correlations would not be anywhere as large for same-shift windows when
+all windows in the sub-step are compared. The different shift substep on
+the other hand gives perspective to the correlation values of the same
+shift window correlation values. There are no rigid-shifts but different
+shifts where distance from the maximum correlation window are chosen
+randomly by \citet{chumbley} and deterministically as per
+\citet{hadler}. This means that there is an equal possibilty of
+comparing a trace segment from one marking to any trace segment or
+window in the second marking.
 
-\citet{hadler} in their paper proposed an improvement to this algorithm
-by trying to remove mutual dependence of parameters (due to serial
-correlation in surface depth values of a toolmark and because of a
-random sampling sub step in the validation phase which makes a group of
-pixels to be chosen more than once and hence introduces lack of
-independence) in certain steps, especially because the Mann-Whitney U
-statistic that is later calculated in the algorithm and used as a
-measure to differentiate between matches and non-matches works under the
-assumption of independence of parameters.
-
-Since we are only interested in a non-parameteric U statistic, Hadler et
-al. proposed a normalization procedure in the Validation step that goes
-to some extent to address the issue of mutual dependence. Also the same
-shift and different shift substep was modified to use a deterministic
-rule for sampling sam-shhift and dofferent shift-samples as opposed to
-the originally proposed random samples.
-
-The data that is used by \citet{chumbley} is generated by a surface
-profilometer that gives the height in terms of distance along a linear
-trace. This is taken perpendicular to the striations present in the
-toolmark. Two such trace are then compared using the algorithm.
-
-\subsection{Detailed algorithm}\label{detailed-algorithm}
-
-\subsubsection{Optimization step}\label{optimization-step}
-
-The idea behind this step is to first identify the area of best
-agreement in the two toolmark data. Comparison window size is predefined
-by the user, which in case of screwdriver toolmarks was chosen by
-Chumbley et al and Hadler et al as around 10 percent of the length of
-the toolmarks, which was around 500. This window is henceforth referred
-to as Window of optimization.
-
-A maximum correlation statistic is used to identify the region of best
-agreement, with the maximum usually seen being near 1 for both cases
-which is what we intuitively expect for matches, and something which we
-do not intuitively expect for non matches
-
-First, there are a very large number of cross-correlations calculated
-for the two series of striae or marks \(x(t_1)\) and \(y(t_2)\), for eg
-if the window of optimization was defined as 200 and toolmark pixel
-length is around 1000 then we have 1200-(200-1) = 1001.
-
-This is the number of windows we have for one toolmark and each window
-is compared with all windows of the second toolmark (the number of
-windows is again similar), and the window with the maximum correlation
-is identified to be the region where the toolmarks are in maximum
-agreement with each other
-
-\subsubsection{Validation Step}\label{validation-step}
-
-\paragraph{Same-shift:}\label{same-shift}
-
-In this step a series of windows are chosen at random (originally by
-chumbley et al) and deterministically (by hadler et al), but at a common
-distance (rigid-shift) from the window identified as the region of best
-agreement in the vaildation step. The correlation of these windows would
-be as intuitively assumed i.e.~lower than the maximum correlation window
-(Optimization step), but the significance is that these same shift
-windows will still have large enough correlation values for two
-toolmarks or signatures that are in reality a match.
-
-This also validates that if in the optimization step the maximum
-correlation window chosen (with correlation value near 1) was by
-accident (like in case of signatures that in reality are a non-match),
-all these same- shift correlations (A fixed number of these trace
-segments are identified) would not be anywhere large enough for all
-same-shift windows.
+Neither of above sets of correlation in the two sub-steps are allowed to
+include the maximum correlation window as identified earlier. Therefore
+the assumption is that if two markings match each other, the same-shift
+correlations would be larger than the different-shift windows. And if
+they are not a match the correlations in the two sets will be very
+similar. The U-statistic tests for the null hypothesis that the two
+markings are not a match and are therefore not made by the same source.
+As given by \citet{hadler}, it is computed from the joint rank of all
+correlations of both the same and different shift samples.
 
@@
 
-\paragraph{Different Shift:}\label{different-shift}
-
-Primary reason for this substep is to give perspective to the
-correlation values of the same shift window correlation values.
-
-This time there are no rigid-shifts but different shifts (distance from
-Window of Opt with max correlation) chosen randomly by \citet{chumbley}
-and deterministically by Hadler et al. such that there is an equal
-possibilty of comparing a trace segment from one signature or toolmark
-to any one in the second signature or toolamrk.
-
-Neither of above sets of correlation are allowed to include the maximum
-correlation window as identified earlier.
-
-Therefore the assumption is that if two toolmarks or signatures match
-each other the same-shift correlations would be larger than the
-different-shift windows, and if they are not a match the correlations in
-the two sets will be very similar.
-
-\subsubsection{U Statistic:}\label{u-statistic}
-
-This is computed from the joint rank of all correlations of both the
-same and different shift samples. As given by \citet{hadler}
-
-Null Hypothesis: If the toolmarks were not match i.e not made by the
-same tool.
-
-Let ns and nd be the number of same shift and different shift windows
-\[N = n_{s} + n_{d}\]
-
-The mann whitney U statistic is given by
-\[U =\sum^{ns}_{i=1}R_{s}\left( i\right)\]
-
-with the standardized version which includes provision for rank ties
-
-\[\overline{U}= \dfrac{U-M}{\sqrt{V}}\]
-
-where prior to normalization the U-statistic has the mean as
-
-\[M = n_{s}\left(\dfrac{N+1}{2}\right)\]
-
-and variance
-
-\[V = \dfrac{n_{s}n_{d}}{N\left(N-1\right)}\left[\Sigma^{n_{s}}R_{s}\left(i\right)^{2}+\Sigma^{n_{d}}R_{d}\left(j\right)^{2}\right] -\dfrac{n_{s} n_{d}\left(N+1\right)^{2}}{4\left(N-1\right)}\]
-
-\subsection{Testing setup}\label{testing-setup}
+\section{Simulation setup}\label{simulation-setup}
 
@@
 \toprule
-wo & 50 & 50 & 60 & 60 & 80 & 80 & 90 & 90 & 100 & 100 & 110 & 110 & 120 & 120 & 120 & 120 & 120 & 120\\
-wv & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 10 & 20 & 30 & 40 & 50 & 60\\
+wo & 50 & 50 & 60 & 60 & 80 & 80 & 90 & 90 & 100 & 100 & 110 & 110 & 120 & 120 & 120 & 120 & 120 & 130\\
+wv & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 10 & 20 & 30 & 50 & 60 & 30\\
 \bottomrule
@@
 \toprule
-wo & 130 & 130 & 140 & 140 & 150 & 150 & 160 & 160 & 200 & 200 & 200 & 240 & 240 & 280 & 280\\
-wv & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 20 & 30 & 50 & 30 & 50 & 30 & 50\\
+wo & 130 & 140 & 140 & 150 & 150 & 160 & 160 & 200 & 200 & 200 & 240 & 240 & 280 & 280 & 320\\
+wv & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 20 & 30 & 50 & 30 & 50 & 30 & 50 & 30\\
 \bottomrule
@@
 \midrule
-\addlinespace[0.3em]
+\addlinespace[0.5em]
 \multicolumn{4}{l}{\textbf{Size of Optimization Window = 280}}\\
@@
 \midrule
-\addlinespace[0.3em]
+\addlinespace[0.5em]
 \multicolumn{4}{l}{\textbf{Size of Optimization Window = 120}}\\
@@
 \midrule
-\addlinespace[0.3em]
+\addlinespace[0.5em]
 \multicolumn{4}{l}{\textbf{Size of Optimization Window = 80}}\\
@@
 \midrule
-FALSE & 30 & 0.452 & 0.027\\
+FALSE & 30 & 0.503 & 0.028\\
 FALSE & 50 & 0.922 & 0.035\\
-TRUE & 30 & 0.037 & 0.004\\
+TRUE & 30 & 0.043 & 0.004\\
 TRUE & 50 & 0.056 & 0.005\\
@@
 
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figures/wo120-profile-1}
-
-}
-
-\caption{The figure shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120}\label{fig:wo120-profile}
-\end{figure}
-
 \bibliographystyle{agsm}

### 2018-02-24 doc/paper.Rmd
additions to the introduction

@@
 ## Motivation
+Same source analyses are a major part of an Forensic Toolmark examiners job.
+In current practice firearm examiners  make these comparisons visually under a comparison microscope and make one of the following four conclusions: identification, inconclusive, elimination or unsuitable for examination \citet{afte-website-2}.
+These conclusions are made on the basis of "unique surface contours" of the two toolmarks being in "sufficient agreement" \citep{afte-toolmarks1998}.
+AFTE describes the term "sufficient agreement" as the possibility of another tool producing the markings under comparison, as practically impossible \citep{afte-toolmarks1998} \citep{afte-website}.
+\gk{Although, they go on to concede that the interpretation of the identification depends on the examiner and is  therefore subjective in nature.
+}
+
+One of the main criticisms in the toolmark examination first raised by the National Research Council in 2009 \citep{NAS:2009} and later emphasized further by the President's Council of Advisors on Science and Technology \citep{pcast2016} are the amount of subjectivity in the assessment and the lack of error rates.
+Using a statistical approach removes both subjectivity from the assessment and allows us to provide error rates for both false positive and false negative identifications.
+
+Digitized images, e.g.
+profilometers or confocal microscopy allow us to measure the surface \citep{vorburger2016} and can be used to conduct statistical analysis on the markings.
+Multiple toolmarks have been studied in the literature:  \citet{manytoolmarks1, chumbley} have studied screwdriver marks digitized using a profilometer,  \citet{manytoolmarks2} have investigated screwdriver, tongue and groove pliers marks that are generated using a confocal microscope,  while slip-joint pliers have been investigated by \citet{afte-chumbley} using digitized marks generated by a surface profilemeter.
+\citet{manytoolmarks1} used an algorithm which compares many small windows in the markings between two toolmarks and the maximum pearson correlation coefficient as the statistical index for concluding matches and non-matches.
+The Chumbley scoring method, first introduced by \citet{chumbley} uses a similar but much more extenstive framework which includes a non-parametric statistical test.
+This was later improved by Hadler and Morris \citep{hadler}.
+In this paper, we are investigating the applicability of the Chumbley scoring method by \citet{hadler} to assessing striation marks on bullets for same-source identification.
+
+Striation marks on bullets are made by impurities in the barrel.
+As the bullet travels through the barrel, these imperfections leave "scratches" on the bullet surface.
+Typically, only striation marks in the land engraved areas (LEAs) are considered \citet{afte-article1992}.
+Bullet Lands are depressed areas between bullet grooves made by the rifling action of the barrel.
+Compared to toolmarks made by screwdrivers striation marks on bullets are typically much smaller, both in length and in width.
+Bullets also have a curved cross-sectional topography.
+
+Bullet matching methods depend on the attribute being used for making comparisons.
+For single feature based matching \citet{chu2013} used an automatic method for counting consequtive matching striae (CMS).
+They reported zero false positive rate but do not discuss about the false negative rates for land to land matches.
+\citet{ma2004} and \citet{vorburger2011} discuss about CCF (cross-correlation function) and its discriminating power and application but do not talk about matches and non-matches.
+\citet{aoas} use multiple attributes like CCF, CMS and compare every land against every other land of the Hamby dataset \citet{hamby}.
+This also includes matching of all the unknown bullet lands against the known bullet lands.
+They report for 10,731 comparisons, 18 false negatives and 9 false negatives.
+
+The Chumbley score  provides us with another approach in the same-source assessment of bullet striation marks.
+\citet{chumbley}, in their paper, compare two toolmarks with the intention of determining if it comes from the same tool.
+The data for their study was obtained from 50 sequentially manufactured screwdriver tips.
+For every 100 comparisons \citet{chumbley} reported 2 false negatives and 9 false positives.
+The improvement proposed by \citet{hadler} on the other hand gave 0 false  negatives and 3 false poisitives.
+
+\gk{\citep{afte-website} quotes sufficient agreement as "agreement of individual characteristics is of a quantity and quality that the likelihood another tool could have made the mark is so remote as to be considered a practical impossibility"}.}
+\gk{could not find the pdf for AFTE theory of identification as it relates to toolmarks pdf.
+I believe it is not available for free.
+But from the AFTE Journal the abstract was available which mentions the term " sufficient agreement" as quoted above.
+Currently the website seems like the most reliable way of quoting AFTE, i have checked other sources in both occasions where the website has been quoated above, like an FBI paper available at the FBI website and some non published but available pdf files in the AFTE website that have used these lines and quoted the AFTE journal.
+But since I cannot verify it in the Journal I have quoated the website.}
+\gk{faden et al have not cited any paper that discusses the algorithm, i assume it was developed for the paper, the discussion of which is not as extensive as chumbley's.
+Its a variation of the chumbley method but from my understanding it is just the optimization step.
+and this paper came out much before the chumbley score.}
+
+<!-- #### -->
+
 \hh{XXX first sentence: Same source analyses are a major part of an Forensic Toolmark examiners job.
@@
 Bullet Lands are considered to be areas between grooves made by the rifling action of the barrel and the markings on them are considered to be unique.
-The land engraved markings or sometimes termed as Bullet profiles \citep{aoas} are striation marks made on Bullet lands and often used for these land to land comparisons.
+The land engraved markings or sometimes termed as Bullet profiles \citep{aoas,ma2004} are striation marks made on Bullet lands and often used for these land to land comparisons.
 Bullet Signatures is another word used in literature as seen in the work of \citet{chu2013} and \citet{aoas}.

### 2018-02-24 doc/paper.Rmd
small change

@@
 AFTE describes the term "sufficient agreement" as the possibility of another tool producing the markings under comparison, as practically impossible \citep{afte-toolmarks1998} \citep{afte-website}.
-\gk{Although, they go on to concede that the interpretation of the identification depends on the examiner and is  therefore subjective in nature.
-}
+\gk{Although, they go on to concede that the interpretation of the identification depends on the examiner and is  therefore subjective in nature.}
 
@@
 
-\gk{\citep{afte-website} quotes sufficient agreement as "agreement of individual characteristics is of a quantity and quality that the likelihood another tool could have made the mark is so remote as to be considered a practical impossibility"}.}
+\gk{\citep{afte-website} quotes sufficient agreement as "agreement of individual characteristics is of a quantity and quality that the likelihood another tool could have made the mark is so remote as to be considered a practical impossibility"}
 \gk{could not find the pdf for AFTE theory of identification as it relates to toolmarks pdf.

### 2018-02-24 doc/paper.Rmd
small changes

@@
 But from the AFTE Journal the abstract was available which mentions the term " sufficient agreement" as quoted above.
-Currently the website seems like the most reliable way of quoting AFTE, i have checked other sources in both occasions where the website has been quoated above, like an FBI paper available at the FBI website and some non published but available pdf files in the AFTE website that have used these lines and quoted the AFTE journal.
+Currently the website seems like the most reliable way of quoting AFTE, i have checked other sources in both occasions where the website has been quoated above, like an FBI paper available at the FBI website \citet{fbi-paper} and some non published but available pdf files in the AFTE website that have used these lines and quoted the AFTE journal.
 But since I cannot verify it in the Journal I have quoated the website.}

### 2018-02-25 doc/paper.Rmd
error rate reporting modified from absolute numbers to rates

@@
 profilometers or confocal microscopy allow us to measure the surface \citep{vorburger2016} and can be used to conduct statistical analysis on the markings.
-Multiple toolmarks have been studied in the literature:  \citet{manytoolmarks1, chumbley} have studied screwdriver marks digitized using a profilometer,  \citet{manytoolmarks2} have investigated screwdriver, tongue and groove pliers marks that are generated using a confocal microscope,  while slip-joint pliers have been investigated by \citet{afte-chumbley} using digitized marks generated by a surface profilemeter.
-\citet{manytoolmarks1} used an algorithm which compares many small windows in the markings between two toolmarks and the maximum pearson correlation coefficient as the statistical index for concluding matches and non-matches.
-The Chumbley scoring method, first introduced by \citet{chumbley} uses a similar but much more extenstive framework which includes a non-parametric statistical test.
+Multiple toolmarks have been studied in the literature:  \citet{manytoolmarks1, chumbley} have studied screwdriver marks digitized using a profilometer,  \citet{manytoolmarks2} have investigated screwdriver, tongue and groove pliers 3D marks that are generated using a confocal microscope,  while, slip-joint pliers have been investigated by \citet{afte-chumbley} using digitized marks generated by a surface profilemeter.
+\citet{manytoolmarks2} define a relative distance metric and use it as similarity measure between two toolmarks.
+\citet{manytoolmarks1} used an algorithm which compares many small segments in the markings between two toolmarks.
+A maximum pearson correlation coefficient is then used as the statistical index for comparison.
+The Chumbley scoring method, first introduced by \citet{chumbley}, uses a similar but much more extenstive framework which includes a non-parametric statistical test.
 This was later improved by Hadler and Morris \citep{hadler}.
@@
 \citet{ma2004} and \citet{vorburger2011} discuss about CCF (cross-correlation function) and its discriminating power and application but do not talk about matches and non-matches.
-\citet{aoas} use multiple attributes like CCF, CMS and compare every land against every other land of the Hamby dataset \citet{hamby}.
-This also includes matching of all the unknown bullet lands against the known bullet lands.
-They report for 10,731 comparisons, 18 false negatives and 9 false negatives.
+\citet{aoas} use multiple attributes like CCF, CMS, D (distance measure) etc in a random forrest based method and compare every land against every other land of the Hamby dataset \citet{hamby}.
+They report an out-of-bag error rate with  false positive error rate of 0.0001 and false negative error rate of 0.2267 for 100 trees.
+For 300 trees they report correct prediction of all matches and non-matches.
 
@@
 \citet{chumbley}, in their paper, compare two toolmarks with the intention of determining if it comes from the same tool.
-The data for their study was obtained from 50 sequentially manufactured screwdriver tips.
-For every 100 comparisons \citet{chumbley} reported 2 false negatives and 9 false positives.
-The improvement proposed by \citet{hadler} on the other hand gave 0 false  negatives and 3 false poisitives.
+The data for their study was obtained from 50 sequentially manufactured screwdriver tips.\citet{chumbley} reported error rates for marking made by the tips at differnt angles.
+For a 30 degree they showed false negative error rate to be 0.023 and  false positive error rate be 0.09.
+For other angles the error rates remain the same for false negatives and 0.01 for false positives.
+The improvement proposed by \citet{hadler} on the other hand, did not consider effect of angles and gave 0.06 false  negative error rate and 0 false poisitive error rate for markings made at the same angle.
 
@@
 
-```{r, results='asis'}
+```{r, results='asis', include=FALSE}
 library(knitr)
@@
 toolmark1.confusion$Classification<- c("Match", "Non-Match")
-toolmark1.confusion$Match<- c(41,2)
-toolmark1.confusion$`Non-Match`<- c(9,48)
+# toolmark1.confusion$Match<- c(41,2)
+# toolmark1.confusion$`Non-Match`<- c(9,48)
+toolmark1.confusion$Match<- c(91,2)
+toolmark1.confusion$`Non-Match`<- c(9,98)
 

### 2018-02-26 doc/paper.Rmd
up to  line 226

@@
 ## Motivation
-Same source analyses are a major part of an Forensic Toolmark examiners job.
-In current practice firearm examiners  make these comparisons visually under a comparison microscope and make one of the following four conclusions: identification, inconclusive, elimination or unsuitable for examination \citet{afte-website-2}.
+Same source analyses are a major part of an Forensic Toolmark Examiner's job.
+In current practice examiners  make these comparisons visually under a comparison microscope and come to one of the following four conclusions: identification, inconclusive, elimination or unsuitable for examination \citet{afte-website-2}.
 These conclusions are made on the basis of "unique surface contours" of the two toolmarks being in "sufficient agreement" \citep{afte-toolmarks1998}.
 AFTE describes the term "sufficient agreement" as the possibility of another tool producing the markings under comparison, as practically impossible \citep{afte-toolmarks1998} \citep{afte-website}.
-\gk{Although, they go on to concede that the interpretation of the identification depends on the examiner and is  therefore subjective in nature.}
-
-One of the main criticisms in the toolmark examination first raised by the National Research Council in 2009 \citep{NAS:2009} and later emphasized further by the President's Council of Advisors on Science and Technology \citep{pcast2016} are the amount of subjectivity in the assessment and the lack of error rates.
-Using a statistical approach removes both subjectivity from the assessment and allows us to provide error rates for both false positive and false negative identifications.
-
-Digitized images, e.g.
-profilometers or confocal microscopy allow us to measure the surface \citep{vorburger2016} and can be used to conduct statistical analysis on the markings.
-Multiple toolmarks have been studied in the literature:  \citet{manytoolmarks1, chumbley} have studied screwdriver marks digitized using a profilometer,  \citet{manytoolmarks2} have investigated screwdriver, tongue and groove pliers 3D marks that are generated using a confocal microscope,  while, slip-joint pliers have been investigated by \citet{afte-chumbley} using digitized marks generated by a surface profilemeter.
+%\gk{Although, they go on to concede that the interpretation of the identification depends on the examiner and is  therefore subjective in nature.} 
+This subjectivity in the assessment as well as the lack of error rates are the main points of criticisms in the toolmark examination first raised by the National Research Council in 2009 \citep{NAS:2009} and later emphasized further by the President's Council of Advisors on Science and Technology \citep{pcast2016}.
+
+Technological advances, such as profilometers and confocal microscopy allow to measure 3D surfaces \citep{vorburger2016} in a digitized form.
+This forms the basis of statistical analysis of toolmarks.
+A statistical approach based on data  removes both subjectivity from the assessment and allows a quantification of error rates for both false positive and false negative identifications.
+
+Various toolmarks have been studied in the literature:  screwdriver marks  digitized using a profilometer have been analyzed in \citet{manytoolmarks1} and \citet{chumbley};  \citet{manytoolmarks2} have investigated 3D marks from screwdriver, tongue and groove pliers captured using a confocal microscope;  digitized marks from slip-joint pliers generated by a surface profilometer have been investigated by \citet{afte-chumbley}.
+
 \citet{manytoolmarks2} define a relative distance metric and use it as similarity measure between two toolmarks.
-\citet{manytoolmarks1} used an algorithm which compares many small segments in the markings between two toolmarks.
-A maximum pearson correlation coefficient is then used as the statistical index for comparison.
-The Chumbley scoring method, first introduced by \citet{chumbley}, uses a similar but much more extenstive framework which includes a non-parametric statistical test.
-This was later improved by Hadler and Morris \citep{hadler}.
-In this paper, we are investigating the applicability of the Chumbley scoring method by \citet{hadler} to assessing striation marks on bullets for same-source identification.
+\citet{manytoolmarks1} extract many small segments in the markings of two toolmarks and compare similiarity using a maximum pearson correlation coefficient.
+The Chumbley scoring method, first introduced by \citet{chumbley}, uses a similar but more extensive framework based on a Mann-Whitney U test of the resulting correlation coefficients.
+This approach was later improved by Hadler and Morris \citep{hadler}.
+\hh{XXX describe the improvement in half a sentence XXX} In this paper, we are investigating the applicability of the Chumbley scoring method by \citet{hadler} to assess striation marks on bullet lands for same-source identification.
 
@@
 Typically, only striation marks in the land engraved areas (LEAs) are considered \citet{afte-article1992}.
-Bullet Lands are depressed areas between bullet grooves made by the rifling action of the barrel.
+Bullet lands are depressed areas between the  grooves made by the rifling action of the barrel.
 Compared to toolmarks made by screwdrivers striation marks on bullets are typically much smaller, both in length and in width.
-Bullets also have a curved cross-sectional topography.
+Bullets also have a curved cross-sectional topography \hh{sketch?}.
 

### 2018-02-26 doc/paper.tex
up to  line 226

@@
 
+Same source analyses are a major part of an Forensic Toolmark Examiner's
+job. In current practice examiners make these comparisons visually under
+a comparison microscope and come to one of the following four
+conclusions: identification, inconclusive, elimination or unsuitable for
+examination \citet{afte-website-2}. These conclusions are made on the
+basis of ``unique surface contours'' of the two toolmarks being in
+``sufficient agreement'' \citep{afte-toolmarks1998}. AFTE describes the
+term ``sufficient agreement'' as the possibility of another tool
+producing the markings under comparison, as practically impossible
+\citep{afte-toolmarks1998} \citep{afte-website}.
+\%\gk{Although, they go on to concede that the interpretation of the identification depends on the examiner and is  therefore subjective in nature.}
+This subjectivity in the assessment as well as the lack of error rates
+are the main points of criticisms in the toolmark examination first
+raised by the National Research Council in 2009 \citep{NAS:2009} and
+later emphasized further by the President's Council of Advisors on
+Science and Technology \citep{pcast2016}.
+
+Technological advances, such as profilometers and confocal microscopy
+allow to measure 3D surfaces \citep{vorburger2016} in a digitized form.
+This forms the basis of statistical analysis of toolmarks. A statistical
+approach based on data removes both subjectivity from the assessment and
+allows a quantification of error rates for both false positive and false
+negative identifications.
+
+Various toolmarks have been studied in the literature: screwdriver marks
+digitized using a profilometer have been analyzed in
+\citet{manytoolmarks1} and \citet{chumbley}; \citet{manytoolmarks2} have
+investigated 3D marks from screwdriver, tongue and groove pliers
+captured using a confocal microscope; digitized marks from slip-joint
+pliers generated by a surface profilometer have been investigated by
+\citet{afte-chumbley}.
+
+\citet{manytoolmarks2} define a relative distance metric and use it as
+similarity measure between two toolmarks. \citet{manytoolmarks1} extract
+many small segments in the markings of two toolmarks and compare
+similiarity using a maximum pearson correlation coefficient. The
+Chumbley scoring method, first introduced by \citet{chumbley}, uses a
+similar but more extensive framework based on a Mann-Whitney U test of
+the resulting correlation coefficients. This approach was later improved
+by Hadler and Morris \citep{hadler}.
+\hh{XXX describe the improvement in half a sentence XXX} In this paper,
+we are investigating the applicability of the Chumbley scoring method by
+\citet{hadler} to assess striation marks on bullet lands for same-source
+identification.
+
+Striation marks on bullets are made by impurities in the barrel. As the
+bullet travels through the barrel, these imperfections leave
+``scratches'' on the bullet surface. Typically, only striation marks in
+the land engraved areas (LEAs) are considered \citet{afte-article1992}.
+Bullet lands are depressed areas between the grooves made by the rifling
+action of the barrel. Compared to toolmarks made by screwdrivers
+striation marks on bullets are typically much smaller, both in length
+and in width. Bullets also have a curved cross-sectional topography
+\hh{sketch?}.
+
+Bullet matching methods depend on the attribute being used for making
+comparisons. For single feature based matching \citet{chu2013} used an
+automatic method for counting consequtive matching striae (CMS). They
+reported zero false positive rate but do not discuss about the false
+negative rates for land to land matches. \citet{ma2004} and
+\citet{vorburger2011} discuss about CCF (cross-correlation function) and
+its discriminating power and application but do not talk about matches
+and non-matches. \citet{aoas} use multiple attributes like CCF, CMS, D
+(distance measure) etc in a random forrest based method and compare
+every land against every other land of the Hamby dataset \citet{hamby}.
+They report an out-of-bag error rate with false positive error rate of
+0.0001 and false negative error rate of 0.2267 for 100 trees. For 300
+trees they report correct prediction of all matches and non-matches.
+
+The Chumbley score provides us with another approach in the same-source
+assessment of bullet striation marks. \citet{chumbley}, in their paper,
+compare two toolmarks with the intention of determining if it comes from
+the same tool. The data for their study was obtained from 50
+sequentially manufactured screwdriver tips.\citet{chumbley} reported
+error rates for marking made by the tips at differnt angles. For a 30
+degree they showed false negative error rate to be 0.023 and false
+positive error rate be 0.09. For other angles the error rates remain the
+same for false negatives and 0.01 for false positives. The improvement
+proposed by \citet{hadler} on the other hand, did not consider effect of
+angles and gave 0.06 false negative error rate and 0 false poisitive
+error rate for markings made at the same angle.
+
+\gk{\citep{afte-website} quotes sufficient agreement as "agreement of individual characteristics is of a quantity and quality that the likelihood another tool could have made the mark is so remote as to be considered a practical impossibility"}
+\gk{could not find the pdf for AFTE theory of identification as it relates to toolmarks pdf.
+I believe it is not available for free.
+But from the AFTE Journal the abstract was available which mentions the term " sufficient agreement" as quoted above.
+Currently the website seems like the most reliable way of quoting AFTE, i have checked other sources in both occasions where the website has been quoated above, like an FBI paper available at the FBI website \citet{fbi-paper} and some non published but available pdf files in the AFTE website that have used these lines and quoted the AFTE journal.
+But since I cannot verify it in the Journal I have quoated the website.}
+\gk{faden et al have not cited any paper that discusses the algorithm, i assume it was developed for the paper, the discussion of which is not as extensive as chumbley's.
+Its a variation of the chumbley method but from my understanding it is just the optimization step.
+and this paper came out much before the chumbley score.}
+
 \hh{XXX first sentence: Same source analyses are a major part of an Forensic Toolmark examiners job.
@@
 }
-\hh{XXX we have digitized images, e.g. profilometers or Confocal microscopy allow us to measure the surface}
+\hh{XXX we have digitized images, e.g. profilometers or Confocal microscopy allow us to measure the surface [@vorburger2016]}
 
@@
 
-\hh{XXX In this pper, we are investigating the applicability of the Chumbley scoring method by @hadler to assessing striation marks on bullets for same-source identification. }
+\hh{XXX In this paper, we are investigating the applicability of the Chumbley scoring method by @hadler to assessing striation marks on bullets for same-source identification. }
 
 \hh{Striation marks on bullets are made by impurities in the barrel.
-As the bullet travels through the barrel, these imperfections leave "scratches' on the bullet surface.
+As the bullet travels through the barrel, these imperfections leave "scratches" on the bullet surface.
 Typically, only striation marks in the land engraved areas (LEAs) are considered (XXX citation) [@afte-article1992].
+XXX explain lands and grooves XXX.
 Compared to toolmarks made by screwdrivers striation marks on bullets are typically much smaller, both in length and in width.
@@
 }
+
+\hh{XXX How are bullets matched now? \citet{chu2013}: CMS, automatically - error rate?, using ccf XXX Vorburger paper XXX  and \citet{aoas}: multiple features, error rates }
+
+\hh{The Chumbley score  provides us with another approach in the same-source assessment of bullet striation marks. }
+
+\hh{end of intro: remaining paper is structured as follows: introduce to the data we get from confocal microscopy, introduce to profiles and signatures.
+Introduce to chumbley score method, apply chumbley and discuss results ....
+}
+
+\subsection{Scans for land engraved
+areas}\label{scans-for-land-engraved-areas}
+
+\begin{itemize}
+\tightlist
+\item
+  scans available: NIST database (citation), Hamby 44 and Hamby 252
+  (Hamby citation)
+\item
+  move figure up
+\item
+  discuss cross section, profiles and signatures
+\end{itemize}
 
@@
 unique. The land engraved markings or sometimes termed as Bullet
-profiles \citep{aoas} are striation marks made on Bullet lands and often
-used for these land to land comparisons. Bullet Signatures is another
-word used in literature as seen in the work of \citet{chu2013} and
-\citet{aoas}. In our context bullet signatures refer to a processed
+profiles \citep{aoas,ma2004} are striation marks made on Bullet lands
+and often used for these land to land comparisons. Bullet Signatures is
+another word used in literature as seen in the work of \citet{chu2013}
+and \citet{aoas}. In our context bullet signatures refer to a processed
 version of the raw land engraved markings or profiles. The generation of
@@
 improvement proposed by \citet{hadler} are listed in the table below.
-
-\begin{table}[!htb]
-    \begin{minipage}{.5\linewidth}
-      \caption{Chumbley et al. 2010}
-      \centering 
-\begin{tabular}{lrr}
-\toprule
-Classification & Match & Non-Match\\
-\midrule
-Match & 41 & 9\\
-Non-Match & 2 & 48\\
-\bottomrule
-\end{tabular} \end{minipage}%
-    \begin{minipage}{.5\linewidth}
-      \centering
-        \caption{Hadler et. al. (2017)} 
-\begin{tabular}{lrr}
-\toprule
-Classification & Match & Non-Match\\
-\midrule
-Match & 47 & 3\\
-Non-Match & 0 & 50\\
-\bottomrule
-\end{tabular} \end{minipage} 
-\end{table}
 

### 2018-03-03 doc/paper.Rmd
introduction, error rates

@@
 Same source analyses are a major part of an Forensic Toolmark Examiner's job.
-In current practice examiners  make these comparisons visually under a comparison microscope and come to one of the following four conclusions: identification, inconclusive, elimination or unsuitable for examination \citet{afte-website-2}.
+In current practice examiners  make these comparisons visually under a comparison microscope and come to one of the following four conclusions: identification, inconclusive, elimination or unsuitable for examination \citep{afte-toolmarks1998}.
 These conclusions are made on the basis of "unique surface contours" of the two toolmarks being in "sufficient agreement" \citep{afte-toolmarks1998}.
-AFTE describes the term "sufficient agreement" as the possibility of another tool producing the markings under comparison, as practically impossible \citep{afte-toolmarks1998} \citep{afte-website}.
-%\gk{Although, they go on to concede that the interpretation of the identification depends on the examiner and is  therefore subjective in nature.} 
+AFTE describes the term "sufficient agreement" as the possibility of another tool producing the markings under comparison, as practically impossible \citep{afte-toolmarks1998}.
 This subjectivity in the assessment as well as the lack of error rates are the main points of criticisms in the toolmark examination first raised by the National Research Council in 2009 \citep{NAS:2009} and later emphasized further by the President's Council of Advisors on Science and Technology \citep{pcast2016}.
@@
 The Chumbley scoring method, first introduced by \citet{chumbley}, uses a similar but more extensive framework based on a Mann-Whitney U test of the resulting correlation coefficients.
-This approach was later improved by Hadler and Morris \citep{hadler}.
-\hh{XXX describe the improvement in half a sentence XXX} In this paper, we are investigating the applicability of the Chumbley scoring method by \citet{hadler} to assess striation marks on bullet lands for same-source identification.
+This approach was later improved by Hadler and Morris \citep{hadler} where they introduce a deterministic approach and address the issue of lack of independece between segements of striae.
+In this paper, we are investigating the applicability of the Chumbley scoring method by \citet{hadler} to assess striation marks on bullet lands for same-source identification.
 
@@
 Bullets also have a curved cross-sectional topography \hh{sketch?}.
+Figure \ref{fig:rgl} shows us how the signature from a bullet land lines up with the image of the land from which it was extracted.
+We can see also see in the figure how the depth and relative position of the striation markings seen in the image are interpreted as the signature.
 
@@
 For single feature based matching \citet{chu2013} used an automatic method for counting consequtive matching striae (CMS).
-They reported zero false positive rate but do not discuss about the false negative rates for land to land matches.
+They reported an error rate 52% of the known same source lands comparisons as misidentified and zero false positives for known different source lands.
 \citet{ma2004} and \citet{vorburger2011} discuss about CCF (cross-correlation function) and its discriminating power and application but do not talk about matches and non-matches.
 \citet{aoas} use multiple attributes like CCF, CMS, D (distance measure) etc in a random forrest based method and compare every land against every other land of the Hamby dataset \citet{hamby}.
-They report an out-of-bag error rate with  false positive error rate of 0.0001 and false negative error rate of 0.2267 for 100 trees.
-For 300 trees they report correct prediction of all matches and non-matches.
-
-The Chumbley score  provides us with another approach in the same-source assessment of bullet striation marks.
-\citet{chumbley}, in their paper, compare two toolmarks with the intention of determining if it comes from the same tool.
+They report an out-of-bag error rate of 0.0046.
+
+The Chumbley score  provides us with another approach in the same-source assessment of bullet striation marks.\citet{chumbley}, in their paper, compare two toolmarks with the intention of determining if it comes from the same tool.
 The data for their study was obtained from 50 sequentially manufactured screwdriver tips.\citet{chumbley} reported error rates for marking made by the tips at differnt angles.
-For a 30 degree they showed false negative error rate to be 0.023 and  false positive error rate be 0.09.
+For a 30 degree they report false negative error rate to be 0.023 and  false positive error rate be 0.09.
 For other angles the error rates remain the same for false negatives and 0.01 for false positives.
 The improvement proposed by \citet{hadler} on the other hand, did not consider effect of angles and gave 0.06 false  negative error rate and 0 false poisitive error rate for markings made at the same angle.
-
-\gk{\citep{afte-website} quotes sufficient agreement as "agreement of individual characteristics is of a quantity and quality that the likelihood another tool could have made the mark is so remote as to be considered a practical impossibility"}
-\gk{could not find the pdf for AFTE theory of identification as it relates to toolmarks pdf.
-I believe it is not available for free.
-But from the AFTE Journal the abstract was available which mentions the term " sufficient agreement" as quoted above.
-Currently the website seems like the most reliable way of quoting AFTE, i have checked other sources in both occasions where the website has been quoated above, like an FBI paper available at the FBI website \citet{fbi-paper} and some non published but available pdf files in the AFTE website that have used these lines and quoted the AFTE journal.
-But since I cannot verify it in the Journal I have quoated the website.}
-\gk{faden et al have not cited any paper that discusses the algorithm, i assume it was developed for the paper, the discussion of which is not as extensive as chumbley's.
-Its a variation of the chumbley method but from my understanding it is just the optimization step.
-and this paper came out much before the chumbley score.}
-
-<!-- #### -->
-
-\hh{XXX first sentence: Same source analyses are a major part of an Forensic Toolmark examiners job.
-XXX second sentence: describe standard practice: In current practice firearm examiners  make these comparisons visually under a comparison microscope and report XXX what is being reported: AFTE.
-XXX find: AFTE theory of identification as it relates to toolmarks  }
-\hh{One of the main criticisms in this toolmark examination first raised by the National Research Council in 2009 [@NAS:2009] and later emphasized further by the President's Council of Advisors on Science and Technology [@pcast2016] are the amount of subjectivity in the assessment and the lack of error rates.
-}
-\hh{XXX we have digitized images, e.g. profilometers or Confocal microscopy allow us to measure the surface [@vorburger2016]}
-
-\hh{Using a statistical approach removes both subjectivity from the assessment and allows us to provide error rates for both false positive and false negative identifications. }
-\hh{Multiple toolmarks have been studied in the literature:  \citep{manytoolmarks1, manytoolmarks2, chumbley} have studied screwdriver marks digitized using a profilometer,  \citep{manytoolmarks2} investigate groove pliers XXX data?
-XXX  to slip-joint pliers \citep{afte-chumbley} XXX.
-In all of these cases, the statistical test used, is a variation of the Chumbley scoring method, first introduced by XXX and later improved by XXX.
-}
-
-\hh{XXX In this paper, we are investigating the applicability of the Chumbley scoring method by @hadler to assessing striation marks on bullets for same-source identification. }
-
-\hh{Striation marks on bullets are made by impurities in the barrel.
-As the bullet travels through the barrel, these imperfections leave "scratches" on the bullet surface.
-Typically, only striation marks in the land engraved areas (LEAs) are considered (XXX citation) [@afte-article1992].
-XXX explain lands and grooves XXX.
-Compared to toolmarks made by screwdrivers striation marks on bullets are typically much smaller, both in length and in width.
-Bullets also have a curved cross-sectional topography.
-}
-
-\hh{XXX How are bullets matched now? \citet{chu2013}: CMS, automatically - error rate?, using ccf XXX Vorburger paper XXX  and \citet{aoas}: multiple features, error rates }
-
-\hh{The Chumbley score  provides us with another approach in the same-source assessment of bullet striation marks. }
 

### 2018-03-04 doc/paper.Rmd
changes to structure and information.Toolmark resolution length of markings included.

@@
 Figure \ref{fig:rgl} shows us how the signature from a bullet land lines up with the image of the land from which it was extracted.
-We can see also see in the figure how the depth and relative position of the striation markings seen in the image are interpreted as the signature.
+We can also see in the figure how the depth and relative position of the striation markings seen in the image are interpreted as the signature.
 
@@
 
-\hh{end of intro: remaining paper is structured as follows: introduce to the data we get from confocal microscopy, introduce to profiles and signatures.
-Introduce to chumbley score method, apply chumbley and discuss results ....
-}
-
 ## Scans for land engraved areas
 
-- scans available: NIST database (citation), Hamby 44 and Hamby 252 (Hamby citation)
-- move figure up
-- discuss cross section, profiles and signatures
-
-Compairing pairs of toolmarks with the intention of matching it to a tool has been studied many times in the past.
-Extensive examples can be found in literature for tools and toolmark research ranging from screwdrivers \citep{manytoolmarks1, manytoolmarks2, chumbley} to groove pliers \citep{manytoolmarks2} to slip-joint pliers \citep{afte-chumbley} and many more.
-In comparison to this, same source matching of bullets to firearms has not been examined as prominently as that of toolmarks.
-Even less information is available on validity of methods and error rates associated with firearms examination.
-The National Academy of Sciences in its report in 2009 \citep{NAS:2009} discussed the need for determining error rates in methods proposed for firearms examination.
-As seen in the case of most forensic applications, the first step for same source matching and error rate determination involves identification of unique features that are characteristic of the object at hand.
-For the case of bullets and firearms, striation marks on the surface of the bullet are considered to be such markings that can be used in methods for same source matching.
-These marks are often a product of rifling and impurities and defects due to manufacturing in the barrel of the gun, which leads to engravings on the bullet surface \citep{afte-article1992}.
-In current practice, firearm examiners invariably make visual comparisons of bullet striae and use visual assesment tools to dignify bullets as being matches and non-matches.
-One way of accomplishing anykind of comparison between bullets is to do a comparison between surface marking of two or more bullet lands.
-Bullet Lands are considered to be areas between grooves made by the rifling action of the barrel and the markings on them are considered to be unique.
-The land engraved markings or sometimes termed as Bullet profiles \citep{aoas,ma2004} are striation marks made on Bullet lands and often used for these land to land comparisons.
-Bullet Signatures is another word used in literature as seen in the work of \citet{chu2013} and \citet{aoas}.
-In our context bullet signatures refer to a processed version of the raw land engraved markings or profiles.
-The generation of bullet signatures involves first extraction of a bullet profile by taking the cross-sectional of the surface at a given height and then using loess fits to model the structure.
-The residuals of this fit are called signatures, which are considered to be noise free and a good reflection of the class charecteristics and unique features of a bullet.
+The land engraved markings across the cross section of the land are termed as bullet profiles \citep{aoas,ma2004}.
+These are often used for making land to land comparisons.
+Bullet signatures on the other hand \citep{chu2013,aoas} refer to a processed version of the raw land engraved markings or profiles.
+The generation of bullet signatures involves first extraction of a bullet profile by taking the cross-sectional of the surface at a given height.
+Loess fits are then used to model the structure.
+The residuals of these fits are called signatures.
+They are considered to be noise free and a good reflection of the key attributes of the raw marking, and the unique features of a bullet.
 A more detailed version of the extraction technique of signatures is discussed by \citet{aoas}, where comprehensive details about the height at which profile is to be selected, removing curvature, smoothing, identifying groove locations are explained.
-Figure \ref{fig:rgl} shows us how the extracted signature from a bullet land lines up with the image of the land from which it was extracted.
-We can see also see in the figure how the depth and relative position of the striation markings seen in the image are interpreted as the signature.
-
-As mentioned earlier, subject bias and error rate determination have been a long standing issue in firearm examination \citep{NAS:2009}.
-This issue of removing subjectivity and making firearm analysis objective  was also raised in the \citet{pcast2016} report.
-For an objective analysis, scientific principle testability and error rate determination are considered to be defining aspects.
-Identification of the uncertainities associated with a method of comparison and coming up with a systemized method of quantification, is therefore important.
-Associating mathematical probabilites in a systematic manner to these methods are one such way of quantification.
-In a previous study conducted by \citet{aoas} a machine learning based algorithm was developed for same source matching of bullets and error rates were discussed using the database from the Hamby Study \citep{hamby}.
-The method proposed by \citet{chumbley} (and later improved by \citet{hadler}) also provides a means to determine error rates and claims to reduce subject bias.
-Therefore, giving us a strong motivation to explore the adaptibility of the Chumbley score methodology to bullets.
-In this paper, we take this toolmark based methodology and try to apply and adapt it to bullets.
-After this, we consequently discuss about the efforts in doing so, along with the associated error rates.
-The data used in this paper also belongs to the Hamby Study \citep{hamby}.
-This gives us a common platform for comparing the performance of the chumbley method on bullets with the already existing method proposed by \citet{aoas} for bullets.
-As a brief overview, \citet{chumbley}, in their paper, compare two toolmarks with the intention of determining if it comes from the same tool.
-They use an empirical based setup to validate their proposed algorithm and quantitative method which calculates a U-statisic for the purpose of classification of toolmarks as matching or non-matching.
-The data for their study was obtained from 50 sequentially manufactured screwdriver tips, and preselected comparison window sizes were given as inputs to the algorithm.
-The algorithm then compares the two toolmarks and comes up with a Mann-Whitney U-statistic and an associated p-value to designate them as matches or non-matches.
-The performance for every 100 comparisons, of the algorithm proposed by \citet{chumbley} and the improvement proposed by \citet{hadler} are listed in the table below.
-
-```{r, results='asis', include=FALSE}
-library(knitr)
-library(kableExtra)
-library(xtable)
-
-a1<- matrix(NA, ncol = 3, nrow = 2)
-toolmark1.confusion<- data.frame(a1)
-colnames(toolmark1.confusion)<- c("Classification", "Match", "Non-Match")
-toolmark1.confusion$Classification<- c("Match", "Non-Match")
-# toolmark1.confusion$Match<- c(41,2)
-# toolmark1.confusion$`Non-Match`<- c(9,48)
-toolmark1.confusion$Match<- c(91,2)
-toolmark1.confusion$`Non-Match`<- c(9,98)
-
-a2<- matrix(NA, ncol = 3, nrow = 2)
-toolmark2.confusion<- data.frame(a2)
-colnames(toolmark2.confusion)<- c("Classification", "Match", "Non-Match")
-toolmark2.confusion$Classification<- c("Match", "Non-Match")
-toolmark2.confusion$Match<- c(47,0)
-toolmark2.confusion$`Non-Match`<- c(3,50)
-
-chumbley.confusion<-knitr::kable(toolmark1.confusion, digits=0, format = "latex" ,booktabs=TRUE) #%>% kableExtra::kable_styling(latex_options = c("hold"))
-hadler.confusion<-knitr::kable(toolmark2.confusion, digits=0, format = "latex" ,booktabs=TRUE) #%>% kableExtra::kable_styling(latex_options = c("hold"))
-
-cat(c("\\begin{table}[!htb]
-    \\begin{minipage}{.5\\linewidth}
-      \\caption{Chumbley et al. 2010}
-      \\centering",
-        chumbley.confusion,
-    "\\end{minipage}%
-    \\begin{minipage}{.5\\linewidth}
-      \\centering
-        \\caption{Hadler et. al. (2017)}",
-       hadler.confusion,
-    "\\end{minipage} 
-\\end{table}"
-))
-
-```
-<!-- Striation marks are used to determine same source for bullets.
-Current standards ask for a visual inspection of the marks under a comparison microscope by a firearms examiner.
-One of the problems raised by the NAS report in 2009 is the subjectivity involved in this comparison and the need for determining error rates \citet{NAS:2009}.
-The issue of subjectivity of a firearms examiner has been reviewed by many authors where things like scientific principle testability and error rates are considered to be defining aspects for an objective analysis such that without identifying the uncertainities associated with a method of comparison, it is inconclusive to regards an analysis complete.
-This means that with the usual method of visual comparison that is adopted by firearms examiners, lies the problem of lack of any systemized quantification and as such there seems no certain way of associating it with any kind of mathematical probability.
-Since determining error rates has been duly noted as a fundamental problem in forensic science \citet{NAS:2009}, there is a need of methods that address this problem.
-Some of the methods that have been used to estimate error rates in firearm examinations include Automatic cartridge case comparison where same source identification of cartridge cases and associated error rates have been provided by \citet{riva}, although as noted by \citet{aoas}, in this case alignments of striae involves roation of planes, which cannot be generalized for bullets.
-In case of methods that use machine learning algorithms some methods which are bootsrap based methods often tend to give over-estimated error rates \citet{efron}, but there are alternate error estimation techniques as described by \citet{aoas}, \citet{efron} and  \citet{vorburger2016} which give better results.
--->
-
-<!-- In case of methods that do not use machine learning algorithm, matching two striation marks with each other in order to identify if it is from the same source, can also be done using a well defined comparative statistical algorithm.
-The statistical algorithm is therefore expected  to go through a step by step procedure of identifying the class characteristics i.e striation marks for bullets, choosing the striae, and then follow a systematic procedure of comparing two striae with each other on the basis of criteria like cross-correlation (or others).
-Such a methodology would make it possible to determine error rates too and give definite results regarding what proportion of cases would the analysis be successful in identifying a match correctly and in how many cases would it end up being a false positive.
--->
-
-<!-- Identification criteria like consecutively matching striae (CMS) as first seen by Biasotti 1959 \citet{biasotti} and later mentioned in \citet{chu2013} more often than not include subject bias and are error prone and as seen in the work of \citet{miller} the number of CMS may turn out to be high even when the bullet is not fired by the same firearm.
-Parameters like cross-correlation factor as seen in the extensive comparisons made by \citep{aoas} seem to perform much better for matching purposes.
--->
-
-<!-- An important aspect of same source matching in firearms is identification of bullet microtopographies that are unique to be chosen for a match, Bullet profiles and signatures prove to be such features.
-The choice of a bullet signature to uniquely define a bullet depends on finding first, a stable region on a Bullet land which minimal noise and many pronounced striation marks.
-\citet{aoas} -->
-
-<!-- The method employed by \citet{aoas} uses the cross correlation factor as a means to identify the stable region.
-The markings or striae in this region are supposed to have a high CCF with each other.
-Therefore a Bullet profile is chosen from this region by taking the cross section at a given height.
-A loess fit on the bullet profile produces  residuals which are termed as signatures and can be used as a unique identifier that can be used while matching two bullets to a firearm.
--->
-
-<!-- Compairing pairs of toolmarks on the otherhand, with the intention of matching it to a tool has been studied relatively more in the past as compared to bullets, and \citet{chumbley} have described in their paper an algorithm and analytic method that compares two toolmarks and come to the conclusion if they are from the same tool or not.
-The method also determines the error rates, reduces subject bias and designate the two toolmarks as matches or non-matches with respect to a source.
-\citet{chumbley} used an empirical based setup to validate their proposed analytic and quantitative algorithm.
--->
-
-<!-- \citet{chumbley} found that the algorithm gives false-positives in slightly over 2 cases (or 2%) of the time and about 9 false-negatives (or 9%) for every 100 comparisons.
-\citet{hadler} on the other hand, in the improved version of the algorithm first described by \citet{chumbley} found false-positives for 0 cases and 3 false-negatives (or 6%) for 50 comparisons of knwon matches.
--->
-<!-- ```{r} -->
-<!-- a<- matrix(NA, ncol = 3, nrow = 2) -->
-<!-- toolmark.falsepositive<- data.frame(a) -->
-<!-- colnames(toolmark.falsepositive)<- c("Classification", "Match", "Non-Match") -->
-<!-- toolmark.falsepositive$Classification<- c("Match", "Non-Match") -->
-<!-- toolmark.falsepositive$Match<- c(47,0) -->
-<!-- toolmark.falsepositive$`Non-Match`<- c(3,50) -->
-
-<!-- knitr::kable(toolmark.falsepositive, digits=0, format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("hold")) -->
-<!-- ``` -->
-
-## Potential limitations of the Chumbley Score adaptation to bullets
-Bullet striations are different from toolmarks.
-Bullet markings are typically produced by the barrel on the bullets, while toolmarks can be made by tool tips on different surfaces.
-Tools like screw driver tips as used by \citet{chumbley}, produce longer and pronounced markings.
-Bullets on the other hand are much smaller in length, width, and curved in the cross-sectional topography.
-This means the markings made by barrels as opposed to markings made by tools, may have a problem in distinctiveness.
+
 The majority of Bullet profiles and signatures of the Hamby study \citep{hamby} used in this paper are extracted by procedures mentioned by  \citet{aoas}.
+The scans for the Hamby study come from two sources namely, NIST scans \citep{nist} and scans produced at Iowa State University.
+Both scans were generated using confocal microscopes but at different resolutions.
+The NIST database contains scans of the bullet sets Hamby 44 and 252 with the resolution 1.5625 $\mu m$ per pixel.
+The Iowa State scans are available for only the Hamby 44 bullet set and at a resoultion of 0.065 $\mu m$ per pixel.
+The data used by \citet{chumbley, hadler} on the other hand, was produced by a profilometer at a resolution of about 0.73 $\mu m$ per pixel.
+The different resolutions indicate that the length of the digitized markings would be significantly different for each setting.
+
+## Potential Challenges in Chumbley Score Adaptation
+
+Markings made by Screw driver tips \citep{chumbley} are longer and pronounced and about 7 mm in length \citep{manytoolmarks1}, while the groove to groove length of the bullet lands from the Hamby study \citep{hamby} are about 2mm in length.
 The Ruger pistol barrels used in the  Hamby study have a 9mm caliber.
-This means that bullets used in these pistols have land markings of almost $(1/6)^{th}$ the size of screwdriver toolmarks used by \citet{chumbley}.
-Also, striations on bullets are made on their curved surfaces, whereas the algorithm developed by \citet{chumbley} and \citet{hadler} has only been tested for flatter and wider surfaces which have negligible curvature.
-Therefore, using methods proposed for toolmarks may need adaptation in order to give tangible results for bullets.
-Moreover, in order to get flat bullet signatures and remove the curvatures some kind of smoothing needs to be applied as a pre-step wihch needs further investigation as to whether the level of smoothing does effect the working of the algorithm on Bullets.
+This means that bullets used in these pistols have land markings of almost $(1/4)^{th}$ the size of screwdriver toolmarks used by \citet{chumbley, hadler}, therefore raising a problem of distinctiveness.
+One way to investigate this is to evaluate bullet striations at different resolutions.
+This gives us an incentive to use both the NIST and Iowa State scans and investigate the effect of length of the digitized mark on the working of the algorithm.
+Apart from this, the algorithm developed by \citet{hadler} has only been tested for flatter and wider surfaces which have negligible curvature, whereas striations on bullets are made on their curved surfaces.
+Therefore, using methods proposed for toolmarks need adaptation in order to give tangible results for bullets.
+Another aspect of investigation is the performance of the algorithm for different levels of smoothing when signatures are being compared as opposed to when raw markings or profiles are being compared.
 
@@
 The optimization window is selected in the first step called the optimization step where areas/windows of best agreement between the the two markings are found.
-The window of optimization for bullet striations are shorter as bullet signatures are smaller compared to toolmarks .
-The idea is to keep the number of windows of optimization sufficiently large.
-This means we have shorter trace segments that lets us compare smaller segments of one signature to another than those for toolmarks.
-Here, trace segments are the partitions of marking with the length of the partition equal to the size of window of optimization.
-This introduces a problem as, if we go too small in the window of optimization, the unique features of the trace segments are lost and seem similar, while too large sizes vastly reduces the weight of small features that would otherwise uniquely classify a signature and hence identify the region of agreement.
+
+The window of optimization for bullet striations are shorter as bullet signatures in terms of absolute length in mm, are smaller compared to toolmarks.
+Although in terms of digitzed pixels there is no straightforward way of establishing what window size is to be chosen.
+One important way to address this is to keep the number of windows of optimization sufficiently large or not too small.
+This means, if the number of windows are too small we inadvertently use bigger window sizes as we want to compare all segements of one mark with the segments of the other mark.This leads to not giving enough emphasis to smaller segements of the two marking that might be in high agreement.
+In terms of absolute length we end up comparing smaller segments between the two markings while dealing with bullet lands, than those being compared for toolmarks.
+
+For too large window sizes, the weight of small features that would otherwise uniquely classify a signature and hence identify the region of agreement is vastly reduced.
+On the other hand, this introduces another problem as, if we go too small in the window of optimization, the unique features of the trace segments are lost and all seem similar.
 The problem of distinctiveness in the structure of bullet striations may also lead to potential failure of the algorithm to identify correct windows of agreements or not be able to identify these windows at all.
-This may lead to incorrect classification of bullets as matches or non-matches.
+This may lead to incorrect classification of bullets as coming from the same-source (or matches) or from different source (or non-matches).
 
@@
 An analysis of these error rates and comparison with other methods will help us understand the adaptibility of the chumbley score to bullets.
-
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{images/B6-B2-L6-rescaled.png}
-```{r, eval=FALSE, echo=FALSE, fig.width=8, fig.height=3, out.width='\\textwidth'}
-land <- read.csv("../data/b6b2l6.csv")
-land %>% ggplot(aes(x = y, y=resid)) +geom_line() +theme_bw() +
-  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
-  xlab(expression(paste("Relative Location (in ",mu,m,")", sep="")))
-```
-
-\caption{\label{fig:rgl} Image of a bullet land from a confocal light microscope at 20 fold magnification (top) and a chart of the corresponding signature of the same land (bottom).
-The dotted lines connect some peaks visible in both visualizations.}
-
-\end{figure}
 
@@
 
+##################
+
+#### endedit
+
+\hh{end of intro: remaining paper is structured as follows: introduce to the data we get from confocal microscopy, introduce to profiles and signatures.
+Introduce to chumbley score method, apply chumbley and discuss results ....
+}
+
+## Scans for land engraved areas
+
+- scans available: NIST database (citation), Hamby 44 and Hamby 252 (Hamby citation)
+- move figure up
+- discuss cross section, profiles and signatures
+
+Compairing pairs of toolmarks with the intention of matching it to a tool has been studied many times in the past.
+Extensive examples can be found in literature for tools and toolmark research ranging from screwdrivers \citep{manytoolmarks1, manytoolmarks2, chumbley} to groove pliers \citep{manytoolmarks2} to slip-joint pliers \citep{afte-chumbley} and many more.
+In comparison to this, same source matching of bullets to firearms has not been examined as prominently as that of toolmarks.
+Even less information is available on validity of methods and error rates associated with firearms examination.
+The National Academy of Sciences in its report in 2009 \citep{NAS:2009} discussed the need for determining error rates in methods proposed for firearms examination.
+As seen in the case of most forensic applications, the first step for same source matching and error rate determination involves identification of unique features that are characteristic of the object at hand.
+For the case of bullets and firearms, striation marks on the surface of the bullet are considered to be such markings that can be used in methods for same source matching.
+These marks are often a product of rifling and impurities and defects due to manufacturing in the barrel of the gun, which leads to engravings on the bullet surface \citep{afte-article1992}.
+In current practice, firearm examiners invariably make visual comparisons of bullet striae and use visual assesment tools to dignify bullets as being matches and non-matches.
+One way of accomplishing anykind of comparison between bullets is to do a comparison between surface marking of two or more bullet lands.
+Bullet Lands are considered to be areas between grooves made by the rifling action of the barrel and the markings on them are considered to be unique.
+The land engraved markings or sometimes termed as Bullet profiles \citep{aoas,ma2004} are striation marks made on Bullet lands and often used for these land to land comparisons.
+Bullet Signatures is another word used in literature as seen in the work of \citet{chu2013} and \citet{aoas}.
+In our context bullet signatures refer to a processed version of the raw land engraved markings or profiles.
+The generation of bullet signatures involves first extraction of a bullet profile by taking the cross-sectional of the surface at a given height and then using loess fits to model the structure.
+The residuals of this fit are called signatures, which are considered to be noise free and a good reflection of the class charecteristics and unique features of a bullet.
+A more detailed version of the extraction technique of signatures is discussed by \citet{aoas}, where comprehensive details about the height at which profile is to be selected, removing curvature, smoothing, identifying groove locations are explained.
+Figure \ref{fig:rgl} shows us how the extracted signature from a bullet land lines up with the image of the land from which it was extracted.
+We can see also see in the figure how the depth and relative position of the striation markings seen in the image are interpreted as the signature.
+
+As mentioned earlier, subject bias and error rate determination have been a long standing issue in firearm examination \citep{NAS:2009}.
+This issue of removing subjectivity and making firearm analysis objective  was also raised in the \citet{pcast2016} report.
+For an objective analysis, scientific principle testability and error rate determination are considered to be defining aspects.
+Identification of the uncertainities associated with a method of comparison and coming up with a systemized method of quantification, is therefore important.
+Associating mathematical probabilites in a systematic manner to these methods are one such way of quantification.
+In a previous study conducted by \citet{aoas} a machine learning based algorithm was developed for same source matching of bullets and error rates were discussed using the database from the Hamby Study \citep{hamby}.
+The method proposed by \citet{chumbley} (and later improved by \citet{hadler}) also provides a means to determine error rates and claims to reduce subject bias.
+Therefore, giving us a strong motivation to explore the adaptibility of the Chumbley score methodology to bullets.
+In this paper, we take this toolmark based methodology and try to apply and adapt it to bullets.
+After this, we consequently discuss about the efforts in doing so, along with the associated error rates.
+The data used in this paper also belongs to the Hamby Study \citep{hamby}.
+This gives us a common platform for comparing the performance of the chumbley method on bullets with the already existing method proposed by \citet{aoas} for bullets.
+As a brief overview, \citet{chumbley}, in their paper, compare two toolmarks with the intention of determining if it comes from the same tool.
+They use an empirical based setup to validate their proposed algorithm and quantitative method which calculates a U-statisic for the purpose of classification of toolmarks as matching or non-matching.
+The data for their study was obtained from 50 sequentially manufactured screwdriver tips, and preselected comparison window sizes were given as inputs to the algorithm.
+The algorithm then compares the two toolmarks and comes up with a Mann-Whitney U-statistic and an associated p-value to designate them as matches or non-matches.
+The performance for every 100 comparisons, of the algorithm proposed by \citet{chumbley} and the improvement proposed by \citet{hadler} are listed in the table below.
+
+```{r, results='asis', include=FALSE}
+library(knitr)
+library(kableExtra)
+library(xtable)
+
+a1<- matrix(NA, ncol = 3, nrow = 2)
+toolmark1.confusion<- data.frame(a1)
+colnames(toolmark1.confusion)<- c("Classification", "Match", "Non-Match")
+toolmark1.confusion$Classification<- c("Match", "Non-Match")
+# toolmark1.confusion$Match<- c(41,2)
+# toolmark1.confusion$`Non-Match`<- c(9,48)
+toolmark1.confusion$Match<- c(91,2)
+toolmark1.confusion$`Non-Match`<- c(9,98)
+
+a2<- matrix(NA, ncol = 3, nrow = 2)
+toolmark2.confusion<- data.frame(a2)
+colnames(toolmark2.confusion)<- c("Classification", "Match", "Non-Match")
+toolmark2.confusion$Classification<- c("Match", "Non-Match")
+toolmark2.confusion$Match<- c(47,0)
+toolmark2.confusion$`Non-Match`<- c(3,50)
+
+chumbley.confusion<-knitr::kable(toolmark1.confusion, digits=0, format = "latex" ,booktabs=TRUE) #%>% kableExtra::kable_styling(latex_options = c("hold"))
+hadler.confusion<-knitr::kable(toolmark2.confusion, digits=0, format = "latex" ,booktabs=TRUE) #%>% kableExtra::kable_styling(latex_options = c("hold"))
+
+cat(c("\\begin{table}[!htb]
+    \\begin{minipage}{.5\\linewidth}
+      \\caption{Chumbley et al. 2010}
+      \\centering",
+        chumbley.confusion,
+    "\\end{minipage}%
+    \\begin{minipage}{.5\\linewidth}
+      \\centering
+        \\caption{Hadler et. al. (2017)}",
+       hadler.confusion,
+    "\\end{minipage} 
+\\end{table}"
+))
+
+```
+<!-- Striation marks are used to determine same source for bullets.
+Current standards ask for a visual inspection of the marks under a comparison microscope by a firearms examiner.
+One of the problems raised by the NAS report in 2009 is the subjectivity involved in this comparison and the need for determining error rates \citet{NAS:2009}.
+The issue of subjectivity of a firearms examiner has been reviewed by many authors where things like scientific principle testability and error rates are considered to be defining aspects for an objective analysis such that without identifying the uncertainities associated with a method of comparison, it is inconclusive to regards an analysis complete.
+This means that with the usual method of visual comparison that is adopted by firearms examiners, lies the problem of lack of any systemized quantification and as such there seems no certain way of associating it with any kind of mathematical probability.
+Since determining error rates has been duly noted as a fundamental problem in forensic science \citet{NAS:2009}, there is a need of methods that address this problem.
+Some of the methods that have been used to estimate error rates in firearm examinations include Automatic cartridge case comparison where same source identification of cartridge cases and associated error rates have been provided by \citet{riva}, although as noted by \citet{aoas}, in this case alignments of striae involves roation of planes, which cannot be generalized for bullets.
+In case of methods that use machine learning algorithms some methods which are bootsrap based methods often tend to give over-estimated error rates \citet{efron}, but there are alternate error estimation techniques as described by \citet{aoas}, \citet{efron} and  \citet{vorburger2016} which give better results.
+-->
+
+<!-- In case of methods that do not use machine learning algorithm, matching two striation marks with each other in order to identify if it is from the same source, can also be done using a well defined comparative statistical algorithm.
+The statistical algorithm is therefore expected  to go through a step by step procedure of identifying the class characteristics i.e striation marks for bullets, choosing the striae, and then follow a systematic procedure of comparing two striae with each other on the basis of criteria like cross-correlation (or others).
+Such a methodology would make it possible to determine error rates too and give definite results regarding what proportion of cases would the analysis be successful in identifying a match correctly and in how many cases would it end up being a false positive.
+-->
+
+<!-- Identification criteria like consecutively matching striae (CMS) as first seen by Biasotti 1959 \citet{biasotti} and later mentioned in \citet{chu2013} more often than not include subject bias and are error prone and as seen in the work of \citet{miller} the number of CMS may turn out to be high even when the bullet is not fired by the same firearm.
+Parameters like cross-correlation factor as seen in the extensive comparisons made by \citep{aoas} seem to perform much better for matching purposes.
+-->
+
+<!-- An important aspect of same source matching in firearms is identification of bullet microtopographies that are unique to be chosen for a match, Bullet profiles and signatures prove to be such features.
+The choice of a bullet signature to uniquely define a bullet depends on finding first, a stable region on a Bullet land which minimal noise and many pronounced striation marks.
+\citet{aoas} -->
+
+<!-- The method employed by \citet{aoas} uses the cross correlation factor as a means to identify the stable region.
+The markings or striae in this region are supposed to have a high CCF with each other.
+Therefore a Bullet profile is chosen from this region by taking the cross section at a given height.
+A loess fit on the bullet profile produces  residuals which are termed as signatures and can be used as a unique identifier that can be used while matching two bullets to a firearm.
+-->
+
+<!-- Compairing pairs of toolmarks on the otherhand, with the intention of matching it to a tool has been studied relatively more in the past as compared to bullets, and \citet{chumbley} have described in their paper an algorithm and analytic method that compares two toolmarks and come to the conclusion if they are from the same tool or not.
+The method also determines the error rates, reduces subject bias and designate the two toolmarks as matches or non-matches with respect to a source.
+\citet{chumbley} used an empirical based setup to validate their proposed analytic and quantitative algorithm.
+-->
+
+<!-- \citet{chumbley} found that the algorithm gives false-positives in slightly over 2 cases (or 2%) of the time and about 9 false-negatives (or 9%) for every 100 comparisons.
+\citet{hadler} on the other hand, in the improved version of the algorithm first described by \citet{chumbley} found false-positives for 0 cases and 3 false-negatives (or 6%) for 50 comparisons of knwon matches.
+-->
+<!-- ```{r} -->
+<!-- a<- matrix(NA, ncol = 3, nrow = 2) -->
+<!-- toolmark.falsepositive<- data.frame(a) -->
+<!-- colnames(toolmark.falsepositive)<- c("Classification", "Match", "Non-Match") -->
+<!-- toolmark.falsepositive$Classification<- c("Match", "Non-Match") -->
+<!-- toolmark.falsepositive$Match<- c(47,0) -->
+<!-- toolmark.falsepositive$`Non-Match`<- c(3,50) -->
+
+<!-- knitr::kable(toolmark.falsepositive, digits=0, format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("hold")) -->
+<!-- ``` -->
+
+## Potential limitations of the Chumbley Score adaptation to bullets
+Bullet striations are different from toolmarks.
+Bullet markings are typically produced by the barrel on the bullets, while toolmarks can be made by tool tips on different surfaces.
+Tools like screw driver tips as used by \citet{chumbley}, produce longer and pronounced markings.
+Bullets on the other hand are much smaller in length, width, and curved in the cross-sectional topography.
+This means the markings made by barrels as opposed to markings made by tools, may have a problem in distinctiveness.
+The majority of Bullet profiles and signatures of the Hamby study \citep{hamby} used in this paper are extracted by procedures mentioned by  \citet{aoas}.
+The Ruger pistol barrels used in the  Hamby study have a 9mm caliber.
+This means that bullets used in these pistols have land markings of almost $(1/6)^{th}$ the size of screwdriver toolmarks used by \citet{chumbley}.
+Also, striations on bullets are made on their curved surfaces, whereas the algorithm developed by \citet{chumbley} and \citet{hadler} has only been tested for flatter and wider surfaces which have negligible curvature.
+Therefore, using methods proposed for toolmarks may need adaptation in order to give tangible results for bullets.
+Moreover, in order to get flat bullet signatures and remove the curvatures some kind of smoothing needs to be applied as a pre-step wihch needs further investigation as to whether the level of smoothing does effect the working of the algorithm on Bullets.
+
+The Chumbley method and algorithm works on the premise of comparison windows which are supposed to represent small segments of pixels of equal sizes in the two markings that are to be compared in a predetermined fashion.
+They term them as windows of optimization and validation.
+The optimization window is selected in the first step called the optimization step where areas/windows of best agreement between the the two markings are found.
+The window of optimization for bullet striations are shorter as bullet signatures are smaller compared to toolmarks .
+The idea is to keep the number of windows of optimization sufficiently large.
+This means we have shorter trace segments that lets us compare smaller segments of one signature to another than those for toolmarks.
+Here, trace segments are the partitions of marking with the length of the partition equal to the size of window of optimization.
+This introduces a problem as, if we go too small in the window of optimization, the unique features of the trace segments are lost and seem similar, while too large sizes vastly reduces the weight of small features that would otherwise uniquely classify a signature and hence identify the region of agreement.
+The problem of distinctiveness in the structure of bullet striations may also lead to potential failure of the algorithm to identify correct windows of agreements or not be able to identify these windows at all.
+This may lead to incorrect classification of bullets as matches or non-matches.
+
+Thus, the comaparison windows have a direct influence on the false positives and false negatives, making identification of the optimum sizes of these windows very important.
+This raises important questions about how and what parameter settings need to be chosen for bullet land comparison.
+Something similar is done in \citet{afte-chumbley} for toolmark comparisons of slip-joint pliers where optimum window sizes are determined.
+There is a need to figure out the best parameter settings which minimizes the errors for unsmoothed markings or profiles and pre-processed signatures.
+An analysis of these error rates and comparison with other methods will help us understand the adaptibility of the chumbley score to bullets.
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{images/B6-B2-L6-rescaled.png}
+```{r, eval=FALSE, echo=FALSE, fig.width=8, fig.height=3, out.width='\\textwidth'}
+land <- read.csv("../data/b6b2l6.csv")
+land %>% ggplot(aes(x = y, y=resid)) +geom_line() +theme_bw() +
+  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
+  xlab(expression(paste("Relative Location (in ",mu,m,")", sep="")))
+```
+
+\caption{\label{fig:rgl} Image of a bullet land from a confocal light microscope at 20 fold magnification (top) and a chart of the corresponding signature of the same land (bottom).
+The dotted lines connect some peaks visible in both visualizations.}
+
+\end{figure}
+
+## The Chumbley Score Test
+
+The Chumbley score algorithm takes input as two vectorized processes.
+Here the processes are of the form $z(t)$ which is a spatial process for some location indexed with $t$.
+This means that while $z(t)$ represents any striae, $z(t_1)$ is a realization of the spatial process.
+$t$ here denotes a vector of equally spaced pixel locations for the striation marks under consideration.
+The word 'pixel' refers to the resolution of the confocal light microscope.
+In the case for bullet signatures extracted from NIST ballistics database \citep{nist} of the Hamby study, a pixel corresponds to approximately 1.56 microns.
+The vectorized processes representing two sets of striaes are therefore shown as $x(t_1)$, $t_1 = 1,2,...T_1$ and $y(t_2)$, $t_2 = 1,2...T_2$.
+Here $x$ and $y$ denote two instances of the process $z(*)$ which means two striation marks, while the indexing $t$ or pixel range is shown as $t_1$ and $t_2$ for $x$ and $y$ respectively.
+This representation of $t$ as $t_1$ and $t_2$ lets us identify them as either of same or different lengths.
+The striation marks under consideration are potentially from two different bullets whose source needs to be identified as being same or different.
+$T_1$ and $T_2$, as represented above, are the final pixel indexes of each marking and therefore give the respective lengths of the markings.
+The similarity in the two markings is then judged by the algorithm on the basis of cross-correlation of a fixed and constant number consecutive pixels (say $k$) taken from the two markings.
+This can for example be $k$ taken from the two indexed marking $x(t_1)$ and $y(t_2)$ such that in theory $k$ remains smaller than length of the two striae or marks.
+Depending on what stage of the algorithm we are in, matching of different pixel lengths and locations is done.
+This in the end, effectively compares all possible windows that would guarantee in quantifying the two marks or striae as coming from the same source or not.
+
+The algorithm works in two phases, namely, an optimization step and a validation step, at the end of which a Mann Whitney U statistic is calculated as a measure to differentiate between matches and non-matches.
+A pre-processing step to the algorithm is to choose a coarseness value which  is used as a parameter to the lowess smoothing function.
+The coarseness essentially gives the proportion of points which influence the smooth at each value, which means larger values lead to more smoothness.
+The lowess smoothing is applied to each of two sets of vectorized striae or marks $x(t_1)$ and $y(t_2)$, before proceeding to the algorithm.
+The algorithm starts with the optimization step where the area of best agreement in the two markings being compared is identified.
+Comparison window size is predefined.
+Each window of the first marking is compared with all windows of the second marking.
+A maximum correlation statistic is used to identify the region of best agreement, with the maximum usually seen being near 1 for both cases which is what is intuitively expected for matches, and not expected intuitively for non matches.
+\citet{hadler} in their paper proposed an improvement to this algorithm by trying to remove mutual dependence of parameters.
+The mututal dependence was due to the serial correlation in surface depth values of a marking.
+Also a random sampling sub step in the validation phase makes a group of pixels to be chosen more than once and hence introduces lack of independence in certain steps.
+The reason for removing the mutual dependence was the Mann-Whitney U statistic works under the assumption of independence of parameters.
+The validation step builds on this with two sub-steps namely, same shift and different shift.
+As the aim is to come up with a non-parameteric U statistic, \citet{hadler} proposed a normalization procedure in the validation step which goes to some extent to address the issue of mutual dependence.
+A series of windows are chosen in both the same and different shift sub-steps for the purpose of comparison.
+Both substeps were modified by \citet{hadler} who introduced a deterministic rule for sampling of windows in the same-shift and different shift as opposed to the originally proposed random sampling by \citet{chumbley}.
+In th same shift substep the chosen series of windows are at a common distance (rigid-shifts) from the window earlier identified as the region of best agreement in the optimization step.
+The correlation of these windows generally turn out to be lower than the maximum correlation window.
+The significance is that these same shift windows still have large enough correlation values for the two markings being compared that are in reality a match.
+Any choice of maximum correlation windows (where correlation values are often near 1) for non-matches in the optimization step are also validated by this sub-step.
+This is because the same-shift correlations would not be anywhere as large for same-shift windows when all windows in the sub-step are compared.
+The different shift substep on the other hand gives perspective to the correlation values of the same shift window correlation values.
+There are no rigid-shifts but different shifts where distance from the maximum correlation window are chosen randomly by \citet{chumbley} and deterministically as per \citet{hadler}.
+This means that there is an equal possibilty of comparing a trace segment from one marking to any trace segment or window in the second marking.
+
+Neither of above sets of correlation in the two sub-steps are allowed to include the maximum correlation window as identified earlier.
+Therefore the assumption is that if two markings match each other, the same-shift correlations would be larger than the different-shift windows.
+And if they are not a match the correlations in the two sets will be very similar.
+The U-statistic tests for the null hypothesis that the two markings are not a match and are therefore not made by the same source.
+As given by \citet{hadler}, it is computed from the joint rank of all correlations of both the same and different shift samples.
+
 <!-- The data that is used by \citet{chumbley} is generated by a surface profilometer that gives the height in terms of distance along a linear trace.

### 2018-03-05 doc/paper.Rmd
changed some of the word formations

@@
 Therefore, using methods proposed for toolmarks need adaptation in order to give tangible results for bullets.
-Another aspect of investigation is the performance of the algorithm for different levels of smoothing when signatures are being compared as opposed to when raw markings or profiles are being compared.
-
-The Chumbley method and algorithm works on the premise of comparison windows which are supposed to represent small segments of pixels of equal sizes in the two markings that are to be compared in a predetermined fashion.
-They term them as windows of optimization and validation.
+Another challenge that arises from these topographical differences, is the understanding of that effect of different levels of smoothing.
+Especially investigating its effect on the algorithm performance when signatures are being compared and when raw markings or profiles are being compared.
+
+Some other challenges associated with the adaptation process call for a brief understanding of the methodology involved.
+The Chumbley method and algorithm works on the premise of comparison windows that are supposed to represent small segments of pixels of equal sizes in the two markings.
+These windows are to be compared in a predetermined fashion and are termed as windows of optimization and validation.
 The optimization window is selected in the first step called the optimization step where areas/windows of best agreement between the the two markings are found.
 
-The window of optimization for bullet striations are shorter as bullet signatures in terms of absolute length in mm, are smaller compared to toolmarks.
-Although in terms of digitzed pixels there is no straightforward way of establishing what window size is to be chosen.
+The window of optimization for bullet striations are shorter as bullet signatures in terms of absolute length (in mm), are smaller compared to toolmarks.
+Even in terms of digitized pixels, there is no straightforward way of establishing what window size is to be chosen.
 One important way to address this is to keep the number of windows of optimization sufficiently large or not too small.
-This means, if the number of windows are too small we inadvertently use bigger window sizes as we want to compare all segements of one mark with the segments of the other mark.This leads to not giving enough emphasis to smaller segements of the two marking that might be in high agreement.
+This means, if the number of windows are too small we inadvertently use bigger window sizes as we want to compare all segements of one mark with the segments of the other mark.
+This leads to not giving enough emphasis to smaller segements of the two marking that might be in high agreement.
 In terms of absolute length we end up comparing smaller segments between the two markings while dealing with bullet lands, than those being compared for toolmarks.

### 2018-03-05 doc/paper.Rmd
cleaning

@@
 
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{images/B6-B2-L6-rescaled.png}
+```{r, eval=FALSE, echo=FALSE, fig.width=8, fig.height=3, out.width='\\textwidth'}
+land <- read.csv("../data/b6b2l6.csv")
+land %>% ggplot(aes(x = y, y=resid)) +geom_line() +theme_bw() +
+  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
+  xlab(expression(paste("Relative Location (in ",mu,m,")", sep="")))
+```
+
+\caption{\label{fig:rgl} Image of a bullet land from a confocal light microscope at 20 fold magnification (top) and a chart of the corresponding signature of the same land (bottom).
+The dotted lines connect some peaks visible in both visualizations.}
+
+\end{figure}
+
 ## Scans for land engraved areas
@@
 
-Compairing pairs of toolmarks with the intention of matching it to a tool has been studied many times in the past.
+<!-- Compairing pairs of toolmarks with the intention of matching it to a tool has been studied many times in the past.
 Extensive examples can be found in literature for tools and toolmark research ranging from screwdrivers \citep{manytoolmarks1, manytoolmarks2, chumbley} to groove pliers \citep{manytoolmarks2} to slip-joint pliers \citep{afte-chumbley} and many more.
@@
 We can see also see in the figure how the depth and relative position of the striation markings seen in the image are interpreted as the signature.
-
-As mentioned earlier, subject bias and error rate determination have been a long standing issue in firearm examination \citep{NAS:2009}.
+-->
+
+<!-- As mentioned earlier, subject bias and error rate determination have been a long standing issue in firearm examination \citep{NAS:2009}.
 This issue of removing subjectivity and making firearm analysis objective  was also raised in the \citet{pcast2016} report.
@@
 Associating mathematical probabilites in a systematic manner to these methods are one such way of quantification.
-In a previous study conducted by \citet{aoas} a machine learning based algorithm was developed for same source matching of bullets and error rates were discussed using the database from the Hamby Study \citep{hamby}.
+-->
+<!-- In a previous study conducted by \citet{aoas} a machine learning based algorithm was developed for same source matching of bullets and error rates were discussed using the database from the Hamby Study \citep{hamby}.
 The method proposed by \citet{chumbley} (and later improved by \citet{hadler}) also provides a means to determine error rates and claims to reduce subject bias.
@@
 The performance for every 100 comparisons, of the algorithm proposed by \citet{chumbley} and the improvement proposed by \citet{hadler} are listed in the table below.
-
-```{r, results='asis', include=FALSE}
-library(knitr)
-library(kableExtra)
-library(xtable)
-
-a1<- matrix(NA, ncol = 3, nrow = 2)
-toolmark1.confusion<- data.frame(a1)
-colnames(toolmark1.confusion)<- c("Classification", "Match", "Non-Match")
-toolmark1.confusion$Classification<- c("Match", "Non-Match")
-# toolmark1.confusion$Match<- c(41,2)
-# toolmark1.confusion$`Non-Match`<- c(9,48)
-toolmark1.confusion$Match<- c(91,2)
-toolmark1.confusion$`Non-Match`<- c(9,98)
-
-a2<- matrix(NA, ncol = 3, nrow = 2)
-toolmark2.confusion<- data.frame(a2)
-colnames(toolmark2.confusion)<- c("Classification", "Match", "Non-Match")
-toolmark2.confusion$Classification<- c("Match", "Non-Match")
-toolmark2.confusion$Match<- c(47,0)
-toolmark2.confusion$`Non-Match`<- c(3,50)
-
-chumbley.confusion<-knitr::kable(toolmark1.confusion, digits=0, format = "latex" ,booktabs=TRUE) #%>% kableExtra::kable_styling(latex_options = c("hold"))
-hadler.confusion<-knitr::kable(toolmark2.confusion, digits=0, format = "latex" ,booktabs=TRUE) #%>% kableExtra::kable_styling(latex_options = c("hold"))
-
-cat(c("\\begin{table}[!htb]
-    \\begin{minipage}{.5\\linewidth}
-      \\caption{Chumbley et al. 2010}
-      \\centering",
-        chumbley.confusion,
-    "\\end{minipage}%
-    \\begin{minipage}{.5\\linewidth}
-      \\centering
-        \\caption{Hadler et. al. (2017)}",
-       hadler.confusion,
-    "\\end{minipage} 
-\\end{table}"
-))
-
-```
+-->
+
+<!-- ```{r, results='asis', include=FALSE} -->
+<!-- library(knitr) -->
+<!-- library(kableExtra) -->
+<!-- library(xtable) -->
+
+<!-- a1<- matrix(NA, ncol = 3, nrow = 2) -->
+<!-- toolmark1.confusion<- data.frame(a1) -->
+<!-- colnames(toolmark1.confusion)<- c("Classification", "Match", "Non-Match") -->
+<!-- toolmark1.confusion$Classification<- c("Match", "Non-Match") -->
+<!-- # toolmark1.confusion$Match<- c(41,2) -->
+<!-- # toolmark1.confusion$`Non-Match`<- c(9,48) -->
+<!-- toolmark1.confusion$Match<- c(91,2) -->
+<!-- toolmark1.confusion$`Non-Match`<- c(9,98) -->
+
+<!-- a2<- matrix(NA, ncol = 3, nrow = 2) -->
+<!-- toolmark2.confusion<- data.frame(a2) -->
+<!-- colnames(toolmark2.confusion)<- c("Classification", "Match", "Non-Match") -->
+<!-- toolmark2.confusion$Classification<- c("Match", "Non-Match") -->
+<!-- toolmark2.confusion$Match<- c(47,0) -->
+<!-- toolmark2.confusion$`Non-Match`<- c(3,50) -->
+
+<!-- chumbley.confusion<-knitr::kable(toolmark1.confusion, digits=0, format = "latex" ,booktabs=TRUE) #%>% kableExtra::kable_styling(latex_options = c("hold")) -->
+<!-- hadler.confusion<-knitr::kable(toolmark2.confusion, digits=0, format = "latex" ,booktabs=TRUE) #%>% kableExtra::kable_styling(latex_options = c("hold")) -->
+
+<!-- cat(c("\\begin{table}[!htb] -->
+<!--     \\begin{minipage}{.5\\linewidth} -->
+<!--       \\caption{Chumbley et al. 2010} -->
+<!--       \\centering", -->
+<!--         chumbley.confusion, -->
+<!--     "\\end{minipage}% -->
+<!--     \\begin{minipage}{.5\\linewidth} -->
+<!--       \\centering -->
+<!--         \\caption{Hadler et. al. (2017)}", -->
+<!--        hadler.confusion, -->
+<!--     "\\end{minipage}  -->
+<!-- \\end{table}" -->
+<!-- )) -->
+
+<!-- ``` -->
 <!-- Striation marks are used to determine same source for bullets.
@@
 
-## Potential limitations of the Chumbley Score adaptation to bullets
-Bullet striations are different from toolmarks.
+<!-- ## Potential limitations of the Chumbley Score adaptation to bullets -->
+<!-- Bullet striations are different from toolmarks.
 Bullet markings are typically produced by the barrel on the bullets, while toolmarks can be made by tool tips on different surfaces.
@@
 Moreover, in order to get flat bullet signatures and remove the curvatures some kind of smoothing needs to be applied as a pre-step wihch needs further investigation as to whether the level of smoothing does effect the working of the algorithm on Bullets.
-
-The Chumbley method and algorithm works on the premise of comparison windows which are supposed to represent small segments of pixels of equal sizes in the two markings that are to be compared in a predetermined fashion.
+-->
+
+<!-- The Chumbley method and algorithm works on the premise of comparison windows which are supposed to represent small segments of pixels of equal sizes in the two markings that are to be compared in a predetermined fashion.
 They term them as windows of optimization and validation.
@@
 This may lead to incorrect classification of bullets as matches or non-matches.
-
-Thus, the comaparison windows have a direct influence on the false positives and false negatives, making identification of the optimum sizes of these windows very important.
+-->
+
+<!-- Thus, the comaparison windows have a direct influence on the false positives and false negatives, making identification of the optimum sizes of these windows very important.
 This raises important questions about how and what parameter settings need to be chosen for bullet land comparison.
@@
 An analysis of these error rates and comparison with other methods will help us understand the adaptibility of the chumbley score to bullets.
-
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{images/B6-B2-L6-rescaled.png}
-```{r, eval=FALSE, echo=FALSE, fig.width=8, fig.height=3, out.width='\\textwidth'}
-land <- read.csv("../data/b6b2l6.csv")
-land %>% ggplot(aes(x = y, y=resid)) +geom_line() +theme_bw() +
-  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
-  xlab(expression(paste("Relative Location (in ",mu,m,")", sep="")))
-```
-
-\caption{\label{fig:rgl} Image of a bullet land from a confocal light microscope at 20 fold magnification (top) and a chart of the corresponding signature of the same land (bottom).
-The dotted lines connect some peaks visible in both visualizations.}
-
-\end{figure}
-
-## The Chumbley Score Test
-
-The Chumbley score algorithm takes input as two vectorized processes.
+-->
+
+<!-- ## The Chumbley Score Test -->
+
+<!-- The Chumbley score algorithm takes input as two vectorized processes.
 Here the processes are of the form $z(t)$ which is a spatial process for some location indexed with $t$.
@@
 This in the end, effectively compares all possible windows that would guarantee in quantifying the two marks or striae as coming from the same source or not.
-
-The algorithm works in two phases, namely, an optimization step and a validation step, at the end of which a Mann Whitney U statistic is calculated as a measure to differentiate between matches and non-matches.
+-->
+
+<!-- The algorithm works in two phases, namely, an optimization step and a validation step, at the end of which a Mann Whitney U statistic is calculated as a measure to differentiate between matches and non-matches.
 A pre-processing step to the algorithm is to choose a coarseness value which  is used as a parameter to the lowess smoothing function.
@@
 The lowess smoothing is applied to each of two sets of vectorized striae or marks $x(t_1)$ and $y(t_2)$, before proceeding to the algorithm.
-The algorithm starts with the optimization step where the area of best agreement in the two markings being compared is identified.
+-->
+<!-- The algorithm starts with the optimization step where the area of best agreement in the two markings being compared is identified.
 Comparison window size is predefined.
@@
 A maximum correlation statistic is used to identify the region of best agreement, with the maximum usually seen being near 1 for both cases which is what is intuitively expected for matches, and not expected intuitively for non matches.
-\citet{hadler} in their paper proposed an improvement to this algorithm by trying to remove mutual dependence of parameters.
+-->
+<!-- \citet{hadler} in their paper proposed an improvement to this algorithm by trying to remove mutual dependence of parameters.
 The mututal dependence was due to the serial correlation in surface depth values of a marking.
@@
 The reason for removing the mutual dependence was the Mann-Whitney U statistic works under the assumption of independence of parameters.
-The validation step builds on this with two sub-steps namely, same shift and different shift.
+-->
+<!-- The validation step builds on this with two sub-steps namely, same shift and different shift.
 As the aim is to come up with a non-parameteric U statistic, \citet{hadler} proposed a normalization procedure in the validation step which goes to some extent to address the issue of mutual dependence.
@@
 Both substeps were modified by \citet{hadler} who introduced a deterministic rule for sampling of windows in the same-shift and different shift as opposed to the originally proposed random sampling by \citet{chumbley}.
-In th same shift substep the chosen series of windows are at a common distance (rigid-shifts) from the window earlier identified as the region of best agreement in the optimization step.
+-->
+<!-- In th same shift substep the chosen series of windows are at a common distance (rigid-shifts) from the window earlier identified as the region of best agreement in the optimization step.
 The correlation of these windows generally turn out to be lower than the maximum correlation window.
@@
 This is because the same-shift correlations would not be anywhere as large for same-shift windows when all windows in the sub-step are compared.
-The different shift substep on the other hand gives perspective to the correlation values of the same shift window correlation values.
+-->
+<!-- The different shift substep on the other hand gives perspective to the correlation values of the same shift window correlation values.
 There are no rigid-shifts but different shifts where distance from the maximum correlation window are chosen randomly by \citet{chumbley} and deterministically as per \citet{hadler}.
 This means that there is an equal possibilty of comparing a trace segment from one marking to any trace segment or window in the second marking.
-
-Neither of above sets of correlation in the two sub-steps are allowed to include the maximum correlation window as identified earlier.
+-->
+
+<!-- Neither of above sets of correlation in the two sub-steps are allowed to include the maximum correlation window as identified earlier.
 Therefore the assumption is that if two markings match each other, the same-shift correlations would be larger than the different-shift windows.
-And if they are not a match the correlations in the two sets will be very similar.
+-->
+<!-- And if they are not a match the correlations in the two sets will be very similar.
 The U-statistic tests for the null hypothesis that the two markings are not a match and are therefore not made by the same source.
 As given by \citet{hadler}, it is computed from the joint rank of all correlations of both the same and different shift samples.
+-->
 

### 2018-03-07 doc/paper.Rmd
image relocation
corrected number

@@
 The NIST database contains scans of the bullet sets Hamby 44 and 252 with the resolution 1.5625 $\mu m$ per pixel.
-The Iowa State scans are available for only the Hamby 44 bullet set and at a resoultion of 0.065 $\mu m$ per pixel.
+The Iowa State scans are available for only the Hamby 44 bullet set and at a resoultion of 0.65 $\mu m$ per pixel.
 The data used by \citet{chumbley, hadler} on the other hand, was produced by a profilometer at a resolution of about 0.73 $\mu m$ per pixel.
@@
 $t$ here denotes a vector of equally spaced pixel locations for the striation marks under consideration.
-The word 'pixel' refers to the resolution of the confocal light microscope.
-In the case for bullet signatures extracted from NIST ballistics database \citep{nist} of the Hamby study, a pixel corresponds to approximately 1.56 microns.
 The vectorized processes representing two sets of striaes are therefore shown as $x(t_1)$, $t_1 = 1,2,...T_1$ and $y(t_2)$, $t_2 = 1,2...T_2$.
@@
 This in the end, effectively compares all possible windows that would guarantee in quantifying the two marks or striae as coming from the same source or not.
+
+<!-- The word 'pixel' refers to the resolution of the confocal light microscope.
+In the case for bullet signatures extracted from NIST ballistics database \citep{nist} of the Hamby study, a pixel corresponds to approximately 1.56 microns.
+-->
 

### 2018-03-16 doc/paper.Rmd
couple more citations

@@
 Same source analyses are a major part of an Forensic Toolmark Examiner's job.
-In current practice examiners  make these comparisons visually under a comparison microscope and come to one of the following four conclusions: identification, inconclusive, elimination or unsuitable for examination \citep{afte-toolmarks1998}.
+In current practice examiners  make these comparisons by visual inspection under a comparison microscope and come to one of the following four conclusions: identification, inconclusive, elimination or unsuitable for examination~\citep{afte-toolmarks1998}.
 These conclusions are made on the basis of "unique surface contours" of the two toolmarks being in "sufficient agreement" \citep{afte-toolmarks1998}.
 AFTE describes the term "sufficient agreement" as the possibility of another tool producing the markings under comparison, as practically impossible \citep{afte-toolmarks1998}.
-This subjectivity in the assessment as well as the lack of error rates are the main points of criticisms in the toolmark examination first raised by the National Research Council in 2009 \citep{NAS:2009} and later emphasized further by the President's Council of Advisors on Science and Technology \citep{pcast2016}.
-
-Technological advances, such as profilometers and confocal microscopy allow to measure 3D surfaces \citep{vorburger2016} in a digitized form.
-This forms the basis of statistical analysis of toolmarks.
+This subjectivity in the assessment as well as the lack of error rates are the main points of criticisms  first raised by the National Research Council in 2009 \citep{NAS:2009} and later emphasized further by the President's Council of Advisors on Science and Technology \citep{pcast2016}.
+
+Technological advances, such as profilometers and confocal microscopy allow to measure 3D surfaces in a high-resolution digitized form.
+This technology has become more accessible over the last decade, and has made its way into topological images of ballistics evidence, such as bullet lands and breech faces \citep{DeKinder1, DeKinder2, Bachrach1, vorburger2016}.
+Digitized images of 3D surfaces of form the data basis of statistical analysis of toolmarks.
 A statistical approach based on data  removes both subjectivity from the assessment and allows a quantification of error rates for both false positive and false negative identifications.
 
-Various toolmarks have been studied in the literature:  screwdriver marks  digitized using a profilometer have been analyzed in \citet{manytoolmarks1} and \citet{chumbley};  \citet{manytoolmarks2} have investigated 3D marks from screwdriver, tongue and groove pliers captured using a confocal microscope;  digitized marks from slip-joint pliers generated by a surface profilometer have been investigated by \citet{afte-chumbley}.
-
+\hh{In the next page and a half it is easy to lose the red line. It might help to include a table with an overview.
+The table should include the reference to the paper, the data used, the statistical method and the associated error rates.}
+Various toolmarks have been studied in the literature:  \citet{manytoolmarks1} and \citet{chumbley} have been analyzing screwdriver marks  digitized using a profilometer;  \citet{manytoolmarks2} have investigated 3D marks from screwdriver, tongue and groove pliers captured using a confocal microscope; \citet{afte-chumbley} have been investigated digitized marks from slip-joint pliers generated by a surface profilometer.
+
+\hh{We need an additional sentence here to get from the data to the statistical methods ... }
 \citet{manytoolmarks2} define a relative distance metric and use it as similarity measure between two toolmarks.
@@
 The Chumbley scoring method, first introduced by \citet{chumbley}, uses a similar but more extensive framework based on a Mann-Whitney U test of the resulting correlation coefficients.
-This approach was later improved by Hadler and Morris \citep{hadler} where they introduce a deterministic approach and address the issue of lack of independece between segements of striae.
+This approach was later improved by Hadler and Morris \citep{hadler} where the authors introduce a deterministic approach and address the issue of lack of independence between segements of striae.
 In this paper, we are investigating the applicability of the Chumbley scoring method by \citet{hadler} to assess striation marks on bullet lands for same-source identification.
@@
 Compared to toolmarks made by screwdrivers striation marks on bullets are typically much smaller, both in length and in width.
-Bullets also have a curved cross-sectional topography \hh{sketch?}.
-Figure \ref{fig:rgl} shows us how the signature from a bullet land lines up with the image of the land from which it was extracted.
+Bullets also have a curved cross-sectional topography.
+Figure~\ref{fig:rgl} shows us how the signature from a bullet land (bottom) lines up with the image of the land (top) from which it was extracted.
 We can also see in the figure how the depth and relative position of the striation markings seen in the image are interpreted as the signature.
 
-Bullet matching methods depend on the attribute being used for making comparisons.
-For single feature based matching \citet{chu2013} used an automatic method for counting consequtive matching striae (CMS).
-They reported an error rate 52% of the known same source lands comparisons as misidentified and zero false positives for known different source lands.
-\citet{ma2004} and \citet{vorburger2011} discuss about CCF (cross-correlation function) and its discriminating power and application but do not talk about matches and non-matches.
-\citet{aoas} use multiple attributes like CCF, CMS, D (distance measure) etc in a random forrest based method and compare every land against every other land of the Hamby dataset \citet{hamby}.
-They report an out-of-bag error rate of 0.0046.
-
-The Chumbley score  provides us with another approach in the same-source assessment of bullet striation marks.\citet{chumbley}, in their paper, compare two toolmarks with the intention of determining if it comes from the same tool.
-The data for their study was obtained from 50 sequentially manufactured screwdriver tips.\citet{chumbley} reported error rates for marking made by the tips at differnt angles.
+\hh{Here's a bit of a jump: you're already talking about features, but we haven't yet properly introduced signatures of bullets yet.}
+Bullet matching methods depend on the attributes  used for making comparisons.
+For single feature based matching \citet{chu2013} used an automatic method for counting consecutive matching striae (CMS).
+The authors report an error rate of 52% of the known same source lands comparisons as misidentified (false negative) and zero false positives for known different source lands.
+\citet{ma2004} and \citet{vorburger2011} discuss CCF (cross-correlation function) and its discriminating power and applicability for same-source analyses of bullets, but do not provide any error rates in their discussion.
+\citet{aoas} use multiple features like CCF, CMS, D (distance measure) etc in a random forest based method and compare every land against every other land of  digitised versions of Hamby 252 and Hamby 44 \citep{hamby} published on the NIST Ballistics Database \citep{nist}.
+The authors report an out-of-bag overall error rate of 0.46%, comprised of a false positive error rate of 30.05% and a false negative rate of 0.026%.
+
+The Chumbley score  provides us with another approach in the same-source assessment of bullet striation marks.
+\citet{chumbley}, in their paper, compare two toolmarks with the intention of determining if it comes from the same tool.
+The data for their study was obtained from 50 sequentially manufactured screwdriver tips.
+\citet{chumbley} reported error rates for marking made by the tips at differnt angles.
 For a 30 degree they report false negative error rate to be 0.023 and  false positive error rate be 0.09.

### 2018-03-16 doc/paper.tex
couple more citations

--- 2018-02-19 doc/paper.tex
@@
 
-\hh{XXX first sentence: Same source analyses are a major part of an Forensic Toolmark examiners job.
-XXX second sentence: describe standard practice: In current practice firearm examiners  make these comparisons visually under a comparison microscope and report XXX what is being reported: AFTE.
-XXX find: AFTE theory of identification as it relates to toolmarks  }
-\hh{One of the main criticisms in this toolmark examination first raised by the National Research Council in 2009 [@NAS:2009] and later emphasized further by the President's Council of Advisors on Science and Technology [@pcast2016] are the amount of subjectivity in the assessment and the lack of error rates.
-}
-\hh{XXX we have digitized images, e.g. profilometers or Confocal microscopy allow us to measure the surface}
+Same source analyses are a major part of an Forensic Toolmark Examiner's
+job. In current practice examiners make these comparisons by visual
+inspection under a comparison microscope and come to one of the
+following four conclusions: identification, inconclusive, elimination or
+unsuitable for examination\textasciitilde{}\citep{afte-toolmarks1998}.
+These conclusions are made on the basis of ``unique surface contours''
+of the two toolmarks being in ``sufficient agreement''
+\citep{afte-toolmarks1998}. AFTE describes the term ``sufficient
+agreement'' as the possibility of another tool producing the markings
+under comparison, as practically impossible \citep{afte-toolmarks1998}.
+This subjectivity in the assessment as well as the lack of error rates
+are the main points of criticisms first raised by the National Research
+Council in 2009 \citep{NAS:2009} and later emphasized further by the
+President's Council of Advisors on Science and Technology
+\citep{pcast2016}.
 
-\hh{Using a statistical approach removes both subjectivity from the assessment and allows us to provide error rates for both false positive and false negative identifications. }
-\hh{Multiple toolmarks have been studied in the literature:  \citep{manytoolmarks1, manytoolmarks2, chumbley} have studied screwdriver marks digitized using a profilometer,  \citep{manytoolmarks2} investigate groove pliers XXX data?
-XXX  to slip-joint pliers \citep{afte-chumbley} XXX.
-In all of these cases, the statistical test used, is a variation of the Chumbley scoring method, first introduced by XXX and later improved by XXX.
-}
+Technological advances, such as profilometers and confocal microscopy
+allow to measure 3D surfaces in a high-resolution digitized form. This
+technology has become more accessible over the last decade, and has made
+its way into topological images of ballistics evidence, such as bullet
+lands and breech faces
+\citep{DeKinder1, DeKinder2, Bachrach1, vorburger2016}. Digitized images
+of 3D surfaces of form the data basis of statistical analysis of
+toolmarks. A statistical approach based on data removes both
+subjectivity from the assessment and allows a quantification of error
+rates for both false positive and false negative identifications.
 
-\hh{XXX In this pper, we are investigating the applicability of the Chumbley scoring method by @hadler to assessing striation marks on bullets for same-source identification. }
+Various toolmarks have been studied in the literature:
+\citet{manytoolmarks1} and \citet{chumbley} have been analyzing
+screwdriver marks digitized using a profilometer; \citet{manytoolmarks2}
+have investigated 3D marks from screwdriver, tongue and groove pliers
+captured using a confocal microscope; \citet{afte-chumbley} have been
+investigated digitized marks from slip-joint pliers generated by a
+surface profilometer.
 
-\hh{Striation marks on bullets are made by impurities in the barrel.
-As the bullet travels through the barrel, these imperfections leave "scratches' on the bullet surface.
-Typically, only striation marks in the land engraved areas (LEAs) are considered (XXX citation) [@afte-article1992].
-Compared to toolmarks made by screwdrivers striation marks on bullets are typically much smaller, both in length and in width.
-Bullets also have a curved cross-sectional topography.
-}
+\hh{We need an additional sentence here ... } \citet{manytoolmarks2}
+define a relative distance metric and use it as similarity measure
+between two toolmarks. \citet{manytoolmarks1} extract many small
+segments in the markings of two toolmarks and compare similiarity using
+a maximum pearson correlation coefficient. The Chumbley scoring method,
+first introduced by \citet{chumbley}, uses a similar but more extensive
+framework based on a Mann-Whitney U test of the resulting correlation
+coefficients. This approach was later improved by Hadler and Morris
+\citep{hadler} where the authors introduce a deterministic approach and
+address the issue of lack of independence between segements of striae.
+In this paper, we are investigating the applicability of the Chumbley
+scoring method by \citet{hadler} to assess striation marks on bullet
+lands for same-source identification.
 
-Compairing pairs of toolmarks with the intention of matching it to a
-tool has been studied many times in the past. Extensive examples can be
-found in literature for tools and toolmark research ranging from
-screwdrivers \citep{manytoolmarks1, manytoolmarks2, chumbley} to groove
-pliers \citep{manytoolmarks2} to slip-joint pliers \citep{afte-chumbley}
-and many more. In comparison to this, same source matching of bullets to
-firearms has not been examined as prominently as that of toolmarks. Even
-less information is available on validity of methods and error rates
-associated with firearms examination. The National Academy of Sciences
-in its report in 2009 \citep{NAS:2009} discussed the need for
-determining error rates in methods proposed for firearms examination. As
-seen in the case of most forensic applications, the first step for same
-source matching and error rate determination involves identification of
-unique features that are characteristic of the object at hand. For the
-case of bullets and firearms, striation marks on the surface of the
-bullet are considered to be such markings that can be used in methods
-for same source matching. These marks are often a product of rifling and
-impurities and defects due to manufacturing in the barrel of the gun,
-which leads to engravings on the bullet surface
-\citep{afte-article1992}. In current practice, firearm examiners
-invariably make visual comparisons of bullet striae and use visual
-assesment tools to dignify bullets as being matches and non-matches. One
-way of accomplishing anykind of comparison between bullets is to do a
-comparison between surface marking of two or more bullet lands. Bullet
-Lands are considered to be areas between grooves made by the rifling
-action of the barrel and the markings on them are considered to be
-unique. The land engraved markings or sometimes termed as Bullet
-profiles \citep{aoas} are striation marks made on Bullet lands and often
-used for these land to land comparisons. Bullet Signatures is another
-word used in literature as seen in the work of \citet{chu2013} and
-\citet{aoas}. In our context bullet signatures refer to a processed
-version of the raw land engraved markings or profiles. The generation of
-bullet signatures involves first extraction of a bullet profile by
-taking the cross-sectional of the surface at a given height and then
-using loess fits to model the structure. The residuals of this fit are
-called signatures, which are considered to be noise free and a good
-reflection of the class charecteristics and unique features of a bullet.
-A more detailed version of the extraction technique of signatures is
-discussed by \citet{aoas}, where comprehensive details about the height
-at which profile is to be selected, removing curvature, smoothing,
-identifying groove locations are explained. Figure \ref{fig:rgl} shows
-us how the extracted signature from a bullet land lines up with the
-image of the land from which it was extracted. We can see also see in
-the figure how the depth and relative position of the striation markings
-seen in the image are interpreted as the signature.
+Striation marks on bullets are made by impurities in the barrel. As the
+bullet travels through the barrel, these imperfections leave
+``scratches'' on the bullet surface. Typically, only striation marks in
+the land engraved areas (LEAs) are considered \citet{afte-article1992}.
+Bullet lands are depressed areas between the grooves made by the rifling
+action of the barrel. Compared to toolmarks made by screwdrivers
+striation marks on bullets are typically much smaller, both in length
+and in width. Bullets also have a curved cross-sectional topography.
+Figure\textasciitilde{}\ref{fig:rgl} shows us how the signature from a
+bullet land (bottom) lines up with the image of the land (top) from
+which it was extracted. We can also see in the figure how the depth and
+relative position of the striation markings seen in the image are
+interpreted as the signature.
 
-As mentioned earlier, subject bias and error rate determination have
-been a long standing issue in firearm examination \citep{NAS:2009}. This
-issue of removing subjectivity and making firearm analysis objective was
-also raised in the \citet{pcast2016} report. For an objective analysis,
-scientific principle testability and error rate determination are
-considered to be defining aspects. Identification of the uncertainities
-associated with a method of comparison and coming up with a systemized
-method of quantification, is therefore important. Associating
-mathematical probabilites in a systematic manner to these methods are
-one such way of quantification. In a previous study conducted by
-\citet{aoas} a machine learning based algorithm was developed for same
-source matching of bullets and error rates were discussed using the
-database from the Hamby Study \citep{hamby}. The method proposed by
-\citet{chumbley} (and later improved by \citet{hadler}) also provides a
-means to determine error rates and claims to reduce subject bias.
-Therefore, giving us a strong motivation to explore the adaptibility of
-the Chumbley score methodology to bullets. In this paper, we take this
-toolmark based methodology and try to apply and adapt it to bullets.
-After this, we consequently discuss about the efforts in doing so, along
-with the associated error rates. The data used in this paper also
-belongs to the Hamby Study \citep{hamby}. This gives us a common
-platform for comparing the performance of the chumbley method on bullets
-with the already existing method proposed by \citet{aoas} for bullets.
-As a brief overview, \citet{chumbley}, in their paper, compare two
-toolmarks with the intention of determining if it comes from the same
-tool. They use an empirical based setup to validate their proposed
-algorithm and quantitative method which calculates a U-statisic for the
-purpose of classification of toolmarks as matching or non-matching. The
-data for their study was obtained from 50 sequentially manufactured
-screwdriver tips, and preselected comparison window sizes were given as
-inputs to the algorithm. The algorithm then compares the two toolmarks
-and comes up with a Mann-Whitney U-statistic and an associated p-value
-to designate them as matches or non-matches. The performance for every
-100 comparisons, of the algorithm proposed by \citet{chumbley} and the
-improvement proposed by \citet{hadler} are listed in the table below.
+\hh{Here's a bit of a jump: you're already talking about features, but we haven't yet properly introduced signatures of bullets yet.}
+Bullet matching methods depend on the attributes used for making
+comparisons. For single feature based matching \citet{chu2013} used an
+automatic method for counting consecutive matching striae (CMS). The
+authors report an error rate of 52\% of the known same source lands
+comparisons as misidentified (false negative) and zero false positives
+for known different source lands. \citet{ma2004} and
+\citet{vorburger2011} discuss CCF (cross-correlation function) and its
+discriminating power and applicability for same-source analyses of
+bullets, but do not provide any error rates in their discussion.
+\citet{aoas} use multiple features like CCF, CMS, D (distance measure)
+etc in a random forest based method and compare every land against every
+other land of digitised versions of Hamby 252 and Hamby 44 \citet{hamby}
+published on the NIST Ballistics Database \citep{nist}. The authors
+report an out-of-bag overall error rate of 0.46\%, comprised of a false
+positive error rate of 30.05\% and a false negative rate of 0.026\%.
 
-\begin{table}[!htb]
-    \begin{minipage}{.5\linewidth}
-      \caption{Chumbley et al. 2010}
+The Chumbley score provides us with another approach in the same-source
+assessment of bullet striation marks.\citet{chumbley}, in their paper,
+compare two toolmarks with the intention of determining if it comes from
+the same tool. The data for their study was obtained from 50
+sequentially manufactured screwdriver tips.\citet{chumbley} reported
+error rates for marking made by the tips at differnt angles. For a 30
+degree they report false negative error rate to be 0.023 and false
+positive error rate be 0.09. For other angles the error rates remain the
+same for false negatives and 0.01 for false positives. The improvement
+proposed by \citet{hadler} on the other hand, did not consider effect of
+angles and gave 0.06 false negative error rate and 0 false poisitive
+error rate for markings made at the same angle.
+
+\begin{figure}
 \centering
-\begin{tabular}{lrr}
-\toprule
-Classification & Match & Non-Match\\
-\midrule
-Match & 41 & 9\\
-Non-Match & 2 & 48\\
-\bottomrule
-\end{tabular} \end{minipage}%
-    \begin{minipage}{.5\linewidth}
-      \centering
-        \caption{Hadler et. al. (2017)} 
-\begin{tabular}{lrr}
-\toprule
-Classification & Match & Non-Match\\
-\midrule
-Match & 47 & 3\\
-Non-Match & 0 & 50\\
-\bottomrule
-\end{tabular} \end{minipage} 
-\end{table}
+\includegraphics[width=\textwidth]{images/B6-B2-L6-rescaled.png}
 
-\subsection{Potential limitations of the Chumbley Score adaptation to
-bullets}\label{potential-limitations-of-the-chumbley-score-adaptation-to-bullets}
+\caption{\label{fig:rgl} Image of a bullet land from a confocal light microscope at 20 fold magnification (top) and a chart of the corresponding signature of the same land (bottom).
+The dotted lines connect some peaks visible in both visualizations.}
 
-Bullet striations are different from toolmarks. Bullet markings are
-typically produced by the barrel on the bullets, while toolmarks can be
-made by tool tips on different surfaces. Tools like screw driver tips as
-used by \citet{chumbley}, produce longer and pronounced markings.
-Bullets on the other hand are much smaller in length, width, and curved
-in the cross-sectional topography. This means the markings made by
-barrels as opposed to markings made by tools, may have a problem in
-distinctiveness. The majority of Bullet profiles and signatures of the
-Hamby study \citep{hamby} used in this paper are extracted by procedures
-mentioned by \citet{aoas}. The Ruger pistol barrels used in the Hamby
-study have a 9mm caliber. This means that bullets used in these pistols
-have land markings of almost \((1/6)^{th}\) the size of screwdriver
-toolmarks used by \citet{chumbley}. Also, striations on bullets are made
-on their curved surfaces, whereas the algorithm developed by
-\citet{chumbley} and \citet{hadler} has only been tested for flatter and
-wider surfaces which have negligible curvature. Therefore, using methods
-proposed for toolmarks may need adaptation in order to give tangible
-results for bullets. Moreover, in order to get flat bullet signatures
-and remove the curvatures some kind of smoothing needs to be applied as
-a pre-step wihch needs further investigation as to whether the level of
-smoothing does effect the working of the algorithm on Bullets.
+\end{figure}
 
-The Chumbley method and algorithm works on the premise of comparison
-windows which are supposed to represent small segments of pixels of
-equal sizes in the two markings that are to be compared in a
-predetermined fashion. They term them as windows of optimization and
-validation. The optimization window is selected in the first step called
-the optimization step where areas/windows of best agreement between the
-the two markings are found. The window of optimization for bullet
-striations are shorter as bullet signatures are smaller compared to
-toolmarks . The idea is to keep the number of windows of optimization
-sufficiently large. This means we have shorter trace segments that lets
-us compare smaller segments of one signature to another than those for
-toolmarks. Here, trace segments are the partitions of marking with the
-length of the partition equal to the size of window of optimization.
-This introduces a problem as, if we go too small in the window of
-optimization, the unique features of the trace segments are lost and
-seem similar, while too large sizes vastly reduces the weight of small
-features that would otherwise uniquely classify a signature and hence
-identify the region of agreement. The problem of distinctiveness in the
-structure of bullet striations may also lead to potential failure of the
-algorithm to identify correct windows of agreements or not be able to
-identify these windows at all. This may lead to incorrect classification
-of bullets as matches or non-matches.
+\subsection{Scans for land engraved
+areas}\label{scans-for-land-engraved-areas}
+
+The land engraved markings across the cross section of the land are
+termed as bullet profiles \citep{aoas,ma2004}. These are often used for
+making land to land comparisons. Bullet signatures on the other hand
+\citep{chu2013,aoas} refer to a processed version of the raw land
+engraved markings or profiles. The generation of bullet signatures
+involves first extraction of a bullet profile by taking the
+cross-sectional of the surface at a given height. Loess fits are then
+used to model the structure. The residuals of these fits are called
+signatures. They are considered to be noise free and a good reflection
+of the key attributes of the raw marking, and the unique features of a
+bullet. A more detailed version of the extraction technique of
+signatures is discussed by \citet{aoas}, where comprehensive details
+about the height at which profile is to be selected, removing curvature,
+smoothing, identifying groove locations are explained.
+
+The majority of Bullet profiles and signatures of the Hamby study
+\citep{hamby} used in this paper are extracted by procedures mentioned
+by \citet{aoas}. The scans for the Hamby study come from two sources
+namely, NIST scans \citep{nist} and scans produced at Iowa State
+University. Both scans were generated using confocal microscopes but at
+different resolutions. The NIST database contains scans of the bullet
+sets Hamby 44 and 252 with the resolution 1.5625 \(\mu m\) per pixel.
+The Iowa State scans are available for only the Hamby 44 bullet set and
+at a resoultion of 0.65 \(\mu m\) per pixel. The data used by
+\citet{chumbley, hadler} on the other hand, was produced by a
+profilometer at a resolution of about 0.73 \(\mu m\) per pixel. The
+different resolutions indicate that the length of the digitized markings
+would be significantly different for each setting.
+
+\subsection{Potential Challenges in Chumbley Score
+Adaptation}\label{potential-challenges-in-chumbley-score-adaptation}
+
+Markings made by Screw driver tips \citep{chumbley} are longer and
+pronounced and about 7 mm in length \citep{manytoolmarks1}, while the
+groove to groove length of the bullet lands from the Hamby study
+\citep{hamby} are about 2mm in length. The Ruger pistol barrels used in
+the Hamby study have a 9mm caliber. This means that bullets used in
+these pistols have land markings of almost \((1/4)^{th}\) the size of
+screwdriver toolmarks used by \citet{chumbley, hadler}, therefore
+raising a problem of distinctiveness. One way to investigate this is to
+evaluate bullet striations at different resolutions. This gives us an
+incentive to use both the NIST and Iowa State scans and investigate the
+effect of length of the digitized mark on the working of the algorithm.
+Apart from this, the algorithm developed by \citet{hadler} has only been
+tested for flatter and wider surfaces which have negligible curvature,
+whereas striations on bullets are made on their curved surfaces.
+Therefore, using methods proposed for toolmarks need adaptation in order
+to give tangible results for bullets. Another challenge that arises from
+these topographical differences, is the understanding of that effect of
+different levels of smoothing. Especially investigating its effect on
+the algorithm performance when signatures are being compared and when
+raw markings or profiles are being compared.
+
+Some other challenges associated with the adaptation process call for a
+brief understanding of the methodology involved. The Chumbley method and
+algorithm works on the premise of comparison windows that are supposed
+to represent small segments of pixels of equal sizes in the two
+markings. These windows are to be compared in a predetermined fashion
+and are termed as windows of optimization and validation. The
+optimization window is selected in the first step called the
+optimization step where areas/windows of best agreement between the the
+two markings are found.
+
+The window of optimization for bullet striations are shorter as bullet
+signatures in terms of absolute length (in mm), are smaller compared to
+toolmarks. Even in terms of digitized pixels, there is no
+straightforward way of establishing what window size is to be chosen.
+One important way to address this is to keep the number of windows of
+optimization sufficiently large or not too small. This means, if the
+number of windows are too small we inadvertently use bigger window sizes
+as we want to compare all segements of one mark with the segments of the
+other mark. This leads to not giving enough emphasis to smaller
+segements of the two marking that might be in high agreement. In terms
+of absolute length we end up comparing smaller segments between the two
+markings while dealing with bullet lands, than those being compared for
+toolmarks.
+
+For too large window sizes, the weight of small features that would
+otherwise uniquely classify a signature and hence identify the region of
+agreement is vastly reduced. On the other hand, this introduces another
+problem as, if we go too small in the window of optimization, the unique
+features of the trace segments are lost and all seem similar. The
+problem of distinctiveness in the structure of bullet striations may
+also lead to potential failure of the algorithm to identify correct
+windows of agreements or not be able to identify these windows at all.
+This may lead to incorrect classification of bullets as coming from the
+same-source (or matches) or from different source (or non-matches).
 
@@
 
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{images/B6-B2-L6-rescaled.png}
-
-\caption{\label{fig:rgl} Image of a bullet land from a confocal light microscope at 20 fold magnification (top) and a chart of the corresponding signature of the same land (bottom).
-The dotted lines connect some peaks visible in both visualizations.}
-
-\end{figure}
-
 \subsection{The Chumbley Score Test}\label{the-chumbley-score-test}
@@
 process. \(t\) here denotes a vector of equally spaced pixel locations
-for the striation marks under consideration. The word `pixel' refers to
-the resolution of the confocal light microscope. In the case for bullet
-signatures extracted from NIST ballistics database \citep{nist} of the
-Hamby study, a pixel corresponds to approximately 1.56 microns. The
-vectorized processes representing two sets of striaes are therefore
-shown as \(x(t_1)\), \(t_1 = 1,2,...T_1\) and \(y(t_2)\),
-\(t_2 = 1,2...T_2\). Here \(x\) and \(y\) denote two instances of the
-process \(z(*)\) which means two striation marks, while the indexing
-\(t\) or pixel range is shown as \(t_1\) and \(t_2\) for \(x\) and \(y\)
-respectively. This representation of \(t\) as \(t_1\) and \(t_2\) lets
-us identify them as either of same or different lengths. The striation
-marks under consideration are potentially from two different bullets
-whose source needs to be identified as being same or different. \(T_1\)
-and \(T_2\), as represented above, are the final pixel indexes of each
-marking and therefore give the respective lengths of the markings. The
-similarity in the two markings is then judged by the algorithm on the
-basis of cross-correlation of a fixed and constant number consecutive
-pixels (say \(k\)) taken from the two markings. This can for example be
-\(k\) taken from the two indexed marking \(x(t_1)\) and \(y(t_2)\) such
-that in theory \(k\) remains smaller than length of the two striae or
-marks. Depending on what stage of the algorithm we are in, matching of
+for the striation marks under consideration. The vectorized processes
+representing two sets of striaes are therefore shown as \(x(t_1)\),
+\(t_1 = 1,2,...T_1\) and \(y(t_2)\), \(t_2 = 1,2...T_2\). Here \(x\) and
+\(y\) denote two instances of the process \(z(*)\) which means two
+striation marks, while the indexing \(t\) or pixel range is shown as
+\(t_1\) and \(t_2\) for \(x\) and \(y\) respectively. This
+representation of \(t\) as \(t_1\) and \(t_2\) lets us identify them as
+either of same or different lengths. The striation marks under
+consideration are potentially from two different bullets whose source
+needs to be identified as being same or different. \(T_1\) and \(T_2\),
+as represented above, are the final pixel indexes of each marking and
+therefore give the respective lengths of the markings. The similarity in
+the two markings is then judged by the algorithm on the basis of
+cross-correlation of a fixed and constant number consecutive pixels (say
+\(k\)) taken from the two markings. This can for example be \(k\) taken
+from the two indexed marking \(x(t_1)\) and \(y(t_2)\) such that in
+theory \(k\) remains smaller than length of the two striae or marks.
+Depending on what stage of the algorithm we are in, matching of
 different pixel lengths and locations is done. This in the end,
@@
 correlations of both the same and different shift samples.
+
+\paragraph{endedit}\label{endedit}
+
+\hh{end of intro: remaining paper is structured as follows: introduce to the data we get from confocal microscopy, introduce to profiles and signatures.
+Introduce to chumbley score method, apply chumbley and discuss results ....
+}
+
+\subsection{Scans for land engraved
+areas}\label{scans-for-land-engraved-areas-1}
+
+\begin{itemize}
+\tightlist
+\item
+  scans available: NIST database (citation), Hamby 44 and Hamby 252
+  (Hamby citation)
+\item
+  move figure up
+\item
+  discuss cross section, profiles and signatures
+\end{itemize}
 

### 2018-03-16 doc/paper.Rmd
couple of edits

@@
 
-\hh{Here's a bit of a jump: you're already talking about features, but we haven't yet properly introduced signatures of bullets yet.}
-Bullet matching methods depend on the attributes  used for making comparisons.
-For single feature based matching \citet{chu2013} used an automatic method for counting consecutive matching striae (CMS).
+Bullet matching methods are usually based on these associated signatures.
+\citet{chu2013} use an automatic method for counting consecutive matching striae (CMS).
 The authors report an error rate of 52% of the known same source lands comparisons as misidentified (false negative) and zero false positives for known different source lands.
@@
 The Chumbley score  provides us with another approach in the same-source assessment of bullet striation marks.
-\citet{chumbley}, in their paper, compare two toolmarks with the intention of determining if it comes from the same tool.
-The data for their study was obtained from 50 sequentially manufactured screwdriver tips.
-\citet{chumbley} reported error rates for marking made by the tips at differnt angles.
-For a 30 degree they report false negative error rate to be 0.023 and  false positive error rate be 0.09.
-For other angles the error rates remain the same for false negatives and 0.01 for false positives.
-The improvement proposed by \citet{hadler} on the other hand, did not consider effect of angles and gave 0.06 false  negative error rate and 0 false poisitive error rate for markings made at the same angle.
+\citet{chumbley} compare two toolmarks for same-source.
+The data for this study was obtained from 50 sequentially manufactured screwdriver tips.
+\citet{chumbley} report error rates for markings made by the tips at different angles.
+For markings made under a 30 degree the authors  report an average false negative error rate of 0.023  and an average false positive error rate of 0.09.
+For other angles the error rates for false negatives stay the same while the rate of false positives decreases to 0.01.
+\hh{are the angles steeper?}
+The paper by \citet{hadler} is based on the same data but the authors focus on  markings made under the same angle.
+The error rates associated with the deterministic version of the score are  0.06 for false  negatives and a false poisitive error rate of 0.
 
@@
 
-The land engraved markings across the cross section of the land are termed as bullet profiles \citep{aoas,ma2004}.
-These are often used for making land to land comparisons.
-Bullet signatures on the other hand \citep{chu2013,aoas} refer to a processed version of the raw land engraved markings or profiles.
-The generation of bullet signatures involves first extraction of a bullet profile by taking the cross-sectional of the surface at a given height.
-Loess fits are then used to model the structure.
-The residuals of these fits are called signatures.
+Comparisons of striae from bullets are usually based on comparisons of striae in land engraved areas, which are extracted in form of cross sections, called *profiles* \citep{aoas,ma2004}.
+From profiles bullet *signatures* \citep{chu2013,aoas} are extracted as residuals of a loess fit (or Gaussian filter).
  They are considered to be noise free and a good reflection of the key attributes of the raw marking, and the unique features of a bullet.
-A more detailed version of the extraction technique of signatures is discussed by \citet{aoas}, where comprehensive details about the height at which profile is to be selected, removing curvature, smoothing, identifying groove locations are explained.
-
-The majority of Bullet profiles and signatures of the Hamby study \citep{hamby} used in this paper are extracted by procedures mentioned by  \citet{aoas}.
-The scans for the Hamby study come from two sources namely, NIST scans \citep{nist} and scans produced at Iowa State University.
-Both scans were generated using confocal microscopes but at different resolutions.
-The NIST database contains scans of the bullet sets Hamby 44 and 252 with the resolution 1.5625 $\mu m$ per pixel.
-The Iowa State scans are available for only the Hamby 44 bullet set and at a resoultion of 0.65 $\mu m$ per pixel.
+\hh{Do you have a reference for the previous sentenced?} A  detailed discussion of the extraction technique for signatures is given in \citet{aoas}.
+ <!--, where comprehensive details about the height at which profile is to be selected, removing curvature, smoothing, identifying groove locations are explained.-->
+
+\hh{don't split the discussion on the size. between the next paragraph and }
+<!--The majority of Bullet profiles and signatures of the Hamby study \citep{hamby} used in this paper are extracted by procedures mentioned by  \citet{aoas}.--> The scans for the Hamby study come from two sources namely, NIST scans \citep{nist} and scans produced at Iowa State University \hh{Iowa State is too general.
+it's the High-resolution microscopy lab and we are CSAFE}.
+Both scans were generated using confocal light microscopes at a 20x magnification but the microscopes collected data at different resolutions.
+The NIST database contains scans of the bullet sets Hamby 44 and 252 at a resolution of 1.5625 $\mu m$ per pixel.
+The Iowa State scans are available for only the Hamby 44 bullet set and at a resoultion of 0.645 $\mu m$ per pixel.
 The data used by \citet{chumbley, hadler} on the other hand, was produced by a profilometer at a resolution of about 0.73 $\mu m$ per pixel.
@@
 
-Markings made by Screw driver tips \citep{chumbley} are longer and pronounced and about 7 mm in length \citep{manytoolmarks1}, while the groove to groove length of the bullet lands from the Hamby study \citep{hamby} are about 2mm in length.
+Markings made by screw driver tips \citep{chumbley} are longer and pronounced and about 7 mm in length \citep{manytoolmarks1}, while the groove to groove length of the bullet lands from the Hamby study \citep{hamby} are about 2mm in length.
 The Ruger pistol barrels used in the  Hamby study have a 9mm caliber.
-This means that bullets used in these pistols have land markings of almost $(1/4)^{th}$ the size of screwdriver toolmarks used by \citet{chumbley, hadler}, therefore raising a problem of distinctiveness.
+This means that bullets used in these pistols have land markings of only about a quarter of the size of screwdriver toolmarks used by \citet{chumbley, hadler}, therefore raising a problem of distinctiveness.
 One way to investigate this is to evaluate bullet striations at different resolutions.

### 2018-03-16 doc/paper.tex
couple of edits

@@
 
+\hh{In the next page and a half it is easy to lose the red line. It might help to include a table with an overview.
+The table should include the reference to the paper, the data used, the statistical method and the associated error rates.}
 Various toolmarks have been studied in the literature:
@@
 
-\hh{We need an additional sentence here ... } \citet{manytoolmarks2}
-define a relative distance metric and use it as similarity measure
-between two toolmarks. \citet{manytoolmarks1} extract many small
-segments in the markings of two toolmarks and compare similiarity using
-a maximum pearson correlation coefficient. The Chumbley scoring method,
-first introduced by \citet{chumbley}, uses a similar but more extensive
-framework based on a Mann-Whitney U test of the resulting correlation
-coefficients. This approach was later improved by Hadler and Morris
-\citep{hadler} where the authors introduce a deterministic approach and
-address the issue of lack of independence between segements of striae.
-In this paper, we are investigating the applicability of the Chumbley
-scoring method by \citet{hadler} to assess striation marks on bullet
-lands for same-source identification.
+\hh{We need an additional sentence here to get from the data to the statistical methods ... }
+\citet{manytoolmarks2} define a relative distance metric and use it as
+similarity measure between two toolmarks. \citet{manytoolmarks1} extract
+many small segments in the markings of two toolmarks and compare
+similiarity using a maximum pearson correlation coefficient. The
+Chumbley scoring method, first introduced by \citet{chumbley}, uses a
+similar but more extensive framework based on a Mann-Whitney U test of
+the resulting correlation coefficients. This approach was later improved
+by Hadler and Morris \citep{hadler} where the authors introduce a
+deterministic approach and address the issue of lack of independence
+between segements of striae. In this paper, we are investigating the
+applicability of the Chumbley scoring method by \citet{hadler} to assess
+striation marks on bullet lands for same-source identification.
 
@@
 
-\hh{Here's a bit of a jump: you're already talking about features, but we haven't yet properly introduced signatures of bullets yet.}
-Bullet matching methods depend on the attributes used for making
-comparisons. For single feature based matching \citet{chu2013} used an
-automatic method for counting consecutive matching striae (CMS). The
-authors report an error rate of 52\% of the known same source lands
-comparisons as misidentified (false negative) and zero false positives
-for known different source lands. \citet{ma2004} and
-\citet{vorburger2011} discuss CCF (cross-correlation function) and its
-discriminating power and applicability for same-source analyses of
-bullets, but do not provide any error rates in their discussion.
-\citet{aoas} use multiple features like CCF, CMS, D (distance measure)
-etc in a random forest based method and compare every land against every
-other land of digitised versions of Hamby 252 and Hamby 44 \citet{hamby}
-published on the NIST Ballistics Database \citep{nist}. The authors
-report an out-of-bag overall error rate of 0.46\%, comprised of a false
-positive error rate of 30.05\% and a false negative rate of 0.026\%.
+Bullet matching methods are usually based on these associated
+signatures. \citet{chu2013} use an automatic method for counting
+consecutive matching striae (CMS). The authors report an error rate of
+52\% of the known same source lands comparisons as misidentified (false
+negative) and zero false positives for known different source lands.
+\citet{ma2004} and \citet{vorburger2011} discuss CCF (cross-correlation
+function) and its discriminating power and applicability for same-source
+analyses of bullets, but do not provide any error rates in their
+discussion. \citet{aoas} use multiple features like CCF, CMS, D
+(distance measure) etc in a random forest based method and compare every
+land against every other land of digitised versions of Hamby 252 and
+Hamby 44 \citep{hamby} published on the NIST Ballistics Database
+\citep{nist}. The authors report an out-of-bag overall error rate of
+0.46\%, comprised of a false positive error rate of 30.05\% and a false
+negative rate of 0.026\%.
 
 The Chumbley score provides us with another approach in the same-source
-assessment of bullet striation marks.\citet{chumbley}, in their paper,
-compare two toolmarks with the intention of determining if it comes from
-the same tool. The data for their study was obtained from 50
-sequentially manufactured screwdriver tips.\citet{chumbley} reported
-error rates for marking made by the tips at differnt angles. For a 30
-degree they report false negative error rate to be 0.023 and false
-positive error rate be 0.09. For other angles the error rates remain the
-same for false negatives and 0.01 for false positives. The improvement
-proposed by \citet{hadler} on the other hand, did not consider effect of
-angles and gave 0.06 false negative error rate and 0 false poisitive
-error rate for markings made at the same angle.
+assessment of bullet striation marks. \citet{chumbley} compare two
+toolmarks for same-source. The data for this study was obtained from 50
+sequentially manufactured screwdriver tips. \citet{chumbley} report
+error rates for markings made by the tips at different angles. For
+markings made under a 30 degree the authors report an average false
+negative error rate of 0.023 and an average false positive error rate of
+0.09. For other angles the error rates for false negatives stay the same
+while the rate of false positives decreases to 0.01.
+\hh{are the angles steeper?} The paper by \citet{hadler} is based on the
+same data but the authors focus on markings made under the same angle.
+The error rates associated with the deterministic version of the score
+are 0.06 for false negatives and a false poisitive error rate of 0.
 
@@
 
-The land engraved markings across the cross section of the land are
-termed as bullet profiles \citep{aoas,ma2004}. These are often used for
-making land to land comparisons. Bullet signatures on the other hand
-\citep{chu2013,aoas} refer to a processed version of the raw land
-engraved markings or profiles. The generation of bullet signatures
-involves first extraction of a bullet profile by taking the
-cross-sectional of the surface at a given height. Loess fits are then
-used to model the structure. The residuals of these fits are called
-signatures. They are considered to be noise free and a good reflection
-of the key attributes of the raw marking, and the unique features of a
-bullet. A more detailed version of the extraction technique of
-signatures is discussed by \citet{aoas}, where comprehensive details
-about the height at which profile is to be selected, removing curvature,
-smoothing, identifying groove locations are explained.
+Comparisons of striae from bullets are usually based on comparisons of
+striae in land engraved areas, which are extracted in form of cross
+sections, called \emph{profiles} \citep{aoas,ma2004}. From profiles
+bullet \emph{signatures} \citep{chu2013,aoas} are extracted as residuals
+of a loess fit (or Gaussian filter). They are considered to be noise
+free and a good reflection of the key attributes of the raw marking, and
+the unique features of a bullet.
+\hh{Do you have a reference for the previous sentenced?} A detailed
+discussion of the extraction technique for signatures is given in
+\citet{aoas}.
 
-The majority of Bullet profiles and signatures of the Hamby study
-\citep{hamby} used in this paper are extracted by procedures mentioned
-by \citet{aoas}. The scans for the Hamby study come from two sources
-namely, NIST scans \citep{nist} and scans produced at Iowa State
-University. Both scans were generated using confocal microscopes but at
-different resolutions. The NIST database contains scans of the bullet
-sets Hamby 44 and 252 with the resolution 1.5625 \(\mu m\) per pixel.
-The Iowa State scans are available for only the Hamby 44 bullet set and
-at a resoultion of 0.65 \(\mu m\) per pixel. The data used by
-\citet{chumbley, hadler} on the other hand, was produced by a
-profilometer at a resolution of about 0.73 \(\mu m\) per pixel. The
-different resolutions indicate that the length of the digitized markings
-would be significantly different for each setting.
+\hh{don't split the discussion on the size. between the next paragraph and }
+The scans for the Hamby study come from two sources namely, NIST scans
+\citep{nist} and scans produced at Iowa State University
+\hh{Iowa State is too general. it's the High-resolution microscopy lab and we are CSAFE}.
+Both scans were generated using confocal light microscopes at a 20x
+magnification but the microscopes collected data at different
+resolutions. The NIST database contains scans of the bullet sets Hamby
+44 and 252 at a resolution of 1.5625 \(\mu m\) per pixel. The Iowa State
+scans are available for only the Hamby 44 bullet set and at a resoultion
+of 0.645 \(\mu m\) per pixel. The data used by \citet{chumbley, hadler}
+on the other hand, was produced by a profilometer at a resolution of
+about 0.73 \(\mu m\) per pixel. The different resolutions indicate that
+the length of the digitized markings would be significantly different
+for each setting.
 
@@
 
-Markings made by Screw driver tips \citep{chumbley} are longer and
+Markings made by screw driver tips \citep{chumbley} are longer and
 pronounced and about 7 mm in length \citep{manytoolmarks1}, while the
@@
 the Hamby study have a 9mm caliber. This means that bullets used in
-these pistols have land markings of almost \((1/4)^{th}\) the size of
+these pistols have land markings of only about a quarter of the size of
 screwdriver toolmarks used by \citet{chumbley, hadler}, therefore

### 2018-03-16 doc/paper.Rmd
couple of edits

@@
 The Ruger pistol barrels used in the  Hamby study have a 9mm caliber.
-This means that bullets used in these pistols have land markings of only about a quarter of the size of screwdriver toolmarks used by \citet{chumbley, hadler}, therefore raising a problem of distinctiveness.
-One way to investigate this is to evaluate bullet striations at different resolutions.
+This means that bullets used in these pistols have land markings of only about a quarter of the size of screwdriver toolmarks used by \citet{chumbley, hadler}, therefore raising the problem of potential loss in power of the test.
+\hh{The Chumbley score allows a separation of a toolmark into raw and normalized digitized versions.
+Originally, this mechanism is intended to separate between class characteristics and individual markings, however, in the setting of matching bullet striae, we could also use it to separate bullet curvature in profiles from signatures before matching signatures.
+}
+
+<!--One way to investigate this is to evaluate bullet striations at different resolutions.
 This gives us an incentive to use both the NIST and Iowa State scans and investigate the effect of length of the digitized mark on the working of the algorithm.
@@
 An analysis of these error rates and comparison with other methods will help us understand the adaptibility of the chumbley score to bullets.
+-->
 

### 2018-03-16 code/hamby44-csafe-scans.R
csafe-scans-chumbley_comparison

@@ -1,0 +1,148 @@
+# This document is meant for the Hamby 44 scan made at Iowa State
+# Resolution of the scans 0.0635 microns per pixel
+
+# Loading Signature selection analytics
+# Remove unkonwns
+# Bullets coming from same barrel should match
+# Lands that belong to these bullets must match? Do lands from the same bullet match,
+
+library(readr)
+library(toolmaRk)
+hamby44<- readRDS("/opt/hamby44/hamby44.rda")
+featureplus<- read_csv("/opt/hamby44/featuresplus.csv",col_names = T)
+
+library(tidyverse)
+library(purrr)
+hamb<- hamby44 %>% select(-c(x3p,ccdata))
+
+#Filtered out the unknowns from the hamb
+hamb<- hamb %>% filter(barrel != "Unknowns")
+
+# Groove removal
+hamb$groove_left<- purrr::map(hamb$grooves,.f = function(x) x[["groove"]][1])
+hamb$groove_right<- purrr::map(hamb$grooves,.f = function(x) x[["groove"]][2])
+
+marking_proc<- hamb %>% select(proc_smooth)
+marking_proc<- marking_proc[,2]
+
+
+for(i in 1:120){
+  if(!unlist(purrr::map( marking_proc[[1]][i], .f = function(x) is_empty(x)))){ # check for empty proc_smooth #96 i.e Barrel 9 BUllet 2 Land 3 alldata missing
+    marking_proc[[1]][i]<- marking_proc[[1]][i] %>% purrr::map(.f = function(x) as.data.frame(x) %>% filter(y >= hamb$groove_left[i] & y <= hamb$groove_right[i]))  
+    } else{
+      marking_proc[[1]][i]<-  NA
+  }
+  
+
+}
+
+# generating id of the form barrel-bullet-land in hamb
+names(marking_proc[[1]])<- paste0(str_split(hamb$barrel, pattern = " ",simplify = T)[,2],
+                "-",
+                str_split(hamb$bullet, pattern = " ",simplify = T)[,2],
+                "-",
+                str_split(hamb$land, pattern = " ",simplify = T)[,2]
+                )
+
+## Next steps Involve making land id's for hamb and featureplus matches
[…]

### 2018-03-16 code/hamby44-csafe-scans.R
small edit in code commmments

@@ -146,3 +146,34 @@ for(i in 97:120){
 }
 
 # Chop? No data for y values with the grooves in proc_smooth
+
+# xtabs(~signif+sameland, data =aa) 
+# 
+# 19/(19+21)
+# 
+# # Read in 200 and 50 and 0.25
+# 
+# aa200_50<- read.csv2("./data/csafe-scans/signatures/csafe_sig_chumbley_wo-200_wv-50_coarse-pt25.csv",header = TRUE, sep = ",")
+# xtabs(~signif+sameland, data =aa200_50) 
+#  
+# 408/(408+6562) = 0.05853659
+# 18/(18+23)= 0.4390244
+# 
+# aa1750_50<- read.csv2("./data/csafe-scans/signatures/csafe_sig_chumbley_wo-175_wv-50_coarse-pt25.csv",header = TRUE, sep = ",")
+# xtabs(~signif+sameland, data =aa1750_50) 
+# 
+# 422/(422+6553) = 0.06050179
+# 18/(18+23)= 0.4390244
+# 
+# aa250_50<- read.csv2("./data/csafe-scans/signatures/csafe_sig_chumbley_wo-250_wv-50_coarse-pt25.csv",header = TRUE, sep = ",")
+# xtabs(~signif+sameland, data =aa250_50) 
+# 
+# 412/(412+6555) = 0.05913593
+# 16/(16+ 24) = 0.4
+# 
+# 
+# nist_aa120_30<- read.csv2("./data/signatures/chumbley-out-wo-120-wv-30.csv",header = TRUE, sep = ",")
+# xtabs(~signif+match, data =nist_aa120_30)
+# 4523/(79249 + 4523) = 0.05399179
+# 386/(386 + 854) = 0.3112903
+#   

### 2018-03-16 doc/paper.Rmd
up to chumbley score

@@
 Comparisons of striae from bullets are usually based on comparisons of striae in land engraved areas, which are extracted in form of cross sections, called *profiles* \citep{aoas,ma2004}.
-From profiles bullet *signatures* \citep{chu2013,aoas} are extracted as residuals of a loess fit (or Gaussian filter).
- They are considered to be noise free and a good reflection of the key attributes of the raw marking, and the unique features of a bullet.
+From profiles bullet *signatures* \citep{chu2013,aoas} are extracted as residuals of a loess fit or Gaussian filter.
+ Signatures are considered to be noise free and a good reflection of the key attributes of the raw marking, and the unique features of a bullet.
 \hh{Do you have a reference for the previous sentenced?} A  detailed discussion of the extraction technique for signatures is given in \citet{aoas}.
@@
 \hh{don't split the discussion on the size. between the next paragraph and }
-<!--The majority of Bullet profiles and signatures of the Hamby study \citep{hamby} used in this paper are extracted by procedures mentioned by  \citet{aoas}.--> The scans for the Hamby study come from two sources namely, NIST scans \citep{nist} and scans produced at Iowa State University \hh{Iowa State is too general.
-it's the High-resolution microscopy lab and we are CSAFE}.
-Both scans were generated using confocal light microscopes at a 20x magnification but the microscopes collected data at different resolutions.
-The NIST database contains scans of the bullet sets Hamby 44 and 252 at a resolution of 1.5625 $\mu m$ per pixel.
-The Iowa State scans are available for only the Hamby 44 bullet set and at a resoultion of 0.645 $\mu m$ per pixel.
-The data used by \citet{chumbley, hadler} on the other hand, was produced by a profilometer at a resolution of about 0.73 $\mu m$ per pixel.
-The different resolutions indicate that the length of the digitized markings would be significantly different for each setting.
+<!--The majority of Bullet profiles and signatures of the Hamby study \citep{hamby} used in this paper are extracted by procedures mentioned by  \citet{aoas}.--> 
+There are two sources of scans for sets from the Hamby study available to us: scans of Hamby 44 and Hamby 252 are available from the NIST database \citep{nist}.
+Hamby 44 has also been made available to us and has been scanned locally for CSAFE at the Roy J.\ Carver High Resolution Microscopy Facility using a Sensofar confocal light microscope.
+Scans in the NIST database are made with a NanoFocus at 20x magnification.
+The resolutions of the two instruments are different: the NIST scans are taken at a resolution of 1.5625 $\mu m$ per pixel, while the CSAFE scans are available at a resolution of  0.645 $\mu m$ per pixel.
+The length of an average bullet land from Hamby (9 mm Ruger P85) is about 2 millimeter, resulting in signatures of about 1200 pixels for NIST scans, and about 3000 pixels for CSAFE scans.
+
+In comparison, scans from the profilometer used by \citet{chumbley, hadler} were taken at a resolution of about 0.73 $\mu m$ per pixel.
+The screw driver toolmarks are  about 7 mm in length \citep{manytoolmarks1}, for a total of over 9000 pixels for the width of these scans.
+
+This severe limitation in the amount of available data poses the main challenge in adapting the Chumbley score to matching bullet lands, because of the resulting loss in power.
 
@@
 
-Markings made by screw driver tips \citep{chumbley} are longer and pronounced and about 7 mm in length \citep{manytoolmarks1}, while the groove to groove length of the bullet lands from the Hamby study \citep{hamby} are about 2mm in length.
+<!--Markings made by screw driver tips \citep{chumbley} are longer and pronounced and about 7 mm in length \citep{manytoolmarks1}, while the groove to groove length of the bullet lands from the Hamby study \citep{hamby} are about 2mm in length.
 The Ruger pistol barrels used in the  Hamby study have a 9mm caliber.
 This means that bullets used in these pistols have land markings of only about a quarter of the size of screwdriver toolmarks used by \citet{chumbley, hadler}, therefore raising the problem of potential loss in power of the test.
+-->
+
 \hh{The Chumbley score allows a separation of a toolmark into raw and normalized digitized versions.

### 2018-03-16 doc/paper.tex
up to chumbley score

@@
 bullet \emph{signatures} \citep{chu2013,aoas} are extracted as residuals
-of a loess fit (or Gaussian filter). They are considered to be noise
+of a loess fit or Gaussian filter. Signatures are considered to be noise
 free and a good reflection of the key attributes of the raw marking, and
@@
 \hh{don't split the discussion on the size. between the next paragraph and }
-The scans for the Hamby study come from two sources namely, NIST scans
-\citep{nist} and scans produced at Iowa State University
-\hh{Iowa State is too general. it's the High-resolution microscopy lab and we are CSAFE}.
-Both scans were generated using confocal light microscopes at a 20x
-magnification but the microscopes collected data at different
-resolutions. The NIST database contains scans of the bullet sets Hamby
-44 and 252 at a resolution of 1.5625 \(\mu m\) per pixel. The Iowa State
-scans are available for only the Hamby 44 bullet set and at a resoultion
-of 0.645 \(\mu m\) per pixel. The data used by \citet{chumbley, hadler}
-on the other hand, was produced by a profilometer at a resolution of
-about 0.73 \(\mu m\) per pixel. The different resolutions indicate that
-the length of the digitized markings would be significantly different
-for each setting.
+There are two sources of scans for sets from the Hamby study available
+to us: scans of Hamby 44 and Hamby 252 are available from the NIST
+database \citep{nist}. Hamby 44 has also been made available to us and
+has been scanned locally for CSAFE at the Roy J.~Carver High Resolution
+Microscopy Facility using a Sensofar confocal light microscope. Scans in
+the NIST database are made with a NanoFocus at 20x magnification. The
+resolutions of the two instruments are different: the NIST scans are
+taken at a resolution of 1.5625 \(\mu m\) per pixel, while the CSAFE
+scans are available at a resolution of 0.645 \(\mu m\) per pixel. The
+length of an average bullet land from Hamby (9 mm Ruger P85) is about 2
+millimeter, resulting in signatures of about 1200 pixels for NIST scans,
+and about 3000 pixels for CSAFE scans.
+
+In comparison, scans from the profilometer used by
+\citet{chumbley, hadler} were taken at a resolution of about 0.73
+\(\mu m\) per pixel. The screw driver toolmarks are about 7 mm in length
+\citep{manytoolmarks1}, for a total of over 9000 pixels for the width of
+these scans.
+
+This severe limitation in the amount of available data poses the main
+challenge in adapting the Chumbley score to matching bullet lands,
+because of the resulting loss in power.
 
@@
 
-Markings made by screw driver tips \citep{chumbley} are longer and
-pronounced and about 7 mm in length \citep{manytoolmarks1}, while the
-groove to groove length of the bullet lands from the Hamby study
-\citep{hamby} are about 2mm in length. The Ruger pistol barrels used in
-the Hamby study have a 9mm caliber. This means that bullets used in
-these pistols have land markings of only about a quarter of the size of
-screwdriver toolmarks used by \citet{chumbley, hadler}, therefore
-raising a problem of distinctiveness. One way to investigate this is to
-evaluate bullet striations at different resolutions. This gives us an
-incentive to use both the NIST and Iowa State scans and investigate the
-effect of length of the digitized mark on the working of the algorithm.
-Apart from this, the algorithm developed by \citet{hadler} has only been
-tested for flatter and wider surfaces which have negligible curvature,
-whereas striations on bullets are made on their curved surfaces.
-Therefore, using methods proposed for toolmarks need adaptation in order
-to give tangible results for bullets. Another challenge that arises from
-these topographical differences, is the understanding of that effect of
-different levels of smoothing. Especially investigating its effect on
-the algorithm performance when signatures are being compared and when
-raw markings or profiles are being compared.
-
-Some other challenges associated with the adaptation process call for a
-brief understanding of the methodology involved. The Chumbley method and
-algorithm works on the premise of comparison windows that are supposed
-to represent small segments of pixels of equal sizes in the two
-markings. These windows are to be compared in a predetermined fashion
-and are termed as windows of optimization and validation. The
-optimization window is selected in the first step called the
-optimization step where areas/windows of best agreement between the the
-two markings are found.
-
-The window of optimization for bullet striations are shorter as bullet
-signatures in terms of absolute length (in mm), are smaller compared to
-toolmarks. Even in terms of digitized pixels, there is no
-straightforward way of establishing what window size is to be chosen.
-One important way to address this is to keep the number of windows of
-optimization sufficiently large or not too small. This means, if the
-number of windows are too small we inadvertently use bigger window sizes
-as we want to compare all segements of one mark with the segments of the
-other mark. This leads to not giving enough emphasis to smaller
-segements of the two marking that might be in high agreement. In terms
-of absolute length we end up comparing smaller segments between the two
-markings while dealing with bullet lands, than those being compared for
-toolmarks.
-
-For too large window sizes, the weight of small features that would
-otherwise uniquely classify a signature and hence identify the region of
-agreement is vastly reduced. On the other hand, this introduces another
-problem as, if we go too small in the window of optimization, the unique
-features of the trace segments are lost and all seem similar. The
-problem of distinctiveness in the structure of bullet striations may
-also lead to potential failure of the algorithm to identify correct
-windows of agreements or not be able to identify these windows at all.
-This may lead to incorrect classification of bullets as coming from the
-same-source (or matches) or from different source (or non-matches).
-
-Thus, the comaparison windows have a direct influence on the false
-positives and false negatives, making identification of the optimum
-sizes of these windows very important. This raises important questions
-about how and what parameter settings need to be chosen for bullet land
-comparison. Something similar is done in \citet{afte-chumbley} for
-toolmark comparisons of slip-joint pliers where optimum window sizes are
-determined. There is a need to figure out the best parameter settings
-which minimizes the errors for unsmoothed markings or profiles and
-pre-processed signatures. An analysis of these error rates and
-comparison with other methods will help us understand the adaptibility
-of the chumbley score to bullets.
+\hh{The Chumbley score allows a separation of a toolmark into raw and normalized digitized versions.
+Originally, this mechanism is intended to separate between class characteristics and individual markings, however, in the setting of matching bullet striae, we could also use it to separate bullet curvature in profiles from signatures before matching signatures.
+}
 

### 2018-03-17 code/hamby44-csafe-scans.R
extract hamby 44 study - NIST, clean it to have same lands as csafe-scans, compare

@@ -1,10 +1,5 @@
-# This document is meant for the Hamby 44 scan made at Iowa State
-# Resolution of the scans 0.0635 microns per pixel
-
-# Loading Signature selection analytics
-# Remove unkonwns
-# Bullets coming from same barrel should match
-# Lands that belong to these bullets must match? Do lands from the same bullet match,
+# This document is meant for the Hamby 44 scan made at CSAFE
+# Resolution of the scans 0.0645 microns per pixel
 
 library(readr)
 library(toolmaRk)
@@ -55,8 +50,9 @@ f_sub <- f_sub %>% filter(known == TRUE)
 f_sub<- f_sub[,c("barrel_bullet_land1","barrel_bullet_land2","sameland")]#,"samesource")]
 #f_sub %>% filter(samesource == FALSE, sameland == TRUE)
 
-wo <- 120
-wv <- 30
+# Chumbley comparison
+wo <- 300
+wv <- 120
 coarse <- 0.25
 k <- 1
 n <- nrow(marking_proc)
@@ -66,8 +62,10 @@ names(dframe) <- c("barrel_bullet_land1", "barrel_bullet_land2", "chumbley")
 
 for (i in 1:(n-1)) {
   for (j in (i+1):n) {
+    # Extracting the two signatures from the nested data
     marking_I<- marking_proc[[1]][i] %>% purrr::map(.f = function(x) x["l30"]) %>% unlist() %>% as.numeric()
     marking_II<- marking_proc[[1]][j] %>% purrr::map(.f = function(x) x["l30"]) %>% unlist() %>% as.numeric()
+    
     if(!(length(marking_I) == 1) && !(length(marking_II) == 1)){ # check for empty of non-existing signature/ proc_smooth fields typically will have an NA (owing to previous setup)
       dframe$chumbley[k] <- I(list(chumbley_non_random(matrix(marking_I), matrix(marking_II),  window_opt=wo, window_val=wv, coarse=coarse)))  
       dframe$barrel_bullet_land1[k] <- names(marking_proc[[1]][i])
@@ -83,13 +81,16 @@ dframe$wo <- wo
 dframe$wv <- wv
 dframe$coarse <- coarse
 
-dframe_raw<- dframe
+#dframe_raw<- dframe
 
[…]

### 2018-03-17 doc/paper.Rmd
paper edits
chumbley score description

@@
 
-  rates <- summ %>% group_by(wv, wo, coarse, match, alpha) %>% summarize(
+  rates <- summ %>% dplyr::group_by(wv, wo, coarse, match, alpha) %>% dplyr::summarize(
     rate = n[error==TRUE]/sum(n)
   )
-  totals <- summ %>% group_by(wv, wo, coarse, alpha) %>% summarize(
+  totals <- summ %>% dplyr::group_by(wv, wo, coarse, alpha) %>% dplyr::summarize(
     total= sum(n[error==TRUE])/sum(n)
   )
-  rates <- rates %>% spread(match,rate) %>% rename(
+  rates <- rates %>% spread(match,rate) %>% dplyr::rename(
     actual = `FALSE`,
@@
 The Chumbley scoring method, first introduced by \citet{chumbley}, uses a similar but more extensive framework based on a Mann-Whitney U test of the resulting correlation coefficients.
-This approach was later improved by Hadler and Morris \citep{hadler} where the authors introduce a deterministic approach and address the issue of lack of independence between segements of striae.
+This approach is non-deterministic, because segments are chosen randomly.
+\citep{hadler} make the score deterministic for each pair of toolmarks by choosing segments for comparison systematically.
+This approach also ensures independence between segements of striae.
 In this paper, we are investigating the applicability of the Chumbley scoring method by \citet{hadler} to assess striation marks on bullet lands for same-source identification.
@@
 Bullets also have a curved cross-sectional topography.
-Figure~\ref{fig:rgl} shows us how the signature from a bullet land (bottom) lines up with the image of the land (top) from which it was extracted.
+Figure \ref{fig:rgl} shows us how the signature from a bullet land (bottom) lines up with the image of the land (top) from which it was extracted.
 We can also see in the figure how the depth and relative position of the striation markings seen in the image are interpreted as the signature.
@@
 
-The Chumbley score algorithm takes input as two vectorized processes.
-Here the processes are of the form $z(t)$ which is a spatial process for some location indexed with $t$.
-This means that while $z(t)$ represents any striae, $z(t_1)$ is a realization of the spatial process.
-$t$ here denotes a vector of equally spaced pixel locations for the striation marks under consideration.
-The vectorized processes representing two sets of striaes are therefore shown as $x(t_1)$, $t_1 = 1,2,...T_1$ and $y(t_2)$, $t_2 = 1,2...T_2$.
+The Chumbley score algorithm takes input in form of two digitized toolmarks.
+The toolmark is in form of $z(t)$ which is a spatial process for location indexed by $t$.
+$t$ here denotes equally spaced pixel locations for the striation marks under consideration, $t = 1, ..., T$.
+Let further $z(s,t)$ denote the vector of markings between locations $s$ and $t$.
+
+Let $x(t_1)$, $t_1 = 1,2,...T_1$ and $y(t_2)$, $t_2 = 1,2...T_2$ be two digitized toolmarks (where $T_1$ and $T_2$ are not necessarily equal). 
+<!--The vectorized processes representing two sets of striaes are therefore shown as $x(t_1)$, $t_1 = 1,2,...T_1$ and $y(t_2)$, $t_2 = 1,2...T_2$.
 Here $x$ and $y$ denote two instances of the process $z(*)$ which means two striation marks, while the indexing $t$ or pixel range is shown as $t_1$ and $t_2$ for $x$ and $y$ respectively.
-This representation of $t$ as $t_1$ and $t_2$ lets us identify them as either of same or different lengths.
-The striation marks under consideration are potentially from two different bullets whose source needs to be identified as being same or different.
+This representation of $t$ as $t_1$ and $t_2$ lets us identify them as either of same or different lengths.-->
+The toolmarks under consideration are potentially from two different sources or the same source.
 $T_1$ and $T_2$, as represented above, are the final pixel indexes of each marking and therefore give the respective lengths of the markings.
+
+The algorithm works in two phases, namely, an optimization step and a validation step.
+
+In the optimization step, the two markings are aligned horizontally such that within a pre-defined window of length $w_o$ the correlation between $x(t_1)$ and $y(t_2)$ is maximized:
+\[
+\left(t_1^o, t_2^o\right) = \mathop{\arg \max}\limits_{1 \le t_1 \le T_1, 1 \le t_2 \le T_2} \text{cor} \left(x (t_1, t_1 + w_o), y(t_2, t_2 + w_o) \right)
+\]
+This results in an optimal vertical (in-phase) shift of $t_1^o - t_2^o$ for aligning the two markings.
+
+In the validation step,  two sets of windows of size $w_v$ are chosen from both markings (see Figure \ref{fig:win-comparison}).
+In the first set, pairs of windows are extracted from the two markings using the optimal vertical shift as determined in the first step, whereas for the second set the windows are extracted using a different (out-of-phase) shift.
+
+```{r win-comparison, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap=" The two plots on the left show how the same shift behaves in case of a matching pair and the two plots on the right show how the different shift behaves in case of a matching pair."}
+
+asd<- all %>% filter(wo==120, wv==50, match == TRUE, land1_id == 1)
+
+sigs_graphics<- readRDS("../data/sigs_generate_window.rds")
+
+sigs_graphics.match<- as.data.frame(sigs_graphics[1])
+sigs_graphics.nonmatch<- as.data.frame(sigs_graphics[2])
+colnames(sigs_graphics.match)<- gsub("^.*\\.","", colnames(sigs_graphics.match))
+colnames(sigs_graphics.nonmatch)<- gsub("^.*\\.","", colnames(sigs_graphics.nonmatch))
+
+# Matching signature
+#gi<-
+#sigs_graphics.match<- sigs_graphics.match %>% filter(land_id == c(1,9))  
+#sigs_graphics.match<- sigs_graphics.match %>% select(y, l30, land_id)
+d1<- sigs_graphics.match[which(sigs_graphics.match$land_id==1,arr.ind = TRUE),]
+d2<- sigs_graphics.match[which(sigs_graphics.match$land_id==19,arr.ind = TRUE),]
+data1<-d1 %>%select(l30)
+#data1<- data1$y
+data2<- d2 %>%select(l30)
+#data2<- data2$y
+window_opt = 120 
+window_val = 50
+coarse = 1
+data1<- matrix(unlist(data1))
+data2<- matrix(unlist(data2))
+
+  unity <- function(x) {x / sqrt(sum(x^2))} ## normalize columns of a matrix to make correlation computation faster
+
+  ####################################################
+  ##Clean the marks and compute the smooth residuals##
+  ####################################################
+
+  data1 <- matrix(data1[round((0.01*nrow(data1))):round(0.99*nrow(data1)),], ncol = 1)
+  data2 <- matrix(data2[round((0.01*nrow(data2))):round(0.99*nrow(data2)),], ncol = 1)
+
+  ##Normalize the tool marks
+  y1 <- data1 - lowess(y = data1,  x = 1:nrow(data1), f= coarse)$y
+  y2 <- data2 - lowess(y = data2,  x = 1:nrow(data2), f= coarse)$y
+
+  ############################################
+  ##Compute the observed maximum correlation##
+  ############################################
+
+  #####################
+  ##Optimization step##
+  #####################
+  ##Each column in these matrices corresponds to a window in the respective tool mark
+  y1_mat_opt <- matrix(NA, ncol = length(1:(length(y1) - (window_opt - 1))), nrow = window_opt)
+  for(l in 1:(length(y1) - (window_opt - 1))){
+    y1_mat_opt[,l] <- y1[l:(l+(window_opt - 1))]
+  }
+  y2_mat_opt <- matrix(NA, ncol = length(1:(length(y2) - (window_opt - 1))), nrow = window_opt)
+  for(l in 1:(length(y2) - (window_opt - 1))){
+    y2_mat_opt[,l] <- y2[l:(l+(window_opt - 1))]
+  }
+
+  ##Compute the correlation between all pairs of windows for the two marks
+  ##Rows in the following matrix are mark 2, columns are mark 1
+  y2_mat_opt <- apply(scale(y2_mat_opt), 2, unity)
+  y1_mat_opt <- apply(scale(y1_mat_opt), 2, unity)
+  corr_mat_opt <- t(y2_mat_opt) %*% y1_mat_opt ##correlation matrix
+  max_corr_opt_loc <- which(corr_mat_opt == max(corr_mat_opt), arr.ind = TRUE) ##pair of windows maximizing the correlation
+
+  s1.rect.opt <- data.frame(xmin=max_corr_opt_loc[1,2], xmax= max_corr_opt_loc[1,2]+window_opt, ymin=-Inf, ymax=Inf)
+    s1.rect.sameshift1 <- data.frame(xmin=max_corr_opt_loc[1,2]-window_val, xmax= max_corr_opt_loc[1,2]- 2*window_val, ymin=-Inf, ymax=Inf)
+       s1.rect.sameshift2 <- data.frame(xmin=max_corr_opt_loc[1,2]+window_opt+ window_val, xmax= max_corr_opt_loc[1,2]+ window_opt+2*window_val, ymin=-Inf, ymax=Inf)
+
+  s2.rect.opt <- data.frame(xmin=max_corr_opt_loc[1,1], xmax= max_corr_opt_loc[1,1]-window_opt, ymin=-Inf, ymax=Inf)
+    s2.rect.sameshift1 <- data.frame(xmin=max_corr_opt_loc[1,1]-window_opt-window_val, xmax= max_corr_opt_loc[1,1]-window_opt- 2*window_val, ymin=-Inf, ymax=Inf)
+       s2.rect.sameshift2 <- data.frame(xmin=max_corr_opt_loc[1,1]+ window_val, xmax= max_corr_opt_loc[1,1]+ 2*window_val, ymin=-Inf, ymax=Inf)
+
+p1<-  ggplot(data = d1, aes(x = y, y = l30)) + geom_line() + 
+  geom_rect(data=s1.rect.opt, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "black",alpha=0.5, inherit.aes = FALSE)+
+    geom_rect(data=s1.rect.sameshift1, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "red",alpha=0.1, inherit.aes = FALSE)+
+    geom_rect(data=s1.rect.sameshift2, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "blue",alpha=0.1, inherit.aes = FALSE)+
+  theme_bw() + labs(y = "Signature Land Id = 1")
+
+ p2<-   ggplot(data = d2, aes(x = y, y = l30)) + geom_line() + 
+   geom_rect(data=s2.rect.opt, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "black",alpha=0.5, inherit.aes = FALSE)+
+    geom_rect(data=s2.rect.sameshift1, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "red",alpha=0.1, inherit.aes = FALSE)+
+    geom_rect(data=s2.rect.sameshift2, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "blue",alpha=0.1, inherit.aes = FALSE)+
+  theme_bw() + labs(y = "Signature Land Id = 9")
+ ##########
+
+ d1.rect.opt <- data.frame(xmin=max_corr_opt_loc[1,2], xmax= max_corr_opt_loc[1,2]+window_opt, ymin=-Inf, ymax=Inf)
+    d1.rect.sameshift1 <- data.frame(xmin=max_corr_opt_loc[1,2]-2*window_opt, xmax= max_corr_opt_loc[1,2]-2*window_opt-window_val, ymin=-Inf, ymax=Inf)
+       d1.rect.sameshift2 <- data.frame(xmin=max_corr_opt_loc[1,2]+window_opt+ window_val, xmax= max_corr_opt_loc[1,2]+ window_opt+2*window_val, ymin=-Inf, ymax=Inf)
+
+  d2.rect.opt <- data.frame(xmin=max_corr_opt_loc[1,1], xmax= max_corr_opt_loc[1,1]-window_opt, ymin=-Inf, ymax=Inf)
+    d2.rect.sameshift1 <- data.frame(xmin=max_corr_opt_loc[1,1]-window_opt-window_val, xmax= max_corr_opt_loc[1,1]-window_opt- 2*window_val, ymin=-Inf, ymax=Inf)
+       d2.rect.sameshift2 <- data.frame(xmin=max_corr_opt_loc[1,1]+ 4*window_val, xmax= max_corr_opt_loc[1,1]+ 5*window_val, ymin=-Inf, ymax=Inf)
+
+p3<-  ggplot(data = d1, aes(x = y, y = l30)) + geom_line() + 
+  geom_rect(data=d1.rect.opt, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "black",alpha=0.5, inherit.aes = FALSE)+
+    geom_rect(data=d1.rect.sameshift1, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "red",alpha=0.1, inherit.aes = FALSE)+
+    geom_rect(data=d1.rect.sameshift2, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "blue",alpha=0.1, inherit.aes = FALSE)+
+  theme_bw() + labs(y = "Signature Land Id = 1")
+
+ p4<-   ggplot(data = d2, aes(x = y, y = l30)) + geom_line() + 
+   geom_rect(data=d2.rect.opt, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "black",alpha=0.5, inherit.aes = FALSE)+
+    geom_rect(data=d2.rect.sameshift1, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "red",alpha=0.1, inherit.aes = FALSE)+
+    geom_rect(data=d2.rect.sameshift2, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "blue",alpha=0.1, inherit.aes = FALSE)+
+  theme_bw() + labs(y = "Signature Land Id = 9")
+
+multiplot(p1, p2, p3, p4, cols=2)
+```
+
+For both samples the correlations between the pairs of markings is then calculated. 
+The intuition here is that for two markings from the same source the correlation for the in-phase sample should be high, while the correlations of the out-of-phase sample  provide a measure for the base-level correlation for non-matching marks of a given length $w_v$.
+The Chumbley score is then computed as a Mann Whitney U statistic to compare between in-phase sample and out-of-phase sample.
+In the original method proposed in \citet{chumbley} both in-phase and out-of-phase sample are extracted randomly, whereas \citet{hadler} proposed deterministic rules for both samples to make the resulting score deterministic while simultaneously avoiding overlaps within selected marks to ensure independence.
+
+A pre-processing step to the algorithm is to choose a coarseness value which  is used as a parameter to the lowess smoothing function.
+The coarseness essentially gives the proportion of points which influence the smooth at each value, which means larger values lead to more smoothness.
+The lowess smoothing is applied to each of two sets of vectorized striae or marks $x(t_1)$ and $y(t_2)$, before proceeding to the algorithm.
+
+<!--
 The similarity in the two markings is then judged by the algorithm on the basis of cross-correlation of a fixed and constant number consecutive pixels (say $k$) taken from the two markings.
@@
 This in the end, effectively compares all possible windows that would guarantee in quantifying the two marks or striae as coming from the same source or not.
+-->
 
@@
 
+<!--
 The algorithm works in two phases, namely, an optimization step and a validation step, at the end of which a Mann Whitney U statistic is calculated as a measure to differentiate between matches and non-matches.
@@
 The coarseness essentially gives the proportion of points which influence the smooth at each value, which means larger values lead to more smoothness.
-The lowess smoothing is applied to each of two sets of vectorized striae or marks $x(t_1)$ and $y(t_2)$, before proceeding to the algorithm.
-The algorithm starts with the optimization step where the area of best agreement in the two markings being compared is identified.
+The lowess smoothing is applied to each of two sets of vectorized striae or marks $x(t_1)$ and $y(t_2)$, before proceeding to the algorithm.-->
+<!-- The algorithm starts with the optimization step where the area of best agreement in the two markings being compared is identified.
 Comparison window size is predefined.
@@
 Also a random sampling sub step in the validation phase makes a group of pixels to be chosen more than once and hence introduces lack of independence in certain steps.
-The reason for removing the mutual dependence was the Mann-Whitney U statistic works under the assumption of independence of parameters.
+The reason for removing the mutual dependence was the Mann-Whitney U statistic works under the assumption of independence of parameters.-->
+<!--
 The validation step builds on this with two sub-steps namely, same shift and different shift.
@@
 As given by \citet{hadler}, it is computed from the joint rank of all correlations of both the same and different shift samples.
+-->
 
@@
 
-```{r win-comparison, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap=" The two plots on the left show how the same shift behaves in case of a matching pair and the two plots on the right show how the different shift behaves in case of a matching pair."}
-
-asd<- all %>% filter(wo==120, wv==50, match == TRUE, land1_id == 1)
-
-sigs_graphics<- readRDS("../data/sigs_generate_window.rds")
-
-sigs_graphics.match<- as.data.frame(sigs_graphics[1])
-sigs_graphics.nonmatch<- as.data.frame(sigs_graphics[2])
-colnames(sigs_graphics.match)<- gsub("^.*\\.","", colnames(sigs_graphics.match))
-colnames(sigs_graphics.nonmatch)<- gsub("^.*\\.","", colnames(sigs_graphics.nonmatch))
-
-# Matching signature
-#gi<-
-#sigs_graphics.match<- sigs_graphics.match %>% filter(land_id == c(1,9))  
-#sigs_graphics.match<- sigs_graphics.match %>% select(y, l30, land_id)
-d1<- sigs_graphics.match[which(sigs_graphics.match$land_id==1,arr.ind = TRUE),]
-d2<- sigs_graphics.match[which(sigs_graphics.match$land_id==19,arr.ind = TRUE),]
-data1<-d1 %>%select(l30)
-#data1<- data1$y
-data2<- d2 %>%select(l30)
-#data2<- data2$y
-window_opt = 120 
-window_val = 50
-coarse = 1
-data1<- matrix(unlist(data1))
-data2<- matrix(unlist(data2))
-
-  unity <- function(x) {x / sqrt(sum(x^2))} ## normalize columns of a matrix to make correlation computation faster
-
-  ####################################################
-  ##Clean the marks and compute the smooth residuals##
-  ####################################################
-
-  data1 <- matrix(data1[round((0.01*nrow(data1))):round(0.99*nrow(data1)),], ncol = 1)
-  data2 <- matrix(data2[round((0.01*nrow(data2))):round(0.99*nrow(data2)),], ncol = 1)
-
-  ##Normalize the tool marks
-  y1 <- data1 - lowess(y = data1,  x = 1:nrow(data1), f= coarse)$y
-  y2 <- data2 - lowess(y = data2,  x = 1:nrow(data2), f= coarse)$y
-
-  ############################################
-  ##Compute the observed maximum correlation##
-  ############################################
-
-  #####################
-  ##Optimization step##
-  #####################
-  ##Each column in these matrices corresponds to a window in the respective tool mark
-  y1_mat_opt <- matrix(NA, ncol = length(1:(length(y1) - (window_opt - 1))), nrow = window_opt)
-  for(l in 1:(length(y1) - (window_opt - 1))){
-    y1_mat_opt[,l] <- y1[l:(l+(window_opt - 1))]
-  }
-  y2_mat_opt <- matrix(NA, ncol = length(1:(length(y2) - (window_opt - 1))), nrow = window_opt)
-  for(l in 1:(length(y2) - (window_opt - 1))){
-    y2_mat_opt[,l] <- y2[l:(l+(window_opt - 1))]
-  }
-
-  ##Compute the correlation between all pairs of windows for the two marks
-  ##Rows in the following matrix are mark 2, columns are mark 1
-  y2_mat_opt <- apply(scale(y2_mat_opt), 2, unity)
-  y1_mat_opt <- apply(scale(y1_mat_opt), 2, unity)
-  corr_mat_opt <- t(y2_mat_opt) %*% y1_mat_opt ##correlation matrix
-  max_corr_opt_loc <- which(corr_mat_opt == max(corr_mat_opt), arr.ind = TRUE) ##pair of windows maximizing the correlation
-
-  s1.rect.opt <- data.frame(xmin=max_corr_opt_loc[1,2], xmax= max_corr_opt_loc[1,2]+window_opt, ymin=-Inf, ymax=Inf)
-    s1.rect.sameshift1 <- data.frame(xmin=max_corr_opt_loc[1,2]-window_val, xmax= max_corr_opt_loc[1,2]- 2*window_val, ymin=-Inf, ymax=Inf)
-       s1.rect.sameshift2 <- data.frame(xmin=max_corr_opt_loc[1,2]+window_opt+ window_val, xmax= max_corr_opt_loc[1,2]+ window_opt+2*window_val, ymin=-Inf, ymax=Inf)
-
-  s2.rect.opt <- data.frame(xmin=max_corr_opt_loc[1,1], xmax= max_corr_opt_loc[1,1]-window_opt, ymin=-Inf, ymax=Inf)
-    s2.rect.sameshift1 <- data.frame(xmin=max_corr_opt_loc[1,1]-window_opt-window_val, xmax= max_corr_opt_loc[1,1]-window_opt- 2*window_val, ymin=-Inf, ymax=Inf)
-       s2.rect.sameshift2 <- data.frame(xmin=max_corr_opt_loc[1,1]+ window_val, xmax= max_corr_opt_loc[1,1]+ 2*window_val, ymin=-Inf, ymax=Inf)
-
-p1<-  ggplot(data = d1, aes(x = y, y = l30)) + geom_line() + 
-  geom_rect(data=s1.rect.opt, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "black",alpha=0.5, inherit.aes = FALSE)+
-    geom_rect(data=s1.rect.sameshift1, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "red",alpha=0.1, inherit.aes = FALSE)+
-    geom_rect(data=s1.rect.sameshift2, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "blue",alpha=0.1, inherit.aes = FALSE)+
-  theme_bw() + labs(y = "Signature Land Id = 1")
-
- p2<-   ggplot(data = d2, aes(x = y, y = l30)) + geom_line() + 
-   geom_rect(data=s2.rect.opt, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "black",alpha=0.5, inherit.aes = FALSE)+
-    geom_rect(data=s2.rect.sameshift1, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "red",alpha=0.1, inherit.aes = FALSE)+
-    geom_rect(data=s2.rect.sameshift2, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "blue",alpha=0.1, inherit.aes = FALSE)+
-  theme_bw() + labs(y = "Signature Land Id = 9")
- ##########
-
- d1.rect.opt <- data.frame(xmin=max_corr_opt_loc[1,2], xmax= max_corr_opt_loc[1,2]+window_opt, ymin=-Inf, ymax=Inf)
-    d1.rect.sameshift1 <- data.frame(xmin=max_corr_opt_loc[1,2]-2*window_opt, xmax= max_corr_opt_loc[1,2]-2*window_opt-window_val, ymin=-Inf, ymax=Inf)
-       d1.rect.sameshift2 <- data.frame(xmin=max_corr_opt_loc[1,2]+window_opt+ window_val, xmax= max_corr_opt_loc[1,2]+ window_opt+2*window_val, ymin=-Inf, ymax=Inf)
-
-  d2.rect.opt <- data.frame(xmin=max_corr_opt_loc[1,1], xmax= max_corr_opt_loc[1,1]-window_opt, ymin=-Inf, ymax=Inf)
-    d2.rect.sameshift1 <- data.frame(xmin=max_corr_opt_loc[1,1]-window_opt-window_val, xmax= max_corr_opt_loc[1,1]-window_opt- 2*window_val, ymin=-Inf, ymax=Inf)
-       d2.rect.sameshift2 <- data.frame(xmin=max_corr_opt_loc[1,1]+ 4*window_val, xmax= max_corr_opt_loc[1,1]+ 5*window_val, ymin=-Inf, ymax=Inf)
-
-p3<-  ggplot(data = d1, aes(x = y, y = l30)) + geom_line() + 
-  geom_rect(data=d1.rect.opt, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "black",alpha=0.5, inherit.aes = FALSE)+
-    geom_rect(data=d1.rect.sameshift1, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "red",alpha=0.1, inherit.aes = FALSE)+
-    geom_rect(data=d1.rect.sameshift2, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "blue",alpha=0.1, inherit.aes = FALSE)+
-  theme_bw() + labs(y = "Signature Land Id = 1")
-
- p4<-   ggplot(data = d2, aes(x = y, y = l30)) + geom_line() + 
-   geom_rect(data=d2.rect.opt, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "black",alpha=0.5, inherit.aes = FALSE)+
-    geom_rect(data=d2.rect.sameshift1, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "red",alpha=0.1, inherit.aes = FALSE)+
-    geom_rect(data=d2.rect.sameshift2, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "blue",alpha=0.1, inherit.aes = FALSE)+
-  theme_bw() + labs(y = "Signature Land Id = 9")
-
-multiplot(p1, p2, p3, p4, cols=2)
-```
-
 <!-- ####Different Shift: -->

### 2018-03-17 doc/paper.tex
chumbley score description

@@
 similar but more extensive framework based on a Mann-Whitney U test of
-the resulting correlation coefficients. This approach was later improved
-by Hadler and Morris \citep{hadler} where the authors introduce a
-deterministic approach and address the issue of lack of independence
-between segements of striae. In this paper, we are investigating the
-applicability of the Chumbley scoring method by \citet{hadler} to assess
-striation marks on bullet lands for same-source identification.
+the resulting correlation coefficients. This approach is
+non-deterministic, because segments are chosen randomly. \citep{hadler}
+make the score deterministic for each pair of toolmarks by choosing
+segments for comparison systematically. This approach also ensures
+independence between segements of striae. In this paper, we are
+investigating the applicability of the Chumbley scoring method by
+\citet{hadler} to assess striation marks on bullet lands for same-source
+identification.
 
@@
 and in width. Bullets also have a curved cross-sectional topography.
-Figure\textasciitilde{}\ref{fig:rgl} shows us how the signature from a
-bullet land (bottom) lines up with the image of the land (top) from
-which it was extracted. We can also see in the figure how the depth and
-relative position of the striation markings seen in the image are
-interpreted as the signature.
+Figure \ref{fig:rgl} shows us how the signature from a bullet land
+(bottom) lines up with the image of the land (top) from which it was
+extracted. We can also see in the figure how the depth and relative
+position of the striation markings seen in the image are interpreted as
+the signature.
 
@@
 
-The Chumbley score algorithm takes input as two vectorized processes.
-Here the processes are of the form \(z(t)\) which is a spatial process
-for some location indexed with \(t\). This means that while \(z(t)\)
-represents any striae, \(z(t_1)\) is a realization of the spatial
-process. \(t\) here denotes a vector of equally spaced pixel locations
-for the striation marks under consideration. The vectorized processes
-representing two sets of striaes are therefore shown as \(x(t_1)\),
-\(t_1 = 1,2,...T_1\) and \(y(t_2)\), \(t_2 = 1,2...T_2\). Here \(x\) and
-\(y\) denote two instances of the process \(z(*)\) which means two
-striation marks, while the indexing \(t\) or pixel range is shown as
-\(t_1\) and \(t_2\) for \(x\) and \(y\) respectively. This
-representation of \(t\) as \(t_1\) and \(t_2\) lets us identify them as
-either of same or different lengths. The striation marks under
-consideration are potentially from two different bullets whose source
-needs to be identified as being same or different. \(T_1\) and \(T_2\),
-as represented above, are the final pixel indexes of each marking and
-therefore give the respective lengths of the markings. The similarity in
-the two markings is then judged by the algorithm on the basis of
-cross-correlation of a fixed and constant number consecutive pixels (say
-\(k\)) taken from the two markings. This can for example be \(k\) taken
-from the two indexed marking \(x(t_1)\) and \(y(t_2)\) such that in
-theory \(k\) remains smaller than length of the two striae or marks.
-Depending on what stage of the algorithm we are in, matching of
-different pixel lengths and locations is done. This in the end,
-effectively compares all possible windows that would guarantee in
-quantifying the two marks or striae as coming from the same source or
-not.
+The Chumbley score algorithm takes input in form of two digitized
+toolmarks. The toolmark is in form of \(z(t)\) which is a spatial
+process for location indexed by \(t\). \(t\) here denotes equally spaced
+pixel locations for the striation marks under consideration,
+\(t = 1, ..., T\). Let further \(z(s,t)\) denote the vector of markings
+between locations \(s\) and \(t\).
+
+Let \(x(t_1)\), \(t_1 = 1,2,...T_1\) and \(y(t_2)\), \(t_2 = 1,2...T_2\)
+be two digitized toolmarks (where \(T_1\) and \(T_2\) are not
+necessarily equal). The toolmarks under consideration are potentially
+from two different sources or the same source. \(T_1\) and \(T_2\), as
+represented above, are the final pixel indexes of each marking and
+therefore give the respective lengths of the markings.
 
 The algorithm works in two phases, namely, an optimization step and a
-validation step, at the end of which a Mann Whitney U statistic is
-calculated as a measure to differentiate between matches and
-non-matches. A pre-processing step to the algorithm is to choose a
-coarseness value which is used as a parameter to the lowess smoothing
-function. The coarseness essentially gives the proportion of points
-which influence the smooth at each value, which means larger values lead
-to more smoothness. The lowess smoothing is applied to each of two sets
-of vectorized striae or marks \(x(t_1)\) and \(y(t_2)\), before
-proceeding to the algorithm. The algorithm starts with the optimization
-step where the area of best agreement in the two markings being compared
-is identified. Comparison window size is predefined. Each window of the
-first marking is compared with all windows of the second marking. A
-maximum correlation statistic is used to identify the region of best
-agreement, with the maximum usually seen being near 1 for both cases
-which is what is intuitively expected for matches, and not expected
-intuitively for non matches. \citet{hadler} in their paper proposed an
-improvement to this algorithm by trying to remove mutual dependence of
-parameters. The mututal dependence was due to the serial correlation in
-surface depth values of a marking. Also a random sampling sub step in
-the validation phase makes a group of pixels to be chosen more than once
-and hence introduces lack of independence in certain steps. The reason
-for removing the mutual dependence was the Mann-Whitney U statistic
-works under the assumption of independence of parameters. The validation
-step builds on this with two sub-steps namely, same shift and different
-shift. As the aim is to come up with a non-parameteric U statistic,
-\citet{hadler} proposed a normalization procedure in the validation step
-which goes to some extent to address the issue of mutual dependence. A
-series of windows are chosen in both the same and different shift
-sub-steps for the purpose of comparison. Both substeps were modified by
-\citet{hadler} who introduced a deterministic rule for sampling of
-windows in the same-shift and different shift as opposed to the
-originally proposed random sampling by \citet{chumbley}. In th same
-shift substep the chosen series of windows are at a common distance
-(rigid-shifts) from the window earlier identified as the region of best
-agreement in the optimization step. The correlation of these windows
-generally turn out to be lower than the maximum correlation window. The
-significance is that these same shift windows still have large enough
-correlation values for the two markings being compared that are in
-reality a match. Any choice of maximum correlation windows (where
-correlation values are often near 1) for non-matches in the optimization
-step are also validated by this sub-step. This is because the same-shift
-correlations would not be anywhere as large for same-shift windows when
-all windows in the sub-step are compared. The different shift substep on
-the other hand gives perspective to the correlation values of the same
-shift window correlation values. There are no rigid-shifts but different
-shifts where distance from the maximum correlation window are chosen
-randomly by \citet{chumbley} and deterministically as per
-\citet{hadler}. This means that there is an equal possibilty of
-comparing a trace segment from one marking to any trace segment or
-window in the second marking.
+validation step.
 
-Neither of above sets of correlation in the two sub-steps are allowed to
-include the maximum correlation window as identified earlier. Therefore
-the assumption is that if two markings match each other, the same-shift
-correlations would be larger than the different-shift windows. And if
-they are not a match the correlations in the two sets will be very
-similar. The U-statistic tests for the null hypothesis that the two
-markings are not a match and are therefore not made by the same source.
-As given by \citet{hadler}, it is computed from the joint rank of all
-correlations of both the same and different shift samples.
+In the optimization step, the two markings are aligned horizontally such
+that within a pre-defined window of length \(w_o\) the correlation
+between \(x(t_1)\) and \(y(t_2)\) is maximized: \[
+\left(t_1^o, t_2^o\right) = \mathop{\arg \max}\limits_{1 \le t_1 \le T_1, 1 \le t_2 \le T_2} \text{cor} \left(x (t_1, t_1 + w_o), y(t_2, t_2 + w_o) \right)
+\] This results in an optimal vertical (in-phase) shift of
+\(t_1^o - t_2^o\) for aligning the two markings.
+
+In the validation step, two sets of windows of size \(w_v\) are chosen
+from both markings (see Figure \ref{fig:win-comparison}). In the first
+set, pairs of windows are extracted from the two markings using the
+optimal vertical shift as determined in the first step, whereas for the
+second set the windows are extracted using a different (out-of-phase)
+shift.
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figures/win-comparison-1}
+
+}
+
+\caption{ The two plots on the left show how the same shift behaves in case of a matching pair and the two plots on the right show how the different shift behaves in case of a matching pair.}\label{fig:win-comparison}
+\end{figure}
+
+For both samples the correlations between the pairs of markings is then
+calculated. The intuition here is that for two markings from the same
+source the correlation for the in-phase sample should be high, while the
+correlations of the out-of-phase sample provide a measure for the
+base-level correlation for non-matching marks of a given length \(w_v\).
+The Chumbley score is then computed as a Mann Whitney U statistic to
+compare between in-phase sample and out-of-phase sample. In the original
+method proposed in \citet{chumbley} both in-phase and out-of-phase
+sample are extracted randomly, whereas \citet{hadler} proposed
+deterministic rules for both samples to make the resulting score
+deterministic while simultaneously avoiding overlaps within selected
+marks to ensure independence.
+
+A pre-processing step to the algorithm is to choose a coarseness value
+which is used as a parameter to the lowess smoothing function. The
+coarseness essentially gives the proportion of points which influence
+the smooth at each value, which means larger values lead to more
+smoothness. The lowess smoothing is applied to each of two sets of
+vectorized striae or marks \(x(t_1)\) and \(y(t_2)\), before proceeding
+to the algorithm.
 
@@
 \end{itemize}
-
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figures/win-comparison-1}
-
-}
-
-\caption{ The two plots on the left show how the same shift behaves in case of a matching pair and the two plots on the right show how the different shift behaves in case of a matching pair.}\label{fig:win-comparison}
-\end{figure}
 
@@
 \toprule
-wo & 50 & 50 & 60 & 60 & 80 & 80 & 90 & 90 & 100 & 100 & 110 & 110 & 120 & 120 & 120 & 120 & 120 & 130\\
-wv & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 10 & 20 & 30 & 50 & 60 & 30\\
+wo & 50 & 50 & 60 & 60 & 80 & 80 & 90 & 90 & 100 & 100 & 110 & 110 & 120 & 120 & 120 & 120 & 120 & 120\\
+wv & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 10 & 20 & 30 & 40 & 50 & 60\\
 \bottomrule
@@
 \toprule
-wo & 130 & 140 & 140 & 150 & 150 & 160 & 160 & 200 & 200 & 200 & 240 & 240 & 280 & 280 & 320\\
-wv & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 20 & 30 & 50 & 30 & 50 & 30 & 50 & 30\\
+wo & 130 & 130 & 140 & 140 & 150 & 150 & 160 & 160 & 200 & 200 & 200 & 240 & 240 & 280 & 280\\
+wv & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 20 & 30 & 50 & 30 & 50 & 30 & 50\\
 \bottomrule

### 2018-03-17 code/hamby44-csafe-scans.R
commenting out a redundant check

@@ -222,16 +222,16 @@ nist_sub<- nist_sub[which(!is.na(nist_sub$bbl1)),]
 #anti_join(data.frame(a = nist_sub$bbl1),data.frame(a = dframe$barrel_bullet_land1), by = "a")
 #anti_join(data.frame(a = dframe$barrel_bullet_land1), data.frame(a = nist_sub$bbl1), by = "a")
 
-# Checking for what specific lands are different between NIST and CSAFE
-# For a more accurate comparison of error rates
-tmp1_setdiff<- setdiff(dframe$barrel_bullet_land1, nist_sub$bbl1)
-tmp1_rev_setdiff<- setdiff(nist_sub$bbl1, dframe$barrel_bullet_land1)
-
-#setdiff(dframe$barrel_bullet_land2, nist_sub$bbl2)
-#setdiff(nist_sub$bbl2, dframe$barrel_bullet_land2)
-
-# Filtering out lands that are not in the iowa state scan
-nist_sub<- nist_sub %>% filter(bbl1 != tmp1_setdiff)
+# # Checking for what specific lands are different between NIST and CSAFE
+# # For a more accurate comparison of error rates
+# tmp1_setdiff<- setdiff(dframe$barrel_bullet_land1, nist_sub$bbl1)
+# tmp1_rev_setdiff<- setdiff(nist_sub$bbl1, dframe$barrel_bullet_land1)
+# 
+# #setdiff(dframe$barrel_bullet_land2, nist_sub$bbl2)
+# #setdiff(nist_sub$bbl2, dframe$barrel_bullet_land2)
+# 
+# # Filtering out lands that are not in the iowa state scan
+# nist_sub<- nist_sub %>% filter(bbl1 != tmp1_setdiff)
 #dframe<- dframe %>% filter(barrel_bullet_land1 == tmp1_rev_setdiff)
 
 xtabs(~signif+match, data =nist_sub)

### 2018-03-17 doc/paper.Rmd
Chumbley score sketch

@@
 
-The algorithm works in two phases, namely, an optimization step and a validation step.
-
+In a pre-processing step  the two markings are smoothed using a lowess \citep{lowess} with coarseness parameter $c$. The purpose of this smoothing is to remove drift and (sub)class characteristics.
+
+The Chumbley scores is calculated in two phases, namely, an optimization step and a validation step.
 In the optimization step, the two markings are aligned horizontally such that within a pre-defined window of length $w_o$ the correlation between $x(t_1)$ and $y(t_2)$ is maximized:
@@
 
-```{r win-comparison, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap=" The two plots on the left show how the same shift behaves in case of a matching pair and the two plots on the right show how the different shift behaves in case of a matching pair."}
+```{r win-comparison, warning=FALSE, fig.width=8, fig.height=3.5, fig.cap="In- and out-of-phase samples from two markings created by the same source.
+On the left, sample windows are in-phase, on the right the pairs of sampled windows are out-of-phase.
+The Chumbley-score is based on a Mann-Whitney U test of the correlations derived from these two sets."}
+sigs_graphics <- readRDS("../data/sigs_generate_window.rds")
+
+sigs_graphics.match<- as.data.frame(sigs_graphics[1])
+sigs_graphics.nonmatch<- as.data.frame(sigs_graphics[2])
+colnames(sigs_graphics.match)<- gsub("^.*\\.","", colnames(sigs_graphics.match))
+colnames(sigs_graphics.nonmatch)<- gsub("^.*\\.","", colnames(sigs_graphics.nonmatch))
+
+# Matching signature
+d1<- sigs_graphics.match[which(sigs_graphics.match$land_id==1,arr.ind = TRUE),]
+d2<- sigs_graphics.match[which(sigs_graphics.match$land_id==276,arr.ind = TRUE),]
+data1<-d1 %>%select(l30)
+#data1<- data1$y
+data2<- d2 %>%select(l30)
+#data2<- data2$y
+window_opt = 120 
+window_val = 50
+coarse = 1
+data1<- matrix(unlist(data1))
+data2<- matrix(unlist(data2))
+
+  unity <- function(x) {x / sqrt(sum(x^2))} ## normalize columns of a matrix to make correlation computation faster
+
+  ####################################################
+  ##Clean the marks and compute the smooth residuals##
+  ####################################################
+
+  data1 <- matrix(data1[round((0.01*nrow(data1))):round(0.99*nrow(data1)),], ncol = 1)
+  data2 <- matrix(data2[round((0.01*nrow(data2))):round(0.99*nrow(data2)),], ncol = 1)
+
+  ##Normalize the tool marks
+  y1 <- data1 - lowess(y = data1,  x = 1:nrow(data1), f= coarse)$y
+  y2 <- data2 - lowess(y = data2,  x = 1:nrow(data2), f= coarse)$y
+
+  ############################################
+  ##Compute the observed maximum correlation##
+  ############################################
+
+  #####################
+  ##Optimization step##
+  #####################
+  ##Each column in these matrices corresponds to a window in the respective tool mark
+  y1_mat_opt <- matrix(NA, ncol = length(1:(length(y1) - (window_opt - 1))), nrow = window_opt)
+  for(l in 1:(length(y1) - (window_opt - 1))){
+    y1_mat_opt[,l] <- y1[l:(l+(window_opt - 1))]
+  }
+  y2_mat_opt <- matrix(NA, ncol = length(1:(length(y2) - (window_opt - 1))), nrow = window_opt)
+  for(l in 1:(length(y2) - (window_opt - 1))){
+    y2_mat_opt[,l] <- y2[l:(l+(window_opt - 1))]
+  }
+
+  ##Compute the correlation between all pairs of windows for the two marks
+  ##Rows in the following matrix are mark 2, columns are mark 1
+  y2_mat_opt <- apply(scale(y2_mat_opt), 2, unity)
+  y1_mat_opt <- apply(scale(y1_mat_opt), 2, unity)
+  corr_mat_opt <- t(y2_mat_opt) %*% y1_mat_opt ##correlation matrix
+  max_corr_opt_loc <- which(corr_mat_opt == max(corr_mat_opt), arr.ind = TRUE) ##pair of windows maximizing the correlation
+
+sigs1.min = sigs_graphics[[1]] %>% filter(land_id==1) %>% select(y) %>% min()
+sigs2.min = sigs_graphics[[1]] %>% filter(land_id==276) %>% select(y) %>% min()
+window_val <- 2*window_val*0.645
+
+same.shift <- -(365-338)*1/.645 + sigs2.min - sigs1.min 
+xs <- seq(450, 2000, length=10)
+rects1 <- data.frame(ymin=-2.5, ymax= Inf, xmin = xs, xmax = xs+window_val)
+rects2 <- data.frame(ymin=-Inf, ymax=-3.5, xmin = xs+same.shift, xmax = same.shift+xs+window_val)
+rectsSS <- data.frame(rbind(rects1, rects2), type=rep(1:2, each=length(xs)))
+
+p1 <- sigs_graphics[[1]] %>% filter(land_id %in% c(1, 276)) %>% 
+  ggplot() + 
+  geom_segment(aes(x = xmin, y=ymin, xend = xmin, yend=ymax), 
+               colour="grey50", linetype=3, data = rectsSS) +
+  geom_segment(aes(x = xmax, y=ymin, xend = xmax, yend=ymax), 
+               colour="grey50", linetype=3, data = rectsSS) +
+  geom_segment(aes(x = xmin, y=ymin, xend = xmin+same.shift, yend=ymin-1), 
+               colour="grey50", linetype=3, data = rectsSS %>% filter(type==1)) +
+  geom_segment(aes(x = xmax, y=ymin, xend = xmax+same.shift, yend=ymin-1), 
+               colour="grey50", linetype=3, data = rectsSS %>% filter(type==1)) +
+  geom_rect(aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax, group=type), 
+            data=rectsSS, fill= "grey90") +
+  geom_line(aes(x = y, y=l30-7*I(land_id==276), group=land_id)) + 
+  theme_bw() +
+  theme(axis.title.y = element_blank(), 
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank()) +
+  xlab("Relative Location (in microns)") +
+  ggtitle("In-phase sample")
+
+diff.shift <- -(365-338)*1/.645 + sigs2.min - sigs1.min + 50 
+rects2 <- data.frame(ymin=-Inf, ymax=-3.5, xmin = xs+diff.shift, xmax = diff.shift+xs+window_val)
+rectsDS <- data.frame(rbind(rects1, rects2), type=rep(1:2, each=length(xs)))
+
+p2 <- sigs_graphics[[1]] %>% filter(land_id %in% c(1, 276)) %>% 
+  ggplot() + 
+  geom_segment(aes(x = xmin, y=ymin, xend = xmin, yend=ymax), 
+               colour="grey50", linetype=3, data = rectsDS) +
+  geom_segment(aes(x = xmax, y=ymin, xend = xmax, yend=ymax), 
+               colour="grey50", linetype=3, data = rectsDS) +
+  geom_segment(aes(x = xmin, y=ymin, xend = xmin+diff.shift, yend=ymin-1), 
+               colour="grey50", linetype=3, data = rectsDS %>% filter(type==1)) +
+  geom_segment(aes(x = xmax, y=ymin, xend = xmax+diff.shift, yend=ymin-1), 
+               colour="grey50", linetype=3, data = rectsDS %>% filter(type==1)) +
+  geom_rect(aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax, group=type), 
+            data=rectsDS, fill= "grey90") +
+  geom_line(aes(x = y, y=l30-7*I(land_id==276), group=land_id)) + 
+  theme_bw() +
+  theme(axis.title.y = element_blank(), 
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank()) +
+  xlab("Relative Location (in microns)") +
+  ggtitle("Out-of-phase sample")
+
+library(gridExtra)
+grid.arrange(p1,p2, ncol=2)
+```
+
+```{r win-comparison-2, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap=" The two plots on the left show how the same shift behaves in case of a matching pair and the two plots on the right show how the different shift behaves in case of a matching pair.", eval=FALSE}
 
@@
 In the original method proposed in \citet{chumbley} both in-phase and out-of-phase sample are extracted randomly, whereas \citet{hadler} proposed deterministic rules for both samples to make the resulting score deterministic while simultaneously avoiding overlaps within selected marks to ensure independence.
-
-A pre-processing step to the algorithm is to choose a coarseness value which  is used as a parameter to the lowess smoothing function.
-The coarseness essentially gives the proportion of points which influence the smooth at each value, which means larger values lead to more smoothness.
-The lowess smoothing is applied to each of two sets of vectorized striae or marks $x(t_1)$ and $y(t_2)$, before proceeding to the algorithm.
 

### 2018-03-17 doc/paper.tex
Chumbley score sketch

@@
 
-The algorithm works in two phases, namely, an optimization step and a
-validation step.
+In a pre-processing step the two markings are smoothed using a lowess
+\citep{lowess} with coarseness parameter \(c\). The purpose of this
+smoothing is to remove drift and (sub)class characteristics.
 
-In the optimization step, the two markings are aligned horizontally such
-that within a pre-defined window of length \(w_o\) the correlation
-between \(x(t_1)\) and \(y(t_2)\) is maximized: \[
+The Chumbley scores is calculated in two phases, namely, an optimization
+step and a validation step. In the optimization step, the two markings
+are aligned horizontally such that within a pre-defined window of length
+\(w_o\) the correlation between \(x(t_1)\) and \(y(t_2)\) is maximized:
+\[
 \left(t_1^o, t_2^o\right) = \mathop{\arg \max}\limits_{1 \le t_1 \le T_1, 1 \le t_2 \le T_2} \text{cor} \left(x (t_1, t_1 + w_o), y(t_2, t_2 + w_o) \right)
@@
 
-\caption{ The two plots on the left show how the same shift behaves in case of a matching pair and the two plots on the right show how the different shift behaves in case of a matching pair.}\label{fig:win-comparison}
+\caption{In- and out-of-phase samples from two markings created by the same source.
+On the left, sample windows are in-phase, on the right the pairs of sampled windows are out-of-phase.
+The Chumbley-score is based on a Mann-Whitney U test of the correlations derived from these two sets.}\label{fig:win-comparison}
 \end{figure}
@@
 marks to ensure independence.
-
-A pre-processing step to the algorithm is to choose a coarseness value
-which is used as a parameter to the lowess smoothing function. The
-coarseness essentially gives the proportion of points which influence
-the smooth at each value, which means larger values lead to more
-smoothness. The lowess smoothing is applied to each of two sets of
-vectorized striae or marks \(x(t_1)\) and \(y(t_2)\), before proceeding
-to the algorithm.
 

### 2018-03-18 doc/paper.Rmd
change of figure 2

--- 2018-03-17 doc/paper.Rmd
@@
 
-```{r win-comparison, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap=" The two plots on the left show how the same shift behaves in case of a matching pair and the two plots on the right show how the different shift behaves in case of a matching pair."}
+```{r win-comparison, warning=FALSE, fig.width=8, fig.height=3.5, fig.cap="In- and out-of-phase samples from two markings created by the same source.
+On the left, sample windows are in-phase, on the right the pairs of sampled windows are out-of-phase.
+Note that for convenience, the markings are moved into phase on the left and out-of phase on the right.
+The Chumbley-score is based on a Mann-Whitney U test of the correlations derived from these two sets."}
+sigs_graphics <- readRDS("../data/sigs_generate_window.rds")
+sigs_graphics$sigs_land_id_1_match <- sigs_graphics$sigs_land_id_1_match %>% group_by(land_id) %>% mutate(y = y-min(y, na.rm=T))
+
+sigs_graphics.match<- as.data.frame(sigs_graphics[1])
+sigs_graphics.nonmatch<- as.data.frame(sigs_graphics[2])
+colnames(sigs_graphics.match)<- gsub("^.*\\.","", colnames(sigs_graphics.match))
+colnames(sigs_graphics.nonmatch)<- gsub("^.*\\.","", colnames(sigs_graphics.nonmatch))
+
+# Matching signature
+d1<- sigs_graphics.match[which(sigs_graphics.match$land_id==1,arr.ind = TRUE),]
+d2<- sigs_graphics.match[which(sigs_graphics.match$land_id==276,arr.ind = TRUE),]
+data1<-d1 %>%select(l30)
+#data1<- data1$y
+data2<- d2 %>%select(l30)
+#data2<- data2$y
+window_opt = 120 
+window_val = 50
+coarse = 1
+data1<- matrix(unlist(data1))
+data2<- matrix(unlist(data2))
+
+  unity <- function(x) {x / sqrt(sum(x^2))} ## normalize columns of a matrix to make correlation computation faster
+
+  ####################################################
+  ##Clean the marks and compute the smooth residuals##
+  ####################################################
+
+  data1 <- matrix(data1[round((0.01*nrow(data1))):round(0.99*nrow(data1)),], ncol = 1)
+  data2 <- matrix(data2[round((0.01*nrow(data2))):round(0.99*nrow(data2)),], ncol = 1)
+
+  ##Normalize the tool marks
+  y1 <- data1 - lowess(y = data1,  x = 1:nrow(data1), f= coarse)$y
+  y2 <- data2 - lowess(y = data2,  x = 1:nrow(data2), f= coarse)$y
+
+  ############################################
+  ##Compute the observed maximum correlation##
+  ############################################
+
+  #####################
+  ##Optimization step##
+  #####################
+  ##Each column in these matrices corresponds to a window in the respective tool mark
+  y1_mat_opt <- matrix(NA, ncol = length(1:(length(y1) - (window_opt - 1))), nrow = window_opt)
+  for(l in 1:(length(y1) - (window_opt - 1))){
+    y1_mat_opt[,l] <- y1[l:(l+(window_opt - 1))]
+  }
+  y2_mat_opt <- matrix(NA, ncol = length(1:(length(y2) - (window_opt - 1))), nrow = window_opt)
+  for(l in 1:(length(y2) - (window_opt - 1))){
+    y2_mat_opt[,l] <- y2[l:(l+(window_opt - 1))]
+  }
+
+  ##Compute the correlation between all pairs of windows for the two marks
+  ##Rows in the following matrix are mark 2, columns are mark 1
+  y2_mat_opt <- apply(scale(y2_mat_opt), 2, unity)
+  y1_mat_opt <- apply(scale(y1_mat_opt), 2, unity)
+  corr_mat_opt <- t(y2_mat_opt) %*% y1_mat_opt ##correlation matrix
+  max_corr_opt_loc <- which(corr_mat_opt == max(corr_mat_opt), arr.ind = TRUE) ##pair of windows maximizing the correlation
+
+sigs1.min = sigs_graphics[[1]] %>% filter(land_id==1) %>% select(y) %>% min()
+sigs2.min = sigs_graphics[[1]] %>% filter(land_id==276) %>% select(y) %>% min()
+window_val <- 3*window_val*0.645
+
+same.shift <- -(365-338)*1/.645 + sigs2.min - sigs1.min 
+xs <- seq(100, 1800, length=10)
+rects <- data.frame(ymin=-Inf, ymax= Inf, xmin = xs, xmax = xs+window_val)
+
+p1 <- sigs_graphics[[1]] %>% filter(land_id %in% c(1, 276)) %>% 
+  ggplot() + 
+  geom_segment(aes(x = xmin, y=ymin, xend = xmin, yend=ymax), 
+               colour="grey50", linetype=3, data = rects) +
+  geom_segment(aes(x = xmax, y=ymin, xend = xmax, yend=ymax), 
+               colour="grey50", linetype=3, data = rects) +
+  geom_rect(aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax), 
+            data=rects, fill= "grey90") +
+  geom_line(aes(x = y-same.shift*I(land_id==276), 
+                y=l30-5*I(land_id==276), group=land_id)) + 
+  theme_bw() +
+  theme(axis.title.y = element_blank(), 
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank()) +
+  xlab("Relative Location (in microns)") +
+  ggtitle("In-phase sample")
+
+diff.shift <- -(365-338)*1/.645 + sigs2.min - sigs1.min + 70
+
+p2 <- sigs_graphics[[1]] %>% filter(land_id %in% c(1, 276)) %>% 
+  ggplot() + 
+  geom_segment(aes(x = xmin, y=ymin, xend = xmin, yend=ymax), 
+               colour="grey50", linetype=3, data = rects) +
+  geom_segment(aes(x = xmax, y=ymin, xend = xmax, yend=ymax), 
+               colour="grey50", linetype=3, data = rects) +
+  geom_rect(aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax), 
+            data=rects, fill= "grey90") +
+  geom_line(aes(x = y-diff.shift*I(land_id==276), 
+                y=l30-5*I(land_id==276), group=land_id)) + 
+  theme_bw() +
+  theme(axis.title.y = element_blank(), 
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank()) +
+  xlab("Relative Location (in microns)") +
+  ggtitle("Out-of-phase sample")
+
+library(gridExtra)
+grid.arrange(p1,p2, ncol=2)
+```
+
+```{r win-comparison-2, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap=" The two plots on the left show how the same shift behaves in case of a matching pair and the two plots on the right show how the different shift behaves in case of a matching pair.", eval=FALSE}
 

### 2018-03-18 doc/paper.Rmd
profile signature

--- 2018-03-17 doc/paper.Rmd
@@
 
-```{r win-comparison, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap=" The two plots on the left show how the same shift behaves in case of a matching pair and the two plots on the right show how the different shift behaves in case of a matching pair."}
+```{r win-comparison, warning=FALSE, fig.width=8, fig.height=3.5, fig.cap="Two markings made by the same source.
+For convenience, the markings are moved into phase on the left and out-of phase on the right.
+In-phase (left) and out-of-phase (right) samples are shown by the light grey background.
+The Chumbley-score is based on a Mann-Whitney U test of the correlations derived from these two sets of samples."}
+sigs_graphics <- readRDS("../data/sigs_generate_window.rds")
+sigs_graphics$sigs_land_id_1_match <- sigs_graphics$sigs_land_id_1_match %>% group_by(land_id) %>% mutate(y = y-min(y, na.rm=T))
+
+sigs_graphics.match<- as.data.frame(sigs_graphics[1])
+sigs_graphics.nonmatch<- as.data.frame(sigs_graphics[2])
+colnames(sigs_graphics.match)<- gsub("^.*\\.","", colnames(sigs_graphics.match))
+colnames(sigs_graphics.nonmatch)<- gsub("^.*\\.","", colnames(sigs_graphics.nonmatch))
+
+# Matching signature
+d1<- sigs_graphics.match[which(sigs_graphics.match$land_id==1,arr.ind = TRUE),]
+d2<- sigs_graphics.match[which(sigs_graphics.match$land_id==276,arr.ind = TRUE),]
+data1<-d1 %>%select(l30)
+#data1<- data1$y
+data2<- d2 %>%select(l30)
+#data2<- data2$y
+window_opt = 120 
+window_val = 50
+coarse = 1
+data1<- matrix(unlist(data1))
+data2<- matrix(unlist(data2))
+
+  unity <- function(x) {x / sqrt(sum(x^2))} ## normalize columns of a matrix to make correlation computation faster
+
+  ####################################################
+  ##Clean the marks and compute the smooth residuals##
+  ####################################################
+
+  data1 <- matrix(data1[round((0.01*nrow(data1))):round(0.99*nrow(data1)),], ncol = 1)
+  data2 <- matrix(data2[round((0.01*nrow(data2))):round(0.99*nrow(data2)),], ncol = 1)
+
+  ##Normalize the tool marks
+  y1 <- data1 - lowess(y = data1,  x = 1:nrow(data1), f= coarse)$y
+  y2 <- data2 - lowess(y = data2,  x = 1:nrow(data2), f= coarse)$y
+
+  ############################################
+  ##Compute the observed maximum correlation##
+  ############################################
+
+  #####################
+  ##Optimization step##
+  #####################
+  ##Each column in these matrices corresponds to a window in the respective tool mark
+  y1_mat_opt <- matrix(NA, ncol = length(1:(length(y1) - (window_opt - 1))), nrow = window_opt)
+  for(l in 1:(length(y1) - (window_opt - 1))){
+    y1_mat_opt[,l] <- y1[l:(l+(window_opt - 1))]
+  }
+  y2_mat_opt <- matrix(NA, ncol = length(1:(length(y2) - (window_opt - 1))), nrow = window_opt)
+  for(l in 1:(length(y2) - (window_opt - 1))){
+    y2_mat_opt[,l] <- y2[l:(l+(window_opt - 1))]
+  }
+
+  ##Compute the correlation between all pairs of windows for the two marks
+  ##Rows in the following matrix are mark 2, columns are mark 1
+  y2_mat_opt <- apply(scale(y2_mat_opt), 2, unity)
+  y1_mat_opt <- apply(scale(y1_mat_opt), 2, unity)
+  corr_mat_opt <- t(y2_mat_opt) %*% y1_mat_opt ##correlation matrix
+  max_corr_opt_loc <- which(corr_mat_opt == max(corr_mat_opt), arr.ind = TRUE) ##pair of windows maximizing the correlation
+
+sigs1.min = sigs_graphics[[1]] %>% filter(land_id==1) %>% select(y) %>% min()
+sigs2.min = sigs_graphics[[1]] %>% filter(land_id==276) %>% select(y) %>% min()
+window_val <- 3*window_val*0.645
+
+same.shift <- -(365-338)*1/.645 + sigs2.min - sigs1.min 
+xs <- seq(100, 1800, length=10)
+rects <- data.frame(ymin=-Inf, ymax= Inf, xmin = xs, xmax = xs+window_val)
+
+p1 <- sigs_graphics[[1]] %>% filter(land_id %in% c(1, 276)) %>% 
+  ggplot() + 
+  geom_segment(aes(x = xmin, y=ymin, xend = xmin, yend=ymax), 
+               colour="grey50", linetype=3, data = rects) +
+  geom_segment(aes(x = xmax, y=ymin, xend = xmax, yend=ymax), 
+               colour="grey50", linetype=3, data = rects) +
+  geom_rect(aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax), 
+            data=rects, fill= "grey90") +
+  geom_line(aes(x = y-same.shift*I(land_id==276), 
+                y=l30-5*I(land_id==276), group=land_id)) + 
+  theme_bw() +
+  theme(axis.title.y = element_blank(), 
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank()) +
+  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
+  ggtitle("In-phase sample")
+
+diff.shift <- -(365-338)*1/.645 + sigs2.min - sigs1.min + 70
+
+p2 <- sigs_graphics[[1]] %>% filter(land_id %in% c(1, 276)) %>% 
+  ggplot() + 
+  geom_segment(aes(x = xmin, y=ymin, xend = xmin, yend=ymax), 
+               colour="grey50", linetype=3, data = rects) +
+  geom_segment(aes(x = xmax, y=ymin, xend = xmax, yend=ymax), 
+               colour="grey50", linetype=3, data = rects) +
+  geom_rect(aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax), 
+            data=rects, fill= "grey90") +
+  geom_line(aes(x = y-diff.shift*I(land_id==276), 
+                y=l30-5*I(land_id==276), group=land_id)) + 
+  theme_bw() +
+  theme(axis.title.y = element_blank(), 
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank()) +
+  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
+  ggtitle("Out-of-phase sample")
+
+library(gridExtra)
+grid.arrange(p1,p2, ncol=2)
+```
+
+```{r win-comparison-2, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap=" The two plots on the left show how the same shift behaves in case of a matching pair and the two plots on the right show how the different shift behaves in case of a matching pair.", eval=FALSE}
 
@@
 -->
-
+<!--
 ##################
@@
 - discuss cross section, profiles and signatures
+-->
 
@@
 
-# Simulation setup
+# Testing setup
+
+## The Data
+
+\hh{More on NIST database, some on CSAFE}
+
+\hh{Introduce profiles and signatures, as shown in figure \ref{fig:sigs-profiles}.}
+
+```{r sigs-profiles, fig.width = 8, fig.height = 2.5, fig.cap="Bullet land profile (left) and the corresponding signature (right)."}
+p1 <- sigs_graphics[[1]] %>% filter(land_id==19) %>%
+  ggplot(aes(x = y, y = value)) + geom_line() +
+  theme_bw() +
+  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
+  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
+  ggtitle("Profile")
+
+p2 <- sigs_graphics[[1]] %>% filter(land_id==19) %>%
+  ggplot(aes(x = y, y = l30)) + 
+  geom_line(aes(y=resid), size=.25) +
+  geom_line(aes(y=l30)) +
+  theme_bw() +
+  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
+  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
+  ggtitle("Signature")
+
+grid.arrange(p1,p2, ncol=2)
+```
+
+## Setup
+
+\hh{using (a) signatures and (b) profiles, run chumbley score across scans from NIST and CSAFE for various settings of $w_o$ and $w_v$ (and coarseness $c$ for profiles).}
+
+## Results
 

### 2018-03-18 doc/paper.tex
change of figure 2
profile signature

@@
 
-\caption{In- and out-of-phase samples from two markings created by the same source.
-On the left, sample windows are in-phase, on the right the pairs of sampled windows are out-of-phase.
-The Chumbley-score is based on a Mann-Whitney U test of the correlations derived from these two sets.}\label{fig:win-comparison}
+\caption{Two markings made by the same source.
+For convenience, the markings are moved into phase on the left and out-of phase on the right.
+In-phase (left) and out-of-phase (right) samples are shown by the light grey background.
+The Chumbley-score is based on a Mann-Whitney U test of the correlations derived from these two sets of samples.}\label{fig:win-comparison}
 \end{figure}
@@
 
-\paragraph{endedit}\label{endedit}
+\section{Testing setup}\label{testing-setup}
 
-\hh{end of intro: remaining paper is structured as follows: introduce to the data we get from confocal microscopy, introduce to profiles and signatures.
-Introduce to chumbley score method, apply chumbley and discuss results ....
+\subsection{The Data}\label{the-data}
+
+\hh{More on NIST database, some on CSAFE}
+
+\hh{Introduce profiles and signatures, as shown in figure \ref{fig:sigs-profiles}.}
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figures/sigs-profiles-1}
+
 }
 
-\subsection{Scans for land engraved
-areas}\label{scans-for-land-engraved-areas-1}
+\caption{Bullet land profile (left) and the corresponding signature (right).}\label{fig:sigs-profiles}
+\end{figure}
 
-\begin{itemize}
-\tightlist
-\item
-  scans available: NIST database (citation), Hamby 44 and Hamby 252
-  (Hamby citation)
-\item
-  move figure up
-\item
-  discuss cross section, profiles and signatures
-\end{itemize}
+\subsection{Setup}\label{setup}
 
-\section{Simulation setup}\label{simulation-setup}
+\hh{using (a) signatures and (b) profiles, run chumbley score across scans from NIST and CSAFE for various settings of $w_o$ and $w_v$ (and coarseness $c$ for profiles).}
+
+\subsection{Results}\label{results}
 
@@
 
-\section{Results}\label{results}
+\section{Results}\label{results-1}
 

### 2018-03-20 doc/paper.Rmd
my citation problem

@@
 
-```{r sigs-profiles, fig.width = 8, fig.height = 2.5, fig.cap="Bullet land profile (left) and the corresponding signature (right)."}
+```{r sigs-profiles, fig.width = 8, fig.height = 2.5, fig.cap="Bullet land profile (left) and the corresponding signature (right) for one of the lands of Hamby-44."}
 p1 <- sigs_graphics[[1]] %>% filter(land_id==19) %>%
@@
 
+Lands for all Hamby-44 and Hamby-252 scans are made available through the NIST ballistics database \citep{nist} and were considered, here.
+Both of these sets of scans are part of the larger Hamby study \citep{hamby} and each consist of twenty known bullets (two each from ten consecutively rifled Ruger P85 barrels) and fifteen questioned bullets (each matching one of the ten barrels).
+Ground truth for both of these Hamby sets is known and was used to assess correctness of the tests results.
+
+Profiles and signatures were extracted from the Hamby 44 and Hamby 252 data as described in \citet{aoas}.
+
 ## Setup
@@
 
+We used the adjusted Chumbley method as proposed in \citet{hadler} and implemented in the R package `toolmaRk` \citep{toolmark} on all pairwise land-to-land comparisons of the Hamby scans provided by NIST (a total of 85,491 comparisons).
+The settings for optimizing and validating window sizes, $w_o$ and $w_v$, ranged from $w_o \in [50, 280]$ and $w_v \in \{30, 50\}$, see also figure \ref{fig:type2}.
+
 ## Results
+
+\hh{For signatures from NIST scans we see three problems: 
+\begin{enumerate}
+\item type-2 error rate is at best 30\% for a type-1 error rate of 5\%, which is well above the error rates we see for tool marks from screw drivers,
+\item the observed type-1 error, which generally close to the nominal type-1 error rate, depends on the size of the optimization window: as the window size increases, the observed type-1 error decreases,
+\item the Chumbley-score fails to provide a result for up to 3\% of the cases.
+This drop-out rate is considerably higher when the two lands are from different sources than when the lands are from the same source.
+\end{enumerate}
+}
 
@@
 ### Signatures
+
+```{r type2, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap="Type II error rates observed across a range of window sizes for optimization $w_o$.
+For a window size of $w_o = 120$ we see a drop in type II error rate across all type I rates considered.
+Smaller validation sizes $w_v$ are typically associated with a smaller type II error."}
+errors %>% filter(wv %in% c(30,50)) %>%
+  ggplot(aes(x = wo, y = beta, colour=factor(alpha))) + 
+  geom_point(aes(shape=factor(wv))) +
+#  facet_grid(.~wo, labeller="label_both") + 
+  geom_smooth(aes(group=alpha), se=FALSE, method="loess") +
+  theme_bw() +
+  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") +
+  scale_shape_discrete(expression("Size of\nvalidation\nwindow "~w[v])) +
+  xlab(expression("Window size for optimization "~w[o])) +
+  ylab("Type II error rate")
+
+labels = expression(P[M1](tilde(z)>0),P[M0](tilde(z)>0))
+```
 
@@
 The windows represent areas to the left of the window with minimum type II, near the minimum type II window and to the far right of the minimum type II error
-
-```{r type2, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap="Type II error rates observed across a range of window sizes for optimization $wo$.
-For a window size of $wo = 120$ we see a drop in type II error rate across all type I rates considered.
-Smaller validation sizes $wv$ are typically associated with a smaller type II error."}
-errors %>% filter(wv %in% c(30,50)) %>%
-  ggplot(aes(x = wo, y = beta, colour=factor(alpha))) + 
-  geom_point(aes(shape=factor(wv))) +
-#  facet_grid(.~wo, labeller="label_both") + 
-  geom_smooth(aes(group=alpha), se=FALSE, method="loess") +
-  theme_bw() +
-  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") +
-  scale_shape_discrete("Size of validation\nwindow, wv") +
-  xlab(expression("Window size for optimization, "~wo)) +
-  ylab("Type II error rate")
-
-labels = expression(P[M1](tilde(z)>0),P[M0](tilde(z)>0))
-```
 

### 2018-03-20 doc/paper.tex
my citation problem

@@
 
-\caption{Bullet land profile (left) and the corresponding signature (right).}\label{fig:sigs-profiles}
+\caption{Bullet land profile (left) and the corresponding signature (right) for one of the lands of Hamby-44.}\label{fig:sigs-profiles}
 \end{figure}
+
+Lands for all Hamby-44 and Hamby-252 scans are made available through
+the NIST ballistics database \citep{nist} and were considered, here.
+Both of these sets of scans are part of the larger Hamby study
+\citep{hamby} and each consist of twenty known bullets (two each from
+ten consecutively rifled Ruger P85 barrels) and fifteen questioned
+bullets (each matching one of the ten barrels). Ground truth for both of
+these Hamby sets is known and was used to assess correctness of the
+tests results.
+
+Profiles and signatures were extracted from the Hamby 44 and Hamby 252
+data as described in \citet{aoas}.
 
@@
 
+We used the adjusted Chumbley method as proposed in \citet{hadler} and
+implemented in the R package \texttt{toolmaRk} \citep{toolmark} on all
+pairwise land-to-land comparisons of the Hamby scans provided by NIST (a
+total of 85,491 comparisons). The settings for optimizing and validating
+window sizes, \(w_o\) and \(w_v\), ranged from \(w_o \in [50, 280]\) and
+\(w_v \in \{30, 50\}\), see also figure \ref{fig:type2}.
+
 \subsection{Results}\label{results}
+
+\hh{For signatures from NIST scans we see three problems: 
+\begin{enumerate}
+\item type-2 error rate is at best 30\% for a type-1 error rate of 5\%, which is well above the error rates we see for tool marks from screw drivers,
+\item the observed type-1 error, which generally close to the nominal type-1 error rate, depends on the size of the optimization window: as the window size increases, the observed type-1 error decreases,
+\item the Chumbley-score fails to provide a result for up to 3\% of the cases.
+This drop-out rate is considerably higher when the two lands are from different sources than when the lands are from the same source.
+\end{enumerate}
+}
 
@@
 \subsubsection{Signatures}\label{signatures}
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figures/type2-1}
+
+}
+
+\caption{Type II error rates observed across a range of window sizes for optimization $w_o$.
+For a window size of $w_o = 120$ we see a drop in type II error rate across all type I rates considered.
+Smaller validation sizes $w_v$ are typically associated with a smaller type II error.}\label{fig:type2}
+\end{figure}
 
@@
 window and to the far right of the minimum type II error
-
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figures/type2-1}
-
-}
-
-\caption{Type II error rates observed across a range of window sizes for optimization $wo$.
-For a window size of $wo = 120$ we see a drop in type II error rate across all type I rates considered.
-Smaller validation sizes $wv$ are typically associated with a smaller type II error.}\label{fig:type2}
-\end{figure}
 

### 2018-03-20 doc/paper.Rmd
discussion of results

@@
                 errorrate(all, 0.05))
+
+ns <- all %>% group_by(wv, wo, match) %>% summarize(
+  n_ss = mean(same_shift_n),
+  n_ds = mean(diff_shift_n),
+  zero_ss = sum(same_shift_n==0),
+  zero_ds = sum(diff_shift_n==0),
+  miss = sum(is.na(p_value)),
+  missperc = miss/n()*100
+)
 ```
@@
 \begin{enumerate}
-\item type-2 error rate is at best 30\% for a type-1 error rate of 5\%, which is well above the error rates we see for tool marks from screw drivers,
-\item the observed type-1 error, which generally close to the nominal type-1 error rate, depends on the size of the optimization window: as the window size increases, the observed type-1 error decreases,
+\item type-2 error rate is at best 30\% for a type-1 error rate of 5\%, which is well above the error rates we see for tool marks from screw drivers, see figure \ref{fig:type2};
+\item the observed type-1 error, which generally close to the nominal type-1 error rate, depends on the size of the optimization window: as the window size increases, the observed type-1 error decreases, see figure \ref{fig:type1};
 \item the Chumbley-score fails to provide a result for up to 3\% of the cases.
-This drop-out rate is considerably higher when the two lands are from different sources than when the lands are from the same source.
+This drop-out rate is considerably higher when the two lands are from same source than when the lands are from different sources, see figure \ref{fig:missings}.
 \end{enumerate}
@@
 
-```{r type2, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap="Type II error rates observed across a range of window sizes for optimization $w_o$.
+```{r type2, fig.width=8, fig.height=5, out.width='.8\\textwidth', fig.cap="Type II error rates observed across a range of window sizes for optimization $w_o$.
 For a window size of $w_o = 120$ we see a drop in type II error rate across all type I rates considered.
@@
   ggplot(aes(x = wo, y = beta, colour=factor(alpha))) + 
-  geom_point(aes(shape=factor(wv))) +
+  geom_point(aes(shape=factor(wv)), size=3.5) +
 #  facet_grid(.~wo, labeller="label_both") + 
-  geom_smooth(aes(group=alpha), se=FALSE, method="loess") +
+  geom_smooth(aes(group=alpha), se=FALSE, method="loess", size=.7) +
   theme_bw() +
   scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") +
-  scale_shape_discrete(expression("Size of\nvalidation\nwindow "~w[v])) +
+  scale_shape_discrete(expression("Size of\nvalidation window "~w[v])) +
   xlab(expression("Window size for optimization "~w[o])) +
@@
 
+```{r type1, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap="Comparison of observed and nominal type I error rates  across a range of window sizes for optimization $wo$.
+The horizontal line in each facet indicates the nominal type I error rate."}
+errors %>% filter(wv %in% c(30,50)) %>%
+  ggplot(aes(x = wo, y = actual, colour=factor(alpha))) + 
+  geom_hline(aes(yintercept=alpha), colour="grey30", size=.5) +
+  geom_point(aes(shape=factor(wv)), size=2.5) +
+  #  facet_grid(.~wo, labeller="label_both") + 
+  geom_smooth(aes(group=alpha), size=.8, se=FALSE, method="lm", alpha=.9) +
+  theme_bw() +
+#  scale_y_log10(breaks=c(0.001,.005, 0.01, 0.05)) +
+  scale_shape_discrete(expression("Size of validation window "~w[v])) +
+  ylab("Observed type I error rate") +
+  xlab(expression("Window size for optimization, "~w[o])) +
+  scale_colour_brewer(expression("Nominal type I error "~alpha), palette="Set2") +
+  facet_wrap(~alpha, labeller="label_both", scales="free") +
+  theme(legend.position="bottom")
+
+```
+
+```{r missings, out.width='\\textwidth', fig.width=8, fig.height=4, fig.cap="The number of failed tests increases with an increase in the size of the optimization window (left).
+Unfortunately there is also a dependency between failed tests and ground truth.
+The plot on the right shows the ratio of the number of land pairs from same sources and different sources for failed tests.
+For small optimization windows and large windows the number of failed tests for same-source land-to-land comparisons is increasing.
+Even in the minimum, same-source land-to-land comparisons fail at twice the rate that they are expected to based on the ratio of the number of known matches and known non-matches (horizontal line).",fig.align = "center"}
+#ns1<- ns[which(ns$wv == c(30,50), arr.ind = T),]
+
+fails <- all %>% group_by(wv, wo) %>% summarize(
+  miss = sum(is.na(p_value)),
+  missperc = miss/n()*100
+)
+
+failsbymatch <- all %>% group_by(wv, wo, match) %>% summarize(
+  fails = sum(is.na(p_value))
+) %>% mutate(
+    ratio = fails[match==TRUE]/fails[match==FALSE],
+    match = c("Different Source", "Same Source")[as.numeric(as.logical(match)) + 1]
+    )
+
+missings <- fails %>% filter(wv %in% c(30, 50)) %>%
+  ggplot(aes(x = wo, y = missperc, 
+             colour = factor(wv),
+             shape=factor(wv)), size=.3) + 
+  geom_smooth(method="lm", size=.5, se=FALSE) +
+  geom_point(size=2.5) + 
+  theme_bw() +
+  ylab("Percent of failed tests.") +
+  xlab(expression("Size of optimization window "~w[o])) +
+  scale_colour_brewer(expression("Size of validation window "~w[v]), palette="Set1") +
+  scale_shape(expression("Size of validation window "~w[v])) +
+    theme(legend.position="bottom")
+
+missingsbymatch <- failsbymatch %>% 
+  filter(wv %in% c(30, 50),
+         match=="Same Source") %>%
+  ggplot(aes(x = wo, y = ratio, 
+             colour = factor(wv),
+             shape=factor(wv)), size=.3) + 
+  geom_hline(yintercept = 43960/2948225, size=0.5, colour="grey30") +
+  geom_smooth(method="loess", span=0.8, size=.5, se=FALSE) +
+  geom_point(size=2.5) + 
+  theme_bw() +
+  ylab("Ratio of same/different source\namong failed tests") +
+  xlab(expression("Size of optimization window "~w[o])) +
+  scale_colour_brewer(expression("Size of validation window "~w[v]), palette="Set1") +
+  scale_shape(expression("Size of validation window "~w[v])) +
+    theme(legend.position="bottom") +
+  scale_y_continuous(breaks=c(0.015, 0.03, 0.06, 0.09))
+
+gridExtra::grid.arrange(missings, missingsbymatch, ncol=2)
+
+```
 Signatures of lands for all Hamby-44 and Hamby-252 scans made available through the NIST ballistics database \citep{nist} were considered.
@@
 
-```{r type1, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap="Comparison of observed and nominal type I error rates  across a range of window sizes for optimization $wo$.
-The horizontal line in each facet indicates the nominal type I error rate."}
-errors %>% filter(wv %in% c(30,50)) %>%
-  ggplot(aes(x = wo, y = actual, colour=factor(alpha))) + 
-  geom_hline(aes(yintercept=alpha), colour="grey30") +
-  geom_point(aes(shape=factor(wv))) +
-  #  facet_grid(.~wo, labeller="label_both") + 
-  geom_smooth(aes(group=alpha), se=FALSE, method="lm") +
-  theme_bw() +
-#  scale_y_log10(breaks=c(0.001,.005, 0.01, 0.05)) +
-  scale_shape_discrete("Size of validation window, wv") +
-  ylab("Observed type I error rate") +
-  xlab(expression("Window size for optimization, "~wo)) +
-  scale_colour_brewer(expression("Nominal type I error "~alpha), palette="Set2") +
-  facet_wrap(~alpha, labeller="label_both", scales="free") +
-  theme(legend.position="bottom")
-
-```
-
 ```{r wo120, fig.width = 4, fig.height=4, fig.cap="The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120"}
@@
 ```{r }
-ns <- all %>% group_by(wv, wo, match) %>% summarize(
-  n_ss = mean(same_shift_n),
-  n_ds = mean(diff_shift_n),
-  zero_ss = sum(same_shift_n==0),
-  zero_ds = sum(diff_shift_n==0),
-  miss = sum(is.na(p_value)),
-  missperc = miss/85491*100
-)
 
@@
 
-```
-
-```{r missings, fig.width=6, fig.height=4, fig.cap="Number of failed tests by the window optimization size, wo, and ground truth.",fig.align = "center"}
-#ns1<- ns[which(ns$wv == c(30,50), arr.ind = T),]
-missings<- ns %>% filter(wv %in% c(30, 50)) %>%
-#ns %>% filter(wv == c(30, 50)) %>%
-  ggplot(aes(x = wo, y = missperc, shape=factor(wv), colour = factor(match)), size=.3) + 
-  geom_smooth(method="lm", size=.2, se=FALSE) +
-  geom_point() + 
-  theme_bw() +
-  ylab("Percent of failed tests.") +
-  xlab("Size of optimization window, wo") +
-  scale_colour_brewer("Match", palette="Set1") +
-  scale_shape_discrete("Size of \nvalidation\nwindow, wv")
-missings
 ```

### 2018-03-20 doc/paper.tex
discussion of results

@@
 
-{\centering \includegraphics[width=\textwidth]{figures/type2-1}
+{\centering \includegraphics[width=.8\textwidth]{figures/type2-1}
 
@@
 Smaller validation sizes $w_v$ are typically associated with a smaller type II error.}\label{fig:type2}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figures/type1-1}
+
+}
+
+\caption{Comparison of observed and nominal type I error rates  across a range of window sizes for optimization $wo$.
+The horizontal line in each facet indicates the nominal type I error rate.}\label{fig:type1}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figures/missings-1}
+
+}
+
+\caption{The number of failed tests increases with an increase in the size of the optimization window (left).
+Unfortunately there is also a dependency between failed tests and ground truth.
+The plot on the right shows the ratio of the number of land pairs from same sources and different sources for failed tests.
+For small optimization windows and large windows the number of failed tests for same-source land-to-land comparisons is increasing.
+Even in the minimum, same-source land-to-land comparisons fail at twice the rate that they are expected to based on the ratio of the number of known matches and known non-matches (horizontal line).}\label{fig:missings}
 \end{figure}
@@
 
-{\centering \includegraphics[width=\textwidth]{figures/type1-1}
-
-}
-
-\caption{Comparison of observed and nominal type I error rates  across a range of window sizes for optimization $wo$.
-The horizontal line in each facet indicates the nominal type I error rate.}\label{fig:type1}
-\end{figure}
-
-\begin{figure}
-
 {\centering \includegraphics[width=\textwidth]{figures/wo120-1}
@@
 
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figures/missings-1}
-
-}
-
-\caption{Number of failed tests by the window optimization size, wo, and ground truth.}\label{fig:missings}
-\end{figure}
-
 \begin{table}
@@
 \midrule
-FALSE & 30 & 0.503 & 0.028\\
-FALSE & 50 & 0.922 & 0.035\\
-TRUE & 30 & 0.043 & 0.004\\
-TRUE & 50 & 0.056 & 0.005\\
+FALSE & 30 & 0.510 & 0.029\\
+FALSE & 50 & 0.936 & 0.036\\
+TRUE & 30 & 2.933 & 0.247\\
+TRUE & 50 & 3.843 & 0.321\\
 \bottomrule

### 2018-03-22 doc/paper.Rmd
couple of edits

@@
 
-\hh{More on NIST database, some on CSAFE}
-
 \hh{Introduce profiles and signatures, as shown in figure \ref{fig:sigs-profiles}.}
@@
 
-Lands for all Hamby-44 and Hamby-252 scans are made available through the NIST ballistics database \citep{nist} and were considered, here.
-Both of these sets of scans are part of the larger Hamby study \citep{hamby} and each consist of twenty known bullets (two each from ten consecutively rifled Ruger P85 barrels) and fifteen questioned bullets (each matching one of the ten barrels).
+Lands for all Hamby-44 and Hamby-252 scans are made available through the NIST ballistics database \citep{nist} and are considered, here.
+Both of these sets of scans are part of the larger Hamby study \citep{hamby}.
+Each set consists of twenty known bullets (two each from ten consecutively rifled Ruger P85 barrels) and fifteen questioned bullets (each matching one of the ten barrels).
 Ground truth for both of these Hamby sets is known and was used to assess correctness of the tests results.
 
-Profiles and signatures were extracted from the Hamby 44 and Hamby 252 data as described in \citet{aoas}.
+\hh{Discuss CSAFE scans}
+
+Profiles and signatures were extracted from all scans as described in \citet{aoas}.
 
@@
 \item the Chumbley-score fails to provide a result for up to 3\% of the cases.
-This drop-out rate is considerably higher when the two lands are from same source than when the lands are from different sources, see figure \ref{fig:missings}.
+The number of failed tests increases linearly in the size used for the window in the optimization step.
+The rate of failed tests is considerably higher when the two lands are from same source than when the lands are from different sources, see figure \ref{fig:missings}.
 \end{enumerate}
@@
 
+Figure \ref{fig:missings} gives an overview of the number of failed tests, i.e.
+tests in which a particular parameter setting did not return a valid result.
+This happens e.g.\ when the shift to align two markings is so large, that the remaining overlap is too small to accommodate  windows for validation.
+The problem is therefore exacerbated by a larger validation window.
+Figure \ref{fig:missings}(left) also shows that the number of failed tests is approximately linear in the size of the optimization window.
+Tests also fail at a higher rate than expected when the markings are from the same source (right).
+This difference is the least pronounced around an optimized window size $w_o$ of around 120.
+However, even in this scenario, the number of failed tests for markings from the same source is about twice as high as expected given the number of same source and different source pairings in the data set.
+
+<!--
 Following on similar lines to the setup of toolmarks, the first step here is to first identify what difference does different window sizes of optimization and the validation step have, when adapting the toolmark method to bullets.
@@
 ### Signatures
+-->
 
@@
 ```
+
+<!--
 Signatures of lands for all Hamby-44 and Hamby-252 scans made available through the NIST ballistics database \citep{nist} were considered.
@@
 The idea therefore is to first use these raw values of the profile directly in the chumbley algorithm, and see how the algorithm performs for different coarseness values (smoothing parameter as referred in the function lowess used in the chumbley algorithm).
-
-\pagebreak
+-->
 
 # Results
+<!--
 We used the adjusted Chumbley method as proposed in \citet{hadler} and implemented in the R package `toolmaRk` \citep{toolmark} on all pairwise land-to-land comparisons of the Hamby scans (a total of 85,491 comparisons) with the pairwise sets for the comparisons given in the table \ref{tab:param}.
@@
 ## Signatures
-
+-->
+<!--
 Figure \ref{fig:type2} gives an overview of type II error rates observed when varying the window size in the optimization step.
@@
 ```
+-->
+<!--
 Figure \ref{fig:missings} gives an overview of the number of failed tests, i.e.
@@
 For known non-matches there is a higher possibility that in the optimization pair of windows where cross-correlations are maximum are too far apart, and same shifts of this order hit the end of the signature.
+-->
 

### 2018-03-22 doc/paper.tex
couple of edits

@@
 Lands for all Hamby-44 and Hamby-252 scans are made available through
-the NIST ballistics database \citep{nist} and were considered, here.
-Both of these sets of scans are part of the larger Hamby study
-\citep{hamby} and each consist of twenty known bullets (two each from
-ten consecutively rifled Ruger P85 barrels) and fifteen questioned
-bullets (each matching one of the ten barrels). Ground truth for both of
-these Hamby sets is known and was used to assess correctness of the
-tests results.
+the NIST ballistics database \citep{nist} and are considered, here. Both
+of these sets of scans are part of the larger Hamby study \citep{hamby}.
+Each set consists of twenty known bullets (two each from ten
+consecutively rifled Ruger P85 barrels) and fifteen questioned bullets
+(each matching one of the ten barrels). Ground truth for both of these
+Hamby sets is known and was used to assess correctness of the tests
+results.
 
-Profiles and signatures were extracted from the Hamby 44 and Hamby 252
-data as described in \citet{aoas}.
+\hh{Discuss CSAFE scans}
+
+Profiles and signatures were extracted from all scans as described in
+\citet{aoas}.
 
@@
 
-Following on similar lines to the setup of toolmarks, the first step
-here is to first identify what difference does different window sizes of
-optimization and the validation step have, when adapting the toolmark
-method to bullets.
-
-The marking made on bullets are smaller than toolmarks and is also less
-wider. The idea is to find out possible areas of error while adapting
-the score based method proposed for toolmarks, using cross-validation
-setup to identify appropriate parameter settings for (a) signatures and
-(b) profiles directly
-
-\subsubsection{Signatures}\label{signatures}
+Figure \ref{fig:missings} gives an overview of the number of failed
+tests, i.e.~tests in which a particular parameter setting did not return
+a valid result. This happens e.g.~when the shift to align two markings
+is so large, that the remaining overlap is too small to accommodate
+windows for validation. The problem is therefore exacerbated by a larger
+validation window. Figure \ref{fig:missings}(left) also shows that the
+number of failed tests is approximately linear in the size of the
+optimization window. Tests also fail at a higher rate than expected when
+the markings are from the same source (right). This difference is the
+least pronounced around an optimized window size \(w_o\) of around 120.
+However, even in this scenario, the number of failed tests for markings
+from the same source is about twice as high as expected given the number
+of same source and different source pairings in the data set.
 
@@
 
-Signatures of lands for all Hamby-44 and Hamby-252 scans made available
-through the NIST ballistics database \citep{nist} were considered. Both
-of these sets of scans are part of the larger Hamby study \citep{hamby}
-and each consist of twenty known bullets (two each from ten
-consecutively rifled Ruger P85 barrels) and fifteen questioned bullets
-(each matching one of the ten barrels). Ground truth for both of these
-Hamby sets is known and was used to assess correctness of the tests
-results.
-
-Bullet signatures being compared at this time are therefore from the
-Hamby 44 and Hamby 252 data. The database setup and pre-processing
-system used for choosing the Bullet signatures are as described by
-\citet{aoas}. In order to choose the bullet signatures we first filter
-out Land\_id for Profiles from the Hamby 44 and Hamby 252 data and
-remove all NA values. Then run\_id = 3 is chosen as the signatures
-generated from this run\_id give the closest match. Different run\_id's
-have some different settings for generating the signatures.(The level of
-smoothing does not seem to be one of them)
-
-The bullet signatures when generated by this process already includes a
-loess smoothing. Therefore, the coarseness factor is set to 1 while
-running the chumbley non random algorithm for comparing different
-optimization windows.The algorithm generates the same\_shift,
-different\_shift, U-Stat and P\_value parameters which are then used to
-calculate the errors associated with different sets of window sizes.
-
-\subsubsection{Profiles}\label{profiles}
-
-The profiles are cross-sectional values of the the bullet striation mark
-which are chosen at an optimum height (x as used by \citet{aoas}). This
-x or height is not a randomly chosen level. The rationale behind the
-choice has been explained by \citet{aoas}. A region is first chosen
-where the cross-correlation seems to change very less and in this region
-an optimum height is chosen. The profiles generally resemble a curve
-which is more or less similar to a quadratic curve (a quadratic fit to
-the raw data values of the profile is not an exact fit but it does show
-a similar trend). Profiles are the set of raw values representing the
-striation marks, and signatures are generated from these by removal of
-the inherent curvature and applying some smoothing (the signatures
-generated by \citet{aoas} use a loess function for smoothing).
-
-Similar to signatures the run\_id = 3 was used when applying the
-chumbley algorithm using the database setup given by \citet{aoas} of
-Hamby-44 and Hamby-252 datasets, on the profiles. The run\_id not only
-defines the level of smoothing but also signifies the chosen height at
-which the profiles were selected initially. Another important aspect is
-the range of horizontal values (which is referred to as the y values in
-\citet{aoas}) in the signatures. These have already been pre-processed
-in the database to not include any grooves.
-
-Therefore for the sake of comparison the run\_id = 3 is still chosen so
-as to ensure that the horizontal values remain the same as that of the
-signatures. This also gives us profiles with the grooves removed.
-
-The idea therefore is to first use these raw values of the profile
-directly in the chumbley algorithm, and see how the algorithm performs
-for different coarseness values (smoothing parameter as referred in the
-function lowess used in the chumbley algorithm).
-
-\pagebreak
-
 \section{Results}\label{results-1}
 
-We used the adjusted Chumbley method as proposed in \citet{hadler} and
-implemented in the R package \texttt{toolmaRk} \citep{toolmark} on all
-pairwise land-to-land comparisons of the Hamby scans (a total of 85,491
-comparisons) with the pairwise sets for the comparisons given in the
-table \ref{tab:param}.
-
-\begin{table}[!h][!h]
-\caption{\label{tab:param}Overview of parameter settings used for optimization and validation windows for bullet land signatures.}
-
-\centering
-\resizebox{\linewidth}{!}{\resizebox{\linewidth}{!}{\begin{tabular}[t]{lrrrrrrrrrrrrrrrrrr}
-\toprule
-wo & 50 & 50 & 60 & 60 & 80 & 80 & 90 & 90 & 100 & 100 & 110 & 110 & 120 & 120 & 120 & 120 & 120 & 120\\
-wv & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 10 & 20 & 30 & 40 & 50 & 60\\
-\bottomrule
-\end{tabular}}}
-\centering
-\begin{tabular}[t]{lrrrrrrrrrrrrrrr}
-\toprule
-wo & 130 & 130 & 140 & 140 & 150 & 150 & 160 & 160 & 200 & 200 & 200 & 240 & 240 & 280 & 280\\
-wv & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 20 & 30 & 50 & 30 & 50 & 30 & 50\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\begin{table}[!h]
-\caption{\label{tab:confusion}Confusion Table for different optimization window sizes with validation window size as 30.}
-
-\centering
-\begin{tabular}[t]{llrl}
-\toprule
-\hspace{1em}\hspace{1em}signif & match & Freq & Type\\
-\midrule
-\addlinespace[0.5em]
-\multicolumn{4}{l}{\textbf{Size of Optimization Window = 280}}\\
-\hspace{1em}FALSE & FALSE & 78909 & True Negative\\
-\hspace{1em}TRUE & FALSE & 4192 & False Positive (Type I)\\
-\hspace{1em}FALSE & TRUE & 446 & False Negative (Type II)\\
-\hspace{1em}TRUE & TRUE & 731 & True Positive\\
-\bottomrule
-\end{tabular}
-\centering
-\begin{tabular}[t]{llrl}
-\toprule
-signif & match & Freq & Type\\
-\midrule
-\addlinespace[0.5em]
-\multicolumn{4}{l}{\textbf{Size of Optimization Window = 120}}\\
-\hspace{1em}FALSE & FALSE & 79249 & True Negative\\
-\hspace{1em}TRUE & FALSE & 4523 & False Positive (Type I)\\
-\hspace{1em}FALSE & TRUE & 386 & False Negative (Type II)\\
-\hspace{1em}TRUE & TRUE & 854 & True Positive\\
-\bottomrule
-\end{tabular}
-\centering
-\begin{tabular}[t]{llrl}
-\toprule
-signif & match & Freq & Type\\
-\midrule
-\addlinespace[0.5em]
-\multicolumn{4}{l}{\textbf{Size of Optimization Window = 80}}\\
-\hspace{1em}FALSE & FALSE & 79477 & True Negative\\
-\hspace{1em}TRUE & FALSE & 4503 & False Positive (Type I)\\
-\hspace{1em}FALSE & TRUE & 463 & False Negative (Type II)\\
-\hspace{1em}TRUE & TRUE & 781 & True Positive\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\subsection{Signatures}\label{signatures-1}
-
-Figure \ref{fig:type2} gives an overview of type II error rates observed
-when varying the window size in the optimization step. Two levels of
-validation window size 30 and 50 were chosen as to compare the error
-rates for different nominal type I errors. We notice that the trends for
-these nominal type I errors are similar and in most cases a validation
-window of 50 has higher type II error than for 30. A change in this
-trend is seen for a 0.05 \(\alpha\) level, although the difference
-between the two windows is very small for this case. We can also notice
-an obvious trend of increase in the Type II error as the window of
-optimization increases and see a minimum around the optimization window
-size of 120 pixels. Hence we are inclined to choose a smaller validation
-window size and optimization window as 120.
-
-Table \ref{tab:confusion} shows the confusion tables with the
-classification of type I and type II errors and how the numbers change
-with a change in the optimization window. The windows represent areas to
-the left of the window with minimum type II, near the minimum type II
-window and to the far right of the minimum type II error
-
-Figure \ref{fig:type1} compares nominal (fixed) type I error and
-actually observed type I errors for the parameter settings in table
-\ref{tab:param}. With an increasing size of the window used in the
-optimization step the observed type I error rate decreases (slighty).
-This means as the optimization window increase the observed type I error
-rate gets smaller. A smaller validation window on the other hand, tends
-to be associated with a higher type I error rate. This can be better
-imagined for a given window of optimization, where the actual Type I
-error is comparable to the nominal level for only a select few
-validation window sizes. For these comparable validation window sizes of
-30 and 50 as done here, the actual type I error increases very slightly
-and can be seen in Figure \ref{fig:type1}. This increase is not as much
-when compared to the variation seen with the optimization window sizes.
-This effect might be related to the increasing number of tests that fail
-for larger optimization window sizes, in particular for non-matching
-striae (see fig \ref{fig:missings}).
-
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figures/wo120-1}
-
-}
-
-\caption{The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120}\label{fig:wo120}
-\end{figure}
-
-The actual type I error and type II error for signatures were also
-compared for different validation window sizes. Figure \ref{fig:wo120}
-shows the actual rates for different nominal \(\alpha\) levels. We can
-see that the type II error rises with higher validation windows for the
-smaller nominal \(\alpha\) levels while for the nominal \(\alpha\) =
-0.05 its almost constant.
-
-\begin{table}
-
-\caption{\label{tab:lms}Estimates of the increase in percent of failed tests corresponding to a 100 point increase in the optimization window.}
-\centering
-\begin{tabular}[t]{lrrr}
-\toprule
-match & wv & estimate & std.error\\
-\midrule
-FALSE & 30 & 0.510 & 0.029\\
-FALSE & 50 & 0.936 & 0.036\\
-TRUE & 30 & 2.933 & 0.247\\
-TRUE & 50 & 3.843 & 0.321\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-Figure \ref{fig:missings} gives an overview of the number of failed
-tests, i.e.~tests in which a particular parameter setting did not return
-a valid result. This happens, when the shift to align two signatures is
-so large, that the remaining overlap is too small to accommodate windows
-for validation. The problem is therefore exacerbated by a larger
-validation window. Figure \ref{fig:missings} also shows that the number
-of failed tests is approximately linear in the size of the optimization
-window. Test results from different sources have a much higher chance to
-fail, raising the question, whether failed tests should be treated as
-rejections of the null hypothesis of same source. For known non-matches
-there is a higher possibility that in the optimization pair of windows
-where cross-correlations are maximum are too far apart, and same shifts
-of this order hit the end of the signature.
-
-\subsection{Profiles}\label{profiles-1}
+\subsection{Profiles}\label{profiles}
 

### 2018-03-23 doc/paper.Rmd
couple of wrong assumptions uncovered

@@
 library(kableExtra)
+library(xtable)
 ```
@@
                 errorrate(all, 0.05))
+errors$type <- "signatures"
 
@@
                 errorrate(allp, 0.05))
+errors2$type <- "profiles"
 
@@
 
-Figure \ref{fig:prof_missings} (a) shows the type II error rates for profiles for the optimization window 120 and validation window 30 with varying level of coarseness.
-We can see that the type II error for all the nominal $\alpha$ levels is lowest in the range of 0.20 to 0.35.
-Therefore, a value of 0.25 can be used keeping in mind it keeps the type II error lowest while runnning simulations.
-Thus for comparisons of different window sizes etc as seen in the differnt parts of Figure \ref{fig:prof_missings} this coarseness value is used.
-
-On the other hand Figure \ref{fig:prof_missings} (b) shows if the coarseness level set in the chumbley agorithm has any effect on the signatures, which are pre-processed and already smoothed to a certain extent.
-From Figure \ref{fig:prof_missings} (b) we can notice that for different nominal $\alpha$ levels, the type II error fluctuates slightly but does not change much, thereby helping us conclude that the coarseness levels set in the LOWESS smoothing in the chumbley alggorithm does effect the type II error much for signatures.
-
-### Comparison of profiles and signatures
-
-Another reason for failed tests can be incorrect identification of maximum correlation windows in the optimization step as seen in figure \ref{fig:prof_missings}(d) because of the level of smoothing, as too much smoothing would subdue intricate features that might otherwise help in the correlation calculations and correct identification of maximum correlation windows irrespective of the size.
-This would again cause a simiar effect as explained for figure \ref{fig:missings} with validation windows, irrespective of size, during the shifts end up at the ends of the markings resulting in an invalid calculation and failed comparison attempt.
-
-In figure \ref{fig:prof_missings}(d) and (f), we compare profiles and signatures on the basis of number of failed tests.
-The profiles chosen for figure \ref{fig:prof_missings}(f) have a constant coarseness of 0.25 and window of optimization as 120.
-The signatures in this case are not smoothed using the chumbley algorithm step of LOWESS smoothing.
-Instead signatures are used as calculated by \citet{aoas}.
-The smoothing in these signatures were determined and fixed on the basis of their performance in the random forest based algorithm proposed by \citet{aoas}.
-The comparison of profiles and signatures with variation of validation window size therefore is made on even footing.
-The trends are similar to figure \ref{fig:missings} in the sense that for known non-matches the number of failed tests are more for both signatures and profiles and increasing linearly with the validation window size.
-The problem is however, worse for profiles which has higher number of failed tests than signatures for all validation windows.
-
-The total error for different validation window sizes for signatures and profiles can be seen in figure \ref{fig:prof_missings} (e).The optimization window size is 120 and profiles are calculated at a default 0.25 coarseness level while signatures as before are not smoothed again in the modified chumbley algorithm.
-We can see that the total error is always higher for profiles as compared to signatures for all sizes of validation window.
-
-```{r, fig.cap="Type II errors for different levels of coarseness."}
+\hh{When dealing with profiles, coarseness is an additional parameter that has to be considered in the matching.}
+
+\hh{XXX We have been operating under a false assumption: smoothing the signature at f=1 DOES change it - see figure \ref{fig:profile-sketch}.
+We need to re-run some of the results to compare the results against a }
+
+```{r profile-sketch, fig.height = 3, fig.width='\\textwidth'}
+profile <- sigs_graphics[[1]] %>% filter(land_id==19)
+profile$lw25 <- lowess(x = profile$y, y = profile$value, f=.25)$y
+profile$lw50 <- lowess(x = profile$y, y = profile$value, f=.5)$y
+profile$lw75 <- lowess(x = profile$y, y = profile$value, f=.75)$y
+profile$lw100 <- lowess(x = profile$y, y = profile$value, f=1)$y
+
+profile$l25 <- predict(loess(data = profile, value~y, span=.25))
+profile$l50 <- predict(loess(data = profile, value~y, span=.5))
+profile$l75 <- predict(loess(data = profile, value~y, span=.75))
+profile$l100 <- predict(loess(data = profile, value~y, span=2))
+
+pw1 <- profile %>%
+  ggplot(aes(x = y, y = value-lw25)) + geom_line() +
+  theme_bw() +
+  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
+  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
+  ggtitle("Coarseness 0.25")
+
+pw2 <- profile %>%
+  ggplot(aes(x = y, y = value-lw50)) + geom_line() +
+  theme_bw() +
+  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
+  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
+  ggtitle("Coarseness 0.5")
+
+pw3 <- profile %>%
+  ggplot(aes(x = y, y = value-lw75)) + geom_line() +
+  theme_bw() +
+  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
+  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
+  ggtitle("Coarseness 0.75")
+
+pw4 <- profile %>%
+  ggplot(aes(x = y, y = value-lw100)) + geom_line() +
+  theme_bw() +
+  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
+  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
+  ggtitle("Coarseness 1")
+
+p1 <- profile %>%
+  ggplot(aes(x = y, y = value-l25)) + geom_line() +
+  theme_bw() +
+  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
+  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
+  ggtitle("Coarseness 0.25")
+
+p2 <- profile %>%
+  ggplot(aes(x = y, y = value-l50)) + geom_line() +
+  theme_bw() +
+  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
+  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
+  ggtitle("Coarseness 0.5")
+
+p3 <- profile %>%
+  ggplot(aes(x = y, y = value-l75)) + geom_line() +
+  theme_bw() +
+  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
+  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
+  ggtitle("Coarseness 0.75")
+
+p4 <- profile %>%
+  ggplot(aes(x = y, y = value-l100)) + geom_line() +
+  theme_bw() +
+  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
+  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
+  ggtitle("Coarseness 1")
+
+prof  <- profile %>% ggplot(aes (x = y, y = value)) + geom_line() +
+  theme_bw() +
+  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
+  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
+  ggtitle("Profile")
+grid.arrange(pw1, pw2, pw3, pw4, p1, p2, p3, p4, prof, ncol=4)
+```
+
+\hh{Figure \ref{fig:coarse} shows the type II error rates for profiles using an optimization window $w_o= 120$ and a validation window $w_v = 30$ for varying level of coarseness.
+The type II error for all nominal levels of  $\alpha$ are the lowest for a coarseness range of 0.20 to 0.35.
+For the remainder of the analysis we use a fixed value of coarseness of 0.25.
+}
+
+```{r coarse, fig.cap="Type II error with respect to coarseness parameter over profiles, $w_o = 120$, $w_v = 30$.
+Optimal values for coarseness are around $c \\approx 0.30$", fig.width=5, fig.height = 3, out.width='.6\\textwidth'}
 e1<-errors2 %>% filter(wv==30)%>% ggplot(aes(x = coarse, y=beta, colour=factor(alpha))) +
@@
   scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + xlab("Profile Coarseness")+
-  ylab("Type II error") + ggtitle(label = "(a)")
-
-e2<- errors_sig_c %>% filter(wv==30)%>% ggplot(aes(x = coarse, y=beta, colour=factor(alpha))) +
-  geom_point() + geom_line() +
-  theme_bw() +
-  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2")+ xlab("Signature Coarseness") + ylab("Type II error") + ggtitle(label = "(b)")
-
-#multiplot(e1,e2, cols = 1)
-
-```
-
-```{r}
-#errors %>% ggplot(aes(x = wv, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + facet_grid(~wo) +geom_point(data=errors2, aes(shape="profiles"))
-
-total1<- errors_sig_c %>% ggplot(aes(x = coarse, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + geom_line()  + geom_point(data= errors2 %>% filter(wv==30), aes(shape="profiles")) + geom_line(data=errors2 %>% filter(wv==30), aes(shape="profiles")) + theme_bw() +
-  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + xlab("Coarseness") +ylab("Total error") + ggtitle(label = "(c)")
-
-total2<- wo120 %>% filter(wv>10) %>% ggplot(aes(x = wv, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + geom_line() +geom_point(data=errors2 %>% filter(wv>10) %>% filter(coarse==0.25), aes(shape="profiles")) + geom_line(data=errors2 %>% filter(wv>10)%>% filter(coarse==0.25), aes(shape="profiles")) + theme_bw() +
-  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + ylab("Total error") + ggtitle(label = "(e)")
-
-```
-
-```{r prof_missings, fig.width=10, fig.height=10, fig.cap="Row 3:  Total error and Number of failed tests by the window validation size, wv, and ground truth, Row 2: Total error and Number of failed tests with Coarseness for both profiles and signatures, Row 1: Type II error for different coarseness levels as used in the modified chumbley algorithm for profiles and signatures"}
-
-#ns1<- ns[which(ns$wv == c(30,50), arr.ind = T),]
-miss_1<- nsp %>% filter(wv == 30) %>%
-#ns %>% filter(wv == c(30, 50)) %>%
-  ggplot(aes(x = coarse, y = missperc, shape= "profile", colour = factor(match))) + 
-  #geom_smooth(method="auto", size=.5, se=FALSE) +
-  geom_point(size = 2) + 
-  theme_bw() +
-  ylab("Percent of failed tests.") +
-  xlab("Coarseness") +
-  scale_colour_brewer("Match", palette="Set1") + geom_point(data=ns_sig_c  %>% filter(wv %in% c(30,50)) , aes(shape="signature"), size = 2) + ggtitle(label = "(d)")
-# + scale_colour_brewer("Match", palette="Set2") #+geom_smooth(method="auto", size=.5, se=FALSE) 
-  #scale_shape_discrete("Size of \nvalidation\nwindow, #wv")
-
-miss_2<- nsp %>% filter(coarse == 0.25) %>%
-#ns %>% filter(wv == c(30, 50)) %>%
-  ggplot(aes(x = wv, y = missperc, shape= "profile", colour = factor(match))) + 
-  #geom_smooth(method="auto", size=.5, se=FALSE) +
-  geom_point(size = 2) + 
-  theme_bw() +
-  ylab("Percent of failed tests.") +
-  xlab("wv") +
-  scale_colour_brewer("Match", palette="Set1") + geom_point(data=ns %>% filter(wo == 120)  %>% filter(coarse == 1) , aes(shape="signature"), size = 2) + ggtitle(label = "(f)")#+ scale_colour_brewer("Match", palette="Set2")
-
-#multiplot(plotlist = list(total1, total2, miss_1, miss_2, e1, e2), cols = 2)#, matrix(c(1,2,3,3), nrow=2, byrow=TRUE))
-multiplot(plotlist = list(e1, total1, total2,e2, miss_1, miss_2), cols = 2)
-```
- \pagebreak
-
-## Conclusion
+  ylab("Type II error") 
+e1
+```
+
+```{r, include =F}
+wo120_sig <- errors %>% filter(wo==120, wv %in% c(30, 50))
+wo120_pr <- errors2 %>% filter(wo==120, wv %in% c(30, 50), coarse == .25)
+
+dt <- rbind(wo120_sig, wo120_pr) %>% 
+  ungroup() %>% 
+  mutate(beta = round(100*beta, 1)) %>%
+  select(wv, type, alpha, beta) %>% 
+  spread(alpha, beta)
+print(xtable(dt), include.rownames=FALSE)
+```
+
+\begin{table}[ht]
+\caption{\label{tab:type2} Type II error rates for profiles and signatures of bullet lands. For profiles, a coarseness value of $c = 0.25$ is used to remove bullet curvature.}
+\centering
+\begin{tabular}{rlrrrr}
+  \hline
+optimization && \multicolumn{4}{l}{Nominal type I error rate $\alpha$}\\  
+window $w_o$ & source & 0.001 & 0.005 & 0.01 & 0.05 \\ 
+  \hline
+ 30 & profiles & 54.40 & 43.80 & 39.70 & 30.00 \\ 
+   30 & signatures & 55.00 & 45.40 & 41.40 & 31.10 \\ \hline
+   50 & profiles & 58.50 & 44.40 & 40.70 & 28.70 \\ 
+   50 & signatures & 62.60 & 49.60 & 44.20 & 30.50 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+# Conclusion
 
@@
 -->
+
+# Appendix
+\begin{appendix}
+
+On the other hand Figure \ref{fig:prof_missings} (b) shows if the coarseness level set in the chumbley agorithm has any effect on the signatures, which are pre-processed and already smoothed to a certain extent.
+From Figure \ref{fig:prof_missings} (b) we can notice that for different nominal $\alpha$ levels, the type II error fluctuates slightly but does not change much, thereby helping us conclude that the coarseness levels set in the LOWESS smoothing in the chumbley alggorithm does effect the type II error much for signatures.
+
+\subsection{Comparison of profiles and signatures}
+
+Another reason for failed tests can be incorrect identification of maximum correlation windows in the optimization step as seen in figure \ref{fig:prof_missings}(d) because of the level of smoothing, as too much smoothing would subdue intricate features that might otherwise help in the correlation calculations and correct identification of maximum correlation windows irrespective of the size.
+This would again cause a simiar effect as explained for figure \ref{fig:missings} with validation windows, irrespective of size, during the shifts end up at the ends of the markings resulting in an invalid calculation and failed comparison attempt.
+
+In figure \ref{fig:prof_missings}(d) and (f), we compare profiles and signatures on the basis of number of failed tests.
+The profiles chosen for figure \ref{fig:prof_missings}(f) have a constant coarseness of 0.25 and window of optimization as 120.
+The signatures in this case are not smoothed using the chumbley algorithm step of LOWESS smoothing.
+Instead signatures are used as calculated by \citet{aoas}.
+The smoothing in these signatures were determined and fixed on the basis of their performance in the random forest based algorithm proposed by \citet{aoas}.
+The comparison of profiles and signatures with variation of validation window size therefore is made on even footing.
+The trends are similar to figure \ref{fig:missings} in the sense that for known non-matches the number of failed tests are more for both signatures and profiles and increasing linearly with the validation window size.
+The problem is however, worse for profiles which has higher number of failed tests than signatures for all validation windows.
+
+The total error for different validation window sizes for signatures and profiles can be seen in figure \ref{fig:prof_missings} (e).The optimization window size is 120 and profiles are calculated at a default 0.25 coarseness level while signatures as before are not smoothed again in the modified chumbley algorithm.
+We can see that the total error is always higher for profiles as compared to signatures for all sizes of validation window.
+
+```{r, fig.cap="Type II errors for different levels of coarseness."}
+e1<-errors2 %>% filter(wv==30)%>% ggplot(aes(x = coarse, y=beta, colour=factor(alpha))) +
+  geom_point() + geom_line() +
+  theme_bw() +
+  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + xlab("Profile Coarseness")+
+  ylab("Type II error") + ggtitle(label = "(a)")
+
+e2<- errors_sig_c %>% filter(wv==30)%>% ggplot(aes(x = coarse, y=beta, colour=factor(alpha))) +
+  geom_point() + geom_line() +
+  theme_bw() +
+  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2")+ xlab("Signature Coarseness") + ylab("Type II error") + ggtitle(label = "(b)")
+
+#multiplot(e1,e2, cols = 1)
+
+```
+
+```{r}
+#errors %>% ggplot(aes(x = wv, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + facet_grid(~wo) +geom_point(data=errors2, aes(shape="profiles"))
+
+total1<- errors_sig_c %>% ggplot(aes(x = coarse, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + geom_line()  + geom_point(data= errors2 %>% filter(wv==30), aes(shape="profiles")) + geom_line(data=errors2 %>% filter(wv==30), aes(shape="profiles")) + theme_bw() +
+  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + xlab("Coarseness") +ylab("Total error") + ggtitle(label = "(c)")
+
+total2<- wo120 %>% filter(wv>10) %>% ggplot(aes(x = wv, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + geom_line() +geom_point(data=errors2 %>% filter(wv>10) %>% filter(coarse==0.25), aes(shape="profiles")) + geom_line(data=errors2 %>% filter(wv>10)%>% filter(coarse==0.25), aes(shape="profiles")) + theme_bw() +
+  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + ylab("Total error") + ggtitle(label = "(e)")
+
+```
+
+```{r prof_missings, fig.width=10, fig.height=10, fig.cap="Row 3:  Total error and Number of failed tests by the window validation size, wv, and ground truth, Row 2: Total error and Number of failed tests with Coarseness for both profiles and signatures, Row 1: Type II error for different coarseness levels as used in the modified chumbley algorithm for profiles and signatures"}
+
+#ns1<- ns[which(ns$wv == c(30,50), arr.ind = T),]
+miss_1<- nsp %>% filter(wv == 30) %>%
+#ns %>% filter(wv == c(30, 50)) %>%
+  ggplot(aes(x = coarse, y = missperc, shape= "profile", colour = factor(match))) + 
+  #geom_smooth(method="auto", size=.5, se=FALSE) +
+  geom_point(size = 2) + 
+  theme_bw() +
+  ylab("Percent of failed tests.") +
+  xlab("Coarseness") +
+  scale_colour_brewer("Match", palette="Set1") + geom_point(data=ns_sig_c  %>% filter(wv %in% c(30,50)) , aes(shape="signature"), size = 2) + ggtitle(label = "(d)")
+# + scale_colour_brewer("Match", palette="Set2") #+geom_smooth(method="auto", size=.5, se=FALSE) 
+  #scale_shape_discrete("Size of \nvalidation\nwindow, #wv")
+
+miss_2<- nsp %>% filter(coarse == 0.25) %>%
+#ns %>% filter(wv == c(30, 50)) %>%
+  ggplot(aes(x = wv, y = missperc, shape= "profile", colour = factor(match))) + 
+  #geom_smooth(method="auto", size=.5, se=FALSE) +
+  geom_point(size = 2) + 
+  theme_bw() +
+  ylab("Percent of failed tests.") +
+  xlab("wv") +
+  scale_colour_brewer("Match", palette="Set1") + geom_point(data=ns %>% filter(wo == 120)  %>% filter(coarse == 1) , aes(shape="signature"), size = 2) + ggtitle(label = "(f)")#+ scale_colour_brewer("Match", palette="Set2")
+
+#multiplot(plotlist = list(total1, total2, miss_1, miss_2, e1, e2), cols = 2)#, matrix(c(1,2,3,3), nrow=2, byrow=TRUE))
+multiplot(plotlist = list(e1, total1, total2,e2, miss_1, miss_2), cols = 2)
+```
+\end{appendix}

### 2018-03-23 doc/paper.tex
couple of wrong assumptions uncovered

@@
 
-Figure \ref{fig:prof_missings} (a) shows the type II error rates for
-profiles for the optimization window 120 and validation window 30 with
-varying level of coarseness. We can see that the type II error for all
-the nominal \(\alpha\) levels is lowest in the range of 0.20 to 0.35.
-Therefore, a value of 0.25 can be used keeping in mind it keeps the type
-II error lowest while runnning simulations. Thus for comparisons of
-different window sizes etc as seen in the differnt parts of Figure
-\ref{fig:prof_missings} this coarseness value is used.
+\hh{When dealing with profiles, coarseness is an additional parameter that has to be considered in the matching.}
 
-On the other hand Figure \ref{fig:prof_missings} (b) shows if the
-coarseness level set in the chumbley agorithm has any effect on the
-signatures, which are pre-processed and already smoothed to a certain
-extent. From Figure \ref{fig:prof_missings} (b) we can notice that for
-different nominal \(\alpha\) levels, the type II error fluctuates
-slightly but does not change much, thereby helping us conclude that the
-coarseness levels set in the LOWESS smoothing in the chumbley alggorithm
-does effect the type II error much for signatures.
+\hh{XXX We have been operating under a false assumption: smoothing the signature at f=1 DOES change it - see figure \ref{fig:profile-sketch}.
+We need to re-run some of the results to compare the results against a }
 
-\subsubsection{Comparison of profiles and
-signatures}\label{comparison-of-profiles-and-signatures}
+\begin{center}\includegraphics[width=\textwidth]{figures/profile-sketch-1} \end{center}
 
-Another reason for failed tests can be incorrect identification of
-maximum correlation windows in the optimization step as seen in figure
-\ref{fig:prof_missings}(d) because of the level of smoothing, as too
-much smoothing would subdue intricate features that might otherwise help
-in the correlation calculations and correct identification of maximum
-correlation windows irrespective of the size. This would again cause a
-simiar effect as explained for figure \ref{fig:missings} with validation
-windows, irrespective of size, during the shifts end up at the ends of
-the markings resulting in an invalid calculation and failed comparison
-attempt.
-
-In figure \ref{fig:prof_missings}(d) and (f), we compare profiles and
-signatures on the basis of number of failed tests. The profiles chosen
-for figure \ref{fig:prof_missings}(f) have a constant coarseness of 0.25
-and window of optimization as 120. The signatures in this case are not
-smoothed using the chumbley algorithm step of LOWESS smoothing. Instead
-signatures are used as calculated by \citet{aoas}. The smoothing in
-these signatures were determined and fixed on the basis of their
-performance in the random forest based algorithm proposed by
-\citet{aoas}. The comparison of profiles and signatures with variation
-of validation window size therefore is made on even footing. The trends
-are similar to figure \ref{fig:missings} in the sense that for known
-non-matches the number of failed tests are more for both signatures and
-profiles and increasing linearly with the validation window size. The
-problem is however, worse for profiles which has higher number of failed
-tests than signatures for all validation windows.
-
-The total error for different validation window sizes for signatures and
-profiles can be seen in figure \ref{fig:prof_missings} (e).The
-optimization window size is 120 and profiles are calculated at a default
-0.25 coarseness level while signatures as before are not smoothed again
-in the modified chumbley algorithm. We can see that the total error is
-always higher for profiles as compared to signatures for all sizes of
-validation window.
+\hh{Figure \ref{fig:coarse} shows the type II error rates for profiles using an optimization window $w_o= 120$ and a validation window $w_v = 30$ for varying level of coarseness.
+The type II error for all nominal levels of  $\alpha$ are the lowest for a coarseness range of 0.20 to 0.35.
+For the remainder of the analysis we use a fixed value of coarseness of 0.25.
+}
 
@@
 
-{\centering \includegraphics[width=\textwidth]{figures/prof_missings-1}
+{\centering \includegraphics[width=.6\textwidth]{figures/coarse-1}
 
@@
 
-\caption{Row 3:  Total error and Number of failed tests by the window validation size, wv, and ground truth, Row 2: Total error and Number of failed tests with Coarseness for both profiles and signatures, Row 1: Type II error for different coarseness levels as used in the modified chumbley algorithm for profiles and signatures}\label{fig:prof_missings}
+\caption{Type II error with respect to coarseness parameter over profiles, $w_o = 120$, $w_v = 30$. Optimal values for coarseness are around $c \approx 0.30$}\label{fig:coarse}
 \end{figure}
 
-\pagebreak
+\begin{table}[ht]
+\caption{\label{tab:type2} Type II error rates for profiles and signatures of bullet lands. For profiles, a coarseness value of $c = 0.25$ is used to remove bullet curvature.}
+\centering
+\begin{tabular}{rlrrrr}
+  \hline
+optimization && \multicolumn{4}{l}{Nominal type I error rate $\alpha$}\\  
+window $w_o$ & source & 0.001 & 0.005 & 0.01 & 0.05 \\ 
+  \hline
+ 30 & profiles & 54.40 & 43.80 & 39.70 & 30.00 \\ 
+   30 & signatures & 55.00 & 45.40 & 41.40 & 31.10 \\ \hline
+   50 & profiles & 58.50 & 44.40 & 40.70 & 28.70 \\ 
+   50 & signatures & 62.60 & 49.60 & 44.20 & 30.50 \\ 
+   \hline
+\end{tabular}
+\end{table}
 
-\subsection{Conclusion}\label{conclusion}
+\section{Conclusion}\label{conclusion}
 
@@
 
+\section{Appendix}\label{appendix}
+
+\begin{appendix}
+
+On the other hand Figure \ref{fig:prof_missings} (b) shows if the coarseness level set in the chumbley agorithm has any effect on the signatures, which are pre-processed and already smoothed to a certain extent.
+From Figure \ref{fig:prof_missings} (b) we can notice that for different nominal $\alpha$ levels, the type II error fluctuates slightly but does not change much, thereby helping us conclude that the coarseness levels set in the LOWESS smoothing in the chumbley alggorithm does effect the type II error much for signatures.
+
+\subsection{Comparison of profiles and signatures}
+
+Another reason for failed tests can be incorrect identification of maximum correlation windows in the optimization step as seen in figure \ref{fig:prof_missings}(d) because of the level of smoothing, as too much smoothing would subdue intricate features that might otherwise help in the correlation calculations and correct identification of maximum correlation windows irrespective of the size.
+This would again cause a simiar effect as explained for figure \ref{fig:missings} with validation windows, irrespective of size, during the shifts end up at the ends of the markings resulting in an invalid calculation and failed comparison attempt.
+
+In figure \ref{fig:prof_missings}(d) and (f), we compare profiles and signatures on the basis of number of failed tests.
+The profiles chosen for figure \ref{fig:prof_missings}(f) have a constant coarseness of 0.25 and window of optimization as 120.
+The signatures in this case are not smoothed using the chumbley algorithm step of LOWESS smoothing.
+Instead signatures are used as calculated by \citet{aoas}.
+The smoothing in these signatures were determined and fixed on the basis of their performance in the random forest based algorithm proposed by \citet{aoas}.
+The comparison of profiles and signatures with variation of validation window size therefore is made on even footing.
+The trends are similar to figure \ref{fig:missings} in the sense that for known non-matches the number of failed tests are more for both signatures and profiles and increasing linearly with the validation window size.
+The problem is however, worse for profiles which has higher number of failed tests than signatures for all validation windows.
+
+The total error for different validation window sizes for signatures and profiles can be seen in figure \ref{fig:prof_missings} (e).The optimization window size is 120 and profiles are calculated at a default 0.25 coarseness level while signatures as before are not smoothed again in the modified chumbley algorithm.
+We can see that the total error is always higher for profiles as compared to signatures for all sizes of validation window.
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figures/prof_missings-1}
+
+}
+
+\caption{Row 3:  Total error and Number of failed tests by the window validation size, wv, and ground truth, Row 2: Total error and Number of failed tests with Coarseness for both profiles and signatures, Row 1: Type II error for different coarseness levels as used in the modified chumbley algorithm for profiles and signatures}\label{fig:prof_missings}
+\end{figure}
+\end{appendix}
+
 \bibliographystyle{agsm}

### 2018-03-30 doc/paper.Rmd
mini edits

@@
 
-```{r profile-sketch, fig.height = 3, fig.width='\\textwidth'}
+```{r profile-sketch, fig.height = 2.5, fig.width='\\textwidth', fig.cap="Overview of the effect of different coarseness parameters $c$ on the profile shown in Figure \\ref{fig:sigs-profiles}. "}
 profile <- sigs_graphics[[1]] %>% filter(land_id==19)
@@
 
-profile$l25 <- predict(loess(data = profile, value~y, span=.25))
-profile$l50 <- predict(loess(data = profile, value~y, span=.5))
-profile$l75 <- predict(loess(data = profile, value~y, span=.75))
-profile$l100 <- predict(loess(data = profile, value~y, span=2))
-
-pw1 <- profile %>%
-  ggplot(aes(x = y, y = value-lw25)) + geom_line() +
+profiles <- tidyr::gather(profile, coarseness, values, lw25:lw75)
+profiles <- profiles %>% mutate(
+  c = readr::parse_number(coarseness)/100
+)
+profiles %>% 
+  ggplot(aes(x = y, y = value-values)) + geom_line() +
   theme_bw() +
@@
   xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) +
-  ggtitle("Coarseness 0.25")
-
-pw2 <- profile %>%
-  ggplot(aes(x = y, y = value-lw50)) + geom_line() +
-  theme_bw() +
-  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
-  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
-  ggtitle("Coarseness 0.5")
-
-pw3 <- profile %>%
-  ggplot(aes(x = y, y = value-lw75)) + geom_line() +
-  theme_bw() +
-  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
-  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
-  ggtitle("Coarseness 0.75")
-
-pw4 <- profile %>%
-  ggplot(aes(x = y, y = value-lw100)) + geom_line() +
-  theme_bw() +
-  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
-  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
-  ggtitle("Coarseness 1")
-
-p1 <- profile %>%
-  ggplot(aes(x = y, y = value-l25)) + geom_line() +
-  theme_bw() +
-  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
-  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
-  ggtitle("Coarseness 0.25")
-
-p2 <- profile %>%
-  ggplot(aes(x = y, y = value-l50)) + geom_line() +
-  theme_bw() +
-  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
-  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
-  ggtitle("Coarseness 0.5")
-
-p3 <- profile %>%
-  ggplot(aes(x = y, y = value-l75)) + geom_line() +
-  theme_bw() +
-  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
-  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
-  ggtitle("Coarseness 0.75")
-
-p4 <- profile %>%
-  ggplot(aes(x = y, y = value-l100)) + geom_line() +
-  theme_bw() +
-  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
-  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
-  ggtitle("Coarseness 1")
-
-prof  <- profile %>% ggplot(aes (x = y, y = value)) + geom_line() +
-  theme_bw() +
-  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
-  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
-  ggtitle("Profile")
-grid.arrange(pw1, pw2, pw3, pw4, p1, p2, p3, p4, prof, ncol=4)
+  facet_grid(~c, labeller="label_both")
+
 ```

### 2018-04-06 doc/paper.Rmd
re-phrasing error rates

@@
 \citet{aoas} use multiple features like CCF, CMS, D (distance measure) etc in a random forest based method and compare every land against every other land of  digitised versions of Hamby 252 and Hamby 44 \citep{hamby} published on the NIST Ballistics Database \citep{nist}.
-The authors report an out-of-bag overall error rate of 0.46%, comprised of a false positive error rate of 30.05% and a false negative rate of 0.026%.
+The authors report an out-of-bag overall error rate of 0.46%, comprised of an error rate of 30.05% of  same-source pairs that were not identified and an error rate of 0.026% of different-source pairs that were incorrectly identified as same-source.
 
@@
 The paper by \citet{hadler} is based on the same data but the authors focus on  markings made under the same angle.
-The error rates associated with the deterministic version of the score are  0.06 for false  negatives and a false poisitive error rate of 0.
+The error rates associated with the deterministic version of the score are  0.06 for false  negatives and a false positive error rate of 0.
 

### 2018-04-09 doc/paper.Rmd
some edits to definition profiles

@@
 \citet{chumbley} report error rates for markings made by the tips at different angles.
-For markings made under a 30 degree the authors  report an average false negative error rate of 0.023  and an average false positive error rate of 0.09.
-For other angles the error rates for false negatives stay the same while the rate of false positives decreases to 0.01.
-\hh{are the angles steeper?}
+For markings made at 30 degree the authors  report an average false negative error rate of 0.089  and an average false positive error rate of 0.023.
+For other angles of 60 and 85 degrees the false negatives error rate is 0.09 while the rate of false positives decreases to 0.01.
 The paper by \citet{hadler} is based on the same data but the authors focus on  markings made under the same angle.
@@
 From profiles bullet *signatures* \citep{chu2013,aoas} are extracted as residuals of a loess fit or Gaussian filter.
- Signatures are considered to be noise free and a good reflection of the key attributes of the raw marking, and the unique features of a bullet.
+This effectively removes topographic structure from the data in the attempt to increase the signal to noise ratio.
+The span of the loess fit was found using cross-validation, as described by \citet{aoas}.
+ <!-- Signatures are considered to be noise free and a good reflection of the key attributes of the raw marking, and the unique features of a bullet.
 \hh{Do you have a reference for the previous sentenced?} A  detailed discussion of the extraction technique for signatures is given in \citet{aoas}.
+-->
  <!--, where comprehensive details about the height at which profile is to be selected, removing curvature, smoothing, identifying groove locations are explained.-->

### 2018-04-12 code/evaluate-csafe-chumbley.R
adding data-profile-runs

@@ -1,0 +1,45 @@
+if (!file.exists("/opt/hamby44/csafe-sigs.rds")) {
+  files <- dir("/opt/hamby44/", pattern="csv")
+  files <- grep("^H44-", files, value = TRUE)
+  
+  csafe <- data.frame()
+  for (file in files) {
+    tmp <- read_csv(file=file.path(path="/opt/hamby44/", file)) 
+ #   if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% select(-x)
+    csafe <- rbind(csafe, tmp)
+  }
+  saveRDS(csafe, file="/opt/hamby44/csafe-sigs.rds")
+} else {
+  csafe <- readRDS("/opt/hamby44/csafe-sigs.rds")
+}
+
+csafe$coarse <- 1
+
+errorrate <- function(data, alpha) {
+  summ <- data %>% filter(!is.na(p_value)) %>%
+    mutate(signif = p_value < alpha) %>%
+    group_by(wv, wo, match, coarse, signif) %>% tally()
+  summ$error <- with(summ, match != signif)
+  summ$alpha <- alpha
+  
+  rates <- summ %>% dplyr::group_by(wv, wo, coarse, match, alpha) %>% dplyr::summarize(
+    rate = n[error==TRUE]/sum(n)
+  )
+  totals <- summ %>% dplyr::group_by(wv, wo, coarse, alpha) %>% dplyr::summarize(
+    total= sum(n[error==TRUE])/sum(n)
+  )
+  rates <- rates %>% spread(match,rate) %>% dplyr::rename(
+    actual = `FALSE`,
+    beta = `TRUE`
+  )
+  left_join(rates, totals)
+}
+
+
+
+errors <- rbind(errorrate(csafe, 0.001), 
+                errorrate(csafe, 0.005), 
+                errorrate(csafe, 0.01), 
+                errorrate(csafe, 0.05))
+errors %>% ggplot(aes(x = wo, y = beta, colour=factor(alpha), shape=factor(wv))) +
+  geom_point() + geom_smooth(aes(group=alpha), se=FALSE)

### 2018-04-12 code/evaluate-profiles.R

@@ -1,0 +1,83 @@
+library(tidyverse)
+if (!file.exists("/opt/hamby44/hamby-profiles.rds")) {
+  files <- dir("/opt/hamby44/", pattern="rds")
+  files <- grep("^chumbley-csafe-profiles", files, value = TRUE)
+  
+  profiles <- data.frame()
+  for (file in files) {
+    tmp <- readRDS(file=file.path(path="/opt/hamby44/", file)) 
+    #   if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% select(-x)
+    profiles <- rbind(profiles, tmp)
+  }
+  saveRDS(profiles, file="/opt/hamby44/hamby-profiles.rds")
+} else {
+  profiles <- readRDS("/opt/hamby44/hamby-profiles.rds")
+}
+
+profiles$coarse <- 0.3
+
+errorrate <- function(data, alpha) {
+  summ <- data %>% filter(!is.na(p_value)) %>%
+    mutate(signif = p_value < alpha) %>%
+    group_by(wv, wo, match, coarse, signif) %>% tally()
+  summ$error <- with(summ, match != signif)
+  summ$alpha <- alpha
+  
+  rates <- summ %>% dplyr::group_by(wv, wo, coarse, match, alpha) %>% dplyr::summarize(
+    rate = n[error==TRUE]/sum(n)
+  )
+  totals <- summ %>% dplyr::group_by(wv, wo, coarse, alpha) %>% dplyr::summarize(
+    total= sum(n[error==TRUE])/sum(n)
+  )
+  rates <- rates %>% spread(match,rate) %>% dplyr::rename(
+    actual = `FALSE`,
+    beta = `TRUE`
+  )
+  left_join(rates, totals)
+}
+
+
+
+errors <- rbind(errorrate(profiles, 0.001), 
+                errorrate(profiles, 0.005), 
+                errorrate(profiles, 0.01), 
+                errorrate(profiles, 0.05))
+errors %>% ggplot(aes(x = wo, y = beta, colour=factor(alpha), shape=factor(wv))) +
+  geom_point() + geom_smooth(aes(group=alpha), se=FALSE)
+
+
+testfailed <- profiles %>% group_by(wo, wv, match) %>% summarize(
+  total = n(), 
+  failed = sum(is.na(p_value)))
+testfailed %>% ggplot(aes(x = wo, y = failed/total, colour=factor(match), shape=factor(wv))) + geom_point() +
+  ylim(0, NA)
+
[…]

### 2018-04-13 doc/paper.Rmd
including locations

@@
 library(xtable)
+library(gridExtra)
 ```
@@
 
+<!--
 ## Potential Challenges in Chumbley Score Adaptation
 
-<!--Markings made by screw driver tips \citep{chumbley} are longer and pronounced and about 7 mm in length \citep{manytoolmarks1}, while the groove to groove length of the bullet lands from the Hamby study \citep{hamby} are about 2mm in length.
+Markings made by screw driver tips \citep{chumbley} are longer and pronounced and about 7 mm in length \citep{manytoolmarks1}, while the groove to groove length of the bullet lands from the Hamby study \citep{hamby} are about 2mm in length.
 The Ruger pistol barrels used in the  Hamby study have a 9mm caliber.
@@
 -->
-
-\hh{The Chumbley score allows a separation of a toolmark into raw and normalized digitized versions.
-Originally, this mechanism is intended to separate between class characteristics and individual markings, however, in the setting of matching bullet striae, we could also use it to separate bullet curvature in profiles from signatures before matching signatures.
-}
 
@@
 
-In a pre-processing step  the two markings are smoothed using a lowess \citep{lowess} with coarseness parameter $c$. The purpose of this smoothing is to remove drift and (sub)class characteristics.
-
-The Chumbley scores is calculated in two phases, namely, an optimization step and a validation step.
+In a pre-processing step  the two markings are smoothed using a lowess \citep{lowess} with coarseness parameter $c$. 
+ Originally, this smoothing is intended to remove drift and (sub)class characteristics from individual markings, however, in the setting of matching bullet striae, we can also make use of this mechanism can be used to separate bullet curvature in profiles from signatures before matching signatures.
+
+After removing sub-class structure, 
+the Chumbley scores is calculated in two steps: an optimization step and a validation step.
 In the optimization step, the two markings are aligned horizontally such that within a pre-defined window of length $w_o$ the correlation between $x(t_1)$ and $y(t_2)$ is maximized:
@@
 
-library(gridExtra)
 grid.arrange(p1,p2, ncol=2)
@@
 
-```{r profile-sketch, fig.height = 2.5, fig.width='\\textwidth', fig.cap="Overview of the effect of different coarseness parameters $c$ on the profile shown in Figure \\ref{fig:sigs-profiles}. "}
+```{r profile-sketch, fig.height = 5, fig.width='\\textwidth', fig.cap="Overview of the effect of different coarseness parameters $c$ on the profile shown in Figure \\ref{fig:sigs-profiles}. "}
 profile <- sigs_graphics[[1]] %>% filter(land_id==19)
+profile$lw20 <- lowess(x = profile$y, y = profile$value, f=.20)$y
 profile$lw25 <- lowess(x = profile$y, y = profile$value, f=.25)$y
-profile$lw50 <- lowess(x = profile$y, y = profile$value, f=.5)$y
-profile$lw75 <- lowess(x = profile$y, y = profile$value, f=.75)$y
-profile$lw100 <- lowess(x = profile$y, y = profile$value, f=1)$y
-
-profiles <- tidyr::gather(profile, coarseness, values, lw25:lw75)
+profile$lw30 <- lowess(x = profile$y, y = profile$value, f=.30)$y
+#profile$lw50 <- lowess(x = profile$y, y = profile$value, f=.5)$y
+#profile$lw75 <- lowess(x = profile$y, y = profile$value, f=.75)$y
+#profile$lw100 <- lowess(x = profile$y, y = profile$value, f=1)$y
+
+profiles <- tidyr::gather(profile, coarseness, values, lw20:lw30)
 profiles <- profiles %>% mutate(
@@
 )
-profiles %>% 
+p1 <- profiles %>% 
   ggplot(aes(x = y, y = value-values)) + geom_line() +
@@
 
+profc20 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo:160-wv:30-c:20.rds")
+profc20$coarseness <- 0.2
+profc25 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo:160-wv:30-c:25.rds")
+profc30 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo:160-wv:30.rds")
+profc30$coarseness <- .30
+profc <- bind_rows(profc20, profc25, profc30)
+profc <- profc %>% mutate(
+  locations = chumbley %>% purrr::map_int(.f = function(x) {
+    res <- NULL
+    try(res <- x$locations[1], silent = TRUE)
+    res
+  })
+)
+profsumm <- read.csv("../data/profiles-summary.csv")
+profsumm <- profsumm %>% select(-run_id)
+profc <- profc %>% left_join(profsumm, by=c("land2_id"="land_id", "profile2_id"="profile_id"))
+
+p2 <- profc %>% 
+  ggplot(aes(x = locations/(length-wo))) +
+  geom_histogram(binwidth = 0.025) +
+  facet_grid(.~coarseness, labeller="label_both") + 
+  theme_bw()
+
+grid.arrange(p1, p2, ncol=1)
 ```

### 2018-04-16 doc/paper.tex
mini edits
i am not sure what changes i am commiting

@@
 \citep{nist}. The authors report an out-of-bag overall error rate of
-0.46\%, comprised of a false positive error rate of 30.05\% and a false
-negative rate of 0.026\%.
+0.46\%, comprised of an error rate of 30.05\% of same-source pairs that
+were not identified and an error rate of 0.026\% of different-source
+pairs that were incorrectly identified as same-source.
 
@@
 error rates for markings made by the tips at different angles. For
-markings made under a 30 degree the authors report an average false
-negative error rate of 0.023 and an average false positive error rate of
-0.09. For other angles the error rates for false negatives stay the same
-while the rate of false positives decreases to 0.01.
-\hh{are the angles steeper?} The paper by \citet{hadler} is based on the
-same data but the authors focus on markings made under the same angle.
-The error rates associated with the deterministic version of the score
-are 0.06 for false negatives and a false poisitive error rate of 0.
+markings made at 30 degree the authors report an average false negative
+error rate of 0.089 and an average false positive error rate of 0.023.
+For other angles of 60 and 85 degrees the false negatives error rate is
+0.09 while the rate of false positives decreases to 0.01. The paper by
+\citet{hadler} is based on the same data but the authors focus on
+markings made under the same angle. The error rates associated with the
+deterministic version of the score are 0.06 for false negatives and a
+false positive error rate of 0.
 
@@
 bullet \emph{signatures} \citep{chu2013,aoas} are extracted as residuals
-of a loess fit or Gaussian filter. Signatures are considered to be noise
-free and a good reflection of the key attributes of the raw marking, and
-the unique features of a bullet.
-\hh{Do you have a reference for the previous sentenced?} A detailed
-discussion of the extraction technique for signatures is given in
-\citet{aoas}.
+of a loess fit or Gaussian filter. This effectively removes topographic
+structure from the data in the attempt to increase the signal to noise
+ratio. The span of the loess fit was found using cross-validation, as
+described by \citet{aoas}.
 
@@
 
-\subsection{Potential Challenges in Chumbley Score
-Adaptation}\label{potential-challenges-in-chumbley-score-adaptation}
-
-\hh{The Chumbley score allows a separation of a toolmark into raw and normalized digitized versions.
-Originally, this mechanism is intended to separate between class characteristics and individual markings, however, in the setting of matching bullet striae, we could also use it to separate bullet curvature in profiles from signatures before matching signatures.
-}
-
 \subsection{The Chumbley Score Test}\label{the-chumbley-score-test}
@@
 In a pre-processing step the two markings are smoothed using a lowess
-\citep{lowess} with coarseness parameter \(c\). The purpose of this
-smoothing is to remove drift and (sub)class characteristics.
+\citep{lowess} with coarseness parameter \(c\). Originally, this
+smoothing is intended to remove drift and (sub)class characteristics
+from individual markings, however, in the setting of matching bullet
+striae, we can also make use of this mechanism can be used to separate
+bullet curvature in profiles from signatures before matching signatures.
 
-The Chumbley scores is calculated in two phases, namely, an optimization
-step and a validation step. In the optimization step, the two markings
-are aligned horizontally such that within a pre-defined window of length
-\(w_o\) the correlation between \(x(t_1)\) and \(y(t_2)\) is maximized:
-\[
+After removing sub-class structure, the Chumbley scores is calculated in
+two steps: an optimization step and a validation step. In the
+optimization step, the two markings are aligned horizontally such that
+within a pre-defined window of length \(w_o\) the correlation between
+\(x(t_1)\) and \(y(t_2)\) is maximized: \[
 \left(t_1^o, t_2^o\right) = \mathop{\arg \max}\limits_{1 \le t_1 \le T_1, 1 \le t_2 \le T_2} \text{cor} \left(x (t_1, t_1 + w_o), y(t_2, t_2 + w_o) \right)
@@
 
-\begin{center}\includegraphics[width=\textwidth]{figures/profile-sketch-1} \end{center}
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figures/profile-sketch-1}
+
+}
+
+\caption{Overview of the effect of different coarseness parameters $c$ on the profile shown in Figure \ref{fig:sigs-profiles}. }\label{fig:profile-sketch}
+\end{figure}
 

### 2018-04-16 doc/paper.Rmd
commenting out section :: profile-summary.csv file missing

@@
 
-profc20 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo:160-wv:30-c:20.rds")
+profc20 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo_160-wv_30-c_20.rds")
 profc20$coarseness <- 0.2
-profc25 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo:160-wv:30-c:25.rds")
-profc30 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo:160-wv:30.rds")
+profc25 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo_160-wv_30-c_25.rds")
+profc30 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo_160-wv_30.rds")
 profc30$coarseness <- .30
@@
 )
-profsumm <- read.csv("../data/profiles-summary.csv")
-profsumm <- profsumm %>% select(-run_id)
-profc <- profc %>% left_join(profsumm, by=c("land2_id"="land_id", "profile2_id"="profile_id"))
-
-p2 <- profc %>% 
-  ggplot(aes(x = locations/(length-wo))) +
-  geom_histogram(binwidth = 0.025) +
-  facet_grid(.~coarseness, labeller="label_both") + 
-  theme_bw()
-
-grid.arrange(p1, p2, ncol=1)
+# profsumm <- read.csv("../data/profiles-summary.csv")
+# profsumm <- profsumm %>% select(-run_id)
+# profc <- profc %>% left_join(profsumm, by=c("land2_id"="land_id", "profile2_id"="profile_id"))
+# 
+# p2 <- profc %>% 
+#   ggplot(aes(x = locations/(length-wo))) +
+#   geom_histogram(binwidth = 0.025) +
+#   facet_grid(.~coarseness, labeller="label_both") + 
+#   theme_bw()
+#  
+# grid.arrange(p1, p2, ncol=1)
 ```
@@
   \hline
-optimization && \multicolumn{4}{l}{Nominal type I error rate $\alpha$}\\  
-window $w_o$ & source & 0.001 & 0.005 & 0.01 & 0.05 \\ 
+validation && \multicolumn{4}{l}{Nominal type I error rate $\alpha$}\\  
+window $w_v$ & source & 0.001 & 0.005 & 0.01 & 0.05 \\ 
   \hline

### 2018-04-16 doc/paper.Rmd
explanation of histograms

@@
 In a pre-processing step  the two markings are smoothed using a lowess \citep{lowess} with coarseness parameter $c$. 
- Originally, this smoothing is intended to remove drift and (sub)class characteristics from individual markings, however, in the setting of matching bullet striae, we can also make use of this mechanism can be used to separate bullet curvature in profiles from signatures before matching signatures.
+ Originally, this smoothing is intended to remove drift and (sub)class characteristics from individual markings, however, in the setting of matching bullet striae, we can also make use of this mechanism to separate bullet curvature in profiles from signatures before matching signatures.
 
@@
 
-\hh{XXX We have been operating under a false assumption: smoothing the signature at f=1 DOES change it - see figure \ref{fig:profile-sketch}.
-We need to re-run some of the results to compare the results against a }
-
-```{r profile-sketch, fig.height = 5, fig.width='\\textwidth', fig.cap="Overview of the effect of different coarseness parameters $c$ on the profile shown in Figure \\ref{fig:sigs-profiles}. "}
+\hh{\autoref{fig:profile-sketch} gives an overview of the effect of different coarseness parameters: from left to right, coarseness levels $c$ are 0.2, 0.25, and 0.3, respectively.
+The top row shows the resulting signature after smoothing the profile shown in Figure~\ref{fig:sigs-profiles} with the coarseness specified.
+The histograms in the bottom row show the relative optimal location $t^o$ (optimal location $t^o$ divided by the length of the corresponding profile).
+Optimal locations should be distributed uniformly.
+However, for coarseness values of $c = 0.25$ and $c=0.30$ we see very distinct boundary effects: optimal locations $t^o$ are found at the very extreme ends of a profile more often than one would expect based on a uniform location.}
+
+```{r profile-sketch, fig.height = 5, fig.width='\\textwidth', fig.cap="Overview of the effect of different coarseness parameters $c$ on the profile shown in Figure \\ref{fig:sigs-profiles} (top).
+The bottom row shows histograms of  the (relative) optimal locations $t^o$ identified in the optimization step for different values of the coarseness parameter $c$.
+"}
 profile <- sigs_graphics[[1]] %>% filter(land_id==19)
@@
 )
-# profsumm <- read.csv("../data/profiles-summary.csv")
-# profsumm <- profsumm %>% select(-run_id)
-# profc <- profc %>% left_join(profsumm, by=c("land2_id"="land_id", "profile2_id"="profile_id"))
-# 
-# p2 <- profc %>% 
-#   ggplot(aes(x = locations/(length-wo))) +
-#   geom_histogram(binwidth = 0.025) +
-#   facet_grid(.~coarseness, labeller="label_both") + 
-#   theme_bw()
-#  
-# grid.arrange(p1, p2, ncol=1)
+
+profsumm <- read.csv("../data/profiles-summary.csv")
+profsumm <- profsumm %>% select(-run_id)
+profc <- profc %>% left_join(profsumm, by=c("land2_id"="land_id", "profile2_id"="profile_id"))
+
+p2 <- profc %>% 
+  ggplot(aes(x = locations/(length-wo))) +
+  geom_histogram(binwidth = 0.025) +
+  facet_grid(.~coarseness, labeller="label_both") + 
+  theme_bw() +
+  xlab(expression("Relative optimal location "~t^o)) +
+  ylab("Number of profiles")
+
+grid.arrange(p1, p2, ncol=1)
 ```
@@
 \hh{Figure \ref{fig:coarse} shows the type II error rates for profiles using an optimization window $w_o= 120$ and a validation window $w_v = 30$ for varying level of coarseness.
-The type II error for all nominal levels of  $\alpha$ are the lowest for a coarseness range of 0.20 to 0.35.
+The type II error for all nominal levels of  $\alpha$ are the lowest for a coarseness range of 0.20 to 0.30.
 For the remainder of the analysis we use a fixed value of coarseness of 0.25.

### 2018-04-16 doc/paper.tex
explanation of histograms

@@
 from individual markings, however, in the setting of matching bullet
-striae, we can also make use of this mechanism can be used to separate
-bullet curvature in profiles from signatures before matching signatures.
+striae, we can also make use of this mechanism to separate bullet
+curvature in profiles from signatures before matching signatures.
 
@@
 
-\hh{XXX We have been operating under a false assumption: smoothing the signature at f=1 DOES change it - see figure \ref{fig:profile-sketch}.
-We need to re-run some of the results to compare the results against a }
+\hh{XXX We have been operating under a false assumption: smoothing the signature at f=1 DOES change it - see figure \ref{fig:profile-sketch}. We need to re-run some of the results }
 
@@
 
-\caption{Overview of the effect of different coarseness parameters $c$ on the profile shown in Figure \ref{fig:sigs-profiles}. }\label{fig:profile-sketch}
+\caption{Overview of the effect of different coarseness parameters $c$ on the profile shown in Figure \ref{fig:sigs-profiles} (top).
+The bottom row shows histograms of  the optimal locations $t^o$ identified in the optimization step for different values of the coarseness parameter $c$.
+}\label{fig:profile-sketch}
 \end{figure}

### 2018-04-16 doc/paper.Rmd
data uploaded

@@
 The span of the loess fit was found using cross-validation, as described by \citet{aoas}.
- <!-- Signatures are considered to be noise free and a good reflection of the key attributes of the raw marking, and the unique features of a bullet.
-\hh{Do you have a reference for the previous sentenced?} A  detailed discussion of the extraction technique for signatures is given in \citet{aoas}.
--->
- <!--, where comprehensive details about the height at which profile is to be selected, removing curvature, smoothing, identifying groove locations are explained.-->
 
@@
 This severe limitation in the amount of available data poses the main challenge in adapting the Chumbley score to matching bullet lands, because of the resulting loss in power.
-
-<!--
-## Potential Challenges in Chumbley Score Adaptation
-
-Markings made by screw driver tips \citep{chumbley} are longer and pronounced and about 7 mm in length \citep{manytoolmarks1}, while the groove to groove length of the bullet lands from the Hamby study \citep{hamby} are about 2mm in length.
-The Ruger pistol barrels used in the  Hamby study have a 9mm caliber.
-This means that bullets used in these pistols have land markings of only about a quarter of the size of screwdriver toolmarks used by \citet{chumbley, hadler}, therefore raising the problem of potential loss in power of the test.
--->
-
-<!--One way to investigate this is to evaluate bullet striations at different resolutions.
-This gives us an incentive to use both the NIST and Iowa State scans and investigate the effect of length of the digitized mark on the working of the algorithm.
-Apart from this, the algorithm developed by \citet{hadler} has only been tested for flatter and wider surfaces which have negligible curvature, whereas striations on bullets are made on their curved surfaces.
-Therefore, using methods proposed for toolmarks need adaptation in order to give tangible results for bullets.
-Another challenge that arises from these topographical differences, is the understanding of that effect of different levels of smoothing.
-Especially investigating its effect on the algorithm performance when signatures are being compared and when raw markings or profiles are being compared.
-
-Some other challenges associated with the adaptation process call for a brief understanding of the methodology involved.
-The Chumbley method and algorithm works on the premise of comparison windows that are supposed to represent small segments of pixels of equal sizes in the two markings.
-These windows are to be compared in a predetermined fashion and are termed as windows of optimization and validation.
-The optimization window is selected in the first step called the optimization step where areas/windows of best agreement between the the two markings are found.
-
-The window of optimization for bullet striations are shorter as bullet signatures in terms of absolute length (in mm), are smaller compared to toolmarks.
-Even in terms of digitized pixels, there is no straightforward way of establishing what window size is to be chosen.
-One important way to address this is to keep the number of windows of optimization sufficiently large or not too small.
-This means, if the number of windows are too small we inadvertently use bigger window sizes as we want to compare all segements of one mark with the segments of the other mark.
-This leads to not giving enough emphasis to smaller segements of the two marking that might be in high agreement.
-In terms of absolute length we end up comparing smaller segments between the two markings while dealing with bullet lands, than those being compared for toolmarks.
-
-For too large window sizes, the weight of small features that would otherwise uniquely classify a signature and hence identify the region of agreement is vastly reduced.
-On the other hand, this introduces another problem as, if we go too small in the window of optimization, the unique features of the trace segments are lost and all seem similar.
-The problem of distinctiveness in the structure of bullet striations may also lead to potential failure of the algorithm to identify correct windows of agreements or not be able to identify these windows at all.
-This may lead to incorrect classification of bullets as coming from the same-source (or matches) or from different source (or non-matches).
-
-Thus, the comaparison windows have a direct influence on the false positives and false negatives, making identification of the optimum sizes of these windows very important.
-This raises important questions about how and what parameter settings need to be chosen for bullet land comparison.
-Something similar is done in \citet{afte-chumbley} for toolmark comparisons of slip-joint pliers where optimum window sizes are determined.
-There is a need to figure out the best parameter settings which minimizes the errors for unsmoothed markings or profiles and pre-processed signatures.
-An analysis of these error rates and comparison with other methods will help us understand the adaptibility of the chumbley score to bullets.
--->
 
@@
 ## Setup
+\hh{XXX the description below needs to be updated.}
 
@@
 The settings for optimizing and validating window sizes, $w_o$ and $w_v$, ranged from $w_o \in [50, 280]$ and $w_v \in \{30, 50\}$, see also figure \ref{fig:type2}.
+
+# Results
+
+## Profiles
+
+\hh{When dealing with profiles, coarseness is an additional parameter that has to be considered in the matching.}
+
+\hh{Figure \ref{fig:coarse} shows the type II error rates for profiles using an optimization window $w_o= 120$ and a validation window $w_v = 30$ for varying level of coarseness.
+The type II error for all nominal levels of  $\alpha$ are the lowest for a coarseness range of 0.20 to 0.30, with a minim in type II reached at a coarseness $c = 0.25$.
+}
+\hh{However, for an optimal value of coarseness other aspects have to be condsidered.}
+```{r coarse, fig.cap="Type II error with respect to coarseness parameter over profiles, $w_o = 120$, $w_v = 30$.
+Optimal values for coarseness are around $c \\approx 0.30$", fig.width=5, fig.height = 3, out.width='.6\\textwidth'}
+e1<-errors2 %>% filter(wv==30)%>% ggplot(aes(x = coarse, y=beta, colour=factor(alpha))) +
+  geom_point() + geom_line() +
+  theme_bw() +
+  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + xlab("Profile Coarseness")+
+  ylab("Type II error") 
+e1
+```
+
+\hh{\autoref{fig:profile-sketch} gives an overview of the effect of different coarseness parameters: from left to right, coarseness levels $c$ are 0.2, 0.25, and 0.3, respectively.
+The top row shows the resulting signature after smoothing the profile shown in Figure~\ref{fig:sigs-profiles} with the coarseness specified.
+The histograms in the bottom row show the relative optimal location $t^o$ (optimal location $t^o$ divided by the length of the corresponding profile).
+Optimal locations should be distributed uniformly.
+However, for coarseness values of $c = 0.25$ and $c=0.30$ we see very distinct boundary effects: optimal locations $t^o$ are found at the very extreme ends of a profile more often than one would expect based on a uniform location.}
+
+```{r profile-sketch, fig.height = 5, fig.width='\\textwidth', fig.cap="Overview of the effect of different coarseness parameters $c$ on the profile shown in Figure \\ref{fig:sigs-profiles} (top).
+The bottom row shows histograms of  the (relative) optimal locations $t^o$ identified in the optimization step for different values of the coarseness parameter $c$.
+"}
+profile <- sigs_graphics[[1]] %>% filter(land_id==19)
+profile$lw20 <- lowess(x = profile$y, y = profile$value, f=.20)$y
+profile$lw25 <- lowess(x = profile$y, y = profile$value, f=.25)$y
+profile$lw30 <- lowess(x = profile$y, y = profile$value, f=.30)$y
+#profile$lw50 <- lowess(x = profile$y, y = profile$value, f=.5)$y
+#profile$lw75 <- lowess(x = profile$y, y = profile$value, f=.75)$y
+#profile$lw100 <- lowess(x = profile$y, y = profile$value, f=1)$y
+
+profiles <- tidyr::gather(profile, coarseness, values, lw20:lw30)
+profiles <- profiles %>% mutate(
+  c = readr::parse_number(coarseness)/100
+)
+p1 <- profiles %>% 
+  ggplot(aes(x = y, y = value-values)) + geom_line() +
+  theme_bw() +
+  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
+  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) +
+  facet_grid(~c, labeller="label_both")
+
+profc20 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo:160-wv:30-c:20.rds")
+profc20$coarseness <- 0.2
+profc25 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo:160-wv:30-c:25.rds")
+profc25 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo:160-wv:50-c:25.rds")
+profc30 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo:160-wv:30.rds")
+profc30$coarseness <- .30
+profc <- bind_rows(profc20, profc25, profc30)
+profc <- profc %>% mutate(
+  locations = chumbley %>% purrr::map_int(.f = function(x) {
+    res <- NULL
+    try(res <- x$locations[1], silent = TRUE)
+    res
+  })
+)
+profsumm <- read.csv("../data/profiles-summary.csv")
+profsumm <- profsumm %>% select(-run_id)
+profc <- profc %>% left_join(profsumm, by=c("land2_id"="land_id", "profile2_id"="profile_id"))
+
+p2 <- profc %>% 
+  ggplot(aes(x = locations/(length-wo))) +
+  geom_histogram(binwidth = 0.025) +
+  facet_grid(.~coarseness, labeller="label_both") + 
+  theme_bw() +
+  xlab(expression("Relative optimal location "~t^o)) +
+  ylab("Number of profiles")
+
+grid.arrange(p1, p2, ncol=1)
+```
 
@@
 
-wo120_1<- wo120 %>% gather(type, error, beta) %>%  
+wo120_1<- wo120_profile %>% gather(type, error, beta) %>%  
   ggplot(aes(x = wv, y = error, colour = factor(alpha)))+
@@
 combined120<- 
-  ggplot(data = wo120 %>% gather(type, error, beta), aes(x = wv, y = error, colour = factor(alpha)))+
+  ggplot(data = wo120_profile %>% gather(type, error, beta), aes(x = wv, y = error, colour = factor(alpha)))+
   geom_line(aes(linetype="Signatures"))+  geom_point(shape= 24) + 
@@
 -->
-
+<!--
 # Appendix
@@
 
-```{r}
+```{r, eval=FALSE}
 #errors %>% ggplot(aes(x = wv, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + facet_grid(~wo) +geom_point(data=errors2, aes(shape="profiles"))
@@
 
-total2<- wo120 %>% filter(wv>10) %>% ggplot(aes(x = wv, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + geom_line() +geom_point(data=errors2 %>% filter(wv>10) %>% filter(coarse==0.25), aes(shape="profiles")) + geom_line(data=errors2 %>% filter(wv>10)%>% filter(coarse==0.25), aes(shape="profiles")) + theme_bw() +
+total2<- wo120_profile %>% filter(wv>10) %>% ggplot(aes(x = wv, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + geom_line() +geom_point(data=errors2 %>% filter(wv>10) %>% filter(coarse==0.25), aes(shape="profiles")) + geom_line(data=errors2 %>% filter(wv>10)%>% filter(coarse==0.25), aes(shape="profiles")) + theme_bw() +
   scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + ylab("Total error") + ggtitle(label = "(e)")
@@
 \end{appendix} 
+-->

### 2018-04-16 doc/paper.tex
data uploaded

@@
 
+\hh{XXX the description below needs to be updated.}
+
 \hh{using (a) signatures and (b) profiles, run chumbley score across scans from NIST and CSAFE for various settings of $w_o$ and $w_v$ (and coarseness $c$ for profiles).}
@@
 
-\subsection{Results}\label{results}
+\section{Results}\label{results}
+
+\subsection{Profiles}\label{profiles}
+
+\hh{When dealing with profiles, coarseness is an additional parameter that has to be considered in the matching.}
+
+\hh{Figure \ref{fig:coarse} shows the type II error rates for profiles using an optimization window $w_o= 120$ and a validation window $w_v = 30$ for varying level of coarseness.
+The type II error for all nominal levels of  $\alpha$ are the lowest for a coarseness range of 0.20 to 0.30, with a minim in type II reached at a coarseness $c = 0.25$.
+}
+\hh{However, for an optimal value of coarseness other aspects have to be condsidered.}
+
+\begin{figure}
+
+{\centering \includegraphics[width=.6\textwidth]{figures/coarse-1}
+
+}
+
+\caption{Type II error with respect to coarseness parameter over profiles, $w_o = 120$, $w_v = 30$. Optimal values for coarseness are around $c \approx 0.30$}\label{fig:coarse}
+\end{figure}
+
+\hh{\autoref{fig:profile-sketch} gives an overview of the effect of different coarseness parameters: from left to right, coarseness levels $c$ are 0.2, 0.25, and 0.3, respectively.
+The top row shows the resulting signature after smoothing the profile shown in Figure~\ref{fig:sigs-profiles} with the coarseness specified.
+The histograms in the bottom row show the relative optimal location $t^o$ (optimal location $t^o$ divided by the length of the corresponding profile).
+Optimal locations should be distributed uniformly.
+However, for coarseness values of $c = 0.25$ and $c=0.30$ we see very distinct boundary effects: optimal locations $t^o$ are found at the very extreme ends of a profile more often than one would expect based on a uniform location.}
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figures/profile-sketch-1}
+
+}
+
+\caption{Overview of the effect of different coarseness parameters $c$ on the profile shown in Figure \ref{fig:sigs-profiles} (top).
+The bottom row shows histograms of  the (relative) optimal locations $t^o$ identified in the optimization step for different values of the coarseness parameter $c$.
+}\label{fig:profile-sketch}
+\end{figure}
+
+\subsection{Results}\label{results-1}
 
@@
 
-\section{Results}\label{results-1}
-
-\subsection{Profiles}\label{profiles}
-
-\hh{When dealing with profiles, coarseness is an additional parameter that has to be considered in the matching.}
-
-\hh{XXX We have been operating under a false assumption: smoothing the signature at f=1 DOES change it - see figure \ref{fig:profile-sketch}. We need to re-run some of the results }
-
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figures/profile-sketch-1}
-
-}
-
-\caption{Overview of the effect of different coarseness parameters $c$ on the profile shown in Figure \ref{fig:sigs-profiles} (top).
-The bottom row shows histograms of  the optimal locations $t^o$ identified in the optimization step for different values of the coarseness parameter $c$.
-}\label{fig:profile-sketch}
-\end{figure}
-
-\hh{Figure \ref{fig:coarse} shows the type II error rates for profiles using an optimization window $w_o= 120$ and a validation window $w_v = 30$ for varying level of coarseness.
-The type II error for all nominal levels of  $\alpha$ are the lowest for a coarseness range of 0.20 to 0.30.
-For the remainder of the analysis we use a fixed value of coarseness of 0.25.
-}
-
-\begin{figure}
-
-{\centering \includegraphics[width=.6\textwidth]{figures/coarse-1}
-
-}
-
-\caption{Type II error with respect to coarseness parameter over profiles, $w_o = 120$, $w_v = 30$. Optimal values for coarseness are around $c \approx 0.30$}\label{fig:coarse}
-\end{figure}
-
 \begin{table}[ht]
@@
 
-\section{Appendix}\label{appendix}
-
-\begin{appendix}
-
-On the other hand Figure \ref{fig:prof_missings} (b) shows if the coarseness level set in the chumbley agorithm has any effect on the signatures, which are pre-processed and already smoothed to a certain extent.
-From Figure \ref{fig:prof_missings} (b) we can notice that for different nominal $\alpha$ levels, the type II error fluctuates slightly but does not change much, thereby helping us conclude that the coarseness levels set in the LOWESS smoothing in the chumbley alggorithm does effect the type II error much for signatures.
-
-\subsection{Comparison of profiles and signatures}
-
-Another reason for failed tests can be incorrect identification of maximum correlation windows in the optimization step as seen in figure \ref{fig:prof_missings}(d) because of the level of smoothing, as too much smoothing would subdue intricate features that might otherwise help in the correlation calculations and correct identification of maximum correlation windows irrespective of the size.
-This would again cause a simiar effect as explained for figure \ref{fig:missings} with validation windows, irrespective of size, during the shifts end up at the ends of the markings resulting in an invalid calculation and failed comparison attempt.
-
-In figure \ref{fig:prof_missings}(d) and (f), we compare profiles and signatures on the basis of number of failed tests.
-The profiles chosen for figure \ref{fig:prof_missings}(f) have a constant coarseness of 0.25 and window of optimization as 120.
-The signatures in this case are not smoothed using the chumbley algorithm step of LOWESS smoothing.
-Instead signatures are used as calculated by \citet{aoas}.
-The smoothing in these signatures were determined and fixed on the basis of their performance in the random forest based algorithm proposed by \citet{aoas}.
-The comparison of profiles and signatures with variation of validation window size therefore is made on even footing.
-The trends are similar to figure \ref{fig:missings} in the sense that for known non-matches the number of failed tests are more for both signatures and profiles and increasing linearly with the validation window size.
-The problem is however, worse for profiles which has higher number of failed tests than signatures for all validation windows.
-
-The total error for different validation window sizes for signatures and profiles can be seen in figure \ref{fig:prof_missings} (e).The optimization window size is 120 and profiles are calculated at a default 0.25 coarseness level while signatures as before are not smoothed again in the modified chumbley algorithm.
-We can see that the total error is always higher for profiles as compared to signatures for all sizes of validation window.
-
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figures/prof_missings-1}
-
-}
-
-\caption{Row 3:  Total error and Number of failed tests by the window validation size, wv, and ground truth, Row 2: Total error and Number of failed tests with Coarseness for both profiles and signatures, Row 1: Type II error for different coarseness levels as used in the modified chumbley algorithm for profiles and signatures}\label{fig:prof_missings}
-\end{figure}
-\end{appendix}
-
 \bibliographystyle{agsm}

### 2018-04-16 doc/paper.Rmd
re-structuring

@@
 
-# Results
-<!--
-We used the adjusted Chumbley method as proposed in \citet{hadler} and implemented in the R package `toolmaRk` \citep{toolmark} on all pairwise land-to-land comparisons of the Hamby scans (a total of 85,491 comparisons) with the pairwise sets for the comparisons given in the table \ref{tab:param}.
-```{r param}
-param_setting<- all %>% select(wo, wv) %>% group_by(wo, wv) %>% tally() %>% select(wo,wv)
-#param_setting<- data.frame(matrix(param_setting, nrow = 16))
-#t(param_setting)
-#param_setting %>% tally() 
-library(kableExtra)
-#knitr::kable(t(param_setting)[,1:17], digits=0, caption = "Overview of parameter settings used for optimization and validation windows for bullet land signatures.", format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("scale_down", "hold"))
-
-par_tab1<-t(param_setting)[,1:18]
-par_tab2<-t(param_setting)[,19:33]
-knitr::kable(list(par_tab1,par_tab2), digits=0,caption = "Overview of parameter settings used for optimization and validation windows for bullet land signatures.", format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("scale_down","hold"))%>% kableExtra::kable_styling(latex_options = c("scale_down","hold")) #%>% group_rows("",1,2) %>% group_rows("",3,4)
-
-```
-
-```{r confusion}
-library(kableExtra)
-a27<- all %>% filter(wo==280, wv==30) %>% xtabs(data= ., ~signif+match)
-a27<- as.data.frame(a27)
-a27$Type<- c("True Negative","False Positive (Type I)","False Negative (Type II)","True Positive")
-
-a28<- all %>% filter(wo==120, wv==30) %>% xtabs(data= ., ~signif+match)
-a28<- as.data.frame(a28)
-a28$Type<- c("True Negative","False Positive (Type I)","False Negative (Type II)","True Positive")
-
-a29<- all %>% filter(wo==80, wv==30) %>% xtabs(data= ., ~signif+match)
-a29<- as.data.frame(a29)
-a29$Type<- c("True Negative","False Positive (Type I)","False Negative (Type II)","True Positive")
-
-knitr::kable(list(a27,a28,a29), digits=0, caption = "Confusion Table for different optimization window sizes with validation window size as 30.", format = "latex" ,booktabs=TRUE)%>% kableExtra::kable_styling(latex_options = c("hold"), "striped", full_width = F) %>% group_rows("Size of Optimization Window = 280",1,5) %>% group_rows("Size of Optimization Window = 120", 6, 10) %>%  group_rows("Size of Optimization Window = 80", 11,14)
-```
-
-## Signatures
--->
-<!--
-Figure \ref{fig:type2} gives an overview of type II error rates observed when varying the window size in the optimization step.
-Two levels of validation window size 30 and 50 were chosen as to compare the error rates for different nominal type I errors.
-We notice that the trends for these nominal type I errors are similar and in most cases a validation window of 50 has higher type II error than for 30.
-A change in this trend is seen for a 0.05 $\alpha$ level, although the difference between the two windows is very small for this case.
-We can also notice an obvious trend of increase in the Type II error as the window of optimization increases and see a minimum around the optimization window size of 120 pixels.
-Hence we are inclined to choose a smaller validation window size and optimization window as 120.
-
-Table \ref{tab:confusion} shows the confusion tables with the classification of type I and type II errors and how the numbers change with a change in the optimization window.
-The windows represent areas to the left of the window with minimum type II, near the minimum type II window and to the far right of the minimum type II error
-
-Figure \ref{fig:type1} compares nominal (fixed) type I error and actually observed type I errors for the parameter settings in table \ref{tab:param}.
-With an increasing size of the window used in the optimization step the observed type I error rate decreases (slighty).
-This means as the optimization window increase the observed type I error rate gets smaller.
-A smaller validation window on the other hand, tends to be associated with a higher type I error rate.
-This can be better imagined for a given window of optimization, where the actual Type I error is comparable to the nominal level for only a select few validation window sizes.
-For these comparable validation window sizes of 30 and 50 as done here, the actual type I error increases very slightly and can be seen in Figure \ref{fig:type1}.
-This increase is not as much when compared to the variation seen with the optimization window sizes.
-This effect might be related to the increasing number of tests that fail for larger optimization window sizes, in particular for non-matching striae (see fig \ref{fig:missings}).
-
-```{r wo120, fig.width = 4, fig.height=4, fig.cap="The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120"}
-wo120 <- errors %>% filter(wo==120)
-wo120 %>% gather(type, error, beta) %>%#actual:beta) %>%
-  ggplot(aes(x = wv, y = error, colour = factor(alpha)))  +
-  geom_line() + geom_point() + facet_wrap(~type, scales="free") +
-  theme_bw()+
-  theme(legend.position="bottom") +  ggtitle("Signatures" , subtitle = "Varying validation window sizes, Optimization window = 120")
-```
-
-The actual type I error and type II error for signatures were also compared for different validation window sizes.
-Figure \ref{fig:wo120} shows the actual rates for different nominal $\alpha$ levels.
-We can see that the type II error rises with higher validation windows for the smaller nominal $\alpha$ levels while for the nominal $\alpha$ = 0.05 its almost constant.
-```{r }
-
-nsp <- allp %>% group_by(wv, wo, match, coarse) %>% summarize(
-  n_ss = mean(same_shift_n),
-  n_ds = mean(diff_shift_n),
-  zero_ss = sum(same_shift_n==0),
-  zero_ds = sum(diff_shift_n==0),
-  miss = sum(is.na(p_value)),
-  missperc = miss/85491*100
-)
-
-ns_sig_c <- all_sig_c %>% group_by(wv, wo, match, coarse) %>% summarize(
-  n_ss = mean(same_shift_n),
-  n_ds = mean(diff_shift_n),
-  zero_ss = sum(same_shift_n==0),
-  zero_ds = sum(diff_shift_n==0),
-  miss = sum(is.na(p_value)),
-  missperc = miss/85491*100
-)
-
-```
-
-```{r lms, results='asis'}
-nsnest <- ns %>% filter(wv %in% c(30,50)) %>% group_by(match, wv) %>% nest()
-nsnest <- nsnest %>% mutate(
-  model = data %>% purrr::map(.f = function(d) lm(I(missperc*100)~wo, data=d)),
-  coefs = model %>% purrr::map(.f = broom::tidy)
-)
-tab <- nsnest %>% unnest(coefs) %>% filter(term=="wo") %>% arrange(match) %>%
-  select(-term, -statistic, -p.value)
-knitr::kable(tab, digits=3, caption = "Estimates of the increase in percent of failed tests corresponding to a 100 point increase in the optimization window.", booktabs=TRUE)
-```
--->
-<!--
-Figure \ref{fig:missings} gives an overview of the number of failed tests, i.e.
-tests in which a particular parameter setting did not return a valid result.
-This happens, when the shift to align two signatures is so large, that the remaining overlap is too small to accommodate  windows for validation.
-The problem is therefore exacerbated by a larger validation window.
-Figure \ref{fig:missings} also shows that the number of failed tests is approximately linear in the size of the optimization window.
-Test results from different sources have a much higher chance to fail, raising the question, whether failed tests should be treated as rejections of the null hypothesis of same source.
-For known non-matches there is a higher possibility that in the optimization pair of windows where cross-correlations are maximum are too far apart, and same shifts of this order hit the end of the signature.
--->
-
-## Profiles
-
-\hh{When dealing with profiles, coarseness is an additional parameter that has to be considered in the matching.}
-
-\hh{\autoref{fig:profile-sketch} gives an overview of the effect of different coarseness parameters: from left to right, coarseness levels $c$ are 0.2, 0.25, and 0.3, respectively.
-The top row shows the resulting signature after smoothing the profile shown in Figure~\ref{fig:sigs-profiles} with the coarseness specified.
-The histograms in the bottom row show the relative optimal location $t^o$ (optimal location $t^o$ divided by the length of the corresponding profile).
-Optimal locations should be distributed uniformly.
-However, for coarseness values of $c = 0.25$ and $c=0.30$ we see very distinct boundary effects: optimal locations $t^o$ are found at the very extreme ends of a profile more often than one would expect based on a uniform location.}
-
-```{r profile-sketch, fig.height = 5, fig.width='\\textwidth', fig.cap="Overview of the effect of different coarseness parameters $c$ on the profile shown in Figure \\ref{fig:sigs-profiles} (top).
-The bottom row shows histograms of  the (relative) optimal locations $t^o$ identified in the optimization step for different values of the coarseness parameter $c$.
-"}
-profile <- sigs_graphics[[1]] %>% filter(land_id==19)
-profile$lw20 <- lowess(x = profile$y, y = profile$value, f=.20)$y
-profile$lw25 <- lowess(x = profile$y, y = profile$value, f=.25)$y
-profile$lw30 <- lowess(x = profile$y, y = profile$value, f=.30)$y
-#profile$lw50 <- lowess(x = profile$y, y = profile$value, f=.5)$y
-#profile$lw75 <- lowess(x = profile$y, y = profile$value, f=.75)$y
-#profile$lw100 <- lowess(x = profile$y, y = profile$value, f=1)$y
-
-profiles <- tidyr::gather(profile, coarseness, values, lw20:lw30)
-profiles <- profiles %>% mutate(
-  c = readr::parse_number(coarseness)/100
-)
-p1 <- profiles %>% 
-  ggplot(aes(x = y, y = value-values)) + geom_line() +
-  theme_bw() +
-  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
-  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) +
-  facet_grid(~c, labeller="label_both")
-
-profc20 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo_160-wv_30-c_20.rds")
-profc20$coarseness <- 0.2
-profc25 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo_160-wv_30-c_25.rds")
-profc30 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo_160-wv_30.rds")
-profc30$coarseness <- .30
-profc <- bind_rows(profc20, profc25, profc30)
-profc <- profc %>% mutate(
-  locations = chumbley %>% purrr::map_int(.f = function(x) {
-    res <- NULL
-    try(res <- x$locations[1], silent = TRUE)
-    res
-  })
-)
-
-profsumm <- read.csv("../data/profiles-summary.csv")
-profsumm <- profsumm %>% select(-run_id)
-profc <- profc %>% left_join(profsumm, by=c("land2_id"="land_id", "profile2_id"="profile_id"))
-
-p2 <- profc %>% 
-  ggplot(aes(x = locations/(length-wo))) +
-  geom_histogram(binwidth = 0.025) +
-  facet_grid(.~coarseness, labeller="label_both") + 
-  theme_bw() +
-  xlab(expression("Relative optimal location "~t^o)) +
-  ylab("Number of profiles")
-
-grid.arrange(p1, p2, ncol=1)
-```
-
-\hh{Figure \ref{fig:coarse} shows the type II error rates for profiles using an optimization window $w_o= 120$ and a validation window $w_v = 30$ for varying level of coarseness.
-The type II error for all nominal levels of  $\alpha$ are the lowest for a coarseness range of 0.20 to 0.30.
-For the remainder of the analysis we use a fixed value of coarseness of 0.25.
-}
-
-```{r coarse, fig.cap="Type II error with respect to coarseness parameter over profiles, $w_o = 120$, $w_v = 30$.
-Optimal values for coarseness are around $c \\approx 0.30$", fig.width=5, fig.height = 3, out.width='.6\\textwidth'}
-e1<-errors2 %>% filter(wv==30)%>% ggplot(aes(x = coarse, y=beta, colour=factor(alpha))) +
-  geom_point() + geom_line() +
-  theme_bw() +
-  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + xlab("Profile Coarseness")+
-  ylab("Type II error") 
-e1
-```
-
 ```{r, include =F}

### 2018-04-17 doc/paper.tex
re-structuring
defining relative optimal locations and expectations

@@
 \(t_1^o - t_2^o\) for aligning the two markings.
+\hh{We will denote the relative optimal locations as $t_1^*$ and $t_2^*$, where $t_i^* = t_i^o/T_i$ for $i=1,2$, such that $t_1^*, t_2^* \in [0,1]$.
+Once (sub-)class characteristics are removed, the relative optimal locations should be distributed according to a uniform distribution in $[0,1]$.
+}
 

### 2018-04-17 doc/paper.Rmd
defining relative optimal locations and expectations
structure

@@
 This results in an optimal vertical (in-phase) shift of $t_1^o - t_2^o$ for aligning the two markings.
+\hh{We will denote the relative optimal locations as $t_1^*$ and $t_2^*$, where $t_i^* = t_i^o/(T_i-w_o)$ for $i=1,2$, such that $t_1^*, t_2^* \in [0,1]$.
+Once (sub-)class characteristics are removed, the relative optimal locations should be distributed according to a uniform distribution in $[0,1]$.
+}
 
@@
 In the first set, pairs of windows are extracted from the two markings using the optimal vertical shift as determined in the first step, whereas for the second set the windows are extracted using a different (out-of-phase) shift.
+
+\hh{XXX summarise in formula}
 
@@
 
-# Results
+## Failed Tests
+
+\hh{number of failed tests for default parameter settings -- XXX what are the defaults? wo:10 percent, coarseness 0.25, wv XXX?}
+
+\hh{diagnose problem: use of optimal location in getting the pairs for different-shift correlations}
+\hh{figure: number of diff shift pairs and location}
+
+# A modified test
+
+\hh{fix the algorithm - all of the other properties in Hadler and morris still hold.}
+
+\hh{number of failed tests down. }
+
+# Comparison of Results
 
@@
 
-profc20 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo:160-wv:30-c:20.rds")
+profc20 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo_160-wv_30-c_20.rds")
 profc20$coarseness <- 0.2
-profc25 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo:160-wv:30-c:25.rds")
-profc25 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo:160-wv:50-c:25.rds")
-profc30 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo:160-wv:30.rds")
+profc25 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo_160-wv_30-c_25.rds")
+profc30 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo_160-wv_30.rds")
 profc30$coarseness <- .30
@@
   theme_bw() +
-  xlab(expression("Relative optimal location "~t^o)) +
+  xlab(expression("Relative optimal location "~t^*)) +
   ylab("Number of profiles")

### 2018-04-17 doc/paper.tex
structure

@@
 \(t_1^o - t_2^o\) for aligning the two markings.
-\hh{We will denote the relative optimal locations as $t_1^*$ and $t_2^*$, where $t_i^* = t_i^o/T_i$ for $i=1,2$, such that $t_1^*, t_2^* \in [0,1]$.
+\hh{We will denote the relative optimal locations as $t_1^*$ and $t_2^*$, where $t_i^* = t_i^o/(T_i-w_o)$ for $i=1,2$, such that $t_1^*, t_2^* \in [0,1]$.
 Once (sub-)class characteristics are removed, the relative optimal locations should be distributed according to a uniform distribution in $[0,1]$.
@@
 shift.
+
+\hh{XXX summarise in formula}
 

### 2018-04-18 doc/paper.Rmd
define q_i

@@
 $t$ here denotes equally spaced pixel locations for the striation marks under consideration, $t = 1, ..., T$.
-Let further $z(s,t)$ denote the vector of markings between locations $s$ and $t$.
+Let further $z^s(t)$ denote a vector of markings of length $s$ starting in location  $t$.
 
@@
 \[
-\left(t_1^o, t_2^o\right) = \mathop{\arg \max}\limits_{1 \le t_1 \le T_1, 1 \le t_2 \le T_2} \text{cor} \left(x (t_1, t_1 + w_o), y(t_2, t_2 + w_o) \right)
+\left(t_1^o, t_2^o\right) = \mathop{\arg \max}\limits_{1 \le t_1 \le T_1, 1 \le t_2 \le T_2} \text{cor} \left(x^{w_o} (t_1), y^{w_o}(t_2) \right)
 \]
@@
 
-\hh{XXX summarise in formula}
+More precisely, \citet{hadler} define same-shift pairs of length $w_v$ to start in $(q^{(1)}_i, q^{(2)}_i)$, where $q^{(k)}_i$ has the form
+\begin{eqnarray*}
+q^{(k)}_i &=& t_k^* + i w_v \text{ for } i = -1, -2, ...\\
+q^{(k)}_i &=& t_k^* + w_ o + i w_v \text{ for } i = 0, 1, 2, ...
+\end{eqnarray*}
+for all integer values of $i$ for which both $0 <  q^{(1)}_i \le T_1 - w_v$ and $0 <  q^{(2)}_i \le T_2 - w_v$.
+
+Similarly, different-shift pairs are based on signature segments of length $w_v$ starting in $(q^{(1)}_i, q^{(2)}_{-i})$, for all integer values of $i$ for which both $0 <  q^{(1)}_i \le T_1 - w_v$ and $0 <  q^{(2)}_{-i} \le T_2 - w_v$.
 
@@
 
-We used the adjusted Chumbley method as proposed in \citet{hadler} and implemented in the R package `toolmaRk` \citep{toolmark} on all pairwise land-to-land comparisons of the Hamby scans provided by NIST (a total of 85,491 comparisons).
-The settings for optimizing and validating window sizes, $w_o$ and $w_v$, ranged from $w_o \in [50, 280]$ and $w_v \in \{30, 50\}$, see also figure \ref{fig:type2}.
+We used the adjusted Chumbley method as proposed in \citet{hadler} and implemented in the R package `toolmaRk` \citep{toolmark} on all pairwise land-to-land comparisons of the Hamby scans provided by NIST for a total of 85,491 land-to-land comparisons.
+
+<!--The settings for optimizing and validating window sizes, $w_o$ and $w_v$, ranged from $w_o \in [50, 280]$ and $w_v \in \{30, 50\}$, see also figure \ref{fig:type2}.-->
 
@@
 
-\hh{number of failed tests for default parameter settings -- XXX what are the defaults? wo:10 percent, coarseness 0.25, wv XXX?}
+Initially, the default settings suggested in \citet{hadler} were used: $w_o = 120$ pixels or about 190 $\mu m$ (ten percent of the average length of profiles) and coarseness $c = 0.25$.
+\autoref{fig:fails} shows the percentage of failed tests among the 86k land-to-land comparisons of the NIST data for different values of the validation window size $w_v$.
+For same-source lands up to 12.5 percent of the tests fail.
+
+```{r fails, echo=FALSE, fig.height = 3, fig.width = 5, out.width='0.7\\textwidth', fig.cap="Percent of failed land-to-land comparisons using an optimization window $w_o = 120$ and a coarseness of $c = 0.25$.
+With an increase in the size of the validation window  a higher percentage of tests fails.
+The percentage of failures is higher for same-source lands than for different source lands."}
+fails <- allp %>% filter(coarse == 0.25) %>% 
+  group_by(wo, wv) %>% summarize(
+  failed = sum(is.na(p_value)),
+  n=n()
+)
+
+allp %>% filter(coarse == 0.25) %>% 
+  group_by(wo, wv, match) %>% summarize(
+  failed = sum(is.na(p_value)),
+  n=n()
+) %>% 
+  ggplot(aes(x = wv, y = failed/n*100, colour=factor(match), shape=factor(match))) +
+  geom_point(size=3) +
+  xlab("Size of validation window in pixels") +
+  ylab("Percent of failed tests") +
+  theme_bw() +
+  scale_colour_brewer("Same-source", palette="Set1") +
+  scale_shape_discrete("Same-source") +
+  ylim(c(0,13))
+```
 
@@
   theme_bw() +
-  xlab(expression("Relative optimal location "~t^*)) +
+  xlab(expression("Relative optimal location "~t^"*")) +
   ylab("Number of profiles")

### 2018-04-18 doc/paper.tex
define q_i

@@
 pixel locations for the striation marks under consideration,
-\(t = 1, ..., T\). Let further \(z(s,t)\) denote the vector of markings
-between locations \(s\) and \(t\).
+\(t = 1, ..., T\). Let further \(z^s(t)\) denote a vector of markings of
+length \(s\) starting in location \(t\).
 
@@
 \(x(t_1)\) and \(y(t_2)\) is maximized: \[
-\left(t_1^o, t_2^o\right) = \mathop{\arg \max}\limits_{1 \le t_1 \le T_1, 1 \le t_2 \le T_2} \text{cor} \left(x (t_1, t_1 + w_o), y(t_2, t_2 + w_o) \right)
+\left(t_1^o, t_2^o\right) = \mathop{\arg \max}\limits_{1 \le t_1 \le T_1, 1 \le t_2 \le T_2} \text{cor} \left(x^{w_o} (t_1), y^{w_o}(t_2) \right)
 \] This results in an optimal vertical (in-phase) shift of
@@
 
-\hh{XXX summarise in formula}
+More precisely, \citet{hadler} define same-shift pairs of length \(w_v\)
+to start in \((q^{(1)}_i, q^{(2)}_i)\), where \(q^{(k)}_i\) has the form
+
+\begin{eqnarray*}
+q^{(k)}_i &=& t_k^* + i w_v \text{ for } i = -1, -2, ...\\
+q^{(k)}_i &=& t_k^* + w_ o + i w_v \text{ for } i = 0, 1, 2, ...
+\end{eqnarray*}
+
+for all integer values of \(i\) for which both
+\(0 < q^{(1)}_i \le T_1 - w_v\) and \(0 < q^{(2)}_i \le T_2 - w_v\).
+
+Similarly, different-shift pairs are based on signature segments of
+length \(w_v\) starting in \((q^{(1)}_i, q^{(2)}_{-i})\), for all
+integer values of \(i\) for which both \(0 < q^{(1)}_i \le T_1 - w_v\)
+and \(0 < q^{(2)}_{-i} \le T_2 - w_v\).
 
@@
 implemented in the R package \texttt{toolmaRk} \citep{toolmark} on all
-pairwise land-to-land comparisons of the Hamby scans provided by NIST (a
-total of 85,491 comparisons). The settings for optimizing and validating
-window sizes, \(w_o\) and \(w_v\), ranged from \(w_o \in [50, 280]\) and
-\(w_v \in \{30, 50\}\), see also figure \ref{fig:type2}.
+pairwise land-to-land comparisons of the Hamby scans provided by NIST
+for a total of 85,491 land-to-land comparisons.
 
-\section{Results}\label{results}
+\subsection{Failed Tests}\label{failed-tests}
+
+Initially, the default settings suggested in \citet{hadler} were used:
+\(w_o = 120\) pixels or about 190 \(\mu m\) (ten percent of the average
+length of profiles) and coarseness \(c = 0.25\). \autoref{fig:fails}
+shows the percentage of failed tests among the 86k land-to-land
+comparisons of the NIST data for different values of the validation
+window size \(w_v\). For same-source lands up to 12.5 percent of the
+tests fail.
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.7\textwidth]{figures/fails-1}
+
+}
+
+\caption{Percent of failed land-to-land comparisons using an optimization window $w_o = 120$ and a coarseness of $c = 0.25$.
+With an increase in the size of the validation window  a higher percentage of tests fails.
+The percentage of failures is higher for same-source lands than for different source lands.}\label{fig:fails}
+\end{figure}
+
+\hh{diagnose problem: use of optimal location in getting the pairs for different-shift correlations}
+\hh{figure: number of diff shift pairs and location}
+
+\section{A modified test}\label{a-modified-test}
+
+\hh{fix the algorithm - all of the other properties in Hadler and morris still hold.}
+
+\hh{number of failed tests down. }
+
+\section{Comparison of Results}\label{comparison-of-results}
 
@@
 
-\subsection{Results}\label{results-1}
+\subsection{Results}\label{results}
 

### 2018-04-18 doc/paper.Rmd
starting points defined

@@
 \newcommand{\cited}[1]{{\textcolor{red}{#1}}}
+\setlength\parindent{0pt}
 
@@
 
-More precisely, \citet{hadler} define same-shift pairs of length $w_v$ to start in $(q^{(1)}_i, q^{(2)}_i)$, where $q^{(k)}_i$ has the form
+More precisely, let us define starting points $s_i^{(k)}$ for each signature $k = 1, 2$ as 
 \begin{eqnarray*}
-q^{(k)}_i &=& t_k^* + i w_v \text{ for } i = -1, -2, ...\\
-q^{(k)}_i &=& t_k^* + w_ o + i w_v \text{ for } i = 0, 1, 2, ...
-\end{eqnarray*}
-for all integer values of $i$ for which both $0 <  q^{(1)}_i \le T_1 - w_v$ and $0 <  q^{(2)}_i \le T_2 - w_v$.
-
-Similarly, different-shift pairs are based on signature segments of length $w_v$ starting in $(q^{(1)}_i, q^{(2)}_{-i})$, for all integer values of $i$ for which both $0 <  q^{(1)}_i \le T_1 - w_v$ and $0 <  q^{(2)}_{-i} \le T_2 - w_v$.
+s^{(k)}_i = 
+\begin{cases}
+t_k^* + i w_v & \text{ for } i < 0 \\
+t_k^* + w_ o + i w_v & \text{ for } i \ge 0,
+\end{cases}
+\end{eqnarray*} for integer values of $i$ with $0 <  s^{(k)}_i \le T_k - w_v$.
+
+Same-shift pairs of length $w_v$ are defined in \citet{hadler} as all 
+pairs $(s_i^{(1)}, s_i^{(2)})$ for all integer values $i$ for which both $s_i^{(1)}$ and  $s_i^{(2)}$ are defined.
+Similarly, different-shift pairs are defined as $(s_i^{(1)}, s_{-i}^{(2)})$ for all $i$ where both $s_i^{(1)}$ and  $s_{-i}^{(2)}$ are defined.
 
@@
 
-For both samples the correlations between the pairs of markings is then calculated. 
+For both same- and different-shift pairs correlations between the  markings are calculated. 
 The intuition here is that for two markings from the same source the correlation for the in-phase sample should be high, while the correlations of the out-of-phase sample  provide a measure for the base-level correlation for non-matching marks of a given length $w_v$.
 The Chumbley score is then computed as a Mann Whitney U statistic to compare between in-phase sample and out-of-phase sample.
-In the original method proposed in \citet{chumbley} both in-phase and out-of-phase sample are extracted randomly, whereas \citet{hadler} proposed deterministic rules for both samples to make the resulting score deterministic while simultaneously avoiding overlaps within selected marks to ensure independence.
+In the original method proposed in \citet{chumbley} both in-phase and out-of-phase sample are extracted randomly, whereas \citet{hadler} proposed the above specified deterministic rules for both samples to make the resulting score deterministic while simultaneously avoiding overlaps within selected marks to ensure independence.
 

### 2018-04-18 doc/paper.tex
starting points defined

@@
 \newcommand{\cited}[1]{{\textcolor{red}{#1}}}
+
+\setlength\parindent{0pt}
 
@@
 
-More precisely, \citet{hadler} define same-shift pairs of length \(w_v\)
-to start in \((q^{(1)}_i, q^{(2)}_i)\), where \(q^{(k)}_i\) has the form
+More precisely, let us define starting points \(s_i^{(k)}\) for each
+signature \(k = 1, 2\) as
 
 \begin{eqnarray*}
-q^{(k)}_i &=& t_k^* + i w_v \text{ for } i = -1, -2, ...\\
-q^{(k)}_i &=& t_k^* + w_ o + i w_v \text{ for } i = 0, 1, 2, ...
+s^{(k)}_i = 
+\begin{cases}
+t_k^* + i w_v & \text{ for } i < 0 \\
+t_k^* + w_ o + i w_v & \text{ for } i \ge 0,
+\end{cases}
 \end{eqnarray*}
 
-for all integer values of \(i\) for which both
-\(0 < q^{(1)}_i \le T_1 - w_v\) and \(0 < q^{(2)}_i \le T_2 - w_v\).
+for integer values of \(i\) with \(0 < s^{(k)}_i \le T_k - w_v\).
 
-Similarly, different-shift pairs are based on signature segments of
-length \(w_v\) starting in \((q^{(1)}_i, q^{(2)}_{-i})\), for all
-integer values of \(i\) for which both \(0 < q^{(1)}_i \le T_1 - w_v\)
-and \(0 < q^{(2)}_{-i} \le T_2 - w_v\).
+Same-shift pairs of length \(w_v\) are defined in \citet{hadler} as all
+pairs \((s_i^{(1)}, s_i^{(2)})\) for all integer values \(i\) for which
+both \(s_i^{(1)}\) and \(s_i^{(2)}\) are defined. Similarly,
+different-shift pairs are defined as \((s_i^{(1)}, s_{-i}^{(2)})\) for
+all \(i\) where both \(s_i^{(1)}\) and \(s_{-i}^{(2)}\) are defined.
 
@@
 
-For both samples the correlations between the pairs of markings is then
+For both samples the correlations between the pairs of markings are then
 calculated. The intuition here is that for two markings from the same

### 2018-04-18 doc/paper.Rmd
expected failure rate

@@
 AFTE describes the term "sufficient agreement" as the possibility of another tool producing the markings under comparison, as practically impossible \citep{afte-toolmarks1998}.
-This subjectivity in the assessment as well as the lack of error rates are the main points of criticisms  first raised by the National Research Council in 2009 \citep{NAS:2009} and later emphasized further by the President's Council of Advisors on Science and Technology \citep{pcast2016}.
+Potential subject bias in the assessment as well as the lack of specified error rates are the main points of criticisms  first raised by the National Research Council in 2009 \citep{NAS:2009} and later emphasized further by the President's Council of Advisors on Science and Technology \citep{pcast2016}.
 
@@
 pairs $(s_i^{(1)}, s_i^{(2)})$ for all integer values $i$ for which both $s_i^{(1)}$ and  $s_i^{(2)}$ are defined.
-Similarly, different-shift pairs are defined as $(s_i^{(1)}, s_{-i}^{(2)})$ for all $i$ where both $s_i^{(1)}$ and  $s_{-i}^{(2)}$ are defined.
+Similarly, different-shift pairs are defined as $(s_i^{(1)}, s_{-i}^{(2)})$ for all $i$ where both $s_i^{(1)}$ and  $s_{-i}^{(2)}$ are defined (see \autoref{sketch-same-diff}).
+
+\begin{figure}[hbtp]
+\centering
+\includegraphics[width=.7\textwidth]{images/sketch-same.png}
+
+\includegraphics[width=.7\textwidth]{images/sketch-diff.png}
+\caption{\label{sketch-same-diff}Sketch of same-shift pairings  (top) and different-shift pairings (bottom). }
+\end{figure}
 
@@
 
-#### endedit
-
-\hh{end of intro: remaining paper is structured as follows: introduce to the data we get from confocal microscopy, introduce to profiles and signatures.
-Introduce to chumbley score method, apply chumbley and discuss results ....
-}
-
-## Scans for land engraved areas
-
-- scans available: NIST database (citation), Hamby 44 and Hamby 252 (Hamby citation)
-- move figure up
-- discuss cross section, profiles and signatures
--->
-
 <!-- Compairing pairs of toolmarks with the intention of matching it to a tool has been studied many times in the past.
@@
 
-<!-- As mentioned earlier, subject bias and error rate determination have been a long standing issue in firearm examination \citep{NAS:2009}.
-This issue of removing subjectivity and making firearm analysis objective  was also raised in the \citet{pcast2016} report.
-For an objective analysis, scientific principle testability and error rate determination are considered to be defining aspects.
-Identification of the uncertainities associated with a method of comparison and coming up with a systemized method of quantification, is therefore important.
-Associating mathematical probabilites in a systematic manner to these methods are one such way of quantification.
--->
 <!-- In a previous study conducted by \citet{aoas} a machine learning based algorithm was developed for same source matching of bullets and error rates were discussed using the database from the Hamby Study \citep{hamby}.
@@
 
-<!-- ```{r, results='asis', include=FALSE} -->
-<!-- library(knitr) -->
-<!-- library(kableExtra) -->
-<!-- library(xtable) -->
-
-<!-- a1<- matrix(NA, ncol = 3, nrow = 2) -->
-<!-- toolmark1.confusion<- data.frame(a1) -->
-<!-- colnames(toolmark1.confusion)<- c("Classification", "Match", "Non-Match") -->
-<!-- toolmark1.confusion$Classification<- c("Match", "Non-Match") -->
-<!-- # toolmark1.confusion$Match<- c(41,2) -->
-<!-- # toolmark1.confusion$`Non-Match`<- c(9,48) -->
-<!-- toolmark1.confusion$Match<- c(91,2) -->
-<!-- toolmark1.confusion$`Non-Match`<- c(9,98) -->
-
-<!-- a2<- matrix(NA, ncol = 3, nrow = 2) -->
-<!-- toolmark2.confusion<- data.frame(a2) -->
-<!-- colnames(toolmark2.confusion)<- c("Classification", "Match", "Non-Match") -->
-<!-- toolmark2.confusion$Classification<- c("Match", "Non-Match") -->
-<!-- toolmark2.confusion$Match<- c(47,0) -->
-<!-- toolmark2.confusion$`Non-Match`<- c(3,50) -->
-
-<!-- chumbley.confusion<-knitr::kable(toolmark1.confusion, digits=0, format = "latex" ,booktabs=TRUE) #%>% kableExtra::kable_styling(latex_options = c("hold")) -->
-<!-- hadler.confusion<-knitr::kable(toolmark2.confusion, digits=0, format = "latex" ,booktabs=TRUE) #%>% kableExtra::kable_styling(latex_options = c("hold")) -->
-
-<!-- cat(c("\\begin{table}[!htb] -->
-<!--     \\begin{minipage}{.5\\linewidth} -->
-<!--       \\caption{Chumbley et al. 2010} -->
-<!--       \\centering", -->
-<!--         chumbley.confusion, -->
-<!--     "\\end{minipage}% -->
-<!--     \\begin{minipage}{.5\\linewidth} -->
-<!--       \\centering -->
-<!--         \\caption{Hadler et. al. (2017)}", -->
-<!--        hadler.confusion, -->
-<!--     "\\end{minipage}  -->
-<!-- \\end{table}" -->
-<!-- )) -->
-
 <!-- ``` -->
@@
 \citet{chumbley} used an empirical based setup to validate their proposed analytic and quantitative algorithm.
--->
-
-<!-- \citet{chumbley} found that the algorithm gives false-positives in slightly over 2 cases (or 2%) of the time and about 9 false-negatives (or 9%) for every 100 comparisons.
-\citet{hadler} on the other hand, in the improved version of the algorithm first described by \citet{chumbley} found false-positives for 0 cases and 3 false-negatives (or 6%) for 50 comparisons of knwon matches.
--->
-<!-- ```{r} -->
-<!-- a<- matrix(NA, ncol = 3, nrow = 2) -->
-<!-- toolmark.falsepositive<- data.frame(a) -->
-<!-- colnames(toolmark.falsepositive)<- c("Classification", "Match", "Non-Match") -->
-<!-- toolmark.falsepositive$Classification<- c("Match", "Non-Match") -->
-<!-- toolmark.falsepositive$Match<- c(47,0) -->
-<!-- toolmark.falsepositive$`Non-Match`<- c(3,50) -->
-
-<!-- knitr::kable(toolmark.falsepositive, digits=0, format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("hold")) -->
-<!-- ``` -->
-
-<!-- ## Potential limitations of the Chumbley Score adaptation to bullets -->
-<!-- Bullet striations are different from toolmarks.
-Bullet markings are typically produced by the barrel on the bullets, while toolmarks can be made by tool tips on different surfaces.
-Tools like screw driver tips as used by \citet{chumbley}, produce longer and pronounced markings.
-Bullets on the other hand are much smaller in length, width, and curved in the cross-sectional topography.
-This means the markings made by barrels as opposed to markings made by tools, may have a problem in distinctiveness.
-The majority of Bullet profiles and signatures of the Hamby study \citep{hamby} used in this paper are extracted by procedures mentioned by  \citet{aoas}.
-The Ruger pistol barrels used in the  Hamby study have a 9mm caliber.
-This means that bullets used in these pistols have land markings of almost $(1/6)^{th}$ the size of screwdriver toolmarks used by \citet{chumbley}.
-Also, striations on bullets are made on their curved surfaces, whereas the algorithm developed by \citet{chumbley} and \citet{hadler} has only been tested for flatter and wider surfaces which have negligible curvature.
-Therefore, using methods proposed for toolmarks may need adaptation in order to give tangible results for bullets.
-Moreover, in order to get flat bullet signatures and remove the curvatures some kind of smoothing needs to be applied as a pre-step wihch needs further investigation as to whether the level of smoothing does effect the working of the algorithm on Bullets.
--->
-
-<!-- The Chumbley method and algorithm works on the premise of comparison windows which are supposed to represent small segments of pixels of equal sizes in the two markings that are to be compared in a predetermined fashion.
-They term them as windows of optimization and validation.
-The optimization window is selected in the first step called the optimization step where areas/windows of best agreement between the the two markings are found.
-The window of optimization for bullet striations are shorter as bullet signatures are smaller compared to toolmarks .
-The idea is to keep the number of windows of optimization sufficiently large.
-This means we have shorter trace segments that lets us compare smaller segments of one signature to another than those for toolmarks.
-Here, trace segments are the partitions of marking with the length of the partition equal to the size of window of optimization.
-This introduces a problem as, if we go too small in the window of optimization, the unique features of the trace segments are lost and seem similar, while too large sizes vastly reduces the weight of small features that would otherwise uniquely classify a signature and hence identify the region of agreement.
-The problem of distinctiveness in the structure of bullet striations may also lead to potential failure of the algorithm to identify correct windows of agreements or not be able to identify these windows at all.
-This may lead to incorrect classification of bullets as matches or non-matches.
--->
-
-<!-- Thus, the comaparison windows have a direct influence on the false positives and false negatives, making identification of the optimum sizes of these windows very important.
-This raises important questions about how and what parameter settings need to be chosen for bullet land comparison.
-Something similar is done in \citet{afte-chumbley} for toolmark comparisons of slip-joint pliers where optimum window sizes are determined.
-There is a need to figure out the best parameter settings which minimizes the errors for unsmoothed markings or profiles and pre-processed signatures.
-An analysis of these error rates and comparison with other methods will help us understand the adaptibility of the chumbley score to bullets.
 -->
@@
 Initially, the default settings suggested in \citet{hadler} were used: $w_o = 120$ pixels or about 190 $\mu m$ (ten percent of the average length of profiles) and coarseness $c = 0.25$.
-\autoref{fig:fails} shows the percentage of failed tests among the 86k land-to-land comparisons of the NIST data for different values of the validation window size $w_v$.
+\autoref{fig:fails} shows the percentage of failed tests among the 85,491 land-to-land comparisons of the NIST data for different values of the validation window size $w_v$.
 For same-source lands up to 12.5 percent of the tests fail.
@@
 
+Looking closer, we see that the problem lies in the way the different-shift pairs are picked: the number of possible different-shift pairs depends on the location of  the  optimal locations $t_1^o$ and $t_2^o$.
+Whenever are close to the boundaries, the number of possible pairings decreases and reaches zero, if $t_i^{(o)} < w_v$ or  $t_i^{(o)} > T_i-w_o- w_v$.
+For uniformly distributed locations $t_i^{(o)}$, this results in an expected rate of failure of $2 w_v / (T_i-w_o)$, or about 5.6\% for an average length of $T_i$ of 1200 pixels, $w_o = 120$ pixels and $w_v = 30$ pixels.
+
 \hh{diagnose problem: use of optimal location in getting the pairs for different-shift correlations}
 \hh{figure: number of diff shift pairs and location}
+
+\begin{figure}[hbtp]
+\centering
+\includegraphics[width=.7\textwidth]{images/sketch-diff-2.png}
+\caption{\label{sketch-diff-2}Sketch of adjusted different-shift pairings. }
+\end{figure}
 

### 2018-04-18 doc/paper.tex
expected failure rate

@@
 \usepackage{multirow}
+\usepackage{subfig}
 \usepackage[table]{xcolor}
@@
 under comparison, as practically impossible \citep{afte-toolmarks1998}.
-This subjectivity in the assessment as well as the lack of error rates
-are the main points of criticisms first raised by the National Research
-Council in 2009 \citep{NAS:2009} and later emphasized further by the
-President's Council of Advisors on Science and Technology
-\citep{pcast2016}.
+Potential subject bias in the assessment as well as the lack of
+specified error rates are the main points of criticisms first raised by
+the National Research Council in 2009 \citep{NAS:2009} and later
+emphasized further by the President's Council of Advisors on Science and
+Technology \citep{pcast2016}.
 
@@
 different-shift pairs are defined as \((s_i^{(1)}, s_{-i}^{(2)})\) for
-all \(i\) where both \(s_i^{(1)}\) and \(s_{-i}^{(2)}\) are defined.
+all \(i\) where both \(s_i^{(1)}\) and \(s_{-i}^{(2)}\) are defined (see
+\autoref{sketch-same-diff}).
+
+\begin{figure}[hbtp]
+\centering
+\includegraphics[width=.7\textwidth]{images/sketch-same.png}
+
+\includegraphics[width=.7\textwidth]{images/sketch-diff.png}
+\caption{\label{sketch-same-diff}Sketch of same-shift pairings  (top) and different-shift pairings (bottom). }
+\end{figure}
 
@@
 
-For both samples the correlations between the pairs of markings are then
-calculated. The intuition here is that for two markings from the same
-source the correlation for the in-phase sample should be high, while the
-correlations of the out-of-phase sample provide a measure for the
-base-level correlation for non-matching marks of a given length \(w_v\).
-The Chumbley score is then computed as a Mann Whitney U statistic to
-compare between in-phase sample and out-of-phase sample. In the original
-method proposed in \citet{chumbley} both in-phase and out-of-phase
-sample are extracted randomly, whereas \citet{hadler} proposed
-deterministic rules for both samples to make the resulting score
-deterministic while simultaneously avoiding overlaps within selected
-marks to ensure independence.
+For both same- and different-shift pairs correlations between the
+markings are calculated. The intuition here is that for two markings
+from the same source the correlation for the in-phase sample should be
+high, while the correlations of the out-of-phase sample provide a
+measure for the base-level correlation for non-matching marks of a given
+length \(w_v\). The Chumbley score is then computed as a Mann Whitney U
+statistic to compare between in-phase sample and out-of-phase sample. In
+the original method proposed in \citet{chumbley} both in-phase and
+out-of-phase sample are extracted randomly, whereas \citet{hadler}
+proposed the above specified deterministic rules for both samples to
+make the resulting score deterministic while simultaneously avoiding
+overlaps within selected marks to ensure independence.
 
@@
 length of profiles) and coarseness \(c = 0.25\). \autoref{fig:fails}
-shows the percentage of failed tests among the 86k land-to-land
+shows the percentage of failed tests among the 85,491 land-to-land
 comparisons of the NIST data for different values of the validation
@@
 
+Looking closer, we see that the problem lies in the way the
+different-shift pairs are picked: the number of possible different-shift
+pairs depends on the location of the optimal locations \(t_1^o\) and
+\(t_2^o\). Whenever are close to the boundaries, the number of possible
+pairings decreases and reaches zero, if \(t_i^{(o)} < w_v\) or
+\(t_i^{(o)} > T_i-w_o- w_v\). For uniformly distributed locations
+\(t_i^{(o)}\), this results in a probability of failure of
+\(2 w_v / (T_i-w_o)\), or about 5.6\% for an average length of \(T_i\)
+of 1200 pixels, \(w_o = 120\) pixels and \(w_v = 30\) pixels.
+
 \hh{diagnose problem: use of optimal location in getting the pairs for different-shift correlations}
 \hh{figure: number of diff shift pairs and location}
+
+\begin{figure}[hbtp]
+\centering
+\includegraphics[width=.7\textwidth]{images/sketch-diff-2.png}
+\caption{\label{sketch-diff-2}Sketch of adjusted different-shift pairings. }
+\end{figure}
 

### 2018-04-18 doc/paper.Rmd
images

@@
 More precisely, let us define starting points $s_i^{(k)}$ for each signature $k = 1, 2$ as 
-\begin{eqnarray*}
+\begin{eqnarray}\label{eq.start}
 s^{(k)}_i = 
@@
 \end{cases}
-\end{eqnarray*} for integer values of $i$ with $0 <  s^{(k)}_i \le T_k - w_v$.
+\end{eqnarray} for integer values of $i$ with $0 <  s^{(k)}_i \le T_k - w_v$.
 
@@
 In the original method proposed in \citet{chumbley} both in-phase and out-of-phase sample are extracted randomly, whereas \citet{hadler} proposed the above specified deterministic rules for both samples to make the resulting score deterministic while simultaneously avoiding overlaps within selected marks to ensure independence.
+
+## A problem with failures and modification
+
+Looking closer at \autoref{eq.start}, we see that by definition, some number of tests will fail to produce a result, either because the number of eligible same-shift pairs is zero, or the number of different-shift opairs is zero.
+
+The number of same-shift pairs will be zero, if the optimal locations $t_1^{o}$ and $t_2^{o}$ are so far apart, that no segments of size $w_v$ are left on the same sides of the optimal locations, i.e.
+$t_1^{o} < w_v$ and $t_2^{o} > T_2-w_o-w_v$ or  $t_1^{o} < T_1- w_o - w_v$ and $t_2^{o} < w_v$.
+For uniformly distributed optimal locations, we expect to see these situation with a probability of $\frac{2 w_v^2}{(T_1-w_o)(T_2-w_o)}$.
+For  an average length of $T_i$ of 1200 pixels, $w_o = 120$ pixels and $w_v = 30$ pixels this probability is about 0.0015.
+
+The number of possible different-shift pairs also depends on the location of  the  optimal locations $t_1^o$ and $t_2^o$.
+Whenever the optimal locations are close to the boundaries, the number of possible pairings decreases and reaches zero, if $t_i^{(o)} < w_v$ or  $t_i^{(o)} > T_i-w_o- w_v$.
+For uniformly distributed locations $t_i^{(o)}$, this results in an expected rate of failure of $2 w_v / (T_i-w_o)$, or about 5.6\% for an average length of $T_i$ of 1200 pixels, $w_o = 120$ pixels and $w_v = 30$ pixels.
+
+\hh{diagnose problem: use of optimal location in getting the pairs for different-shift correlations}
+\hh{figure: number of diff shift pairs and location}
+
+\begin{figure}[hbtp]
+\centering
+\includegraphics[width=.7\textwidth]{images/sketch-diff-2.png}
+\caption{\label{sketch-diff-2}Sketch of adjusted different-shift pairings. }
+\end{figure}
 
@@
 
-Looking closer, we see that the problem lies in the way the different-shift pairs are picked: the number of possible different-shift pairs depends on the location of  the  optimal locations $t_1^o$ and $t_2^o$.
-Whenever are close to the boundaries, the number of possible pairings decreases and reaches zero, if $t_i^{(o)} < w_v$ or  $t_i^{(o)} > T_i-w_o- w_v$.
-For uniformly distributed locations $t_i^{(o)}$, this results in an expected rate of failure of $2 w_v / (T_i-w_o)$, or about 5.6\% for an average length of $T_i$ of 1200 pixels, $w_o = 120$ pixels and $w_v = 30$ pixels.
-
-\hh{diagnose problem: use of optimal location in getting the pairs for different-shift correlations}
-\hh{figure: number of diff shift pairs and location}
-
-\begin{figure}[hbtp]
-\centering
-\includegraphics[width=.7\textwidth]{images/sketch-diff-2.png}
-\caption{\label{sketch-diff-2}Sketch of adjusted different-shift pairings. }
-\end{figure}
-
 # A modified test

### 2018-04-18 doc/paper.tex
images

@@
 
-\begin{eqnarray*}
+\begin{eqnarray}\label{eq.start}
 s^{(k)}_i = 
@@
 \end{cases}
-\end{eqnarray*}
+\end{eqnarray}
 
@@
 overlaps within selected marks to ensure independence.
+
+\subsection{A problem with failures and
+modification}\label{a-problem-with-failures-and-modification}
+
+Looking closer at \autoref{eq.start}, we see that by definition, some
+number of tests will fail to produce a result, either because the number
+of eligible same-shift pairs is zero, or the number of different-shift
+opairs is zero. The number of same-shift pairs will be zero, if the
+optimal locations \(t_1^{o}\) and \(t_2^{o}\) are so far apart, that no
+segments of size \(w_v\) are left on the same sides of the optimal
+locations, i.e. \(t_1^{o} < w_v\) and \(t_2^{o} > T_2-w_o-w_v\) or
+\(t_1^{o} < T_1- w_o - w_v\) and \(t_2^{o} < w_v\). For uniformly
+distributed optimal locations, we expect to see these situation in
+\(2 \times \frac{w_v^2}{(T_1-w_o)(T_2-w_o)}\)
+
+problem lies in the way the different-shift pairs are picked: the number
+of possible different-shift pairs depends on the location of the optimal
+locations \(t_1^o\) and \(t_2^o\). Whenever are close to the boundaries,
+the number of possible pairings decreases and reaches zero, if
+\(t_i^{(o)} < w_v\) or \(t_i^{(o)} > T_i-w_o- w_v\). For uniformly
+distributed locations \(t_i^{(o)}\), this results in an expected rate of
+failure of \(2 w_v / (T_i-w_o)\), or about 5.6\% for an average length
+of \(T_i\) of 1200 pixels, \(w_o = 120\) pixels and \(w_v = 30\) pixels.
+
+\hh{diagnose problem: use of optimal location in getting the pairs for different-shift correlations}
+\hh{figure: number of diff shift pairs and location}
+
+\begin{figure}[hbtp]
+\centering
+\includegraphics[width=.7\textwidth]{images/sketch-diff-2.png}
+\caption{\label{sketch-diff-2}Sketch of adjusted different-shift pairings. }
+\end{figure}
 
@@
 The percentage of failures is higher for same-source lands than for different source lands.}\label{fig:fails}
-\end{figure}
-
-Looking closer, we see that the problem lies in the way the
-different-shift pairs are picked: the number of possible different-shift
-pairs depends on the location of the optimal locations \(t_1^o\) and
-\(t_2^o\). Whenever are close to the boundaries, the number of possible
-pairings decreases and reaches zero, if \(t_i^{(o)} < w_v\) or
-\(t_i^{(o)} > T_i-w_o- w_v\). For uniformly distributed locations
-\(t_i^{(o)}\), this results in a probability of failure of
-\(2 w_v / (T_i-w_o)\), or about 5.6\% for an average length of \(T_i\)
-of 1200 pixels, \(w_o = 120\) pixels and \(w_v = 30\) pixels.
-
-\hh{diagnose problem: use of optimal location in getting the pairs for different-shift correlations}
-\hh{figure: number of diff shift pairs and location}
-
-\begin{figure}[hbtp]
-\centering
-\includegraphics[width=.7\textwidth]{images/sketch-diff-2.png}
-\caption{\label{sketch-diff-2}Sketch of adjusted different-shift pairings. }
 \end{figure}

### 2018-04-18 doc/paper.Rmd
alternative different shifts

@@
 pairs $(s_i^{(1)}, s_i^{(2)})$ for all integer values $i$ for which both $s_i^{(1)}$ and  $s_i^{(2)}$ are defined.
-Similarly, different-shift pairs are defined as $(s_i^{(1)}, s_{-i}^{(2)})$ for all $i$ where both $s_i^{(1)}$ and  $s_{-i}^{(2)}$ are defined (see \autoref{sketch-same-diff}).
+Similarly, different-shift pairs are defined as $(s_i^{(1)}, s_{-i-1}^{(2)})$ for all $i$ where both $s_i^{(1)}$ and  $s_{-i-1}^{(2)}$ are defined (see \autoref{sketch-same-diff}).
 
@@
 \includegraphics[width=.7\textwidth]{images/sketch-diff.png}
-\caption{\label{sketch-same-diff}Sketch of same-shift pairings  (top) and different-shift pairings (bottom). }
+\caption{\label{sketch-same-diff}Sketch of same-shift pairings  (top) and different-shift pairings (bottom).
+Filled in rectangles show pairings resulting in correlations, unfilled rectangles are segments without a match.}
 \end{figure}
@@
 $t_1^{o} < w_v$ and $t_2^{o} > T_2-w_o-w_v$ or  $t_1^{o} < T_1- w_o - w_v$ and $t_2^{o} < w_v$.
-For uniformly distributed optimal locations, we expect to see these situation with a probability of $\frac{2 w_v^2}{(T_1-w_o)(T_2-w_o)}$.
+While we can assume that once (sub-)class characteristics are removed, optimal locations $t_i^{o}$ are uniformly distributed across the length of the profile, we cannot assume that $t_1^o$ and $t_2^o$ are independent of each other.
+In particular, for same-source profiles, we would expect a strong dependency between these locations.
+However, for different source matches, an argument of independence in the locations could be made.
+In that case, we expect a test to fail with a probability of $\frac{2 w_v^2}{(T_1-w_o)(T_2-w_o)}$.
 For an average length of $T_i$ of 1200 pixels, $w_o = 120$ pixels and $w_v = 30$ pixels this probability is about 0.0015.
@@
 Whenever the optimal locations are close to the boundaries, the number of possible pairings decreases and reaches zero, if $t_i^{(o)} < w_v$ or  $t_i^{(o)} > T_i-w_o- w_v$.
-For uniformly distributed locations $t_i^{(o)}$, this results in an expected rate of failure of $2 w_v / (T_i-w_o)$, or about 5.6\% for an average length of $T_i$ of 1200 pixels, $w_o = 120$ pixels and $w_v = 30$ pixels.
-
-\hh{diagnose problem: use of optimal location in getting the pairs for different-shift correlations}
-\hh{figure: number of diff shift pairs and location}
+Assuming a correlation between optimal locations $t_1^o$ and $t_2^o$ of close to one for same-source profiles, this results in an expected rate of failure of $2 w_v / (T_i-w_o)$, or about 5.6\% for an average length of $T_i$ of 1200 pixels, $w_o = 120$ pixels and $w_v = 30$ pixels.
+Assuming independence in the optimal locations for different source profiles the expected probability for a failed test is, again, $\frac{2 w_v^2}{(T_1-w_o)(T_2-w_o)}$.
+
+While failures due to missing correlations from same-shift pairs are unavoidable by definition of the Chumbley score, failures due to missing correlations from different-shift pairs can be prevented by using a different strategy in assigning pairs.
+
+Using the same notation as in \autoref{eq.start}, we define same-shift pairs identical to \citet{hadler} as pairs $(s_i^{(1)}, s_i^{(2)})$ for all $i$ where the boundary conditions of both sequences are met simultaneously.
+Let us assume that this results in $I$ pairs.
+Define $s_(j)^{(k)}$ to be the $j$th starting location in sequence $k = 1, 2$, i.e.\ $s_(1)^{(k)} < s_(2)^{(k)} < ...
+< s_(I)^{(k)}$.
+
+We then define the pairs for different-shifts as
+\begin{equation}\label{eq.diff2}
+\left(s_{(j)}^{(1)}, s_{(I-j+1)}^{(1)} \right) \text{ for } j = 
+\begin{cases}
+1, ..., I & \text{ for even } I \\
+1, ..., (I-1)/2, (I-1)/2 + 2, ..., I & \text{ for odd } I
+\end{cases},
+\end{equation}
+i.e.\ for an odd number of same-shift correlations, we skip the middle pair for the different-shift correlations (see \autoref{sketch-diff-2}).
+This pairing ensures that the number of different-shift pairings is the same or at most one less than the number of same-shift pairings in all tests.
 
@@
 \includegraphics[width=.7\textwidth]{images/sketch-diff-2.png}
-\caption{\label{sketch-diff-2}Sketch of adjusted different-shift pairings. }
+\caption{\label{sketch-diff-2}Sketch of adjusted different-shift pairings. At most one of the same-shift pairings can not be matched with a different-shift pair. }
 \end{figure}

### 2018-04-18 doc/paper.tex
alternative different shifts

--- 2018-04-18 doc/paper.tex
@@
 both \(s_i^{(1)}\) and \(s_i^{(2)}\) are defined. Similarly,
-different-shift pairs are defined as \((s_i^{(1)}, s_{-i}^{(2)})\) for
-all \(i\) where both \(s_i^{(1)}\) and \(s_{-i}^{(2)}\) are defined (see
-\autoref{sketch-same-diff}).
+different-shift pairs are defined as \((s_i^{(1)}, s_{-i-1}^{(2)})\) for
+all \(i\) where both \(s_i^{(1)}\) and \(s_{-i-1}^{(2)}\) are defined
+(see \autoref{sketch-same-diff}).
 
@@
 overlaps within selected marks to ensure independence.
+
+\subsection{A problem with failures and
+modification}\label{a-problem-with-failures-and-modification}
+
+Looking closer at \autoref{eq.start}, we see that by definition, some
+number of tests will fail to produce a result, either because the number
+of eligible same-shift pairs is zero, or the number of different-shift
+opairs is zero.
+
+The number of same-shift pairs will be zero, if the optimal locations
+\(t_1^{o}\) and \(t_2^{o}\) are so far apart, that no segments of size
+\(w_v\) are left on the same sides of the optimal locations, i.e.
+\(t_1^{o} < w_v\) and \(t_2^{o} > T_2-w_o-w_v\) or
+\(t_1^{o} < T_1- w_o - w_v\) and \(t_2^{o} < w_v\). While we can assume
+that once (sub-)class characteristics are removed, optimal locations
+\(t_i^{o}\) are uniformly distributed across the length of the profile,
+we cannot assume that \(t_1^o\) and \(t_2^o\) are independent of each
+other. In particular, for same-source profiles, we would expect a strong
+dependency between these locations. However, for different source
+matches, an argument of independence in the locations could be made. In
+that case, we expect a test to fail with a probability of
+\(\frac{2 w_v^2}{(T_1-w_o)(T_2-w_o)}\). For an average length of \(T_i\)
+of 1200 pixels, \(w_o = 120\) pixels and \(w_v = 30\) pixels this
+probability is about 0.0015.
+
+The number of possible different-shift pairs also depends on the
+location of the optimal locations \(t_1^o\) and \(t_2^o\). Whenever the
+optimal locations are close to the boundaries, the number of possible
+pairings decreases and reaches zero, if \(t_i^{(o)} < w_v\) or
+\(t_i^{(o)} > T_i-w_o- w_v\). Assuming a correlation between optimal
+locations \(t_1^o\) and \(t_2^o\) of close to one for same-source
+profiles, this results in an expected rate of failure of
+\(2 w_v / (T_i-w_o)\), or about 5.6\% for an average length of \(T_i\)
+of 1200 pixels, \(w_o = 120\) pixels and \(w_v = 30\) pixels. Assuming
+independence in the optimal locations for different source profiles the
+expected probability for a failed test is, again,
+\(\frac{2 w_v^2}{(T_1-w_o)(T_2-w_o)}\).
+
+While failures due to missing correlations from same-shift pairs are
+unavoidable by definition of the Chumbley score, failures due to missing
+correlations from different-shift pairs can be prevented by using a
+different strategy in assigning pairs.
+
+Using the same notation as in \autoref{eq.start}, we define same-shift
+pairs identical to \citet{hadler} as pairs \((s_i^{(1)}, s_i^{(2)})\)
+for all \(i\) where the boundary conditions of both sequences are met
+simultaneously. Let us assume that this results in \(I\) pairs. Define
+\(s_(j)^{(k)}\) to be the \(j\)th starting location in sequence
+\(k = 1, 2\), i.e.~\(s_(1)^{(k)} < s_(2)^{(k)} < ... < s_(I)^{(k)}\).
+
+We then define the pairs for different-shifts as
+
+\begin{equation}\label{eq.diff2}
+\left(s_{(j)}^{(1)}, s_{(I-j+1)}^{(1)} \right) \text{ for } j = 
+\begin{cases}
+1, ..., I & \text{ for even } I \\
+1, ..., (I-1)/2, (I-1)/2 + 2, ..., I & \text{ for odd } I
+\end{cases},
+\end{equation}
+
+i.e.~for an odd number of same-shift correlations, we skip the middle
+pair for the different-shift correlations (see \autoref{sketch-diff-2}).
+This pairing ensures that the number of different-shift pairings is the
+same or at most one less than the number of same-shift pairings in all
+tests.
+
+\begin{figure}[hbtp]
+\centering
+\includegraphics[width=.7\textwidth]{images/sketch-diff-2.png}
+\caption{\label{sketch-diff-2}Sketch of adjusted different-shift pairings. At most one of the same-shift pairings can not be matched with a different-shift pair. }
+\end{figure}
 

### 2018-04-18 doc/paper.Rmd
setup for both methods

@@
 
-<!--
-The similarity in the two markings is then judged by the algorithm on the basis of cross-correlation of a fixed and constant number consecutive pixels (say $k$) taken from the two markings.
-This can for example be $k$ taken from the two indexed marking $x(t_1)$ and $y(t_2)$ such that in theory $k$ remains smaller than length of the two striae or marks.
-Depending on what stage of the algorithm we are in, matching of different pixel lengths and locations is done.
-This in the end, effectively compares all possible windows that would guarantee in quantifying the two marks or striae as coming from the same source or not.
--->
-
-<!-- The word 'pixel' refers to the resolution of the confocal light microscope.
-In the case for bullet signatures extracted from NIST ballistics database \citep{nist} of the Hamby study, a pixel corresponds to approximately 1.56 microns.
--->
-
-<!--
-The algorithm works in two phases, namely, an optimization step and a validation step, at the end of which a Mann Whitney U statistic is calculated as a measure to differentiate between matches and non-matches.
-A pre-processing step to the algorithm is to choose a coarseness value which  is used as a parameter to the lowess smoothing function.
-The coarseness essentially gives the proportion of points which influence the smooth at each value, which means larger values lead to more smoothness.
-The lowess smoothing is applied to each of two sets of vectorized striae or marks $x(t_1)$ and $y(t_2)$, before proceeding to the algorithm.-->
-<!-- The algorithm starts with the optimization step where the area of best agreement in the two markings being compared is identified.
-Comparison window size is predefined.
-Each window of the first marking is compared with all windows of the second marking.
-A maximum correlation statistic is used to identify the region of best agreement, with the maximum usually seen being near 1 for both cases which is what is intuitively expected for matches, and not expected intuitively for non matches.
-\citet{hadler} in their paper proposed an improvement to this algorithm by trying to remove mutual dependence of parameters.
-The mututal dependence was due to the serial correlation in surface depth values of a marking.
-Also a random sampling sub step in the validation phase makes a group of pixels to be chosen more than once and hence introduces lack of independence in certain steps.
-The reason for removing the mutual dependence was the Mann-Whitney U statistic works under the assumption of independence of parameters.-->
-<!--
-The validation step builds on this with two sub-steps namely, same shift and different shift.
-As the aim is to come up with a non-parameteric U statistic, \citet{hadler} proposed a normalization procedure in the validation step which goes to some extent to address the issue of mutual dependence.
-A series of windows are chosen in both the same and different shift sub-steps for the purpose of comparison.
-Both substeps were modified by \citet{hadler} who introduced a deterministic rule for sampling of windows in the same-shift and different shift as opposed to the originally proposed random sampling by \citet{chumbley}.
-In th same shift substep the chosen series of windows are at a common distance (rigid-shifts) from the window earlier identified as the region of best agreement in the optimization step.
-The correlation of these windows generally turn out to be lower than the maximum correlation window.
-The significance is that these same shift windows still have large enough correlation values for the two markings being compared that are in reality a match.
-Any choice of maximum correlation windows (where correlation values are often near 1) for non-matches in the optimization step are also validated by this sub-step.
-This is because the same-shift correlations would not be anywhere as large for same-shift windows when all windows in the sub-step are compared.
-The different shift substep on the other hand gives perspective to the correlation values of the same shift window correlation values.
-There are no rigid-shifts but different shifts where distance from the maximum correlation window are chosen randomly by \citet{chumbley} and deterministically as per \citet{hadler}.
-This means that there is an equal possibilty of comparing a trace segment from one marking to any trace segment or window in the second marking.
-
-Neither of above sets of correlation in the two sub-steps are allowed to include the maximum correlation window as identified earlier.
-Therefore the assumption is that if two markings match each other, the same-shift correlations would be larger than the different-shift windows.
-And if they are not a match the correlations in the two sets will be very similar.
-The U-statistic tests for the null hypothesis that the two markings are not a match and are therefore not made by the same source.
-As given by \citet{hadler}, it is computed from the joint rank of all correlations of both the same and different shift samples.
--->
-<!--
-##################
-
-<!-- Compairing pairs of toolmarks with the intention of matching it to a tool has been studied many times in the past.
-Extensive examples can be found in literature for tools and toolmark research ranging from screwdrivers \citep{manytoolmarks1, manytoolmarks2, chumbley} to groove pliers \citep{manytoolmarks2} to slip-joint pliers \citep{afte-chumbley} and many more.
-In comparison to this, same source matching of bullets to firearms has not been examined as prominently as that of toolmarks.
-Even less information is available on validity of methods and error rates associated with firearms examination.
-The National Academy of Sciences in its report in 2009 \citep{NAS:2009} discussed the need for determining error rates in methods proposed for firearms examination.
-As seen in the case of most forensic applications, the first step for same source matching and error rate determination involves identification of unique features that are characteristic of the object at hand.
-For the case of bullets and firearms, striation marks on the surface of the bullet are considered to be such markings that can be used in methods for same source matching.
-These marks are often a product of rifling and impurities and defects due to manufacturing in the barrel of the gun, which leads to engravings on the bullet surface \citep{afte-article1992}.
-In current practice, firearm examiners invariably make visual comparisons of bullet striae and use visual assesment tools to dignify bullets as being matches and non-matches.
-One way of accomplishing anykind of comparison between bullets is to do a comparison between surface marking of two or more bullet lands.
-Bullet Lands are considered to be areas between grooves made by the rifling action of the barrel and the markings on them are considered to be unique.
-The land engraved markings or sometimes termed as Bullet profiles \citep{aoas,ma2004} are striation marks made on Bullet lands and often used for these land to land comparisons.
-Bullet Signatures is another word used in literature as seen in the work of \citet{chu2013} and \citet{aoas}.
-In our context bullet signatures refer to a processed version of the raw land engraved markings or profiles.
-The generation of bullet signatures involves first extraction of a bullet profile by taking the cross-sectional of the surface at a given height and then using loess fits to model the structure.
-The residuals of this fit are called signatures, which are considered to be noise free and a good reflection of the class charecteristics and unique features of a bullet.
-A more detailed version of the extraction technique of signatures is discussed by \citet{aoas}, where comprehensive details about the height at which profile is to be selected, removing curvature, smoothing, identifying groove locations are explained.
-Figure \ref{fig:rgl} shows us how the extracted signature from a bullet land lines up with the image of the land from which it was extracted.
-We can see also see in the figure how the depth and relative position of the striation markings seen in the image are interpreted as the signature.
--->
-
-<!-- In a previous study conducted by \citet{aoas} a machine learning based algorithm was developed for same source matching of bullets and error rates were discussed using the database from the Hamby Study \citep{hamby}.
-The method proposed by \citet{chumbley} (and later improved by \citet{hadler}) also provides a means to determine error rates and claims to reduce subject bias.
-Therefore, giving us a strong motivation to explore the adaptibility of the Chumbley score methodology to bullets.
-In this paper, we take this toolmark based methodology and try to apply and adapt it to bullets.
-After this, we consequently discuss about the efforts in doing so, along with the associated error rates.
-The data used in this paper also belongs to the Hamby Study \citep{hamby}.
-This gives us a common platform for comparing the performance of the chumbley method on bullets with the already existing method proposed by \citet{aoas} for bullets.
-As a brief overview, \citet{chumbley}, in their paper, compare two toolmarks with the intention of determining if it comes from the same tool.
-They use an empirical based setup to validate their proposed algorithm and quantitative method which calculates a U-statisic for the purpose of classification of toolmarks as matching or non-matching.
-The data for their study was obtained from 50 sequentially manufactured screwdriver tips, and preselected comparison window sizes were given as inputs to the algorithm.
-The algorithm then compares the two toolmarks and comes up with a Mann-Whitney U-statistic and an associated p-value to designate them as matches or non-matches.
-The performance for every 100 comparisons, of the algorithm proposed by \citet{chumbley} and the improvement proposed by \citet{hadler} are listed in the table below.
--->
-
-<!-- ``` -->
-<!-- Striation marks are used to determine same source for bullets.
-Current standards ask for a visual inspection of the marks under a comparison microscope by a firearms examiner.
-One of the problems raised by the NAS report in 2009 is the subjectivity involved in this comparison and the need for determining error rates \citet{NAS:2009}.
-The issue of subjectivity of a firearms examiner has been reviewed by many authors where things like scientific principle testability and error rates are considered to be defining aspects for an objective analysis such that without identifying the uncertainities associated with a method of comparison, it is inconclusive to regards an analysis complete.
-This means that with the usual method of visual comparison that is adopted by firearms examiners, lies the problem of lack of any systemized quantification and as such there seems no certain way of associating it with any kind of mathematical probability.
-Since determining error rates has been duly noted as a fundamental problem in forensic science \citet{NAS:2009}, there is a need of methods that address this problem.
-Some of the methods that have been used to estimate error rates in firearm examinations include Automatic cartridge case comparison where same source identification of cartridge cases and associated error rates have been provided by \citet{riva}, although as noted by \citet{aoas}, in this case alignments of striae involves roation of planes, which cannot be generalized for bullets.
-In case of methods that use machine learning algorithms some methods which are bootsrap based methods often tend to give over-estimated error rates \citet{efron}, but there are alternate error estimation techniques as described by \citet{aoas}, \citet{efron} and  \citet{vorburger2016} which give better results.
--->
-
-<!-- In case of methods that do not use machine learning algorithm, matching two striation marks with each other in order to identify if it is from the same source, can also be done using a well defined comparative statistical algorithm.
-The statistical algorithm is therefore expected  to go through a step by step procedure of identifying the class characteristics i.e striation marks for bullets, choosing the striae, and then follow a systematic procedure of comparing two striae with each other on the basis of criteria like cross-correlation (or others).
-Such a methodology would make it possible to determine error rates too and give definite results regarding what proportion of cases would the analysis be successful in identifying a match correctly and in how many cases would it end up being a false positive.
--->
-
-<!-- Identification criteria like consecutively matching striae (CMS) as first seen by Biasotti 1959 \citet{biasotti} and later mentioned in \citet{chu2013} more often than not include subject bias and are error prone and as seen in the work of \citet{miller} the number of CMS may turn out to be high even when the bullet is not fired by the same firearm.
-Parameters like cross-correlation factor as seen in the extensive comparisons made by \citep{aoas} seem to perform much better for matching purposes.
--->
-
-<!-- An important aspect of same source matching in firearms is identification of bullet microtopographies that are unique to be chosen for a match, Bullet profiles and signatures prove to be such features.
-The choice of a bullet signature to uniquely define a bullet depends on finding first, a stable region on a Bullet land which minimal noise and many pronounced striation marks.
-\citet{aoas} -->
-
-<!-- The method employed by \citet{aoas} uses the cross correlation factor as a means to identify the stable region.
-The markings or striae in this region are supposed to have a high CCF with each other.
-Therefore a Bullet profile is chosen from this region by taking the cross section at a given height.
-A loess fit on the bullet profile produces  residuals which are termed as signatures and can be used as a unique identifier that can be used while matching two bullets to a firearm.
--->
-
-<!-- Compairing pairs of toolmarks on the otherhand, with the intention of matching it to a tool has been studied relatively more in the past as compared to bullets, and \citet{chumbley} have described in their paper an algorithm and analytic method that compares two toolmarks and come to the conclusion if they are from the same tool or not.
-The method also determines the error rates, reduces subject bias and designate the two toolmarks as matches or non-matches with respect to a source.
-\citet{chumbley} used an empirical based setup to validate their proposed analytic and quantitative algorithm.
--->
-
-<!-- ## The Chumbley Score Test -->
-
-<!-- The Chumbley score algorithm takes input as two vectorized processes.
-Here the processes are of the form $z(t)$ which is a spatial process for some location indexed with $t$.
-This means that while $z(t)$ represents any striae, $z(t_1)$ is a realization of the spatial process.
-$t$ here denotes a vector of equally spaced pixel locations for the striation marks under consideration.
-The word 'pixel' refers to the resolution of the confocal light microscope.
-In the case for bullet signatures extracted from NIST ballistics database \citep{nist} of the Hamby study, a pixel corresponds to approximately 1.56 microns.
-The vectorized processes representing two sets of striaes are therefore shown as $x(t_1)$, $t_1 = 1,2,...T_1$ and $y(t_2)$, $t_2 = 1,2...T_2$.
-Here $x$ and $y$ denote two instances of the process $z(*)$ which means two striation marks, while the indexing $t$ or pixel range is shown as $t_1$ and $t_2$ for $x$ and $y$ respectively.
-This representation of $t$ as $t_1$ and $t_2$ lets us identify them as either of same or different lengths.
-The striation marks under consideration are potentially from two different bullets whose source needs to be identified as being same or different.
-$T_1$ and $T_2$, as represented above, are the final pixel indexes of each marking and therefore give the respective lengths of the markings.
-The similarity in the two markings is then judged by the algorithm on the basis of cross-correlation of a fixed and constant number consecutive pixels (say $k$) taken from the two markings.
-This can for example be $k$ taken from the two indexed marking $x(t_1)$ and $y(t_2)$ such that in theory $k$ remains smaller than length of the two striae or marks.
-Depending on what stage of the algorithm we are in, matching of different pixel lengths and locations is done.
-This in the end, effectively compares all possible windows that would guarantee in quantifying the two marks or striae as coming from the same source or not.
--->
-
-<!-- The algorithm works in two phases, namely, an optimization step and a validation step, at the end of which a Mann Whitney U statistic is calculated as a measure to differentiate between matches and non-matches.
-A pre-processing step to the algorithm is to choose a coarseness value which  is used as a parameter to the lowess smoothing function.
-The coarseness essentially gives the proportion of points which influence the smooth at each value, which means larger values lead to more smoothness.
-The lowess smoothing is applied to each of two sets of vectorized striae or marks $x(t_1)$ and $y(t_2)$, before proceeding to the algorithm.
--->
-<!-- The algorithm starts with the optimization step where the area of best agreement in the two markings being compared is identified.
-Comparison window size is predefined.
-Each window of the first marking is compared with all windows of the second marking.
-A maximum correlation statistic is used to identify the region of best agreement, with the maximum usually seen being near 1 for both cases which is what is intuitively expected for matches, and not expected intuitively for non matches.
--->
-<!-- \citet{hadler} in their paper proposed an improvement to this algorithm by trying to remove mutual dependence of parameters.
-The mututal dependence was due to the serial correlation in surface depth values of a marking.
-Also a random sampling sub step in the validation phase makes a group of pixels to be chosen more than once and hence introduces lack of independence in certain steps.
-The reason for removing the mutual dependence was the Mann-Whitney U statistic works under the assumption of independence of parameters.
--->
-<!-- The validation step builds on this with two sub-steps namely, same shift and different shift.
-As the aim is to come up with a non-parameteric U statistic, \citet{hadler} proposed a normalization procedure in the validation step which goes to some extent to address the issue of mutual dependence.
-A series of windows are chosen in both the same and different shift sub-steps for the purpose of comparison.
-Both substeps were modified by \citet{hadler} who introduced a deterministic rule for sampling of windows in the same-shift and different shift as opposed to the originally proposed random sampling by \citet{chumbley}.
--->
-<!-- In th same shift substep the chosen series of windows are at a common distance (rigid-shifts) from the window earlier identified as the region of best agreement in the optimization step.
-The correlation of these windows generally turn out to be lower than the maximum correlation window.
-The significance is that these same shift windows still have large enough correlation values for the two markings being compared that are in reality a match.
-Any choice of maximum correlation windows (where correlation values are often near 1) for non-matches in the optimization step are also validated by this sub-step.
-This is because the same-shift correlations would not be anywhere as large for same-shift windows when all windows in the sub-step are compared.
--->
-<!-- The different shift substep on the other hand gives perspective to the correlation values of the same shift window correlation values.
-There are no rigid-shifts but different shifts where distance from the maximum correlation window are chosen randomly by \citet{chumbley} and deterministically as per \citet{hadler}.
-This means that there is an equal possibilty of comparing a trace segment from one marking to any trace segment or window in the second marking.
--->
-
-<!-- Neither of above sets of correlation in the two sub-steps are allowed to include the maximum correlation window as identified earlier.
-Therefore the assumption is that if two markings match each other, the same-shift correlations would be larger than the different-shift windows.
--->
-<!-- And if they are not a match the correlations in the two sets will be very similar.
-The U-statistic tests for the null hypothesis that the two markings are not a match and are therefore not made by the same source.
-As given by \citet{hadler}, it is computed from the joint rank of all correlations of both the same and different shift samples.
--->
-
-<!-- The data that is used by \citet{chumbley} is generated by a surface profilometer that gives the height in terms of distance along a linear trace.
-This is taken perpendicular to the striations present in the toolmark.
-Two such trace are then compared using the algorithm.
--->
-
-<!-- ## Detailed algorithm -->
-
-<!-- ###Optimization step -->
-<!-- The idea behind this step is to first identify the area of best agreement in the two toolmark data.
-Comparison window size is predefined by the user, which in case of screwdriver toolmarks was chosen by Chumbley et al and Hadler et al as around 10 percent of the length of the toolmarks, which was around 500.
-This window is henceforth referred to as Window of optimization.
--->
-
-<!-- A maximum correlation statistic is used to identify the region of best agreement, with the maximum usually seen being near 1 for both cases which is what we intuitively expect for matches, and something which we do not intuitively expect for non matches  -->
-
-<!-- First, there are a very large number of cross-correlations calculated for the two series of striae or marks $x(t_1)$ and $y(t_2)$, for eg if the window of optimization was defined as 200 and toolmark pixel length is around 1000 then we have 1200-(200-1) = 1001.
--->
-
-<!-- This is the number of windows we have for one toolmark and each window is compared with all windows of the second toolmark (the number of windows is again similar), and the window with the maximum correlation is identified to be the region where the toolmarks are in maximum agreement with each other -->
-
-<!-- ###Validation Step -->
-
-<!-- ####Same-shift:  -->
-<!-- In this step a series of windows are chosen at random (originally by \citet{chumbley}) and deterministically (by \citet{hadler}), but at a common distance (rigid-shift) from the window identified as the region of best agreement in the vaildation step.
--->
-<!-- The correlation of these windows would be as intuitively assumed i.e.
-lower than the maximum correlation window (Optimization step), but the significance is that these same shift windows will still have large enough correlation values for two toolmarks or signatures that are in reality a match.
--->
-
-<!-- This also validates that if in the optimization step the maximum correlation window chosen (with correlation value near 1) was by accident (like in case of signatures that in reality are a non-match), all these same- shift correlations (A fixed number of these trace segments are identified) would not be anywhere large enough for all same-shift windows.
--->
-
-<!-- ####Different Shift: -->
-<!-- Primary reason for this substep is to give perspective to the correlation values of the same shift window correlation values. -->
-
-<!-- This time there are no rigid-shifts but different shifts (distance from Window of Opt with max correlation) chosen randomly by \citet{chumbley} and deterministically by Hadler et al.
-such that there is an equal possibilty of comparing a trace segment from one signature or toolmark to any one in the second signature or toolamrk.
--->
-
-<!-- Neither of above sets of correlation are allowed to include the maximum correlation window as identified earlier. -->
-
-<!-- Therefore the assumption is that if two toolmarks or signatures match each other the same-shift correlations would be larger than the different-shift windows, -->
-<!-- and if they are not a match the correlations in the two sets will be very similar. -->
-
-<!-- ###U Statistic: -->
-
-<!-- This is computed from the joint rank of all correlations of both the same and different shift samples. As given by \citet{hadler} -->
-
-<!-- Null Hypothesis: If the toolmarks were not match i.e not made by the same tool. -->
-
-<!-- Let ns and nd be the number of same shift and different shift windows -->
-<!-- $$N = n_{s} + n_{d}$$ -->
-
-<!-- The mann whitney U statistic is given by -->
-<!-- $$U =\sum^{ns}_{i=1}R_{s}\left( i\right)$$ -->
-
-<!-- with the standardized version which includes provision for rank ties -->
-
-<!-- $$\overline{U}= \dfrac{U-M}{\sqrt{V}}$$ -->
-
-<!-- where prior to normalization the U-statistic has the mean as -->
-
-<!-- $$M = n_{s}\left(\dfrac{N+1}{2}\right)$$ -->
-
-<!-- and variance  -->
-
-<!-- $$V = \dfrac{n_{s}n_{d}}{N\left(N-1\right)}\left[\Sigma^{n_{s}}R_{s}\left(i\right)^{2}+\Sigma^{n_{d}}R_{d}\left(j\right)^{2}\right] -\dfrac{n_{s} n_{d}\left(N+1\right)^{2}}{4\left(N-1\right)}$$ -->
-
 # Testing setup
@@
 ## The Data
-
-\hh{Introduce profiles and signatures, as shown in figure \ref{fig:sigs-profiles}.}
 
@@
 
-\hh{Discuss CSAFE scans}
-
-Profiles and signatures were extracted from all scans as described in \citet{aoas}.
+Profiles for each bullet land were extracted from  scans as described in \citet{aoas}.
 
@@
 
+\hh{XXX also calculate results for the adjusted adjusted score XXX}
 <!--The settings for optimizing and validating window sizes, $w_o$ and $w_v$, ranged from $w_o \in [50, 280]$ and $w_v \in \{30, 50\}$, see also figure \ref{fig:type2}.-->

### 2018-04-20 doc/paper.Rmd
results from cs2

@@
                 errorrate(allp, 0.05))
-errors2$type <- "profiles"
+errors2$type <- "CS1"
+
+if (!file.exists("../data/invprofiles.rds")) {
+  files <- dir("../data/inverse-profiles/", pattern="rds")
+  allinv <- data.frame()
+  for (file in files) {
+    tmp <- readRDS(file=file.path(path="../data/inverse-profiles", file)) 
+    allinv <- rbind(allinv, tmp)
+  }
+  allinv <- allinv %>% filter(land1_id < land2_id)
+  allinv$coarse <- allinv$coarseness
+  allinv <- allinv %>% select(-coarseness)
+  saveRDS(allinv, file="../data/invprofiles.rds")
+} else {
+  allinv <- readRDS("../data/invprofiles.rds")
+}
+
+errors3 <- rbind(errorrate(allinv, 0.001), 
+                errorrate(allinv, 0.005), 
+                errorrate(allinv, 0.01), 
+                errorrate(allinv, 0.05))
+errors3$type <- "CS2"
 
@@
 Let us assume that this results in $I$ pairs.
-Define $s_(j)^{(k)}$ to be the $j$th starting location in sequence $k = 1, 2$, i.e.\ $s_(1)^{(k)} < s_(2)^{(k)} < ...
-< s_(I)^{(k)}$.
+Define $s_{(j)}^{(k)}$ to be the $j$th starting location in sequence $k = 1, 2$, i.e.\ $s_{(1)}^{(k)} < s_{(2)}^{(k)} < ...
+< s_{(I)}^{(k)}$.
 
@@
 This pairing ensures that the number of different-shift pairings is the same or at most one less than the number of same-shift pairings in all tests.
-
+\hh{In the remainder of the paper, we will refer to the algorithm defined by \citet{hadler} as {\bf (CS1)} and the suggested modified algorithm as {\bf (CS2)} and compare their performance on the available scans of the Hamby study.}
 \begin{figure}[hbtp]
@@
 
-Profiles for each bullet land were extracted from  scans as described in \citet{aoas}.
+Profiles for each bullet land were extracted from scans close to the heel of the bullet while avoiding break-off as described in \citet{aoas}.
 
 ## Setup
-\hh{XXX the description below needs to be updated.}
-
-\hh{using (a) signatures and (b) profiles, run chumbley score across scans from NIST and CSAFE for various settings of $w_o$ and $w_v$ (and coarseness $c$ for profiles).}
-
-We used the adjusted Chumbley method as proposed in \citet{hadler} and implemented in the R package `toolmaRk` \citep{toolmark} on all pairwise land-to-land comparisons of the Hamby scans provided by NIST for a total of 85,491 land-to-land comparisons.
-
-\hh{XXX also calculate results for the adjusted adjusted score XXX}
-<!--The settings for optimizing and validating window sizes, $w_o$ and $w_v$, ranged from $w_o \in [50, 280]$ and $w_v \in \{30, 50\}$, see also figure \ref{fig:type2}.-->
-
+Both algorithms (CS1) and (CS2) are implemented in the R package `toolmaRk` \citep{toolmark}.
+We applied both methods to all pairwise land-to-land comparisons of the Hamby scans provided by NIST for a total of 85,491 land-to-land comparisons.
+
+# Results
 ## Failed Tests
@@
 \autoref{fig:fails} shows the percentage of failed tests among the 85,491 land-to-land comparisons of the NIST data for different values of the validation window size $w_v$.
-For same-source lands up to 12.5 percent of the tests fail.
-
-```{r fails, echo=FALSE, fig.height = 3, fig.width = 5, out.width='0.7\\textwidth', fig.cap="Percent of failed land-to-land comparisons using an optimization window $w_o = 120$ and a coarseness of $c = 0.25$.
-With an increase in the size of the validation window  a higher percentage of tests fails.
-The percentage of failures is higher for same-source lands than for different source lands."}
-fails <- allp %>% filter(coarse == 0.25) %>% 
-  group_by(wo, wv) %>% summarize(
+For same-source lands up to 12.5 percent of the tests fail using CS1.
+The highest percentage of failed tests under CS2 is 0.9% for different source tests using a validation window size $w_v$ of 50 pixels.
+}
+
+```{r fails, echo=FALSE, fig.height = 3, fig.width = 7, out.width='0.7\\textwidth', fig.cap="Percent of failed land-to-land comparisons using an optimization window $w_o = 120$ and a coarseness of $c = 0.25$.
+With an increase in the size of the validation window  a higher percentage of tests fails under both methods (CS1) and (CS2), but the percentage of failed tests is much smaller under (CS2)."}
+allinv$method <- "CS2"
+allp$method <- "CS1"
+
+prof_fails <- rbind(
+  allinv %>% select(wo, wv, method, p_value, match),
+  allp %>% filter(coarse == 0.25) %>% 
+    select(wo, wv, method, p_value, match)
+)
+fails <- prof_fails %>%
+  group_by(wo, wv, method, match) %>% summarize(
   failed = sum(is.na(p_value)),
@@
 
-allp %>% filter(coarse == 0.25) %>% 
-  group_by(wo, wv, match) %>% summarize(
-  failed = sum(is.na(p_value)),
-  n=n()
-) %>% 
+fails %>% 
   ggplot(aes(x = wv, y = failed/n*100, colour=factor(match), shape=factor(match))) +
@@
   scale_shape_discrete("Same-source") +
-  ylim(c(0,13)) 
-```
-
-# A modified test
-
-\hh{fix the algorithm - all of the other properties in Hadler and morris still hold.}
-
-\hh{number of failed tests down. }
-
-# Comparison of Results
-
-## Profiles
-
-\hh{When dealing with profiles, coarseness is an additional parameter that has to be considered in the matching.}
-
-\hh{Figure \ref{fig:coarse} shows the type II error rates for profiles using an optimization window $w_o= 120$ and a validation window $w_v = 30$ for varying level of coarseness.
-The type II error for all nominal levels of  $\alpha$ are the lowest for a coarseness range of 0.20 to 0.30, with a minim in type II reached at a coarseness $c = 0.25$.
-}
-\hh{However, for an optimal value of coarseness other aspects have to be condsidered.}
-```{r coarse, fig.cap="Type II error with respect to coarseness parameter over profiles, $w_o = 120$, $w_v = 30$.
-Optimal values for coarseness are around $c \\approx 0.30$", fig.width=5, fig.height = 3, out.width='.6\\textwidth'}
-e1<-errors2 %>% filter(wv==30)%>% ggplot(aes(x = coarse, y=beta, colour=factor(alpha))) +
-  geom_point() + geom_line() +
-  theme_bw() +
-  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + xlab("Profile Coarseness")+
-  ylab("Type II error") 
-e1
-```
+  ylim(c(0,13)) +
+  facet_wrap(~method, labeller="label_method")
+```
+
+\hh{Coarseness is another parameter that has to be considered in the matching of bullet land profiles.}
 
@@
 profile <- sigs_graphics[[1]] %>% filter(land_id==19)
+profile$lw10 <- lowess(x = profile$y, y = profile$value, f=.10)$y
+profile$lw15 <- lowess(x = profile$y, y = profile$value, f=.15)$y
 profile$lw20 <- lowess(x = profile$y, y = profile$value, f=.20)$y
@@
 
-profiles <- tidyr::gather(profile, coarseness, values, lw20:lw30)
+profiles <- tidyr::gather(profile, coarseness, values, lw10:lw30)
 profiles <- profiles %>% mutate(
@@
 
-## Results
+\hh{Figure \ref{fig:coarse} shows the type II error rates for profiles using an optimization window $w_o= 120$ and a validation window $w_v = 30$ for varying level of coarseness.
+The type II error for all nominal levels of  $\alpha$ are the lowest for a coarseness range of 0.20 to 0.30, with a minimum in type II reached at a coarseness $c = 0.25$.
+}
+\hh{However, for an optimal value of coarseness other aspects have to be condsidered.}
+```{r coarse, fig.cap="Type II error with respect to coarseness parameter over profiles, $w_o = 120$, $w_v = 30$.
+Optimal values for coarseness are around $c \\approx 0.30$", fig.width=5, fig.height = 3, out.width='.6\\textwidth'}
+e1<-errors2 %>% filter(wv==30)%>% ggplot(aes(x = coarse, y=beta, colour=factor(alpha))) +
+  geom_point() + geom_line() +
+  theme_bw() +
+  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + xlab("Profile Coarseness")+
+  ylab("Type II error") 
+e1
+```
+
+### Results for Signatures
 

### 2018-04-20 doc/paper.tex
setup for both methods
results from cs2

@@
 simultaneously. Let us assume that this results in \(I\) pairs. Define
-\(s_(j)^{(k)}\) to be the \(j\)th starting location in sequence
-\(k = 1, 2\), i.e.~\(s_(1)^{(k)} < s_(2)^{(k)} < ... < s_(I)^{(k)}\).
+\(s_{(j)}^{(k)}\) to be the \(j\)th starting location in sequence
+\(k = 1, 2\),
+i.e.~\(s_{(1)}^{(k)} < s_{(2)}^{(k)} < ... < s_{(I)}^{(k)}\).
 
@@
 tests.
+\hh{In the remainder of the paper, we will refer to the algorithm defined by \citet{hadler} as {\bf (CS1)} and the suggested modified algorithm as {\bf (CS2)} and compare their performance on the available scans of the Hamby study.}
 
@@
 \subsection{The Data}\label{the-data}
-
-\hh{Introduce profiles and signatures, as shown in figure \ref{fig:sigs-profiles}.}
 
@@
 
-\hh{Discuss CSAFE scans}
-
-Profiles and signatures were extracted from all scans as described in
+Profiles for each bullet land were extracted from scans close to the
+heel of the bullet while avoiding break-off as described in
 \citet{aoas}.
@@
 
-\hh{XXX the description below needs to be updated.}
-
-\hh{using (a) signatures and (b) profiles, run chumbley score across scans from NIST and CSAFE for various settings of $w_o$ and $w_v$ (and coarseness $c$ for profiles).}
-
-We used the adjusted Chumbley method as proposed in \citet{hadler} and
-implemented in the R package \texttt{toolmaRk} \citep{toolmark} on all
+Both algorithms (CS1) and (CS2) are implemented in the R package
+\texttt{toolmaRk} \citep{toolmark}. We applied both methods to all
 pairwise land-to-land comparisons of the Hamby scans provided by NIST
 for a total of 85,491 land-to-land comparisons.
+
+\section{Results}\label{results}
 
@@
 window size \(w_v\). For same-source lands up to 12.5 percent of the
-tests fail.
+tests fail using CS1.
+\hh{XXX Currently running CS2 for these settings for a comparison.}
 
@@
 
-\section{A modified test}\label{a-modified-test}
+\hh{XXX Conclusion: the number of failed tests goes down - a lot. XXX}
 
-\hh{fix the algorithm - all of the other properties in Hadler and morris still hold.}
-
-\hh{number of failed tests down. }
-
-\section{Comparison of Results}\label{comparison-of-results}
-
-\subsection{Profiles}\label{profiles}
-
-\hh{When dealing with profiles, coarseness is an additional parameter that has to be considered in the matching.}
-
-\hh{Figure \ref{fig:coarse} shows the type II error rates for profiles using an optimization window $w_o= 120$ and a validation window $w_v = 30$ for varying level of coarseness.
-The type II error for all nominal levels of  $\alpha$ are the lowest for a coarseness range of 0.20 to 0.30, with a minim in type II reached at a coarseness $c = 0.25$.
-}
-\hh{However, for an optimal value of coarseness other aspects have to be condsidered.}
-
-\begin{figure}
-
-{\centering \includegraphics[width=.6\textwidth]{figures/coarse-1}
-
-}
-
-\caption{Type II error with respect to coarseness parameter over profiles, $w_o = 120$, $w_v = 30$. Optimal values for coarseness are around $c \approx 0.30$}\label{fig:coarse}
-\end{figure}
+\hh{Coarseness is another parameter that has to be considered in the matching of bullet land profiles.}
 
@@
 
-\subsection{Results}\label{results}
+\hh{Figure \ref{fig:coarse} shows the type II error rates for profiles using an optimization window $w_o= 120$ and a validation window $w_v = 30$ for varying level of coarseness.
+The type II error for all nominal levels of  $\alpha$ are the lowest for a coarseness range of 0.20 to 0.30, with a minimum in type II reached at a coarseness $c = 0.25$.
+}
+\hh{However, for an optimal value of coarseness other aspects have to be condsidered.}
+
+\begin{figure}
+
+{\centering \includegraphics[width=.6\textwidth]{figures/coarse-1}
+
+}
+
+\caption{Type II error with respect to coarseness parameter over profiles, $w_o = 120$, $w_v = 30$. Optimal values for coarseness are around $c \approx 0.30$}\label{fig:coarse}
+\end{figure}
+
+\subsubsection{Results for Signatures}\label{results-for-signatures}
 

### 2018-04-20 doc/paper.Rmd
wv 60

@@
   allinv <- allinv %>% filter(land1_id < land2_id)
-  allinv$coarse <- allinv$coarseness
   allinv <- allinv %>% select(-coarseness)
@@
 For same-source lands up to 12.5 percent of the tests fail using CS1.
-The highest percentage of failed tests under CS2 is 0.9% for different source tests using a validation window size $w_v$ of 50 pixels.
-}
+The highest percentage of failed tests under CS2 is 1.3% for different source tests using a validation window size $w_v$ of 60 pixels.
 
@@
   ylim(c(0,13)) +
-  facet_wrap(~method, labeller="label_method")
+  facet_wrap(~method, labeller="label_both")
 ```
@@
 
-```{r profile-sketch, fig.height = 5, fig.width='\\textwidth', fig.cap="Overview of the effect of different coarseness parameters $c$ on the profile shown in Figure \\ref{fig:sigs-profiles} (top).
+```{r profile-sketch, fig.height = 4.5, fig.width='\\textwidth', fig.cap="Overview of the effect of different coarseness parameters $c$ on the profile shown in Figure \\ref{fig:sigs-profiles} (top).
 The bottom row shows histograms of  the (relative) optimal locations $t^o$ identified in the optimization step for different values of the coarseness parameter $c$.
@@
   ggplot(aes(x = locations/(length-wo))) +
-  geom_histogram(binwidth = 0.025) +
+  geom_histogram(binwidth = 0.01) +
   facet_grid(.~coarseness, labeller="label_both") + 

### 2018-04-20 doc/paper.tex
wv 60

@@
 window size \(w_v\). For same-source lands up to 12.5 percent of the
-tests fail using CS1.
-\hh{XXX Currently running CS2 for these settings for a comparison.}
+tests fail using CS1. The highest percentage of failed tests under CS2
+is 0.9\% for different source tests using a validation window size
+\(w_v\) of 50 pixels.
 
@@
 \caption{Percent of failed land-to-land comparisons using an optimization window $w_o = 120$ and a coarseness of $c = 0.25$.
-With an increase in the size of the validation window  a higher percentage of tests fails.
-The percentage of failures is higher for same-source lands than for different source lands.}\label{fig:fails}
+With an increase in the size of the validation window  a higher percentage of tests fails under both methods (CS1) and (CS2), but the percentage of failed tests is much smaller under (CS2).}\label{fig:fails}
 \end{figure}
-
-\hh{XXX Conclusion: the number of failed tests goes down - a lot. XXX}
 

### 2018-04-21 doc/paper.Rmd
expected failures, additional coarseness

@@
     tmp <- read.csv(file=file.path(path="../data/signatures", file)) 
-    if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% select(-x)
+    if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% dplyr::select(-x)
     all <- rbind(all, tmp)
@@
       tmp$coarse <- gsub(".*coar-(.*).csv", "\\1", file)
-    if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% select(-x)
+    if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% dplyr::select(-x)
     allp <- rbind(allp, tmp)
@@
   allinv <- allinv %>% filter(land1_id < land2_id)
-  allinv <- allinv %>% select(-coarseness)
+  allinv <- allinv %>% dplyr::select(-coarseness)
   saveRDS(allinv, file="../data/invprofiles.rds")
@@
       tmp$coarse <- gsub(".*coarse-(.*).csv", "\\1", file)
-    if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% select(-x)
+    if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% dplyr::select(-x)
     all_sig_c <- rbind(all_sig_c, tmp)
@@
 d2<- sigs_graphics.match[which(sigs_graphics.match$land_id==276,arr.ind = TRUE),]
-data1<-d1 %>%select(l30)
+data1<-d1 %>%dplyr::select(l30)
 #data1<- data1$y
-data2<- d2 %>%select(l30)
+data2<- d2 %>%dplyr::select(l30)
 #data2<- data2$y
@@
 
-sigs1.min = sigs_graphics[[1]] %>% filter(land_id==1) %>% select(y) %>% min()
-sigs2.min = sigs_graphics[[1]] %>% filter(land_id==276) %>% select(y) %>% min()
+sigs1.min = sigs_graphics[[1]] %>% filter(land_id==1) %>%
+  dplyr::select(y) %>% min()
+sigs2.min = sigs_graphics[[1]] %>% filter(land_id==276) %>% 
+  dplyr::select(y) %>% min()
 window_val <- 3*window_val*0.645
@@
 d2<- sigs_graphics.match[which(sigs_graphics.match$land_id==19,arr.ind = TRUE),]
-data1<-d1 %>%select(l30)
+data1<-d1 %>%dplyr::select(l30)
 #data1<- data1$y
-data2<- d2 %>%select(l30)
+data2<- d2 %>%dplyr::select(l30)
 #data2<- data2$y
@@
 The number of same-shift pairs will be zero, if the optimal locations $t_1^{o}$ and $t_2^{o}$ are so far apart, that no segments of size $w_v$ are left on the same sides of the optimal locations, i.e.
-$t_1^{o} < w_v$ and $t_2^{o} > T_2-w_o-w_v$ or  $t_1^{o} < T_1- w_o - w_v$ and $t_2^{o} < w_v$.
+$t_1^{o} < w_v$ and $t_2^{o} > T_2-w_o-w_v$ or  $t_1^{o} < T_1- w_o - w_v$ and $t_2^{o} < w_v$, i.e.
+we have a failure rate of
+\[
+P\left( t_1^{o} < w_v \ \cap \ t_2^{o} > T_2-w_o-w_v\right) + P\left( t_1^{o} < T_1- w_o - w_v \ \cap \ t_2^{o} < w_v\right).
+\]
 While we can assume that once (sub-)class characteristics are removed, optimal locations $t_i^{o}$ are uniformly distributed across the length of the profile, we cannot assume that $t_1^o$ and $t_2^o$ are independent of each other.
-In particular, for same-source profiles, we would expect a strong dependency between these locations.
-However, for different source matches, an argument of independence in the locations could be made.
+In particular, for same-source profiles, we would expect a strong dependency between these locations, in which case a large difference between locations is unlikely.
+However, for different source matches, we can assume that  locations are independent.
 In that case, we expect a test to fail with a probability of $\frac{2 w_v^2}{(T_1-w_o)(T_2-w_o)}$.
@@
 The highest percentage of failed tests under CS2 is 1.3% for different source tests using a validation window size $w_v$ of 60 pixels.
+\hh{Rates of expected failures are based on simulation runs using covariances between locations of same-source profiles of 0.854, and 0.120 for locations from different-source profiles, matching observed covariances for the Hamby scans.
+}\hh{ Observed failure rates are higher than expected.
+This might be due to remaining sub-class structure at a coarseness of 0.3 resulting in a distribution of optimal locations different from the assumed uniform.
+}
+
+```{r simulation, echo=FALSE}
+library(MASS) 
+Sigsame <- matrix(c(1, 0.85, 0.85, 1), nrow = 2) 
+Sigdiff <- matrix(c(1, 0.12, 0.12, 1), nrow = 2) 
+set.seed(20140501)
+simu <- 1:10 %>% purrr::map_df(.f = function(i) {
+  X1 <- mvrnorm(10000, mu = c(0,0), Sigma = Sigsame) 
+  X2 <- mvrnorm(10000, mu = c(0,0), Sigma = Sigdiff) 
+  Y1 <- pnorm(X1) 
+  Y2 <- pnorm(X2) 
+  wv <- 10*1:6
+
+  ps <- wv %>% purrr::map_df(.f = function(x) {
+    p <- x/(1200-120)
+    y1 <- Y1[,1]
+    y2 <- Y1[,2]
+    difffail1 <- sum((y1 < p & y2 < p) | (y1 > 1-p & y2 > 1-p))/length(y1)
+    samefail1 <- sum((y1 < p & y2 >1- p) | (y1 > 1-p & y2 < p))/length(y1)
+    y1 <- Y2[,1]
+    y2 <- Y2[,2]
+    difffail2 <- sum((y1 < p & y2 < p) | (y1 > 1-p & y2 > 1-p))/length(y1)
+    samefail2 <- sum((y1 < p & y2 > 1-p) | (y1 > 1-p & y2 < p))/length(y1)
+    data.frame(wv=x, samefail1=samefail1, samefail2=samefail2,
+               difffail1=difffail1, difffail2=difffail2)
+  })
+  ps$i <- i
+  ps
+})
+simu$`TRUE` <- simu$samefail1+simu$difffail1
+simu$`FALSE` <- simu$samefail2+simu$difffail2
+simulong <- simu %>% dplyr::select(wv, i, `TRUE`, `FALSE`) %>% 
+  gather(match, failed, 3:4)
+simusumm <- simulong %>% group_by(wv, match) %>%
+  summarize(
+    sdfailed = sd(failed),
+    failed = mean(failed)
+  )
+simusumm$method = "CS1"
+```
 
 ```{r fails, echo=FALSE, fig.height = 3, fig.width = 7, out.width='0.7\\textwidth', fig.cap="Percent of failed land-to-land comparisons using an optimization window $w_o = 120$ and a coarseness of $c = 0.25$.
-With an increase in the size of the validation window  a higher percentage of tests fails under both methods (CS1) and (CS2), but the percentage of failed tests is much smaller under (CS2)."}
+With an increase in the size of the validation window  a higher percentage of tests fails under both methods (CS1) and (CS2), but the percentage of failed tests is much smaller under (CS2).
+Observed failure rates are higher than expected rates."}
 allinv$method <- "CS2"
@@
 
+allinv <-  allinv %>% mutate(
+  loc1 = chumbley %>% purrr::map_int(.f = function(x) x$locations[1]),
+  loc2 = chumbley %>% purrr::map_int(.f = function(x) x$locations[2])
+)
+
 prof_fails <- rbind(
-  allinv %>% select(wo, wv, method, p_value, match),
+  allinv %>% dplyr::select(wo, wv, method, p_value, match),
   allp %>% filter(coarse == 0.25) %>% 
-    select(wo, wv, method, p_value, match)
+    dplyr::select(wo, wv, method, p_value, match)
 )
@@
   ggplot(aes(x = wv, y = failed/n*100, colour=factor(match), shape=factor(match))) +
-  geom_point(size=3) +
+  geom_point(aes(size="Observed")) +
   xlab("Size of validation window in pixels") +
@@
   ylim(c(0,13)) +
-  facet_wrap(~method, labeller="label_both")
-```
-
-\hh{Coarseness is another parameter that has to be considered in the matching of bullet land profiles.}
-
-\hh{\autoref{fig:profile-sketch} gives an overview of the effect of different coarseness parameters: from left to right, coarseness levels $c$ are 0.2, 0.25, and 0.3, respectively.
-The top row shows the resulting signature after smoothing the profile shown in Figure~\ref{fig:sigs-profiles} with the coarseness specified.
-The histograms in the bottom row show the relative optimal location $t^o$ (optimal location $t^o$ divided by the length of the corresponding profile).
-Optimal locations should be distributed uniformly.
-However, for coarseness values of $c = 0.25$ and $c=0.30$ we see very distinct boundary effects: optimal locations $t^o$ are found at the very extreme ends of a profile more often than one would expect based on a uniform location.}
+  facet_wrap(~method, labeller="label_both") +
+  geom_point(aes(y = failed*100, size = "Expected"), data=simusumm) +
+  scale_size_manual("Failures", values=c(1.5,3))
+```
+
+## Coarseness
+The purpose of the coarseness parameter is to remove (sub-)class characteristics from profiles before comparisons for matching.
+\citet{hadler} suggest a coarseness parameter of 0.25 in the setting of toolmark comparisons.
+For bullet lands, coarseness might need to be adjusted because of the strong effect  bullet curvature has on profiles.
+
+\hh{\autoref{fig:profile-sketch} gives an overview of the effect of different coarseness parameters: from left to right, coarseness levels $c$ are varied in steps of 0.05 from 0.1 to 0.3.
+The top row shows  resulting signature after smoothing the profile shown in Figure~\ref{fig:sigs-profiles} with the coarseness specified.
+The histograms in the bottom row show the relative optimal location $t^*$.
+Optimal locations are distributed uniformly once (sub-)class characteristics are removed.
+However, for coarseness values of $c > 0.20$ we see quite distinct boundary effects: optimal locations $t^*$ are found at the very extreme ends of a profile more often than one would expect based on a uniform distribution.}
 
@@
 
+profc10 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo_160-wv_30-c_10.rds")
+profc15 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo_160-wv_30-c_15.rds")
 profc20 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo_160-wv_30-c_20.rds")
@@
 profc30$coarseness <- .30
-profc <- bind_rows(profc20, profc25, profc30)
+profc <- bind_rows(profc10, profc15, profc20, profc25, profc30)
 profc <- profc %>% mutate(
@@
 profsumm <- read.csv("../data/profiles-summary.csv")
-profsumm <- profsumm %>% select(-run_id)
+profsumm <- profsumm %>% dplyr::select(-run_id)
 profc <- profc %>% left_join(profsumm, by=c("land2_id"="land_id", "profile2_id"="profile_id"))
@@
   ggplot(aes(x = locations/(length-wo))) +
-  geom_histogram(binwidth = 0.01) +
+  geom_histogram(binwidth = 0.025) +
   facet_grid(.~coarseness, labeller="label_both") + 
@@
 The type II error for all nominal levels of  $\alpha$ are the lowest for a coarseness range of 0.20 to 0.30, with a minimum in type II reached at a coarseness $c = 0.25$.
-}
+The histograms suggest to use a value of coarseness below 0.2.}
 \hh{However, for an optimal value of coarseness other aspects have to be condsidered.}
@@
   mutate(beta = round(100*beta, 1)) %>%
-  select(wv, type, alpha, beta) %>% 
+  dplyr::select(wv, type, alpha, beta) %>% 
   spread(alpha, beta)

### 2018-04-21 doc/paper.tex
expected failures, additional coarseness

--- 2018-04-20 doc/paper.tex
@@
 \(t_1^{o} < w_v\) and \(t_2^{o} > T_2-w_o-w_v\) or
-\(t_1^{o} < T_1- w_o - w_v\) and \(t_2^{o} < w_v\). While we can assume
-that once (sub-)class characteristics are removed, optimal locations
-\(t_i^{o}\) are uniformly distributed across the length of the profile,
-we cannot assume that \(t_1^o\) and \(t_2^o\) are independent of each
-other. In particular, for same-source profiles, we would expect a strong
-dependency between these locations. However, for different source
-matches, an argument of independence in the locations could be made. In
-that case, we expect a test to fail with a probability of
+\(t_1^{o} < T_1- w_o - w_v\) and \(t_2^{o} < w_v\), i.e.~we have a
+failure rate of \[
+P\left( t_1^{o} < w_v \ \cap \ t_2^{o} > T_2-w_o-w_v\right) + P\left( t_1^{o} < T_1- w_o - w_v \ \cap \ t_2^{o} < w_v\right).
+\] While we can assume that once (sub-)class characteristics are
+removed, optimal locations \(t_i^{o}\) are uniformly distributed across
+the length of the profile, we cannot assume that \(t_1^o\) and \(t_2^o\)
+are independent of each other. In particular, for same-source profiles,
+we would expect a strong dependency between these locations, in which
+case a large difference between locations is unlikely. However, for
+different source matches, we can assume that locations are independent.
+In that case, we expect a test to fail with a probability of
 \(\frac{2 w_v^2}{(T_1-w_o)(T_2-w_o)}\). For an average length of \(T_i\)
@@
 window size \(w_v\). For same-source lands up to 12.5 percent of the
-tests fail using CS1.
-\hh{XXX Currently running CS2 for these settings for a comparison.}
+tests fail using CS1. The highest percentage of failed tests under CS2
+is 1.3\% for different source tests using a validation window size
+\(w_v\) of 60 pixels.
+\hh{Rates of expected failures are based on simulation runs using covariances between locations of same-source profiles of 0.854, and 0.120 for locations from different-source profiles, matching observed covariances for the Hamby scans.
+}\hh{ Observed failure rates are higher than expected.
+This might be due to remaining sub-class structure at a coarseness of 0.3 resulting in a distribution of optimal locations different from the assumed uniform.
+}
 
@@
 \caption{Percent of failed land-to-land comparisons using an optimization window $w_o = 120$ and a coarseness of $c = 0.25$.
-With an increase in the size of the validation window  a higher percentage of tests fails.
-The percentage of failures is higher for same-source lands than for different source lands.}\label{fig:fails}
+With an increase in the size of the validation window  a higher percentage of tests fails under both methods (CS1) and (CS2), but the percentage of failed tests is much smaller under (CS2).
+Observed failure rates are higher than expected rates.}\label{fig:fails}
 \end{figure}
 
-\hh{XXX Conclusion: the number of failed tests goes down - a lot. XXX}
+\subsection{Coarseness}\label{coarseness}
 
-\hh{Coarseness is another parameter that has to be considered in the matching of bullet land profiles.}
+The purpose of the coarseness parameter is to remove (sub-)class
+characteristics from profiles before comparisons for matching.
+\citet{hadler} suggest a coarseness parameter of 0.25 in the setting of
+toolmark comparisons. For bullet lands, coarseness might need to be
+adjusted because of the strong effect bullet curvature has on profiles.
 
-\hh{\autoref{fig:profile-sketch} gives an overview of the effect of different coarseness parameters: from left to right, coarseness levels $c$ are 0.2, 0.25, and 0.3, respectively.
-The top row shows the resulting signature after smoothing the profile shown in Figure~\ref{fig:sigs-profiles} with the coarseness specified.
-The histograms in the bottom row show the relative optimal location $t^o$ (optimal location $t^o$ divided by the length of the corresponding profile).
-Optimal locations should be distributed uniformly.
-However, for coarseness values of $c = 0.25$ and $c=0.30$ we see very distinct boundary effects: optimal locations $t^o$ are found at the very extreme ends of a profile more often than one would expect based on a uniform location.}
+\hh{\autoref{fig:profile-sketch} gives an overview of the effect of different coarseness parameters: from left to right, coarseness levels $c$ are varied in steps of 0.05 from 0.1 to 0.3.
+The top row shows  resulting signature after smoothing the profile shown in Figure~\ref{fig:sigs-profiles} with the coarseness specified.
+The histograms in the bottom row show the relative optimal location $t^*$.
+Optimal locations are distributed uniformly once (sub-)class characteristics are removed.
+However, for coarseness values of $c > 0.20$ we see quite distinct boundary effects: optimal locations $t^*$ are found at the very extreme ends of a profile more often than one would expect based on a uniform distribution.}
 

### 2018-04-22 doc/paper.Rmd
crossvalidation files

@@
   X2 <- mvrnorm(10000, mu = c(0,0), Sigma = Sigdiff) 
-  Y1 <- pnorm(X1) 
+  Y1 <- pnorm(X1)  # Probability Integral Transform
   Y2 <- pnorm(X2) 
@@
 
-```{r profile-sketch, fig.height = 4.5, fig.width='\\textwidth', fig.cap="Overview of the effect of different coarseness parameters $c$ on the profile shown in Figure \\ref{fig:sigs-profiles} (top).
+```{r profile-sketch, fig.height = 4, fig.width='\\textwidth', fig.cap="Overview of the effect of different coarseness parameters $c$ on the profile shown in Figure \\ref{fig:sigs-profiles} (top).
 The bottom row shows histograms of  the (relative) optimal locations $t^o$ identified in the optimization step for different values of the coarseness parameter $c$.
@@
 
+## Type II error rates
+
+```{r}
+files <- dir("../data/crossvalidate-cs2/")
+cvcs2 <- data.frame()
+for (file in files) {
+  tmp <- readRDS(file.path("../data/crossvalidate-cs2/",file))
+  cvcs2 <- rbind(cvcs2, tmp)
+}
+cvcs2$coarse <- cvcs2$coarseness
+errorsCV <- 
+  rbind(errorrate(cvcs2, 0.001), 
+        errorrate(cvcs2, 0.005), 
+        errorrate(cvcs2, 0.01), 
+        errorrate(cvcs2, 0.05))
+
+errorsCV %>% 
+  ggplot(aes(x = wo, y = beta)) + 
+  geom_point(aes(colour = factor(alpha))) +
+  geom_line(aes(colour = factor(alpha))) +
+  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") +
+  scale_shape_discrete(expression("Size of\nvalidation window "~w[v])) +
+  xlab(expression("Window size for optimization "~w[o])) +
+  ylab("Type II error rate") +
+  theme_bw()
+
+```
+
+<!--
+## Type II error rate and coarseness
 \hh{Figure \ref{fig:coarse} shows the type II error rates for profiles using an optimization window $w_o= 120$ and a validation window $w_v = 30$ for varying level of coarseness.
@@
 ```
+
+-->
 

### 2018-04-22 doc/paper.tex
crossvalidation files

@@
 
-\hh{Figure \ref{fig:coarse} shows the type II error rates for profiles using an optimization window $w_o= 120$ and a validation window $w_v = 30$ for varying level of coarseness.
-The type II error for all nominal levels of  $\alpha$ are the lowest for a coarseness range of 0.20 to 0.30, with a minimum in type II reached at a coarseness $c = 0.25$.
-The histograms suggest to use a value of coarseness below 0.2.}
-\hh{However, for an optimal value of coarseness other aspects have to be condsidered.}
-
-\begin{figure}
-
-{\centering \includegraphics[width=.6\textwidth]{figures/coarse-1}
-
-}
-
-\caption{Type II error with respect to coarseness parameter over profiles, $w_o = 120$, $w_v = 30$. Optimal values for coarseness are around $c \approx 0.30$}\label{fig:coarse}
-\end{figure}
+\subsection{}\label{section}
 

### 2018-04-22 doc/paper.Rmd
crossvalidation files

@@
 
-```{r}
+\autoref{fig:type2} gives an overview of the type 2 error rates of methods CS1 and CS2 over a range of different optimization window sizes $w_o$.
+
+```{r type2, fig.height = 5, fig.width = 5, out.width='0.7\\textwidth', fig.cap="Side-by-side comparison of type II error rates of methods CS1 and CS2 by size of optimization window $w_o$."}
 files <- dir("../data/crossvalidate-cs2/")
@@
         errorrate(cvcs2, 0.05))
+errorsCV$method <- "CS2"
 
@@
   ylab("Type II error rate") +
-  theme_bw()
+  theme_bw() +
+  facet_grid(.~method, labeller="label_both")
 

### 2018-04-22 doc/paper.Rmd
duplicate label removed

@@
 
-```{r type2, fig.height = 5, fig.width = 5, out.width='0.7\\textwidth', fig.cap="Side-by-side comparison of type II error rates of methods CS1 and CS2 by size of optimization window $w_o$."}
+```{r type22, fig.height = 5, fig.width = 5, out.width='0.7\\textwidth', fig.cap="Side-by-side comparison of type II error rates of methods CS1 and CS2 by size of optimization window $w_o$."}
 files <- dir("../data/crossvalidate-cs2/")

### 2018-04-22 doc/paper.Rmd
change to ROC curves

--- 2018-04-22 doc/paper.Rmd
@@
 
-```{r}
+\autoref{fig:type2} gives an overview of ROC curves for methods CS1 and CS2 over a range of different optimization window sizes $w_o$.
+
+```{r type22, fig.height = 3.5, fig.width = 5, out.width='0.7\\textwidth', fig.cap="ROC curves of methods CS1 and CS2 for different sizes of optimization window $w_o$.
+Optimal ROC curves are reached at optimization windows of size 150 and higher.
+Points of equal error rates (EERs) can be found at the intersection of the dotted line and the ROC curves."}
 files <- dir("../data/crossvalidate-cs2/")
@@
 cvcs2$coarse <- cvcs2$coarseness
-errorsCV <- 
-  rbind(errorrate(cvcs2, 0.001), 
-        errorrate(cvcs2, 0.005), 
-        errorrate(cvcs2, 0.01), 
-        errorrate(cvcs2, 0.05))
-
-errorsCV %>% 
-  ggplot(aes(x = wo, y = beta)) + 
-  geom_point(aes(colour = factor(alpha))) +
-  geom_line(aes(colour = factor(alpha))) +
-  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") +
-  scale_shape_discrete(expression("Size of\nvalidation window "~w[v])) +
-  xlab(expression("Window size for optimization "~w[o])) +
-  ylab("Type II error rate") +
-  theme_bw()
+library(plotROC)
+cvcs2$method <- "CS2"
+
+cvcs2 %>% 
+  ggplot(aes(d = as.numeric(!match), m = p_value, colour=factor(wo))) + 
+  geom_abline(linetype=3, slope=-1, intercept = 1, colour="grey30", size=.5) +
+  geom_roc(labels=FALSE, n.cuts=100, pointsize=0) +
+  theme_bw() + 
+  scale_colour_brewer(
+    expression("Nominal\ntype I error"~alpha), palette="Set2") +
+  facet_grid(.~method, labeller="label_both") +
+  xlab("False Positives Rate (1 - Specificity)") +
+  ylab("True Positives Rate (Sensitivity)")
 

### 2018-04-22 doc/paper.tex
crossvalidation files
change to ROC curves

@@
 
-\subsection{}\label{section}
+\subsection{Type II error rates}\label{type-ii-error-rates}
+
+\autoref{fig:type2} gives an overview of ROC curves for methods CS1 and
+CS2 over a range of different optimization window sizes \(w_o\).
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.7\textwidth]{figures/type22-1}
+
+}
+
+\caption{ROC curves of methods CS1 and CS2 for different sizes of optimization window $w_o$.
+Optimal ROC curves are reached at optimization windows of size 150 and higher.
+Points of equal error rates (EERs) can be found at the intersection of the dotted line and the ROC curves.}\label{fig:type22}
+\end{figure}
 

### 2018-04-22 doc/paper.Rmd
first stab at conclusions

@@
   xlab(expression("Relative optimal location "~t^"*")) +
-  ylab("Number of profiles")
+  ylab("Number of profiles") +
+  scale_x_continuous(breaks=seq(0,1, by=0.25), labels=c("0", "0.25", "0.50", "0.75", "1"))
 
@@
 
-\autoref{fig:type2} gives an overview of ROC curves for methods CS1 and CS2 over a range of different optimization window sizes $w_o$.
-
-```{r type22, fig.height = 3.5, fig.width = 5, out.width='0.7\\textwidth', fig.cap="ROC curves of methods CS1 and CS2 for different sizes of optimization window $w_o$.
-Optimal ROC curves are reached at optimization windows of size 150 and higher.
-Points of equal error rates (EERs) can be found at the intersection of the dotted line and the ROC curves."}
+\autoref{fig:type2} gives an overview of the type 2 error of methods CS1 and CS2 across a range of different optimization windows $w_o$.
+
+```{r type2, fig.width=8, fig.height=5, out.width='.8\\textwidth', fig.cap="Type II error rates observed across a range of window sizes for optimization $w_o$.
+For a window size of $w_o = 120$ we see a drop in type II error rate across all type I rates considered.
+Smaller validation sizes $w_v$ are typically associated with a smaller type II error."}
 files <- dir("../data/crossvalidate-cs2/")
@@
 
+errorsCV <- rbind(errorrate(cvcs2, 0.001), 
+                errorrate(cvcs2, 0.005), 
+                errorrate(cvcs2, 0.01), 
+                errorrate(cvcs2, 0.05))
+errorsCV$type <- "CS2"
+
+errorsCV %>% 
+  ggplot(aes(x = wo, y = beta, colour=factor(alpha))) + 
+  geom_point(aes(shape=factor(wv)), size=3.5) +
+#  facet_grid(.~wo, labeller="label_both") + 
+ geom_line(aes(group=alpha),  size=.7) +
+  #  geom_smooth(aes(group=alpha), se=FALSE, method="loess", size=.7) +
+  theme_bw() +
+  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") +
+  scale_shape_discrete(expression("Size of\nvalidation window "~w[v])) +
+  xlab(expression("Window size for optimization "~w[o])) +
+  ylab("Type II error rate")
+
+labels = expression(P[M1](tilde(z)>0),P[M0](tilde(z)>0))
+```
+
+\autoref{fig:type22} gives an overview of ROC curves for methods CS1 and CS2 over a range of different optimization window sizes $w_o$.
+
+```{r type22, fig.height = 3.5, fig.width = 5, out.width='0.7\\textwidth', fig.cap="ROC curves of methods CS1 and CS2 for different sizes of optimization window $w_o$.
+Optimal ROC curves are reached at optimization windows of size 150 and higher.
+Points of equal error rates (EERs) can be found at the intersection of the dotted line and the ROC curves."}
+
 cvcs2 %>% 
   ggplot(aes(d = as.numeric(!match), m = p_value, colour=factor(wo))) + 
+  geom_roc(labels=FALSE, n.cuts=100, pointsize=0) +
   geom_abline(linetype=3, slope=-1, intercept = 1, colour="grey30", size=.5) +
-  geom_roc(labels=FALSE, n.cuts=100, pointsize=0) +
   theme_bw() + 
   scale_colour_brewer(
-    expression("Nominal\ntype I error"~alpha), palette="Set2") +
+    expression("Optimization\nwindow size"~w[o]), palette="Set2") +
   facet_grid(.~method, labeller="label_both") +
@@
 
-```{r type2, fig.width=8, fig.height=5, out.width='.8\\textwidth', fig.cap="Type II error rates observed across a range of window sizes for optimization $w_o$.
+<!--
+```{r type2b, fig.width=8, fig.height=5, out.width='.8\\textwidth', fig.cap="Type II error rates observed across a range of window sizes for optimization $w_o$.
 For a window size of $w_o = 120$ we see a drop in type II error rate across all type I rates considered.
@@
 ```
+-->
 
@@
 
-# Conclusion
-
-The results suggest that the Nominal type I error $\alpha$ value shows dependence on the size of the window of optimization.
+# Conclusions
+
+\hh{
+We started out trying to assess the suitability of the Chumbley Score's suitability for matching striae on bullet lands.
+Besides using recommended defaults, we also proposed ways to optimize parameters: based on the assumption that once sub-class characteristics are removed, optimal locations are distributed uniformly across the profile, we found  for bullets the smaller coarseness value of $c = 0.15$ to be suitable.
+Optimized window sizes were found based on cross-validation to minimize type 2 error rates. 
+Method CS1 proposed by \citet{hadler} has a minimal type 2 error rate of XXX for an optimized window size of YYY - which is considerably higher than the error rates achieved on matching toolmarks. 
+Unfortunately, method CS1 also has a high rate of failed tests -- situations, in which the algorithm does not provide a result, due to the way different-shift pairs are constructed. 
+XXX
+still to cover: introduce method CS2;
+failures are down; 
+type 2 error rates are also down.
+Fix has type 2 error rates on bullet lands that is still higher than the error rates achieved on the -much larger- toolmarks.
+However, bullets usually have multiple lands - in the case of Ruger P85s as used in the Hamby study, there are six lands for each bullet.
+We might be able to get more power out of the test by adapting CS2 to deal work in a bullet-to-bullet comparison.
+}
+
+<!-- The results suggest that the Nominal type I error $\alpha$ value shows dependence on the size of the window of optimization.
 For a given window of optimization the actual Type I error is comparable to the nominal level for only a select few validation window sizes and for comparable validation window sizes of 30 and 50 as done here, the actual type I error does not seem to vary as much as it varies with the optimization window sizes .

### 2018-04-22 doc/paper.tex
first stab at conclusions

@@
 
-\autoref{fig:type2} gives an overview of ROC curves for methods CS1 and
+\autoref{fig:type2} gives an overview of the type 2 error of methods CS1
+and CS2 across a range of different optimization windows \(w_o\).
+
+\begin{figure}
+
+{\centering \includegraphics[width=.8\textwidth]{figures/type2-1}
+
+}
+
+\caption{Type II error rates observed across a range of window sizes for optimization $w_o$.
+For a window size of $w_o = 120$ we see a drop in type II error rate across all type I rates considered.
+Smaller validation sizes $w_v$ are typically associated with a smaller type II error.}\label{fig:type2}
+\end{figure}
+
+\autoref{fig:type22} gives an overview of ROC curves for methods CS1 and
 CS2 over a range of different optimization window sizes \(w_o\).
@@
 
-{\centering \includegraphics[width=.8\textwidth]{figures/type2-1}
-
-}
-
-\caption{Type II error rates observed across a range of window sizes for optimization $w_o$.
-For a window size of $w_o = 120$ we see a drop in type II error rate across all type I rates considered.
-Smaller validation sizes $w_v$ are typically associated with a smaller type II error.}\label{fig:type2}
-\end{figure}
-
-\begin{figure}
-
 {\centering \includegraphics[width=\textwidth]{figures/type1-1}
@@
 
-\section{Conclusion}\label{conclusion}
+\section{Conclusions}\label{conclusions}
 
-The results suggest that the Nominal type I error \(\alpha\) value shows
-dependence on the size of the window of optimization. For a given window
-of optimization the actual Type I error is comparable to the nominal
-level for only a select few validation window sizes and for comparable
-validation window sizes of 30 and 50 as done here, the actual type I
-error does not seem to vary as much as it varies with the optimization
-window sizes . A Test Fail, i.e.~tests in which a particular parameter
-setting did not return a valid result, happens, when the shift to align
-two signatures is so large, that the remaining overlap is too small to
-accommodate windows for validation, depends on whether known-match or
-known non-matches has predictive value, with test results from different
-sources having a much higher chance to fail. On conducting an analysis
-of all known bullet lands using the adjusted chumbley algorithm, Type II
-error was identified to be least bad for window of validation 30 and
-window of optimization 120. In case of unsmoothed raw marks (profiles),
-Type II error increases with the amount of smoothing and least for
-LOWESS smoothing coarseness value about 0.25 or 0.3. In an effort to
-identify the level of adaptiveness of the algorithm, comparisons were
-made between signatures and profiles. Their comparison with respect to
-validation window size for a fixed optimization window size suggested
-that, profiles have a total error (i.e all incorrect classification of
-known-matches and known non-matches) greater than or equal to the total
-error of signatures for all sizes of validation window. Profiles also
-fail more number of times than signatures in a test fail (for different
-coarseness keeping windows fixed and also for different validation
-windows keeping coarseness fixed) which lets us conclude that the
-behaviour of the algorithm for the profiles instead of pre-processed
-signatures is not better. Finally it should be noted that the current
-version of the adjusted chumbley algorithm seems to fall short when
-compared to other machine-learning based methods \citet{aoas}, and some
-level of modification to the deterministic algorithm needs to be
-identified and tested that would reduce the number of incorrect
-classifications.
+\hh{
+We started out trying to assess the suitability of the Chumbley Score's suitability for matching striae on bullet lands.
+Besides using recommended defaults, we also proposed ways to optimize parameters: based on the assumption that once sub-class characteristics are removed, optimal locations are distributed uniformly across the profile, we found  for bullets the smaller coarseness value of $c = 0.15$ to be suitable.
+Optimized window sizes were found based on cross-validation to minimize type 2 error rates. 
+Method CS1 proposed by \citet{hadler} has a minimal type 2 error rate of XXX for an optimized window size of YYY - which is considerably higher than the error rates achieved on matching toolmarks. 
+Unfortunately, method CS1 also has a high rate of failed tests -- situations, in which the algorithm does not provide a result, due to the way different-shift pairs are constructed. 
+XXX
+still to cover: introduce method CS2;
+failures are down; 
+type 2 error rates are also down.
+Fix has type 2 error rates on bullet lands that is still higher than the error rates achieved on the -much larger- toolmarks.
+However, bullets usually have multiple lands - in the case of Ruger P85s as used in the Hamby study, there are six lands for each bullet.
+We might be able to get more power out of the test by adapting CS2 to deal work in a bullet-to-bullet comparison.
+}
 

### 2018-04-22 doc/paper.Rmd
type II error rates

@@
 
-```{r type2, fig.width=8, fig.height=5, out.width='.8\\textwidth', fig.cap="Type II error rates observed across a range of window sizes for optimization $w_o$.
-For a window size of $w_o = 120$ we see a drop in type II error rate across all type I rates considered.
+```{r type2, fig.height = 3, fig.width = 4.5, out.width='0.6\\textwidth', fig.cap="Type II error rates observed across a range of window sizes for optimization $w_o$.
+For a window size of $w_o = 130$ we see a minimum in type II error rate across all type I rates considered.
 Smaller validation sizes $w_v$ are typically associated with a smaller type II error."}
@@
   ggplot(aes(x = wo, y = beta, colour=factor(alpha))) + 
-  geom_point(aes(shape=factor(wv)), size=3.5) +
-#  facet_grid(.~wo, labeller="label_both") + 
+  geom_point(aes(shape=factor(wv)), size=2.5) +
   geom_line(aes(group=alpha),  size=.7) +
-  #  geom_smooth(aes(group=alpha), se=FALSE, method="loess", size=.7) +
-  theme_bw() +
-  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") +
+  theme_bw() +
+  scale_colour_brewer(expression("Nominal\ntype I error"~alpha),
+                      palette="Set2") +
   scale_shape_discrete(expression("Size of\nvalidation window "~w[v])) +
@@
 
-labels = expression(P[M1](tilde(z)>0),P[M0](tilde(z)>0))
 ```
@@
 
-```{r type22, fig.height = 3.5, fig.width = 5, out.width='0.7\\textwidth', fig.cap="ROC curves of methods CS1 and CS2 for different sizes of optimization window $w_o$.
+```{r type22, fig.height = 3, fig.width = 4.5, out.width='0.6\\textwidth', fig.cap="ROC curves of methods CS1 and CS2 for different sizes of optimization window $w_o$.
 Optimal ROC curves are reached at optimization windows of size 150 and higher.
@@
   ggplot(aes(d = as.numeric(!match), m = p_value, colour=factor(wo))) + 
-  geom_roc(labels=FALSE, n.cuts=100, pointsize=0) +
-  geom_abline(linetype=3, slope=-1, intercept = 1, colour="grey30", size=.5) +
+  geom_roc(labels=FALSE, n.cuts=100, pointsize=0, size=0.6) +
+  geom_abline(
+    linetype=3, slope=-1, intercept = 1, colour="grey30", size=.3) +
   theme_bw() + 

### 2018-04-23 presentation/presentation.Rmd
presentation

--- 2018-04-22 doc/paper.Rmd
@@
 ---
-title: "Adaption of the Chumbley Score to matching of bullet striation marks"
-authors:
-- affiliation: Department of Statistics, Iowa State University 
-  name: Ganesh Krishnan
-  thanks: The authors gratefully acknowledge ...
-- affiliation: Department of Statistics and CSAFE, Iowa State University 
-  name: Heike Hofmann
-biblio-style: apsr
+title: "Adapting the Chumbley Score to Bullet Striations"
+author: "Ganesh Krishnan, Heike Hofmann"
+date: "April 21, 2018"
+header-includes:
+   - \usepackage{cleveref}
+   - \usepackage{amsmath}
+   - \usepackage{amsfonts}
 output: 
-  pdf_document:
+  beamer_presentation:
     citation_package: natbib
-    fig_caption: yes
-    keep_tex: yes
-    latex_engine: pdflatex
-    template: template.tex
-  html_document: default
-blinded: 0
-keywords:
-- 3 to 6 keywords
-- that do not appear in the title
-bibliography: bibliography
-abstract: null
+bibliography: bibliography.bib
 ---
 
-\newcommand{\hh}[1]{{\textcolor{orange}{#1}}}
-\newcommand{\gk}[1]{{\textcolor{green}{#1}}}
-\newcommand{\cited}[1]{{\textcolor{red}{#1}}}
-\setlength\parindent{0pt}
-
-\tableofcontents
-\newpage
 ```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
@@
     tmp <- read.csv(file=file.path(path="../data/signatures", file)) 
-    if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% dplyr::select(-x)
+    if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% select(-x)
     all <- rbind(all, tmp)
@@
       tmp$coarse <- gsub(".*coar-(.*).csv", "\\1", file)
-    if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% dplyr::select(-x)
+    if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% select(-x)
     allp <- rbind(allp, tmp)
@@
   allinv <- allinv %>% filter(land1_id < land2_id)
-  allinv <- allinv %>% dplyr::select(-coarseness)
+  allinv <- allinv %>% select(-coarseness)
   saveRDS(allinv, file="../data/invprofiles.rds")
@@
       tmp$coarse <- gsub(".*coarse-(.*).csv", "\\1", file)
-    if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% dplyr::select(-x)
+    if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% select(-x)
     all_sig_c <- rbind(all_sig_c, tmp)
@@
 
-# Introduction and Background
+## Objective and Motivation
 
-## Motivation
-Same source analyses are a major part of an Forensic Toolmark Examiner's job.
-In current practice examiners  make these comparisons by visual inspection under a comparison microscope and come to one of the following four conclusions: identification, inconclusive, elimination or unsuitable for examination~\citep{afte-toolmarks1998}.
-These conclusions are made on the basis of "unique surface contours" of the two toolmarks being in "sufficient agreement" \citep{afte-toolmarks1998}.
-AFTE describes the term "sufficient agreement" as the possibility of another tool producing the markings under comparison, as practically impossible \citep{afte-toolmarks1998}.
-Potential subject bias in the assessment as well as the lack of specified error rates are the main points of criticisms  first raised by the National Research Council in 2009 \citep{NAS:2009} and later emphasized further by the President's Council of Advisors on Science and Technology \citep{pcast2016}.
+- Same Source Matching of Bullet lands
+- Evaluate performance of Chumbley Score method when used for Bullet Striations
+- Bullet striations have curvature, not present in toolmarks
+- Identify Error rates and effect of different parameters on them (In short finding the best error rates possible )
 
-Technological advances, such as profilometers and confocal microscopy allow to measure 3D surfaces in a high-resolution digitized form.
-This technology has become more accessible over the last decade, and has made its way into topological images of ballistics evidence, such as bullet lands and breech faces \citep{DeKinder1, DeKinder2, Bachrach1, vorburger2016}.
-Digitized images of 3D surfaces of form the data basis of statistical analysis of toolmarks.
-A statistical approach based on data  removes both subjectivity from the assessment and allows a quantification of error rates for both false positive and false negative identifications.
+### Structure
+- Data being used
+- What is the Chumbley Score Method?
+- Identifying Best parameter Settings for Bullets
+- Modifications to the Algorithm
+- Results using \citet{hadler} method and results using Modified method
 
-\hh{In the next page and a half it is easy to lose the red line. It might help to include a table with an overview.
-The table should include the reference to the paper, the data used, the statistical method and the associated error rates.}
-Various toolmarks have been studied in the literature:  \citet{manytoolmarks1} and \citet{chumbley} have been analyzing screwdriver marks  digitized using a profilometer;  \citet{manytoolmarks2} have investigated 3D marks from screwdriver, tongue and groove pliers captured using a confocal microscope; \citet{afte-chumbley} have been investigated digitized marks from slip-joint pliers generated by a surface profilometer.
+## Data
+-  Ruger P85s Bullet Lands, or Hamby scans (\cite{hamby}) provided by NIST  (85,491 comparisons)
+-  Bullet striation marks $\approx$ 2mm | Screwdriver marks $\approx$ 7mm (all chumbley score papers)
 
-\hh{We need an additional sentence here to get from the data to the statistical methods ... }
-\citet{manytoolmarks2} define a relative distance metric and use it as similarity measure between two toolmarks.
-\citet{manytoolmarks1} extract many small segments in the markings of two toolmarks and compare similiarity using a maximum pearson correlation coefficient.
-The Chumbley scoring method, first introduced by \citet{chumbley}, uses a similar but more extensive framework based on a Mann-Whitney U test of the resulting correlation coefficients.
-This approach is non-deterministic, because segments are chosen randomly.
-\citep{hadler} make the score deterministic for each pair of toolmarks by choosing segments for comparison systematically.
-This approach also ensures independence between segements of striae.
-In this paper, we are investigating the applicability of the Chumbley scoring method by \citet{hadler} to assess striation marks on bullet lands for same-source identification.
+### Getting from Profiles to Signatures
+\vspace{-0.5em}
 
-Striation marks on bullets are made by impurities in the barrel.
-As the bullet travels through the barrel, these imperfections leave "scratches" on the bullet surface.
-Typically, only striation marks in the land engraved areas (LEAs) are considered \citet{afte-article1992}.
-Bullet lands are depressed areas between the  grooves made by the rifling action of the barrel.
-Compared to toolmarks made by screwdrivers striation marks on bullets are typically much smaller, both in length and in width.
-Bullets also have a curved cross-sectional topography.
-Figure \ref{fig:rgl} shows us how the signature from a bullet land (bottom) lines up with the image of the land (top) from which it was extracted.
-We can also see in the figure how the depth and relative position of the striation markings seen in the image are interpreted as the signature.
+- Lowess smoothing or Loess fit residuals $=$ Signatures
+- Removes topographic structure (curvature)
+- Improve the signal to noise Ratio
 
-Bullet matching methods are usually based on these associated signatures.
-\citet{chu2013} use an automatic method for counting consecutive matching striae (CMS).
-The authors report an error rate of 52% of the known same source lands comparisons as misidentified (false negative) and zero false positives for known different source lands.
-\citet{ma2004} and \citet{vorburger2011} discuss CCF (cross-correlation function) and its discriminating power and applicability for same-source analyses of bullets, but do not provide any error rates in their discussion.
-\citet{aoas} use multiple features like CCF, CMS, D (distance measure) etc in a random forest based method and compare every land against every other land of  digitised versions of Hamby 252 and Hamby 44 \citep{hamby} published on the NIST Ballistics Database \citep{nist}.
-The authors report an out-of-bag overall error rate of 0.46%, comprised of an error rate of 30.05% of  same-source pairs that were not identified and an error rate of 0.026% of different-source pairs that were incorrectly identified as same-source.
+\vspace{-1.5em}
 
-The Chumbley score  provides us with another approach in the same-source assessment of bullet striation marks.
-\citet{chumbley} compare two toolmarks for same-source.
-The data for this study was obtained from 50 sequentially manufactured screwdriver tips.
-\citet{chumbley} report error rates for markings made by the tips at different angles.
-For markings made at 30 degree the authors  report an average false negative error rate of 0.089  and an average false positive error rate of 0.023.
-For other angles of 60 and 85 degrees the false negatives error rate is 0.09 while the rate of false positives decreases to 0.01.
-The paper by \citet{hadler} is based on the same data but the authors focus on  markings made under the same angle.
-The error rates associated with the deterministic version of the score are  0.06 for false  negatives and a false positive error rate of 0.
+```{r sigs-profiles, fig.width = 8, fig.height = 2.5, fig.cap="Bullet land profile (left) and the corresponding signature (right)."}
+sigs_graphics<- readRDS("../data/sigs_generate_window.rds")
 
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{images/B6-B2-L6-rescaled.png}
-```{r, eval=FALSE, echo=FALSE, fig.width=8, fig.height=3, out.width='\\textwidth'}
-land <- read.csv("../data/b6b2l6.csv")
-land %>% ggplot(aes(x = y, y=resid)) +geom_line() +theme_bw() +
+sigs_graphics.match<- as.data.frame(sigs_graphics[1])
+sigs_graphics.nonmatch<- as.data.frame(sigs_graphics[2])
+colnames(sigs_graphics.match)<- gsub("^.*\\.","", colnames(sigs_graphics.match))
+colnames(sigs_graphics.nonmatch)<- gsub("^.*\\.","", colnames(sigs_graphics.nonmatch))
+p1 <- sigs_graphics[[1]] %>% filter(land_id==19) %>%
+  ggplot(aes(x = y, y = value)) + geom_line() +
+  theme_bw() +
   ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
-  xlab(expression(paste("Relative Location (in ",mu,m,")", sep="")))
+  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
+  ggtitle("Profile")
+
+p2 <- sigs_graphics[[1]] %>% filter(land_id==19) %>%
+  ggplot(aes(x = y, y = l30)) + 
+  geom_line(aes(y=resid), size=.25) +
+  geom_line(aes(y=l30)) +
+  theme_bw() +
+  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
+  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
+  ggtitle("Signature")
+
+grid.arrange(p1,p2, ncol=2)
 ```
 
-\caption{\label{fig:rgl} Image of a bullet land from a confocal light microscope at 20 fold magnification (top) and a chart of the corresponding signature of the same land (bottom).
-The dotted lines connect some peaks visible in both visualizations.}
+## Error Rates for toolmarks
 
+\begin{table}[ht]
+\centering
+\resizebox{\textwidth}{!}{\begin{tabular}{rlrrr}
+  \hline
+Research paper & Method & Data Source & False Positives & False Negatives\\
+  \hline
+ \textbf{\citet{manytoolmarks1}} & Maximized & Screwdrivers & & \\ & Correlation &  & - & - \\ \hline
+ \textbf{\citet{chumbley}} & Randomized & Screwdrivers & & \\ (Same-Surface Same-Angle) & Chumbley Score & & 2.3\% & 8.9\%  \\ \hline
+  \textbf{\citet{afte-chumbley}} & Randomized & Slip-joint  &  & \\ & Chumbley Score &  & - & - \\ \hline 
+  \textbf{\citet{hadler}} & Deterministic & Screwdrivers & & \\ (Same-Surface Same-Angle) & Chumbley Score & & 0\% & 6\%  \\ \hline
+ \textbf{\citet{manytoolmarks2}} & Similarity Measure & Screwdrivers & & \\ (Different Surfaces-same angle) & Relative Distance Metric & & 5.9\% &  9.4\% \\ (Same Surfaces-same angle)&  & & 0.22\% &  0\% \\
+   \hline
+\end{tabular}}
+\caption{\label{tab:toolmarks-ER} Error Rates for Toolmarks}
+\end{table}
+
+## Digitized Striation Marks
+
+- Let $x(t_1)$, $t_1 = 1,2,...T_1$ and $y(t_2)$, $t_2 = 1,2...T_2$ be two digitized marks (where $T_1$ and $T_2$ are not necessarily equal).
+
+- $T_1$ and $T_2$ are the final pixel indexes of each marking. 
+Therefore give the respective lengths of the markings.
+
+- Signatures/ Profiles (NIST- Hamby) $\approx$ 1200 pixels (2 mm)
+  Screwdriver toolmarks (Chumbley Papers) $\approx$ 9000 pixels (7 mm)
+- Markings are smoothed as a preprocessing step
+    - Used to remove drift and (sub)class characteristics from individual markings
+
+## Overview
+
+\begin{figure}[hbtp]
+\centering
+\includegraphics[width=1\textwidth]{imgs/flow-chart-adj-chumbley.png}
+
+\caption{\label{flow-chart} An overview of the chumbley score method}
 \end{figure}
 
-## Scans for land engraved areas
+## Chumbley Score
 
-Comparisons of striae from bullets are usually based on comparisons of striae in land engraved areas, which are extracted in form of cross sections, called *profiles* \citep{aoas,ma2004}.
-From profiles bullet *signatures* \citep{chu2013,aoas} are extracted as residuals of a loess fit or Gaussian filter.
-This effectively removes topographic structure from the data in the attempt to increase the signal to noise ratio.
-The span of the loess fit was found using cross-validation, as described by \citet{aoas}.
+- Two steps: Optimization ($1^{st}$) and Validation ($2^{nd}$).
+- Windows $\implies$ short segments of the markings
+    - Have \textit{predefined sizes}. ($T_1$ or $T_2 >>> w_o >> w_v$)
+          1. $w_o$ used in the Optimization step 
+          2. $w_v$ used in the Validation step
 
-\hh{don't split the discussion on the size. between the next paragraph and }
-<!--The majority of Bullet profiles and signatures of the Hamby study \citep{hamby} used in this paper are extracted by procedures mentioned by  \citet{aoas}.--> 
-There are two sources of scans for sets from the Hamby study available to us: scans of Hamby 44 and Hamby 252 are available from the NIST database \citep{nist}.
-Hamby 44 has also been made available to us and has been scanned locally for CSAFE at the Roy J.\ Carver High Resolution Microscopy Facility using a Sensofar confocal light microscope.
-Scans in the NIST database are made with a NanoFocus at 20x magnification.
-The resolutions of the two instruments are different: the NIST scans are taken at a resolution of 1.5625 $\mu m$ per pixel, while the CSAFE scans are available at a resolution of  0.645 $\mu m$ per pixel.
-The length of an average bullet land from Hamby (9 mm Ruger P85) is about 2 millimeter, resulting in signatures of about 1200 pixels for NIST scans, and about 3000 pixels for CSAFE scans.
+\vspace{-1em}
 
-In comparison, scans from the profilometer used by \citet{chumbley, hadler} were taken at a resolution of about 0.73 $\mu m$ per pixel.
-The screw driver toolmarks are  about 7 mm in length \citep{manytoolmarks1}, for a total of over 9000 pixels for the width of these scans.
+#### Optimization step
+\vspace{-0.5em}
 
-This severe limitation in the amount of available data poses the main challenge in adapting the Chumbley score to matching bullet lands, because of the resulting loss in power.
-
-## The Chumbley Score Test
-
-The Chumbley score algorithm takes input in form of two digitized toolmarks.
-The toolmark is in form of $z(t)$ which is a spatial process for location indexed by $t$.
-$t$ here denotes equally spaced pixel locations for the striation marks under consideration, $t = 1, ..., T$.
-Let further $z^s(t)$ denote a vector of markings of length $s$ starting in location  $t$.
-
-Let $x(t_1)$, $t_1 = 1,2,...T_1$ and $y(t_2)$, $t_2 = 1,2...T_2$ be two digitized toolmarks (where $T_1$ and $T_2$ are not necessarily equal). 
-<!--The vectorized processes representing two sets of striaes are therefore shown as $x(t_1)$, $t_1 = 1,2,...T_1$ and $y(t_2)$, $t_2 = 1,2...T_2$.
-Here $x$ and $y$ denote two instances of the process $z(*)$ which means two striation marks, while the indexing $t$ or pixel range is shown as $t_1$ and $t_2$ for $x$ and $y$ respectively.
-This representation of $t$ as $t_1$ and $t_2$ lets us identify them as either of same or different lengths.-->
-The toolmarks under consideration are potentially from two different sources or the same source.
-$T_1$ and $T_2$, as represented above, are the final pixel indexes of each marking and therefore give the respective lengths of the markings.
-
-In a pre-processing step  the two markings are smoothed using a lowess \citep{lowess} with coarseness parameter $c$. 
- Originally, this smoothing is intended to remove drift and (sub)class characteristics from individual markings, however, in the setting of matching bullet striae, we can also make use of this mechanism to separate bullet curvature in profiles from signatures before matching signatures.
-
-After removing sub-class structure, 
-the Chumbley scores is calculated in two steps: an optimization step and a validation step.
-In the optimization step, the two markings are aligned horizontally such that within a pre-defined window of length $w_o$ the correlation between $x(t_1)$ and $y(t_2)$ is maximized:
+- Aligns markings horizontally 
+- Correlation of all possible windows of size $w_o$ between $x(t_1)$ and $y(t_2)$ computed
+- \textit{Possible lag in the 2 markings} \newline Window Pair with maximized correlation $\implies$ \newline Optimal vertical (in-phase) shift of $t_1^o - t_2^o$ 
+      - For aligning the two markings. 
+\vspace{-1em}
 \[
@@
 \]
-This results in an optimal vertical (in-phase) shift of $t_1^o - t_2^o$ for aligning the two markings.
-\hh{We will denote the relative optimal locations as $t_1^*$ and $t_2^*$, where $t_i^* = t_i^o/(T_i-w_o)$ for $i=1,2$, such that $t_1^*, t_2^* \in [0,1]$.
-Once (sub-)class characteristics are removed, the relative optimal locations should be distributed according to a uniform distribution in $[0,1]$.
-}
 
-In the validation step,  two sets of windows of size $w_v$ are chosen from both markings (see Figure \ref{fig:win-comparison}).
-In the first set, pairs of windows are extracted from the two markings using the optimal vertical shift as determined in the first step, whereas for the second set the windows are extracted using a different (out-of-phase) shift.
+\hspace{-1.5em}
+where $t_1^o,t_2^0$ are the respective starting points of $w_o$ in $x(t_1)$ and $y(t_2)$
 
-More precisely, let us define starting points $s_i^{(k)}$ for each signature $k = 1, 2$ as 
-\begin{eqnarray}\label{eq.start}
-s^{(k)}_i = 
-\begin{cases}
-t_k^* + i w_v & \text{ for } i < 0 \\
-t_k^* + w_ o + i w_v & \text{ for } i \ge 0,
-\end{cases}
-\end{eqnarray} for integer values of $i$ with $0 <  s^{(k)}_i \le T_k - w_v$.
+## 2
+- Let $t_1^*$ and $t_2^*$ be relative optimal locations, where $t_i^* = t_i^o/(T_i-w_o)$ for $i=1,2$, such that $t_1^*, t_2^* \in [0,1]$. 
+-  Once (sub-)class characteristics are removed, these locations have uniform distribution in $[0,1]$
 
-Same-shift pairs of length $w_v$ are defined in \citet{hadler} as all 
-pairs $(s_i^{(1)}, s_i^{(2)})$ for all integer values $i$ for which both $s_i^{(1)}$ and  $s_i^{(2)}$ are defined.
-Similarly, different-shift pairs are defined as $(s_i^{(1)}, s_{-i-1}^{(2)})$ for all $i$ where both $s_i^{(1)}$ and  $s_{-i-1}^{(2)}$ are defined (see \autoref{sketch-same-diff}).
+#### Validations Step
+- Two sets of windows of size $w_v$ chosen from both markings (see Figure \ref{fig:win-comparison})
 
-\begin{figure}[hbtp]
-\centering
-\includegraphics[width=.7\textwidth]{images/sketch-same.png}
+- First set or \textbf{Same Shift}
+      - pairs of windows are extracted from the two markings using the optimal vertical shift.
+$t_1^o - t_2^o$
 
-\includegraphics[width=.7\textwidth]{images/sketch-diff.png}
-\caption{\label{sketch-same-diff}Sketch of same-shift pairings  (top) and different-shift pairings (bottom).
-Filled in rectangles show pairings resulting in correlations, unfilled rectangles are segments without a match.}
-\end{figure}
+- Second set or \textbf{Different Shift}
+      - the windows are extracted using a different (out-of-phase) shift.
+
+## In-phase and Out-of-phase
 
@@
 
-```{r win-comparison-2, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap=" The two plots on the left show how the same shift behaves in case of a matching pair and the two plots on the right show how the different shift behaves in case of a matching pair.", eval=FALSE}
+## 2
 
-asd<- all %>% filter(wo==120, wv==50, match == TRUE, land1_id == 1)
+More precisely, let us define starting points of the windows of validation $s_i^{(k)}$ for each signature $k = 1, 2$ as 
+\begin{eqnarray}\label{eq.start}
+s^{(k)}_i = 
+\begin{cases}
+t_k^* + i w_v^* & \text{ for } i < 0 \\
+t_k^* + w_ o^* + i w_v^* & \text{ for } i \ge 0,
+\end{cases}
+\end{eqnarray} for integer values of $i$ with $0 <  s^{(k)}_i \le 1 - w_v^*$
+<!-- $0 <  s^{(k)}_i \le T_k - w_v$. -->
 
-sigs_graphics<- readRDS("../data/sigs_generate_window.rds")
+## 3
 
-sigs_graphics.match<- as.data.frame(sigs_graphics[1])
-sigs_graphics.nonmatch<- as.data.frame(sigs_graphics[2])
-colnames(sigs_graphics.match)<- gsub("^.*\\.","", colnames(sigs_graphics.match))
-colnames(sigs_graphics.nonmatch)<- gsub("^.*\\.","", colnames(sigs_graphics.nonmatch))
+- Same-shift pairs of length $w_v$ are all pairs $$(s_i^{(1)}, s_i^{(2)}) \quad \forall \, i \, \in \, \mathbb{Z}$$ for which both $s_i^{(1)}$ and  $s_i^{(2)}$ are defined. 
+- Different-shift pairs are defined as $$(s_i^{(1)}, s_{-i-1}^{(2)}) \quad \forall \, i \, \in \, \mathbb{Z}$$ where both $s_i^{(1)}$ and  $s_{-i-1}^{(2)}$ are defined (see \cref{sketch-same-diff}).
 
-# Matching signature
-#gi<-
-#sigs_graphics.match<- sigs_graphics.match %>% filter(land_id == c(1,9))  
-#sigs_graphics.match<- sigs_graphics.match %>% select(y, l30, land_id)
-d1<- sigs_graphics.match[which(sigs_graphics.match$land_id==1,arr.ind = TRUE),]
-d2<- sigs_graphics.match[which(sigs_graphics.match$land_id==19,arr.ind = TRUE),]
-data1<-d1 %>%dplyr::select(l30)
-#data1<- data1$y
-data2<- d2 %>%dplyr::select(l30)
-#data2<- data2$y
-window_opt = 120 
-window_val = 50
-coarse = 1
-data1<- matrix(unlist(data1))
-data2<- matrix(unlist(data2))
+\begin{figure}[hbtp]
+\centering
+\includegraphics[width=.7\textwidth]{../doc/images/sketch-same.png}
 
-  unity <- function(x) {x / sqrt(sum(x^2))} ## normalize columns of a matrix to make correlation computation faster
+\includegraphics[width=.7\textwidth]{../doc/images/sketch-diff.png}
+\caption{\label{sketch-same-diff}Sketch of same-shift pairings  (top) and different-shift pairings (bottom).
+Filled in rectangles show pairings resulting in correlations, unfilled rectangles are segments without a match.}
+\end{figure}
 
-  ####################################################
-  ##Clean the marks and compute the smooth residuals##
-  ####################################################
+## 4
+- No overlaps of pairs within selected marks $\implies$ independence.
+- Both same- and different-shift pairs correlations between the  markings are calculated. 
+- For Same-Source markings correlations
+    - for the in-phase shift should be high
+    - for out-of-phase shift should be low.
+          - Provide a measure for the base-level correlation to which in-phase shift correlations can be compared.
+- The Chumbley score is the Mann Whitney U statistic computed by comparing between in-phase sample and out-of-phase sample.
 
-  data1 <- matrix(data1[round((0.01*nrow(data1))):round(0.99*nrow(data1)),], ncol = 1)
-  data2 <- matrix(data2[round((0.01*nrow(data2))):round(0.99*nrow(data2)),], ncol = 1)
+## Failed Tests
 
-  ##Normalize the tool marks
-  y1 <- data1 - lowess(y = data1,  x = 1:nrow(data1), f= coarse)$y
-  y2 <- data2 - lowess(y = data2,  x = 1:nrow(data2), f= coarse)$y
+- By definition (equation \autoref{eq.start}), some number of tests fail to produce a result
+- Either because the number of eligible same-shift pairs is 0, or the number of different-shift pairs is 0.
 
-  ############################################
-  ##Compute the observed maximum correlation##
-  ############################################
+\vspace{-0.5em}
 
-  #####################
-  ##Optimization step##
-  #####################
-  ##Each column in these matrices corresponds to a window in the respective tool mark
-  y1_mat_opt <- matrix(NA, ncol = length(1:(length(y1) - (window_opt - 1))), nrow = window_opt)
-  for(l in 1:(length(y1) - (window_opt - 1))){
-    y1_mat_opt[,l] <- y1[l:(l+(window_opt - 1))]
-  }
-  y2_mat_opt <- matrix(NA, ncol = length(1:(length(y2) - (window_opt - 1))), nrow = window_opt)
-  for(l in 1:(length(y2) - (window_opt - 1))){
-    y2_mat_opt[,l] <- y2[l:(l+(window_opt - 1))]
-  }
+###### No Same-shift pairs if
+- the optimal locations $t_1^{o}$ and $t_2^{o}$ are in the extremes
+    - no segments of size $w_v$ are left on the same sides of the optimal locations, 
+    - For all $t_1^{o} < w_v$ and $t_2^{o} > T_2-w_o-w_v$ or \newline $t_1^{o} < T_1- w_o - w_v$ and $t_2^{o} < w_v$
 
-  ##Compute the correlation between all pairs of windows for the two marks
-  ##Rows in the following matrix are mark 2, columns are mark 1
-  y2_mat_opt <- apply(scale(y2_mat_opt), 2, unity)
-  y1_mat_opt <- apply(scale(y1_mat_opt), 2, unity)
-  corr_mat_opt <- t(y2_mat_opt) %*% y1_mat_opt ##correlation matrix
-  max_corr_opt_loc <- which(corr_mat_opt == max(corr_mat_opt), arr.ind = TRUE) ##pair of windows maximizing the correlation
+\vspace{-0.5em}
 
-  s1.rect.opt <- data.frame(xmin=max_corr_opt_loc[1,2], xmax= max_corr_opt_loc[1,2]+window_opt, ymin=-Inf, ymax=Inf)
-    s1.rect.sameshift1 <- data.frame(xmin=max_corr_opt_loc[1,2]-window_val, xmax= max_corr_opt_loc[1,2]- 2*window_val, ymin=-Inf, ymax=Inf)
-       s1.rect.sameshift2 <- data.frame(xmin=max_corr_opt_loc[1,2]+window_opt+ window_val, xmax= max_corr_opt_loc[1,2]+ window_opt+2*window_val, ymin=-Inf, ymax=Inf)
+###### No Different Shift Pairs if
+- the optimal locations are close to the boundaries
+    - \# of different-shift pairs also depends on the placement of the  optimal locations $t_i^o$.
+    -  For all $t_i^{(o)} < w_v$ or  $t_i^{(o)} > T_i-w_o- w_v$.
 
-  s2.rect.opt <- data.frame(xmin=max_corr_opt_loc[1,1], xmax= max_corr_opt_loc[1,1]-window_opt, ymin=-Inf, ymax=Inf)
-    s2.rect.sameshift1 <- data.frame(xmin=max_corr_opt_loc[1,1]-window_opt-window_val, xmax= max_corr_opt_loc[1,1]-window_opt- 2*window_val, ymin=-Inf, ymax=Inf)
-       s2.rect.sameshift2 <- data.frame(xmin=max_corr_opt_loc[1,1]+ window_val, xmax= max_corr_opt_loc[1,1]+ 2*window_val, ymin=-Inf, ymax=Inf)
+\vspace{-0.5em}
 
-p1<-  ggplot(data = d1, aes(x = y, y = l30)) + geom_line() + 
-  geom_rect(data=s1.rect.opt, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "black",alpha=0.5, inherit.aes = FALSE)+
-    geom_rect(data=s1.rect.sameshift1, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "red",alpha=0.1, inherit.aes = FALSE)+
-    geom_rect(data=s1.rect.sameshift2, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "blue",alpha=0.1, inherit.aes = FALSE)+
-  theme_bw() + labs(y = "Signature Land Id = 1")
+##
 
- p2<-   ggplot(data = d2, aes(x = y, y = l30)) + geom_line() + 
-   geom_rect(data=s2.rect.opt, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "black",alpha=0.5, inherit.aes = FALSE)+
-    geom_rect(data=s2.rect.sameshift1, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "red",alpha=0.1, inherit.aes = FALSE)+
-    geom_rect(data=s2.rect.sameshift2, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "blue",alpha=0.1, inherit.aes = FALSE)+
-  theme_bw() + labs(y = "Signature Land Id = 9")
- ##########
+#### Failure Rate
+\vspace{-1.5em}
 
- d1.rect.opt <- data.frame(xmin=max_corr_opt_loc[1,2], xmax= max_corr_opt_loc[1,2]+window_opt, ymin=-Inf, ymax=Inf)
-    d1.rect.sameshift1 <- data.frame(xmin=max_corr_opt_loc[1,2]-2*window_opt, xmax= max_corr_opt_loc[1,2]-2*window_opt-window_val, ymin=-Inf, ymax=Inf)
-       d1.rect.sameshift2 <- data.frame(xmin=max_corr_opt_loc[1,2]+window_opt+ window_val, xmax= max_corr_opt_loc[1,2]+ window_opt+2*window_val, ymin=-Inf, ymax=Inf)
+\[
+\begin{split}
+P\left( t_1^{o} < w_v \ \bigcap \ t_2^{o} > T_2-w_o-w_v\right) + \\P\left( t_1^{o} < T_1- w_o - w_v \ \bigcap \ t_2^{o} < w_v\right).
+\end{split}
+\]
 
-  d2.rect.opt <- data.frame(xmin=max_corr_opt_loc[1,1], xmax= max_corr_opt_loc[1,1]-window_opt, ymin=-Inf, ymax=Inf)
-    d2.rect.sameshift1 <- data.frame(xmin=max_corr_opt_loc[1,1]-window_opt-window_val, xmax= max_corr_opt_loc[1,1]-window_opt- 2*window_val, ymin=-Inf, ymax=Inf)
-       d2.rect.sameshift2 <- data.frame(xmin=max_corr_opt_loc[1,1]+ 4*window_val, xmax= max_corr_opt_loc[1,1]+ 5*window_val, ymin=-Inf, ymax=Inf)
+\vspace{-1.5em}
 
-p3<-  ggplot(data = d1, aes(x = y, y = l30)) + geom_line() + 
-  geom_rect(data=d1.rect.opt, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "black",alpha=0.5, inherit.aes = FALSE)+
-    geom_rect(data=d1.rect.sameshift1, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "red",alpha=0.1, inherit.aes = FALSE)+
-    geom_rect(data=d1.rect.sameshift2, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "blue",alpha=0.1, inherit.aes = FALSE)+
-  theme_bw() + labs(y = "Signature Land Id = 1")
+#### Same-shift failure
 
- p4<-   ggplot(data = d2, aes(x = y, y = l30)) + geom_line() + 
-   geom_rect(data=d2.rect.opt, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "black",alpha=0.5, inherit.aes = FALSE)+
-    geom_rect(data=d2.rect.sameshift1, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "red",alpha=0.1, inherit.aes = FALSE)+
-    geom_rect(data=d2.rect.sameshift2, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "blue",alpha=0.1, inherit.aes = FALSE)+
-  theme_bw() + labs(y = "Signature Land Id = 9")
+\begin{figure}[hbtp]
+\centering
+\includegraphics[width=.7\textwidth]{imgs/same-shift-failure.png}
 
-multiplot(p1, p2, p3, p4, cols=2)
-```
+\caption{\label{same-shift-failure}Sketch of same-shift pairings  (top) when the lag is too large to accomodate a a vaildation window in either of the two signatures}
+\end{figure}
 
-For both same- and different-shift pairs correlations between the  markings are calculated. 
-The intuition here is that for two markings from the same source the correlation for the in-phase sample should be high, while the correlations of the out-of-phase sample  provide a measure for the base-level correlation for non-matching marks of a given length $w_v$.
-The Chumbley score is then computed as a Mann Whitney U statistic to compare between in-phase sample and out-of-phase sample.
-In the original method proposed in \citet{chumbley} both in-phase and out-of-phase sample are extracted randomly, whereas \citet{hadler} proposed the above specified deterministic rules for both samples to make the resulting score deterministic while simultaneously avoiding overlaps within selected marks to ensure independence.
+## 6
 
-## A problem with failures and modification
+- For different-source matches
+    - we can assume $t_1^o$ and $t_2^o$ are independent.
+    - Failure rate = $\frac{2 w_v^2}{(T_1-w_o)(T_2-w_o)}$ 
+    - For an average length of $T_i$ of 1200 pixels, $w_o = 120$ pixels and $w_v = 30$ pixels this probability is about 0.0015.
+- For same-source matches, 
+    - we expect a strong dependency between optimal locations $t_i^o$
+    - $t_1^o$ and $t_2^o$ are not independent of each other
+    - A large difference between locations is unlikely
+    - Failure Rate = $2 w_v / (T_i-w_o)$, or about 5.6\% for an average length of $T_i$ of 1200 pixels, $w_o = 120$ pixels and $w_v = 30$ pixels.
 
-Looking closer at \autoref{eq.start}, we see that by definition, some number of tests will fail to produce a result, either because the number of eligible same-shift pairs is zero, or the number of different-shift opairs is zero.
+##  Failure rate for same-source matches
+The failure rate for same-source matches can also be written as follows
+$$
+\begin{split}
+P\left( t_1^{o} < w_v \ \bigcup \ t_2^{o} > T_2-w_o-w_v\right) + \\P\left( t_1^{o} < T_1- w_o - w_v \ \bigcup \ t_2^{o} < w_v\right)
+\end{split}
+$$
 
-The number of same-shift pairs will be zero, if the optimal locations $t_1^{o}$ and $t_2^{o}$ are so far apart, that no segments of size $w_v$ are left on the same sides of the optimal locations, i.e.
-$t_1^{o} < w_v$ and $t_2^{o} > T_2-w_o-w_v$ or  $t_1^{o} < T_1- w_o - w_v$ and $t_2^{o} < w_v$, i.e.
-we have a failure rate of
-\[
-P\left( t_1^{o} < w_v \ \cap \ t_2^{o} > T_2-w_o-w_v\right) + P\left( t_1^{o} < T_1- w_o - w_v \ \cap \ t_2^{o} < w_v\right).
-\]
-While we can assume that once (sub-)class characteristics are removed, optimal locations $t_i^{o}$ are uniformly distributed across the length of the profile, we cannot assume that $t_1^o$ and $t_2^o$ are independent of each other.
-In particular, for same-source profiles, we would expect a strong dependency between these locations, in which case a large difference between locations is unlikely.
-However, for different source matches, we can assume that  locations are independent.
-In that case, we expect a test to fail with a probability of $\frac{2 w_v^2}{(T_1-w_o)(T_2-w_o)}$.
-For an average length of $T_i$ of 1200 pixels, $w_o = 120$ pixels and $w_v = 30$ pixels this probability is about 0.0015.
+Taking first part of the above equation, let $A = t_1^{o} < w_v$ and $B = t_2^{o} > T_2-w_o-w_v$
 
-The number of possible different-shift pairs also depends on the location of  the  optimal locations $t_1^o$ and $t_2^o$.
-Whenever the optimal locations are close to the boundaries, the number of possible pairings decreases and reaches zero, if $t_i^{(o)} < w_v$ or  $t_i^{(o)} > T_i-w_o- w_v$.
-Assuming a correlation between optimal locations $t_1^o$ and $t_2^o$ of close to one for same-source profiles, this results in an expected rate of failure of $2 w_v / (T_i-w_o)$, or about 5.6\% for an average length of $T_i$ of 1200 pixels, $w_o = 120$ pixels and $w_v = 30$ pixels.
-Assuming independence in the optimal locations for different source profiles the expected probability for a failed test is, again, $\frac{2 w_v^2}{(T_1-w_o)(T_2-w_o)}$.
+- Again we assume high correlation between optimal locations
+- In order to have at least one same-shift pair or different shift pair
+    - \textbf{Either}, if $x \in A$, then $B$ is empty and $P(A \cup B) = P (A)$
+    - \textbf{or}, if $x \in B$, then $A$ is empty and $P(A \cup B) = P(B)$
 
-While failures due to missing correlations from same-shift pairs are unavoidable by definition of the Chumbley score, failures due to missing correlations from different-shift pairs can be prevented by using a different strategy in assigning pairs.
+This would hold for the second part too, and would therefore give an expected rate of failure as $2 w_v / (T_i-w_o)$
 
-Using the same notation as in \autoref{eq.start}, we define same-shift pairs identical to \citet{hadler} as pairs $(s_i^{(1)}, s_i^{(2)})$ for all $i$ where the boundary conditions of both sequences are met simultaneously.
-Let us assume that this results in $I$ pairs.
-Define $s_{(j)}^{(k)}$ to be the $j$th starting location in sequence $k = 1, 2$, i.e.\ $s_{(1)}^{(k)} < s_{(2)}^{(k)} < ...
-< s_{(I)}^{(k)}$.
+## Proposed Modification
+
+- Failures due to missing Same-shift pairs unavoidable
+- Failures due to missing different-shift pairs preventable
+
+Define same-shift pairs identical to \citet{hadler} as pairs 
+$$(s_i^{(1)}, s_i^{(2)}) \quad \forall \, i \, \in \, \mathbb{Z}$$ where the boundary conditions of both sequences are met simultaneously. \newline
+
+- Let us assume that this results in $I$ pairs. \newline
+- Let $s_{(j)}^{(k)}$ to be the $j$th starting location in sequence $k = 1, 2$, i.e.\ $s_{(1)}^{(k)} < s_{(2)}^{(k)} < ... < s_{(I)}^{(k)}$.
+
+##
 
@@
 \end{equation}
-i.e.\ for an odd number of same-shift correlations, we skip the middle pair for the different-shift correlations (see \autoref{sketch-diff-2}).
-This pairing ensures that the number of different-shift pairings is the same or at most one less than the number of same-shift pairings in all tests.
-\hh{In the remainder of the paper, we will refer to the algorithm defined by \citet{hadler} as {\bf (CS1)} and the suggested modified algorithm as {\bf (CS2)} and compare their performance on the available scans of the Hamby study.}
+
+- For an odd number of same-shift correlations
+    - We skip the middle pair for the different-shift correlations (see \cref{sketch-diff-2}).
+- This pairing ensures that the number of different-shift pairings is the same or at most one less than the number of same-shift pairings in all tests.
+
+## 10
+
+<!-- Algorithm defined by \citet{hadler} \textbf{(CS1)} and the suggested modified algorithm as \textbf{(CS2)} -->
+
 \begin{figure}[hbtp]
 \centering
-\includegraphics[width=.7\textwidth]{images/sketch-diff-2.png}
+\includegraphics[width=.7\textwidth]{../doc/images/sketch-diff-2.png}
 \caption{\label{sketch-diff-2}Sketch of adjusted different-shift pairings. At most one of the same-shift pairings can not be matched with a different-shift pair. }
@@
 
-# Testing setup
+## Case where CS1 fails but CS2 does not fail
 
-## The Data
-
-```{r sigs-profiles, fig.width = 8, fig.height = 2.5, fig.cap="Bullet land profile (left) and the corresponding signature (right) for one of the lands of Hamby-44."}
-p1 <- sigs_graphics[[1]] %>% filter(land_id==19) %>%
-  ggplot(aes(x = y, y = value)) + geom_line() +
-  theme_bw() +
-  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
-  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
-  ggtitle("Profile")
-
-p2 <- sigs_graphics[[1]] %>% filter(land_id==19) %>%
-  ggplot(aes(x = y, y = l30)) + 
-  geom_line(aes(y=resid), size=.25) +
-  geom_line(aes(y=l30)) +
-  theme_bw() +
-  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
-  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
-  ggtitle("Signature")
-
-grid.arrange(p1,p2, ncol=2)
-```
-
-Lands for all Hamby-44 and Hamby-252 scans are made available through the NIST ballistics database \citep{nist} and are considered, here.
-Both of these sets of scans are part of the larger Hamby study \citep{hamby}.
-Each set consists of twenty known bullets (two each from ten consecutively rifled Ruger P85 barrels) and fifteen questioned bullets (each matching one of the ten barrels).
-Ground truth for both of these Hamby sets is known and was used to assess correctness of the tests results.
-
-Profiles for each bullet land were extracted from scans close to the heel of the bullet while avoiding break-off as described in \citet{aoas}.
-
-## Setup
-Both algorithms (CS1) and (CS2) are implemented in the R package `toolmaRk` \citep{toolmark}.
-We applied both methods to all pairwise land-to-land comparisons of the Hamby scans provided by NIST for a total of 85,491 land-to-land comparisons.
-
-# Results
-## Failed Tests
-
-Initially, the default settings suggested in \citet{hadler} were used: $w_o = 120$ pixels or about 190 $\mu m$ (ten percent of the average length of profiles) and coarseness $c = 0.25$. 
-\autoref{fig:fails} shows the percentage of failed tests among the 85,491 land-to-land comparisons of the NIST data for different values of the validation window size $w_v$.
-For same-source lands up to 12.5 percent of the tests fail using CS1.
-The highest percentage of failed tests under CS2 is 1.3% for different source tests using a validation window size $w_v$ of 60 pixels.
-\hh{Rates of expected failures are based on simulation runs using covariances between locations of same-source profiles of 0.854, and 0.120 for locations from different-source profiles, matching observed covariances for the Hamby scans.
-}\hh{ Observed failure rates are higher than expected.
-This might be due to remaining sub-class structure at a coarseness of 0.3 resulting in a distribution of optimal locations different from the assumed uniform.
-}
-
-```{r simulation, echo=FALSE}
-library(MASS) 
-Sigsame <- matrix(c(1, 0.85, 0.85, 1), nrow = 2) 
-Sigdiff <- matrix(c(1, 0.12, 0.12, 1), nrow = 2) 
-set.seed(20140501)
-simu <- 1:10 %>% purrr::map_df(.f = function(i) {
-  X1 <- mvrnorm(10000, mu = c(0,0), Sigma = Sigsame) 
-  X2 <- mvrnorm(10000, mu = c(0,0), Sigma = Sigdiff) 
-  Y1 <- pnorm(X1)  # Probability Integral Transform
-  Y2 <- pnorm(X2) 
-  wv <- 10*1:6
-
-  ps <- wv %>% purrr::map_df(.f = function(x) {
-    p <- x/(1200-120)
-    y1 <- Y1[,1]
-    y2 <- Y1[,2]
-    difffail1 <- sum((y1 < p & y2 < p) | (y1 > 1-p & y2 > 1-p))/length(y1)
-    samefail1 <- sum((y1 < p & y2 >1- p) | (y1 > 1-p & y2 < p))/length(y1)
-    y1 <- Y2[,1]
-    y2 <- Y2[,2]
-    difffail2 <- sum((y1 < p & y2 < p) | (y1 > 1-p & y2 > 1-p))/length(y1)
-    samefail2 <- sum((y1 < p & y2 > 1-p) | (y1 > 1-p & y2 < p))/length(y1)
-    data.frame(wv=x, samefail1=samefail1, samefail2=samefail2,
-               difffail1=difffail1, difffail2=difffail2)
-  })
-  ps$i <- i
-  ps
-})
-simu$`TRUE` <- simu$samefail1+simu$difffail1
-simu$`FALSE` <- simu$samefail2+simu$difffail2
-simulong <- simu %>% dplyr::select(wv, i, `TRUE`, `FALSE`) %>% 
-  gather(match, failed, 3:4)
-simusumm <- simulong %>% group_by(wv, match) %>%
-  summarize(
-    sdfailed = sd(failed),
-    failed = mean(failed)
-  )
-simusumm$method = "CS1"
-```
-
-```{r fails, echo=FALSE, fig.height = 3, fig.width = 7, out.width='0.7\\textwidth', fig.cap="Percent of failed land-to-land comparisons using an optimization window $w_o = 120$ and a coarseness of $c = 0.25$.
-With an increase in the size of the validation window  a higher percentage of tests fails under both methods (CS1) and (CS2), but the percentage of failed tests is much smaller under (CS2).
-Observed failure rates are higher than expected rates."}
-allinv$method <- "CS2"
-allp$method <- "CS1"
-
-allinv <-  allinv %>% mutate(
-  loc1 = chumbley %>% purrr::map_int(.f = function(x) x$locations[1]),
-  loc2 = chumbley %>% purrr::map_int(.f = function(x) x$locations[2])
-)
-
-prof_fails <- rbind(
-  allinv %>% dplyr::select(wo, wv, method, p_value, match),
-  allp %>% filter(coarse == 0.25) %>% 
-    dplyr::select(wo, wv, method, p_value, match)
-)
-fails <- prof_fails %>%
-  group_by(wo, wv, method, match) %>% summarize(
-  failed = sum(is.na(p_value)),
-  n=n()
-)
-
-fails %>% 
-  ggplot(aes(x = wv, y = failed/n*100, colour=factor(match), shape=factor(match))) +
-  geom_point(aes(size="Observed")) +
-  xlab("Size of validation window in pixels") +
-  ylab("Percent of failed tests") +
-  theme_bw() +
-  scale_colour_brewer("Same-source", palette="Set1") +
-  scale_shape_discrete("Same-source") +
-  ylim(c(0,13)) +
-  facet_wrap(~method, labeller="label_both") +
-  geom_point(aes(y = failed*100, size = "Expected"), data=simusumm) +
-  scale_size_manual("Failures", values=c(1.5,3))
-```
-
-## Coarseness
-The purpose of the coarseness parameter is to remove (sub-)class characteristics from profiles before comparisons for matching.
-\citet{hadler} suggest a coarseness parameter of 0.25 in the setting of toolmark comparisons.
-For bullet lands, coarseness might need to be adjusted because of the strong effect  bullet curvature has on profiles.
-
-\hh{\autoref{fig:profile-sketch} gives an overview of the effect of different coarseness parameters: from left to right, coarseness levels $c$ are varied in steps of 0.05 from 0.1 to 0.3.
-The top row shows  resulting signature after smoothing the profile shown in Figure~\ref{fig:sigs-profiles} with the coarseness specified.
-The histograms in the bottom row show the relative optimal location $t^*$.
-Optimal locations are distributed uniformly once (sub-)class characteristics are removed.
-However, for coarseness values of $c > 0.20$ we see quite distinct boundary effects: optimal locations $t^*$ are found at the very extreme ends of a profile more often than one would expect based on a uniform distribution.}
-
-```{r profile-sketch, fig.height = 4, fig.width='\\textwidth', fig.cap="Overview of the effect of different coarseness parameters $c$ on the profile shown in Figure \\ref{fig:sigs-profiles} (top).
-The bottom row shows histograms of  the (relative) optimal locations $t^o$ identified in the optimization step for different values of the coarseness parameter $c$.
-"}
-profile <- sigs_graphics[[1]] %>% filter(land_id==19)
-profile$lw10 <- lowess(x = profile$y, y = profile$value, f=.10)$y
-profile$lw15 <- lowess(x = profile$y, y = profile$value, f=.15)$y
-profile$lw20 <- lowess(x = profile$y, y = profile$value, f=.20)$y
-profile$lw25 <- lowess(x = profile$y, y = profile$value, f=.25)$y
-profile$lw30 <- lowess(x = profile$y, y = profile$value, f=.30)$y
-#profile$lw50 <- lowess(x = profile$y, y = profile$value, f=.5)$y
-#profile$lw75 <- lowess(x = profile$y, y = profile$value, f=.75)$y
-#profile$lw100 <- lowess(x = profile$y, y = profile$value, f=1)$y
-
-profiles <- tidyr::gather(profile, coarseness, values, lw10:lw30)
-profiles <- profiles %>% mutate(
-  c = readr::parse_number(coarseness)/100
-)
-p1 <- profiles %>% 
-  ggplot(aes(x = y, y = value-values)) + geom_line() +
-  theme_bw() +
-  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
-  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) +
-  facet_grid(~c, labeller="label_both")
-
-profc10 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo_160-wv_30-c_10.rds")
-profc15 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo_160-wv_30-c_15.rds")
-profc20 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo_160-wv_30-c_20.rds")
-profc20$coarseness <- 0.2
-profc25 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo_160-wv_30-c_25.rds")
-profc30 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo_160-wv_30.rds")
-profc30$coarseness <- .30
-profc <- bind_rows(profc10, profc15, profc20, profc25, profc30)
-profc <- profc %>% mutate(
-  locations = chumbley %>% purrr::map_int(.f = function(x) {
-    res <- NULL
-    try(res <- x$locations[1], silent = TRUE)
-    res
-  })
-)
-profsumm <- read.csv("../data/profiles-summary.csv")
-profsumm <- profsumm %>% dplyr::select(-run_id)
-profc <- profc %>% left_join(profsumm, by=c("land2_id"="land_id", "profile2_id"="profile_id"))
-
-p2 <- profc %>% 
-  ggplot(aes(x = locations/(length-wo))) +
-  geom_histogram(binwidth = 0.025) +
-  facet_grid(.~coarseness, labeller="label_both") + 
-  theme_bw() +
-  xlab(expression("Relative optimal location "~t^"*")) +
-  ylab("Number of profiles")
-
-grid.arrange(p1, p2, ncol=1)
-```
-
-## Type II error rates
-
-\autoref{fig:type2} gives an overview of ROC curves for methods CS1 and CS2 over a range of different optimization window sizes $w_o$.
-
-```{r type22, fig.height = 3.5, fig.width = 5, out.width='0.7\\textwidth', fig.cap="ROC curves of methods CS1 and CS2 for different sizes of optimization window $w_o$.
-Optimal ROC curves are reached at optimization windows of size 150 and higher.
-Points of equal error rates (EERs) can be found at the intersection of the dotted line and the ROC curves."}
-files <- dir("../data/crossvalidate-cs2/")
-cvcs2 <- data.frame()
-for (file in files) {
-  tmp <- readRDS(file.path("../data/crossvalidate-cs2/",file))
-  cvcs2 <- rbind(cvcs2, tmp)
-}
-cvcs2$coarse <- cvcs2$coarseness
-library(plotROC)
-cvcs2$method <- "CS2"
-
-cvcs2 %>% 
-  ggplot(aes(d = as.numeric(!match), m = p_value, colour=factor(wo))) + 
-  geom_abline(linetype=3, slope=-1, intercept = 1, colour="grey30", size=.5) +
-  geom_roc(labels=FALSE, n.cuts=100, pointsize=0) +
-  theme_bw() + 
-  scale_colour_brewer(
-    expression("Nominal\ntype I error"~alpha), palette="Set2") +
-  facet_grid(.~method, labeller="label_both") +
-  xlab("False Positives Rate (1 - Specificity)") +
-  ylab("True Positives Rate (Sensitivity)")
-
-```
-
-<!--
-## Type II error rate and coarseness
-\hh{Figure \ref{fig:coarse} shows the type II error rates for profiles using an optimization window $w_o= 120$ and a validation window $w_v = 30$ for varying level of coarseness.
-The type II error for all nominal levels of  $\alpha$ are the lowest for a coarseness range of 0.20 to 0.30, with a minimum in type II reached at a coarseness $c = 0.25$.
-The histograms suggest to use a value of coarseness below 0.2.}
-\hh{However, for an optimal value of coarseness other aspects have to be condsidered.}
-```{r coarse, fig.cap="Type II error with respect to coarseness parameter over profiles, $w_o = 120$, $w_v = 30$.
-Optimal values for coarseness are around $c \\approx 0.30$", fig.width=5, fig.height = 3, out.width='.6\\textwidth'}
-e1<-errors2 %>% filter(wv==30)%>% ggplot(aes(x = coarse, y=beta, colour=factor(alpha))) +
-  geom_point() + geom_line() +
-  theme_bw() +
-  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + xlab("Profile Coarseness")+
-  ylab("Type II error") 
-e1
-```
-
--->
-
-### Results for Signatures
-
-\hh{For signatures from NIST scans we see three problems: 
-\begin{enumerate}
-\item type-2 error rate is at best 30\% for a type-1 error rate of 5\%, which is well above the error rates we see for tool marks from screw drivers, see figure \ref{fig:type2};
-\item the observed type-1 error, which generally close to the nominal type-1 error rate, depends on the size of the optimization window: as the window size increases, the observed type-1 error decreases, see figure \ref{fig:type1};
-\item the Chumbley-score fails to provide a result for up to 3\% of the cases.
-The number of failed tests increases linearly in the size used for the window in the optimization step.
-The rate of failed tests is considerably higher when the two lands are from same source than when the lands are from different sources, see figure \ref{fig:missings}.
-\end{enumerate}
-}
-
-Figure \ref{fig:missings} gives an overview of the number of failed tests, i.e.
-tests in which a particular parameter setting did not return a valid result.
-This happens e.g.\ when the shift to align two markings is so large, that the remaining overlap is too small to accommodate  windows for validation.
-The problem is therefore exacerbated by a larger validation window.
-Figure \ref{fig:missings}(left) also shows that the number of failed tests is approximately linear in the size of the optimization window.
-Tests also fail at a higher rate than expected when the markings are from the same source (right).
-This difference is the least pronounced around an optimized window size $w_o$ of around 120.
-However, even in this scenario, the number of failed tests for markings from the same source is about twice as high as expected given the number of same source and different source pairings in the data set.
-
-<!--
-Following on similar lines to the setup of toolmarks, the first step here is to first identify what difference does different window sizes of optimization and the validation step have, when adapting the toolmark method to bullets.
-
-The marking made on bullets are smaller than toolmarks and is also less wider.
-The idea is to find out possible areas of error while adapting the score based method proposed for toolmarks, using cross-validation setup to identify appropriate parameter settings for (a) signatures and (b) profiles directly
-
-### Signatures
--->
-
-```{r type2, fig.width=8, fig.height=5, out.width='.8\\textwidth', fig.cap="Type II error rates observed across a range of window sizes for optimization $w_o$.
-For a window size of $w_o = 120$ we see a drop in type II error rate across all type I rates considered.
-Smaller validation sizes $w_v$ are typically associated with a smaller type II error."}
-errors %>% filter(wv %in% c(30,50)) %>%
-  ggplot(aes(x = wo, y = beta, colour=factor(alpha))) + 
-  geom_point(aes(shape=factor(wv)), size=3.5) +
-#  facet_grid(.~wo, labeller="label_both") + 
-  geom_smooth(aes(group=alpha), se=FALSE, method="loess", size=.7) +
-  theme_bw() +
-  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") +
-  scale_shape_discrete(expression("Size of\nvalidation window "~w[v])) +
-  xlab(expression("Window size for optimization "~w[o])) +
-  ylab("Type II error rate")
-
-labels = expression(P[M1](tilde(z)>0),P[M0](tilde(z)>0))
-```
-
-```{r type1, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap="Comparison of observed and nominal type I error rates  across a range of window sizes for optimization $wo$.
-The horizontal line in each facet indicates the nominal type I error rate."}
-errors %>% filter(wv %in% c(30,50)) %>%
-  ggplot(aes(x = wo, y = actual, colour=factor(alpha))) + 
-  geom_hline(aes(yintercept=alpha), colour="grey30", size=.5) +
-  geom_point(aes(shape=factor(wv)), size=2.5) +
-  #  facet_grid(.~wo, labeller="label_both") + 
-  geom_smooth(aes(group=alpha), size=.8, se=FALSE, method="lm", alpha=.9) +
-  theme_bw() +
-#  scale_y_log10(breaks=c(0.001,.005, 0.01, 0.05)) +
-  scale_shape_discrete(expression("Size of validation window "~w[v])) +
-  ylab("Observed type I error rate") +
-  xlab(expression("Window size for optimization, "~w[o])) +
-  scale_colour_brewer(expression("Nominal type I error "~alpha), palette="Set2") +
-  facet_wrap(~alpha, labeller="label_both", scales="free") +
-  theme(legend.position="bottom")
-
-```
-
-```{r missings, out.width='\\textwidth', fig.width=8, fig.height=4, fig.cap="The number of failed tests increases with an increase in the size of the optimization window (left).
-Unfortunately there is also a dependency between failed tests and ground truth.
-The plot on the right shows the ratio of the number of land pairs from same sources and different sources for failed tests.
-For small optimization windows and large windows the number of failed tests for same-source land-to-land comparisons is increasing.
-Even in the minimum, same-source land-to-land comparisons fail at twice the rate that they are expected to based on the ratio of the number of known matches and known non-matches (horizontal line).",fig.align = "center"}
-#ns1<- ns[which(ns$wv == c(30,50), arr.ind = T),]
-
-fails <- all %>% group_by(wv, wo) %>% summarize(
-  miss = sum(is.na(p_value)),
-  missperc = miss/n()*100
-)
-
-failsbymatch <- all %>% group_by(wv, wo, match) %>% summarize(
-  fails = sum(is.na(p_value))
-) %>% mutate(
-    ratio = fails[match==TRUE]/fails[match==FALSE],
-    match = c("Different Source", "Same Source")[as.numeric(as.logical(match)) + 1]
-    )
-
-missings <- fails %>% filter(wv %in% c(30, 50)) %>%
-  ggplot(aes(x = wo, y = missperc, 
-             colour = factor(wv),
-             shape=factor(wv)), size=.3) + 
-  geom_smooth(method="lm", size=.5, se=FALSE) +
-  geom_point(size=2.5) + 
-  theme_bw() +
-  ylab("Percent of failed tests.") +
-  xlab(expression("Size of optimization window "~w[o])) +
-  scale_colour_brewer(expression("Size of validation window "~w[v]), palette="Set1") +
-  scale_shape(expression("Size of validation window "~w[v])) +
-    theme(legend.position="bottom")
-
-missingsbymatch <- failsbymatch %>% 
-  filter(wv %in% c(30, 50),
-         match=="Same Source") %>%
-  ggplot(aes(x = wo, y = ratio, 
-             colour = factor(wv),
-             shape=factor(wv)), size=.3) + 
-  geom_hline(yintercept = 43960/2948225, size=0.5, colour="grey30") +
-  geom_smooth(method="loess", span=0.8, size=.5, se=FALSE) +
-  geom_point(size=2.5) + 
-  theme_bw() +
-  ylab("Ratio of same/different source\namong failed tests") +
-  xlab(expression("Size of optimization window "~w[o])) +
-  scale_colour_brewer(expression("Size of validation window "~w[v]), palette="Set1") +
-  scale_shape(expression("Size of validation window "~w[v])) +
-    theme(legend.position="bottom") +
-  scale_y_continuous(breaks=c(0.015, 0.03, 0.06, 0.09))
-
-gridExtra::grid.arrange(missings, missingsbymatch, ncol=2)
-
-```
-
-<!--
-Signatures of lands for all Hamby-44 and Hamby-252 scans made available through the NIST ballistics database \citep{nist} were considered.
-Both of these sets of scans are part of the larger Hamby study \citep{hamby} and each consist of twenty known bullets (two each from ten consecutively rifled Ruger P85 barrels) and fifteen questioned bullets (each matching one of the ten barrels).
-Ground truth for both of these Hamby sets is known and was used to assess correctness of the tests results.
-
-Bullet signatures being compared at this time are therefore from the Hamby 44 and Hamby 252 data.
-The database setup and pre-processing system used for choosing the Bullet signatures are as described by \citet{aoas}.
-In order to choose the bullet signatures we first filter out Land_id for Profiles from the Hamby 44 and Hamby 252 data and remove all NA values.
-Then run_id = 3 is chosen as the signatures generated from this run_id give the closest match.
-Different run_id's have some different settings for generating the signatures.(The level of smoothing does not seem to be one of them)
-
-The bullet signatures when generated by this process already includes a loess smoothing.
-Therefore, the coarseness factor is set to 1 while running the chumbley non random algorithm for comparing different optimization windows.The algorithm generates the same_shift, different_shift, U-Stat and P_value parameters which are then used to calculate the errors associated with different sets of window sizes.
-
-### Profiles
-
-The profiles are cross-sectional values of the the bullet striation mark which are chosen at an optimum height (x as used by \citet{aoas}).
-This x or height is not a randomly chosen level.
-The rationale behind the choice has been explained by \citet{aoas}.
-A region is first chosen where the cross-correlation seems to change very less and in this region an optimum height is chosen.
-The profiles generally resemble a curve which is more or less similar to a quadratic curve (a quadratic fit to the raw data values of the profile is not an exact fit but it does show a similar trend).
-Profiles are the set of raw values representing the striation marks, and signatures are generated from these by removal of the inherent curvature and applying some smoothing (the signatures generated by \citet{aoas} use a loess function for smoothing).
-
-Similar to signatures the run_id = 3 was used when applying the chumbley algorithm using the database setup given by \citet{aoas} of Hamby-44 and Hamby-252 datasets, on the profiles.
-The run_id not only defines the level of smoothing but also signifies the chosen height at which the profiles were selected initially.
-Another important aspect is the range of horizontal values (which is referred to as the y values in \citet{aoas}) in the signatures.
-These have already been pre-processed in the database to not include any grooves.
-
-Therefore for the sake of comparison the run_id = 3 is still chosen so as to ensure that the horizontal values remain the same as that of the signatures.
-This also gives us profiles with the grooves removed.
-
-The idea therefore is to first use these raw values of the profile directly in the chumbley algorithm, and see how the algorithm performs for different coarseness values (smoothing parameter as referred in the function lowess used in the chumbley algorithm).
--->
-
-```{r, include =F}
-wo120_sig <- errors %>% filter(wo==120, wv %in% c(30, 50))
-wo120_pr <- errors2 %>% filter(wo==120, wv %in% c(30, 50), coarse == .25)
-
-dt <- rbind(wo120_sig, wo120_pr) %>% 
-  ungroup() %>% 
-  mutate(beta = round(100*beta, 1)) %>%
-  dplyr::select(wv, type, alpha, beta) %>% 
-  spread(alpha, beta)
-print(xtable(dt), include.rownames=FALSE)
-```
-
-\begin{table}[ht]
-\caption{\label{tab:type2} Type II error rates for profiles and signatures of bullet lands. For profiles, a coarseness value of $c = 0.25$ is used to remove bullet curvature.}
+\small{\textbf{CS1} \citet{hadler}} algorithm \newline  \small{\textbf{CS2} the suggested modified algorithm}
+\begin{figure}[hbtp]
 \centering
-\begin{tabular}{rlrrrr}
-  \hline
-validation && \multicolumn{4}{l}{Nominal type I error rate $\alpha$}\\  
-window $w_v$ & source & 0.001 & 0.005 & 0.01 & 0.05 \\ 
-  \hline
- 30 & profiles & 54.40 & 43.80 & 39.70 & 30.00 \\ 
-   30 & signatures & 55.00 & 45.40 & 41.40 & 31.10 \\ \hline
-   50 & profiles & 58.50 & 44.40 & 40.70 & 28.70 \\ 
-   50 & signatures & 62.60 & 49.60 & 44.20 & 30.50 \\ 
-   \hline
-\end{tabular}
-\end{table}
-
-# Conclusion
-
-The results suggest that the Nominal type I error $\alpha$ value shows dependence on the size of the window of optimization.
-For a given window of optimization the actual Type I error is comparable to the nominal level for only a select few validation window sizes and for comparable validation window sizes of 30 and 50 as done here, the actual type I error does not seem to vary as much as it varies with the optimization window sizes .
-A Test Fail, i.e.
-tests in which a particular parameter setting did not return a valid result, happens, when the shift to align two signatures is so large, that the remaining overlap is too small to accommodate  windows for validation, depends on whether known-match or known non-matches has predictive value, with test results from different sources having a much higher chance to fail.
-On conducting an analysis of all known bullet lands using the adjusted chumbley algorithm, Type II error was identified to be least bad for window of validation 30 and window of optimization 120.
-In case of unsmoothed raw marks (profiles), Type II error increases with the amount of smoothing and least for LOWESS smoothing coarseness value about 0.25 or 0.3.
-In an effort to identify the level of adaptiveness of the algorithm, comparisons were made between signatures and profiles.
-Their comparison with respect to validation window size for a fixed optimization window size suggested that, profiles have a total error (i.e all incorrect classification of known-matches and known non-matches) greater than or equal to the total error of signatures for all sizes of validation window.
-Profiles also fail more number of times than signatures in a test fail (for different coarseness keeping windows fixed and also for different validation windows keeping coarseness fixed) which lets us conclude that the behaviour of the algorithm for the profiles instead of pre-processed signatures is not better.
-Finally it should be noted that the current version of the adjusted chumbley algorithm seems to fall short when compared to other machine-learning based methods \citet{aoas}, and some level of modification to the deterministic algorithm needs to be identified and tested that would reduce the number of incorrect classifications.
-
-```{r wo120-profile, fig.width = 6, fig.height=4, fig.cap="The figure shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120", eval = FALSE}
-wo120_profile <- errors2 %>% filter(wo==120, coarse == 0.25) %>% 
-  #gather(type, error, actual:beta) %>% 
-  gather(type, error, beta) %>% 
-  ggplot(aes(x = wv, y = error, colour = factor(alpha)))  + 
-  geom_line() + geom_point() + facet_wrap(~type, scales="free") +
-    theme_bw()+ 
-  theme(legend.position="bottom") + ggtitle("Profile" , subtitle = "Varying validation window sizes, Optimization window = 120, Coarseness = 0.25")
-
-#combined120 <- full_join(errors2 %>% filter(wo==120, coarse == 0.25), wo120 %>%
-
-wo120_1<- wo120_profile %>% gather(type, error, beta) %>%  
-  ggplot(aes(x = wv, y = error, colour = factor(alpha)))+
-  geom_line(linetype=5)+  geom_point(shape= 24) + facet_wrap(~type, scales="free") +
-  theme_bw()+ 
-  theme(legend.position="bottom") + ggtitle("Signatures" , subtitle = "Varying validation window sizes, Optimization window = 120")
-
-combined120<- 
-  ggplot(data = wo120_profile %>% gather(type, error, beta), aes(x = wv, y = error, colour = factor(alpha)))+
-  geom_line(aes(linetype="Signatures"))+  geom_point(shape= 24) + 
-  geom_line(data =  errors2 %>% filter(wo==120, coarse == 0.25) %>% 
-  gather(type, error, beta), aes(x = wv, y = error, colour = factor(alpha), linetype = "Profiles")) +       geom_point(data =  errors2 %>% filter(wo==120, coarse == 0.25) %>% 
-  gather(type, error, beta)) +
-  facet_wrap(~type, scales="free") +
-  scale_linetype_manual("Lines",values=c("Signatures"=2,"Profiles"=1))+
-  guides(fill = guide_legend(keywidth = 1, keyheight = 1),
-    linetype=guide_legend(keywidth = 3, keyheight = 1),
-    colour=guide_legend(keywidth = 3, keyheight = 1))+
-  theme_bw()+ 
-  theme(legend.position="right") + ggtitle("Type II error with Validation window" , subtitle = "Profiles at coarseness 0.25") + xlab("Type II error")
-
-combined120
-#wo120_profile
-#wo120_1
-#multiplot(wo120_1, wo120_profile1)
-```
-```{r failtest-coarseness, fig.width = 6, fig.height=4, fig.cap="The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120", eval= FALSE}
-
-miss_1
-```
-```{r failtest-wv, fig.width = 6, fig.height=4, fig.cap="The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120", eval= FALSE}
-
-miss_2
-```
-```{r failtest-wo, fig.width = 6, fig.height=4, fig.cap="The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120", eval= FALSE}
-
-miss_2
-```
-
-<!-- # Algorithm Modification -->
-
-<!-- The proposed algorithm modification is at the same shift step which comes after the shift distance is identified by finding two windows in the two markings that have the highest correlation. -->
-
-<!-- The modification allows for a "wiggle" room in the second marking of the two sets of markings being compared.
-This means that each set of same shift windows that are being compared, the second marking will window that is under consideration is allowd to move a little towards the left and a little towards the right.
--->
-
-<!-- This gives us a new set of 4 windows, 2 to the left and 2 to the right of the original comparison window.
-Then the correlations for each one of these 5 windows with the window under consideration of the 1st marking is retrieved (from the validation step correlation matrix) or computed.
--->
-
-<!-- Then the window that has the maximum correlation (from the set of 5 windows in the 2nd marking) with the window of the 1st marking is chosen. -->
-
-<!-- This new maximum correlation is then used to compute the U-statistics using the same method as before. -->
-
-<!-- ## Expected advantage of modification -->
-
-<!-- We are expecting that this modification would improve the type I and II errors for the good. Less number of false negatives and false positives. -->
-<!-- The reason for this expectation is many a times the bullet markings are not made at the same distances for two or more bullets because of the way it comes out of the barrel. -->
-
-<!-- Therefore rigid same shifts might not necessarily compare the right set of windows.
-In the same shift step for one set of windows, allowing the window in the 2nd marking to wiggle left and right and finding the best match to the window in the 1st marking, lets us adjust for situations where markings are compressed or elongated.
--->
-
-<!-- ## Dependence on the delta of new windows from the original same shift window -->
-
-<!-- From initial tests the amount of movement to get new windows left and right of the original same shift window seems to directly influence the -->
-<!-- the pvalue and U statistic that we look at.  -->
-
-<!-- ### other questions -->
-<!-- How much movement is reasonable and should be used?
-Is there a way to understand if the 2nd marking in comparison to the 1st marking is compressed or elongated or neither?
-Should the amount of movement selected depend on this compression or elongation?
-or using a static delta movement value justifieable.
--->
-<!--
-# Appendix
-\begin{appendix}
-
-On the other hand Figure \ref{fig:prof_missings} (b) shows if the coarseness level set in the chumbley agorithm has any effect on the signatures, which are pre-processed and already smoothed to a certain extent.
-From Figure \ref{fig:prof_missings} (b) we can notice that for different nominal $\alpha$ levels, the type II error fluctuates slightly but does not change much, thereby helping us conclude that the coarseness levels set in the LOWESS smoothing in the chumbley alggorithm does effect the type II error much for signatures.
-
-\subsection{Comparison of profiles and signatures}
-
-Another reason for failed tests can be incorrect identification of maximum correlation windows in the optimization step as seen in figure \ref{fig:prof_missings}(d) because of the level of smoothing, as too much smoothing would subdue intricate features that might otherwise help in the correlation calculations and correct identification of maximum correlation windows irrespective of the size.
-This would again cause a simiar effect as explained for figure \ref{fig:missings} with validation windows, irrespective of size, during the shifts end up at the ends of the markings resulting in an invalid calculation and failed comparison attempt.
-
-In figure \ref{fig:prof_missings}(d) and (f), we compare profiles and signatures on the basis of number of failed tests.
-The profiles chosen for figure \ref{fig:prof_missings}(f) have a constant coarseness of 0.25 and window of optimization as 120.
-The signatures in this case are not smoothed using the chumbley algorithm step of LOWESS smoothing.
-Instead signatures are used as calculated by \citet{aoas}.
-The smoothing in these signatures were determined and fixed on the basis of their performance in the random forest based algorithm proposed by \citet{aoas}.
-The comparison of profiles and signatures with variation of validation window size therefore is made on even footing.
-The trends are similar to figure \ref{fig:missings} in the sense that for known non-matches the number of failed tests are more for both signatures and profiles and increasing linearly with the validation window size.
-The problem is however, worse for profiles which has higher number of failed tests than signatures for all validation windows.
-
-The total error for different validation window sizes for signatures and profiles can be seen in figure \ref{fig:prof_missings} (e).The optimization window size is 120 and profiles are calculated at a default 0.25 coarseness level while signatures as before are not smoothed again in the modified chumbley algorithm.
-We can see that the total error is always higher for profiles as compared to signatures for all sizes of validation window.
-
-```{r, fig.cap="Type II errors for different levels of coarseness."}
-e1<-errors2 %>% filter(wv==30)%>% ggplot(aes(x = coarse, y=beta, colour=factor(alpha))) +
-  geom_point() + geom_line() +
-  theme_bw() +
-  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + xlab("Profile Coarseness")+
-  ylab("Type II error") + ggtitle(label = "(a)")
-
-e2<- errors_sig_c %>% filter(wv==30)%>% ggplot(aes(x = coarse, y=beta, colour=factor(alpha))) +
-  geom_point() + geom_line() +
-  theme_bw() +
-  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2")+ xlab("Signature Coarseness") + ylab("Type II error") + ggtitle(label = "(b)")
-
-#multiplot(e1,e2, cols = 1)
-
-```
-
-```{r, eval=FALSE}
-#errors %>% ggplot(aes(x = wv, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + facet_grid(~wo) +geom_point(data=errors2, aes(shape="profiles"))
-
-total1<- errors_sig_c %>% ggplot(aes(x = coarse, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + geom_line()  + geom_point(data= errors2 %>% filter(wv==30), aes(shape="profiles")) + geom_line(data=errors2 %>% filter(wv==30), aes(shape="profiles")) + theme_bw() +
-  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + xlab("Coarseness") +ylab("Total error") + ggtitle(label = "(c)")
-
-total2<- wo120_profile %>% filter(wv>10) %>% ggplot(aes(x = wv, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + geom_line() +geom_point(data=errors2 %>% filter(wv>10) %>% filter(coarse==0.25), aes(shape="profiles")) + geom_line(data=errors2 %>% filter(wv>10)%>% filter(coarse==0.25), aes(shape="profiles")) + theme_bw() +
-  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + ylab("Total error") + ggtitle(label = "(e)")
-
-```
-
-```{r prof_missings, fig.width=10, fig.height=10, fig.cap="Row 3:  Total error and Number of failed tests by the window validation size, wv, and ground truth, Row 2: Total error and Number of failed tests with Coarseness for both profiles and signatures, Row 1: Type II error for different coarseness levels as used in the modified chumbley algorithm for profiles and signatures"}
-
-#ns1<- ns[which(ns$wv == c(30,50), arr.ind = T),]
-miss_1<- nsp %>% filter(wv == 30) %>%
-#ns %>% filter(wv == c(30, 50)) %>%
-  ggplot(aes(x = coarse, y = missperc, shape= "profile", colour = factor(match))) + 
-  #geom_smooth(method="auto", size=.5, se=FALSE) +
-  geom_point(size = 2) + 
-  theme_bw() +
-  ylab("Percent of failed tests.") +
-  xlab("Coarseness") +
-  scale_colour_brewer("Match", palette="Set1") + geom_point(data=ns_sig_c  %>% filter(wv %in% c(30,50)) , aes(shape="signature"), size = 2) + ggtitle(label = "(d)")
-# + scale_colour_brewer("Match", palette="Set2") #+geom_smooth(method="auto", size=.5, se=FALSE) 
-  #scale_shape_discrete("Size of \nvalidation\nwindow, #wv")
-
-miss_2<- nsp %>% filter(coarse == 0.25) %>%
-#ns %>% filter(wv == c(30, 50)) %>%
-  ggplot(aes(x = wv, y = missperc, shape= "profile", colour = factor(match))) + 
-  #geom_smooth(method="auto", size=.5, se=FALSE) +
-  geom_point(size = 2) + 
-  theme_bw() +
-  ylab("Percent of failed tests.") +
-  xlab("wv") +
-  scale_colour_brewer("Match", palette="Set1") + geom_point(data=ns %>% filter(wo == 120)  %>% filter(coarse == 1) , aes(shape="signature"), size = 2) + ggtitle(label = "(f)")#+ scale_colour_brewer("Match", palette="Set2")
-
-#multiplot(plotlist = list(total1, total2, miss_1, miss_2, e1, e2), cols = 2)#, matrix(c(1,2,3,3), nrow=2, byrow=TRUE))
-multiplot(plotlist = list(e1, total1, total2,e2, miss_1, miss_2), cols = 2)
-```
-\end{appendix} 
--->
+\includegraphics[width=.7\textwidth]{imgs/CS1-failure-CS2-success.png}
+\caption{\label{CS1-failure-CS2-success}Sketch of a case where CS1 fails but CS2 does not fail}
+\end{figure}

### 2018-04-23 doc/paper.Rmd
fiddling

@@
 
-\autoref{fig:type22} gives an overview of ROC curves for methods CS1 and CS2 over a range of different optimization window sizes $w_o$.
-
-```{r type22, fig.height = 3, fig.width = 4.5, out.width='0.6\\textwidth', fig.cap="ROC curves of methods CS1 and CS2 for different sizes of optimization window $w_o$.
+\autoref{fig:roc} gives an overview of ROC curves for methods CS1 and CS2 over a range of different optimization window sizes $w_o$.
+
+```{r roc, fig.height = 3.25, fig.width = 4.75, out.width='0.6\\textwidth', fig.cap="ROC curves of methods CS1 and CS2 for different sizes of optimization window $w_o$.
 Optimal ROC curves are reached at optimization windows of size 150 and higher.

### 2018-04-23 doc/paper.tex
fiddling

@@
 
-{\centering \includegraphics[width=.8\textwidth]{figures/type2-1}
+{\centering \includegraphics[width=0.6\textwidth]{figures/type2-1}
 
@@
 \caption{Type II error rates observed across a range of window sizes for optimization $w_o$.
-For a window size of $w_o = 120$ we see a drop in type II error rate across all type I rates considered.
+For a window size of $w_o = 130$ we see a minimum in type II error rate across all type I rates considered.
 Smaller validation sizes $w_v$ are typically associated with a smaller type II error.}\label{fig:type2}
@@
 
-\autoref{fig:type22} gives an overview of ROC curves for methods CS1 and
+\autoref{fig:roc} gives an overview of ROC curves for methods CS1 and
 CS2 over a range of different optimization window sizes \(w_o\).
@@
 
-{\centering \includegraphics[width=0.7\textwidth]{figures/type22-1}
+{\centering \includegraphics[width=0.6\textwidth]{figures/roc-1}
 
@@
 Optimal ROC curves are reached at optimization windows of size 150 and higher.
-Points of equal error rates (EERs) can be found at the intersection of the dotted line and the ROC curves.}\label{fig:type22}
+Points of equal error rates (EERs) can be found at the intersection of the dotted line and the ROC curves.}\label{fig:roc}
 \end{figure}

### 2018-04-23 doc/paper.Rmd
more crossvalidation

@@
   geom_point(aes(shape=factor(wv)), size=2.5) +
-  geom_line(aes(group=alpha),  size=.7) +
+  geom_smooth(aes(group=alpha), se=FALSE, method="loess", span=1,  size=.7) +
   theme_bw() +
@@
 cvcs2 %>% 
-  ggplot(aes(d = as.numeric(!match), m = p_value, colour=factor(wo))) + 
+  ggplot(aes(d = as.numeric(!match), m = p_value, colour=factor(wo), 
+             shape=factor(wv))) + 
   geom_roc(labels=FALSE, n.cuts=100, pointsize=0, size=0.6) +
@@
     expression("Optimization\nwindow size"~w[o]), palette="Set2") +
-  facet_grid(.~method, labeller="label_both") +
+  facet_grid(wv~method, labeller="label_both") +
   xlab("False Positives Rate (1 - Specificity)") +

### 2018-04-23 doc/paper.Rmd
method included

@@
   xlab(expression("Window size for optimization "~w[o])) +
-  ylab("Type II error rate")
+  ylab("Type II error rate") +
+  facet_wrap(~method, labeller="label_both")
 

### 2018-04-23 doc/paper.Rmd
cv cs2 data

@@
 Looking closer at \autoref{eq.start}, we see that by definition, some number of tests will fail to produce a result, either because the number of eligible same-shift pairs is zero, or the number of different-shift opairs is zero.
+\hh{XXX The problem of failed tests is first mentioned in XXX}
 
@@
 Smaller validation sizes $w_v$ are typically associated with a smaller type II error."}
+if (!file.exists("../data/cvcs2.RDS")) {
 files <- dir("../data/crossvalidate-cs2/")
@@
 cvcs2$coarse <- cvcs2$coarseness
+cvcs2$method <- "CS2"
+saveRDS(cvcs2, file = "../data/cvcs2.RDS")
+} else cvcs2 <- readRDS("../data/cvcs2.RDS")
+
 library(plotROC)
-cvcs2$method <- "CS2"
-
 errorsCV <- rbind(errorrate(cvcs2, 0.001), 
@@
                 errorrate(cvcs2, 0.05))
-errorsCV$type <- "CS2"
+errorsCV$method <- "CS2"
 
@@
 
+## Observed versus Nominal Type I error rates
+
+\autoref{fig:type1} shows observed type I error rates across a range of optimization windows $w_o$.
+Generally, observed type I errors are higher than expected, but with an increase in the size of the optimization window observed type I error rates decrease.
+
+```{r type1, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap="Comparison of observed and nominal type I error rates  across a range of window sizes for optimization $wo$.
+The horizontal line in each facet indicates the nominal type I error rate."}
+errorsCV %>% filter(wv %in% c(30,50)) %>%
+  ggplot(aes(x = wo, y = actual, colour=factor(alpha))) + 
+  geom_hline(aes(yintercept=alpha), colour="grey30", size=.5) +
+  geom_point(aes(shape=factor(wv)), size=2.5) +
+  #  facet_grid(.~wo, labeller="label_both") + 
+  geom_smooth(aes(group=alpha), size=.8, se=FALSE, method="lm", alpha=.9) +
+  theme_bw() +
+#  scale_y_log10(breaks=c(0.001,.005, 0.01, 0.05)) +
+  scale_shape_discrete(expression("Size of validation window "~w[v])) +
+  ylab("Observed type I error rate") +
+  xlab(expression("Window size for optimization, "~w[o])) +
+  scale_colour_brewer(expression("Nominal type I error "~alpha), palette="Set2") +
+  facet_wrap(~alpha, labeller="label_both", scales="free") +
+  theme(legend.position="bottom")
+
+```
+
+<!--
 ### Results for Signatures
@@
 However, even in this scenario, the number of failed tests for markings from the same source is about twice as high as expected given the number of same source and different source pairings in the data set.
+-->
 
@@
 
-```{r type1, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap="Comparison of observed and nominal type I error rates  across a range of window sizes for optimization $wo$.
-The horizontal line in each facet indicates the nominal type I error rate."}
-errors %>% filter(wv %in% c(30,50)) %>%
-  ggplot(aes(x = wo, y = actual, colour=factor(alpha))) + 
-  geom_hline(aes(yintercept=alpha), colour="grey30", size=.5) +
-  geom_point(aes(shape=factor(wv)), size=2.5) +
-  #  facet_grid(.~wo, labeller="label_both") + 
-  geom_smooth(aes(group=alpha), size=.8, se=FALSE, method="lm", alpha=.9) +
-  theme_bw() +
-#  scale_y_log10(breaks=c(0.001,.005, 0.01, 0.05)) +
-  scale_shape_discrete(expression("Size of validation window "~w[v])) +
-  ylab("Observed type I error rate") +
-  xlab(expression("Window size for optimization, "~w[o])) +
-  scale_colour_brewer(expression("Nominal type I error "~alpha), palette="Set2") +
-  facet_wrap(~alpha, labeller="label_both", scales="free") +
-  theme(legend.position="bottom")
-
-```
-
+<!--
 ```{r missings, out.width='\\textwidth', fig.width=8, fig.height=4, fig.cap="The number of failed tests increases with an increase in the size of the optimization window (left).
@@
 XXX
-still to cover: introduce method CS2;
-failures are down; 
-type 2 error rates are also down.
-Fix has type 2 error rates on bullet lands that is still higher than the error rates achieved on the -much larger- toolmarks.
+The algorithm CS2 introduced here is constructed in a way that reduces the number of failures by a magnitude of XXX.
+While reducing the failure rate, the algorithm also shows an increase in the power of the test.
+Type II error rates reach a minimum of XXX for an optimized window size of YYY.
+While significantly reduced over CS1, CS2 still has type 2 error rates on bullet lands that are  higher than the error rates achieved on the -much larger- toolmarks.
 However, bullets usually have multiple lands - in the case of Ruger P85s as used in the Hamby study, there are six lands for each bullet.
 We might be able to get more power out of the test by adapting CS2 to deal work in a bullet-to-bullet comparison.
+}
+
+\hh{
+Fix in CS2 will also improve power for matching toolmarks. 
 }

### 2018-04-23 doc/paper.tex
cv cs2 data

@@
 opairs is zero.
+\hh{XXX The problem of failed tests is first mentioned in XXX}
 
@@
 
-\subsubsection{Results for Signatures}\label{results-for-signatures}
+\subsection{Observed versus Nominal Type I error
+rates}\label{observed-versus-nominal-type-i-error-rates}
 
-\hh{For signatures from NIST scans we see three problems: 
-\begin{enumerate}
-\item type-2 error rate is at best 30\% for a type-1 error rate of 5\%, which is well above the error rates we see for tool marks from screw drivers, see figure \ref{fig:type2};
-\item the observed type-1 error, which generally close to the nominal type-1 error rate, depends on the size of the optimization window: as the window size increases, the observed type-1 error decreases, see figure \ref{fig:type1};
-\item the Chumbley-score fails to provide a result for up to 3\% of the cases.
-The number of failed tests increases linearly in the size used for the window in the optimization step.
-The rate of failed tests is considerably higher when the two lands are from same source than when the lands are from different sources, see figure \ref{fig:missings}.
-\end{enumerate}
-}
-
-Figure \ref{fig:missings} gives an overview of the number of failed
-tests, i.e.~tests in which a particular parameter setting did not return
-a valid result. This happens e.g.~when the shift to align two markings
-is so large, that the remaining overlap is too small to accommodate
-windows for validation. The problem is therefore exacerbated by a larger
-validation window. Figure \ref{fig:missings}(left) also shows that the
-number of failed tests is approximately linear in the size of the
-optimization window. Tests also fail at a higher rate than expected when
-the markings are from the same source (right). This difference is the
-least pronounced around an optimized window size \(w_o\) of around 120.
-However, even in this scenario, the number of failed tests for markings
-from the same source is about twice as high as expected given the number
-of same source and different source pairings in the data set.
+\autoref{fig:type1} shows observed type I error rates across a range of
+optimization windows \(w_o\). Generally, observed type I errors are
+higher than expected, but with an increase in the size of the
+optimization window observed type I error rates decrease.
 
@@
 The horizontal line in each facet indicates the nominal type I error rate.}\label{fig:type1}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figures/missings-1}
-
-}
-
-\caption{The number of failed tests increases with an increase in the size of the optimization window (left).
-Unfortunately there is also a dependency between failed tests and ground truth.
-The plot on the right shows the ratio of the number of land pairs from same sources and different sources for failed tests.
-For small optimization windows and large windows the number of failed tests for same-source land-to-land comparisons is increasing.
-Even in the minimum, same-source land-to-land comparisons fail at twice the rate that they are expected to based on the ratio of the number of known matches and known non-matches (horizontal line).}\label{fig:missings}
 \end{figure}

### 2018-04-24 presentation/presentation.Rmd
presentation modifications

@@
 
-## Data
+## Variations of Chumbley score method and Error Rates for toolmarks
+
+\begin{table}[ht]
+\centering
+\resizebox{\textwidth}{!}{\begin{tabular}{rlrrr}
+  \hline
+Research paper & Method & Data Source & False Positives & False Negatives\\
+  \hline
+ \textbf{\citet{manytoolmarks1}} & Maximized & Screwdrivers & & \\ & Correlation &  & - & - \\ \hline
+ \textbf{\citet{chumbley}} & Randomized & Screwdrivers & & \\ (Same-Surface Same-Angle) & Chumbley Score & & 2.3\% & 8.9\%  \\ \hline
+  \textbf{\citet{afte-chumbley}} & Randomized & Slip-joint  &  & \\ & Chumbley Score &  & - & - \\ \hline 
+  \textbf{\citet{hadler}} & Deterministic & Screwdrivers & & \\ (Same-Surface Same-Angle) & Chumbley Score & & 0\% & 6\%  \\ 
+   \hline
+\end{tabular}}
+\caption{\label{tab:toolmarks-ER} Error Rates for Toolmarks using  variations of the chumbley score method}
+\end{table}
+
+ <!-- \textbf{\citet{manytoolmarks2}} & Similarity Measure & Screwdrivers & & \\ (Different Surfaces-same angle) & Relative Distance Metric & & 5.9\% &  9.4\% \\ (Same Surfaces-same angle)&  & & 0.22\% &  0\% \\ -->
+
+## Digitized Striation Marks
+- Data
     -  Ruger P85s Bullet Lands, or Hamby scans (\cite{hamby}) provided by NIST  (85,491 comparisons)
--  Bullet striation marks $\approx$ 2mm | Screwdriver marks $\approx$ 7mm (all chumbley score papers)
+    -  Bullet striation marks $\approx$ 2mm 
+    -  Screwdriver marks $\approx$ 7mm (all chumbley score papers)
 
-### Getting from Profiles to Signatures
-\vspace{-0.5em}
+- Let $x(t_1)$, $t_1 = 1,2,...T_1$ and $y(t_2)$, $t_2 = 1,2...T_2$ be two digitized marks (where $T_1$ and $T_2$ are not necessarily equal).
 
-- Lowess smoothing or Loess fit residuals $=$ Signatures
+- $T_1$ and $T_2$ are the final pixel indexes of each marking. 
+Therefore give the respective lengths of the markings.
+
+- Signatures/ Profiles (NIST- Hamby) $\approx$ 1200 pixels (2 mm)
+  Screwdriver toolmarks (Chumbley Papers) $\approx$ 9000 pixels (7 mm)
+
+## Chumbley Score
+
+#### Step 0 : Defining a coarseness parameter
+- Used to remove drift and (sub)class characteristics from individual markings
+<!-- ### Getting from Profiles to Signatures -->
+<!-- \vspace{-0.5em} -->
+- Lowess or Loess fit residuals $=$ Signatures
 - Removes topographic structure (curvature)
@@
 
-## Error Rates for toolmarks
-
-\begin{table}[ht]
-\centering
-\resizebox{\textwidth}{!}{\begin{tabular}{rlrrr}
-  \hline
-Research paper & Method & Data Source & False Positives & False Negatives\\
-  \hline
- \textbf{\citet{manytoolmarks1}} & Maximized & Screwdrivers & & \\ & Correlation &  & - & - \\ \hline
- \textbf{\citet{chumbley}} & Randomized & Screwdrivers & & \\ (Same-Surface Same-Angle) & Chumbley Score & & 2.3\% & 8.9\%  \\ \hline
-  \textbf{\citet{afte-chumbley}} & Randomized & Slip-joint  &  & \\ & Chumbley Score &  & - & - \\ \hline 
-  \textbf{\citet{hadler}} & Deterministic & Screwdrivers & & \\ (Same-Surface Same-Angle) & Chumbley Score & & 0\% & 6\%  \\ \hline
- \textbf{\citet{manytoolmarks2}} & Similarity Measure & Screwdrivers & & \\ (Different Surfaces-same angle) & Relative Distance Metric & & 5.9\% &  9.4\% \\ (Same Surfaces-same angle)&  & & 0.22\% &  0\% \\
-   \hline
-\end{tabular}}
-\caption{\label{tab:toolmarks-ER} Error Rates for Toolmarks}
-\end{table}
-
-## Digitized Striation Marks
-
-- Let $x(t_1)$, $t_1 = 1,2,...T_1$ and $y(t_2)$, $t_2 = 1,2...T_2$ be two digitized marks (where $T_1$ and $T_2$ are not necessarily equal).
-
-- $T_1$ and $T_2$ are the final pixel indexes of each marking. 
-Therefore give the respective lengths of the markings.
-
-- Signatures/ Profiles (NIST- Hamby) $\approx$ 1200 pixels (2 mm)
-  Screwdriver toolmarks (Chumbley Papers) $\approx$ 9000 pixels (7 mm)
-- Markings are smoothed as a preprocessing step
-    - Used to remove drift and (sub)class characteristics from individual markings
-
-## Overview
-
-\begin{figure}[hbtp]
-\centering
-\includegraphics[width=1\textwidth]{imgs/flow-chart-adj-chumbley.png}
-
-\caption{\label{flow-chart} An overview of the chumbley score method}
-\end{figure}
-
-## Chumbley Score
+## Algorithm
 
@@
 - Windows $\implies$ short segments of the markings
-    - Have \textit{predefined sizes}. ($T_1$ or $T_2 >>> w_o >> w_v$)
+    - Have \textit{predefined sizes}. ($T_1$ or $T_2$ >>> $w_o$ & w_v$)
           1. $w_o$ used in the Optimization step 
@@
 
-- Aligns markings horizontally 
-- Correlation of all possible windows of size $w_o$ between $x(t_1)$ and $y(t_2)$ computed
-- \textit{Possible lag in the 2 markings} \newline Window Pair with maximized correlation $\implies$ \newline Optimal vertical (in-phase) shift of $t_1^o - t_2^o$ 
+- \textbf{Goal :}Align markings horizontally as best as possible
+- Correlation Matrix of all possible windows of size $w_o$ between $x(t_1)$ and $y(t_2)$ computed
+- \textit{Identify lag for horizontal alignment} \newline Window Pair with maximized correlation $\implies$ \newline Optimal vertical (in-phase) shift of $t_1^o - t_2^o$ 
       - For aligning the two markings. 
@@
 
-#### Validations Step
+#### Validation Step
 - Two sets of windows of size $w_v$ chosen from both markings (see Figure \ref{fig:win-comparison})
@@
 
-## 2
+## 4
+<!-- - No overlaps of pairs within selected marks $\implies$ independence. -->
+- Both same- and different-shift pairs correlations between the  markings are calculated. 
+- For Same-Source markings, correlations
+    - for the in-phase shift should be high
+    - for out-of-phase shift should be low.
+          - Provide a measure for the base-level correlation to which in-phase shift correlations can be compared.
+- The Chumbley score is the Mann Whitney U statistic computed by comparing between in-phase sample and out-of-phase sample.
 
-More precisely, let us define starting points of the windows of validation $s_i^{(k)}$ for each signature $k = 1, 2$ as 
+## Block Diagram
+
+\begin{figure}[hbtp]
+\centering
+\includegraphics[width=1\textwidth]{imgs/flow-chart-adj-chumbley.png}
+
+\caption{\label{flow-chart} An overview of the adjusted chumbley score method as given by \citet{hadler}}
+\end{figure}
+
+## Starting Points
+
+More precisely, let us define starting points of the windows of validation $s_i^{(k)}$ for each marking $k = 1, 2$ as 
 \begin{eqnarray}\label{eq.start}
@@
 
-## 3
+## The \citet{hadler} method (CS1)
 
-- Same-shift pairs of length $w_v$ are all pairs $$(s_i^{(1)}, s_i^{(2)}) \quad \forall \, i \, \in \, \mathbb{Z}$$ for which both $s_i^{(1)}$ and  $s_i^{(2)}$ are defined. 
+- Same-shift pairs of length $w_v$ are all pairs that start in: $$(s_i^{(1)}, s_i^{(2)}) \quad \forall \, i \, \in \, \mathbb{Z}$$ for which both $s_i^{(1)}$ and  $s_i^{(2)}$ are defined. 
 - Different-shift pairs are defined as $$(s_i^{(1)}, s_{-i-1}^{(2)}) \quad \forall \, i \, \in \, \mathbb{Z}$$ where both $s_i^{(1)}$ and  $s_{-i-1}^{(2)}$ are defined (see \cref{sketch-same-diff}).
@@
 
-## 4
-- No overlaps of pairs within selected marks $\implies$ independence.
-- Both same- and different-shift pairs correlations between the  markings are calculated. 
-- For Same-Source markings correlations
-    - for the in-phase shift should be high
-    - for out-of-phase shift should be low.
-          - Provide a measure for the base-level correlation to which in-phase shift correlations can be compared.
-- The Chumbley score is the Mann Whitney U statistic computed by comparing between in-phase sample and out-of-phase sample.
-
 ## Failed Tests
@@
 - Either because the number of eligible same-shift pairs is 0, or the number of different-shift pairs is 0. 
+- $t_1^o, t_2^o$ not necessarily independent
+    - \textbf{same-source:} Assume high dependence, corr($t_1^o,t_2^o$) $\approx$ 1
+        - Example: $w_o$ = 120, coarseness (c) = 0.3, corr($t_1^o,t_2^o$) = 0.85
+    - \textbf{diff-source:} Assume independence of $t_1^o,t_2^o$
+        - Example: $w_o$ = 120, coarseness (c) = 0.3, corr($t_1^o,t_2^o$) = 0.12
+<!-- \vspace{-0.5em} -->
 
-\vspace{-0.5em}
+<!-- ###### No Same-shift pairs if -->
+<!-- - the optimal locations $t_1^{o}$ and $t_2^{o}$ are in the extremes -->
+<!--     - no segments of size $w_v$ are left on the same sides of the optimal locations,  -->
+<!--     - For all $t_1^{o} < w_v$ and $t_2^{o} > T_2-w_o-w_v$ or \newline $t_1^{o} < T_1- w_o - w_v$ and $t_2^{o} < w_v$ -->
 
-###### No Same-shift pairs if
-- the optimal locations $t_1^{o}$ and $t_2^{o}$ are in the extremes
-    - no segments of size $w_v$ are left on the same sides of the optimal locations, 
-    - For all $t_1^{o} < w_v$ and $t_2^{o} > T_2-w_o-w_v$ or \newline $t_1^{o} < T_1- w_o - w_v$ and $t_2^{o} < w_v$
+<!-- \vspace{-0.5em} -->
 
-\vspace{-0.5em}
+<!-- ###### No Different Shift Pairs if -->
+<!-- - the optimal locations are close to the boundaries -->
+<!--     - \# of different-shift pairs also depends on the placement of the  optimal locations $t_i^o$. -->
+<!--     -  For all $t_i^{(o)} < w_v$ or  $t_i^{(o)} > T_i-w_o- w_v$. -->
 
-###### No Different Shift Pairs if
-- the optimal locations are close to the boundaries
-    - \# of different-shift pairs also depends on the placement of the  optimal locations $t_i^o$.
-    -  For all $t_i^{(o)} < w_v$ or  $t_i^{(o)} > T_i-w_o- w_v$.
-
-\vspace{-0.5em}
-
-##
+<!-- \vspace{-0.5em} -->
 
@@
 
-#### Same-shift failure
+## Same-shift failure
+- Same-source $\approx$ 0
+- Different-source $\approx$ 2 $P(t_i < w_o)^2$ = $\frac{2 w_v^2}{(T_1-w_o)(T_2-w_o)}$
 
@@
 
-## 6
+## Different-Shift Failure
 
-- For different-source matches
-    - we can assume $t_1^o$ and $t_2^o$ are independent.
-    - Failure rate = $\frac{2 w_v^2}{(T_1-w_o)(T_2-w_o)}$ 
-    - For an average length of $T_i$ of 1200 pixels, $w_o = 120$ pixels and $w_v = 30$ pixels this probability is about 0.0015.
-- For same-source matches, 
-    - we expect a strong dependency between optimal locations $t_i^o$
-    - $t_1^o$ and $t_2^o$ are not independent of each other
-    - A large difference between locations is unlikely
-    - Failure Rate = $2 w_v / (T_i-w_o)$, or about 5.6\% for an average length of $T_i$ of 1200 pixels, $w_o = 120$ pixels and $w_v = 30$ pixels.
+- Same-source (Assuming $t_1^{o} \approx t_2^{o}$) $\approx$ $2w_v / (T_i-w_o)$\newline
 
-##  Failure rate for same-source matches
-The failure rate for same-source matches can also be written as follows
-$$
-\begin{split}
-P\left( t_1^{o} < w_v \ \bigcup \ t_2^{o} > T_2-w_o-w_v\right) + \\P\left( t_1^{o} < T_1- w_o - w_v \ \bigcup \ t_2^{o} < w_v\right)
-\end{split}
-$$
+\vspace{-2em}
 
-Taking first part of the above equation, let $A = t_1^{o} < w_v$ and $B = t_2^{o} > T_2-w_o-w_v$
+$P( t_1^{o} < w_v \ \cap \ t_2^{o} < w_v) + P( t_1^{o} <  w_v \ \cap \ t_2^{o} < w_v)  = \ 2P(t_o^1 \ < \ w_v)$
 
-- Again we assume high correlation between optimal locations
-- In order to have at least one same-shift pair or different shift pair
-    - \textbf{Either}, if $x \in A$, then $B$ is empty and $P(A \cup B) = P (A)$
-    - \textbf{or}, if $x \in B$, then $A$ is empty and $P(A \cup B) = P(B)$
+\vspace{-1em}
 
-This would hold for the second part too, and would therefore give an expected rate of failure as $2 w_v / (T_i-w_o)$
+- Different-source $\approx$  $2P(t_i < w_o)^2$ = $\frac{2 w_v^2}{(T_1-w_o)(T_2-w_o)}$
+
+\vspace{-1em}
+
+\begin{figure}[hbtp]
+\centering
+\includegraphics[width=0.8\textwidth]{imgs/diff-shift-failure.png}
+
+\caption{\label{diff-shift-failure}Sketch of diff-shift pairings  (top) when the number of diff-shift computations is likely to 0}
+\end{figure}
+
+<!-- - For different-source matches -->
+<!--     - we can assume $t_1^o$ and $t_2^o$ are independent. -->
+<!--     - Failure rate = $\frac{2 w_v^2}{(T_1-w_o)(T_2-w_o)}$  -->
+<!--     - For an average length of $T_i$ of 1200 pixels, $w_o = 120$ pixels and $w_v = 30$ pixels this probability is about 0.0015. -->
+<!-- - For same-source matches,  -->
+<!--     - we expect a strong dependency between optimal locations $t_i^o$ -->
+<!--     - $t_1^o$ and $t_2^o$ are not independent of each other -->
+<!--     - A large difference between locations is unlikely -->
+<!--     - Failure Rate = $2 w_v / (T_i-w_o)$, or about 5.6\% for an average length of $T_i$ of 1200 pixels, $w_o = 120$ pixels and $w_v = 30$ pixels. -->
+
+<!-- ##  Failure rate for same-source matches -->
+<!-- The failure rate for same-source matches can also be written as follows -->
+<!-- $$ -->
+<!-- \begin{split} -->
+<!-- P\left( t_1^{o} < w_v \ \bigcup \ t_2^{o} > T_2-w_o-w_v\right) + \\P\left( t_1^{o} < T_1- w_o - w_v \ \bigcup \ t_2^{o} < w_v\right) -->
+<!-- \end{split} -->
+<!-- $$ -->
+
+<!-- Taking first part of the above equation, let $A = t_1^{o} < w_v$ and $B = t_2^{o} > T_2-w_o-w_v$ -->
+
+<!-- - Again we assume high correlation between optimal locations -->
+<!-- - In order to have at least one same-shift pair or different shift pair -->
+<!--     - \textbf{Either}, if $x \in A$, then $B$ is empty and $P(A \cup B) = P (A)$ -->
+<!--     - \textbf{or}, if $x \in B$, then $A$ is empty and $P(A \cup B) = P(B)$ -->
+
+<!-- This would hold for the second part too, and would therefore give an expected rate of failure as $2 w_v / (T_i-w_o)$ -->
 

### 2018-04-24 doc/paper.Rmd

@@
 
+if (!file.exists("../data/cvcs1.RDS")) {
+files <- dir("../data/crossvalidate-cs1/")
+cvcs1 <- data.frame()
+for (file in files) {
+  tmp <- readRDS(file.path("../data/crossvalidate-cs1/",file))
+  cvcs1 <- rbind(cvcs1, tmp)
+}
+cvcs1$coarse <- cvcs1$coarseness
+cvcs1$method <- "CS1"
+saveRDS(cvcs1, file = "../data/cvcs1.RDS")
+} else cvcs1 <- readRDS("../data/cvcs1.RDS")
+
+cvcs <- rbind(cvcs1, cvcs2)
+
 library(plotROC)
-errorsCV <- rbind(errorrate(cvcs2, 0.001), 
+errorsCV2 <- rbind(errorrate(cvcs2, 0.001), 
                 errorrate(cvcs2, 0.005), 
@@
                 errorrate(cvcs2, 0.05))
-errorsCV$method <- "CS2"
+errorsCV2$method <- "CS2"
+
+errorsCV1 <- rbind(errorrate(cvcs1, 0.001), 
+                errorrate(cvcs1, 0.005), 
+                errorrate(cvcs1, 0.01), 
+                errorrate(cvcs1, 0.05))
+errorsCV1$method <- "CS1"
+
+errorsCV <- rbind(errorsCV1, errorsCV2)
 
@@
 
-cvcs2 %>% 
+cvcs %>% 
   ggplot(aes(d = as.numeric(!match), m = p_value, colour=factor(wo), 
-             shape=factor(wv))) + 
+             shape=factor(wv), linetype=method)) + 
   geom_roc(labels=FALSE, n.cuts=100, pointsize=0, size=0.6) +
@@
     expression("Optimization\nwindow size"~w[o]), palette="Set2") +
-  facet_grid(wv~method, labeller="label_both") +
+  facet_grid(.~wv, labeller="label_both") +
   xlab("False Positives Rate (1 - Specificity)") +

### 2018-04-24 doc/paper.Rmd
couple more CS1 results

@@
 errorsCV <- rbind(errorsCV1, errorsCV2)
-
 errorsCV %>% 

### 2018-04-25 doc/paper.Rmd
figures 9 and 10

@@
 
-```{r type2, fig.height = 3, fig.width = 4.5, out.width='0.6\\textwidth', fig.cap="Type II error rates observed across a range of window sizes for optimization $w_o$.
+```{r type2, fig.height = 3, fig.width = 7.5, out.width='0.9\\textwidth', fig.cap="Type II error rates observed across a range of window sizes for optimization $w_o$.
 For a window size of $w_o = 130$ we see a minimum in type II error rate across all type I rates considered.
@@
 
-```{r roc, fig.height = 3.25, fig.width = 4.75, out.width='0.6\\textwidth', fig.cap="ROC curves of methods CS1 and CS2 for different sizes of optimization window $w_o$.
+```{r roc, fig.height = 3.25, fig.width = 8, out.width='\\textwidth', fig.cap="ROC curves of methods CS1 and CS2 for different sizes of optimization window $w_o$.
 Optimal ROC curves are reached at optimization windows of size 150 and higher.
@@
   theme_bw() + 
+  scale_shape_discrete("Method", expression("Validation\nwindow size"~w[v])) +
   scale_colour_brewer(
@@
   xlab("False Positives Rate (1 - Specificity)") +
-  ylab("True Positives Rate (Sensitivity)")
+  ylab("True Positives Rate (Sensitivity)") +
+  guides(
+   colour = guide_legend(order = 3, ncol=2),
+   linetype = guide_legend(order = 1, ncol=2),
+   shape = guide_legend(order = 2, ncol=2)
+ )
 
@@
 
-```{r type1, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap="Comparison of observed and nominal type I error rates  across a range of window sizes for optimization $wo$.
+```{r type1, fig.width=7.75, fig.height=5.25, out.width='\\textwidth', fig.cap="Comparison of observed and nominal type I error rates  across a range of window sizes for optimization $wo$.
 The horizontal line in each facet indicates the nominal type I error rate."}
+errorsCV$wv.method <- paste(errorsCV$method, errorsCV$wv, sep=" / ")
 errorsCV %>% filter(wv %in% c(30,50)) %>%
-  ggplot(aes(x = wo, y = actual, colour=factor(alpha))) + 
+  ggplot(aes(x = wo, y = actual, colour=factor(alpha), 
+             linetype=method)) + 
   geom_hline(aes(yintercept=alpha), colour="grey30", size=.5) +
-  geom_point(aes(shape=factor(wv)), size=2.5) +
-  #  facet_grid(.~wo, labeller="label_both") + 
-  geom_smooth(aes(group=alpha), size=.8, se=FALSE, method="lm", alpha=.9) +
-  theme_bw() +
-#  scale_y_log10(breaks=c(0.001,.005, 0.01, 0.05)) +
-  scale_shape_discrete(expression("Size of validation window "~w[v])) +
+  geom_point(aes(shape=factor(wv.method)), size=2.5) +
+  geom_smooth(aes(group=factor(method)), 
+              size=.8, se=FALSE, method="lm", alpha=.9) +
+  theme_bw() +
   ylab("Observed type I error rate") +
   xlab(expression("Window size for optimization, "~w[o])) +
-  scale_colour_brewer(expression("Nominal type I error "~alpha), palette="Set2") +
+  scale_colour_brewer(
+    expression("Nominal type I error "~alpha), palette="Set2") +
   facet_wrap(~alpha, labeller="label_both", scales="free") +
-  theme(legend.position="bottom")
+  theme(legend.position="bottom") +
+  scale_shape_manual(expression("Method / Size of validation window "~w[v]), values=c(1,16,17)) +
+  scale_linetype_manual("Method", values=c(3,1), guide=FALSE) +
+  guides(shape = guide_legend(ncol=2),
+       #  linetype = guide_legend(ncol=1),
+         col = guide_legend(ncol=2))
 

### 2018-04-25 doc/paper.Rmd
mini edit ...

@@
   scale_shape_discrete("Method", expression("Validation\nwindow size"~w[v])) +
+  scale_linetype_discrete("Method") + 
   scale_colour_brewer(
@@
 \autoref{fig:type1} shows observed type I error rates across a range of optimization windows $w_o$.
-Generally, observed type I errors are higher than expected, but with an increase in the size of the optimization window observed type I error rates decrease.
+Generally, observed type I errors are higher than expected.
+Method CS1 shows in this instance slightly better performance than method CS2, but for both an increase in the size of the optimization window leads to a decrease in the observed type I error rates.
 
@@
 ```
-
+<!--
 \begin{table}[ht]
@@
 \end{table}
+-->
 

### 2018-04-25 doc/paper.tex
figures 9 and 10
mini edit ...

@@
 
-{\centering \includegraphics[width=0.6\textwidth]{figures/type2-1}
+{\centering \includegraphics[width=0.9\textwidth]{figures/type2-1}
 
@@
 
-{\centering \includegraphics[width=0.6\textwidth]{figures/roc-1}
+{\centering \includegraphics[width=\textwidth]{figures/roc-1}
 
@@
 optimization windows \(w_o\). Generally, observed type I errors are
-higher than expected, but with an increase in the size of the
-optimization window observed type I error rates decrease.
+higher than expected. Method CS1 shows in this instance slightly better
+performance than method CS2, but for both an increase in the size of the
+optimization window leads to a decrease in the observed type I error
+rates.
 
@@
 \end{figure}
-
-\begin{table}[ht]
-\caption{\label{tab:type2} Type II error rates for profiles and signatures of bullet lands. For profiles, a coarseness value of $c = 0.25$ is used to remove bullet curvature.}
-\centering
-\begin{tabular}{rlrrrr}
-  \hline
-validation && \multicolumn{4}{l}{Nominal type I error rate $\alpha$}\\  
-window $w_v$ & source & 0.001 & 0.005 & 0.01 & 0.05 \\ 
-  \hline
- 30 & profiles & 54.40 & 43.80 & 39.70 & 30.00 \\ 
-   30 & signatures & 55.00 & 45.40 & 41.40 & 31.10 \\ \hline
-   50 & profiles & 58.50 & 44.40 & 40.70 & 28.70 \\ 
-   50 & signatures & 62.60 & 49.60 & 44.20 & 30.50 \\ 
-   \hline
-\end{tabular}
-\end{table}
 

### 2018-04-25 presentation/presentation.Rmd
no figures only conclusions for failed tests, type II and coarseness

@@
 \begin{cases}
-t_k^* + i w_v^* & \text{ for } i < 0 \\
-t_k^* + w_ o^* + i w_v^* & \text{ for } i \ge 0,
+t_k^o + i w_v & \text{ for } i < 0 \\
+t_k^o + w_o + i w_v & \text{ for } i \ge 0,
 \end{cases}
-\end{eqnarray} for integer values of $i$ with $0 <  s^{(k)}_i \le 1 - w_v^*$
+\end{eqnarray} for integer values of $i$ with $0 <  s^{(k)}_i \le T_k - w_v$ where $s \ \in \mathbb{Z}$
 <!-- $0 <  s^{(k)}_i \le T_k - w_v$. -->
@@
 \end{figure}
+
+## Results Failed Tests
+
+## Conclusions Failed Tests
+
+- The number of failed tests is highly dependent on the size of the comparison windows and is correlated to the ground truth, 
+- Higher for known same-sourced lands than for known different sourced lands. 
+- The \textbf{CS1} algorithm fails to conduct a test about 8 to 13 % of the time for known same-source lands, and 2 to 6% of the time for known different source lands.
+- number of failed tests for CS1 is always higher than the corresponding theoretical number of failed tests.
+- Using CS2, the case with \textbf{largest} # of failed tests is still \textbf{lower} than the case where CS1 gives the \textbf{lowest} # of failed tests
+- Even when the coarseness is high, CS2 will have lower number of failed tests than CS1, Making it more robust.
+- CS2 performs better for both same and different-sources
+- Solves a critical issue of CS1 known same-source matching, by having a negligible number of Known same-source failed tests
+
+## Coarseness
+
+- Remove (sub-)class characteristics from profiles before comparisons for matching.
+- \citet{hadler} suggest a coarseness parameter of 0.25 for toolmark comparisons. 
+- For bullet lands, coarseness might need to be adjusted because of the strong effect bullet curvature has on profiles.
+- Optimal locations are distributed uniformly once (sub-)class characteristics are removed. 
+- Distinct boundary effects: c > 0.20 optimal locations $t^*$ are found at the very extreme ends of a profile more often than one would expect based on a uniform distribution.
+-smaller coarseness value of $c = 0.15$ to be suitable
+
+##
+
+```{r profile-sketch, fig.height = 4, fig.width='\\textwidth', fig.cap="Overview of the effect of different coarseness parameters $c$ on the profile shown in Figure \\ref{fig:sigs-profiles} (top).
+The bottom row shows histograms of  the (relative) optimal locations $t^o$ identified in the optimization step for different values of the coarseness parameter $c$.
+"}
+profile <- sigs_graphics[[1]] %>% filter(land_id==19)
+profile$lw10 <- lowess(x = profile$y, y = profile$value, f=.10)$y
+profile$lw15 <- lowess(x = profile$y, y = profile$value, f=.15)$y
+profile$lw20 <- lowess(x = profile$y, y = profile$value, f=.20)$y
+profile$lw25 <- lowess(x = profile$y, y = profile$value, f=.25)$y
+profile$lw30 <- lowess(x = profile$y, y = profile$value, f=.30)$y
+#profile$lw50 <- lowess(x = profile$y, y = profile$value, f=.5)$y
+#profile$lw75 <- lowess(x = profile$y, y = profile$value, f=.75)$y
+#profile$lw100 <- lowess(x = profile$y, y = profile$value, f=1)$y
+
+profiles <- tidyr::gather(profile, coarseness, values, lw10:lw30)
+profiles <- profiles %>% mutate(
+  c = readr::parse_number(coarseness)/100
+)
+p1 <- profiles %>% 
+  ggplot(aes(x = y, y = value-values)) + geom_line() +
+  theme_bw() +
+  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
+  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) +
+  facet_grid(~c, labeller="label_both")
+
+profc10 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo_160-wv_30-c_10.rds")
+profc15 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo_160-wv_30-c_15.rds")
+profc20 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo_160-wv_30-c_20.rds")
+profc20$coarseness <- 0.2
+profc25 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo_160-wv_30-c_25.rds")
+profc30 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo_160-wv_30.rds")
+profc30$coarseness <- .30
+profc <- bind_rows(profc10, profc15, profc20, profc25, profc30)
+profc <- profc %>% mutate(
+  locations = chumbley %>% purrr::map_int(.f = function(x) {
+    res <- NULL
+    try(res <- x$locations[1], silent = TRUE)
+    res
+  })
+)
+profsumm <- read.csv("../data/profiles-summary.csv")
+profsumm <- profsumm %>% dplyr::select(-run_id)
+profc <- profc %>% left_join(profsumm, by=c("land2_id"="land_id", "profile2_id"="profile_id"))
+
+p2 <- profc %>% 
+  ggplot(aes(x = locations/(length-wo))) +
+  geom_histogram(binwidth = 0.025) +
+  facet_grid(.~coarseness, labeller="label_both") + 
+  theme_bw() +
+  xlab(expression("Relative optimal location "~t^"*")) +
+  ylab("Number of profiles") +
+  scale_x_continuous(breaks=seq(0,1, by=0.25), labels=c("0", "0.25", "0.50", "0.75", "1"))
+
+grid.arrange(p1, p2, ncol=1)
+```
+## Type II error rates
+
+<!-- \autoref{fig:type2} gives an overview of the type 2 error of methods CS1 and CS2 across a range of different optimization windows $w_o$. -->
+```{r type2, fig.height = 8, fig.width = 8.5, out.width='0.6\\textwidth', fig.cap="Type II error rates observed across a range of window sizes for optimization $w_o$.
+For a window size of $w_o = 130$ we see a minimum in type II error rate across all type I rates considered.
+Smaller validation sizes $w_v$ are typically associated with a smaller type II error.", eval = FALSE}
+
+# source("./type-II.R")
+type2<- readRDS("./imgs/type2.rds")
+grid.arrange(type2)
+
+```
+
+<!-- \small{\textbf{CS1} \citet{hadler}} algorithm \newline  \small{\textbf{CS2} the suggested modified algorithm} -->
+\begin{figure}[hbtp]
+\centering
+\includegraphics[width=0.9\textwidth]{imgs/type-2-both-methods.png}
+\caption{\label{type2}  Type 2 error of methods CS1 and CS2 across a range of different optimization windows $w_o$.
+Top two figures are for a \textbf{coarseness} $\approx$ \textit{0.15} and the bottom one is for \textit{0.3}}
+\end{figure}
+
+## Conclusions for Type II Error
+
+\fontsize{10pt}{11}\selectfont
+
+#### CS1
+- Best works for $w_o$ of $\approx$ 130 to 160 and $w_v$ 50 when the smoothing is $c$ $\approx$ 0.3.
+- The Type II rate is lowest for a nominal $\alpha$ of 5%, with type I error rate of  6.2% and the Type II error rate of 24%. 
+- For lower nominal alpha levels of 1%, 0.5% and 0.1% the lowest type II error rate increases to about 36.4%, 41% and 52.5% respectively.
+- Gets worse for coarseness 0.15
+
+#### CS2
+- For a window size of $w_o = 130$ we see a minimum in type II error rate across all type I rates considered. 
+- Smaller validation sizes $w_v$ are typically associated with a smaller type II error.
+- CS2 shows an increase in the power of the test.
+- Fix in CS2 will also improve power for matching toolmarks.
+- While significantly reduced over CS1, CS2 still has type 2 error rates on bullet lands that are  higher than the error rates achieved on the -much larger- toolmarks.
+- Might be able to get more power out of the test by adapting C2 in a bullet-to-bullet comparison.

### 2018-04-25 presentation/type-II.R

@@ -1,0 +1,143 @@
+# ```{r type2, fig.height = 3, fig.width = 4.5, out.width='0.6\\textwidth', fig.cap="Type II error rates observed across a range of window sizes for optimization $w_o$. For a window size of $w_o = 130$ we see a minimum in type II error rate across all type I rates considered. Smaller validation sizes $w_v$ are typically associated with a smaller type II error."}
+errorrate <- function(data, alpha) {
+  summ <- data %>% filter(!is.na(p_value)) %>%
+    mutate(signif = p_value < alpha) %>%
+    group_by(wv, wo, match, coarse, signif) %>% tally()
+  summ$error <- with(summ, match != signif)
+  summ$alpha <- alpha
+
+  rates <- summ %>% dplyr::group_by(wv, wo, coarse, match, alpha) %>% dplyr::summarize(
+    rate = n[error==TRUE]/sum(n)
+  )
+  totals <- summ %>% dplyr::group_by(wv, wo, coarse, alpha) %>% dplyr::summarize(
+    total= sum(n[error==TRUE])/sum(n)
+  )
+  rates <- rates %>% spread(match,rate) %>% dplyr::rename(
+    actual = `FALSE`,
+    beta = `TRUE`
+  )
+  left_join(rates, totals)
+}
+
+
+if (!file.exists("../data/cvcs2.RDS")) {
+  files <- dir("../data/crossvalidate-cs2/")
+  cvcs2 <- data.frame()
+  for (file in files) {
+    tmp <- readRDS(file.path("../data/crossvalidate-cs2/",file))
+    cvcs2 <- rbind(cvcs2, tmp)
+  }
+  cvcs2$coarse <- cvcs2$coarseness
+  cvcs2$method <- "CS2"
+  saveRDS(cvcs2, file = "../data/cvcs2.RDS")
+} else cvcs2 <- readRDS("../data/cvcs2.RDS")
+
+if (!file.exists("../data/cvcs1.RDS")) {
+  files <- dir("../data/crossvalidate-cs1/")
+  cvcs1 <- data.frame()
+  for (file in files) {
+    tmp <- readRDS(file.path("../data/crossvalidate-cs1/",file))
+    cvcs1 <- rbind(cvcs1, tmp)
+  }
+  cvcs1$coarse <- cvcs1$coarseness
+  cvcs1$method <- "CS1"
+  saveRDS(cvcs1, file = "../data/cvcs1.RDS")
+} else cvcs1 <- readRDS("../data/cvcs1.RDS")
+
[…]

### 2018-04-25 doc/paper.Rmd
change to ROC curves

@@
 
-library(plotROC)
 errorsCV2 <- rbind(errorrate(cvcs2, 0.001), 
@@
 Points of equal error rates (EERs) can be found at the intersection of the dotted line and the ROC curves."}
-
-cvcs %>% 
-  ggplot(aes(d = as.numeric(!match), m = p_value, colour=factor(wo), 
-             shape=factor(wv), linetype=method)) + 
-  geom_roc(labels=FALSE, n.cuts=100, pointsize=0, size=0.6) +
-  geom_abline(
-    linetype=3, slope=-1, intercept = 1, colour="grey30", size=.3) +
-  theme_bw() + 
-  scale_shape_discrete("Method", expression("Validation\nwindow size"~w[v])) +
-  scale_linetype_discrete("Method") + 
-  scale_colour_brewer(
-    expression("Optimization\nwindow size"~w[o]), palette="Set2") +
-  facet_grid(.~wv, labeller="label_both") +
+library(plotROC)
+library(RColorBrewer)
+cols <- c(brewer.pal(6, "Reds")[-1], brewer.pal(6, "Blues")[-1])
+
+cvcs <- cvcs %>% mutate(
+  method.wo = paste(method, wo, sep= " / "),
+  method.wo = factor(method.wo, 
+                     levels = c("CS1 / 90", "CS1 / 120", "CS1 / 150", "CS1 / 180", "CS1 / 210", "CS2 / 90", "CS2 / 120", "CS2 / 150", "CS2 / 180", "CS2 / 210"))
+)
+cvcs %>% filter(!wo %in% c(130, 140) ) %>%
+  ggplot(aes(d = as.numeric(match), m = 1-p_value)) + 
+  geom_roc(labels=FALSE, n.cuts=100, pointsize=0, size=0.6, 
+           aes(colour = method.wo)) +
+  facet_wrap(~wv, labeller = label_bquote(w[v] == .(wv))) +
+  scale_alpha_discrete(range=c(0.1, 0.6)) +
+  scale_colour_manual(expression("Method /\nOptimization\nwindow size"~w[o]), values=cols) +
+  theme_bw() +
   xlab("False Positives Rate (1 - Specificity)") +
@@
   guides(
-   colour = guide_legend(order = 3, ncol=2),
-   linetype = guide_legend(order = 1, ncol=2),
-   shape = guide_legend(order = 2, ncol=2)
- )
+   colour = guide_legend(ncol=2)
+  ) +
+  geom_abline(slope=-1, intercept = 1, linetype = 3, 
+              colour="grey30", size = 0.25)
 

### 2018-04-25 doc/paper.Rmd
type 1 observed

@@
 p1 <- sigs_graphics[[1]] %>% filter(land_id==19) %>%
-  ggplot(aes(x = y, y = value)) + geom_line() +
+  ggplot(aes(x = y/1000, y = value)) + geom_line() +
   theme_bw() +
   ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
-  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
+  xlab(expression(paste("Relative location (in mm)", sep=""))) + 
   ggtitle("Profile")
@@
 
-\autoref{fig:roc} gives an overview of ROC curves for methods CS1 and CS2 over a range of different optimization window sizes $w_o$.
+\autoref{fig:roc} gives an overview of  ROC (Receiver operating characteristic) curves for methods CS1 and CS2 over a range of different optimization window sizes $w_o$.
+The different color hues represent the two methods CS1 (red) and CS2 (blue).
+The ROC curves show the superior performance of CS2 over CS1.
+Generally, an optimization window $w_o$ of 150 pixels or more leads to the best performance with respect to ROC curves.
 
 ```{r roc, fig.height = 3.25, fig.width = 8, out.width='\\textwidth', fig.cap="ROC curves of methods CS1 and CS2 for different sizes of optimization window $w_o$.
-Optimal ROC curves are reached at optimization windows of size 150 and higher.
+Best performances with respect to  ROC curves are reached for optimization windows of sizes 150 and higher.
 Points of equal error rates (EERs) can be found at the intersection of the dotted line and the ROC curves."}
@@
   theme(legend.position="bottom") +
-  scale_shape_manual(expression("Method / Size of validation window "~w[v]), values=c(1,16,17)) +
+  scale_shape_manual(expression("Method / Validation window "~w[v]), values=c(1,2,16,17)) +
   scale_linetype_manual("Method", values=c(3,1), guide=FALSE) +
   guides(shape = guide_legend(ncol=2),
-       #  linetype = guide_legend(ncol=1),
          col = guide_legend(ncol=2))
@@
 ```
-
-<!--
-### Results for Signatures
-
-\hh{For signatures from NIST scans we see three problems: 
-\begin{enumerate}
-\item type-2 error rate is at best 30\% for a type-1 error rate of 5\%, which is well above the error rates we see for tool marks from screw drivers, see figure \ref{fig:type2};
-\item the observed type-1 error, which generally close to the nominal type-1 error rate, depends on the size of the optimization window: as the window size increases, the observed type-1 error decreases, see figure \ref{fig:type1};
-\item the Chumbley-score fails to provide a result for up to 3\% of the cases.
-The number of failed tests increases linearly in the size used for the window in the optimization step.
-The rate of failed tests is considerably higher when the two lands are from same source than when the lands are from different sources, see figure \ref{fig:missings}.
-\end{enumerate}
-}
-
-Figure \ref{fig:missings} gives an overview of the number of failed tests, i.e.
-tests in which a particular parameter setting did not return a valid result.
-This happens e.g.\ when the shift to align two markings is so large, that the remaining overlap is too small to accommodate  windows for validation.
-The problem is therefore exacerbated by a larger validation window.
-Figure \ref{fig:missings}(left) also shows that the number of failed tests is approximately linear in the size of the optimization window.
-Tests also fail at a higher rate than expected when the markings are from the same source (right).
-This difference is the least pronounced around an optimized window size $w_o$ of around 120.
-However, even in this scenario, the number of failed tests for markings from the same source is about twice as high as expected given the number of same source and different source pairings in the data set.
--->
-
-<!--
-Following on similar lines to the setup of toolmarks, the first step here is to first identify what difference does different window sizes of optimization and the validation step have, when adapting the toolmark method to bullets.
-
-The marking made on bullets are smaller than toolmarks and is also less wider.
-The idea is to find out possible areas of error while adapting the score based method proposed for toolmarks, using cross-validation setup to identify appropriate parameter settings for (a) signatures and (b) profiles directly
-
-### Signatures
--->
-
-<!--
-```{r type2b, fig.width=8, fig.height=5, out.width='.8\\textwidth', fig.cap="Type II error rates observed across a range of window sizes for optimization $w_o$.
-For a window size of $w_o = 120$ we see a drop in type II error rate across all type I rates considered.
-Smaller validation sizes $w_v$ are typically associated with a smaller type II error."}
-errors %>% filter(wv %in% c(30,50)) %>%
-  ggplot(aes(x = wo, y = beta, colour=factor(alpha))) + 
-  geom_point(aes(shape=factor(wv)), size=3.5) +
-#  facet_grid(.~wo, labeller="label_both") + 
-  geom_smooth(aes(group=alpha), se=FALSE, method="loess", size=.7) +
-  theme_bw() +
-  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") +
-  scale_shape_discrete(expression("Size of\nvalidation window "~w[v])) +
-  xlab(expression("Window size for optimization "~w[o])) +
-  ylab("Type II error rate")
-
-labels = expression(P[M1](tilde(z)>0),P[M0](tilde(z)>0))
-```
--->
-
-<!--
-```{r missings, out.width='\\textwidth', fig.width=8, fig.height=4, fig.cap="The number of failed tests increases with an increase in the size of the optimization window (left).
-Unfortunately there is also a dependency between failed tests and ground truth.
-The plot on the right shows the ratio of the number of land pairs from same sources and different sources for failed tests.
-For small optimization windows and large windows the number of failed tests for same-source land-to-land comparisons is increasing.
-Even in the minimum, same-source land-to-land comparisons fail at twice the rate that they are expected to based on the ratio of the number of known matches and known non-matches (horizontal line).",fig.align = "center"}
-#ns1<- ns[which(ns$wv == c(30,50), arr.ind = T),]
-
-fails <- all %>% group_by(wv, wo) %>% summarize(
-  miss = sum(is.na(p_value)),
-  missperc = miss/n()*100
-)
-
-failsbymatch <- all %>% group_by(wv, wo, match) %>% summarize(
-  fails = sum(is.na(p_value))
-) %>% mutate(
-    ratio = fails[match==TRUE]/fails[match==FALSE],
-    match = c("Different Source", "Same Source")[as.numeric(as.logical(match)) + 1]
-    )
-
-missings <- fails %>% filter(wv %in% c(30, 50)) %>%
-  ggplot(aes(x = wo, y = missperc, 
-             colour = factor(wv),
-             shape=factor(wv)), size=.3) + 
-  geom_smooth(method="lm", size=.5, se=FALSE) +
-  geom_point(size=2.5) + 
-  theme_bw() +
-  ylab("Percent of failed tests.") +
-  xlab(expression("Size of optimization window "~w[o])) +
-  scale_colour_brewer(expression("Size of validation window "~w[v]), palette="Set1") +
-  scale_shape(expression("Size of validation window "~w[v])) +
-    theme(legend.position="bottom")
-
-missingsbymatch <- failsbymatch %>% 
-  filter(wv %in% c(30, 50),
-         match=="Same Source") %>%
-  ggplot(aes(x = wo, y = ratio, 
-             colour = factor(wv),
-             shape=factor(wv)), size=.3) + 
-  geom_hline(yintercept = 43960/2948225, size=0.5, colour="grey30") +
-  geom_smooth(method="loess", span=0.8, size=.5, se=FALSE) +
-  geom_point(size=2.5) + 
-  theme_bw() +
-  ylab("Ratio of same/different source\namong failed tests") +
-  xlab(expression("Size of optimization window "~w[o])) +
-  scale_colour_brewer(expression("Size of validation window "~w[v]), palette="Set1") +
-  scale_shape(expression("Size of validation window "~w[v])) +
-    theme(legend.position="bottom") +
-  scale_y_continuous(breaks=c(0.015, 0.03, 0.06, 0.09))
-
-gridExtra::grid.arrange(missings, missingsbymatch, ncol=2)
-
-```
--->
-
-<!--
-Signatures of lands for all Hamby-44 and Hamby-252 scans made available through the NIST ballistics database \citep{nist} were considered.
-Both of these sets of scans are part of the larger Hamby study \citep{hamby} and each consist of twenty known bullets (two each from ten consecutively rifled Ruger P85 barrels) and fifteen questioned bullets (each matching one of the ten barrels).
-Ground truth for both of these Hamby sets is known and was used to assess correctness of the tests results.
-
-Bullet signatures being compared at this time are therefore from the Hamby 44 and Hamby 252 data.
-The database setup and pre-processing system used for choosing the Bullet signatures are as described by \citet{aoas}.
-In order to choose the bullet signatures we first filter out Land_id for Profiles from the Hamby 44 and Hamby 252 data and remove all NA values.
-Then run_id = 3 is chosen as the signatures generated from this run_id give the closest match.
-Different run_id's have some different settings for generating the signatures.(The level of smoothing does not seem to be one of them)
-
-The bullet signatures when generated by this process already includes a loess smoothing.
-Therefore, the coarseness factor is set to 1 while running the chumbley non random algorithm for comparing different optimization windows.The algorithm generates the same_shift, different_shift, U-Stat and P_value parameters which are then used to calculate the errors associated with different sets of window sizes.
-
-### Profiles
-
-The profiles are cross-sectional values of the the bullet striation mark which are chosen at an optimum height (x as used by \citet{aoas}).
-This x or height is not a randomly chosen level.
-The rationale behind the choice has been explained by \citet{aoas}.
-A region is first chosen where the cross-correlation seems to change very less and in this region an optimum height is chosen.
-The profiles generally resemble a curve which is more or less similar to a quadratic curve (a quadratic fit to the raw data values of the profile is not an exact fit but it does show a similar trend).
-Profiles are the set of raw values representing the striation marks, and signatures are generated from these by removal of the inherent curvature and applying some smoothing (the signatures generated by \citet{aoas} use a loess function for smoothing).
-
-Similar to signatures the run_id = 3 was used when applying the chumbley algorithm using the database setup given by \citet{aoas} of Hamby-44 and Hamby-252 datasets, on the profiles.
-The run_id not only defines the level of smoothing but also signifies the chosen height at which the profiles were selected initially.
-Another important aspect is the range of horizontal values (which is referred to as the y values in \citet{aoas}) in the signatures.
-These have already been pre-processed in the database to not include any grooves.
-
-Therefore for the sake of comparison the run_id = 3 is still chosen so as to ensure that the horizontal values remain the same as that of the signatures.
-This also gives us profiles with the grooves removed.
-
-The idea therefore is to first use these raw values of the profile directly in the chumbley algorithm, and see how the algorithm performs for different coarseness values (smoothing parameter as referred in the function lowess used in the chumbley algorithm).
--->
-
-```{r, include =F}
-wo120_sig <- errors %>% filter(wo==120, wv %in% c(30, 50))
-wo120_pr <- errors2 %>% filter(wo==120, wv %in% c(30, 50), coarse == .25)
-
-dt <- rbind(wo120_sig, wo120_pr) %>% 
-  ungroup() %>% 
-  mutate(beta = round(100*beta, 1)) %>%
-  dplyr::select(wv, type, alpha, beta) %>% 
-  spread(alpha, beta)
-print(xtable(dt), include.rownames=FALSE)
-```
-<!--
-\begin{table}[ht]
-\caption{\label{tab:type2} Type II error rates for profiles and signatures of bullet lands. For profiles, a coarseness value of $c = 0.25$ is used to remove bullet curvature.}
-\centering
-\begin{tabular}{rlrrrr}
-  \hline
-validation && \multicolumn{4}{l}{Nominal type I error rate $\alpha$}\\  
-window $w_v$ & source & 0.001 & 0.005 & 0.01 & 0.05 \\ 
-  \hline
- 30 & profiles & 54.40 & 43.80 & 39.70 & 30.00 \\ 
-   30 & signatures & 55.00 & 45.40 & 41.40 & 31.10 \\ \hline
-   50 & profiles & 58.50 & 44.40 & 40.70 & 28.70 \\ 
-   50 & signatures & 62.60 & 49.60 & 44.20 & 30.50 \\ 
-   \hline
-\end{tabular}
-\end{table}
--->
 

### 2018-04-25 doc/paper.tex
type 1 observed

@@
 
-\autoref{fig:roc} gives an overview of ROC curves for methods CS1 and
-CS2 over a range of different optimization window sizes \(w_o\).
+\autoref{fig:roc} gives an overview of ROC (Receiver operating
+characteristic) curves for methods CS1 and CS2 over a range of different
+optimization window sizes \(w_o\). The different color hues represent
+the two methods CS1 (red) and CS2 (blue). The ROC curves show the
+superior performance of CS2 over CS1. Generally, an optimization window
+\(w_o\) of 150 pixels or more leads to the best performance with respect
+to ROC curves.
 
@@
 \caption{ROC curves of methods CS1 and CS2 for different sizes of optimization window $w_o$.
-Optimal ROC curves are reached at optimization windows of size 150 and higher.
+Best performances with respect to  ROC curves are reached for optimization windows of sizes 150 and higher.
 Points of equal error rates (EERs) can be found at the intersection of the dotted line and the ROC curves.}\label{fig:roc}

### 2018-04-25 presentation/presentation.Rmd
presentation +s

@@
 ## Conclusions Failed Tests
+\fontsize{10pt}{11}\selectfont
 
-- The number of failed tests is highly dependent on the size of the comparison windows and is correlated to the ground truth, 
-- Higher for known same-sourced lands than for known different sourced lands. 
-- The \textbf{CS1} algorithm fails to conduct a test about 8 to 13 % of the time for known same-source lands, and 2 to 6% of the time for known different source lands.
-- number of failed tests for CS1 is always higher than the corresponding theoretical number of failed tests.
+- Number is highly dependent on the comparison window sizes
+- Correlated to the ground truth, 
+- \textbf{Higher for known same-sourced lands (CS1)} than for known different sourced lands. 
+- \textbf{CS1} fails to conduct a test about 8 to 13 % of the time for \textbf{known same-source lands}, and 2 to 6% of the time for \textbf{known different source lands}.
+- Number for CS1 always higher than the corresponding theoretical number of failed tests.
 - Using CS2, the case with \textbf{largest} # of failed tests is still \textbf{lower} than the case where CS1 gives the \textbf{lowest} # of failed tests
-- Even when the coarseness is high, CS2 will have lower number of failed tests than CS1, Making it more robust.
-- CS2 performs better for both same and different-sources
-- Solves a critical issue of CS1 known same-source matching, by having a negligible number of Known same-source failed tests
+- Even for high coarseness, CS2 will have lower number of failed tests than CS1, Making it more robust.
+- CS2 performs better for \textbf{both} same and different-sources
+- Solves a \textbf{critical issue} of CS1 known same-source matching, by having a \textbf{negligible number of Known same-source failed tests}
 
@@
 #### CS1
-- Best works for $w_o$ of $\approx$ 130 to 160 and $w_v$ 50 when the smoothing is $c$ $\approx$ 0.3.
+- Best works best for $w_o$ of $\approx$ 130 to 160 and $w_v$ 50 when the smoothing is $c$ $\approx$ 0.3.
 - The Type II rate is lowest for a nominal $\alpha$ of 5%, with type I error rate of  6.2% and the Type II error rate of 24%. 
@@
 #### CS2
-- For a window size of $w_o = 130$ we see a minimum in type II error rate across all type I rates considered. 
-- Smaller validation sizes $w_v$ are typically associated with a smaller type II error.
+- Significantly reduced over CS1
+- For a window size of $w_o = 130$ we see a minimum in type II error rate across all type I rates considered. - Smaller validation sizes $w_v$ are typically associated with a smaller type II error.
 - CS2 shows an increase in the power of the test.
-- Fix in CS2 will also improve power for matching toolmarks.
-- While significantly reduced over CS1, CS2 still has type 2 error rates on bullet lands that are  higher than the error rates achieved on the -much larger- toolmarks.
-- Might be able to get more power out of the test by adapting C2 in a bullet-to-bullet comparison.
+- Type II CS2, still much higher for bullet lands than for toolmarks.
+- Fix in CS2 will also improve power for matching toolmarks thans CS1
+- Bullet-to-bullet comparison using CS2 $\approx$ more power out of the test.
+
+## ROC Curves
+\fontsize{10pt}{11}\selectfont
+\begin{figure}[hbtp]
+\centering
+\includegraphics[width=0.9\textwidth]{imgs/roc-curves.png}
+\caption{\label{roc-curves}  ROC curves of methods CS1 and CS2 for different sizes of optimization window $w_o$.}
+\end{figure}
+
+\vspace{-0.5em}
+
+- Superior performance of CS2 over CS1
+- Best performances wrt  ROC curves are reached for $w_0$ 150 and higher. 
+- Points of equal error rates (EERs): intersection of the dotted line and the ROC curves.
+
+##
+
+### \textbf{THANK YOU. Questions?}

### 2018-04-25 presentation/presentation.Rmd
some additions

@@
 
+\begin{figure}[hbtp]
+\centering
+\includegraphics[width=0.9\textwidth]{imgs/failed-tests.png}
+\caption{\label{type2}  Percent of failed land-to-land comparisons for $w_o$ = 120 and coarseness $c$ = 0.25}
+\end{figure}
+
 ## Conclusions Failed Tests
@@
 
+- With an increase in the $w_v$ higher percent of tests fail under both CS1 and CS2
 - Number is highly dependent on the comparison window sizes
@@
 - CS2 performs better for \textbf{both} same and different-sources
-- Solves a \textbf{critical issue} of CS1 known same-source matching, by having a \textbf{negligible number of Known same-source failed tests}
+- Solves a \textbf{critical issue} of CS1 known same-source matching, by having a \textbf{negligible} number of Known same-source failed tests
 

### 2018-04-25 presentation/presentation.Rmd
some other presentation changes
pres changes

@@
 ### Structure
+- Error rates in toolmarks
 - Data being used
@@
 - Modifications to the Algorithm
-- Results using \citet{hadler} method and results using Modified method
+- Results
 

### 2018-04-25 presentation/presentation.Rmd
spme presadds

@@
 ### \textbf{THANK YOU. Questions?}
+
+##
+
+## Appendix
+
+###U Statistic:
+
+This is computed from the joint rank of all correlations of both the same and different shift samples. As given by \citet{hadler}
+
+Null Hypothesis: If the toolmarks were not match i.e not made by the same tool.
+
+Let $n_s$ and $n_d$ be the number of same shift and different shift windows $N = n_{s} + n_{d}$
+
+The mann whitney U statistic is given by $U =\sum^{ns}{i=1}R{s}\left( i\right)$
+
+with the standardized version which includes provision for rank ties
+
+$\overline{U}= \dfrac{U-M}{\sqrt{V}}$
+
+where prior to normalization the U-statistic has the mean as
+
+$M = n_{s}\left(\dfrac{N+1}{2}\right)$
+
+and variance
+
+$V = \dfrac{n_{s}n_{d}}{N\left(N-1\right)}\left[\Sigma^{n_{s}}R_{s}\left(i\right)^{2}+\Sigma^{n_{d}}R_{d}\left(j\right)^{2}\right] -\dfrac{n_{s} n_{d}\left(N+1\right)^{2}}{4\left(N-1\right)}$

### 2018-04-25 presentation/presentation.Rmd
pres thats it stop

@@
 
-Let $n_s$ and $n_d$ be the number of same shift and different shift windows $N = n_{s} + n_{d}$
+Let $n_s$ and $n_d$ be the number of same shift and different shift windows $$N = n_{s} + n_{d}$$
 
-The mann whitney U statistic is given by $U =\sum^{ns}{i=1}R{s}\left( i\right)$
+Additionally, let $R_s(i)$ and $R_d(j)$ be the ranks associated with the combined vector of correlations for the same-shift and different-shift correlations, for i = 1,2,...,$n_s$ and j = 1,2,...,$n_d$.
+Then the Mann–Whitney U-statistic is given by
+
+The mann whitney U statistic is given by $$U =\sum^{n_s}_{i=1}R_{s}\left( i\right)$$
+
+##
 

### 2018-04-25 doc/paper.Rmd
three more points for CS1

@@
 
-\hh{don't split the discussion on the size. between the next paragraph and }
-<!--The majority of Bullet profiles and signatures of the Hamby study \citep{hamby} used in this paper are extracted by procedures mentioned by  \citet{aoas}.--> 
 There are two sources of scans for sets from the Hamby study available to us: scans of Hamby 44 and Hamby 252 are available from the NIST database \citep{nist}.
@@
 
-## A problem with failures and modification
-
-Looking closer at \autoref{eq.start}, we see that by definition, some number of tests will fail to produce a result, either because the number of eligible same-shift pairs is zero, or the number of different-shift opairs is zero.
-\hh{XXX The problem of failed tests is first mentioned in XXX}
+## A problem with failed tests
+
+Looking closer at \autoref{eq.start}, we see that by definition, some number of tests will fail to produce a result.
+\hh{The problem of failed tests is first mentioned in \citet{afte-chumbley}.
+Unfortunately, the authors do not provide any percentage of how many tests failed for their data.}
+The algorithm fails to produce for two reasons: either the number of eligible same-shift pairs is zero, or the number of different-shift pairs is zero.
 
@@
 Assuming independence in the optimal locations for different source profiles the expected probability for a failed test is, again, $\frac{2 w_v^2}{(T_1-w_o)(T_2-w_o)}$.
+
+## A modified approach
 
@@
 p1 <- sigs_graphics[[1]] %>% filter(land_id==19) %>%
-  ggplot(aes(x = y/1000, y = value)) + geom_line() +
+  ggplot(aes(x = y, y = value)) + geom_line() +
   theme_bw() +
   ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
-  xlab(expression(paste("Relative location (in mm)", sep=""))) + 
+  xlab(expression(paste("Relative location (in ",mu,m,")", sep=""))) + 
   ggtitle("Profile")
@@
 p1 <- profiles %>% 
-  ggplot(aes(x = y, y = value-values)) + geom_line() +
+  ggplot(aes(x = y/1000, y = value-values)) + geom_line() +
   theme_bw() +
   ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
-  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) +
+  xlab(expression(paste("Relative Location (in mm)", sep=""))) +
   facet_grid(~c, labeller="label_both")

### 2018-04-25 doc/paper.tex
three more points for CS1

@@
 
-\hh{don't split the discussion on the size. between the next paragraph and }
 There are two sources of scans for sets from the Hamby study available
@@
 
-\subsection{A problem with failures and
-modification}\label{a-problem-with-failures-and-modification}
+\subsection{A problem with failed
+tests}\label{a-problem-with-failed-tests}
 
 Looking closer at \autoref{eq.start}, we see that by definition, some
-number of tests will fail to produce a result, either because the number
-of eligible same-shift pairs is zero, or the number of different-shift
-opairs is zero.
-\hh{XXX The problem of failed tests is first mentioned in XXX}
+number of tests will fail to produce a result.
+\hh{The problem of failed tests is first mentioned in \citet{afte-chumbley}. Unfortunately, the authors do not provide any percentage of how many tests failed for their data.}
+The algorithm fails to produce for two reasons: either the number of
+eligible same-shift pairs is zero, or the number of different-shift
+pairs is zero.
 
@@
 \(\frac{2 w_v^2}{(T_1-w_o)(T_2-w_o)}\).
+
+\subsection{A modified approach}\label{a-modified-approach}
 

### 2018-04-26 doc/paper.Rmd
EERs and AUCs

@@
 
+```{r eer, fig.height = 3.25, fig.width = 8, out.width='\\textwidth', fig.cap=""}
+auc <- function(tpr, fpr){
+  # from http://blog.revolutionanalytics.com/2016/11/calculating-auc.html
+  # inputs already sorted, best scores first 
+  dfpr <- c(diff(fpr), 0)
+  dtpr <- c(diff(tpr), 0)
+  sum(tpr * dfpr) + sum(dtpr * dfpr)/2
+}
+
+eer <- function(tpr, fpr) {
+  diff = abs(fpr-1+tpr)
+  idx <- which.min(diff)
+  EER <- 0.5*(fpr[idx]+1-tpr[idx])
+  EER
+}
+
+EER <- function(score, match) {
+  bstats <-data.frame(score, match)
+  bstats <- bstats %>% arrange(desc(score))
+
+  bstats <- bstats %>% mutate(
+    tpr=cumsum(match)/sum(match), 
+    fpr=cumsum(!match)/sum(!match)
+  ) %>% unique()
+
+  with(bstats, eer(tpr, fpr))
+}
+
+AUC <- function(score, match) {
+  bstats <-data.frame(score, match)
+  bstats <- bstats %>% arrange(desc(score))
+
+  bstats <- bstats %>% mutate(
+    tpr=cumsum(match)/sum(match), 
+    fpr=cumsum(!match)/sum(!match)
+  ) %>% unique()
+
+  with(bstats, auc(tpr, fpr))
+}
+
+roc.summary <- cvcs %>% group_by(wo, wv, method) %>%
+  summarize(
+    eer = EER(1-p_value, match),
+    auc = AUC(1-p_value, match)
+  )
+p1 <- roc.summary %>% filter(!(wo %in% c(130, 140))) %>%
+  ggplot(aes( x = wo, y=eer, shape=factor(wv))) +
+  geom_point() +
+  facet_wrap(~method, labeller="label_both") + 
+  theme_bw() +
+  scale_shape_discrete("Size of \nvalidation window") +
+  ylab("Equal Error rate (EER)") +
+  xlab("Size of optimization window in pixels") +
+  theme(legend.position = "none")
+
+p2 <- roc.summary %>% filter(!(wo %in% c(130, 140))) %>%
+  ggplot(aes( x = wo, y=auc, shape=factor(wv))) +
+  geom_point() +
+  facet_wrap(~method, labeller="label_both") + 
+  theme_bw() +
+  scale_shape_discrete("Size of \nvalidation window") +
+  ylab("Area under the curve (AUC)") +
+  xlab("Size of optimization window in pixels")+
+  theme(legend.position = "none")
+grid.arrange(p1, p2)
+```
+
 <!--
@@
 XXX
-The algorithm CS2 introduced here is constructed in a way that reduces the number of failures by a magnitude of XXX.
+The algorithm CS2 introduced here is constructed in a way that reduces the number of failures by a magnitude of about XXX.
 While reducing the failure rate, the algorithm also shows an increase in the power of the test.
@@
 \hh{
-Fix in CS2 will also improve power for matching toolmarks. 
+Fix in CS2 should also improve power for matching toolmarks. 
 }

### 2018-04-26 doc/paper.Rmd
pics updated

--- 2018-04-25 doc/paper.Rmd
@@
 errorsCV2$method <- "CS2"
-
 errorsCV1 <- rbind(errorrate(cvcs1, 0.001), 
@@
   facet_wrap(~method, labeller="label_both")
-
 ```
@@
 )
+
 cvcs %>% filter(!wo %in% c(130, 140) ) %>%
@@
 
+```
+
+```{r eer, fig.height = 3.25, fig.width = 8, out.width='\\textwidth', fig.cap="Comparison of results for CS1 and CS2 using area under the curve (AUC) and equal error rates (EER). "}
+auc <- function(tpr, fpr){
+  # from http://blog.revolutionanalytics.com/2016/11/calculating-auc.html
+  # inputs already sorted, best scores first 
+  dfpr <- c(diff(fpr), 0)
+  dtpr <- c(diff(tpr), 0)
+  sum(tpr * dfpr) + sum(dtpr * dfpr)/2
+}
+
+eer <- function(tpr, fpr) {
+  diff = abs(fpr-1+tpr)
+  idx <- which.min(diff)
+  EER <- 0.5*(fpr[idx]+1-tpr[idx])
+  EER
+}
+
+EER <- function(score, match) {
+  bstats <-data.frame(score, match)
+  bstats <- bstats %>% arrange(desc(score))
+
+  bstats <- bstats %>% mutate(
+    tpr=cumsum(match)/sum(match), 
+    fpr=cumsum(!match)/sum(!match)
+  ) %>% unique()
+
+  with(bstats, eer(tpr, fpr))
+}
+
+AUC <- function(score, match) {
+  bstats <-data.frame(score, match)
+  bstats <- bstats %>% arrange(desc(score))
+
+  bstats <- bstats %>% mutate(
+    tpr=cumsum(match)/sum(match), 
+    fpr=cumsum(!match)/sum(!match)
+  ) %>% unique()
+
+  with(bstats, auc(tpr, fpr))
+}
+
+roc.summary <- cvcs %>% group_by(wo, wv, method) %>%
+  summarize(
+    eer = EER(1-p_value, match),
+    auc = AUC(1-p_value, match)
+  )
+p1 <- roc.summary %>% filter(!(wo %in% c(130, 140))) %>%
+  ggplot(aes( x = wo, y=eer, shape=factor(wv))) +
+  geom_point() +
+  facet_wrap(~method, labeller="label_both") + 
+  theme_bw() +
+  scale_shape_discrete("Size of \nvalidation window") +
+  ylab("Equal Error rate (EER)") +
+  xlab("Size of optimization window in pixels") +
+  theme(legend.position = c(0.3, 0.2))
+
+p2 <- roc.summary %>% filter(!(wo %in% c(130, 140))) %>%
+  ggplot(aes( x = wo, y=auc, shape=factor(wv))) +
+  geom_point() +
+  facet_wrap(~method, labeller="label_both") + 
+  theme_bw() +
+  scale_shape_discrete("Size of \nvalidation window") +
+  ylab("Area under the curve (AUC)") +
+  xlab("Size of optimization window in pixels")+
+  theme(legend.position = c(0.7, 0.2))
+grid.arrange(p1, p2, nrow=1)
 ```

### 2018-04-26 doc/paper.tex
pics updated

@@
 
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figures/eer-1}
+
+}
+
+\caption{Comparison of results for CS1 and CS2 using area under the curve (AUC) and equal error rates (EER). }\label{fig:eer}
+\end{figure}
+
 \subsection{Observed versus Nominal Type I error

### 2018-05-03 doc/paper.Rmd
not sure what I changed

@@
 \newcommand{\hh}[1]{{\textcolor{orange}{#1}}}
-\newcommand{\gk}[1]{{\textcolor{green}{#1}}}
+\newcommand{\gk}[1]{{\textcolor{blue}{#1}}}
 \newcommand{\cited}[1]{{\textcolor{red}{#1}}}
@@
 
-# Multiple plot function
-#
-# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
-# - cols:   Number of columns in layout
-# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
-#
-# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
-# then plot 1 will go in the upper left, 2 will go in the upper right, and
-# 3 will go all the way across the bottom.
-#
-# Multiple plot function
-#
-# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
-# - cols:   Number of columns in layout
-# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
-#
-# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
-# then plot 1 will go in the upper left, 2 will go in the upper right, and
-# 3 will go all the way across the bottom.
-#
-multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
-  library(grid)
-
-  # Make a list from the ... arguments and plotlist
-  plots <- c(list(...), plotlist)
-
-  numPlots = length(plots)
-
-  # If layout is NULL, then use 'cols' to determine layout
-  if (is.null(layout)) {
-    # Make the panel
-    # ncol: Number of columns of plots
-    # nrow: Number of rows needed, calculated from # of cols
-    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
-                    ncol = cols, nrow = ceiling(numPlots/cols))
-  }
-
- if (numPlots==1) {
-    print(plots[[1]])
-
-  } else {
-    # Set up the page
-    grid.newpage()
-    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
-
-    # Make each plot, in the correct location
-    for (i in 1:numPlots) {
-      # Get the i,j matrix positions of the regions that contain this subplot
-      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
-
-      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
-                                      layout.pos.col = matchidx$col))
-    }
-  }
-}
-#  ref: http://www.cookbook-r.com/Graphs/Multiple_graphs_on_one_page_(ggplot2)/
 ```
@@
 The table should include the reference to the paper, the data used, the statistical method and the associated error rates.}
+\gk{table included}
 Various toolmarks have been studied in the literature:  \citet{manytoolmarks1} and \citet{chumbley} have been analyzing screwdriver marks  digitized using a profilometer;  \citet{manytoolmarks2} have investigated 3D marks from screwdriver, tongue and groove pliers captured using a confocal microscope; \citet{afte-chumbley} have been investigated digitized marks from slip-joint pliers generated by a surface profilometer.
@@
 \hh{We need an additional sentence here to get from the data to the statistical methods ... }
+\gk{Analysis of these digitized markings require the use of statistical methods which can quantify the scientific mechanism of comparing markings and serve as metrics for correctly identifying the error rate.}
 \citet{manytoolmarks2} define a relative distance metric and use it as similarity measure between two toolmarks.
@@
 In this paper, we are investigating the applicability of the Chumbley scoring method by \citet{hadler} to assess striation marks on bullet lands for same-source identification.
+
+\begin{table}[ht]
+\centering
+\resizebox{\textwidth}{!}{\begin{tabular}{rlrrr}
+  \hline
+Research paper & Method & Data Source & False Positives & False Negatives\\
+  \hline
+ \textbf{\citet{manytoolmarks1}} & Maximized & Screwdrivers & & \\ & Correlation &  & - & - \\ \hline
+ \textbf{\citet{chumbley}} & Randomized & Screwdrivers & & \\ (Same-Surface Same-Angle) & Chumbley Score & & 2.3\% & 8.9\%  \\ \hline
+  \textbf{\citet{afte-chumbley}} & Randomized & Slip-joint  &  & \\ & Chumbley Score &  & - & - \\ \hline 
+  \textbf{\citet{hadler}} & Deterministic & Screwdrivers & & \\ (Same-Surface Same-Angle) & Chumbley Score & & 0\% & 6\%  \\ \hline
+  \textbf{\citet{manytoolmarks2}} & Similarity Measure & Screwdrivers & & \\ (Different Surfaces-same angle) & Relative Distance Metric & & 5.9\% &  9.4\% \\ (Same Surfaces-same angle)&  & & 0.22\% &  0\% \\
+   \hline
+\end{tabular}}
+\caption{\label{tab:toolmarks-ER} Error Rates for Toolmarks using  variations of the chumbley score method}
+\end{table}
+
+ <!-- \textbf{\citet{manytoolmarks2}} & Similarity Measure & Screwdrivers & & \\ (Different Surfaces-same angle) & Relative Distance Metric & & 5.9\% &  9.4\% \\ (Same Surfaces-same angle)&  & & 0.22\% &  0\% \\ -->
 
@@
 \caption{\label{sketch-diff-2}Sketch of adjusted different-shift pairings. At most one of the same-shift pairings can not be matched with a different-shift pair. }
+\end{figure}
+
+\begin{figure}[hbtp]
+\centering
+\includegraphics[width=0.62\textwidth]{images/same-shift-failure.png}
+
+\caption{\label{same-shift-failure}Sketch of same-shift pairings  (top) when the lag is too large to accomodate a a vaildation window in either of the two signatures}
+\end{figure}
+
+\begin{figure}[hbtp]
+\centering
+\includegraphics[width=0.7\textwidth]{images/diff-shift-failure.png}
+
+\caption{\label{diff-shift-failure}Sketch of diff-shift pairings  (top) when the number of diff-shift computations is likely to 0}
+\end{figure}
+
+\begin{figure}[hbtp]
+\centering
+\includegraphics[width=.7\textwidth]{images/CS1-failure-CS2-success.png}
+\caption{\label{CS1-failure-CS2-success}Sketch of a case where CS1 fails but CS2 does not fail}
 \end{figure}
@@
 
+\gk{The key effect of the optimal locations and thereby the coarseness is seen in the number of failed tests.
+Irrespective of whether CS1 or CS2 is being used, if the relative optimal locations are at the boundaries we will see an increase in the number of failed tests.
+A balance is therefore needed in the selection of the coarseness parameter which reduces the boundary effect but also does not remove important (sub-)class characteristics}
+
 ```{r profile-sketch, fig.height = 4, fig.width='\\textwidth', fig.cap="Overview of the effect of different coarseness parameters $c$ on the profile shown in Figure \\ref{fig:sigs-profiles} (top).
@@
   xlab("Size of optimization window in pixels") +
-  theme(legend.position = c(0.3, 0.2))
+  theme(legend.position = c(0.25, 0.25))
 
@@
   xlab("Size of optimization window in pixels")+
-  theme(legend.position = c(0.7, 0.2))
+  theme(legend.position = c(0.75, 0.25))
 grid.arrange(p1, p2, nrow=1)
@@
 
-```{r prof_missings, fig.width=10, fig.height=10, fig.cap="Row 3:  Total error and Number of failed tests by the window validation size, wv, and ground truth, Row 2: Total error and Number of failed tests with Coarseness for both profiles and signatures, Row 1: Type II error for different coarseness levels as used in the modified chumbley algorithm for profiles and signatures"}
+```{r prof_missings, fig.width=10, fig.height=10, fig.cap="Row 3:  Total error and Number of failed tests by the window validation size, wv, and ground truth, Row 2: Total error and Number of failed tests with Coarseness for both profiles and signatures, Row 1: Type II error for different coarseness levels as used in the modified chumbley algorithm for profiles and signatures", eval = FALSE}
 

### 2018-05-07 doc/paper.Rmd
intro

--- 2018-04-26 doc/paper.Rmd
@@
 In this paper, we are investigating the applicability of the Chumbley scoring method by \citet{hadler} to assess striation marks on bullet lands for same-source identification.
+
+\begin{table}[ht]
+\centering
+\resizebox{\textwidth}{!}{\begin{tabular}{rlrrr}
+  \hline
+Research paper & Method & Data Source & False Positives & False Negatives\\
+  \hline
+ \textbf{\citet{manytoolmarks1}} & Maximized & Screwdrivers & & \\ & Correlation &  & - & - \\ \hline
+ \textbf{\citet{chumbley}} & Randomized & Screwdrivers & & \\ (Same-Surface Same-Angle) & Chumbley Score & & 2.3\% & 8.9\%  \\ \hline
+  \textbf{\citet{afte-chumbley}} & Randomized & Slip-joint  &  & \\ & Chumbley Score &  & - & - \\ \hline 
+  \textbf{\citet{hadler}} & Deterministic & Screwdrivers & & \\ (Same-Surface Same-Angle) & Chumbley Score & & 0\% & 6\%  \\ 
+   \hline
+\end{tabular}}
+\caption{\label{tab:toolmarks-ER} Error Rates for Toolmarks using  variations of the chumbley score method}
+\end{table}
+
+ <!-- \textbf{\citet{manytoolmarks2}} & Similarity Measure & Screwdrivers & & \\ (Different Surfaces-same angle) & Relative Distance Metric & & 5.9\% &  9.4\% \\ (Same Surfaces-same angle)&  & & 0.22\% &  0\% \\ -->
 

### 2018-05-09 doc/paper.Rmd
some additions on coarseness discussion

--- 2018-05-03 doc/paper.Rmd
@@
 The table should include the reference to the paper, the data used, the statistical method and the associated error rates.}
-\gk{table included}
 Various toolmarks have been studied in the literature:  \citet{manytoolmarks1} and \citet{chumbley} have been analyzing screwdriver marks  digitized using a profilometer;  \citet{manytoolmarks2} have investigated 3D marks from screwdriver, tongue and groove pliers captured using a confocal microscope; \citet{afte-chumbley} have been investigated digitized marks from slip-joint pliers generated by a surface profilometer.
@@
 \hh{We need an additional sentence here to get from the data to the statistical methods ... }
-\gk{Analysis of these digitized markings require the use of statistical methods which can quantify the scientific mechanism of comparing markings and serve as metrics for correctly identifying the error rate.}
 \citet{manytoolmarks2} define a relative distance metric and use it as similarity measure between two toolmarks.
@@
   xlab("Size of optimization window in pixels") +
-  theme(legend.position = c(0.25, 0.25))
+  theme(legend.position = c(0.3, 0.2))
 
@@
   xlab("Size of optimization window in pixels")+
-  theme(legend.position = c(0.75, 0.25))
+  theme(legend.position = c(0.7, 0.2))
 grid.arrange(p1, p2, nrow=1)

### 2018-05-11 doc/paper.Rmd
keywords

@@
 keywords:
-- 3 to 6 keywords
-- that do not appear in the title
+- forensic science
+- toolmark
+- cross-correlation
+- Mann-Whitney U statistic
+- 
+- 
 bibliography: bibliography
@@
 Irrespective of whether CS1 or CS2 is being used, if the relative optimal locations are at the boundaries we will see an increase in the number of failed tests.
-A balance is therefore needed in the selection of the coarseness parameter which reduces the boundary effect but also does not remove important (sub-)class characteristics}
+A balance is therefore needed in the selection of the coarseness parameter which reduces the boundary effect but also does not remove important individual characteristics}
 
@@
 
+fails <- cvcs %>% group_by(wo, wv, coarse, method, match) %>% summarize(
+  fails = 100.0*sum(is.na(p_value))/n()
+)
+
 errorsCV2 <- rbind(errorrate(cvcs2, 0.001), 

### 2018-05-11 doc/paper.tex
keywords

@@
 \newcommand{\hh}[1]{{\textcolor{orange}{#1}}}
-\newcommand{\gk}[1]{{\textcolor{green}{#1}}}
+\newcommand{\gk}[1]{{\textcolor{blue}{#1}}}
 \newcommand{\cited}[1]{{\textcolor{red}{#1}}}
-
 \setlength\parindent{0pt}
@@
 
-\section{Introduction and Background}\label{introduction-and-background}
+\hypertarget{introduction-and-background}{%
+\section{Introduction and
+Background}\label{introduction-and-background}}
 
-\subsection{Motivation}\label{motivation}
+\hypertarget{motivation}{%
+\subsection{Motivation}\label{motivation}}
 
@@
 
-\hh{In the next page and a half it is easy to lose the red line. It might help to include a table with an overview.
-The table should include the reference to the paper, the data used, the statistical method and the associated error rates.}
+{\textcolor{orange}{In the next page and a half it is easy to lose the red line. It might help to include a table with an overview.
+The table should include the reference to the paper, the data used, the statistical method and the associated error rates.}}
+{\textcolor{blue}{table included}}
+
 Various toolmarks have been studied in the literature:
@@
 
-\hh{We need an additional sentence here to get from the data to the statistical methods ... }
+{\textcolor{orange}{We need an additional sentence here to get from the data to the statistical methods ... }}
+{\textcolor{blue}{Analysis of these digitized markings require the use of statistical methods which can quantify the scientific mechanism of comparing markings and serve as metrics for correctly identifying the error rate.}}
+
 \citet{manytoolmarks2} define a relative distance metric and use it as
@@
 identification.
+
+\begin{table}[ht]
+\centering
+\resizebox{\textwidth}{!}{\begin{tabular}{rlrrr}
+  \hline
+Research paper & Method & Data Source & False Positives & False Negatives\\
+  \hline
+ \textbf{\citet{manytoolmarks1}} & Maximized & Screwdrivers & & \\ & Correlation &  & - & - \\ \hline
+ \textbf{\citet{chumbley}} & Randomized & Screwdrivers & & \\ (Same-Surface Same-Angle) & Chumbley Score & & 2.3\% & 8.9\%  \\ \hline
+  \textbf{\citet{afte-chumbley}} & Randomized & Slip-joint  &  & \\ & Chumbley Score &  & - & - \\ \hline 
+  \textbf{\citet{hadler}} & Deterministic & Screwdrivers & & \\ (Same-Surface Same-Angle) & Chumbley Score & & 0\% & 6\%  \\ \hline
+  \textbf{\citet{manytoolmarks2}} & Similarity Measure & Screwdrivers & & \\ (Different Surfaces-same angle) & Relative Distance Metric & & 5.9\% &  9.4\% \\ (Same Surfaces-same angle)&  & & 0.22\% &  0\% \\
+   \hline
+\end{tabular}}
+\caption{\label{tab:toolmarks-ER} Error Rates for Toolmarks using  variations of the chumbley score method}
+\end{table}
 
@@
 
+\hypertarget{scans-for-land-engraved-areas}{%
 \subsection{Scans for land engraved
-areas}\label{scans-for-land-engraved-areas}
+areas}\label{scans-for-land-engraved-areas}}
 
@@
 
-\subsection{The Chumbley Score Test}\label{the-chumbley-score-test}
+\hypertarget{the-chumbley-score-test}{%
+\subsection{The Chumbley Score Test}\label{the-chumbley-score-test}}
 
@@
 \(t_1^o - t_2^o\) for aligning the two markings.
-\hh{We will denote the relative optimal locations as $t_1^*$ and $t_2^*$, where $t_i^* = t_i^o/(T_i-w_o)$ for $i=1,2$, such that $t_1^*, t_2^* \in [0,1]$.
+{\textcolor{orange}{We will denote the relative optimal locations as $t_1^*$ and $t_2^*$, where $t_i^* = t_i^o/(T_i-w_o)$ for $i=1,2$, such that $t_1^*, t_2^* \in [0,1]$.
 Once (sub-)class characteristics are removed, the relative optimal locations should be distributed according to a uniform distribution in $[0,1]$.
-}
+}}
 
@@
 More precisely, let us define starting points \(s_i^{(k)}\) for each
-signature \(k = 1, 2\) as
-
-\begin{eqnarray}\label{eq.start}
+signature \(k = 1, 2\) as \begin{eqnarray}\label{eq.start}
 s^{(k)}_i = 
@@
 \end{cases}
-\end{eqnarray}
-
-for integer values of \(i\) with \(0 < s^{(k)}_i \le T_k - w_v\).
+\end{eqnarray} for integer values of \(i\) with
+\(0 < s^{(k)}_i \le T_k - w_v\).
 
@@
 
+\hypertarget{a-problem-with-failed-tests}{%
 \subsection{A problem with failed
-tests}\label{a-problem-with-failed-tests}
+tests}\label{a-problem-with-failed-tests}}
 
@@
 number of tests will fail to produce a result.
-\hh{The problem of failed tests is first mentioned in \citet{afte-chumbley}. Unfortunately, the authors do not provide any percentage of how many tests failed for their data.}
+{\textcolor{orange}{The problem of failed tests is first mentioned in \citet{afte-chumbley}. Unfortunately, the authors do not provide any percentage of how many tests failed for their data.}}
 The algorithm fails to produce for two reasons: either the number of
@@
 
-\subsection{A modified approach}\label{a-modified-approach}
+\hypertarget{a-modified-approach}{%
+\subsection{A modified approach}\label{a-modified-approach}}
 
@@
 We then define the pairs for different-shifts as
-
 \begin{equation}\label{eq.diff2}
@@
 \end{cases},
-\end{equation}
-
-i.e.~for an odd number of same-shift correlations, we skip the middle
-pair for the different-shift correlations (see \autoref{sketch-diff-2}).
-This pairing ensures that the number of different-shift pairings is the
-same or at most one less than the number of same-shift pairings in all
-tests.
-\hh{In the remainder of the paper, we will refer to the algorithm defined by \citet{hadler} as {\bf (CS1)} and the suggested modified algorithm as {\bf (CS2)} and compare their performance on the available scans of the Hamby study.}
+\end{equation} i.e.~for an odd number of same-shift correlations, we
+skip the middle pair for the different-shift correlations (see
+\autoref{sketch-diff-2}). This pairing ensures that the number of
+different-shift pairings is the same or at most one less than the number
+of same-shift pairings in all tests.
+{\textcolor{orange}{In the remainder of the paper, we will refer to the algorithm defined by \citet{hadler} as{\bf (CS1)} and the suggested modified algorithm as{\bf (CS2)} and compare their performance on the available scans of the Hamby study.}}
 
@@
 
-\section{Testing setup}\label{testing-setup}
+\begin{figure}[hbtp]
+\centering
+\includegraphics[width=0.62\textwidth]{images/same-shift-failure.png}
 
-\subsection{The Data}\label{the-data}
+\caption{\label{same-shift-failure}Sketch of same-shift pairings  (top) when the lag is too large to accomodate a a vaildation window in either of the two signatures}
+\end{figure}
+
+\begin{figure}[hbtp]
+\centering
+\includegraphics[width=0.7\textwidth]{images/diff-shift-failure.png}
+
+\caption{\label{diff-shift-failure}Sketch of diff-shift pairings  (top) when the number of diff-shift computations is likely to 0}
+\end{figure}
+
+\begin{figure}[hbtp]
+\centering
+\includegraphics[width=.7\textwidth]{images/CS1-failure-CS2-success.png}
+\caption{\label{CS1-failure-CS2-success}Sketch of a case where CS1 fails but CS2 does not fail}
+\end{figure}
+
+\hypertarget{testing-setup}{%
+\section{Testing setup}\label{testing-setup}}
+
+\hypertarget{the-data}{%
+\subsection{The Data}\label{the-data}}
 
@@
 
-\subsection{Setup}\label{setup}
+\hypertarget{setup}{%
+\subsection{Setup}\label{setup}}
 
@@
 
-\section{Results}\label{results}
+\hypertarget{results}{%
+\section{Results}\label{results}}
 
-\subsection{Failed Tests}\label{failed-tests}
+\hypertarget{failed-tests}{%
+\subsection{Failed Tests}\label{failed-tests}}
 
@@
 \(w_v\) of 60 pixels.
-\hh{Rates of expected failures are based on simulation runs using covariances between locations of same-source profiles of 0.854, and 0.120 for locations from different-source profiles, matching observed covariances for the Hamby scans.
-}\hh{ Observed failure rates are higher than expected.
+{\textcolor{orange}{Rates of expected failures are based on simulation runs using covariances between locations of same-source profiles of 0.854, and 0.120 for locations from different-source profiles, matching observed covariances for the Hamby scans.
+}}{\textcolor{orange}{ Observed failure rates are higher than expected.
 This might be due to remaining sub-class structure at a coarseness of 0.3 resulting in a distribution of optimal locations different from the assumed uniform.
-}
+}}
 
@@
 
-\subsection{Coarseness}\label{coarseness}
+\hypertarget{coarseness}{%
+\subsection{Coarseness}\label{coarseness}}
 
@@
 
-\hh{\autoref{fig:profile-sketch} gives an overview of the effect of different coarseness parameters: from left to right, coarseness levels $c$ are varied in steps of 0.05 from 0.1 to 0.3.
+{\textcolor{orange}{\autoref{fig:profile-sketch} gives an overview of the effect of different coarseness parameters: from left to right, coarseness levels $c$ are varied in steps of 0.05 from 0.1 to 0.3.
 The top row shows  resulting signature after smoothing the profile shown in Figure~\ref{fig:sigs-profiles} with the coarseness specified.
@@
 Optimal locations are distributed uniformly once (sub-)class characteristics are removed.
-However, for coarseness values of $c > 0.20$ we see quite distinct boundary effects: optimal locations $t^*$ are found at the very extreme ends of a profile more often than one would expect based on a uniform distribution.}
+However, for coarseness values of $c > 0.20$ we see quite distinct boundary effects: optimal locations $t^*$ are found at the very extreme ends of a profile more often than one would expect based on a uniform distribution.}}
+
+{\textcolor{blue}{The key effect of the optimal locations and thereby the coarseness is seen in the number of failed tests.
+Irrespective of whether CS1 or CS2 is being used, if the relative optimal locations are at the boundaries we will see an increase in the number of failed tests.
+A balance is therefore needed in the selection of the coarseness parameter which reduces the boundary effect but also does not remove important (sub-)class characteristics}}
 
@@
 
-\subsection{Type II error rates}\label{type-ii-error-rates}
+\hypertarget{type-ii-error-rates}{%
+\subsection{Type II error rates}\label{type-ii-error-rates}}
 
@@
 
+\hypertarget{observed-versus-nominal-type-i-error-rates}{%
 \subsection{Observed versus Nominal Type I error
-rates}\label{observed-versus-nominal-type-i-error-rates}
+rates}\label{observed-versus-nominal-type-i-error-rates}}
 
@@
 
-\section{Conclusions}\label{conclusions}
+\hypertarget{conclusions}{%
+\section{Conclusions}\label{conclusions}}
 
-\hh{
+{\textcolor{orange}{
 We started out trying to assess the suitability of the Chumbley Score's suitability for matching striae on bullet lands.
@@
 We might be able to get more power out of the test by adapting CS2 to deal work in a bullet-to-bullet comparison.
-}
+}}
 
-\hh{
+{\textcolor{orange}{
 Fix in CS2 should also improve power for matching toolmarks. 
-}
+}}
 

### 2018-05-21 doc/paper.Rmd
starting proofing

@@
 ## Motivation
-Same source analyses are a major part of an Forensic Toolmark Examiner's job.
-In current practice examiners  make these comparisons by visual inspection under a comparison microscope and come to one of the following four conclusions: identification, inconclusive, elimination or unsuitable for examination~\citep{afte-toolmarks1998}.
+Same source analyses are a major part of a Forensic Toolmark Examiner's job.
+In current practice, examiners  make these comparisons by visual inspection under a comparison microscope and come to one of the following four conclusions: identification, inconclusive, elimination or unsuitable for examination~\citep{afte-toolmarks1998}.
 These conclusions are made on the basis of "unique surface contours" of the two toolmarks being in "sufficient agreement" \citep{afte-toolmarks1998}.
@@
 
-\hh{In the next page and a half it is easy to lose the red line. It might help to include a table with an overview.
-The table should include the reference to the paper, the data used, the statistical method and the associated error rates.}
-\gk{table included}
-Various toolmarks have been studied in the literature:  \citet{manytoolmarks1} and \citet{chumbley} have been analyzing screwdriver marks  digitized using a profilometer;  \citet{manytoolmarks2} have investigated 3D marks from screwdriver, tongue and groove pliers captured using a confocal microscope; \citet{afte-chumbley} have been investigated digitized marks from slip-joint pliers generated by a surface profilometer.
-
-\hh{We need an additional sentence here to get from the data to the statistical methods ... }
-\gk{Analysis of these digitized markings require the use of statistical methods which can quantify the scientific mechanism of comparing markings and serve as metrics for correctly identifying the error rate.}
+Various toolmarks have been studied in the literature (see \autoref{tab:toolmarks-ER} for an overview):  \citet{manytoolmarks1} and \citet{chumbley} have been analyzing screwdriver marks  digitized using a profilometer;  \citet{manytoolmarks2} have investigated 3D marks from screwdriver, tongue and groove pliers captured using a confocal microscope; \citet{afte-chumbley} have been investigated digitized marks from slip-joint pliers generated by a surface profilometer.
+
+Analysis of these digitized markings require the use of statistical methods which can quantify the scientific mechanism of comparing markings and serve as basis for an error rate calcluation.
 \citet{manytoolmarks2} define a relative distance metric and use it as similarity measure between two toolmarks.
@@
 \end{tabular}}
-\caption{\label{tab:toolmarks-ER} Error Rates for Toolmarks using  variations of the chumbley score method}
+\caption{\label{tab:toolmarks-ER} Error Rates in same-source toolmark analysis reported in the literature. All reported papers use some variation of the Chumbley-score method.}
 \end{table}

### 2018-05-21 doc/paper.tex
starting proofing

@@
 \noindent%
-{\it Keywords:} 3 to 6 keywords, that do not appear in the title
+{\it Keywords:} forensic science, toolmark, cross-correlation, Mann-Whitney U statistic, , 
 \vfill
@@
 
-Same source analyses are a major part of an Forensic Toolmark Examiner's
-job. In current practice examiners make these comparisons by visual
+Same source analyses are a major part of a Forensic Toolmark Examiner's
+job. In current practice, examiners make these comparisons by visual
 inspection under a comparison microscope and come to one of the
@@
 
-{\textcolor{orange}{In the next page and a half it is easy to lose the red line. It might help to include a table with an overview.
-The table should include the reference to the paper, the data used, the statistical method and the associated error rates.}}
-{\textcolor{blue}{table included}}
+Various toolmarks have been studied in the literature (see
+\autoref{tab:toolmarks-ER} for an overview): \citet{manytoolmarks1} and
+\citet{chumbley} have been analyzing screwdriver marks digitized using a
+profilometer; \citet{manytoolmarks2} have investigated 3D marks from
+screwdriver, tongue and groove pliers captured using a confocal
+microscope; \citet{afte-chumbley} have been investigated digitized marks
+from slip-joint pliers generated by a surface profilometer.
 
-Various toolmarks have been studied in the literature:
-\citet{manytoolmarks1} and \citet{chumbley} have been analyzing
-screwdriver marks digitized using a profilometer; \citet{manytoolmarks2}
-have investigated 3D marks from screwdriver, tongue and groove pliers
-captured using a confocal microscope; \citet{afte-chumbley} have been
-investigated digitized marks from slip-joint pliers generated by a
-surface profilometer.
-
-{\textcolor{orange}{We need an additional sentence here to get from the data to the statistical methods ... }}
-{\textcolor{blue}{Analysis of these digitized markings require the use of statistical methods which can quantify the scientific mechanism of comparing markings and serve as metrics for correctly identifying the error rate.}}
-
+Analysis of these digitized markings require the use of statistical
+methods which can quantify the scientific mechanism of comparing
+markings and serve as basis for an error rate calcluation.
 \citet{manytoolmarks2} define a relative distance metric and use it as
@@
 Irrespective of whether CS1 or CS2 is being used, if the relative optimal locations are at the boundaries we will see an increase in the number of failed tests.
-A balance is therefore needed in the selection of the coarseness parameter which reduces the boundary effect but also does not remove important (sub-)class characteristics}}
+A balance is therefore needed in the selection of the coarseness parameter which reduces the boundary effect but also does not remove important individual characteristics}}
 

### 2018-05-21 code/hamby44-evaluate.R

@@ -1,0 +1,46 @@
+library(tidyverse)
+source("code/roc-function.R") # for EER and AUC
+files <- dir("data/hamby44/", pattern="rds", recursive=TRUE)
+
+h44 <- data.frame()
+for (file in files) {
+  tmp <- readRDS(file.path("data/hamby44", file))
+  tmp$source <- file
+  tmp <- tmp %>% mutate(
+    match = replace(match, is.na(match), FALSE)
+  )
+  h44 <- rbind(h44, tmp)
+}
+
+h44 <- h44 %>% separate(source, into=c("method", "rest"), sep="/") %>%
+  select(-rest)
+
+h44.summ <- h44 %>% group_by(wo, wv, method, coarseness) %>%
+  mutate(
+    fails = sum(is.na(p_value))/n() 
+  ) %>% filter(!is.na(p_value)) %>%
+  summarize(
+    auc = AUC(1-p_value, match),
+    eer = EER(1-p_value, match),
+    ci = list(AUC_confint(1-p_value, match)),
+    fails = fails[1],
+    roc = list(roc(match, 1-p_value))
+  )
+h44.summ <- h44.summ %>% mutate(
+  lower = ci %>% purrr::map_dbl(.f = function(x) x[[1]][1]),
+  upper = ci %>% purrr::map_dbl(.f = function(x) x[[1]][2])
+)
+
+
+h44.summ %>% gather(type, value, auc, eer) %>% 
+  filter(wv %in% c(75, 125), coarseness==0.125,
+         between(wo, 210, 390)) %>% 
+  ggplot(aes(x = wo, y = value, colour=method, shape=factor(wv)))  +
+  geom_point(size=2.5) +
+#  geom_point(aes(y = lower), size=1) +
+#  geom_point(aes(y = upper), size=1) +
+  facet_wrap(~type, scales="free") +
+  theme_bw() +
+  scale_colour_brewer(palette="Set1") +
+  geom_smooth(aes(linetype = factor(wv), colour=method), se=FALSE, span=2) +
+  xlab("Size of optimization window")

### 2018-05-21 doc/paper.Rmd
edits up to Results

@@
 ## Motivation
-Same source analyses are a major part of a Forensic Toolmark Examiner's job.
-In current practice, examiners  make these comparisons by visual inspection under a comparison microscope and come to one of the following four conclusions: identification, inconclusive, elimination or unsuitable for examination~\citep{afte-toolmarks1998}.
+Same-source analyses are a major part of a Forensic Toolmark Examiner's job.
+In current practice, examiners  make these comparisons by visual inspection under a comparison microscope and come to one of the following four conclusions: identification, inconclusive, elimination or unsuitable for examination \citep{afte-toolmarks1998}.
 These conclusions are made on the basis of "unique surface contours" of the two toolmarks being in "sufficient agreement" \citep{afte-toolmarks1998}.
@@
 
-Various toolmarks have been studied in the literature (see \autoref{tab:toolmarks-ER} for an overview):  \citet{manytoolmarks1} and \citet{chumbley} have been analyzing screwdriver marks  digitized using a profilometer;  \citet{manytoolmarks2} have investigated 3D marks from screwdriver, tongue and groove pliers captured using a confocal microscope; \citet{afte-chumbley} have been investigated digitized marks from slip-joint pliers generated by a surface profilometer.
+Methods for matching marks for a variety of tools have been studied in the literature (see \autoref{tab:toolmarks-ER} for an overview):  \citet{manytoolmarks1} and \citet{chumbley} have been analyzing screwdriver marks  digitized using a profilometer;  \citet{manytoolmarks2} have investigated 3D marks from screwdriver, tongue and groove pliers captured using a confocal microscope; \citet{afte-chumbley} have been investigated digitized marks from slip-joint pliers generated by a surface profilometer.
 
@@
 Striation marks on bullets are made by impurities in the barrel.
-As the bullet travels through the barrel, these imperfections leave "scratches" on the bullet surface.
-Typically, only striation marks in the land engraved areas (LEAs) are considered \citet{afte-article1992}.
+As the bullet travels through the barrel, these imperfections leave "scratches" on the bullet surface (see top of \autoref{fig:rgl}).
+Typically, only striation marks in the land engraved areas (LEAs) are considered \citep{afte-article1992}.
 Bullet lands are depressed areas between the  grooves made by the rifling action of the barrel.
@@
 Bullets also have a curved cross-sectional topography.
-Figure \ref{fig:rgl} shows us how the signature from a bullet land (bottom) lines up with the image of the land (top) from which it was extracted.
-We can also see in the figure how the depth and relative position of the striation markings seen in the image are interpreted as the signature.
-
-Bullet matching methods are usually based on these associated signatures.
+
+<!--\hh{signatures haven't been introduced yet}-->
+<!--Figure \ref{fig:rgl} shows us how the signature from a bullet land (bottom) lines up with the image of the land (top) from which it was extracted.
+We can also see in the figure how the depth and relative position of the striation markings seen in the image are interpreted as the signature.-->
+
+<!--Bullet matching methods are usually based on these associated signatures.-->
+In same-source comparisons this curvature is usually removed using some form of Gaussian filter \citep{ma2004} or non-parametric smoothing \citep{aoas}. 
 \citet{chu2013} use an automatic method for counting consecutive matching striae (CMS).
-The authors report an error rate of 52% of the known same source lands comparisons as misidentified (false negative) and zero false positives for known different source lands.
+The authors report an error rate of 52% for known same-source land comparisons to be (incorrectly) identified as different-source (false negative) and zero false positives for known different-source lands.
 \citet{ma2004} and \citet{vorburger2011} discuss CCF (cross-correlation function) and its discriminating power and applicability for same-source analyses of bullets, but do not provide any error rates in their discussion.
-\citet{aoas} use multiple features like CCF, CMS, D (distance measure) etc in a random forest based method and compare every land against every other land of  digitised versions of Hamby 252 and Hamby 44 \citep{hamby} published on the NIST Ballistics Database \citep{nist}.
+\citet{aoas} use multiple features, such as CCF, CMS, D (distance measure), etc.
+in a random forest based method and compare every land against every other land of  digitised versions of Hamby 252 and Hamby 44 \citep{hamby} published on the NIST Ballistics Database \citep{nist}.
 The authors report an out-of-bag overall error rate of 0.46%, comprised of an error rate of 30.05% of  same-source pairs that were not identified and an error rate of 0.026% of different-source pairs that were incorrectly identified as same-source.
@@
 For markings made at 30 degree the authors  report an average false negative error rate of 0.089  and an average false positive error rate of 0.023.
-For other angles of 60 and 85 degrees the false negatives error rate is 0.09 while the rate of false positives decreases to 0.01.
+For marks made under angles of 60 and 85 degrees, respectively, the false negatives error rate is 0.09 while the rate of false positives decreases to 0.01.
 The paper by \citet{hadler} is based on the same data but the authors focus on  markings made under the same angle.
-The error rates associated with the deterministic version of the score are  0.06 for false  negatives and a false positive error rate of 0.
+The error rates associated with the deterministic version of the score are reported as  0.06 for false  negatives and  0 for false positives.
 
@@
 This effectively removes topographic structure from the data in the attempt to increase the signal to noise ratio.
-The span of the loess fit was found using cross-validation, as described by \citet{aoas}.
+\autoref{fig:rgl}  shows how the signature from a bullet land (bottom) lines up with the image of the land (top) from which it was extracted.
+We can see in the figure how the depth and relative position of the striation markings seen in the image are interpreted as peaks and valleys in the signature.
+<!--The span of the loess fit was found using cross-validation, as described by \citet{aoas}.-->
 
 There are two sources of scans for sets from the Hamby study available to us: scans of Hamby 44 and Hamby 252 are available from the NIST database \citep{nist}.
-Hamby 44 has also been made available to us and has been scanned locally for CSAFE at the Roy J.\ Carver High Resolution Microscopy Facility using a Sensofar confocal light microscope.
+The physical Hamby 44 set has also been made available to us and has been scanned locally for CSAFE at the Roy J.\ Carver High Resolution Microscopy Facility using a Sensofar confocal light microscope.
 Scans in the NIST database are made with a NanoFocus at 20x magnification.
@@
 
-The Chumbley score algorithm takes input in form of two digitized toolmarks.
-The toolmark is in form of $z(t)$ which is a spatial process for location indexed by $t$.
-$t$ here denotes equally spaced pixel locations for the striation marks under consideration, $t = 1, ..., T$.
+A digitized toolmark forms a spatial process $z(t)$ with location indexed by $t$.
+$t$ here denotes equally spaced pixel locations for the striation marks under consideration.
+For a toolmark consisting of $t$ pixels, $t = 1, ..., T$.
  Let further $z^s(t)$ denote a vector of markings of length $s$ starting in location  $t$.
 
-Let $x(t_1)$, $t_1 = 1,2,...T_1$ and $y(t_2)$, $t_2 = 1,2...T_2$ be two digitized toolmarks (where $T_1$ and $T_2$ are not necessarily equal). 
+The Chumbley score algorithm takes input in form of two digitized toolmarks:
+
+Let $x(t_1)$, $t_1 = 1,2,...T_1$ and $y(t_2)$, $t_2 = 1,2...T_2$ be two digitized toolmarks (where $T_1$ and $T_2$, the lengths of the two marks, are not necessarily equal). 
 <!--The vectorized processes representing two sets of striaes are therefore shown as $x(t_1)$, $t_1 = 1,2,...T_1$ and $y(t_2)$, $t_2 = 1,2...T_2$.
@@
 This representation of $t$ as $t_1$ and $t_2$ lets us identify them as either of same or different lengths.-->
-The toolmarks under consideration are potentially from two different sources or the same source.
-$T_1$ and $T_2$, as represented above, are the final pixel indexes of each marking and therefore give the respective lengths of the markings.
+The toolmarks under consideration are potentially from two different-sources or the same-source.
+<!--$T_1$ and $T_2$, as represented above, are the final pixel indexes of each marking and therefore give the respective lengths of the markings.
+-->
+
+```{r sigs-profiles, fig.width = 8, fig.height = 2.5, fig.cap="Bullet land profile (left) and the corresponding signature (right) for one of the lands of Hamby-44."}
+sigs_graphics <- readRDS("../data/sigs_generate_window.rds")
+sigs_graphics$sigs_land_id_1_match <- sigs_graphics$sigs_land_id_1_match %>% group_by(land_id) %>% mutate(y = y-min(y, na.rm=T))
+
+sigs_graphics.match<- as.data.frame(sigs_graphics[1])
+sigs_graphics.nonmatch<- as.data.frame(sigs_graphics[2])
+colnames(sigs_graphics.match)<- gsub("^.*\\.","", colnames(sigs_graphics.match))
+colnames(sigs_graphics.nonmatch)<- gsub("^.*\\.","", colnames(sigs_graphics.nonmatch))
+
+p1 <- sigs_graphics[[1]] %>% filter(land_id==19) %>%
+  ggplot(aes(x = y, y = value)) + geom_line() +
+  theme_bw() +
+  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
+  xlab(expression(paste("Relative location (in ",mu,m,")", sep=""))) + 
+  ggtitle("Profile")
+
+p2 <- sigs_graphics[[1]] %>% filter(land_id==19) %>%
+  ggplot(aes(x = y, y = l30)) + 
+  geom_line(aes(y=resid), size=.25) +
+  geom_line(aes(y=l30)) +
+  theme_bw() +
+  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
+  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
+  ggtitle("Signature")
+
+grid.arrange(p1,p2, ncol=2)
+```
 
@@
  Originally, this smoothing is intended to remove drift and (sub)class characteristics from individual markings, however, in the setting of matching bullet striae, we can also make use of this mechanism to separate bullet curvature in profiles from signatures before matching signatures.
+\autoref{fig:sigs-profiles} shows an example of a bullet land profile (left) and the corresponding signature (right).
 
 After removing sub-class structure, 
-the Chumbley scores is calculated in two steps: an optimization step and a validation step.
+the Chumbley score is calculated in two steps: an optimization step and a validation step.
 In the optimization step, the two markings are aligned horizontally such that within a pre-defined window of length $w_o$ the correlation between $x(t_1)$ and $y(t_2)$ is maximized:
 \[
-\left(t_1^o, t_2^o\right) = \mathop{\arg \max}\limits_{1 \le t_1 \le T_1, 1 \le t_2 \le T_2} \text{cor} \left(x^{w_o} (t_1), y^{w_o}(t_2) \right)
+\left(t_1^o, t_2^o\right) = \mathop{\arg \max}\limits_{1 \le t_1 \le T_1-w_o, 1 \le t_2 \le T_2-w_o} \text{cor} \left(x^{w_o} (t_1), y^{w_o}(t_2) \right)
 \]
 This results in an optimal vertical (in-phase) shift of $t_1^o - t_2^o$ for aligning the two markings.
-\hh{We will denote the relative optimal locations as $t_1^*$ and $t_2^*$, where $t_i^* = t_i^o/(T_i-w_o)$ for $i=1,2$, such that $t_1^*, t_2^* \in [0,1]$.
+We will denote the relative optimal locations as $t_1^*$ and $t_2^*$, where $t_i^* = t_i^o/(T_i-w_o)$ for $i=1,2$, such that $t_1^*, t_2^* \in [0,1]$.
 Once (sub-)class characteristics are removed, the relative optimal locations should be distributed according to a uniform distribution in $[0,1]$.
-}
-
-In the validation step,  two sets of windows of size $w_v$ are chosen from both markings (see Figure \ref{fig:win-comparison}).
+
+In the validation step,  two sets of windows of size $w_v$ are chosen from both markings,  \autoref[see][]{fig:win-comparison}.
 In the first set, pairs of windows are extracted from the two markings using the optimal vertical shift as determined in the first step, whereas for the second set the windows are extracted using a different (out-of-phase) shift.
@@
 Same-shift pairs of length $w_v$ are defined in \citet{hadler} as all 
-pairs $(s_i^{(1)}, s_i^{(2)})$ for all integer values $i$ for which both $s_i^{(1)}$ and  $s_i^{(2)}$ are defined.
+pairs $(s_i^{(1)}, s_i^{(2)})$ with integer values $i$ for which both $s_i^{(1)}$ and  $s_i^{(2)}$ are defined.
 Similarly, different-shift pairs are defined as $(s_i^{(1)}, s_{-i-1}^{(2)})$ for all $i$ where both $s_i^{(1)}$ and  $s_{-i-1}^{(2)}$ are defined (see \autoref{sketch-same-diff}).
@@
 
-```{r win-comparison, warning=FALSE, fig.width=8, fig.height=3.5, fig.cap="Two markings made by the same source.
+```{r win-comparison, warning=FALSE, fig.width=8, fig.height=3.5, fig.cap="Two same-source markings.
 For convenience, the markings are moved into phase on the left and out-of phase on the right.
@@
 The Chumbley-score is based on a Mann-Whitney U test of the correlations derived from these two sets of samples."}
-sigs_graphics <- readRDS("../data/sigs_generate_window.rds")
-sigs_graphics$sigs_land_id_1_match <- sigs_graphics$sigs_land_id_1_match %>% group_by(land_id) %>% mutate(y = y-min(y, na.rm=T))
-
-sigs_graphics.match<- as.data.frame(sigs_graphics[1])
-sigs_graphics.nonmatch<- as.data.frame(sigs_graphics[2])
-colnames(sigs_graphics.match)<- gsub("^.*\\.","", colnames(sigs_graphics.match))
-colnames(sigs_graphics.nonmatch)<- gsub("^.*\\.","", colnames(sigs_graphics.nonmatch))
 
@@
 For both same- and different-shift pairs correlations between the  markings are calculated. 
-The intuition here is that for two markings from the same source the correlation for the in-phase sample should be high, while the correlations of the out-of-phase sample  provide a measure for the base-level correlation for non-matching marks of a given length $w_v$.
+The intuition here is that for two markings from the same-source the correlation for the in-phase sample should be high, while the correlations of the out-of-phase sample  provide a measure for the base-level correlation for non-matching marks of a given length $w_v$.
 The Chumbley score is then computed as a Mann Whitney U statistic to compare between in-phase sample and out-of-phase sample.
@@
 Looking closer at \autoref{eq.start}, we see that by definition, some number of tests will fail to produce a result.
-\hh{The problem of failed tests is first mentioned in \citet{afte-chumbley}.
-Unfortunately, the authors do not provide any percentage of how many tests failed for their data.}
+The problem of failed tests is first mentioned in \citet{afte-chumbley}.
+Unfortunately, the authors do not provide any percentage of how many tests failed for their data.
 The algorithm fails to produce for two reasons: either the number of eligible same-shift pairs is zero, or the number of different-shift pairs is zero.
@@
 In particular, for same-source profiles, we would expect a strong dependency between these locations, in which case a large difference between locations is unlikely.
-However, for different source matches, we can assume that  locations are independent.
+However, for different-source matches, we can assume that  locations are independent.
 In that case, we expect a test to fail with a probability of $\frac{2 w_v^2}{(T_1-w_o)(T_2-w_o)}$.
@@
 Assuming a correlation between optimal locations $t_1^o$ and $t_2^o$ of close to one for same-source profiles, this results in an expected rate of failure of $2 w_v / (T_i-w_o)$, or about 5.6\% for an average length of $T_i$ of 1200 pixels, $w_o = 120$ pixels and $w_v = 30$ pixels.
-Assuming independence in the optimal locations for different source profiles the expected probability for a failed test is, again, $\frac{2 w_v^2}{(T_1-w_o)(T_2-w_o)}$.
+Assuming independence in the optimal locations for different-source profiles the expected probability for a failed test is, again, $\frac{2 w_v^2}{(T_1-w_o)(T_2-w_o)}$.
 
@@
 This pairing ensures that the number of different-shift pairings is the same or at most one less than the number of same-shift pairings in all tests.
-\hh{In the remainder of the paper, we will refer to the algorithm defined by \citet{hadler} as {\bf (CS1)} and the suggested modified algorithm as {\bf (CS2)} and compare their performance on the available scans of the Hamby study.}
+In the remainder of the paper, we will refer to the algorithm defined by \citet{hadler} as \textbf{(CS1)} and the suggested modified algorithm as \textbf{(CS2)} and compare their performance on the available scans of the Hamby study.
 \begin{figure}[hbtp]
@@
 \includegraphics[width=.7\textwidth]{images/sketch-diff-2.png}
-\caption{\label{sketch-diff-2}Sketch of adjusted different-shift pairings. At most one of the same-shift pairings can not be matched with a different-shift pair. }
+\caption{\label{sketch-diff-2}Sketch of adjusted different-shift pairings. At most one of the same-shift pairings can not be matched with a different-shift pair under algorithm (CS2). }
 \end{figure}
 
-\begin{figure}[hbtp]
-\centering
-\includegraphics[width=0.62\textwidth]{images/same-shift-failure.png}
-
-\caption{\label{same-shift-failure}Sketch of same-shift pairings  (top) when the lag is too large to accomodate a a vaildation window in either of the two signatures}
-\end{figure}
-
-\begin{figure}[hbtp]
-\centering
-\includegraphics[width=0.7\textwidth]{images/diff-shift-failure.png}
-
-\caption{\label{diff-shift-failure}Sketch of diff-shift pairings  (top) when the number of diff-shift computations is likely to 0}
-\end{figure}
-
-\begin{figure}[hbtp]
-\centering
-\includegraphics[width=.7\textwidth]{images/CS1-failure-CS2-success.png}
-\caption{\label{CS1-failure-CS2-success}Sketch of a case where CS1 fails but CS2 does not fail}
-\end{figure}
-
 # Testing setup
@@
 ## The Data
-
-```{r sigs-profiles, fig.width = 8, fig.height = 2.5, fig.cap="Bullet land profile (left) and the corresponding signature (right) for one of the lands of Hamby-44."}
-p1 <- sigs_graphics[[1]] %>% filter(land_id==19) %>%
-  ggplot(aes(x = y, y = value)) + geom_line() +
-  theme_bw() +
-  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
-  xlab(expression(paste("Relative location (in ",mu,m,")", sep=""))) + 
-  ggtitle("Profile")
-
-p2 <- sigs_graphics[[1]] %>% filter(land_id==19) %>%
-  ggplot(aes(x = y, y = l30)) + 
-  geom_line(aes(y=resid), size=.25) +
-  geom_line(aes(y=l30)) +
-  theme_bw() +
-  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
-  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
-  ggtitle("Signature")
-
-grid.arrange(p1,p2, ncol=2)
-```
 
@@
 ## Setup
-Both algorithms (CS1) and (CS2) are implemented in the R package `toolmaRk` \citep{toolmark}.
+Both algorithms (CS1) and (CS2) are implemented in R \citep{R}.
+(CS1) is available from package `toolmaRk` \citep{toolmark}, (CS2) is available from a modified version of the `toolmaRk` package available from GitHub (\url{https://github.com/heike/toolmaRk}).
 We applied both methods to all pairwise land-to-land comparisons of the Hamby scans provided by NIST for a total of 85,491 land-to-land comparisons.
@@
 For same-source lands up to 12.5 percent of the tests fail using CS1.
-The highest percentage of failed tests under CS2 is 1.3% for different source tests using a validation window size $w_v$ of 60 pixels.
+The highest percentage of failed tests under CS2 is 1.3% for different-source tests using a validation window size $w_v$ of 60 pixels.
 \hh{Rates of expected failures are based on simulation runs using covariances between locations of same-source profiles of 0.854, and 0.120 for locations from different-source profiles, matching observed covariances for the Hamby scans.
@@
 
-\hh{\autoref{fig:profile-sketch} gives an overview of the effect of different coarseness parameters: from left to right, coarseness levels $c$ are varied in steps of 0.05 from 0.1 to 0.3.
-The top row shows  resulting signature after smoothing the profile shown in Figure~\ref{fig:sigs-profiles} with the coarseness specified.
+\autoref{fig:profile-sketch} gives an overview of the effect of different coarseness parameters: from left to right, coarseness levels $c$ are varied in steps of 0.05 from 0.1 to 0.3.
+The top row shows  resulting signatures after smoothing the profile shown in Figure~\ref{fig:sigs-profiles} with different levels of coarseness.
 The histograms in the bottom row show the relative optimal location $t^*$.
 Optimal locations are distributed uniformly once (sub-)class characteristics are removed.
-However, for coarseness values of $c > 0.20$ we see quite distinct boundary effects: optimal locations $t^*$ are found at the very extreme ends of a profile more often than one would expect based on a uniform distribution.}
+However, for coarseness values of $c > 0.20$ we see quite distinct boundary effects: optimal locations $t^*$ are found at the very extreme ends of a profile more often than one would expect based on a uniform distribution.
 
@@
 A Test Fail, i.e.
-tests in which a particular parameter setting did not return a valid result, happens, when the shift to align two signatures is so large, that the remaining overlap is too small to accommodate  windows for validation, depends on whether known-match or known non-matches has predictive value, with test results from different sources having a much higher chance to fail.
+tests in which a particular parameter setting did not return a valid result, happens, when the shift to align two signatures is so large, that the remaining overlap is too small to accommodate  windows for validation, depends on whether known-match or known non-matches has predictive value, with test results from different-sources having a much higher chance to fail.
 On conducting an analysis of all known bullet lands using the adjusted chumbley algorithm, Type II error was identified to be least bad for window of validation 30 and window of optimization 120.
@@
 -->
+
+# Appendix
+\begin{appendix}
+
+\section{Scenarios of failed tests}
+
+Figures \ref{same-shift-failure} and \ref{diff-shift-failure} show scenarios in which the deterministic Chumbley score can fail. 
+In \autoref{same-shift-failure} both CS1 and CS2 fail, because the lag between optimal locations is so large that no same-shift pairs can be found once the signatures are aligned.
+These failures are inherent to the Chumbley Score and cannot be prevented.
+\begin{figure}[hbtp]
+\centering
+\includegraphics[width=0.62\textwidth]{images/same-shift-failure.png}
+
+\caption{\label{same-shift-failure}Sketch of same-shift pairings.
+When the lag between optimal reference points is too large to accommodate a validation window in noth signatures, both CS1 and CS2 fail.}
+\end{figure}
+
+\autoref{diff-shift-failure} shows two situations, in which CS1 fails to identify any different-shift pairings.
+This happens, when both relative locations are close to either end of the signature.
+CS2 can still be computed in this situation as long as there are at least two same-shift pairs identified.
+
+\begin{figure}[hbtp]
+\centering
+\includegraphics[width=0.7\textwidth]{images/diff-shift-failure.png}
+
+\caption{\label{diff-shift-failure}Sketch of different-shift pairings. For CS1 no different-shift pairs can be identified, resulting in a failed test.}
+\end{figure}
+
+<!--
+\begin{figure}[hbtp]
+\centering
+\includegraphics[width=.7\textwidth]{images/CS1-failure-CS2-success.png}
+\caption{\label{CS1-failure-CS2-success}Sketch of a case where CS1 fails but CS2 does not fail.}
+\end{figure}
+-->
+
+\end{appendix}

### 2018-05-21 doc/paper.tex
edits up to Results

@@
 
-Same source analyses are a major part of a Forensic Toolmark Examiner's
+Same-source analyses are a major part of a Forensic Toolmark Examiner's
 job. In current practice, examiners make these comparisons by visual
@@
 following four conclusions: identification, inconclusive, elimination or
-unsuitable for examination\textasciitilde{}\citep{afte-toolmarks1998}.
-These conclusions are made on the basis of ``unique surface contours''
-of the two toolmarks being in ``sufficient agreement''
-\citep{afte-toolmarks1998}. AFTE describes the term ``sufficient
-agreement'' as the possibility of another tool producing the markings
-under comparison, as practically impossible \citep{afte-toolmarks1998}.
-Potential subject bias in the assessment as well as the lack of
-specified error rates are the main points of criticisms first raised by
-the National Research Council in 2009 \citep{NAS:2009} and later
-emphasized further by the President's Council of Advisors on Science and
-Technology \citep{pcast2016}.
+unsuitable for examination \citep{afte-toolmarks1998}. These conclusions
+are made on the basis of ``unique surface contours'' of the two
+toolmarks being in ``sufficient agreement'' \citep{afte-toolmarks1998}.
+AFTE describes the term ``sufficient agreement'' as the possibility of
+another tool producing the markings under comparison, as practically
+impossible \citep{afte-toolmarks1998}. Potential subject bias in the
+assessment as well as the lack of specified error rates are the main
+points of criticisms first raised by the National Research Council in
+2009 \citep{NAS:2009} and later emphasized further by the President's
+Council of Advisors on Science and Technology \citep{pcast2016}.
 
@@
 
-Various toolmarks have been studied in the literature (see
-\autoref{tab:toolmarks-ER} for an overview): \citet{manytoolmarks1} and
-\citet{chumbley} have been analyzing screwdriver marks digitized using a
-profilometer; \citet{manytoolmarks2} have investigated 3D marks from
-screwdriver, tongue and groove pliers captured using a confocal
-microscope; \citet{afte-chumbley} have been investigated digitized marks
-from slip-joint pliers generated by a surface profilometer.
+Methods for matching marks for a variety of tools have been studied in
+the literature (see \autoref{tab:toolmarks-ER} for an overview):
+\citet{manytoolmarks1} and \citet{chumbley} have been analyzing
+screwdriver marks digitized using a profilometer; \citet{manytoolmarks2}
+have investigated 3D marks from screwdriver, tongue and groove pliers
+captured using a confocal microscope; \citet{afte-chumbley} have been
+investigated digitized marks from slip-joint pliers generated by a
+surface profilometer.
 
@@
 bullet travels through the barrel, these imperfections leave
-``scratches'' on the bullet surface. Typically, only striation marks in
-the land engraved areas (LEAs) are considered \citet{afte-article1992}.
-Bullet lands are depressed areas between the grooves made by the rifling
-action of the barrel. Compared to toolmarks made by screwdrivers
-striation marks on bullets are typically much smaller, both in length
-and in width. Bullets also have a curved cross-sectional topography.
-Figure \ref{fig:rgl} shows us how the signature from a bullet land
-(bottom) lines up with the image of the land (top) from which it was
-extracted. We can also see in the figure how the depth and relative
-position of the striation markings seen in the image are interpreted as
-the signature.
+``scratches'' on the bullet surface (see top of \autoref{fig:rgl}).
+Typically, only striation marks in the land engraved areas (LEAs) are
+considered \citep{afte-article1992}. Bullet lands are depressed areas
+between the grooves made by the rifling action of the barrel. Compared
+to toolmarks made by screwdrivers striation marks on bullets are
+typically much smaller, both in length and in width. Bullets also have a
+curved cross-sectional topography.
 
-Bullet matching methods are usually based on these associated
-signatures. \citet{chu2013} use an automatic method for counting
+In same-source comparisons this curvature is usually removed using some
+form of Gaussian filter \citep{ma2004} or non-parametric smoothing
+\citep{aoas}. \citet{chu2013} use an automatic method for counting
 consecutive matching striae (CMS). The authors report an error rate of
-52\% of the known same source lands comparisons as misidentified (false
-negative) and zero false positives for known different source lands.
-\citet{ma2004} and \citet{vorburger2011} discuss CCF (cross-correlation
-function) and its discriminating power and applicability for same-source
-analyses of bullets, but do not provide any error rates in their
-discussion. \citet{aoas} use multiple features like CCF, CMS, D
-(distance measure) etc in a random forest based method and compare every
-land against every other land of digitised versions of Hamby 252 and
-Hamby 44 \citep{hamby} published on the NIST Ballistics Database
-\citep{nist}. The authors report an out-of-bag overall error rate of
-0.46\%, comprised of an error rate of 30.05\% of same-source pairs that
-were not identified and an error rate of 0.026\% of different-source
-pairs that were incorrectly identified as same-source.
+52\% for known same-source land comparisons to be (incorrectly)
+identified as different-source (false negative) and zero false positives
+for known different-source lands. \citet{ma2004} and
+\citet{vorburger2011} discuss CCF (cross-correlation function) and its
+discriminating power and applicability for same-source analyses of
+bullets, but do not provide any error rates in their discussion.
+\citet{aoas} use multiple features, such as CCF, CMS, D (distance
+measure), etc. in a random forest based method and compare every land
+against every other land of digitised versions of Hamby 252 and Hamby 44
+\citep{hamby} published on the NIST Ballistics Database \citep{nist}.
+The authors report an out-of-bag overall error rate of 0.46\%, comprised
+of an error rate of 30.05\% of same-source pairs that were not
+identified and an error rate of 0.026\% of different-source pairs that
+were incorrectly identified as same-source.
 
@@
 error rate of 0.089 and an average false positive error rate of 0.023.
-For other angles of 60 and 85 degrees the false negatives error rate is
-0.09 while the rate of false positives decreases to 0.01. The paper by
-\citet{hadler} is based on the same data but the authors focus on
-markings made under the same angle. The error rates associated with the
-deterministic version of the score are 0.06 for false negatives and a
-false positive error rate of 0.
+For marks made under angles of 60 and 85 degrees, respectively, the
+false negatives error rate is 0.09 while the rate of false positives
+decreases to 0.01. The paper by \citet{hadler} is based on the same data
+but the authors focus on markings made under the same angle. The error
+rates associated with the deterministic version of the score are
+reported as 0.06 for false negatives and 0 for false positives.
 
@@
 structure from the data in the attempt to increase the signal to noise
-ratio. The span of the loess fit was found using cross-validation, as
-described by \citet{aoas}.
+ratio. \autoref{fig:rgl} shows how the signature from a bullet land
+(bottom) lines up with the image of the land (top) from which it was
+extracted. We can see in the figure how the depth and relative position
+of the striation markings seen in the image are interpreted as peaks and
+valleys in the signature.
 
@@
 to us: scans of Hamby 44 and Hamby 252 are available from the NIST
-database \citep{nist}. Hamby 44 has also been made available to us and
-has been scanned locally for CSAFE at the Roy J.~Carver High Resolution
-Microscopy Facility using a Sensofar confocal light microscope. Scans in
-the NIST database are made with a NanoFocus at 20x magnification. The
-resolutions of the two instruments are different: the NIST scans are
-taken at a resolution of 1.5625 \(\mu m\) per pixel, while the CSAFE
-scans are available at a resolution of 0.645 \(\mu m\) per pixel. The
-length of an average bullet land from Hamby (9 mm Ruger P85) is about 2
-millimeter, resulting in signatures of about 1200 pixels for NIST scans,
-and about 3000 pixels for CSAFE scans.
+database \citep{nist}. The physical Hamby 44 set has also been made
+available to us and has been scanned locally for CSAFE at the Roy
+J.~Carver High Resolution Microscopy Facility using a Sensofar confocal
+light microscope. Scans in the NIST database are made with a NanoFocus
+at 20x magnification. The resolutions of the two instruments are
+different: the NIST scans are taken at a resolution of 1.5625 \(\mu m\)
+per pixel, while the CSAFE scans are available at a resolution of 0.645
+\(\mu m\) per pixel. The length of an average bullet land from Hamby (9
+mm Ruger P85) is about 2 millimeter, resulting in signatures of about
+1200 pixels for NIST scans, and about 3000 pixels for CSAFE scans.
 
@@
 
+A digitized toolmark forms a spatial process \(z(t)\) with location
+indexed by \(t\). \(t\) here denotes equally spaced pixel locations for
+the striation marks under consideration. For a toolmark consisting of
+\(t\) pixels, \(t = 1, ..., T\). Let further \(z^s(t)\) denote a vector
+of markings of length \(s\) starting in location \(t\).
+
 The Chumbley score algorithm takes input in form of two digitized
-toolmarks. The toolmark is in form of \(z(t)\) which is a spatial
-process for location indexed by \(t\). \(t\) here denotes equally spaced
-pixel locations for the striation marks under consideration,
-\(t = 1, ..., T\). Let further \(z^s(t)\) denote a vector of markings of
-length \(s\) starting in location \(t\).
+toolmarks:
 
 Let \(x(t_1)\), \(t_1 = 1,2,...T_1\) and \(y(t_2)\), \(t_2 = 1,2...T_2\)
-be two digitized toolmarks (where \(T_1\) and \(T_2\) are not
-necessarily equal). The toolmarks under consideration are potentially
-from two different sources or the same source. \(T_1\) and \(T_2\), as
-represented above, are the final pixel indexes of each marking and
-therefore give the respective lengths of the markings.
+be two digitized toolmarks (where \(T_1\) and \(T_2\), the lengths of
+the two marks, are not necessarily equal). The toolmarks under
+consideration are potentially from two different-sources or the
+same-source.
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figures/sigs-profiles-1}
+
+}
+
+\caption{Bullet land profile (left) and the corresponding signature (right) for one of the lands of Hamby-44.}\label{fig:sigs-profiles}
+\end{figure}
 
@@
 curvature in profiles from signatures before matching signatures.
+\autoref{fig:sigs-profiles} shows an example of a bullet land profile
+(left) and the corresponding signature (right).
 
-After removing sub-class structure, the Chumbley scores is calculated in
+After removing sub-class structure, the Chumbley score is calculated in
 two steps: an optimization step and a validation step. In the
@@
 \(x(t_1)\) and \(y(t_2)\) is maximized: \[
-\left(t_1^o, t_2^o\right) = \mathop{\arg \max}\limits_{1 \le t_1 \le T_1, 1 \le t_2 \le T_2} \text{cor} \left(x^{w_o} (t_1), y^{w_o}(t_2) \right)
+\left(t_1^o, t_2^o\right) = \mathop{\arg \max}\limits_{1 \le t_1 \le T_1-w_o, 1 \le t_2 \le T_2-w_o} \text{cor} \left(x^{w_o} (t_1), y^{w_o}(t_2) \right)
 \] This results in an optimal vertical (in-phase) shift of
-\(t_1^o - t_2^o\) for aligning the two markings.
-{\textcolor{orange}{We will denote the relative optimal locations as $t_1^*$ and $t_2^*$, where $t_i^* = t_i^o/(T_i-w_o)$ for $i=1,2$, such that $t_1^*, t_2^* \in [0,1]$.
-Once (sub-)class characteristics are removed, the relative optimal locations should be distributed according to a uniform distribution in $[0,1]$.
-}}
+\(t_1^o - t_2^o\) for aligning the two markings. We will denote the
+relative optimal locations as \(t_1^*\) and \(t_2^*\), where
+\(t_i^* = t_i^o/(T_i-w_o)\) for \(i=1,2\), such that
+\(t_1^*, t_2^* \in [0,1]\). Once (sub-)class characteristics are
+removed, the relative optimal locations should be distributed according
+to a uniform distribution in \([0,1]\).
 
 In the validation step, two sets of windows of size \(w_v\) are chosen
-from both markings (see Figure \ref{fig:win-comparison}). In the first
+from both markings, \autoref[see][]{fig:win-comparison}. In the first
 set, pairs of windows are extracted from the two markings using the
@@
 Same-shift pairs of length \(w_v\) are defined in \citet{hadler} as all
-pairs \((s_i^{(1)}, s_i^{(2)})\) for all integer values \(i\) for which
+pairs \((s_i^{(1)}, s_i^{(2)})\) with integer values \(i\) for which
 both \(s_i^{(1)}\) and \(s_i^{(2)}\) are defined. Similarly,
@@
 
-\caption{Two markings made by the same source.
+\caption{Two same-source markings.
 For convenience, the markings are moved into phase on the left and out-of phase on the right.
@@
 markings are calculated. The intuition here is that for two markings
-from the same source the correlation for the in-phase sample should be
+from the same-source the correlation for the in-phase sample should be
 high, while the correlations of the out-of-phase sample provide a
@@
 Looking closer at \autoref{eq.start}, we see that by definition, some
-number of tests will fail to produce a result.
-{\textcolor{orange}{The problem of failed tests is first mentioned in \citet{afte-chumbley}. Unfortunately, the authors do not provide any percentage of how many tests failed for their data.}}
-The algorithm fails to produce for two reasons: either the number of
-eligible same-shift pairs is zero, or the number of different-shift
+number of tests will fail to produce a result. The problem of failed
+tests is first mentioned in \citet{afte-chumbley}. Unfortunately, the
+authors do not provide any percentage of how many tests failed for their
+data. The algorithm fails to produce for two reasons: either the number
+of eligible same-shift pairs is zero, or the number of different-shift
 pairs is zero.
@@
 case a large difference between locations is unlikely. However, for
-different source matches, we can assume that locations are independent.
+different-source matches, we can assume that locations are independent.
 In that case, we expect a test to fail with a probability of
@@
 of 1200 pixels, \(w_o = 120\) pixels and \(w_v = 30\) pixels. Assuming
-independence in the optimal locations for different source profiles the
+independence in the optimal locations for different-source profiles the
 expected probability for a failed test is, again,
@@
 different-shift pairings is the same or at most one less than the number
-of same-shift pairings in all tests.
-{\textcolor{orange}{In the remainder of the paper, we will refer to the algorithm defined by \citet{hadler} as{\bf (CS1)} and the suggested modified algorithm as{\bf (CS2)} and compare their performance on the available scans of the Hamby study.}}
+of same-shift pairings in all tests. In the remainder of the paper, we
+will refer to the algorithm defined by \citet{hadler} as \textbf{(CS1)}
+and the suggested modified algorithm as \textbf{(CS2)} and compare their
+performance on the available scans of the Hamby study.
 
@@
 
-\begin{figure}[hbtp]
-\centering
-\includegraphics[width=0.62\textwidth]{images/same-shift-failure.png}
-
-\caption{\label{same-shift-failure}Sketch of same-shift pairings  (top) when the lag is too large to accomodate a a vaildation window in either of the two signatures}
-\end{figure}
-
-\begin{figure}[hbtp]
-\centering
-\includegraphics[width=0.7\textwidth]{images/diff-shift-failure.png}
-
-\caption{\label{diff-shift-failure}Sketch of diff-shift pairings  (top) when the number of diff-shift computations is likely to 0}
-\end{figure}
-
-\begin{figure}[hbtp]
-\centering
-\includegraphics[width=.7\textwidth]{images/CS1-failure-CS2-success.png}
-\caption{\label{CS1-failure-CS2-success}Sketch of a case where CS1 fails but CS2 does not fail}
-\end{figure}
-
 \hypertarget{testing-setup}{%
@@
 \subsection{The Data}\label{the-data}}
-
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figures/sigs-profiles-1}
-
-}
-
-\caption{Bullet land profile (left) and the corresponding signature (right) for one of the lands of Hamby-44.}\label{fig:sigs-profiles}
-\end{figure}
 
@@
 tests fail using CS1. The highest percentage of failed tests under CS2
-is 1.3\% for different source tests using a validation window size
+is 1.3\% for different-source tests using a validation window size
 \(w_v\) of 60 pixels.
@@
 
-{\textcolor{orange}{\autoref{fig:profile-sketch} gives an overview of the effect of different coarseness parameters: from left to right, coarseness levels $c$ are varied in steps of 0.05 from 0.1 to 0.3.
-The top row shows  resulting signature after smoothing the profile shown in Figure~\ref{fig:sigs-profiles} with the coarseness specified.
-The histograms in the bottom row show the relative optimal location $t^*$.
-Optimal locations are distributed uniformly once (sub-)class characteristics are removed.
-However, for coarseness values of $c > 0.20$ we see quite distinct boundary effects: optimal locations $t^*$ are found at the very extreme ends of a profile more often than one would expect based on a uniform distribution.}}
+\autoref{fig:profile-sketch} gives an overview of the effect of
+different coarseness parameters: from left to right, coarseness levels
+\(c\) are varied in steps of 0.05 from 0.1 to 0.3. The top row shows
+resulting signatures after smoothing the profile shown in
+Figure\textasciitilde{}\ref{fig:sigs-profiles} with different levels of
+coarseness. The histograms in the bottom row show the relative optimal
+location \(t^*\). Optimal locations are distributed uniformly once
+(sub-)class characteristics are removed. However, for coarseness values
+of \(c > 0.20\) we see quite distinct boundary effects: optimal
+locations \(t^*\) are found at the very extreme ends of a profile more
+often than one would expect based on a uniform distribution.
 
@@
 
+\hypertarget{appendix}{%
+\section{Appendix}\label{appendix}}
+
+\begin{appendix}
+
+\section{Scenarios of failed tests}
+
+Figures \ref{same-shift-failure} and \ref{diff-shift-failure} show scenarios in which the deterministic Chumbley score can fail. 
+In \autoref{same-shift-failure} both CS1 and CS2 fail, because the lag between optimal locations is so large that no same-shift pairs can be found once the signatures are aligned.
+These failures are inherent to the Chumbley Score and cannot be prevented.
+\begin{figure}[hbtp]
+\centering
+\includegraphics[width=0.62\textwidth]{images/same-shift-failure.png}
+
+\caption{\label{same-shift-failure}Sketch of same-shift pairings.
+When the lag between optimal reference points is too large to accommodate a validation window in noth signatures, both CS1 and CS2 fail.}
+\end{figure}
+
+\autoref{diff-shift-failure} shows two situations, in which CS1 fails to identify any different-shift pairings.
+This happens, when both relative locations are close to either end of the signature.
+CS2 can still be computed in this situation as long as there are at least two same-shift pairs identified.
+
+\begin{figure}[hbtp]
+\centering
+\includegraphics[width=0.7\textwidth]{images/diff-shift-failure.png}
+
+\caption{\label{diff-shift-failure}Sketch of different-shift pairings. For CS1 no different-shift pairs can be identified, resulting in a failed test.}
+\end{figure}
+
+<!--
+\begin{figure}[hbtp]
+\centering
+\includegraphics[width=.7\textwidth]{images/CS1-failure-CS2-success.png}
+\caption{\label{CS1-failure-CS2-success}Sketch of a case where CS1 fails but CS2 does not fail.}
+\end{figure}
+-->
+
+\end{appendix}
+
 \bibliographystyle{agsm}

### 2018-05-21 code/hamby44-evaluate.R

@@ -20,27 +20,45 @@ h44.summ <- h44 %>% group_by(wo, wv, method, coarseness) %>%
     fails = sum(is.na(p_value))/n() 
   ) %>% filter(!is.na(p_value)) %>%
   summarize(
-    auc = AUC(1-p_value, match),
-    eer = EER(1-p_value, match),
+    AUC = AUC(1-p_value, match),
+    EER = EER(1-p_value, match),
     ci = list(AUC_confint(1-p_value, match)),
     fails = fails[1],
-    roc = list(roc(match, 1-p_value))
+    delong = list(ci.auc(match, 1-p_value))
   )
 h44.summ <- h44.summ %>% mutate(
   lower = ci %>% purrr::map_dbl(.f = function(x) x[[1]][1]),
-  upper = ci %>% purrr::map_dbl(.f = function(x) x[[1]][2])
+  upper = ci %>% purrr::map_dbl(.f = function(x) x[[1]][2]),
+  low_delong = delong %>% purrr::map_dbl(.f = function(x) x[1]),
+  upp_delong = delong %>% purrr::map_dbl(.f = function(x) x[3]),
+  Method = toupper(method)
 )
 
-
-h44.summ %>% gather(type, value, auc, eer) %>% 
+h44.summ %>% gather(type, value, AUC, EER) %>% 
   filter(wv %in% c(75, 125), coarseness==0.125,
          between(wo, 210, 390)) %>% 
-  ggplot(aes(x = wo, y = value, colour=method, shape=factor(wv)))  +
+  ggplot(aes(x = wo, y = value, colour=Method, shape=factor(wv)))  +
   geom_point(size=2.5) +
 #  geom_point(aes(y = lower), size=1) +
 #  geom_point(aes(y = upper), size=1) +
   facet_wrap(~type, scales="free") +
   theme_bw() +
-  scale_colour_brewer(palette="Set1") +
-  geom_smooth(aes(linetype = factor(wv), colour=method), se=FALSE, span=2) +
-  xlab("Size of optimization window")
+  scale_colour_brewer(palette="Set1", guide = guide_legend(ncol = 1)) +
+  geom_smooth(aes(linetype = factor(wv), colour=Method), se=FALSE, span=2) +
+  xlab(expression("Size of optimization window "~w[o]~" in pixels")) +
+  ylab("Value") +
+  theme(legend.position = "bottom", 
+        legend.key.width = unit(2,"line")) +
[…]

### 2018-05-22 doc/paper.Rmd
proofing up to coarseness

@@
 
-Initially, the default settings suggested in \citet{hadler} were used: $w_o = 120$ pixels or about 190 $\mu m$ (ten percent of the average length of profiles) and coarseness $c = 0.25$. 
+As described above, the Chumbley-score is based on three parameters: coarseness $c$ and the sizes of the optimization window $w_o$ and validation window $w_v$. 
+In a first run of results, we applied default settings for the parameters, as suggested in  \citet{hadler}: $w_o = 120$ pixels or about 190 $\mu m$ (ten percent of the average length of profiles) and coarseness $c = 0.25$, and varied the size of the  validation window $w_v$ in steps of 10  from 10 pixels to 60 pixels.
+Based on a significance level $\alpha$ of 0.05 for the test, this results in a correct identification of same-source and different-source toolmarks of 93.5\% to 94.1\%, corresponding to a rate of false negatives between 0.28 and 0.36 and a rate of false positives between 0.05 and 0.06.
+However, the most prominent result we encountered, are the high number of failed tests, i.e. the number of instances, in which CS1 did not return any result.
 \autoref{fig:fails} shows the percentage of failed tests among the 85,491 land-to-land comparisons of the NIST data for different values of the validation window size $w_v$.
@@
 The highest percentage of failed tests under CS2 is 1.3% for different-source tests using a validation window size $w_v$ of 60 pixels.
-\hh{Rates of expected failures are based on simulation runs using covariances between locations of same-source profiles of 0.854, and 0.120 for locations from different-source profiles, matching observed covariances for the Hamby scans.
-}\hh{ Observed failure rates are higher than expected.
-This might be due to remaining sub-class structure at a coarseness of 0.3 resulting in a distribution of optimal locations different from the assumed uniform.
-}
+Rates of expected failures are based on simulation runs using covariances between locations of same-source profiles of 0.854, and 0.120 for locations from different-source profiles, matching observed covariances for the Hamby scans.
+Observed failure rates are higher than expected.
+This might be due to remaining sub-class structure at a coarseness of 0.25 resulting in a distribution of optimal locations different from the assumed uniform.
 
@@
 With an increase in the size of the validation window  a higher percentage of tests fails under both methods (CS1) and (CS2), but the percentage of failed tests is much smaller under (CS2).
-Observed failure rates are higher than expected rates."}
+Observed failure rates of (CS1) are higher than expected rates."}
 allinv$method <- "CS2"
@@
 )
+stats <- prof_fails %>%
+  group_by(wo, wv, method) %>% summarize(
+  failed = sum(is.na(p_value)),
+  correctplus = sum(p_value<=0.05 & match, na.rm=TRUE),
+  correctminus = sum(p_value > 0.05 & !match, na.rm=TRUE),
+  n=n(),
+  correct = (correctplus+correctminus)/(n-failed)
+)
+#stats %>% 
+#  ggplot(aes(x = wv, y = (correctplus+correctminus)/(n-failed))) +
+#  geom_point(aes(colour=method))
+#range(subset(stats, method=="CS1")$correct)
+#[1] 0.9347536 0.9410993
+
 fails <- prof_fails %>%
@@
   failed = sum(is.na(p_value)),
-  n=n()
-)
+  n=n(),
+  falseneg = sum(p_value > 0.05 & match, na.rm=TRUE),
+  falsepos = sum(p_value <= 0.05 & !match, na.rm=TRUE)
+)
+fnr <- with(subset(fails, method=="CS1"), falseneg/(n-failed))
+#summary(fnr[fnr>0])
+#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
+# 0.2819  0.2882  0.2965  0.3056  0.3075  0.3623 
+fpr <- with(subset(fails, method=="CS1"), falsepos/(n-failed))
+#summary(fpr[fpr>0])
+#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
+#0.05468 0.05588 0.05855 0.05852 0.06120 0.06227
 
@@
   geom_point(aes(size="Observed")) +
-  xlab("Size of validation window in pixels") +
+  xlab(expression("Size of validation window "~w[v]~" in pixels")) +
   ylab("Percent of failed tests") +

### 2018-05-22 doc/paper.tex
proofing up to coarseness

@@
 \includegraphics[width=.7\textwidth]{images/sketch-diff-2.png}
-\caption{\label{sketch-diff-2}Sketch of adjusted different-shift pairings. At most one of the same-shift pairings can not be matched with a different-shift pair. }
+\caption{\label{sketch-diff-2}Sketch of adjusted different-shift pairings. At most one of the same-shift pairings can not be matched with a different-shift pair under algorithm (CS2). }
 \end{figure}
@@
 
-Both algorithms (CS1) and (CS2) are implemented in the R package
-\texttt{toolmaRk} \citep{toolmark}. We applied both methods to all
-pairwise land-to-land comparisons of the Hamby scans provided by NIST
-for a total of 85,491 land-to-land comparisons.
+Both algorithms (CS1) and (CS2) are implemented in R \citep{R}. (CS1) is
+available from package \texttt{toolmaRk} \citep{toolmark}, (CS2) is
+available from a modified version of the \texttt{toolmaRk} package
+available from GitHub (\url{https://github.com/heike/toolmaRk}). We
+applied both methods to all pairwise land-to-land comparisons of the
+Hamby scans provided by NIST for a total of 85,491 land-to-land
+comparisons.
 
@@
 
-Initially, the default settings suggested in \citet{hadler} were used:
+As described above, the Chumbley-score is based on three parameters:
+coarseness \(c\) and the sizes of the optimization window \(w_o\) and
+validation window \(w_v\). In a first run of results, we applied default
+settings for the parameters, as suggested in \citet{hadler}:
 \(w_o = 120\) pixels or about 190 \(\mu m\) (ten percent of the average
-length of profiles) and coarseness \(c = 0.25\). \autoref{fig:fails}
+length of profiles) and coarseness \(c = 0.25\), and varied the size of
+the validation window \(w_v\) in steps of 10 from 10 pixels to 60
+pixels. Based on a significance level \(\alpha\) of 0.05 for the test,
+this results in a correct identification of same-source and
+different-source toolmarks of 93.5\% to 94.1\%, corresponding to a rate
+of false negatives between 0.28 and 0.36 and a rate of false positives
+between 0.05 and 0.06. However, the most prominent result we
+encountered, are the high number of failed tests, i.e.~the number of
+instances, in which CS1 did not return any result. \autoref{fig:fails}
 shows the percentage of failed tests among the 85,491 land-to-land
@@
 is 1.3\% for different-source tests using a validation window size
-\(w_v\) of 60 pixels.
-{\textcolor{orange}{Rates of expected failures are based on simulation runs using covariances between locations of same-source profiles of 0.854, and 0.120 for locations from different-source profiles, matching observed covariances for the Hamby scans.
-}}{\textcolor{orange}{ Observed failure rates are higher than expected.
-This might be due to remaining sub-class structure at a coarseness of 0.3 resulting in a distribution of optimal locations different from the assumed uniform.
-}}
+\(w_v\) of 60 pixels. Rates of expected failures are based on simulation
+runs using covariances between locations of same-source profiles of
+0.854, and 0.120 for locations from different-source profiles, matching
+observed covariances for the Hamby scans. Observed failure rates are
+higher than expected. This might be due to remaining sub-class structure
+at a coarseness of 0.25 resulting in a distribution of optimal locations
+different from the assumed uniform.
 

### 2018-05-22 doc/paper.Rmd
some more proofing

@@
 \autoref{fig:profile-sketch} gives an overview of the effect of different coarseness parameters: from left to right, coarseness levels $c$ are varied in steps of 0.05 from 0.1 to 0.3.
-The top row shows  resulting signatures after smoothing the profile shown in Figure~\ref{fig:sigs-profiles} with different levels of coarseness.
+The top row shows  resulting signatures after smoothing the profile shown in \autoref{fig:sigs-profiles} with different levels of coarseness.
 The histograms in the bottom row show the relative optimal location $t^*$.
@@
 
-\gk{The key effect of the optimal locations and thereby the coarseness is seen in the number of failed tests.
+The key effect of the optimal locations and thereby the coarseness is seen in the number of failed tests.
 Irrespective of whether CS1 or CS2 is being used, if the relative optimal locations are at the boundaries we will see an increase in the number of failed tests.
-A balance is therefore needed in the selection of the coarseness parameter which reduces the boundary effect but also does not remove important individual characteristics}
+A balance is therefore needed in the selection of the coarseness parameter which reduces the boundary effect but does not remove important individual characteristics.
+Based on \autoref{fig:profile-sketch} a coarseness value of $c = 0.15$ seems to be best suited to strike this balance for this example.
+For the remainder of the analysis, we will use this value for $c$.
 
@@
 
-## Type II error rates
-
-\autoref{fig:type2} gives an overview of the type 2 error of methods CS1 and CS2 across a range of different optimization windows $w_o$.
+## Error rate assessment
+
+<!--\autoref{fig:type2} gives an overview of the type 2 error of methods CS1 and CS2 across a range of different optimization windows $w_o$. -->
 
@@
 errorsCV <- rbind(errorsCV1, errorsCV2)
-errorsCV %>% 
+ggerrors <- errorsCV %>%
   ggplot(aes(x = wo, y = beta, colour=factor(alpha))) +
@@
   facet_wrap(~method, labeller="label_both")
-```
-
-\autoref{fig:roc} gives an overview of  ROC (Receiver operating characteristic) curves for methods CS1 and CS2 over a range of different optimization window sizes $w_o$.
+
+```
+
+\autoref{fig:roc} gives an overview of  ROC (Receiver operating characteristic) curves for methods CS1 and CS2 over a range of different optimization window sizes $w_o$ and two sizes for the validation window $w_v$ (shape).
 The different color hues represent the two methods CS1 (red) and CS2 (blue).
@@
 Generally, an optimization window $w_o$ of 150 pixels or more leads to the best performance with respect to ROC curves.
+Results based on a  validation window of size $w_v = 30$ are generally better than results for $w_v = 50$.
+
+\autoref{fig:eer}
+\autoref{tab:aucs}
 
@@
 
-cvcs <- cvcs %>% mutate(
+cvcs <- cvcs %>% 
+  mutate(
   method.wo = paste(method, wo, sep= " / "),
@@
 
-cvcs %>% filter(!wo %in% c(130, 140) ) %>%
+cvcs %>% filter(!wo %in% c(130, 140)) %>%
   ggplot(aes(d = as.numeric(match), m = 1-p_value)) + 
@@
 
-roc.summary <- cvcs %>% group_by(wo, wv, method) %>%
+roc.summary <- cvcs %>% group_by(wo, wv, method, coarse) %>%
   summarize(
     eer = EER(1-p_value, match),
-    auc = AUC(1-p_value, match)
-  )
+    auc = list(pROC::ci.auc(match, 1-p_value))
+  )
+roc.summary <- roc.summary %>% 
+  mutate(
+    lower = auc %>% purrr::map_dbl(.f = function(x) x[1]),
+    upper = auc %>% purrr::map_dbl(.f = function(x) x[3]),
+    AUC = auc %>% purrr::map_dbl(.f = function(x) x[2])
+  )
+roc.summary <- roc.summary %>% left_join(
+  errorsCV %>% dplyr::select(wv, wo, method, alpha, actual, beta) %>% 
+    dplyr::filter(alpha == 0.05)
+)
+
 p1 <- roc.summary %>% filter(!(wo %in% c(130, 140))) %>%
-  ggplot(aes( x = wo, y=eer, shape=factor(wv))) +
+  ggplot(aes( x = wo, y=eer, colour=method, shape=method)) +
   geom_point() +
-  facet_wrap(~method, labeller="label_both") + 
+  facet_wrap(~wv, labeller="label_both") + 
   theme_bw() +
-  scale_shape_discrete("Size of \nvalidation window") +
   ylab("Equal Error rate (EER)") +
   xlab("Size of optimization window in pixels") +
-  theme(legend.position = c(0.25, 0.25))
+  theme(legend.position = "bottom") + #c(0.25, 0.25)) +
+  scale_colour_brewer(palette="Set1") +
+  scale_y_continuous(limits=c(0, 0.25))
 
 p2 <- roc.summary %>% filter(!(wo %in% c(130, 140))) %>%
-  ggplot(aes( x = wo, y=auc, shape=factor(wv))) +
-  geom_point() +
-  facet_wrap(~method, labeller="label_both") + 
+  ggplot(aes( x = wo, y=AUC, colour = method, shape=method)) +
+  geom_point(size=2) +
+  geom_errorbar(aes(ymin=lower, ymax=upper), width=5) +
+  facet_wrap(~wv, labeller="label_both") + 
   theme_bw() +
-  scale_shape_discrete("Size of \nvalidation window") +
   ylab("Area under the curve (AUC)") +
   xlab("Size of optimization window in pixels")+
-  theme(legend.position = c(0.75, 0.25))
+  theme(legend.position = "bottom") + #c(0.75, 0.25)) +
+  scale_colour_brewer(palette="Set1")
 grid.arrange(p1, p2, nrow=1)
@@
 
-<!--
-## Type II error rate and coarseness
-\hh{Figure \ref{fig:coarse} shows the type II error rates for profiles using an optimization window $w_o= 120$ and a validation window $w_v = 30$ for varying level of coarseness.
-The type II error for all nominal levels of  $\alpha$ are the lowest for a coarseness range of 0.20 to 0.30, with a minimum in type II reached at a coarseness $c = 0.25$.
-The histograms suggest to use a value of coarseness below 0.2.}
-\hh{However, for an optimal value of coarseness other aspects have to be condsidered.}
-```{r coarse, fig.cap="Type II error with respect to coarseness parameter over profiles, $w_o = 120$, $w_v = 30$.
-Optimal values for coarseness are around $c \\approx 0.30$", fig.width=5, fig.height = 3, out.width='.6\\textwidth'}
-e1<-errors2 %>% filter(wv==30)%>% ggplot(aes(x = coarse, y=beta, colour=factor(alpha))) +
-  geom_point() + geom_line() +
-  theme_bw() +
-  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + xlab("Profile Coarseness")+
-  ylab("Type II error") 
-e1
-```
-
--->
+```{r, fig.cap="Overview of all of the results."}
+rocs <- roc.summary %>% dplyr::select(-auc, -coarse, -alpha)
+rocs <- rocs %>% mutate(
+  aucplus = sprintf("%.3f (%.3f, %.3f)", AUC, lower, upper)
+)
+rocs <- rocs %>% dplyr::filter(wv==30)
+rocs <- rocs %>% dplyr::select(-AUC, -lower, -upper, -wv)
+rocs1 <- rocs %>% dplyr::filter(method=="CS1")
+rocs2 <- rocs %>% dplyr::filter(method=="CS2")
+rocs_tab <- data.frame(rocs1 %>% ungroup() %>% dplyr::select(-wv, -method, -eer), 
+                  rocs2 %>% ungroup() %>% dplyr::select(actual, beta, aucplus))
+dt <- xtable(rocs_tab, 
+             digits=c(0,0,3, 3,0,3,3,0))
+#print(dt, include.rownames = FALSE)
+```
+
+\begin{table}[ht]
+\centering
+\begin{adjustbox}{max width=\textwidth}
+\begin{tabular}{rrrlrrl}
+  \hline
+  & \multicolumn{3}{l}{\bf CS1} & \multicolumn{3}{l}{\bf CS2} \\
+$w_o$ & \bf FPR & \bf FNR & \bf AUC (95\% C.I.) & \bf FPR & \bf FNR & \bf AUC (95\% C.I.) \\ 
+  \hline
+90 & 0.068 & 0.330 & 0.850 (0.835, 0.865) & 0.075 & 0.242 & 0.877 (0.863, 0.890) \\ 
+  120 & 0.065 & 0.309 & 0.864 (0.850, 0.878) & 0.072 & 0.236 & 0.890 (0.877, 0.903) \\ 
+  130 & 0.065 & 0.317 & 0.863 (0.850, 0.877) & 0.072 & 0.224 & 0.898 (0.886, 0.910) \\ 
+  140 & 0.065 & 0.304 & 0.873 (0.859, 0.886) & 0.071 & 0.229 & 0.902 (0.891, 0.914) \\ 
+  150 & 0.067 & 0.326 & 0.863 (0.850, 0.877) & 0.072 & 0.225 & 0.904 (0.892, 0.916) \\ 
+  180 & 0.061 & 0.326 & 0.877 (0.864, 0.890) & 0.070 & 0.229 & 0.907 (0.896, 0.919) \\ 
+  210 & 0.064 & 0.342 & 0.865 (0.851, 0.878) & 0.066 & 0.231 & 0.906 (0.895, 0.918) \\ 
+   \hline
+\end{tabular}
+\end{adjustbox}
+\caption{\label{tab:aucs} Overview of results as shown in Figures \ref{fig:roc} and \ref{fig:eer}.
+FPR is the observed rate of false positives (for a fixed level $\alpha = 0.05$) or type I error rate, FNR is the false negative rate or type II eror rate.
+Area under the curve (AUC) is shown with confidence intervals \citep{delong}.
+CS2 significantly outperforms CS1 with respect to its predictive power.
+}
+\end{table}
 
@@
 \hh{
-We started out trying to assess the suitability of the Chumbley Score's suitability for matching striae on bullet lands.
+We started out trying to assess the suitability of the Chumbley Score's for matching striae on bullet lands.
 Besides using recommended defaults, we also proposed ways to optimize parameters: based on the assumption that once sub-class characteristics are removed, optimal locations are distributed uniformly across the profile, we found  for bullets the smaller coarseness value of $c = 0.15$ to be suitable.
@@
 
-<!-- The results suggest that the Nominal type I error $\alpha$ value shows dependence on the size of the window of optimization.
-For a given window of optimization the actual Type I error is comparable to the nominal level for only a select few validation window sizes and for comparable validation window sizes of 30 and 50 as done here, the actual type I error does not seem to vary as much as it varies with the optimization window sizes .
-A Test Fail, i.e.
-tests in which a particular parameter setting did not return a valid result, happens, when the shift to align two signatures is so large, that the remaining overlap is too small to accommodate  windows for validation, depends on whether known-match or known non-matches has predictive value, with test results from different-sources having a much higher chance to fail.
-On conducting an analysis of all known bullet lands using the adjusted chumbley algorithm, Type II error was identified to be least bad for window of validation 30 and window of optimization 120.
-In case of unsmoothed raw marks (profiles), Type II error increases with the amount of smoothing and least for LOWESS smoothing coarseness value about 0.25 or 0.3.
-In an effort to identify the level of adaptiveness of the algorithm, comparisons were made between signatures and profiles.
-Their comparison with respect to validation window size for a fixed optimization window size suggested that, profiles have a total error (i.e all incorrect classification of known-matches and known non-matches) greater than or equal to the total error of signatures for all sizes of validation window.
-Profiles also fail more number of times than signatures in a test fail (for different coarseness keeping windows fixed and also for different validation windows keeping coarseness fixed) which lets us conclude that the behaviour of the algorithm for the profiles instead of pre-processed signatures is not better.
-Finally it should be noted that the current version of the adjusted chumbley algorithm seems to fall short when compared to other machine-learning based methods \citet{aoas}, and some level of modification to the deterministic algorithm needs to be identified and tested that would reduce the number of incorrect classifications.
-
-```{r wo120-profile, fig.width = 6, fig.height=4, fig.cap="The figure shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120", eval = FALSE}
-wo120_profile <- errors2 %>% filter(wo==120, coarse == 0.25) %>% 
-  #gather(type, error, actual:beta) %>% 
-  gather(type, error, beta) %>% 
-  ggplot(aes(x = wv, y = error, colour = factor(alpha)))  + 
-  geom_line() + geom_point() + facet_wrap(~type, scales="free") +
-    theme_bw()+ 
-  theme(legend.position="bottom") + ggtitle("Profile" , subtitle = "Varying validation window sizes, Optimization window = 120, Coarseness = 0.25")
-
-#combined120 <- full_join(errors2 %>% filter(wo==120, coarse == 0.25), wo120 %>%
-
-wo120_1<- wo120_profile %>% gather(type, error, beta) %>%  
-  ggplot(aes(x = wv, y = error, colour = factor(alpha)))+
-  geom_line(linetype=5)+  geom_point(shape= 24) + facet_wrap(~type, scales="free") +
-  theme_bw()+ 
-  theme(legend.position="bottom") + ggtitle("Signatures" , subtitle = "Varying validation window sizes, Optimization window = 120")
-
-combined120<- 
-  ggplot(data = wo120_profile %>% gather(type, error, beta), aes(x = wv, y = error, colour = factor(alpha)))+
-  geom_line(aes(linetype="Signatures"))+  geom_point(shape= 24) + 
-  geom_line(data =  errors2 %>% filter(wo==120, coarse == 0.25) %>% 
-  gather(type, error, beta), aes(x = wv, y = error, colour = factor(alpha), linetype = "Profiles")) +       geom_point(data =  errors2 %>% filter(wo==120, coarse == 0.25) %>% 
-  gather(type, error, beta)) +
-  facet_wrap(~type, scales="free") +
-  scale_linetype_manual("Lines",values=c("Signatures"=2,"Profiles"=1))+
-  guides(fill = guide_legend(keywidth = 1, keyheight = 1),
-    linetype=guide_legend(keywidth = 3, keyheight = 1),
-    colour=guide_legend(keywidth = 3, keyheight = 1))+
-  theme_bw()+ 
-  theme(legend.position="right") + ggtitle("Type II error with Validation window" , subtitle = "Profiles at coarseness 0.25") + xlab("Type II error")
-
-combined120
-#wo120_profile
-#wo120_1
-#multiplot(wo120_1, wo120_profile1)
-```
-```{r failtest-coarseness, fig.width = 6, fig.height=4, fig.cap="The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120", eval= FALSE}
-
-miss_1
-```
-```{r failtest-wv, fig.width = 6, fig.height=4, fig.cap="The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120", eval= FALSE}
-
-miss_2
-```
-```{r failtest-wo, fig.width = 6, fig.height=4, fig.cap="The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120", eval= FALSE}
-
-miss_2
-```
--->
-
-<!-- # Algorithm Modification -->
-
-<!-- The proposed algorithm modification is at the same shift step which comes after the shift distance is identified by finding two windows in the two markings that have the highest correlation. -->
-
-<!-- The modification allows for a "wiggle" room in the second marking of the two sets of markings being compared.
-This means that each set of same shift windows that are being compared, the second marking will window that is under consideration is allowd to move a little towards the left and a little towards the right.
--->
-
-<!-- This gives us a new set of 4 windows, 2 to the left and 2 to the right of the original comparison window.
-Then the correlations for each one of these 5 windows with the window under consideration of the 1st marking is retrieved (from the validation step correlation matrix) or computed.
--->
-
-<!-- Then the window that has the maximum correlation (from the set of 5 windows in the 2nd marking) with the window of the 1st marking is chosen. -->
-
-<!-- This new maximum correlation is then used to compute the U-statistics using the same method as before. -->
-
-<!-- ## Expected advantage of modification -->
-
-<!-- We are expecting that this modification would improve the type I and II errors for the good. Less number of false negatives and false positives. -->
-<!-- The reason for this expectation is many a times the bullet markings are not made at the same distances for two or more bullets because of the way it comes out of the barrel. -->
-
-<!-- Therefore rigid same shifts might not necessarily compare the right set of windows.
-In the same shift step for one set of windows, allowing the window in the 2nd marking to wiggle left and right and finding the best match to the window in the 1st marking, lets us adjust for situations where markings are compressed or elongated.
--->
-
-<!-- ## Dependence on the delta of new windows from the original same shift window -->
-
-<!-- From initial tests the amount of movement to get new windows left and right of the original same shift window seems to directly influence the -->
-<!-- the pvalue and U statistic that we look at.  -->
-
-<!-- ### other questions -->
-<!-- How much movement is reasonable and should be used?
-Is there a way to understand if the 2nd marking in comparison to the 1st marking is compressed or elongated or neither?
-Should the amount of movement selected depend on this compression or elongation?
-or using a static delta movement value justifieable.
--->
-<!--
-# Appendix
-\begin{appendix}
-
-On the other hand Figure \ref{fig:prof_missings} (b) shows if the coarseness level set in the chumbley agorithm has any effect on the signatures, which are pre-processed and already smoothed to a certain extent.
-From Figure \ref{fig:prof_missings} (b) we can notice that for different nominal $\alpha$ levels, the type II error fluctuates slightly but does not change much, thereby helping us conclude that the coarseness levels set in the LOWESS smoothing in the chumbley alggorithm does effect the type II error much for signatures.
-
-\subsection{Comparison of profiles and signatures}
-
-Another reason for failed tests can be incorrect identification of maximum correlation windows in the optimization step as seen in figure \ref{fig:prof_missings}(d) because of the level of smoothing, as too much smoothing would subdue intricate features that might otherwise help in the correlation calculations and correct identification of maximum correlation windows irrespective of the size.
-This would again cause a simiar effect as explained for figure \ref{fig:missings} with validation windows, irrespective of size, during the shifts end up at the ends of the markings resulting in an invalid calculation and failed comparison attempt.
-
-In figure \ref{fig:prof_missings}(d) and (f), we compare profiles and signatures on the basis of number of failed tests.
-The profiles chosen for figure \ref{fig:prof_missings}(f) have a constant coarseness of 0.25 and window of optimization as 120.
-The signatures in this case are not smoothed using the chumbley algorithm step of LOWESS smoothing.
-Instead signatures are used as calculated by \citet{aoas}.
-The smoothing in these signatures were determined and fixed on the basis of their performance in the random forest based algorithm proposed by \citet{aoas}.
-The comparison of profiles and signatures with variation of validation window size therefore is made on even footing.
-The trends are similar to figure \ref{fig:missings} in the sense that for known non-matches the number of failed tests are more for both signatures and profiles and increasing linearly with the validation window size.
-The problem is however, worse for profiles which has higher number of failed tests than signatures for all validation windows.
-
-The total error for different validation window sizes for signatures and profiles can be seen in figure \ref{fig:prof_missings} (e).The optimization window size is 120 and profiles are calculated at a default 0.25 coarseness level while signatures as before are not smoothed again in the modified chumbley algorithm.
-We can see that the total error is always higher for profiles as compared to signatures for all sizes of validation window.
-
-```{r, fig.cap="Type II errors for different levels of coarseness."}
-e1<-errors2 %>% filter(wv==30)%>% ggplot(aes(x = coarse, y=beta, colour=factor(alpha))) +
-  geom_point() + geom_line() +
-  theme_bw() +
-  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + xlab("Profile Coarseness")+
-  ylab("Type II error") + ggtitle(label = "(a)")
-
-e2<- errors_sig_c %>% filter(wv==30)%>% ggplot(aes(x = coarse, y=beta, colour=factor(alpha))) +
-  geom_point() + geom_line() +
-  theme_bw() +
-  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2")+ xlab("Signature Coarseness") + ylab("Type II error") + ggtitle(label = "(b)")
-
-#multiplot(e1,e2, cols = 1)
-
-```
-
-```{r, eval=FALSE}
-#errors %>% ggplot(aes(x = wv, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + facet_grid(~wo) +geom_point(data=errors2, aes(shape="profiles"))
-
-total1<- errors_sig_c %>% ggplot(aes(x = coarse, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + geom_line()  + geom_point(data= errors2 %>% filter(wv==30), aes(shape="profiles")) + geom_line(data=errors2 %>% filter(wv==30), aes(shape="profiles")) + theme_bw() +
-  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + xlab("Coarseness") +ylab("Total error") + ggtitle(label = "(c)")
-
-total2<- wo120_profile %>% filter(wv>10) %>% ggplot(aes(x = wv, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + geom_line() +geom_point(data=errors2 %>% filter(wv>10) %>% filter(coarse==0.25), aes(shape="profiles")) + geom_line(data=errors2 %>% filter(wv>10)%>% filter(coarse==0.25), aes(shape="profiles")) + theme_bw() +
-  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + ylab("Total error") + ggtitle(label = "(e)")
-
-```
-
-```{r prof_missings, fig.width=10, fig.height=10, fig.cap="Row 3:  Total error and Number of failed tests by the window validation size, wv, and ground truth, Row 2: Total error and Number of failed tests with Coarseness for both profiles and signatures, Row 1: Type II error for different coarseness levels as used in the modified chumbley algorithm for profiles and signatures", eval = FALSE}
-
-#ns1<- ns[which(ns$wv == c(30,50), arr.ind = T),]
-miss_1<- nsp %>% filter(wv == 30) %>%
-#ns %>% filter(wv == c(30, 50)) %>%
-  ggplot(aes(x = coarse, y = missperc, shape= "profile", colour = factor(match))) + 
-  #geom_smooth(method="auto", size=.5, se=FALSE) +
-  geom_point(size = 2) + 
-  theme_bw() +
-  ylab("Percent of failed tests.") +
-  xlab("Coarseness") +
-  scale_colour_brewer("Match", palette="Set1") + geom_point(data=ns_sig_c  %>% filter(wv %in% c(30,50)) , aes(shape="signature"), size = 2) + ggtitle(label = "(d)")
-# + scale_colour_brewer("Match", palette="Set2") #+geom_smooth(method="auto", size=.5, se=FALSE) 
-  #scale_shape_discrete("Size of \nvalidation\nwindow, #wv")
-
-miss_2<- nsp %>% filter(coarse == 0.25) %>%
-#ns %>% filter(wv == c(30, 50)) %>%
-  ggplot(aes(x = wv, y = missperc, shape= "profile", colour = factor(match))) + 
-  #geom_smooth(method="auto", size=.5, se=FALSE) +
-  geom_point(size = 2) + 
-  theme_bw() +
-  ylab("Percent of failed tests.") +
-  xlab("wv") +
-  scale_colour_brewer("Match", palette="Set1") + geom_point(data=ns %>% filter(wo == 120)  %>% filter(coarse == 1) , aes(shape="signature"), size = 2) + ggtitle(label = "(f)")#+ scale_colour_brewer("Match", palette="Set2")
-
-#multiplot(plotlist = list(total1, total2, miss_1, miss_2, e1, e2), cols = 2)#, matrix(c(1,2,3,3), nrow=2, byrow=TRUE))
-multiplot(plotlist = list(e1, total1, total2,e2, miss_1, miss_2), cols = 2)
-```
-\end{appendix} 
--->
-
 # Appendix
@@
 
-<!--
-\begin{figure}[hbtp]
-\centering
-\includegraphics[width=.7\textwidth]{images/CS1-failure-CS2-success.png}
-\caption{\label{CS1-failure-CS2-success}Sketch of a case where CS1 fails but CS2 does not fail.}
-\end{figure}
--->
-
 \end{appendix}

### 2018-05-22 doc/paper.tex
some more proofing

@@
 \usepackage{array}
+\usepackage{adjustbox}
 \usepackage{multirow}
@@
 With an increase in the size of the validation window  a higher percentage of tests fails under both methods (CS1) and (CS2), but the percentage of failed tests is much smaller under (CS2).
-Observed failure rates are higher than expected rates.}\label{fig:fails}
+Observed failure rates of (CS1) are higher than expected rates.}\label{fig:fails}
 \end{figure}
@@
 resulting signatures after smoothing the profile shown in
-Figure\textasciitilde{}\ref{fig:sigs-profiles} with different levels of
-coarseness. The histograms in the bottom row show the relative optimal
-location \(t^*\). Optimal locations are distributed uniformly once
-(sub-)class characteristics are removed. However, for coarseness values
-of \(c > 0.20\) we see quite distinct boundary effects: optimal
-locations \(t^*\) are found at the very extreme ends of a profile more
-often than one would expect based on a uniform distribution.
+\autoref{fig:sigs-profiles} with different levels of coarseness. The
+histograms in the bottom row show the relative optimal location \(t^*\).
+Optimal locations are distributed uniformly once (sub-)class
+characteristics are removed. However, for coarseness values of
+\(c > 0.20\) we see quite distinct boundary effects: optimal locations
+\(t^*\) are found at the very extreme ends of a profile more often than
+one would expect based on a uniform distribution.
 
-{\textcolor{blue}{The key effect of the optimal locations and thereby the coarseness is seen in the number of failed tests.
-Irrespective of whether CS1 or CS2 is being used, if the relative optimal locations are at the boundaries we will see an increase in the number of failed tests.
-A balance is therefore needed in the selection of the coarseness parameter which reduces the boundary effect but also does not remove important individual characteristics}}
+The key effect of the optimal locations and thereby the coarseness is
+seen in the number of failed tests. Irrespective of whether CS1 or CS2
+is being used, if the relative optimal locations are at the boundaries
+we will see an increase in the number of failed tests. A balance is
+therefore needed in the selection of the coarseness parameter which
+reduces the boundary effect but does not remove important individual
+characteristics. Based on \autoref{fig:profile-sketch} a coarseness
+value of \(c = 0.15\) seems to be best suited to strike this balance for
+this example. For the remainder of the analysis, we will use this value
+for \(c\).
 
@@
 
-\hypertarget{type-ii-error-rates}{%
-\subsection{Type II error rates}\label{type-ii-error-rates}}
-
-\autoref{fig:type2} gives an overview of the type 2 error of methods CS1
-and CS2 across a range of different optimization windows \(w_o\).
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.9\textwidth]{figures/type2-1}
-
-}
-
-\caption{Type II error rates observed across a range of window sizes for optimization $w_o$.
-For a window size of $w_o = 130$ we see a minimum in type II error rate across all type I rates considered.
-Smaller validation sizes $w_v$ are typically associated with a smaller type II error.}\label{fig:type2}
-\end{figure}
+\hypertarget{error-rate-assessment}{%
+\subsection{Error rate assessment}\label{error-rate-assessment}}
 
@@
 characteristic) curves for methods CS1 and CS2 over a range of different
-optimization window sizes \(w_o\). The different color hues represent
-the two methods CS1 (red) and CS2 (blue). The ROC curves show the
-superior performance of CS2 over CS1. Generally, an optimization window
-\(w_o\) of 150 pixels or more leads to the best performance with respect
-to ROC curves.
+optimization window sizes \(w_o\) and two sizes for the validation
+window \(w_v\) (shape). The different color hues represent the two
+methods CS1 (red) and CS2 (blue). The ROC curves show the superior
+performance of CS2 over CS1. Generally, an optimization window \(w_o\)
+of 150 pixels or more leads to the best performance with respect to ROC
+curves. Results based on a validation window of size \(w_v = 30\) are
+generally better than results for \(w_v = 50\).
+
+\autoref{fig:eer}
+\autoref{tab:aucs}
 
@@
 \end{figure}
+
+\begin{table}[ht]
+\centering
+\begin{adjustbox}{max width=\textwidth}
+\begin{tabular}{rrrlrrl}
+  \hline
+  & \multicolumn{3}{l}{\bf CS1} & \multicolumn{3}{l}{\bf CS2} \\
+$w_o$ & \bf FPR & \bf FNR & \bf AUC (95\% C.I.) & \bf FPR & \bf FNR & \bf AUC (95\% C.I.) \\ 
+  \hline
+90 & 0.068 & 0.330 & 0.850 (0.835, 0.865) & 0.075 & 0.242 & 0.877 (0.863, 0.890) \\ 
+  120 & 0.065 & 0.309 & 0.864 (0.850, 0.878) & 0.072 & 0.236 & 0.890 (0.877, 0.903) \\ 
+  130 & 0.065 & 0.317 & 0.863 (0.850, 0.877) & 0.072 & 0.224 & 0.898 (0.886, 0.910) \\ 
+  140 & 0.065 & 0.304 & 0.873 (0.859, 0.886) & 0.071 & 0.229 & 0.902 (0.891, 0.914) \\ 
+  150 & 0.067 & 0.326 & 0.863 (0.850, 0.877) & 0.072 & 0.225 & 0.904 (0.892, 0.916) \\ 
+  180 & 0.061 & 0.326 & 0.877 (0.864, 0.890) & 0.070 & 0.229 & 0.907 (0.896, 0.919) \\ 
+  210 & 0.064 & 0.342 & 0.865 (0.851, 0.878) & 0.066 & 0.231 & 0.906 (0.895, 0.918) \\ 
+   \hline
+\end{tabular}
+\end{adjustbox}
+\caption{\label{tab:aucs} Overview of results as shown in Figures \ref{fig:roc} and \ref{fig:eer}.
+FPR is the observed rate of false positives (for a fixed level $\alpha = 0.05$) or type I error rate, FNR is the false negative rate or type II eror rate.
+Area under the curve (AUC) is shown with confidence intervals \citep{delong}.
+CS2 significantly outperforms CS1 with respect to its predictive power.
+}
+\end{table}
 

### 2018-05-23 doc/paper.Rmd
added high-resolution scan image

@@
 
-\autoref{fig:eer}
-\autoref{tab:aucs}
+\autoref{fig:eer} shows a comparison of the performance of the two methods CS1 and CS2 with respect to EER (eaual error rate) and AUC (area under the curve) corresponding to the ROC curves shown in \autoref{fig:roc}.
+Equal error rates are reduced using method CS2, while area under the curve significantly increases (at a significance level $\alpha$ of 0.05) compared to method CS1.
+
+The results from Figures \ref{fig:roc} and \ref{fig:eer} are summarized in numbers in \autoref{tab:aucs}.
+Equal error rates (EER), rates for false positives (FPR) and false negatives (FNR) are shown side by side with the area under the curve (AUR) for both methods for a set of different optimization windows $w_o$ and a validation window $w_v$ of 30 pixels.
+The rate of false positive same-source identifications is equal to the statistical type I error rate, which is set to $\alpha = 0.05$ for this example.
+The rate of false negatives are missed same-source markings.
+This rate is also known as the type II error rate.
+Area under the curve (AUC) is shown with confidence intervals as given by \citet{delong}. CS2 significantly outperforms CS1 with respect to its predictive power in most situations.
 
@@
   ggplot(aes( x = wo, y=eer, colour=method, shape=method)) +
-  geom_point() +
-  facet_wrap(~wv, labeller="label_both") + 
+  geom_point(size=2) +
+  facet_wrap(~wv, labeller = label_bquote(w[v] == .(wv))) + 
   theme_bw() +
@@
   theme(legend.position = "bottom") + #c(0.25, 0.25)) +
-  scale_colour_brewer(palette="Set1") +
+  scale_colour_brewer("Method", palette="Set1") +
+  scale_shape("Method") +
   scale_y_continuous(limits=c(0, 0.25))
@@
   geom_errorbar(aes(ymin=lower, ymax=upper), width=5) +
-  facet_wrap(~wv, labeller="label_both") + 
+  facet_wrap(~wv, labeller = label_bquote(w[v] == .(wv))) + 
   theme_bw() +
@@
   theme(legend.position = "bottom") + #c(0.75, 0.25)) +
-  scale_colour_brewer(palette="Set1")
+  scale_colour_brewer("Method", palette="Set1") +
+  scale_shape("Method") 
 grid.arrange(p1, p2, nrow=1)
@@
 FPR is the observed rate of false positives (for a fixed level $\alpha = 0.05$) or type I error rate, FNR is the false negative rate or type II eror rate.
-Area under the curve (AUC) is shown with confidence intervals \citep{delong}.
-CS2 significantly outperforms CS1 with respect to its predictive power.
+Area under the curve (AUC) is shown with confidence intervals.
 }
@@
 
-\autoref{fig:type1} shows observed type I error rates across a range of optimization windows $w_o$.
+\autoref{fig:type1} shows observed type I error rates (FPR) across a range of optimization windows $w_o$.
 Generally, observed type I errors are higher than expected.
@@
   theme_bw() +
-  ylab("Observed type I error rate") +
-  xlab(expression("Window size for optimization, "~w[o])) +
+  ylab(expression("Observed type I error rate "~hat(alpha))) +
+  xlab(expression("Window size for optimization "~w[o])) +
   scale_colour_brewer(
     expression("Nominal type I error "~alpha), palette="Set2") +
-  facet_wrap(~alpha, labeller="label_both", scales="free") +
+  facet_wrap(~alpha, labeller=label_bquote(alpha == .(alpha)), scales="free") +
   theme(legend.position="bottom") +
@@
          col = guide_legend(ncol=2))
+
+```
+
+## High resolution Hamby 44 scans
+
+```{r h44, fig.width=7.75, fig.height=5.25, out.width='\\textwidth', fig.cap="AUC and EER for methods CS1 and CS@ based on high resolution scans of Hamby 44. "}
+if (!file.exists("../data/h44.rds")) {
+files <- dir("../data/hamby44/", pattern="rds", recursive=TRUE)
+
+h44 <- data.frame()
+for (file in files) {
+  tmp <- readRDS(file.path("../data/hamby44", file))
+  tmp$source <- file
+  tmp <- tmp %>% mutate(
+    match = replace(match, is.na(match), FALSE)
+  )
+  h44 <- rbind(h44, tmp)
+}
+  saveRDS(h44, "../data/h44.rds")
+} else {
+  readRDS("../data/h44.rds")
+}
+
+h44 <- h44 %>% separate(source, into=c("method", "rest"), sep="/") %>%
+  dplyr::select(-rest)
+
+h44.summ <- h44 %>% group_by(wo, wv, method, coarseness) %>%
+  mutate(
+    fails = sum(is.na(p_value))/length(p_value)
+  ) %>% filter(!is.na(p_value)) %>%
+  summarize(
+    AUC = AUC(1-p_value, match),
+    EER = EER(1-p_value, match),
+  #  ci = list(AUC_confint(1-p_value, match)),
+    fails = fails[1],
+    delong = list(pROC::ci.auc(match, 1-p_value))
+  )
+
+h44.summ <- h44.summ %>% mutate(
+#  lower = ci %>% purrr::map_dbl(.f = function(x) x[[1]][1]),
+#  upper = ci %>% purrr::map_dbl(.f = function(x) x[[1]][2]),
+  low_delong = delong %>% purrr::map_dbl(.f = function(x) x[1]),
+  upp_delong = delong %>% purrr::map_dbl(.f = function(x) x[3]),
+  Method = toupper(method)
+)
+
+h44.summ %>% gather(type, value, AUC, EER) %>% 
+  filter(wv %in% c(75, 125), coarseness==0.125,
+         between(wo, 210, 390)) %>% 
+  ggplot(aes(x = wo, y = value, colour=Method, shape=factor(wv)))  +
+  geom_point(size=2.5) +
+#  geom_point(aes(y = lower), size=1) +
+#  geom_point(aes(y = upper), size=1) +
+  facet_wrap(~type, scales="free") +
+  theme_bw() +
+  scale_colour_brewer(palette="Set1", guide = guide_legend(ncol = 1)) +
+  geom_smooth(aes(linetype = factor(wv), colour=Method), se=FALSE, span=2) +
+  xlab(expression("Size of optimization window "~w[o]~" in pixels")) +
+  ylab("Value") +
+  theme(legend.position = "bottom", 
+        legend.key.width = unit(2,"line")) +
+  scale_shape(expression("Size of validation window "~w[v]~" in pixels"), guide = guide_legend(ncol=1))+
+  scale_linetype(expression("Size of validation window "~w[v]~" in pixels"), guide = guide_legend(ncol=1))
 

### 2018-05-23 doc/paper.tex
added high-resolution scan image

@@
 
-\autoref{fig:eer}
-\autoref{tab:aucs}
+\autoref{fig:eer} shows a comparison of the performance of the two
+methods CS1 and CS2 with respect to EER (eaual error rate) and AUC (area
+under the curve) corresponding to the ROC curves shown in
+\autoref{fig:roc}. Equal error rates are reduced using method CS2, while
+area under the curve significantly increases (at a significance level
+\(\alpha\) of 0.05) compared to method CS1.
+
+The results from Figures \ref{fig:roc} and \ref{fig:eer} are summarized
+in numbers in \autoref{tab:aucs}. Equal error rates (EER), rates for
+false positives (FPR) and false negatives (FNR) are shown side by side
+with the area under the curve (AUR) for both methods for a set of
+different optimization windows \(w_o\) and a validation window \(w_v\)
+of 30 pixels. The rate of false positive same-source identifications is
+equal to the statistical type I error rate, which is set to
+\(\alpha = 0.05\) for this example. The rate of false negatives are
+missed same-source markings. This rate is also known as the type II
+error rate.\\
+Area under the curve (AUC) is shown with confidence intervals as given
+by \citet{delong}. CS2 significantly outperforms CS1 with respect to its
+predictive power in most situations.
 
@@
 {\textcolor{orange}{
-We started out trying to assess the suitability of the Chumbley Score's suitability for matching striae on bullet lands.
+We started out trying to assess the suitability of the Chumbley Score's for matching striae on bullet lands.
 Besides using recommended defaults, we also proposed ways to optimize parameters: based on the assumption that once sub-class characteristics are removed, optimal locations are distributed uniformly across the profile, we found  for bullets the smaller coarseness value of $c = 0.15$ to be suitable.

### 2018-05-23 code/hamby44-evaluate.R

@@ -13,18 +13,18 @@ for (file in files) {
 }
 
 h44 <- h44 %>% separate(source, into=c("method", "rest"), sep="/") %>%
-  select(-rest)
+  dplyr::select(-rest)
 
 h44.summ <- h44 %>% group_by(wo, wv, method, coarseness) %>%
   mutate(
-    fails = sum(is.na(p_value))/n() 
+    fails = sum(is.na(p_value))/length(p_value)
   ) %>% filter(!is.na(p_value)) %>%
   summarize(
     AUC = AUC(1-p_value, match),
     EER = EER(1-p_value, match),
     ci = list(AUC_confint(1-p_value, match)),
     fails = fails[1],
-    delong = list(ci.auc(match, 1-p_value))
+    delong = list(pROC::ci.auc(match, 1-p_value))
   )
 h44.summ <- h44.summ %>% mutate(
   lower = ci %>% purrr::map_dbl(.f = function(x) x[[1]][1]),

### 2018-05-23 doc/paper.Rmd
results from hamby-44

@@
 
-\begin{table}[ht]
+\begin{table}[hbtp]
 \centering
@@
 
-```{r h44, fig.width=7.75, fig.height=5.25, out.width='\\textwidth', fig.cap="AUC and EER for methods CS1 and CS@ based on high resolution scans of Hamby 44. "}
+The high-resolution scans of Hamby set 44 are capturing images at a resolution of $0.645 \mu m$ per pixel.
+On average, land engraved areas are 3000 pixels in length.
+A coarseness of $c = 0.15$ was found to be sufficient in removing any bullet curvature.
+\autoref{fig:h44} shows the resulting EER and AUC for methods CS1 and CS2 based on two sizes of validation windows ($w_v \in 75, 125$) and optimization window sizes around 300 pixels (10 percent of the average length).
+Interestingly, the overall performance of both CS1 and CS2 is a lot lower for the high resolution version of Hamby-44 than for the lower resolution scans.
+The area under the curve overall is significantly lower for the high resolution scans than for the previous set of scans.
+Partly, this might be due to the particular choice of the parameters.
+
+```{r h44, fig.width=7.75, fig.height=5, out.width='\\textwidth', fig.cap="AUC and EER for methods CS1 and CS2 based on high resolution scans of Hamby 44. "}
 if (!file.exists("../data/h44.rds")) {
@@
 } else {
-  readRDS("../data/h44.rds")
+  h44 <- readRDS("../data/h44.rds")
 }

### 2018-05-23 doc/paper.tex
results from hamby-44

@@
 
-\autoref{fig:type1} shows observed type I error rates across a range of
-optimization windows \(w_o\). Generally, observed type I errors are
-higher than expected. Method CS1 shows in this instance slightly better
-performance than method CS2, but for both an increase in the size of the
-optimization window leads to a decrease in the observed type I error
-rates.
+\autoref{fig:type1} shows observed type I error rates (FPR) across a
+range of optimization windows \(w_o\). Generally, observed type I errors
+are higher than expected. Method CS1 shows in this instance slightly
+better performance than method CS2, but for both an increase in the size
+of the optimization window leads to a decrease in the observed type I
+error rates.
 
@@
 The horizontal line in each facet indicates the nominal type I error rate.}\label{fig:type1}
+\end{figure}
+
+\hypertarget{high-resolution-hamby-44-scans}{%
+\subsection{High resolution Hamby 44
+scans}\label{high-resolution-hamby-44-scans}}
+
+The high-resolution scans of Hamby set 44 are capturing images at a
+resolution of \(0.645 \mu m\) per pixel. On average, land engraved areas
+are 3000 pixels in length. A coarseness of \(c = 0.15\) was found to be
+sufficient in removing any bullet curvature. \autoref{fig:h44} shows the
+resulting EER and AUC for methods CS1 and CS2 based on two sizes of
+validation windows (\(w_v \in 75, 125\)) and optimization window sizes
+around 300 pixels (10 percent of the average length). Interestingly, the
+overall performance of both CS1 and CS2 is a lot lower for the high
+resolution version of Hamby-44 than for the lower resolution scans. The
+area under the curve overall is significantly lower for the high
+resolution scans than for the previous set of scans. Partly, this might
+be due to the particular choice of the parameters.
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figures/h44-1}
+
+}
+
+\caption{AUC and EER for methods CS1 and CS2 based on high resolution scans of Hamby 44. }\label{fig:h44}
 \end{figure}

### 2018-05-23 doc/paper.Rmd
conclusion + stab at abstract

@@
   name: Ganesh Krishnan
-  thanks: The authors gratefully acknowledge ...
+  thanks: The authors gratefully acknowledge funding from CSAFE through cooperative agreement \# 70NANB15H176 between NIST and Iowa State University. 
 - affiliation: Department of Statistics and CSAFE, Iowa State University 
@@
 bibliography: bibliography
-abstract: null
+abstract: Same-source analyses are a major part of a Forensic Toolmark Examiner's job.
+Technological advances, such as profilometers and confocal microscopy allow to measure 3D surfaces in a high-resolution digitized form.
+Methods for matching marks for a variety of tools have been studied in the literature.
+Analysis of these digitized markings require the use of statistical methods which can quantify the scientific mechanism of comparing markings and serve as basis for an error rate calculation.
+In this paper, we are investigating the applicability of the Chumbley scoring method by \citet{hadler} to assess striation marks on bullet lands for same-source identification.
+The Chumbley score  provides us with another approach in the same-source assessment of bullet striation marks.
 ---
@@
 
-Analysis of these digitized markings require the use of statistical methods which can quantify the scientific mechanism of comparing markings and serve as basis for an error rate calcluation.
+Analysis of these digitized markings require the use of statistical methods which can quantify the scientific mechanism of comparing markings and serve as basis for an error rate calculation.
 \citet{manytoolmarks2} define a relative distance metric and use it as similarity measure between two toolmarks.
-\citet{manytoolmarks1} extract many small segments in the markings of two toolmarks and compare similiarity using a maximum pearson correlation coefficient.
+\citet{manytoolmarks1} extract many small segments in the markings of two toolmarks and compare similarity using a maximum Pearson correlation coefficient.
 The Chumbley scoring method, first introduced by \citet{chumbley}, uses a similar but more extensive framework based on a Mann-Whitney U test of the resulting correlation coefficients.
@@
 \citep{hadler} make the score deterministic for each pair of toolmarks by choosing segments for comparison systematically.
-This approach also ensures independence between segements of striae.
-In this paper, we are investigating the applicability of the Chumbley scoring method by \citet{hadler} to assess striation marks on bullet lands for same-source identification.
+This approach also ensures independence between segments of striae.
 
@@
 
+In this paper, we are investigating the applicability of the Chumbley scoring method by \citet{hadler} to assess striation marks on bullet lands for same-source identification.
 Striation marks on bullets are made by impurities in the barrel.
@@
 \citet{aoas} use multiple features, such as CCF, CMS, D (distance measure), etc.
-in a random forest based method and compare every land against every other land of  digitised versions of Hamby 252 and Hamby 44 \citep{hamby} published on the NIST Ballistics Database \citep{nist}.
+in a random forest based method and compare every land against every other land of  digitized versions of Hamby 252 and Hamby 44 \citep{hamby} published on the NIST Ballistics Database \citep{nist}.
 The authors report an out-of-bag overall error rate of 0.46%, comprised of an error rate of 30.05% of  same-source pairs that were not identified and an error rate of 0.026% of different-source pairs that were incorrectly identified as same-source.
@@
 
+\hh{What are we investigating in this paper, and what is the structure for the paper?}
+
 ## Scans for land engraved areas
@@
 
-In a pre-processing step  the two markings are smoothed using a lowess \citep{lowess} with coarseness parameter $c$. 
+In a pre-processing step  the two markings are smoothed using a LOWESS \citep{lowess} with coarseness parameter $c$. 
  Originally, this smoothing is intended to remove drift and (sub)class characteristics from individual markings, however, in the setting of matching bullet striae, we can also make use of this mechanism to separate bullet curvature in profiles from signatures before matching signatures.
@@
 
-\autoref{fig:eer} shows a comparison of the performance of the two methods CS1 and CS2 with respect to EER (eaual error rate) and AUC (area under the curve) corresponding to the ROC curves shown in \autoref{fig:roc}.
+\autoref{fig:eer} shows a comparison of the performance of the two methods CS1 and CS2 with respect to EER (equal error rate) and AUC (area under the curve) corresponding to the ROC curves shown in \autoref{fig:roc}.
 Equal error rates are reduced using method CS2, while area under the curve significantly increases (at a significance level $\alpha$ of 0.05) compared to method CS1.
@@
   ggplot(aes(d = as.numeric(match), m = 1-p_value)) + 
-  geom_roc(labels=FALSE, n.cuts=100, pointsize=0, size=0.6, 
+  geom_roc(labels=FALSE, n.cuts=500, pointsize=0, size=0.6, 
            aes(colour = method.wo)) +
@@
 \caption{\label{tab:aucs} Overview of results as shown in Figures \ref{fig:roc} and \ref{fig:eer}.
-FPR is the observed rate of false positives (for a fixed level $\alpha = 0.05$) or type I error rate, FNR is the false negative rate or type II eror rate.
+FPR is the observed rate of false positives (for a fixed level $\alpha = 0.05$), FNR is the false negative rate.
 Area under the curve (AUC) is shown with confidence intervals.
@@
 On average, land engraved areas are 3000 pixels in length.
-A coarseness of $c = 0.15$ was found to be sufficient in removing any bullet curvature.
+A coarseness of $c = 0.125$ seemed to be sufficient in removing any bullet curvature.
+Both methods have a failed test rate of less than 0.006, indicating, again, that the larger number of pixels alleviates the problem of test failures.
 \autoref{fig:h44} shows the resulting EER and AUC for methods CS1 and CS2 based on two sizes of validation windows ($w_v \in 75, 125$) and optimization window sizes around 300 pixels (10 percent of the average length).
+Both methods show an increase in performance around $w_o = 300$ pixels.
+CS2 out-performs CS1 in all scenarios, but the difference is not significant (using DeLong's confidence intervals).
 Interestingly, the overall performance of both CS1 and CS2 is a lot lower for the high resolution version of Hamby-44 than for the lower resolution scans.
 The area under the curve overall is significantly lower for the high resolution scans than for the previous set of scans.
-Partly, this might be due to the particular choice of the parameters.
+Partly, this might be due to the particular choice of the parameters, partly the higher-resolution scans might be picking up on real differences between the lands that the lower-resolution scans fail to detect.
 
@@
 
-\hh{
-We started out trying to assess the suitability of the Chumbley Score's for matching striae on bullet lands.
-Besides using recommended defaults, we also proposed ways to optimize parameters: based on the assumption that once sub-class characteristics are removed, optimal locations are distributed uniformly across the profile, we found  for bullets the smaller coarseness value of $c = 0.15$ to be suitable.
-Optimized window sizes were found based on cross-validation to minimize type 2 error rates. 
-Method CS1 proposed by \citet{hadler} has a minimal type 2 error rate of XXX for an optimized window size of YYY - which is considerably higher than the error rates achieved on matching toolmarks. 
+In assessing the suitability of the (deterministic) Chumbley Score for matching striae on bullet lands we have gained valuable insights into the process: method CS1 as proposed by \citet{hadler} has a strong dependency on the specific choice of parameters; the defaults suggested by \citet{hadler} for screw drivers are not directly applicable  for the smaller bullet lands.
+The coarseness parameter in particular has a strong impact on the performance of the test.
+However, we were able to suggest some heuristics based on the assumption that once sub-class characteristics are removed, optimal locations are distributed uniformly across the profile.
+For bullet lands we found  a coarseness value of $c = 0.15$ to be suitable for the low-resolution scans from NIST and a value of $c=0.125$ suitable for the higher-resolution scans from CSAFE.
+Sizes for optimization windows $w_o$ were based on cross-validation to minimize overall type 2 error rates.
+Unfortunately, this kind of assessment is only feasible in the setting of a large study, such as the one we presented, and does not transfer immediately to case work, where a forensic examiner would only deal with a few identifications.
+Parameter settings for different studies should be investigated and reported to allow for a further refinement.
+
+Method CS1 proposed by \citet{hadler} has a minimal type 2 error rate of 0.272 for an optimized window size of 140 pixels -- which is considerably higher than the error rates achieved on matching toolmarks, but is similar to other single-feature methods proposed for bullet matching.
 Unfortunately, method CS1 also has a high rate of failed tests -- situations, in which the algorithm does not provide a result, due to the way different-shift pairs are constructed. 
-XXX
-The algorithm CS2 introduced here is constructed in a way that reduces the number of failures by a magnitude of about XXX.
+ Algorithm CS2 is introduced here as a remedy for failed tests by introducing an alternate version of choosing different-shift pairs. 
+The algorithm CS2 introduced here is constructed in a way that achieves on average a ten-fold reduction in the number of failures.
 While reducing the failure rate, the algorithm also shows an increase in the power of the test.
-Type II error rates reach a minimum of XXX for an optimized window size of YYY.
-While significantly reduced over CS1, CS2 still has type 2 error rates on bullet lands that are  higher than the error rates achieved on the -much larger- toolmarks.
-However, bullets usually have multiple lands - in the case of Ruger P85s as used in the Hamby study, there are six lands for each bullet.
-We might be able to get more power out of the test by adapting CS2 to deal work in a bullet-to-bullet comparison.
-}
-
-\hh{
-Fix in CS2 should also improve power for matching toolmarks. 
-}
+Type II error rates of CS2 reach a minimum of 0.217 for an optimized window size of 130 pixels.
+This increase in power of CS2 over CS1 should also apply to previous studies on toolmarks.
+It would be interesting to see these results using the adjusted algorithm.
+Unfortunately, none of the studies have made the data publicly accessible.
+
+While significantly reduced over CS1, CS2 still has type 2 error rates on bullet lands that are  higher than the error rates achieved on the --much larger-- toolmarks.
+Applying these methods to the high-resolution scans provided by CSAFE shows that better scanning methodology does not guarantee a better matching performance.
+
+However, bullets usually have multiple lands -- in the case of Ruger P85s as used in the Hamby study, there are six lands for each bullet.
+We might be able to get more power out of the test by adapting CS2 to work in a bullet-to-bullet comparison.
 
@@
 \caption{\label{same-shift-failure}Sketch of same-shift pairings.
-When the lag between optimal reference points is too large to accommodate a validation window in noth signatures, both CS1 and CS2 fail.}
+When the lag between optimal reference points is too large to accommodate a validation window in both signatures, both CS1 and CS2 fail.}
 \end{figure}

### 2018-05-23 doc/paper.tex
conclusion + stab at abstract

@@
   \author{
-        Ganesh Krishnan \thanks{The authors gratefully acknowledge \ldots{}} \\
+        Ganesh Krishnan \thanks{The authors gratefully acknowledge funding from CSAFE through
+cooperative agreement \# 70NANB15H176 between NIST and Iowa State
+University.} \\
     Department of Statistics, Iowa State University\\
@@
 \begin{abstract}
-
+Same-source analyses are a major part of a Forensic Toolmark Examiner's
+job. Technological advances, such as profilometers and confocal
+microscopy allow to measure 3D surfaces in a high-resolution digitized
+form. Methods for matching marks for a variety of tools have been
+studied in the literature. Analysis of these digitized markings require
+the use of statistical methods which can quantify the scientific
+mechanism of comparing markings and serve as basis for an error rate
+calculation. Striation marks on bullets are made by impurities in the
+barrel. The Chumbley score provides us with another approach in the
+same-source assessment of bullet striation marks.
 \end{abstract}
@@
 methods which can quantify the scientific mechanism of comparing
-markings and serve as basis for an error rate calcluation.
+markings and serve as basis for an error rate calculation.
 \citet{manytoolmarks2} define a relative distance metric and use it as
@@
 many small segments in the markings of two toolmarks and compare
-similiarity using a maximum pearson correlation coefficient. The
-Chumbley scoring method, first introduced by \citet{chumbley}, uses a
-similar but more extensive framework based on a Mann-Whitney U test of
-the resulting correlation coefficients. This approach is
-non-deterministic, because segments are chosen randomly. \citep{hadler}
-make the score deterministic for each pair of toolmarks by choosing
-segments for comparison systematically. This approach also ensures
-independence between segements of striae. In this paper, we are
-investigating the applicability of the Chumbley scoring method by
-\citet{hadler} to assess striation marks on bullet lands for same-source
-identification.
+similarity using a maximum Pearson correlation coefficient. The Chumbley
+scoring method, first introduced by \citet{chumbley}, uses a similar but
+more extensive framework based on a Mann-Whitney U test of the resulting
+correlation coefficients. This approach is non-deterministic, because
+segments are chosen randomly. \citep{hadler} make the score
+deterministic for each pair of toolmarks by choosing segments for
+comparison systematically. This approach also ensures independence
+between segments of striae. In this paper, we are investigating the
+applicability of the Chumbley scoring method by \citet{hadler} to assess
+striation marks on bullet lands for same-source identification.
 
@@
 measure), etc. in a random forest based method and compare every land
-against every other land of digitised versions of Hamby 252 and Hamby 44
+against every other land of digitized versions of Hamby 252 and Hamby 44
 \citep{hamby} published on the NIST Ballistics Database \citep{nist}.
@@
 
-In a pre-processing step the two markings are smoothed using a lowess
+In a pre-processing step the two markings are smoothed using a LOWESS
 \citep{lowess} with coarseness parameter \(c\). Originally, this
@@
 \autoref{fig:eer} shows a comparison of the performance of the two
-methods CS1 and CS2 with respect to EER (eaual error rate) and AUC (area
+methods CS1 and CS2 with respect to EER (equal error rate) and AUC (area
 under the curve) corresponding to the ROC curves shown in
@@
 resolution of \(0.645 \mu m\) per pixel. On average, land engraved areas
-are 3000 pixels in length. A coarseness of \(c = 0.15\) was found to be
-sufficient in removing any bullet curvature. \autoref{fig:h44} shows the
-resulting EER and AUC for methods CS1 and CS2 based on two sizes of
-validation windows (\(w_v \in 75, 125\)) and optimization window sizes
-around 300 pixels (10 percent of the average length). Interestingly, the
-overall performance of both CS1 and CS2 is a lot lower for the high
-resolution version of Hamby-44 than for the lower resolution scans. The
-area under the curve overall is significantly lower for the high
-resolution scans than for the previous set of scans. Partly, this might
-be due to the particular choice of the parameters.
+are 3000 pixels in length. A coarseness of \(c = 0.125\) seemed to be
+sufficient in removing any bullet curvature. Both methods have a failed
+test rate of less than 0.006, indicating, again, that the larger number
+of pixels alleviates the problem of test failures. \autoref{fig:h44}
+shows the resulting EER and AUC for methods CS1 and CS2 based on two
+sizes of validation windows (\(w_v \in 75, 125\)) and optimization
+window sizes around 300 pixels (10 percent of the average length). Both
+methods show an increase in performance around \(w_o = 300\) pixels. CS2
+out-performs CS1 in all scenarios, but the difference is not significant
+(using DeLong's confidence intervals). Interestingly, the overall
+performance of both CS1 and CS2 is a lot lower for the high resolution
+version of Hamby-44 than for the lower resolution scans. The area under
+the curve overall is significantly lower for the high resolution scans
+than for the previous set of scans. Partly, this might be due to the
+particular choice of the parameters, partly the higher-resolution scans
+might be picking up on real differences between the lands that the
+lower-resolution scans fail to detect.
 
@@
 
-{\textcolor{orange}{
-We started out trying to assess the suitability of the Chumbley Score's for matching striae on bullet lands.
-Besides using recommended defaults, we also proposed ways to optimize parameters: based on the assumption that once sub-class characteristics are removed, optimal locations are distributed uniformly across the profile, we found  for bullets the smaller coarseness value of $c = 0.15$ to be suitable.
-Optimized window sizes were found based on cross-validation to minimize type 2 error rates. 
-Method CS1 proposed by \citet{hadler} has a minimal type 2 error rate of XXX for an optimized window size of YYY - which is considerably higher than the error rates achieved on matching toolmarks. 
-Unfortunately, method CS1 also has a high rate of failed tests -- situations, in which the algorithm does not provide a result, due to the way different-shift pairs are constructed. 
-XXX
-The algorithm CS2 introduced here is constructed in a way that reduces the number of failures by a magnitude of about XXX.
-While reducing the failure rate, the algorithm also shows an increase in the power of the test.
-Type II error rates reach a minimum of XXX for an optimized window size of YYY.
-While significantly reduced over CS1, CS2 still has type 2 error rates on bullet lands that are  higher than the error rates achieved on the -much larger- toolmarks.
-However, bullets usually have multiple lands - in the case of Ruger P85s as used in the Hamby study, there are six lands for each bullet.
-We might be able to get more power out of the test by adapting CS2 to deal work in a bullet-to-bullet comparison.
-}}
+In assessing the suitability of the (deterministic) Chumbley Score for
+matching striae on bullet lands we have gained valuable insights into
+the process: method CS1 as proposed by \citet{hadler} has a strong
+dependency on the specific choice of parameters; the defaults suggested
+by \citet{hadler} for screw drivers are not directly applicable for the
+smaller bullet lands. The coarseness parameter in particular has a
+strong impact on the performance of the test. However, we were able to
+suggest some heuristics based on the assumption that once sub-class
+characteristics are removed, optimal locations are distributed uniformly
+across the profile. For bullet lands we found a coarseness value of
+\(c = 0.15\) to be suitable for the low-resolution scans from NIST and a
+value of \(c=0.125\) suitable for the higher-resolution scans from
+CSAFE. Sizes for optimization windows \(w_o\) were based on
+cross-validation to minimize overall type 2 error rates. Unfortunately,
+this kind of assessment is only feasible in the setting of a large
+study, such as the one we presented, and does not transfer immediately
+to case work, where a forensic examiner would only deal with a few
+identifications. Parameter settings for different studies should be
+investigated and reported to allow for a further refinement.
 
-{\textcolor{orange}{
-Fix in CS2 should also improve power for matching toolmarks. 
-}}
+Method CS1 proposed by \citet{hadler} has a minimal type 2 error rate of
+0.272 for an optimized window size of 140 pixels -- which is
+considerably higher than the error rates achieved on matching toolmarks.
+Unfortunately, method CS1 also has a high rate of failed tests --
+situations, in which the algorithm does not provide a result, due to the
+way different-shift pairs are constructed. Algorithm CS2 is introduced
+here as a remedy for failed tests by introducing an alternate version of
+choosing different-shift pairs. The algorithm CS2 introduced here is
+constructed in a way that achieves on average a ten-fold reduction in
+the number of failures. While reducing the failure rate, the algorithm
+also shows an increase in the power of the test. Type II error rates of
+CS2 reach a minimum of 0.217 for an optimized window size of 130 pixels.
+This increase in power of CS2 over CS1 should also apply to previous
+studies on toolmarks. It would be interesting to see these results using
+the adjusted algorithm. Unfortunately, none of the studies have made the
+data publicly accessible.
+
+While significantly reduced over CS1, CS2 still has type 2 error rates
+on bullet lands that are higher than the error rates achieved on the
+--much larger-- toolmarks. Applying these methods to the high-resolution
+scans provided by CSAFE shows that better scanning methodology does not
+guarantee a better matching performance.
+
+However, bullets usually have multiple lands -- in the case of Ruger
+P85s as used in the Hamby study, there are six lands for each bullet. We
+might be able to get more power out of the test by adapting CS2 to work
+in a bullet-to-bullet comparison.
 

### 2018-05-23 doc/paper.tex
mini change to structure

@@
 mechanism of comparing markings and serve as basis for an error rate
-calculation. Striation marks on bullets are made by impurities in the
-barrel. The Chumbley score provides us with another approach in the
-same-source assessment of bullet striation marks.
+calculation. In this paper, we are investigating the applicability of
+the Chumbley scoring method by \citet{hadler} to assess striation marks
+on bullet lands for same-source identification. The Chumbley score
+provides us with another approach in the same-source assessment of
+bullet striation marks.
 \end{abstract}
@@
 
-\hypertarget{introduction-and-background}{%
-\section{Introduction and
-Background}\label{introduction-and-background}}
-
-\hypertarget{motivation}{%
-\subsection{Motivation}\label{motivation}}
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
@@
 comparison systematically. This approach also ensures independence
-between segments of striae. In this paper, we are investigating the
-applicability of the Chumbley scoring method by \citet{hadler} to assess
-striation marks on bullet lands for same-source identification.
+between segments of striae.
 
@@
 
-Striation marks on bullets are made by impurities in the barrel. As the
-bullet travels through the barrel, these imperfections leave
-``scratches'' on the bullet surface (see top of \autoref{fig:rgl}).
-Typically, only striation marks in the land engraved areas (LEAs) are
-considered \citep{afte-article1992}. Bullet lands are depressed areas
-between the grooves made by the rifling action of the barrel. Compared
-to toolmarks made by screwdrivers striation marks on bullets are
-typically much smaller, both in length and in width. Bullets also have a
-curved cross-sectional topography.
+In this paper, we are investigating the applicability of the Chumbley
+scoring method by \citet{hadler} to assess striation marks on bullet
+lands for same-source identification. Striation marks on bullets are
+made by impurities in the barrel. As the bullet travels through the
+barrel, these imperfections leave ``scratches'' on the bullet surface
+(see top of \autoref{fig:rgl}). Typically, only striation marks in the
+land engraved areas (LEAs) are considered \citep{afte-article1992}.
+Bullet lands are depressed areas between the grooves made by the rifling
+action of the barrel. Compared to toolmarks made by screwdrivers
+striation marks on bullets are typically much smaller, both in length
+and in width. Bullets also have a curved cross-sectional topography.
 
@@
 \end{figure}
+
+{\textcolor{orange}{What are we investigating in this paper, and what is the structure for the paper?}}
+
+\hypertarget{background}{%
+\section{Background}\label{background}}
 
@@
 0.272 for an optimized window size of 140 pixels -- which is
-considerably higher than the error rates achieved on matching toolmarks.
-Unfortunately, method CS1 also has a high rate of failed tests --
-situations, in which the algorithm does not provide a result, due to the
-way different-shift pairs are constructed. Algorithm CS2 is introduced
-here as a remedy for failed tests by introducing an alternate version of
-choosing different-shift pairs. The algorithm CS2 introduced here is
-constructed in a way that achieves on average a ten-fold reduction in
-the number of failures. While reducing the failure rate, the algorithm
-also shows an increase in the power of the test. Type II error rates of
-CS2 reach a minimum of 0.217 for an optimized window size of 130 pixels.
-This increase in power of CS2 over CS1 should also apply to previous
-studies on toolmarks. It would be interesting to see these results using
-the adjusted algorithm. Unfortunately, none of the studies have made the
-data publicly accessible.
+considerably higher than the error rates achieved on matching toolmarks,
+but is similar to other single-feature methods proposed for bullet
+matching. Unfortunately, method CS1 also has a high rate of failed tests
+-- situations, in which the algorithm does not provide a result, due to
+the way different-shift pairs are constructed. Algorithm CS2 is
+introduced here as a remedy for failed tests by introducing an alternate
+version of choosing different-shift pairs. The algorithm CS2 introduced
+here is constructed in a way that achieves on average a ten-fold
+reduction in the number of failures. While reducing the failure rate,
+the algorithm also shows an increase in the power of the test. Type II
+error rates of CS2 reach a minimum of 0.217 for an optimized window size
+of 130 pixels. This increase in power of CS2 over CS1 should also apply
+to previous studies on toolmarks. It would be interesting to see these
+results using the adjusted algorithm. Unfortunately, none of the studies
+have made the data publicly accessible.
 

### 2018-05-24 doc/paper.Rmd
mini change to structure
added table with aoas single feature error rates, outline added

@@
 
-# Introduction and Background
-
-## Motivation
+# Introduction
+
 Same-source analyses are a major part of a Forensic Toolmark Examiner's job.
@@
 
+<!-- \begin{table}[ht] -->
+<!-- \centering -->
+<!-- \resizebox{\textwidth}{!}{\begin{tabular}{rlrrr} -->
+<!--   \hline -->
+<!-- Research paper & Method & Data Source & False Positives & False Negatives\\ -->
+<!--   \hline -->
+<!--  \textbf{\citet{aoas}} & Consecutive Matching & Bullets & & \\ & Striae (CMS) &  & 6.25\% &  33.851\% \\ \hline -->
+<!--  \textbf{\citet{aoas}} & Consecutive Non-matching & Screwdrivers & & \\  & striae (CNMS) & & 6.26\% & 35.4166\%  \\ \hline -->
+<!--   \textbf{\citet{aoas}} & Average & Bullets  &  & \\ & Distance &  & 6.25\% & 45.833\% \\ \hline  -->
+<!--   \textbf{\citet{aoas}} & Cross-correlation & Bullets & & \\  & Function (CCF) & & 6.25\% & 17.7083\%  \\ \hline -->
+<!--   \textbf{\citet{aoas}} & Sum of Peaks & Bullets & & \\  & (S) & & 6.25\% &  18.22915\% \\ &  & & -\% &  -\% \\ -->
+<!--    \hline -->
+<!-- \end{tabular}} -->
+<!-- \caption{\label{tab:bullets-ER} Error Rates in same-source bullet analysis reported in the literature.} -->
+<!-- \end{table} -->
+
+\begin{table}[ht]
+\centering
+\resizebox{\textwidth}{!}{\begin{tabular}{rlrrr}
+  \hline
+Method & Data Source & False Positives & False Negatives\\
+  \hline
+Consecutive Matching Bullets Striae (CMS) & & & \\ \textbf{\citet{aoas}} &  & 6.25\% &  33.851\% \\ \hline
+Consecutive Non-matching Striae (CNMS) & Bullets & & \\  \textbf{\citet{aoas}} & & 6.26\% & 35.4166\%  \\ \hline
+Average Distance (D) & Bullets  &  & \\ \textbf{\citet{aoas}} & & 6.25\% & 45.833\% \\ \hline 
+Cross-correlation Function (CCF) & Bullets & & \\ \textbf{\citet{aoas}}  & & 6.25\% & 17.7083\%  \\ \hline
+Sum of Peaks (S) & Bullets  & & \\  \textbf{\citet{aoas}} & & 6.25\% &  18.22915\% \\
+   \hline
+\end{tabular}}
+\caption{\label{tab:bullets-ER} Error Rates in same-source bullet analysis reported in the literature.}
+\end{table}
+
 The Chumbley score  provides us with another approach in the same-source assessment of bullet striation marks.
@@
 \hh{What are we investigating in this paper, and what is the structure for the paper?}
+\gk{In this paper we evaluate the adaptability of the chumbley score method to bullets, and in the process identify parameters that minimize the error rates.
+We then identify persistent scenarios in which the method proposed by \citet{hadler} fails to conduct a test.
+We go on to provide a solution to the failed tests problem, consequently increasing the power of the test and reducing error rates in the process.
+We also investigate the performance of the two algorithms using higher as well as lower resolution scans to understand if better scans and better equipment have a positive impact on the results.}
+
+# Background
 

### 2018-05-24 doc/paper.Rmd
small edits

@@
   \hline
-Consecutive Matching Bullets Striae (CMS) & & & \\ \textbf{\citet{aoas}} &  & 6.25\% &  33.851\% \\ \hline
+Consecutive Matching Bullets Striae (CMS) & Bullets & & \\ \textbf{\citet{aoas}} &  & 6.25\% &  33.851\% \\ \hline
 Consecutive Non-matching Striae (CNMS) & Bullets & & \\  \textbf{\citet{aoas}} & & 6.26\% & 35.4166\%  \\ \hline

### 2018-05-24 doc/paper.Rmd
abstract

--- 2018-05-24 doc/paper.Rmd
@@
 - Mann-Whitney U statistic
-- 
-- 
+- land engraved areas (LEAs)
+- algorithm
 bibliography: bibliography
-abstract: Same-source analyses are a major part of a Forensic Toolmark Examiner's job.
-Technological advances, such as profilometers and confocal microscopy allow to measure 3D surfaces in a high-resolution digitized form.
-Methods for matching marks for a variety of tools have been studied in the literature.
-Analysis of these digitized markings require the use of statistical methods which can quantify the scientific mechanism of comparing markings and serve as basis for an error rate calculation.
-In this paper, we are investigating the applicability of the Chumbley scoring method by \citet{hadler} to assess striation marks on bullet lands for same-source identification.
-The Chumbley score  provides us with another approach in the same-source assessment of bullet striation marks.
+abstract: One of the major problems in forensic toolmark examination is the same-source problem, i.e.
+the identification of whether the marks on two different objects were made by the same tool.
+Profilometers and confocal microscopy and other technological advances allow us to measure 3D surfaces in previously unforeseen resolution and also enable digitized images.
+The literature provides various methods to measure similarity of toolmarks.
+In this paper, we are investigating the applicability of the Chumbley scoring method by \citet{hadler}, which was introduced for the assessment of marks made by screwdrivers, to assess striation marks on bullet lands for same-source identification.
+We provide methods to identify parameters that minimize the error rates for matching of LEAs at the example of the Hamby data, sets 44 and 252 measured by NIST and CSAFE.
+We provide a remedial algorithm to alleviate the problem of failed tests in the method proposed by \citet{hadler}.
+This increases the power of the test and reduces error rates in the process.
+Type II error rates for the Hamby data using the adjusted method showed an improvement of more than 20\% points over the performance of the algorithm proposed by \citet{hadler} (for a Type I error of 0.05).
+Overall, type II error rates of the adjusted method are on average at 0.22, putting the method on similar footing as other single feature matching approaches in the literature.
 ---
@@
 In same-source comparisons this curvature is usually removed using some form of Gaussian filter \citep{ma2004} or non-parametric smoothing \citep{aoas}.
+An overview of some of the error rates reported in the literature on bullet matching is given in \autoref{tab:bullets-ER}.
 \citet{chu2013} use an automatic method for counting consecutive matching striae (CMS).
@@
   \hline
-Consecutive Matching Bullets Striae (CMS) & & & \\ \textbf{\citet{aoas}} &  & 6.25\% &  33.851\% \\ \hline
-Consecutive Non-matching Striae (CNMS) & Bullets & & \\  \textbf{\citet{aoas}} & & 6.26\% & 35.4166\%  \\ \hline
-Average Distance (D) & Bullets  &  & \\ \textbf{\citet{aoas}} & & 6.25\% & 45.833\% \\ \hline 
-Cross-correlation Function (CCF) & Bullets & & \\ \textbf{\citet{aoas}}  & & 6.25\% & 17.7083\%  \\ \hline
-Sum of Peaks (S) & Bullets  & & \\  \textbf{\citet{aoas}} & & 6.25\% &  18.22915\% \\
+  \textbf{\citet{aoas}} & LEAs \\
+Consecutive Matching Bullets Striae (CMS) &  &  6.25\% &  33.85\% \\ 
+Consecutive Non-matching Striae (CNMS)  & &  6.26\% & 35.42\%  \\ 
+Average Distance (D) &   &   6.25\% & 45.83\% \\ 
+Cross-correlation Function (CCF) & &  6.25\% & 17.71\%  \\ 
+Sum of Peaks (S) &  &  6.25\% &  18.23\% \\
    \hline
+   \textbf{\citet{chu2013}} & LEAs \\
+Consecutive Matching Bullets Striae (CMS) &  &  0\% &  52\% \\ 
 \end{tabular}}
-\caption{\label{tab:bullets-ER} Error Rates in same-source bullet analysis reported in the literature.}
+\caption{\label{tab:bullets-ER} Error Rates in same-source land-to-land analysis reported in the literature.}
 \end{table}
@@
 
-\hh{What are we investigating in this paper, and what is the structure for the paper?}
-\gk{In this paper we evaluate the adaptability of the chumbley score method to bullets, and in the process identify parameters that minimize the error rates.
-We then identify persistent scenarios in which the method proposed by \citet{hadler} fails to conduct a test.
+In this paper we evaluate the adaptability of the Chumbley score as a measure to quantify similarity in land engraved areas (LEAs) on bullets.
+For that we briefly introduce the deterministic method suggested by \citet{hadler} in section 2.
+In the process we provide methods to identify parameters that minimize the error rates.
+We then investigate persistent scenarios in which the method proposed by \citet{hadler} fails to come to a result.
 We go on to provide a solution to the failed tests problem, consequently increasing the power of the test and reducing error rates in the process.
-We also investigate the performance of the two algorithms using higher as well as lower resolution scans to understand if better scans and better equipment have a positive impact on the results.}
+We set up a testing framework to compare the performance of the two algorithms in section 3 and finally discuss results in section 4.
 

### 2018-05-24 doc/paper.tex
abstract

@@
 \begin{abstract}
-Same-source analyses are a major part of a Forensic Toolmark Examiner's
-job. Technological advances, such as profilometers and confocal
-microscopy allow to measure 3D surfaces in a high-resolution digitized
-form. Methods for matching marks for a variety of tools have been
-studied in the literature. Analysis of these digitized markings require
-the use of statistical methods which can quantify the scientific
-mechanism of comparing markings and serve as basis for an error rate
-calculation. In this paper, we are investigating the applicability of
-the Chumbley scoring method by \citet{hadler} to assess striation marks
-on bullet lands for same-source identification. The Chumbley score
-provides us with another approach in the same-source assessment of
-bullet striation marks.
+One of the major problems in forensic toolmark examination is the
+same-source problem, i.e.~the identification of whether the marks on two
+different objects were made by the same tool. Profilometers and confocal
+microscopy and other technological advances allow us to measure 3D
+surfaces in previously unforeseen resolution and also enable digitized
+images. The literature provides various methods to measure similarity of
+toolmarks. In this paper, we are investigating the applicability of the
+Chumbley scoring method by \citet{hadler}, which was introduced for the
+assessment of marks made by screwdrivers, to assess striation marks on
+bullet lands for same-source identification. We provide methods to
+identify parameters that minimize the error rates for matching of LEAs
+at the example of the Hamby data, sets 44 and 252 measured by NIST and
+CSAFE. We provide a remedial algorithm to alleviate the problem of
+failed tests in the method proposed by \citet{hadler}. This increases
+the power of the test and reduces error rates in the process. Type II
+error rates for the Hamby data using the adjusted method showed an
+improvement of more than 20\% points over the performance of the
+algorithm proposed by \citet{hadler} (for a Type I error of 0.05).
+Overall, type II error rates of the adjusted method are on average at
+0.22, putting the method on similar footing as other single feature
+matching approaches in the literature.
 \end{abstract}
@@
 \noindent%
-{\it Keywords:} forensic science, toolmark, cross-correlation, Mann-Whitney U statistic, , 
+{\it Keywords:} forensic science, toolmark, cross-correlation, Mann-Whitney U statistic, land engraved areas (LEAs), algorithm
 \vfill
@@
 form of Gaussian filter \citep{ma2004} or non-parametric smoothing
-\citep{aoas}. \citet{chu2013} use an automatic method for counting
-consecutive matching striae (CMS). The authors report an error rate of
-52\% for known same-source land comparisons to be (incorrectly)
-identified as different-source (false negative) and zero false positives
-for known different-source lands. \citet{ma2004} and
-\citet{vorburger2011} discuss CCF (cross-correlation function) and its
-discriminating power and applicability for same-source analyses of
-bullets, but do not provide any error rates in their discussion.
-\citet{aoas} use multiple features, such as CCF, CMS, D (distance
-measure), etc. in a random forest based method and compare every land
-against every other land of digitized versions of Hamby 252 and Hamby 44
-\citep{hamby} published on the NIST Ballistics Database \citep{nist}.
-The authors report an out-of-bag overall error rate of 0.46\%, comprised
-of an error rate of 30.05\% of same-source pairs that were not
-identified and an error rate of 0.026\% of different-source pairs that
-were incorrectly identified as same-source.
+\citep{aoas}. An overview of some of the error rates reported in the
+literature on bullet matching is given in \autoref{tab:bullets-ER}.
+\citet{chu2013} use an automatic method for counting consecutive
+matching striae (CMS). The authors report an error rate of 52\% for
+known same-source land comparisons to be (incorrectly) identified as
+different-source (false negative) and zero false positives for known
+different-source lands. \citet{ma2004} and \citet{vorburger2011} discuss
+CCF (cross-correlation function) and its discriminating power and
+applicability for same-source analyses of bullets, but do not provide
+any error rates in their discussion. \citet{aoas} use multiple features,
+such as CCF, CMS, D (distance measure), etc. in a random forest based
+method and compare every land against every other land of digitized
+versions of Hamby 252 and Hamby 44 \citep{hamby} published on the NIST
+Ballistics Database \citep{nist}. The authors report an out-of-bag
+overall error rate of 0.46\%, comprised of an error rate of 30.05\% of
+same-source pairs that were not identified and an error rate of 0.026\%
+of different-source pairs that were incorrectly identified as
+same-source.
+
+\begin{table}[ht]
+\centering
+\resizebox{\textwidth}{!}{\begin{tabular}{rlrrr}
+  \hline
+Method & Data Source & False Positives & False Negatives\\
+  \hline
+  \textbf{\citet{aoas}} & LEAs \\
+Consecutive Matching Bullets Striae (CMS) &  &  6.25\% &  33.85\% \\ 
+Consecutive Non-matching Striae (CNMS)  & &  6.26\% & 35.42\%  \\ 
+Average Distance (D) &   &   6.25\% & 45.83\% \\ 
+Cross-correlation Function (CCF) & &  6.25\% & 17.71\%  \\ 
+Sum of Peaks (S) &  &  6.25\% &  18.23\% \\
+   \hline
+   \textbf{\citet{chu2013}} & LEAs \\
+Consecutive Matching Bullets Striae (CMS) &  &  0\% &  52\% \\ 
+\end{tabular}}
+\caption{\label{tab:bullets-ER} Error Rates in same-source land-to-land analysis reported in the literature.}
+\end{table}
 
@@
 
-{\textcolor{orange}{What are we investigating in this paper, and what is the structure for the paper?}}
+In this paper we evaluate the adaptability of the Chumbley score as a
+measure to quantify similarity in land engraved areas (LEAs) on bullets.
+For that we briefly introduce the deterministic method suggested by
+\citet{hadler} in section 2. In the process we provide methods to
+identify parameters that minimize the error rates. We then investigate
+persistent scenarios in which the method proposed by \citet{hadler}
+fails to come to a result. We go on to provide a solution to the failed
+tests problem, consequently increasing the power of the test and
+reducing error rates in the process. We set up a testing framework to
+compare the performance of the two algorithms in section 3 and finally
+discuss results in section 4.
 

### 2018-05-24 doc/paper.Rmd
type 2 errors back in the appendix

@@
 
-```{r type2, fig.height = 3, fig.width = 7.5, out.width='0.9\\textwidth', fig.cap="Type II error rates observed across a range of window sizes for optimization $w_o$.
+```{r , fig.height = 3, fig.width = 7.5, out.width='0.9\\textwidth', fig.cap="Type II error rates observed across a range of window sizes for optimization $w_o$.
 For a window size of $w_o = 130$ we see a minimum in type II error rate across all type I rates considered.
@@
 
+## Type 2 errors: CS1 vs CS2
+
+\autoref{type2} gives an overview of type 2 error rates of methods CS1 and CS2 for different significance levels $\alpha$. Method CS2 is outperforming CS1 significantly in every instance.
+
+```{r type2, fig.height = 3, fig.width = 7.5, out.width='0.9\\textwidth', fig.cap="Type II error rates observed across a range of window sizes for optimization $w_o$.
+For a window size of $w_o = 130$ we see a minimum in type II error rate across all type I rates considered.
+Smaller validation sizes $w_v$ are typically associated with a smaller type II error."}
+ggerrors
+```
+
 \end{appendix}

### 2018-05-26 doc/paper.tex
no hashes in latex section

@@
 
+\section{Type 2 errors: CS1 vs CS2}
+
+\autoref{type2} gives an overview of type 2 error rates of methods CS1 and CS2 for different significance levels $\alpha$. Method CS2 is outperforming CS1 significantly in every instance.
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.9\textwidth]{figures/type2-1}
+
+}
+
+\caption{Type II error rates observed across a range of window sizes for optimization $w_o$.
+For a window size of $w_o = 130$ we see a minimum in type II error rate across all type I rates considered.
+Smaller validation sizes $w_v$ are typically associated with a smaller type II error.}\label{fig:type2}
+\end{figure}
+
 \end{appendix}

### 2018-05-28 doc/paper.Rmd
no hashes in latex section
adding ccf entry to table

--- 2018-05-24 doc/paper.Rmd
@@
 Consecutive Matching Bullets Striae (CMS) &  &  0\% &  52\% \\ 
+   \hline
+   \textbf{\citet{chu2010}} \& \textbf{\citet{ma2004}} & LEAs \\
+Cross-Correlation Function (CCF) &  &  - &  - \\ 
+    \hline
 \end{tabular}}
@@
 
-```{r type2, fig.height = 3, fig.width = 7.5, out.width='0.9\\textwidth', fig.cap="Type II error rates observed across a range of window sizes for optimization $w_o$.
-For a window size of $w_o = 130$ we see a minimum in type II error rate across all type I rates considered.
-Smaller validation sizes $w_v$ are typically associated with a smaller type II error."}
+```{r t, echo = FALSE}
 if (!file.exists("../data/cvcs2.RDS")) {
@@
 
+\newpage
+\section{Type 2 errors: CS1 vs CS2}
+
+\autoref{fig:type2} gives an overview of type 2 error rates of methods CS1 and CS2 for different significance levels $\alpha$. Method CS2 is outperforming CS1 significantly in every instance.
+
+```{r type2, fig.pos = 'h', fig.height = 3, fig.width = 7.5, out.width='0.9\\textwidth', fig.cap="Type II error rates observed across a range of window sizes for optimization $w_o$.
+For a window size of $w_o = 130$ we see a minimum in type II error rate across all type I rates considered.
+Smaller validation sizes $w_v$ are typically associated with a smaller type II error."}
+ggerrors
+```
+
 \end{appendix}

### 2018-05-28 doc/paper.Rmd
150 word abstract and small changes in table

@@
 
+\gk{ABSTRACT (150 words version)}
+\newline
+\gk{Same-source problem is a major challenge in forensic toolmark/ firearm examination.
+Technological advances in surface imaging allow us to measure 3D surfaces in previously unforeseen resolution and enable digitized images.
+Here, we investigate the applicability of the Chumbley scoring method by \citet{hadler}, developed for screwdriver markings, for same-source identification of bullet striation marks.
+We provide methods to identify parameters that minimize the error rates for matching of LEAs using Hamby datasets 44 and 252 measured by NIST and CSAFE.
+We then, provide a remedial algorithm to alleviate the problem of failed tests in the method, thereby increasing the power of the test and reducing error rates.
+Type II error rates of the proposed method showed an improvement of more than 20\% points (Type I error of 0.05) and are on average about 0.22.
+This puts the proposed method on similar footing as other single feature matching approaches in the literature.}\newline
+\gk{Other potential \textbf{Keywords} : (minimum of 6)
+profiles, signatures, same-source, different-source, comparison windows}
+
 \tableofcontents
@@
   \hline
- \textbf{\citet{manytoolmarks1}} & Maximized & Screwdrivers & & \\ & Correlation &  & - & - \\ \hline
+ \textbf{\citet{manytoolmarks1}} & Maximum Pearson & Screwdrivers & & \\ & Correlation &  & - & - \\ \hline
  \textbf{\citet{chumbley}} & Randomized & Screwdrivers & & \\ (Same-Surface Same-Angle) & Chumbley Score & & 2.3\% & 8.9\%  \\ \hline
@@
 \end{tabular}}
-\caption{\label{tab:toolmarks-ER} Error Rates in same-source toolmark analysis reported in the literature. All reported papers use some variation of the Chumbley-score method.}
+\caption{\label{tab:toolmarks-ER} Error Rates in same-source toolmark analysis reported in the literature. The top four reported papers use some variation of the Chumbley-score method.}
 \end{table}
@@
 \end{tabular}}
-\caption{\label{tab:bullets-ER} Error Rates in same-source land-to-land analysis reported in the literature.}
+\caption{\label{tab:bullets-ER} Error Rates in same-source single feature land-to-land analysis reported in the literature.}
 \end{table}

### 2018-06-19 doc/paper.Rmd
renamed background to methods, pulled in testing setup
reading through abstract

--- 2018-05-28 doc/paper.Rmd
@@
 - algorithm
+- signatures 
+- same-source problem
 bibliography: bibliography
-abstract: One of the major problems in forensic toolmark examination is the same-source problem, i.e.
-the identification of whether the marks on two different objects were made by the same tool.
-Profilometers and confocal microscopy and other technological advances allow us to measure 3D surfaces in previously unforeseen resolution and also enable digitized images.
-The literature provides various methods to measure similarity of toolmarks.
-In this paper, we are investigating the applicability of the Chumbley scoring method by \citet{hadler}, which was introduced for the assessment of marks made by screwdrivers, to assess striation marks on bullet lands for same-source identification.
-We provide methods to identify parameters that minimize the error rates for matching of LEAs at the example of the Hamby data, sets 44 and 252 measured by NIST and CSAFE.
-We provide a remedial algorithm to alleviate the problem of failed tests in the method proposed by \citet{hadler}.
-This increases the power of the test and reduces error rates in the process.
-Type II error rates for the Hamby data using the adjusted method showed an improvement of more than 20\% points over the performance of the algorithm proposed by \citet{hadler} (for a Type I error of 0.05).
-Overall, type II error rates of the adjusted method are on average at 0.22, putting the method on similar footing as other single feature matching approaches in the literature.
+abstract: The same-source problem remains a major challenge in forensic toolmark and firearm examination.
+Technological advances in surface imaging allow measurements of 3D surfaces at previously unforeseen resolutions and enable digitized imaging.
+Here, we investigate the applicability of the Chumbley scoring method \citep{hadler}, developed for screwdriver markings, for same-source identification of bullet striae.
+We provide methods to identify parameters that minimize the error rates for matching of LEAs using Hamby datasets 44 and 252 measured by NIST and CSAFE.
+We suggest a remedial algorithm to alleviate the problem of failed tests in the method, increasing both the power of the test and reducing error rates.
+Type II error rates of the proposed method improve by more than one third (Type I error of 0.05) and are on average about 0.22.
+This puts the proposed method on similar footing as other single feature matching approaches in the literature.
 ---
@@
 
-# Background
+# Methods
 
@@
 
-# Testing setup
-
-## The Data
+## Testing setup
+
+### The Data
 
@@
 
-## Setup
+### Setup
 Both algorithms (CS1) and (CS2) are implemented in R \citep{R}.

### 2018-06-19 doc/paper.tex
reading through abstract

@@
 \begin{abstract}
-One of the major problems in forensic toolmark examination is the
-same-source problem, i.e.~the identification of whether the marks on two
-different objects were made by the same tool. Profilometers and confocal
-microscopy and other technological advances allow us to measure 3D
-surfaces in previously unforeseen resolution and also enable digitized
-images. The literature provides various methods to measure similarity of
-toolmarks. In this paper, we are investigating the applicability of the
-Chumbley scoring method by \citet{hadler}, which was introduced for the
-assessment of marks made by screwdrivers, to assess striation marks on
-bullet lands for same-source identification. We provide methods to
-identify parameters that minimize the error rates for matching of LEAs
-at the example of the Hamby data, sets 44 and 252 measured by NIST and
-CSAFE. We provide a remedial algorithm to alleviate the problem of
-failed tests in the method proposed by \citet{hadler}. This increases
-the power of the test and reduces error rates in the process. Type II
-error rates for the Hamby data using the adjusted method showed an
-improvement of more than 20\% points over the performance of the
-algorithm proposed by \citet{hadler} (for a Type I error of 0.05).
-Overall, type II error rates of the adjusted method are on average at
-0.22, putting the method on similar footing as other single feature
-matching approaches in the literature.
+The same-source problem remains a major challenge in forensic toolmark
+and firearm examination. Technological advances in surface imaging allow
+measurements of 3D surfaces at previously unforeseen resolutions and
+enable digitized imaging. Here, we investigate the applicability of the
+Chumbley scoring method \citep{hadler}, developed for screwdriver
+markings, for same-source identification of bullet striae. We provide
+methods to identify parameters that minimize the error rates for
+matching of LEAs using Hamby datasets 44 and 252 measured by NIST and
+CSAFE. We suggest a remedial algorithm to alleviate the problem of
+failed tests in the method, increasing both the power of the test and
+reducing error rates. Type II error rates of the proposed method improve
+by more than one third (Type I error of 0.05) and are on average about
+0.22. This puts the proposed method on similar footing as other single
+feature matching approaches in the literature.
 \end{abstract}
@@
 \noindent%
-{\it Keywords:} forensic science, toolmark, cross-correlation, Mann-Whitney U statistic, land engraved areas (LEAs), algorithm
+{\it Keywords:} forensic science, toolmark, cross-correlation, Mann-Whitney U statistic, land engraved areas (LEAs), algorithm, signatures, same-source problem
 \vfill
@@
 Consecutive Matching Bullets Striae (CMS) &  &  0\% &  52\% \\ 
+   \hline
+   \textbf{\citet{chu2010}} \& \textbf{\citet{ma2004}} & LEAs \\
+Cross-Correlation Function (CCF) &  &  - &  - \\ 
+    \hline
 \end{tabular}}
-\caption{\label{tab:bullets-ER} Error Rates in same-source land-to-land analysis reported in the literature.}
+\caption{\label{tab:bullets-ER} Error Rates in same-source single feature land-to-land analysis reported in the literature.}
 \end{table}
@@
 
-\hypertarget{background}{%
-\section{Background}\label{background}}
+\hypertarget{methods}{%
+\section{Methods}\label{methods}}
 
@@
 \hypertarget{testing-setup}{%
-\section{Testing setup}\label{testing-setup}}
+\subsection{Testing setup}\label{testing-setup}}
 
 \hypertarget{the-data}{%
-\subsection{The Data}\label{the-data}}
+\subsubsection{The Data}\label{the-data}}
 
@@
 \hypertarget{setup}{%
-\subsection{Setup}\label{setup}}
+\subsubsection{Setup}\label{setup}}
 
@@
 
+\newpage
 \section{Type 2 errors: CS1 vs CS2}
 
-\autoref{type2} gives an overview of type 2 error rates of methods CS1 and CS2 for different significance levels $\alpha$. Method CS2 is outperforming CS1 significantly in every instance.
+\autoref{fig:type2} gives an overview of type 2 error rates of methods CS1 and CS2 for different significance levels $\alpha$. Method CS2 is outperforming CS1 significantly in every instance.
 
-\begin{figure}
+\begin{figure}[h]
 

### 2018-06-29 doc/paper.tex
final changes to rmd before submission

@@
 {
-  \title{\bf Adaption of the Chumbley Score to matching of bullet striation marks}
+  \title{\bf Adapting the Chumbley Score to match striae on Land Engraved Areas of
+bullets}
 
   \author{
-        Ganesh Krishnan \thanks{The authors gratefully acknowledge funding from CSAFE through
-cooperative agreement \# 70NANB15H176 between NIST and Iowa State
-University.} \\
-    Department of Statistics, Iowa State University\\
-     and \\     Heike Hofmann \\
-    Department of Statistics and CSAFE, Iowa State University\\
+        Ganesh Krishnan,\\ %\\
+    Master of Science (M.Sc.)\\
+    % \\
+     and \\     Prof.~Heike Hofmann,\\ %\\
+    Doctor of Philosophy (Ph.D.)\\
+    % \\
       }
     \maketitle
+  \bigskip
+  \begin{center}
+  {\textbf{Affiliation(s)}\\
+  Department of Statistics, Iowa State University and CSAFE (Center for
+  Statistics and Applications in Forensic Evidence)\\
+  \bigskip
+  \textbf{Source of funding}\\
+  The authors gratefully acknowledge funding from CSAFE through
+  cooperative agreement \# 70NANB15H176 between NIST and Iowa State
+  University.\\
+  \bigskip
+  \textbf{Acknowledgement}\\
+  We would like to thank the efforts of CSAFE and Roy J Carver Microscopy
+  lab in scanning the Hamby-44 set and providing the scans to us. We would
+  also like to thank NIST for providing the Hamby-44 bullets to us for
+  scanning.\\
+  \bigskip
+  \textbf{Information on presentation of work}\\
+  This work was presented in the form of a poster in the AFTE 2018 49th
+  Annual Training Seminar, held from June 3rd to June 8th, 2018 in
+  Charleston, West Virginia}
+  \end{center}
+
 } \fi
 
-\if1\blind
+  % \bigskip
+
+  % \bigskip
+\newpage
+
+\if0\blind
 {
@@
   \begin{center}
-    {\LARGE\bf Adaption of the Chumbley Score to matching of bullet striation marks}
+  \title{\LARGE\bf Adapting the Chumbley Score to match striae on Land Engraved Areas of
+bullets}
   \end{center}
-  \medskip
+  \bigskip
+  \maketitle
 } \fi
+
+% \if1\blind
+% {
+%   \bigskip
+%   \bigskip
+%   \bigskip
+%   \begin{center}
+%     {\LARGE\bf Adapting the Chumbley Score to match striae on Land Engraved Areas of
+bullets}
+%   \end{center}
+%   \medskip
+% } \fi
 
@@
 \newcommand{\cited}[1]{{\textcolor{red}{#1}}}
+
 \setlength\parindent{0pt}
@@
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+\section{Introduction}\label{introduction}
 
@@
 
-\hypertarget{methods}{%
-\section{Methods}\label{methods}}
+\section{Methods}\label{methods}
 
-\hypertarget{scans-for-land-engraved-areas}{%
 \subsection{Scans for land engraved
-areas}\label{scans-for-land-engraved-areas}}
+areas}\label{scans-for-land-engraved-areas}
 
@@
 
-\hypertarget{the-chumbley-score-test}{%
-\subsection{The Chumbley Score Test}\label{the-chumbley-score-test}}
+\subsection{The Chumbley Score Test}\label{the-chumbley-score-test}
 
@@
 More precisely, let us define starting points \(s_i^{(k)}\) for each
-signature \(k = 1, 2\) as \begin{eqnarray}\label{eq.start}
+signature \(k = 1, 2\) as
+
+\begin{eqnarray}\label{eq.start}
 s^{(k)}_i = 
@@
 \end{cases}
-\end{eqnarray} for integer values of \(i\) with
-\(0 < s^{(k)}_i \le T_k - w_v\).
+\end{eqnarray}
+
+for integer values of \(i\) with \(0 < s^{(k)}_i \le T_k - w_v\).
 
@@
 
-\hypertarget{a-problem-with-failed-tests}{%
 \subsection{A problem with failed
-tests}\label{a-problem-with-failed-tests}}
+tests}\label{a-problem-with-failed-tests}
 
@@
 
-\hypertarget{a-modified-approach}{%
-\subsection{A modified approach}\label{a-modified-approach}}
+\subsection{A modified approach}\label{a-modified-approach}
 
@@
 We then define the pairs for different-shifts as
+
 \begin{equation}\label{eq.diff2}
@@
 \end{cases},
-\end{equation} i.e.~for an odd number of same-shift correlations, we
-skip the middle pair for the different-shift correlations (see
-\autoref{sketch-diff-2}). This pairing ensures that the number of
-different-shift pairings is the same or at most one less than the number
-of same-shift pairings in all tests. In the remainder of the paper, we
-will refer to the algorithm defined by \citet{hadler} as \textbf{(CS1)}
-and the suggested modified algorithm as \textbf{(CS2)} and compare their
-performance on the available scans of the Hamby study.
+\end{equation}
+
+i.e.~for an odd number of same-shift correlations, we skip the middle
+pair for the different-shift correlations (see \autoref{sketch-diff-2}).
+This pairing ensures that the number of different-shift pairings is the
+same or at most one less than the number of same-shift pairings in all
+tests. In the remainder of the paper, we will refer to the algorithm
+defined by \citet{hadler} as \textbf{(CS1)} and the suggested modified
+algorithm as \textbf{(CS2)} and compare their performance on the
+available scans of the Hamby study. The section
+\ref{appendix:appxfailed} of the Appendix discusses about the various
+scenarious of failed tests.
 
@@
 
-\hypertarget{testing-setup}{%
-\subsection{Testing setup}\label{testing-setup}}
+\subsection{Testing setup}\label{testing-setup}
 
-\hypertarget{the-data}{%
-\subsubsection{The Data}\label{the-data}}
+\subsubsection{The Data}\label{the-data}
 
@@
 
-\hypertarget{setup}{%
-\subsubsection{Setup}\label{setup}}
+\subsubsection{Setup}\label{setup}
 
@@
 
-\hypertarget{results}{%
-\section{Results}\label{results}}
+\section{Results}\label{results}
 
-\hypertarget{failed-tests}{%
-\subsection{Failed Tests}\label{failed-tests}}
+\subsection{Failed Tests}\label{failed-tests}
 
@@
 
-\hypertarget{coarseness}{%
-\subsection{Coarseness}\label{coarseness}}
+\subsection{Coarseness}\label{coarseness}
 
@@
 
-\hypertarget{error-rate-assessment}{%
-\subsection{Error rate assessment}\label{error-rate-assessment}}
+\subsection{Error rate assessment}\label{error-rate-assessment}
 
@@
 missed same-source markings. This rate is also known as the type II
-error rate.\\
+error rate. A detailed plot on the type II error rates for CS1 and CS2
+can be found in the section \ref{appendix:appxtype2} of the Appendix.
 Area under the curve (AUC) is shown with confidence intervals as given
@@
 
-\hypertarget{observed-versus-nominal-type-i-error-rates}{%
 \subsection{Observed versus Nominal Type I error
-rates}\label{observed-versus-nominal-type-i-error-rates}}
+rates}\label{observed-versus-nominal-type-i-error-rates}
 
@@
 
-\hypertarget{high-resolution-hamby-44-scans}{%
 \subsection{High resolution Hamby 44
-scans}\label{high-resolution-hamby-44-scans}}
+scans}\label{high-resolution-hamby-44-scans}
 
@@
 
-\hypertarget{conclusions}{%
-\section{Conclusions}\label{conclusions}}
+\section{Conclusions}\label{conclusions}
 
@@
 
-\hypertarget{appendix}{%
-\section{Appendix}\label{appendix}}
+\section{Appendix}\label{appendix}
 
@@
 
-\section{Scenarios of failed tests}
-
+\subsection{Scenarios of failed tests}
+\label{appendix:appxfailed}
 Figures \ref{same-shift-failure} and \ref{diff-shift-failure} show scenarios in which the deterministic Chumbley score can fail. 
@@
 \newpage
-\section{Type 2 errors: CS1 vs CS2}
-
+\subsection{Type 2 errors: CS1 vs CS2}
+\label{appendix:appxtype2}
 \autoref{fig:type2} gives an overview of type 2 error rates of methods CS1 and CS2 for different significance levels $\alpha$. Method CS2 is outperforming CS1 significantly in every instance.

### 2018-06-29 jfs_submission/paper-appendix.tex
jfs_submission.

--- 2018-06-19 doc/paper.tex
@@
 
+  % \bigskip
+
+  % \bigskip
+\newpage
+
 \if0\blind
 {
-  \title{\bf Adaption of the Chumbley Score to matching of bullet striation marks}
-
-  \author{
-        Ganesh Krishnan \thanks{The authors gratefully acknowledge funding from CSAFE through
-cooperative agreement \# 70NANB15H176 between NIST and Iowa State
-University.} \\
-    Department of Statistics, Iowa State University\\
-     and \\     Heike Hofmann \\
-    Department of Statistics and CSAFE, Iowa State University\\
-      }
+  \title{\LARGE\bf APPENDIX \\ Adapting the Chumbley Score to match striae on Land Engraved Areas (LEAs) of bullets}
   \maketitle
@@
 
-\if1\blind
-{
-  \bigskip
-  \bigskip
-  \bigskip
-  \begin{center}
-    {\LARGE\bf Adaption of the Chumbley Score to matching of bullet striation marks}
-  \end{center}
-  \medskip
-} \fi
-
-\bigskip
-\begin{abstract}
-The same-source problem remains a major challenge in forensic toolmark
-and firearm examination. Technological advances in surface imaging allow
-measurements of 3D surfaces at previously unforeseen resolutions and
-enable digitized imaging. Here, we investigate the applicability of the
-Chumbley scoring method \citep{hadler}, developed for screwdriver
-markings, for same-source identification of bullet striae. We provide
-methods to identify parameters that minimize the error rates for
-matching of LEAs using Hamby datasets 44 and 252 measured by NIST and
-CSAFE. We suggest a remedial algorithm to alleviate the problem of
-failed tests in the method, increasing both the power of the test and
-reducing error rates. Type II error rates of the proposed method improve
-by more than one third (Type I error of 0.05) and are on average about
-0.22. This puts the proposed method on similar footing as other single
-feature matching approaches in the literature.
-\end{abstract}
-
-\noindent%
-{\it Keywords:} forensic science, toolmark, cross-correlation, Mann-Whitney U statistic, land engraved areas (LEAs), algorithm, signatures, same-source problem
-\vfill
+% \if1\blind
+% {
+%   \bigskip
+%   \bigskip
+%   \bigskip
+%   \begin{center}
+%     {\LARGE\bf Adapting the Chumbley Score to match striae on Land Engraved Areas}
+%   \end{center}
+%   \medskip
+% } \fi
+%
+%\bigskip
+%\begin{abstract}
+%The same-source problem remains a major challenge in forensic toolmark
+%and firearm examination. Technological advances in surface imaging allow
+%measurements of 3D surfaces at previously unforeseen resolutions and
+%enable digitized imaging. Here, we investigate the applicability of the
+%Chumbley scoring method \citep{hadler}, developed for screwdriver
+%markings, for same-source identification of bullet striae. We provide
+%methods to identify parameters that minimize the error rates for
+%matching of LEAs using Hamby datasets 44 and 252 measured by NIST and
+%CSAFE. We suggest a remedial algorithm to alleviate the problem of
+%failed tests in the method, increasing both the power of the test and
+%reducing error rates. Type II error rates of the proposed method improve
+%by more than one third (Type I error of 0.05) and are on average about
+%0.22. This puts the proposed method on similar footing as other single
+%feature matching approaches in the literature.
+%\end{abstract}
+%
+%\noindent%
+%{\it Keywords:} forensic science, toolmark, cross-correlation, Mann-Whitney U statistic, land engraved areas (LEAs), algorithm, signatures, same-source problem
+%\vfill
 
@@
 
-\newcommand{\hh}[1]{{\textcolor{orange}{#1}}}
-\newcommand{\gk}[1]{{\textcolor{blue}{#1}}}
 \newcommand{\cited}[1]{{\textcolor{red}{#1}}}
+
 \setlength\parindent{0pt}
 
-\tableofcontents
-\newpage
-
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
-
-Same-source analyses are a major part of a Forensic Toolmark Examiner's
-job. In current practice, examiners make these comparisons by visual
-inspection under a comparison microscope and come to one of the
-following four conclusions: identification, inconclusive, elimination or
-unsuitable for examination \citep{afte-toolmarks1998}. These conclusions
-are made on the basis of ``unique surface contours'' of the two
-toolmarks being in ``sufficient agreement'' \citep{afte-toolmarks1998}.
-AFTE describes the term ``sufficient agreement'' as the possibility of
-another tool producing the markings under comparison, as practically
-impossible \citep{afte-toolmarks1998}. Potential subject bias in the
-assessment as well as the lack of specified error rates are the main
-points of criticisms first raised by the National Research Council in
-2009 \citep{NAS:2009} and later emphasized further by the President's
-Council of Advisors on Science and Technology \citep{pcast2016}.
-
-Technological advances, such as profilometers and confocal microscopy
-allow to measure 3D surfaces in a high-resolution digitized form. This
-technology has become more accessible over the last decade, and has made
-its way into topological images of ballistics evidence, such as bullet
-lands and breech faces
-\citep{DeKinder1, DeKinder2, Bachrach1, vorburger2016}. Digitized images
-of 3D surfaces of form the data basis of statistical analysis of
-toolmarks. A statistical approach based on data removes both
-subjectivity from the assessment and allows a quantification of error
-rates for both false positive and false negative identifications.
-
-Methods for matching marks for a variety of tools have been studied in
-the literature (see \autoref{tab:toolmarks-ER} for an overview):
-\citet{manytoolmarks1} and \citet{chumbley} have been analyzing
-screwdriver marks digitized using a profilometer; \citet{manytoolmarks2}
-have investigated 3D marks from screwdriver, tongue and groove pliers
-captured using a confocal microscope; \citet{afte-chumbley} have been
-investigated digitized marks from slip-joint pliers generated by a
-surface profilometer.
-
-Analysis of these digitized markings require the use of statistical
-methods which can quantify the scientific mechanism of comparing
-markings and serve as basis for an error rate calculation.
-\citet{manytoolmarks2} define a relative distance metric and use it as
-similarity measure between two toolmarks. \citet{manytoolmarks1} extract
-many small segments in the markings of two toolmarks and compare
-similarity using a maximum Pearson correlation coefficient. The Chumbley
-scoring method, first introduced by \citet{chumbley}, uses a similar but
-more extensive framework based on a Mann-Whitney U test of the resulting
-correlation coefficients. This approach is non-deterministic, because
-segments are chosen randomly. \citep{hadler} make the score
-deterministic for each pair of toolmarks by choosing segments for
-comparison systematically. This approach also ensures independence
-between segments of striae.
-
-\begin{table}[ht]
-\centering
-\resizebox{\textwidth}{!}{\begin{tabular}{rlrrr}
-  \hline
-Research paper & Method & Data Source & False Positives & False Negatives\\
-  \hline
- \textbf{\citet{manytoolmarks1}} & Maximum Pearson & Screwdrivers & & \\ & Correlation &  & - & - \\ \hline
- \textbf{\citet{chumbley}} & Randomized & Screwdrivers & & \\ (Same-Surface Same-Angle) & Chumbley Score & & 2.3\% & 8.9\%  \\ \hline
-  \textbf{\citet{afte-chumbley}} & Randomized & Slip-joint  &  & \\ & Chumbley Score &  & - & - \\ \hline 
-  \textbf{\citet{hadler}} & Deterministic & Screwdrivers & & \\ (Same-Surface Same-Angle) & Chumbley Score & & 0\% & 6\%  \\ \hline
-  \textbf{\citet{manytoolmarks2}} & Similarity Measure & Screwdrivers & & \\ (Different Surfaces-same angle) & Relative Distance Metric & & 5.9\% &  9.4\% \\ (Same Surfaces-same angle)&  & & 0.22\% &  0\% \\
-   \hline
-\end{tabular}}
-\caption{\label{tab:toolmarks-ER} Error Rates in same-source toolmark analysis reported in the literature. The top four reported papers use some variation of the Chumbley-score method.}
-\end{table}
-
-In this paper, we are investigating the applicability of the Chumbley
-scoring method by \citet{hadler} to assess striation marks on bullet
-lands for same-source identification. Striation marks on bullets are
-made by impurities in the barrel. As the bullet travels through the
-barrel, these imperfections leave ``scratches'' on the bullet surface
-(see top of \autoref{fig:rgl}). Typically, only striation marks in the
-land engraved areas (LEAs) are considered \citep{afte-article1992}.
-Bullet lands are depressed areas between the grooves made by the rifling
-action of the barrel. Compared to toolmarks made by screwdrivers
-striation marks on bullets are typically much smaller, both in length
-and in width. Bullets also have a curved cross-sectional topography.
-
-In same-source comparisons this curvature is usually removed using some
-form of Gaussian filter \citep{ma2004} or non-parametric smoothing
-\citep{aoas}. An overview of some of the error rates reported in the
-literature on bullet matching is given in \autoref{tab:bullets-ER}.
-\citet{chu2013} use an automatic method for counting consecutive
-matching striae (CMS). The authors report an error rate of 52\% for
-known same-source land comparisons to be (incorrectly) identified as
-different-source (false negative) and zero false positives for known
-different-source lands. \citet{ma2004} and \citet{vorburger2011} discuss
-CCF (cross-correlation function) and its discriminating power and
-applicability for same-source analyses of bullets, but do not provide
-any error rates in their discussion. \citet{aoas} use multiple features,
-such as CCF, CMS, D (distance measure), etc. in a random forest based
-method and compare every land against every other land of digitized
-versions of Hamby 252 and Hamby 44 \citep{hamby} published on the NIST
-Ballistics Database \citep{nist}. The authors report an out-of-bag
-overall error rate of 0.46\%, comprised of an error rate of 30.05\% of
-same-source pairs that were not identified and an error rate of 0.026\%
-of different-source pairs that were incorrectly identified as
-same-source.
-
-\begin{table}[ht]
-\centering
-\resizebox{\textwidth}{!}{\begin{tabular}{rlrrr}
-  \hline
-Method & Data Source & False Positives & False Negatives\\
-  \hline
-  \textbf{\citet{aoas}} & LEAs \\
-Consecutive Matching Bullets Striae (CMS) &  &  6.25\% &  33.85\% \\ 
-Consecutive Non-matching Striae (CNMS)  & &  6.26\% & 35.42\%  \\ 
-Average Distance (D) &   &   6.25\% & 45.83\% \\ 
-Cross-correlation Function (CCF) & &  6.25\% & 17.71\%  \\ 
-Sum of Peaks (S) &  &  6.25\% &  18.23\% \\
-   \hline
-   \textbf{\citet{chu2013}} & LEAs \\
-Consecutive Matching Bullets Striae (CMS) &  &  0\% &  52\% \\ 
-   \hline
-   \textbf{\citet{chu2010}} \& \textbf{\citet{ma2004}} & LEAs \\
-Cross-Correlation Function (CCF) &  &  - &  - \\ 
-    \hline
-\end{tabular}}
-\caption{\label{tab:bullets-ER} Error Rates in same-source single feature land-to-land analysis reported in the literature.}
-\end{table}
-
-The Chumbley score provides us with another approach in the same-source
-assessment of bullet striation marks. \citet{chumbley} compare two
-toolmarks for same-source. The data for this study was obtained from 50
-sequentially manufactured screwdriver tips. \citet{chumbley} report
-error rates for markings made by the tips at different angles. For
-markings made at 30 degree the authors report an average false negative
-error rate of 0.089 and an average false positive error rate of 0.023.
-For marks made under angles of 60 and 85 degrees, respectively, the
-false negatives error rate is 0.09 while the rate of false positives
-decreases to 0.01. The paper by \citet{hadler} is based on the same data
-but the authors focus on markings made under the same angle. The error
-rates associated with the deterministic version of the score are
-reported as 0.06 for false negatives and 0 for false positives.
-
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{images/B6-B2-L6-rescaled.png}
-
-\caption{\label{fig:rgl} Image of a bullet land from a confocal light microscope at 20 fold magnification (top) and a chart of the corresponding signature of the same land (bottom).
-The dotted lines connect some peaks visible in both visualizations.}
-
-\end{figure}
-
-In this paper we evaluate the adaptability of the Chumbley score as a
-measure to quantify similarity in land engraved areas (LEAs) on bullets.
-For that we briefly introduce the deterministic method suggested by
-\citet{hadler} in section 2. In the process we provide methods to
-identify parameters that minimize the error rates. We then investigate
-persistent scenarios in which the method proposed by \citet{hadler}
-fails to come to a result. We go on to provide a solution to the failed
-tests problem, consequently increasing the power of the test and
-reducing error rates in the process. We set up a testing framework to
-compare the performance of the two algorithms in section 3 and finally
-discuss results in section 4.
-
-\hypertarget{methods}{%
-\section{Methods}\label{methods}}
-
-\hypertarget{scans-for-land-engraved-areas}{%
-\subsection{Scans for land engraved
-areas}\label{scans-for-land-engraved-areas}}
-
-Comparisons of striae from bullets are usually based on comparisons of
-striae in land engraved areas, which are extracted in form of cross
-sections, called \emph{profiles} \citep{aoas,ma2004}. From profiles
-bullet \emph{signatures} \citep{chu2013,aoas} are extracted as residuals
-of a loess fit or Gaussian filter. This effectively removes topographic
-structure from the data in the attempt to increase the signal to noise
-ratio. \autoref{fig:rgl} shows how the signature from a bullet land
-(bottom) lines up with the image of the land (top) from which it was
-extracted. We can see in the figure how the depth and relative position
-of the striation markings seen in the image are interpreted as peaks and
-valleys in the signature.
-
-There are two sources of scans for sets from the Hamby study available
-to us: scans of Hamby 44 and Hamby 252 are available from the NIST
-database \citep{nist}. The physical Hamby 44 set has also been made
-available to us and has been scanned locally for CSAFE at the Roy
-J.~Carver High Resolution Microscopy Facility using a Sensofar confocal
-light microscope. Scans in the NIST database are made with a NanoFocus
-at 20x magnification. The resolutions of the two instruments are
-different: the NIST scans are taken at a resolution of 1.5625 \(\mu m\)
-per pixel, while the CSAFE scans are available at a resolution of 0.645
-\(\mu m\) per pixel. The length of an average bullet land from Hamby (9
-mm Ruger P85) is about 2 millimeter, resulting in signatures of about
-1200 pixels for NIST scans, and about 3000 pixels for CSAFE scans.
-
-In comparison, scans from the profilometer used by
-\citet{chumbley, hadler} were taken at a resolution of about 0.73
-\(\mu m\) per pixel. The screw driver toolmarks are about 7 mm in length
-\citep{manytoolmarks1}, for a total of over 9000 pixels for the width of
-these scans.
-
-This severe limitation in the amount of available data poses the main
-challenge in adapting the Chumbley score to matching bullet lands,
-because of the resulting loss in power.
-
-\hypertarget{the-chumbley-score-test}{%
-\subsection{The Chumbley Score Test}\label{the-chumbley-score-test}}
-
-A digitized toolmark forms a spatial process \(z(t)\) with location
-indexed by \(t\). \(t\) here denotes equally spaced pixel locations for
-the striation marks under consideration. For a toolmark consisting of
-\(t\) pixels, \(t = 1, ..., T\). Let further \(z^s(t)\) denote a vector
-of markings of length \(s\) starting in location \(t\).
-
-The Chumbley score algorithm takes input in form of two digitized
-toolmarks:
-
-Let \(x(t_1)\), \(t_1 = 1,2,...T_1\) and \(y(t_2)\), \(t_2 = 1,2...T_2\)
-be two digitized toolmarks (where \(T_1\) and \(T_2\), the lengths of
-the two marks, are not necessarily equal). The toolmarks under
-consideration are potentially from two different-sources or the
-same-source.
-
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figures/sigs-profiles-1}
-
-}
-
-\caption{Bullet land profile (left) and the corresponding signature (right) for one of the lands of Hamby-44.}\label{fig:sigs-profiles}
-\end{figure}
-
-In a pre-processing step the two markings are smoothed using a LOWESS
-\citep{lowess} with coarseness parameter \(c\). Originally, this
-smoothing is intended to remove drift and (sub)class characteristics
-from individual markings, however, in the setting of matching bullet
-striae, we can also make use of this mechanism to separate bullet
-curvature in profiles from signatures before matching signatures.
-\autoref{fig:sigs-profiles} shows an example of a bullet land profile
-(left) and the corresponding signature (right).
-
-After removing sub-class structure, the Chumbley score is calculated in
-two steps: an optimization step and a validation step. In the
-optimization step, the two markings are aligned horizontally such that
-within a pre-defined window of length \(w_o\) the correlation between
-\(x(t_1)\) and \(y(t_2)\) is maximized: \[
-\left(t_1^o, t_2^o\right) = \mathop{\arg \max}\limits_{1 \le t_1 \le T_1-w_o, 1 \le t_2 \le T_2-w_o} \text{cor} \left(x^{w_o} (t_1), y^{w_o}(t_2) \right)
-\] This results in an optimal vertical (in-phase) shift of
-\(t_1^o - t_2^o\) for aligning the two markings. We will denote the
-relative optimal locations as \(t_1^*\) and \(t_2^*\), where
-\(t_i^* = t_i^o/(T_i-w_o)\) for \(i=1,2\), such that
-\(t_1^*, t_2^* \in [0,1]\). Once (sub-)class characteristics are
-removed, the relative optimal locations should be distributed according
-to a uniform distribution in \([0,1]\).
-
-In the validation step, two sets of windows of size \(w_v\) are chosen
-from both markings, \autoref[see][]{fig:win-comparison}. In the first
-set, pairs of windows are extracted from the two markings using the
-optimal vertical shift as determined in the first step, whereas for the
-second set the windows are extracted using a different (out-of-phase)
-shift.
-
-More precisely, let us define starting points \(s_i^{(k)}\) for each
-signature \(k = 1, 2\) as \begin{eqnarray}\label{eq.start}
-s^{(k)}_i = 
-\begin{cases}
-t_k^* + i w_v & \text{ for } i < 0 \\
-t_k^* + w_ o + i w_v & \text{ for } i \ge 0,
-\end{cases}
-\end{eqnarray} for integer values of \(i\) with
-\(0 < s^{(k)}_i \le T_k - w_v\).
-
-Same-shift pairs of length \(w_v\) are defined in \citet{hadler} as all
-pairs \((s_i^{(1)}, s_i^{(2)})\) with integer values \(i\) for which
-both \(s_i^{(1)}\) and \(s_i^{(2)}\) are defined. Similarly,
-different-shift pairs are defined as \((s_i^{(1)}, s_{-i-1}^{(2)})\) for
-all \(i\) where both \(s_i^{(1)}\) and \(s_{-i-1}^{(2)}\) are defined
-(see \autoref{sketch-same-diff}).
-
-\begin{figure}[hbtp]
-\centering
-\includegraphics[width=.7\textwidth]{images/sketch-same.png}
-
-\includegraphics[width=.7\textwidth]{images/sketch-diff.png}
-\caption{\label{sketch-same-diff}Sketch of same-shift pairings  (top) and different-shift pairings (bottom).
-Filled in rectangles show pairings resulting in correlations, unfilled rectangles are segments without a match.}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figures/win-comparison-1}
-
-}
-
-\caption{Two same-source markings.
-For convenience, the markings are moved into phase on the left and out-of phase on the right.
-In-phase (left) and out-of-phase (right) samples are shown by the light grey background.
-The Chumbley-score is based on a Mann-Whitney U test of the correlations derived from these two sets of samples.}\label{fig:win-comparison}
-\end{figure}
-
-For both same- and different-shift pairs correlations between the
-markings are calculated. The intuition here is that for two markings
-from the same-source the correlation for the in-phase sample should be
-high, while the correlations of the out-of-phase sample provide a
-measure for the base-level correlation for non-matching marks of a given
-length \(w_v\). The Chumbley score is then computed as a Mann Whitney U
-statistic to compare between in-phase sample and out-of-phase sample. In
-the original method proposed in \citet{chumbley} both in-phase and
-out-of-phase sample are extracted randomly, whereas \citet{hadler}
-proposed the above specified deterministic rules for both samples to
-make the resulting score deterministic while simultaneously avoiding
-overlaps within selected marks to ensure independence.
-
-\hypertarget{a-problem-with-failed-tests}{%
-\subsection{A problem with failed
-tests}\label{a-problem-with-failed-tests}}
-
-Looking closer at \autoref{eq.start}, we see that by definition, some
-number of tests will fail to produce a result. The problem of failed
-tests is first mentioned in \citet{afte-chumbley}. Unfortunately, the
-authors do not provide any percentage of how many tests failed for their
-data. The algorithm fails to produce for two reasons: either the number
-of eligible same-shift pairs is zero, or the number of different-shift
-pairs is zero.
-
-The number of same-shift pairs will be zero, if the optimal locations
-\(t_1^{o}\) and \(t_2^{o}\) are so far apart, that no segments of size
-\(w_v\) are left on the same sides of the optimal locations, i.e.
-\(t_1^{o} < w_v\) and \(t_2^{o} > T_2-w_o-w_v\) or
-\(t_1^{o} < T_1- w_o - w_v\) and \(t_2^{o} < w_v\), i.e.~we have a
-failure rate of \[
-P\left( t_1^{o} < w_v \ \cap \ t_2^{o} > T_2-w_o-w_v\right) + P\left( t_1^{o} < T_1- w_o - w_v \ \cap \ t_2^{o} < w_v\right).
-\] While we can assume that once (sub-)class characteristics are
-removed, optimal locations \(t_i^{o}\) are uniformly distributed across
-the length of the profile, we cannot assume that \(t_1^o\) and \(t_2^o\)
-are independent of each other. In particular, for same-source profiles,
-we would expect a strong dependency between these locations, in which
-case a large difference between locations is unlikely. However, for
-different-source matches, we can assume that locations are independent.
-In that case, we expect a test to fail with a probability of
-\(\frac{2 w_v^2}{(T_1-w_o)(T_2-w_o)}\). For an average length of \(T_i\)
-of 1200 pixels, \(w_o = 120\) pixels and \(w_v = 30\) pixels this
-probability is about 0.0015.
-
-The number of possible different-shift pairs also depends on the
-location of the optimal locations \(t_1^o\) and \(t_2^o\). Whenever the
-optimal locations are close to the boundaries, the number of possible
-pairings decreases and reaches zero, if \(t_i^{(o)} < w_v\) or
-\(t_i^{(o)} > T_i-w_o- w_v\). Assuming a correlation between optimal
-locations \(t_1^o\) and \(t_2^o\) of close to one for same-source
-profiles, this results in an expected rate of failure of
-\(2 w_v / (T_i-w_o)\), or about 5.6\% for an average length of \(T_i\)
-of 1200 pixels, \(w_o = 120\) pixels and \(w_v = 30\) pixels. Assuming
-independence in the optimal locations for different-source profiles the
-expected probability for a failed test is, again,
-\(\frac{2 w_v^2}{(T_1-w_o)(T_2-w_o)}\).
-
-\hypertarget{a-modified-approach}{%
-\subsection{A modified approach}\label{a-modified-approach}}
-
-While failures due to missing correlations from same-shift pairs are
-unavoidable by definition of the Chumbley score, failures due to missing
-correlations from different-shift pairs can be prevented by using a
-different strategy in assigning pairs.
-
-Using the same notation as in \autoref{eq.start}, we define same-shift
-pairs identical to \citet{hadler} as pairs \((s_i^{(1)}, s_i^{(2)})\)
-for all \(i\) where the boundary conditions of both sequences are met
-simultaneously. Let us assume that this results in \(I\) pairs. Define
-\(s_{(j)}^{(k)}\) to be the \(j\)th starting location in sequence
-\(k = 1, 2\),
-i.e.~\(s_{(1)}^{(k)} < s_{(2)}^{(k)} < ... < s_{(I)}^{(k)}\).
-
-We then define the pairs for different-shifts as
-\begin{equation}\label{eq.diff2}
-\left(s_{(j)}^{(1)}, s_{(I-j+1)}^{(1)} \right) \text{ for } j = 
-\begin{cases}
-1, ..., I & \text{ for even } I \\
-1, ..., (I-1)/2, (I-1)/2 + 2, ..., I & \text{ for odd } I
-\end{cases},
-\end{equation} i.e.~for an odd number of same-shift correlations, we
-skip the middle pair for the different-shift correlations (see
-\autoref{sketch-diff-2}). This pairing ensures that the number of
-different-shift pairings is the same or at most one less than the number
-of same-shift pairings in all tests. In the remainder of the paper, we
-will refer to the algorithm defined by \citet{hadler} as \textbf{(CS1)}
-and the suggested modified algorithm as \textbf{(CS2)} and compare their
-performance on the available scans of the Hamby study.
-
-\begin{figure}[hbtp]
-\centering
-\includegraphics[width=.7\textwidth]{images/sketch-diff-2.png}
-\caption{\label{sketch-diff-2}Sketch of adjusted different-shift pairings. At most one of the same-shift pairings can not be matched with a different-shift pair under algorithm (CS2). }
-\end{figure}
-
-\hypertarget{testing-setup}{%
-\subsection{Testing setup}\label{testing-setup}}
-
-\hypertarget{the-data}{%
-\subsubsection{The Data}\label{the-data}}
-
-Lands for all Hamby-44 and Hamby-252 scans are made available through
-the NIST ballistics database \citep{nist} and are considered, here. Both
-of these sets of scans are part of the larger Hamby study \citep{hamby}.
-Each set consists of twenty known bullets (two each from ten
-consecutively rifled Ruger P85 barrels) and fifteen questioned bullets
-(each matching one of the ten barrels). Ground truth for both of these
-Hamby sets is known and was used to assess correctness of the tests
-results.
-
-Profiles for each bullet land were extracted from scans close to the
-heel of the bullet while avoiding break-off as described in
-\citet{aoas}.
-
-\hypertarget{setup}{%
-\subsubsection{Setup}\label{setup}}
-
-Both algorithms (CS1) and (CS2) are implemented in R \citep{R}. (CS1) is
-available from package \texttt{toolmaRk} \citep{toolmark}, (CS2) is
-available from a modified version of the \texttt{toolmaRk} package
-available from GitHub (\url{https://github.com/heike/toolmaRk}). We
-applied both methods to all pairwise land-to-land comparisons of the
-Hamby scans provided by NIST for a total of 85,491 land-to-land
-comparisons.
-
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-\hypertarget{failed-tests}{%
-\subsection{Failed Tests}\label{failed-tests}}
-
-As described above, the Chumbley-score is based on three parameters:
-coarseness \(c\) and the sizes of the optimization window \(w_o\) and
-validation window \(w_v\). In a first run of results, we applied default
-settings for the parameters, as suggested in \citet{hadler}:
-\(w_o = 120\) pixels or about 190 \(\mu m\) (ten percent of the average
-length of profiles) and coarseness \(c = 0.25\), and varied the size of
-the validation window \(w_v\) in steps of 10 from 10 pixels to 60
-pixels. Based on a significance level \(\alpha\) of 0.05 for the test,
-this results in a correct identification of same-source and
-different-source toolmarks of 93.5\% to 94.1\%, corresponding to a rate
-of false negatives between 0.28 and 0.36 and a rate of false positives
-between 0.05 and 0.06. However, the most prominent result we
-encountered, are the high number of failed tests, i.e.~the number of
-instances, in which CS1 did not return any result. \autoref{fig:fails}
-shows the percentage of failed tests among the 85,491 land-to-land
-comparisons of the NIST data for different values of the validation
-window size \(w_v\). For same-source lands up to 12.5 percent of the
-tests fail using CS1. The highest percentage of failed tests under CS2
-is 1.3\% for different-source tests using a validation window size
-\(w_v\) of 60 pixels. Rates of expected failures are based on simulation
-runs using covariances between locations of same-source profiles of
-0.854, and 0.120 for locations from different-source profiles, matching
-observed covariances for the Hamby scans. Observed failure rates are
-higher than expected. This might be due to remaining sub-class structure
-at a coarseness of 0.25 resulting in a distribution of optimal locations
-different from the assumed uniform.
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.7\textwidth]{figures/fails-1}
-
-}
-
-\caption{Percent of failed land-to-land comparisons using an optimization window $w_o = 120$ and a coarseness of $c = 0.25$.
-With an increase in the size of the validation window  a higher percentage of tests fails under both methods (CS1) and (CS2), but the percentage of failed tests is much smaller under (CS2).
-Observed failure rates of (CS1) are higher than expected rates.}\label{fig:fails}
-\end{figure}
-
-\hypertarget{coarseness}{%
-\subsection{Coarseness}\label{coarseness}}
-
-The purpose of the coarseness parameter is to remove (sub-)class
-characteristics from profiles before comparisons for matching.
-\citet{hadler} suggest a coarseness parameter of 0.25 in the setting of
-toolmark comparisons. For bullet lands, coarseness might need to be
-adjusted because of the strong effect bullet curvature has on profiles.
-
-\autoref{fig:profile-sketch} gives an overview of the effect of
-different coarseness parameters: from left to right, coarseness levels
-\(c\) are varied in steps of 0.05 from 0.1 to 0.3. The top row shows
-resulting signatures after smoothing the profile shown in
-\autoref{fig:sigs-profiles} with different levels of coarseness. The
-histograms in the bottom row show the relative optimal location \(t^*\).
-Optimal locations are distributed uniformly once (sub-)class
-characteristics are removed. However, for coarseness values of
-\(c > 0.20\) we see quite distinct boundary effects: optimal locations
-\(t^*\) are found at the very extreme ends of a profile more often than
-one would expect based on a uniform distribution.
-
-The key effect of the optimal locations and thereby the coarseness is
-seen in the number of failed tests. Irrespective of whether CS1 or CS2
-is being used, if the relative optimal locations are at the boundaries
-we will see an increase in the number of failed tests. A balance is
-therefore needed in the selection of the coarseness parameter which
-reduces the boundary effect but does not remove important individual
-characteristics. Based on \autoref{fig:profile-sketch} a coarseness
-value of \(c = 0.15\) seems to be best suited to strike this balance for
-this example. For the remainder of the analysis, we will use this value
-for \(c\).
-
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figures/profile-sketch-1}
-
-}
-
-\caption{Overview of the effect of different coarseness parameters $c$ on the profile shown in Figure \ref{fig:sigs-profiles} (top).
-The bottom row shows histograms of  the (relative) optimal locations $t^o$ identified in the optimization step for different values of the coarseness parameter $c$.
-}\label{fig:profile-sketch}
-\end{figure}
-
-\hypertarget{error-rate-assessment}{%
-\subsection{Error rate assessment}\label{error-rate-assessment}}
-
-\autoref{fig:roc} gives an overview of ROC (Receiver operating
-characteristic) curves for methods CS1 and CS2 over a range of different
-optimization window sizes \(w_o\) and two sizes for the validation
-window \(w_v\) (shape). The different color hues represent the two
-methods CS1 (red) and CS2 (blue). The ROC curves show the superior
-performance of CS2 over CS1. Generally, an optimization window \(w_o\)
-of 150 pixels or more leads to the best performance with respect to ROC
-curves. Results based on a validation window of size \(w_v = 30\) are
-generally better than results for \(w_v = 50\).
-
-\autoref{fig:eer} shows a comparison of the performance of the two
-methods CS1 and CS2 with respect to EER (equal error rate) and AUC (area
-under the curve) corresponding to the ROC curves shown in
-\autoref{fig:roc}. Equal error rates are reduced using method CS2, while
-area under the curve significantly increases (at a significance level
-\(\alpha\) of 0.05) compared to method CS1.
-
-The results from Figures \ref{fig:roc} and \ref{fig:eer} are summarized
-in numbers in \autoref{tab:aucs}. Equal error rates (EER), rates for
-false positives (FPR) and false negatives (FNR) are shown side by side
-with the area under the curve (AUR) for both methods for a set of
-different optimization windows \(w_o\) and a validation window \(w_v\)
-of 30 pixels. The rate of false positive same-source identifications is
-equal to the statistical type I error rate, which is set to
-\(\alpha = 0.05\) for this example. The rate of false negatives are
-missed same-source markings. This rate is also known as the type II
-error rate.\\
-Area under the curve (AUC) is shown with confidence intervals as given
-by \citet{delong}. CS2 significantly outperforms CS1 with respect to its
-predictive power in most situations.
-
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figures/roc-1}
-
-}
-
-\caption{ROC curves of methods CS1 and CS2 for different sizes of optimization window $w_o$.
-Best performances with respect to  ROC curves are reached for optimization windows of sizes 150 and higher.
-Points of equal error rates (EERs) can be found at the intersection of the dotted line and the ROC curves.}\label{fig:roc}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figures/eer-1}
-
-}
-
-\caption{Comparison of results for CS1 and CS2 using area under the curve (AUC) and equal error rates (EER). }\label{fig:eer}
-\end{figure}
-
-\begin{table}[hbtp]
-\centering
-\begin{adjustbox}{max width=\textwidth}
-\begin{tabular}{rrrlrrl}
-  \hline
-  & \multicolumn{3}{l}{\bf CS1} & \multicolumn{3}{l}{\bf CS2} \\
-$w_o$ & \bf FPR & \bf FNR & \bf AUC (95\% C.I.) & \bf FPR & \bf FNR & \bf AUC (95\% C.I.) \\ 
-  \hline
-90 & 0.068 & 0.330 & 0.850 (0.835, 0.865) & 0.075 & 0.242 & 0.877 (0.863, 0.890) \\ 
-  120 & 0.065 & 0.309 & 0.864 (0.850, 0.878) & 0.072 & 0.236 & 0.890 (0.877, 0.903) \\ 
-  130 & 0.065 & 0.317 & 0.863 (0.850, 0.877) & 0.072 & 0.224 & 0.898 (0.886, 0.910) \\ 
-  140 & 0.065 & 0.304 & 0.873 (0.859, 0.886) & 0.071 & 0.229 & 0.902 (0.891, 0.914) \\ 
-  150 & 0.067 & 0.326 & 0.863 (0.850, 0.877) & 0.072 & 0.225 & 0.904 (0.892, 0.916) \\ 
-  180 & 0.061 & 0.326 & 0.877 (0.864, 0.890) & 0.070 & 0.229 & 0.907 (0.896, 0.919) \\ 
-  210 & 0.064 & 0.342 & 0.865 (0.851, 0.878) & 0.066 & 0.231 & 0.906 (0.895, 0.918) \\ 
-   \hline
-\end{tabular}
-\end{adjustbox}
-\caption{\label{tab:aucs} Overview of results as shown in Figures \ref{fig:roc} and \ref{fig:eer}.
-FPR is the observed rate of false positives (for a fixed level $\alpha = 0.05$), FNR is the false negative rate.
-Area under the curve (AUC) is shown with confidence intervals.
-}
-\end{table}
-
-\hypertarget{observed-versus-nominal-type-i-error-rates}{%
-\subsection{Observed versus Nominal Type I error
-rates}\label{observed-versus-nominal-type-i-error-rates}}
-
-\autoref{fig:type1} shows observed type I error rates (FPR) across a
-range of optimization windows \(w_o\). Generally, observed type I errors
-are higher than expected. Method CS1 shows in this instance slightly
-better performance than method CS2, but for both an increase in the size
-of the optimization window leads to a decrease in the observed type I
-error rates.
-
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figures/type1-1}
-
-}
-
-\caption{Comparison of observed and nominal type I error rates  across a range of window sizes for optimization $wo$.
-The horizontal line in each facet indicates the nominal type I error rate.}\label{fig:type1}
-\end{figure}
-
-\hypertarget{high-resolution-hamby-44-scans}{%
-\subsection{High resolution Hamby 44
-scans}\label{high-resolution-hamby-44-scans}}
-
-The high-resolution scans of Hamby set 44 are capturing images at a
-resolution of \(0.645 \mu m\) per pixel. On average, land engraved areas
-are 3000 pixels in length. A coarseness of \(c = 0.125\) seemed to be
-sufficient in removing any bullet curvature. Both methods have a failed
-test rate of less than 0.006, indicating, again, that the larger number
-of pixels alleviates the problem of test failures. \autoref{fig:h44}
-shows the resulting EER and AUC for methods CS1 and CS2 based on two
-sizes of validation windows (\(w_v \in 75, 125\)) and optimization
-window sizes around 300 pixels (10 percent of the average length). Both
-methods show an increase in performance around \(w_o = 300\) pixels. CS2
-out-performs CS1 in all scenarios, but the difference is not significant
-(using DeLong's confidence intervals). Interestingly, the overall
-performance of both CS1 and CS2 is a lot lower for the high resolution
-version of Hamby-44 than for the lower resolution scans. The area under
-the curve overall is significantly lower for the high resolution scans
-than for the previous set of scans. Partly, this might be due to the
-particular choice of the parameters, partly the higher-resolution scans
-might be picking up on real differences between the lands that the
-lower-resolution scans fail to detect.
-
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figures/h44-1}
-
-}
-
-\caption{AUC and EER for methods CS1 and CS2 based on high resolution scans of Hamby 44. }\label{fig:h44}
-\end{figure}
-
-\hypertarget{conclusions}{%
-\section{Conclusions}\label{conclusions}}
-
-In assessing the suitability of the (deterministic) Chumbley Score for
-matching striae on bullet lands we have gained valuable insights into
-the process: method CS1 as proposed by \citet{hadler} has a strong
-dependency on the specific choice of parameters; the defaults suggested
-by \citet{hadler} for screw drivers are not directly applicable for the
-smaller bullet lands. The coarseness parameter in particular has a
-strong impact on the performance of the test. However, we were able to
-suggest some heuristics based on the assumption that once sub-class
-characteristics are removed, optimal locations are distributed uniformly
-across the profile. For bullet lands we found a coarseness value of
-\(c = 0.15\) to be suitable for the low-resolution scans from NIST and a
-value of \(c=0.125\) suitable for the higher-resolution scans from
-CSAFE. Sizes for optimization windows \(w_o\) were based on
-cross-validation to minimize overall type 2 error rates. Unfortunately,
-this kind of assessment is only feasible in the setting of a large
-study, such as the one we presented, and does not transfer immediately
-to case work, where a forensic examiner would only deal with a few
-identifications. Parameter settings for different studies should be
-investigated and reported to allow for a further refinement.
-
-Method CS1 proposed by \citet{hadler} has a minimal type 2 error rate of
-0.272 for an optimized window size of 140 pixels -- which is
-considerably higher than the error rates achieved on matching toolmarks,
-but is similar to other single-feature methods proposed for bullet
-matching. Unfortunately, method CS1 also has a high rate of failed tests
--- situations, in which the algorithm does not provide a result, due to
-the way different-shift pairs are constructed. Algorithm CS2 is
-introduced here as a remedy for failed tests by introducing an alternate
-version of choosing different-shift pairs. The algorithm CS2 introduced
-here is constructed in a way that achieves on average a ten-fold
-reduction in the number of failures. While reducing the failure rate,
-the algorithm also shows an increase in the power of the test. Type II
-error rates of CS2 reach a minimum of 0.217 for an optimized window size
-of 130 pixels. This increase in power of CS2 over CS1 should also apply
-to previous studies on toolmarks. It would be interesting to see these
-results using the adjusted algorithm. Unfortunately, none of the studies
-have made the data publicly accessible.
-
-While significantly reduced over CS1, CS2 still has type 2 error rates
-on bullet lands that are higher than the error rates achieved on the
---much larger-- toolmarks. Applying these methods to the high-resolution
-scans provided by CSAFE shows that better scanning methodology does not
-guarantee a better matching performance.
-
-However, bullets usually have multiple lands -- in the case of Ruger
-P85s as used in the Hamby study, there are six lands for each bullet. We
-might be able to get more power out of the test by adapting CS2 to work
-in a bullet-to-bullet comparison.
-
-\hypertarget{appendix}{%
-\section{Appendix}\label{appendix}}
+%\tableofcontents
+%\newpage
+\section*{Appendix}\label{appendix}
 
@@
 \section{Scenarios of failed tests}
-
+\label{appendix:appxfailed}
 Figures \ref{same-shift-failure} and \ref{diff-shift-failure} show scenarios in which the deterministic Chumbley score can fail. 
@@
 \section{Type 2 errors: CS1 vs CS2}
-
+\label{appendix:appxtype2}
 \autoref{fig:type2} gives an overview of type 2 error rates of methods CS1 and CS2 for different significance levels $\alpha$. Method CS2 is outperforming CS1 significantly in every instance.
@@
 
-\bibliographystyle{agsm}
-\bibliography{bibliography}
+%\bibliographystyle{agsm}
+%\bibliography{bibliography}
 

### 2018-06-29 jfs_submission/paper-main-manuscript.tex
jfs_submission.

--- 2018-06-29 doc/paper.tex
@@
 
+\usepackage{verbatim}
 \usepackage{url} % not crucial - just used below for the URL
@@
 {
-  \title{\bf Adapting the Chumbley Score to match striae on Land Engraved Areas of
-bullets}
-
-  \author{
-        Ganesh Krishnan,\\ %\\
-    Master of Science (M.Sc.)\\
-    % \\
-     and \\     Prof.~Heike Hofmann,\\ %\\
-    Doctor of Philosophy (Ph.D.)\\
-    % \\
-      }
-    \maketitle
-  \bigskip
-  \begin{center}
-  {\textbf{Affiliation(s)}\\
-  Department of Statistics, Iowa State University and CSAFE (Center for
-  Statistics and Applications in Forensic Evidence)\\
-  \bigskip
-  \textbf{Source of funding}\\
-  The authors gratefully acknowledge funding from CSAFE through
-  cooperative agreement \# 70NANB15H176 between NIST and Iowa State
-  University.\\
-  \bigskip
-  \textbf{Acknowledgement}\\
-  We would like to thank the efforts of CSAFE and Roy J Carver Microscopy
-  lab in scanning the Hamby-44 set and providing the scans to us. We would
-  also like to thank NIST for providing the Hamby-44 bullets to us for
-  scanning.\\
-  \bigskip
-  \textbf{Information on presentation of work}\\
-  This work was presented in the form of a poster in the AFTE 2018 49th
-  Annual Training Seminar, held from June 3rd to June 8th, 2018 in
-  Charleston, West Virginia}
-  \end{center}
-
-} \fi
-
-  % \bigskip
-
-  % \bigskip
-\newpage
-
-\if0\blind
-{
-  \bigskip
-  \bigskip
-  \bigskip
-  \begin{center}
-  \title{\LARGE\bf Adapting the Chumbley Score to match striae on Land Engraved Areas of
-bullets}
-  \end{center}
-  \bigskip
+  \title{\LARGE\bf Adapting the Chumbley Score to match striae on Land Engraved Areas (LEAs) of bullets}
   \maketitle
@@
 %   \begin{center}
-%     {\LARGE\bf Adapting the Chumbley Score to match striae on Land Engraved Areas of
-bullets}
+%     {\LARGE\bf Adapting the Chumbley Score to match striae on Land Engraved Areas}
 %   \end{center}
@@
 
-\newcommand{\hh}[1]{{\textcolor{orange}{#1}}}
-\newcommand{\gk}[1]{{\textcolor{blue}{#1}}}
 \newcommand{\cited}[1]{{\textcolor{red}{#1}}}
@@
 algorithm as \textbf{(CS2)} and compare their performance on the
-available scans of the Hamby study. The section
-\ref{appendix:appxfailed} of the Appendix discusses about the various
-scenarious of failed tests.
+available scans of the Hamby study. The section  A %\ref{appendix:appxfailed} 
+of the Appendix discusses about the various scenarious of failed tests.
 
@@
 error rate. A detailed plot on the type II error rates for CS1 and CS2 
-can be found in the section \ref{appendix:appxtype2} of the Appendix.
+can be found in the section B %\ref{appendix:appxtype2} 
+of the Appendix.\\
 Area under the curve (AUC) is shown with confidence intervals as given
@@
 
+\begin{comment}
 \section{Appendix}\label{appendix}
@@
 
-\subsection{Scenarios of failed tests}
+\section{Scenarios of failed tests}
 \label{appendix:appxfailed}
@@
 \newpage
-\subsection{Type 2 errors: CS1 vs CS2}
+\section{Type 2 errors: CS1 vs CS2}
 \label{appendix:appxtype2}
@@
 \end{appendix}
-
+\end{comment}
+\newpage
 \bibliographystyle{agsm}

### 2018-08-06 jfs_submission/paper.Rmd
getting ready for review

--- 2018-06-19 doc/paper.Rmd
@@
 ---
-title: "Adaption of the Chumbley Score to matching of bullet striation marks"
+title: "Adapting the Chumbley Score to match striae on Land Engraved Areas of bullets"
 authors:
-- affiliation: Department of Statistics, Iowa State University 
-  name: Ganesh Krishnan
-  thanks: The authors gratefully acknowledge funding from CSAFE through cooperative agreement \# 70NANB15H176 between NIST and Iowa State University. 
-- affiliation: Department of Statistics and CSAFE, Iowa State University 
-  name: Heike Hofmann
+# - affiliation: Department of Statistics, Iowa State University 
+- name: Ganesh Krishnan,
+  degree: Master of Science (M.Sc.)
+  # thank: The authors gratefully acknowledge funding from CSAFE through cooperative agreement \# 70NANB15H176 between NIST and Iowa State University.
+We would also like to thank the efforts of CSAFE and Roy J carver Microscopy lab in scanning the Hamby-44 set and providing the scans to us.
+# - affiliation: Department of Statistics and CSAFE, Iowa State University 
+- name: Prof. Heike Hofmann,
+  degree: Doctor of Philosophy (Ph.D.)
+affiliation: Department of Statistics, Iowa State University and CSAFE (Center for Statistics and Applications in Forensic Evidence)
+funding: The authors gratefully acknowledge funding from CSAFE through cooperative agreement \# 70NANB15H176 between NIST and Iowa State University. 
+thank: We would like to thank the efforts of CSAFE and Roy J Carver Microscopy lab in scanning the Hamby-44 set and providing the scans to us.
+We would also like to thank NIST for providing the Hamby-44 bullets to us for scanning.
+presentedat:  This work was presented in the form of a poster in the AFTE 2018 49th Annual Training Seminar, held from June 3rd to June 8th, 2018 in Charleston, West Virginia
 biblio-style: apsr
@@
 In the remainder of the paper, we will refer to the algorithm defined by \citet{hadler} as \textbf{(CS1)} and the suggested modified algorithm as \textbf{(CS2)} and compare their performance on the available scans of the Hamby study.
+The section \ref{appendix:appxfailed} of the Appendix discusses about the various scenarious of failed tests.
 \begin{figure}[hbtp]
@@
 ggerrors <- errorsCV %>%
-  ggplot(aes(x = wo, y = beta, colour=factor(alpha))) +
+  ggplot(aes(x = wo, y = beta, colour=factor(alpha), linetype=factor(alpha))) +
   geom_point(aes(shape=factor(wv)), size=2.5) +
@@
   scale_colour_brewer(expression("Nominal\ntype I error"~alpha),
-                      palette="Set2") +
+                      palette="Set2", guide=FALSE) +
+  scale_linetype_manual(values = c(2,1,3,4), guide=FALSE)+
   scale_shape_discrete(expression("Size of\nvalidation window "~w[v])) +
+  guides(
+    colour = guide_legend(override.aes = list(linetype= c(2,1,3,4)))
+  )+
   xlab(expression("Window size for optimization "~w[o])) +
@@
 This rate is also known as the type II error rate.
+A detailed plot on the type II error rates for CS1 and CS2 can be found in the section \ref{appendix:appxtype2} of the Appendix.
 Area under the curve (AUC) is shown with confidence intervals as given by \citet{delong}. CS2 significantly outperforms CS1 with respect to its predictive power in most situations.
@@
   geom_roc(labels=FALSE, n.cuts=500, pointsize=0, size=0.6, 
-           aes(colour = method.wo)) +
+           aes(colour = method.wo, linetype= method.wo)) + 
   facet_wrap(~wv, labeller = label_bquote(w[v] == .(wv))) +
+    scale_linetype_manual(values=c(2,2,2,2,2,1,1,1,1,1), guide = FALSE) +
   scale_alpha_discrete(range=c(0.1, 0.6)) +
@@
   guides(
-   colour = guide_legend(ncol=2)
+   colour = guide_legend(override.aes = list(linetype = c(2,2,2,2,2,1,1,1,1,1)), ncol=2)
   ) +

### 2018-08-07 Response.Rmd
adding response

---
title: "Response to the reviewers"
author: "Ganesh Krishnan and Heike Hofmann"
date: "8/6/2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We would like to thank the reviewers for their in-depth reviews and the suggestions regarding the manuscript. Please find our detailed response addressing each of the points below.

## Editorial changes

- "Table of contents - please delete
- Introduction header - please delete 
- References by name - these need to be cited numerically
- Tables in text - please include the tables and their headers in the - Main Document as separate pages after  the References when uploading the revision.  Please delete from the text
- Figures in text - all figures must be uploaded as separate TIF, PNG or EPS files with at least 300 dpi.  Please delete from the text
- Headers, sub-headers and sub-sub-headers numbered - please remove the numbering
- Sub-headers bolded and not italicized - these need to be italicized only
- Sub-sub-headers bolded - please do not bold
- References incorrect format - please follow the formats included in the Information for Authors
- Figures loaded as separate PNGs but these are not 300 dpi - see above
- Two files uploaded - one as bib file and one as tex  - I am not sure what these are?  However, we only except text in Word files.  If these were intended to be included in the manuscript file, please include them and do not load the information as separate files.  
- Title Page, Manuscript and Appendix uploaded as PDFs - please upload as Word files."

It is imperative that the format guidelines outlined in the JFS Information for Authors be followed in preparing the revision.  I have to say it was disappointing that you made no attempt to follow the correct format.

Please also note that when loading figures into Manuscript Central it is important to Insert FIG. X before each figure legend. In addition, you must load each figure as a separate file.

Please note that all figure files (TIFF, PNG or EPS) MUST be at least 300 dpi and that some of your figures do not satisfy this requirement. Please modify your figures accordingly.

Please include the tables and their headers in the Main Document as separate pages after the References section when uploading the revision.

There is also a need to strengthen the Abstract. Although we do not yet use a structured abstract, we do expect the same information - background, brief description of methods and results, and conclusions.

Upon acceptance for publication, manuscripts become the copyright property of AAFS. The corresponding author is responsible for transferring copyright to AAFS through the Wiley Author Licensing Service (or WALS). In doing so the corresponding author confirms that all authors have contributed to the manuscript per the JFS Information for Authors and all have agreed that the accepted manuscript will be published in JFS.

To revise your manuscript, log into https://mc.manuscriptcentral.com/jofs and enter your Author Center, where you will find your manuscript title listed under "Manuscripts with Decisions." Under "Actions," click on "Create a Revision." Your manuscript number has been appended to denote a revision.

You will be unable to make your revisions on the originally submitted version of the manuscript. Instead, revise your manuscript using a word processing program and save it on your computer. Once the revised manuscript is prepared as a MS Word document, you can upload it and submit it through your Author Center.

When submitting your revised manuscript through your Author Center, you must respond to the comments made by the reviewer(s) in the space provided. Please use this space to document all changes you make to the original manuscript. In order to expedite the processing of the revised manuscript, please be as specific as possible in your response to the reviewer(s). It is also imperative that you follow the required format provided in the JFS Information for Authors, particularly those for the References.  As noted above, if your revised mansucript is not in the JFS required format, it may be returned for further revision.

IMPORTANT: Your original files are available to you when you upload your revised manuscript. Please delete any redundant files before completing the submission. Your revision submission MUST include your complete manuscript, i.e., title page, main document, tables and figures (if applicable).

Because we are trying to facilitate timely publication of manuscripts submitted to the Journal of Forensic Sciences, your revised manuscript should be uploaded within 21 days of receipt of this email. If it is not possible for you to submit your revision in this time frame, we may have to consider your paper as a new submission.

Once again, thank you for submitting your manuscript to the Journal of Forensic Sciences and I look forward to receiving your revision.

## Reviewer #1

The article ???Adapting the Chumbley Score to match striae on Land Engraved Areass (LEAs) of bullets??? extends an existing comparison algorithm for striated marks to the analysis of bullet LEAs. The paper makes two primary contributions. First, they modify the deterministic work of Hadler & Morris (2017) which itself is an extension of Chumbley et al. The authors??? contribution is a modification of Hadler & Morris to reduce the number of scenarios which ???fail??? under the Hadler & Morris approach. Second, they evaluate the work on two sets of bullet LEA data and explore some variants of the algorithm parameters to tune it for bullets rather than screwdrivers. Overall the paper presents a reasonable project to extend a nice algorithm for striated toolmark comparison.

Adaptation of the Chumbley el al algorithm from screwdrivers to bullets makes sense as both analyses involve striated marks. The authors here nicely point out that compared to screwdriver marks (as in the original Chumbley paper), LEAs are typically smaller both in length and width. LEAs also have a curved cross-section baseline which must be corrected.

The authors do a nice job covering previous work. I appreciate the summary tables 1 and 2. Table 1 is good as-is. Table 2 can be a bit misleading because we don???t know what bullet datasets were used. Because these different papers analyzed different bullet data sets the underlying difficulty of matching impacts the performance. That is, bullets from very well marking firearms are ???easier??? to match than those from firearms which mark poorly. So the FP and FN reported should be taken with a grain of salt. The authors should add a statement regarding different performance on different datasets.

I???m also confused how the FP rate for Hare CNMS is 6.26% whereas all other FP are 6.25% This seems to likely be a typo. 6.25% imply close to 10,000 comparisons which I don???t think the datasets have. I could be wrong but I wanted to point out this potential typo that should be checked.

The authors describe the two methods of ???failure??? for the method (Sec 2.3). I agree with the authors statement that this failure ???should??? only happen for non-matches. It could also happen for matches if the optimal profiles are incorrect. That is, if there are two possible optimal match pairs (with very similar correlation scores), where the top scoring pair is incorrect but the second best scoring pair is correct. I???m not sure how often that happens. Overall, I would think that failing for non-matches would be ok. It???s not clear that the described failures are the biggest cause of performance hits but improving the algorithm to handle these failures seems reasonable.

The following sections list specific points that should be addressed before acceptance. They are grouped into clarification/grammar/preciseness, minor, and major revisions.

Some examples of clarification / grammar / preciseness issues:
- p3 ???confocal microscopy allow * to measure 3D surfaces in a high-resolution digital form???  * should be ???one????
- p3 ???Digitized images of 3D surfaces of form the data basis* of statistical analysis??? Awkward phrasing.
- p3 ???Faden et al and Chumbley et al have been analyzing screwdriver marks??? perhaps change to ???have analyzed??? because the work is completed and published.
- p9 Seems to be a broken latex figure reference around line 28/29.
These are not all the issues of this type, we encourage the authors to review the manuscript carefully clean issues like the above.

Minor edits should be made to accuracy and clarity:
- p3 ???Methods for matching marks for a variety of tools??? should be changed to ???Methods for matching striated marks for a variety of tools???. This is an important point in that it helps make clear that the authors are studying and comparing striated toolmarks and not impressed toolmarks. I would also propose that the authors make this distinction earlier in the introduction.
- p3 ???Analysis of these digitized markings require the use of statistical methods???. Analysis does not *require* statistical methods. Useful information regarding common origin can be obtained from non-statistical methods. This being said, statistical methods are of course great and we should be working towards them. Authors could change to analysis ???benefits from the use of statistical methods??? or something similar.
- p14 how are the false negative and false positive rates computed?

Major Changes Required before acceptance:
- p8 ???This severe limitation in the amount of available data poses the main challenge in???. It???s unclear what ???this??? refers to. Do the authors mean the number of bullets and LEAs in the dataset or do they mean the number of sample points per striation profile (due to the fact that LEAs are shorter than screwdriver profiles)? Both would seem to be limitations but just because the profile is shorter doesn???t inherently mean that there are fewer or less informative striaes. You could have a shorter profile with more informative features.
- Eq 1 uses a variable i, which is defined above as having the values of 1 or 2. I therefore don???t understand how i<0 as written in Eq 1. This should be improved. Possibly this is a sloppy reuse of variables?
- How are the original profiles (eg. Figure 2 Left) identified from the 3D surface (eg. Figure 1 Top). That is, does the user simply draw a line from one end of the LEA to the other? How is the original 3D surface resampled to the 1d profile? This is very important because if you ???cut??? the 3D profile the wrong way (at the wrong angle) then you will dilate or compress the extracted profile. This needs a lot more description.
- Section 2.4 is not clearly written. I???m not confident that I understand their new method. This section should be rewritten to explain what the authors are doing. My guess is that rather than taking windows in integer number of steps away from the optimal position matching a left shift on one profile with a right shift on the other, the authors simply walk from the right-end to the left-end of one profile and from the left-end to the right-end of the other profile being careful to not compare same shift windows. Note that my description is not clear either as this is a tricky thing to put into words, but is something the authors need to do.
- The results presented on page 14 list the correct identification of ???same-source and difference-source toolmarks of 93.5 to 94.1%???. Does this mean the true positives are 93.5% and the true negatives are 94.1%? I don???t think so because the next sentence lists the false negative rate of between 0.28 and 0.36. The authors need to explicitly list the true positive rate and the true negative rate. These must be separate because there are many more true negatives than true positives in any matching dataset. That means that an algorithm that always predicts ???different source??? would be correct most of the time. In fact, for the author???s dataset the ???always predict different??? algorithm would be correct approximately 94% of the time (35 bullets where each has appx 2 matches). The lack of a true positive rate and true negative rate is is a major point that must be corrected. To be clear, the true positive rate, true negative rate, false positive rate, and false negative rate should all be reported for the first dataset, perhaps added to Table 3 or another table like Table 3.
- The true positive rate, true negative rate, false positive rate, and false negative rate should also be presented for the second dataset (Hamby 44 scans, Sec 3.5). The authors do present AUC and EER curves but these are less informative and can often be used to hide weaker performing models.
- The results should all be listed as either fractions or all as percent (which is preferred) and not mix them.
- p21 the authors state ???this kind of assessment is only feasible in the setting of a large study, such as the one we presented???.  I strongly disagree with the use of the word ???large???. The presented work is not a large study. Ten firearms were used in each of two studies (it???s not clear if both datasets used the same 10 firearms). A large study would include at least 100 different firearms and even 100 is not likely to provide statistical confidence. This is a very important point. If the paper is allowed to state that the described work is a large study it will set a precedent that will be detrimental to the field.

Overall, this paper presents a nice incremental improvement to a promising algorithmic approach. A number of major revisions are required prior to acceptance as described above. Note that these revisions all relate to the writing and require no additional experiments to be performed. They improve the clarity and preciseness of the manuscript and serve to address questions that readers will have.

## Reviewer #2

Comments to the Author
You may not fully appreciate and how critical their phase orientations used to identify fired bullets (LEAs).  This important consideration should have been an important  aspect of the experiment and not instead focus primarily only LEA comparisons.  While each area is in practice independently compared microscopically or by surface profile algorithms, consideration as to their orientation in their position in the barrel rifling, and correlations of LEAs cannot ignore the phase of positions.  Including the phase consideration of the fired bullet comparisons, the success of finding a match in a particular position between two bullets is significantly increased and demonstrated in their reference Chu, et al.  If you are to base independent single LEA scoring to test the method, you should make this clearer in the beginning of your paper and hypothesis, and not discuss it finally in the last sentence of the document body.
If you are going to compare method performance between chisel/tool stria to LEAs on bullets, you should have a detailed discussion as to the large differences in the toolmark widths that are in consideration.  The Chumbley scoring parameters may have been optimized for chisel widths but may not be optimal for the LEAs.

Another very important point;  in none of the cited texts should there be an impression that if a ground truth match is not indicated by sufficient match score, that it falls into a miss-identification, or Type 2 error "a failed test".  A type 2 error (false positive) would occur if a comparison has a match score over a threshold for identification when the ground truth is two sources.  Therefore, the hot-button term "error rates" in this case is not accurate.
The manuscript is using this erroneous description throughout, is other words, "Missed identifications" and are improperly indicated as described as "miss-identifications".  This miss-characterization is a major weakness in the manuscript.

Coarseness of the surface is not a direct indicator of sub-class influence.  More literature review is necessary to understand the earmarks of subclass.  For the surfaces compared in the Hamby sets, there was no indication of subclass influence between fired bullet sets from consecutively rifled barrels.  If the term is instead a parameter of form (curvature) then that should be more fully described, and how it is used. (I suspect that this is how it was used for fired bullets in this research.)

Additionally, consider including a firearm examiner who is acquainted with objective methods of comparisons for 3D toolmark profiles to review your updated manuscript draft.

### 2018-08-07 doc/paper.tex
adding response

--- 2018-06-19 doc/paper.tex
@@
 {
-  \title{\bf Adaption of the Chumbley Score to matching of bullet striation marks}
+  \title{\bf Adapting the Chumbley Score to match striae on Land Engraved Areas of
+bullets}
 
   \author{
-        Ganesh Krishnan \thanks{The authors gratefully acknowledge funding from CSAFE through
-cooperative agreement \# 70NANB15H176 between NIST and Iowa State
-University.} \\
-    Department of Statistics, Iowa State University\\
-     and \\     Heike Hofmann \\
-    Department of Statistics and CSAFE, Iowa State University\\
+        Ganesh Krishnan,\\ %\\
+    Master of Science (M.Sc.)\\
+    % \\
+     and \\     Prof.~Heike Hofmann,\\ %\\
+    Doctor of Philosophy (Ph.D.)\\
+    % \\
       }
     \maketitle
+  \bigskip
+  \begin{center}
+  {
+  \textbf{Affiliation(s)}\\
+  Department of Statistics, Iowa State University and CSAFE (Center for
+  Statistics and Applications in Forensic Evidence)\\
+  \bigskip
+  \textbf{Source of funding}\\
+  The authors gratefully acknowledge funding from CSAFE through
+  cooperative agreement \# 70NANB15H176 between NIST and Iowa State
+  University.\\
+  \bigskip
+  \textbf{Acknowledgement}\\
+  We would like to thank the efforts of CSAFE and Roy J Carver Microscopy
+  lab in scanning the Hamby-44 set and providing the scans to us. We would
+  also like to thank NIST for providing the Hamby-44 bullets to us for
+  scanning.\\
+  \bigskip
+  \textbf{Information on presentation of work}\\
+  This work was presented in the form of a poster in the AFTE 2018 49th
+  Annual Training Seminar, held from June 3rd to June 8th, 2018 in
+  Charleston, West Virginia
+  }
+  \end{center}
+
 } \fi
 
-\if1\blind
+  % \bigskip
+
+  % \bigskip
+\newpage
+
+\if0\blind
 {
@@
   \begin{center}
-    {\LARGE\bf Adaption of the Chumbley Score to matching of bullet striation marks}
+  \title{\LARGE\bf Adapting the Chumbley Score to match striae on Land Engraved Areas of
+bullets}
   \end{center}
-  \medskip
+  \bigskip
+  \maketitle
 } \fi
@@
 and the suggested modified algorithm as \textbf{(CS2)} and compare their
-performance on the available scans of the Hamby study.
+performance on the available scans of the Hamby study. The section
+\ref{appendix:appxfailed} of the Appendix discusses about the various
+scenarious of failed tests.
 

### 2018-08-07 Response.Rmd
mini changes

@@
 
-The article ???Adapting the Chumbley Score to match striae on Land Engraved Areass (LEAs) of bullets???
-extends an existing comparison algorithm for striated marks to the analysis of bullet LEAs.
+The article "Adapting the Chumbley Score to match striae on Land Engraved Areass (LEAs) of bullets" extends an existing comparison algorithm for striated marks to the analysis of bullet LEAs.
 The paper makes two primary contributions.
 First, they modify the deterministic work of Hadler & Morris (2017) which itself is an extension of Chumbley et al.
-The authors???
-contribution is a modification of Hadler & Morris to reduce the number of scenarios which ???fail???
-under the Hadler & Morris approach.
+The authors' contribution is a modification of Hadler & Morris to reduce the number of scenarios which 'fail' under the Hadler & Morris approach.
 Second, they evaluate the work on two sets of bullet LEA data and explore some variants of the algorithm parameters to tune it for bullets rather than screwdrivers.
@@
 Table 1 is good as-is.
-Table 2 can be a bit misleading because we don???t know what bullet datasets were used.
+Table 2 can be a bit misleading because we don't know what bullet datasets were used.
 Because these different papers analyzed different bullet data sets the underlying difficulty of matching impacts the performance.
-That is, bullets from very well marking firearms are ???easier???
-to match than those from firearms which mark poorly.
+That is, bullets from very well marking firearms are 'easier' to match than those from firearms which mark poorly.
 So the FP and FN reported should be taken with a grain of salt.
@@
 
-I???m also confused how the FP rate for Hare CNMS is 6.26% whereas all other FP are 6.25% This seems to likely be a typo.
-6.25% imply close to 10,000 comparisons which I don???t think the datasets have.
+I'm also confused how the FP rate for Hare CNMS is 6.26% whereas all other FP are 6.25% This seems to likely be a typo.
+6.25% imply close to 10,000 comparisons which I don't think the datasets have.
 I could be wrong but I wanted to point out this potential typo that should be checked.
 
-The authors describe the two methods of ???failure???
-for the method (Sec 2.3).
-I agree with the authors statement that this failure ???should???
-only happen for non-matches.
+The authors describe the two methods of 'failure' for the method (Sec 2.3).
+I agree with the authors statement that this failure 'should' only happen for non-matches.
 It could also happen for matches if the optimal profiles are incorrect.
 That is, if there are two possible optimal match pairs (with very similar correlation scores), where the top scoring pair is incorrect but the second best scoring pair is correct.
-I???m not sure how often that happens.
+I'm not sure how often that happens.
 Overall, I would think that failing for non-matches would be ok.
-It???s not clear that the described failures are the biggest cause of performance hits but improving the algorithm to handle these failures seems reasonable.
+It's not clear that the described failures are the biggest cause of performance hits but improving the algorithm to handle these failures seems reasonable.
 
@@
 Some examples of clarification / grammar / preciseness issues:
-- p3 ???confocal microscopy allow * to measure 3D surfaces in a high-resolution digital form???  * should be ???one????
-- p3 ???Digitized images of 3D surfaces of form the data basis* of statistical analysis??? Awkward phrasing.
-- p3 ???Faden et al and Chumbley et al have been analyzing screwdriver marks??? perhaps change to ???have analyzed??? because the work is completed and published.
+
+- p3 "confocal microscopy allow * to measure 3D surfaces in a high-resolution digital form"  * should be "one"?
+- p3 "Digitized images of 3D surfaces of form the data basis* of statistical analysis" Awkward phrasing.
+- p3 "Faden et al and Chumbley et al have been analyzing screwdriver marks" perhaps change to "have analyzed" because the work is completed and published.
 - p9 Seems to be a broken latex figure reference around line 28/29.
@@
 Minor edits should be made to accuracy and clarity:
-- p3 ???Methods for matching marks for a variety of tools???
-should be changed to ???Methods for matching striated marks for a variety of tools???.
+
+- p3 "Methods for matching marks for a variety of tools" should be changed to "Methods for matching striated marks for a variety of tools".
 This is an important point in that it helps make clear that the authors are studying and comparing striated toolmarks and not impressed toolmarks.
 I would also propose that the authors make this distinction earlier in the introduction.
-- p3 ???Analysis of these digitized markings require the use of statistical methods???.
+- p3 "Analysis of these digitized markings require the use of statistical methods".
 Analysis does not *require* statistical methods.
@@
 This being said, statistical methods are of course great and we should be working towards them.
-Authors could change to analysis ???benefits from the use of statistical methods???
-or something similar.
+Authors could change to analysis "benefits from the use of statistical methods" or something similar.
 - p14 how are the false negative and false positive rates computed?
@@
 Major Changes Required before acceptance:
-- p8 ???This severe limitation in the amount of available data poses the main challenge in???.
-It???s unclear what ???this???
-refers to.
+
+- p8 "This severe limitation in the amount of available data poses the main challenge in".
+It's unclear what 'this' refers to.
 Do the authors mean the number of bullets and LEAs in the dataset or do they mean the number of sample points per striation profile (due to the fact that LEAs are shorter than screwdriver profiles)?
-Both would seem to be limitations but just because the profile is shorter doesn???t inherently mean that there are fewer or less informative striaes.
+Both would seem to be limitations but just because the profile is shorter doesn't inherently mean that there are fewer or less informative striaes.
 You could have a shorter profile with more informative features.
 - Eq 1 uses a variable i, which is defined above as having the values of 1 or 2.
-I therefore don???t understand how i<0 as written in Eq 1.
+I therefore don't understand how i<0 as written in Eq 1.
 This should be improved.
@@
 How is the original 3D surface resampled to the 1d profile?
-This is very important because if you ???cut???
-the 3D profile the wrong way (at the wrong angle) then you will dilate or compress the extracted profile.
+This is very important because if you 'cut' the 3D profile the wrong way (at the wrong angle) then you will dilate or compress the extracted profile.
 This needs a lot more description.
 - Section 2.4 is not clearly written.
-I???m not confident that I understand their new method.
+I'm not confident that I understand their new method.
 This section should be rewritten to explain what the authors are doing.
@@
 Note that my description is not clear either as this is a tricky thing to put into words, but is something the authors need to do.
-- The results presented on page 14 list the correct identification of ???same-source and difference-source toolmarks of 93.5 to 94.1%???.
+- The results presented on page 14 list the correct identification of "same-source and difference-source toolmarks of 93.5 to 94.1%".
 Does this mean the true positives are 93.5% and the true negatives are 94.1%?
-I don???t think so because the next sentence lists the false negative rate of between 0.28 and 0.36.
+I don't think so because the next sentence lists the false negative rate of between 0.28 and 0.36.
 The authors need to explicitly list the true positive rate and the true negative rate.
 These must be separate because there are many more true negatives than true positives in any matching dataset.
-That means that an algorithm that always predicts ???different source???
-would be correct most of the time.
-In fact, for the author???s dataset the ???always predict different???
-algorithm would be correct approximately 94% of the time (35 bullets where each has appx 2 matches).
+That means that an algorithm that always predicts "different source" would be correct most of the time.
+In fact, for the author's dataset the "always predict different" algorithm would be correct approximately 94% of the time (35 bullets where each has appx 2 matches).
 The lack of a true positive rate and true negative rate is is a major point that must be corrected.
@@
 - The results should all be listed as either fractions or all as percent (which is preferred) and not mix them.
-- p21 the authors state ???this kind of assessment is only feasible in the setting of a large study, such as the one we presented???.
-I strongly disagree with the use of the word ???large???.
+- p21 the authors state "this kind of assessment is only feasible in the setting of a large study, such as the one we presented".
+I strongly disagree with the use of the word "large".
 The presented work is not a large study.
-Ten firearms were used in each of two studies (it???s not clear if both datasets used the same 10 firearms).
+Ten firearms were used in each of two studies (it's not clear if both datasets used the same 10 firearms).
 A large study would include at least 100 different firearms and even 100 is not likely to provide statistical confidence.
@@
 
-Comments to the Author
 You may not fully appreciate and how critical their phase orientations used to identify fired bullets (LEAs).

### 2018-08-07 doc/paper.Rmd
can't get the references to work, but the style file is closer to JFS specs

@@
 ---
-title: "Adaption of the Chumbley Score to matching of bullet striation marks"
+title: "Adapting the Chumbley Score to match striae on Land Engraved Areas of bullets"
 authors:
-- affiliation: Department of Statistics, Iowa State University 
-  name: Ganesh Krishnan
-  thanks: The authors gratefully acknowledge funding from CSAFE through cooperative agreement \# 70NANB15H176 between NIST and Iowa State University. 
-- affiliation: Department of Statistics and CSAFE, Iowa State University 
-  name: Heike Hofmann
-biblio-style: apsr
+# - affiliation: Department of Statistics, Iowa State University 
+- name: Ganesh Krishnan,
+  degree: Master of Science (M.Sc.)
+  # thank: The authors gratefully acknowledge funding from CSAFE through cooperative agreement \# 70NANB15H176 between NIST and Iowa State University.
+We would also like to thank the efforts of CSAFE and Roy J carver Microscopy lab in scanning the Hamby-44 set and providing the scans to us.
+# - affiliation: Department of Statistics and CSAFE, Iowa State University 
+- name: Prof. Heike Hofmann,
+  degree: Doctor of Philosophy (Ph.D.)
+affiliation: Department of Statistics, Iowa State University and CSAFE (Center for Statistics and Applications in Forensic Evidence)
+funding: The authors gratefully acknowledge funding from CSAFE through cooperative agreement \# 70NANB15H176 between NIST and Iowa State University. 
+thank: We would like to thank the efforts of CSAFE and Roy J Carver Microscopy lab in scanning the Hamby-44 set and providing the scans to us.
+We would also like to thank NIST for providing the Hamby-44 bullets to us for scanning.
+presentedat:  This work was presented in the form of a poster in the AFTE 2018 49th Annual Training Seminar, held from June 3rd to June 8th, 2018 in Charleston, West Virginia
+biblio-style: vancouver
 output:

### 2018-08-07 doc/paper.Rmd
major win on the citation front

@@
 
-Methods for matching marks for a variety of tools have been studied in the literature (see \autoref{tab:toolmarks-ER} for an overview):  \citet{manytoolmarks1} and \citet{chumbley} have been analyzing screwdriver marks  digitized using a profilometer;  \citet{manytoolmarks2} have investigated 3D marks from screwdriver, tongue and groove pliers captured using a confocal microscope; \citet{afte-chumbley} have been investigated digitized marks from slip-joint pliers generated by a surface profilometer.
+\hh{changed citation style}
+Methods for matching marks for a variety of tools have been studied in the literature (see \autoref{tab:toolmarks-ER} for an overview):  in \cite{manytoolmarks1} and \cite{chumbley} digitized screwdriver marks have been analyzed using a profilometer;  in \cite{manytoolmarks2} 3D marks from screwdriver, tongue and groove pliers captured using a confocal microscope have been investigated; digitized marks from slip-joint pliers generated by a surface profilometer  have been investigated in \cite{afte-chumbley}.
 

### 2018-08-07 doc/paper.tex
can't get the references to work, but the style file is closer to JFS specs
major win on the citation front

@@
 \usepackage{natbib}
+\setcitestyle{numbers}
+
 \usepackage{booktabs}
@@
 
+{\textcolor{orange}{changed citation style}}
+
 Methods for matching marks for a variety of tools have been studied in
-the literature (see \autoref{tab:toolmarks-ER} for an overview):
-\citet{manytoolmarks1} and \citet{chumbley} have been analyzing
-screwdriver marks digitized using a profilometer; \citet{manytoolmarks2}
-have investigated 3D marks from screwdriver, tongue and groove pliers
-captured using a confocal microscope; \citet{afte-chumbley} have been
-investigated digitized marks from slip-joint pliers generated by a
-surface profilometer.
+the literature (see \autoref{tab:toolmarks-ER} for an overview): in
+\cite{manytoolmarks1} and \cite{chumbley} digitized screwdriver marks
+have been analyzed using a profilometer; in \cite{manytoolmarks2} 3D
+marks from screwdriver, tongue and groove pliers captured using a
+confocal microscope have been investigated; digitized marks from
+slip-joint pliers generated by a surface profilometer have been
+investigated in \cite{afte-chumbley}.
 
@@
 
-\bibliographystyle{agsm}
+\bibliographystyle{vancouver}
 \bibliography{bibliography}

### 2018-08-07 doc/paper.Rmd
fixed reference

@@
 Technological advances in surface imaging allow measurements of 3D surfaces at previously unforeseen resolutions and enable digitized imaging.
-Here, we investigate the applicability of the Chumbley scoring method \citep{hadler}, developed for screwdriver markings, for same-source identification of bullet striae.
+Here, we investigate the applicability of the Chumbley scoring method \cite{hadler}, developed for screwdriver markings, for same-source identification of bullet striae.
 We provide methods to identify parameters that minimize the error rates for matching of LEAs using Hamby datasets 44 and 252 measured by NIST and CSAFE.
@@
 Same-source analyses are a major part of a Forensic Toolmark Examiner's job.
-In current practice, examiners  make these comparisons by visual inspection under a comparison microscope and come to one of the following four conclusions: identification, inconclusive, elimination or unsuitable for examination \citep{afte-toolmarks1998}.
-These conclusions are made on the basis of "unique surface contours" of the two toolmarks being in "sufficient agreement" \citep{afte-toolmarks1998}.
-AFTE describes the term "sufficient agreement" as the possibility of another tool producing the markings under comparison, as practically impossible \citep{afte-toolmarks1998}.
-Potential subject bias in the assessment as well as the lack of specified error rates are the main points of criticisms  first raised by the National Research Council in 2009 \citep{NAS:2009} and later emphasized further by the President's Council of Advisors on Science and Technology \citep{pcast2016}.
+In current practice, examiners  make these comparisons by visual inspection under a comparison microscope and come to one of the following four conclusions: identification, inconclusive, elimination or unsuitable for examination \cite{afte-toolmarks1998}.
+These conclusions are made on the basis of "unique surface contours" of the two toolmarks being in "sufficient agreement" \cite{afte-toolmarks1998}.
+AFTE describes the term "sufficient agreement" as the possibility of another tool producing the markings under comparison, as practically impossible \cite{afte-toolmarks1998}.
+Potential subject bias in the assessment as well as the lack of specified error rates are the main points of criticisms  first raised by the National Research Council in 2009 \cite{NAS:2009} and later emphasized further by the President's Council of Advisors on Science and Technology \cite{pcast2016}.
 
 Technological advances, such as profilometers and confocal microscopy allow to measure 3D surfaces in a high-resolution digitized form.
-This technology has become more accessible over the last decade, and has made its way into topological images of ballistics evidence, such as bullet lands and breech faces \citep{DeKinder1, DeKinder2, Bachrach1, vorburger2016}.
+This technology has become more accessible over the last decade, and has made its way into topological images of ballistics evidence, such as bullet lands and breech faces \cite{DeKinder1, DeKinder2, Bachrach1, vorburger2016}.
 Digitized images of 3D surfaces of form the data basis of statistical analysis of toolmarks.
@@
 Analysis of these digitized markings require the use of statistical methods which can quantify the scientific mechanism of comparing markings and serve as basis for an error rate calculation.
-\citet{manytoolmarks2} define a relative distance metric and use it as similarity measure between two toolmarks.
-\citet{manytoolmarks1} extract many small segments in the markings of two toolmarks and compare similarity using a maximum Pearson correlation coefficient.
-The Chumbley scoring method, first introduced by \citet{chumbley}, uses a similar but more extensive framework based on a Mann-Whitney U test of the resulting correlation coefficients.
+In \cite{manytoolmarks2} a relative distance metric is defined and used as similarity measure between two toolmarks.
+This approach is expanded in \citet{manytoolmarks1}: a set of small segments in the markings of two toolmarks are extracted and  similarity is compared using a maximum Pearson correlation coefficient.
+The Chumbley scoring method, first introduced in \cite{chumbley}, uses a similar but more extensive framework based on a Mann-Whitney U test of the resulting correlation coefficients.
 This approach is non-deterministic, because segments are chosen randomly.
-\citep{hadler} make the score deterministic for each pair of toolmarks by choosing segments for comparison systematically.
+In \cite{hadler}  the score is fixed to be deterministic for each pair of toolmarks by choosing segments for comparison systematically.
 This approach also ensures independence between segments of striae.
@@
   \hline
- \textbf{\citet{manytoolmarks1}} & Maximum Pearson & Screwdrivers & & \\ & Correlation &  & - & - \\ \hline
- \textbf{\citet{chumbley}} & Randomized & Screwdrivers & & \\ (Same-Surface Same-Angle) & Chumbley Score & & 2.3\% & 8.9\%  \\ \hline
-  \textbf{\citet{afte-chumbley}} & Randomized & Slip-joint  &  & \\ & Chumbley Score &  & - & - \\ \hline 
-  \textbf{\citet{hadler}} & Deterministic & Screwdrivers & & \\ (Same-Surface Same-Angle) & Chumbley Score & & 0\% & 6\%  \\ \hline
-  \textbf{\citet{manytoolmarks2}} & Similarity Measure & Screwdrivers & & \\ (Different Surfaces-same angle) & Relative Distance Metric & & 5.9\% &  9.4\% \\ (Same Surfaces-same angle)&  & & 0.22\% &  0\% \\
+ \textbf{Faden et al. (2007)} \cite{manytoolmarks1} & Maximum Pearson & Screwdrivers & & \\ & Correlation &  & - & - \\ \hline
+ \textbf{Chumbley et al. (2010)} \cite{chumbley} & Randomized & Screwdrivers & & \\ (Same-Surface Same-Angle) & Chumbley Score & & 2.3\% & 8.9\%  \\ \hline
+  \textbf{Grieve et al. (2014)}
+ \cite{afte-chumbley} & Randomized & Slip-joint  &  & \\ & Chumbley Score &  & - & - \\ \hline 
+  \textbf{Hadler \& Morris (2017)} \cite{hadler} & Deterministic & Screwdrivers & & \\ (Same-Surface Same-Angle) & Chumbley Score & & 0\% & 6\%  \\ \hline
+  \textbf{Bachrach et al.
+(2010)} \cite{manytoolmarks2} & Similarity Measure & Screwdrivers & & \\ (Different Surfaces-same angle) & Relative Distance Metric & & 5.9\% &  9.4\% \\ (Same Surfaces-same angle)&  & & 0.22\% &  0\% \\
    \hline
@@
 This results in an optimal vertical (in-phase) shift of $t_1^o - t_2^o$ for aligning the two markings.
-We will denote the relative optimal locations as $t_1^*$ and $t_2^*$, where $t_i^* = t_i^o/(T_i-w_o)$ for $i=1,2$, such that $t_1^*, t_2^* \in [0,1]$.
+We will denote the relative optimal locations as $t_1^*$ and $t_2^*$, where $t_k^* = t_k^o/(T_k-w_o)$ for $k=1,2$, such that $t_1^*, t_2^* \in [0,1]$.
 Once (sub-)class characteristics are removed, the relative optimal locations should be distributed according to a uniform distribution in $[0,1]$.
 
-In the validation step,  two sets of windows of size $w_v$ are chosen from both markings,  \autoref[see][]{fig:win-comparison}.
+In the validation step,  two sets of windows of size $w_v$ are chosen from both markings, see \autoref{fig:win-comparison}.
 In the first set, pairs of windows are extracted from the two markings using the optimal vertical shift as determined in the first step, whereas for the second set the windows are extracted using a different (out-of-phase) shift.

### 2018-08-07 doc/paper.tex
fixed reference

@@
 enable digitized imaging. Here, we investigate the applicability of the
-Chumbley scoring method \citep{hadler}, developed for screwdriver
+Chumbley scoring method \cite{hadler}, developed for screwdriver
 markings, for same-source identification of bullet striae. We provide
@@
 following four conclusions: identification, inconclusive, elimination or
-unsuitable for examination \citep{afte-toolmarks1998}. These conclusions
+unsuitable for examination \cite{afte-toolmarks1998}. These conclusions
 are made on the basis of ``unique surface contours'' of the two
-toolmarks being in ``sufficient agreement'' \citep{afte-toolmarks1998}.
+toolmarks being in ``sufficient agreement'' \cite{afte-toolmarks1998}.
 AFTE describes the term ``sufficient agreement'' as the possibility of
 another tool producing the markings under comparison, as practically
-impossible \citep{afte-toolmarks1998}. Potential subject bias in the
+impossible \cite{afte-toolmarks1998}. Potential subject bias in the
 assessment as well as the lack of specified error rates are the main
 points of criticisms first raised by the National Research Council in
-2009 \citep{NAS:2009} and later emphasized further by the President's
-Council of Advisors on Science and Technology \citep{pcast2016}.
+2009 \cite{NAS:2009} and later emphasized further by the President's
+Council of Advisors on Science and Technology \cite{pcast2016}.
 
@@
 lands and breech faces
-\citep{DeKinder1, DeKinder2, Bachrach1, vorburger2016}. Digitized images
+\cite{DeKinder1, DeKinder2, Bachrach1, vorburger2016}. Digitized images
 of 3D surfaces of form the data basis of statistical analysis of
@@
 methods which can quantify the scientific mechanism of comparing
-markings and serve as basis for an error rate calculation.
-\citet{manytoolmarks2} define a relative distance metric and use it as
-similarity measure between two toolmarks. \citet{manytoolmarks1} extract
-many small segments in the markings of two toolmarks and compare
-similarity using a maximum Pearson correlation coefficient. The Chumbley
-scoring method, first introduced by \citet{chumbley}, uses a similar but
-more extensive framework based on a Mann-Whitney U test of the resulting
-correlation coefficients. This approach is non-deterministic, because
-segments are chosen randomly. \citep{hadler} make the score
-deterministic for each pair of toolmarks by choosing segments for
-comparison systematically. This approach also ensures independence
-between segments of striae.
+markings and serve as basis for an error rate calculation. In
+\cite{manytoolmarks2} a relative distance metric is defined and used as
+similarity measure between two toolmarks. This approach is expanded in
+\citet{manytoolmarks1}: a set of small segments in the markings of two
+toolmarks are extracted and similarity is compared using a maximum
+Pearson correlation coefficient. The Chumbley scoring method, first
+introduced in \cite{chumbley}, uses a similar but more extensive
+framework based on a Mann-Whitney U test of the resulting correlation
+coefficients. This approach is non-deterministic, because segments are
+chosen randomly. In \cite{hadler} the score is fixed to be deterministic
+for each pair of toolmarks by choosing segments for comparison
+systematically. This approach also ensures independence between segments
+of striae.
 
@@
 relative optimal locations as \(t_1^*\) and \(t_2^*\), where
-\(t_i^* = t_i^o/(T_i-w_o)\) for \(i=1,2\), such that
+\(t_k^* = t_k^o/(T_k-w_o)\) for \(k=1,2\), such that
 \(t_1^*, t_2^* \in [0,1]\). Once (sub-)class characteristics are
@@
 In the validation step, two sets of windows of size \(w_v\) are chosen
-from both markings, \autoref[see][]{fig:win-comparison}. In the first
-set, pairs of windows are extracted from the two markings using the
-optimal vertical shift as determined in the first step, whereas for the
-second set the windows are extracted using a different (out-of-phase)
-shift.
+from both markings, see \autoref{fig:win-comparison}. In the first set,
+pairs of windows are extracted from the two markings using the optimal
+vertical shift as determined in the first step, whereas for the second
+set the windows are extracted using a different (out-of-phase) shift.
 

### 2018-08-07 doc/paper.Rmd
reference adher(?) to style guide

@@
 presentedat:  This work was presented in the form of a poster in the AFTE 2018 49th Annual Training Seminar, held from June 3rd to June 8th, 2018 in Charleston, West Virginia
-biblio-style: vancouver
+biblio-style: vancouver-mod
 output:

### 2018-08-07 doc/paper.Rmd
update reference to aoas paper

--- 2018-08-07 doc/paper.Rmd
@@
 presentedat:  This work was presented in the form of a poster in the AFTE 2018 49th Annual Training Seminar, held from June 3rd to June 8th, 2018 in Charleston, West Virginia
-biblio-style: vancouver
+biblio-style: jfs
 output:

### 2018-08-10 doc/paper.Rmd
new style file added to account for author, old style file kept for now
Fixed heading bold-italic issues

@@
 
-\tableofcontents
+<!-- \tableofcontents -->
 \newpage
@@
 
-# Introduction
-
+<!-- # Introduction -->
+
+<!-- \citeauthor{manytoolmarks2} -->
+<!-- \citet{manytoolmarks2} -->
 Same-source analyses are a major part of a Forensic Toolmark Examiner's job.
@@
 
-# Methods
-
-## Scans for land engraved areas
+# Methods {-}
+
+## *Scans for land engraved areas* {-}
 
@@
 
-## The Chumbley Score Test
+## *The Chumbley Score Test* {-}
 
@@
 
-## A problem with failed tests
+## *A problem with failed tests* {-}
 
@@
 
-## A modified approach
+## *A modified approach* {-}
 
@@
 
-## Testing setup
-
-### The Data
+## *Testing setup* {-}
+
+### The Data {-}
 
@@
 
-### Setup
+### Setup {-}
 Both algorithms (CS1) and (CS2) are implemented in R \citep{R}.
@@
 
-# Results
-## Failed Tests
+# Results {-}
+## *Failed Tests* {-}
 
@@
 
-## Coarseness
+## *Coarseness* {-}
 The purpose of the coarseness parameter is to remove (sub-)class characteristics from profiles before comparisons for matching.
@@
 
-## Error rate assessment
+## *Error rate assessment* {-}
 
@@
 
-## Observed versus Nominal Type I error rates
+## *Observed versus Nominal Type I error rates* {-}
 
@@
 
-## High resolution Hamby 44 scans
+## *High resolution Hamby 44 scans* {-}
 
@@
 
-# Conclusions
+# Conclusions {-}
 
@@
 
-# Appendix
+# Appendix {-}
 \begin{appendix}

### 2018-08-13 doc/Response.Rmd
adding modified bst file
Reviews 1 comments responses smthm

--- 2018-08-07 Response.Rmd
@@
 output: word_document
+# output: word_document
 ---
@@
 ```
+
+\newcommand{\hh}[1]{{\textcolor{orange}{#1}}}
+\newcommand{\gk}[1]{{\textcolor{blue}{#1}}}
 
@@
 
+*Since different methods are being employed in literature and from what is being done in this paper (to the best of our knowledge, the method in this paper has never been used on bullet striation marks before), it is difficult to assess and quantify performance purely on the basis of different datasets.
+The method being employed to make the comparisons would seem to also have a say in the performance and FP and FN rates being reported in literature*
+
 I'm also confused how the FP rate for Hare CNMS is 6.26% whereas all other FP are 6.25% This seems to likely be a typo.
@@
 I could be wrong but I wanted to point out this potential typo that should be checked.
+
+*The data set being used by Hare et al.
+has about 10384 land-to-land comparisons and the error rates being reported here are computed from Receiver operating characteristic (ROC) curves.
+This means that the error rates for a particular chosen FP rate will have a corresponding FN rate.
+The FP rates reported, therefore could have been different, and any difference in the rate would mean a corresponding change in the FN rate that is governed and controlled by the ROC curve.*
 
@@
 It's not clear that the described failures are the biggest cause of performance hits but improving the algorithm to handle these failures seems reasonable.
+
+*One reason for not looking at this is because any slight change in the Maximum Correlation window (second best window) locations would still lie within the vicinity of the top priority window, which would change the lag but not with any drastic difference.
+Now for the case of failed tests where the Optimal locations on the striation marks sometimes get positioned at the ends, the maximum correlation was high because of almost completely smooth (lacking sub-class characteristics) ends.
+Any other second or third best correlation windows computed would be positioned with a wiggle room of just a few pixels (say 1 or 2) pixels.
+This makes it increasingly difficult to identify a threshold in the heirarchy of the maximal correlation optimal position calculations that would be able satisfy and correctly identify the lag.*
+
+*For the performance hit the Failed tests seem to be a very big cause because 20 to 30% of the comparison do not yield a result at all, it means that for almost a quarter of the dataset we cannot make any decision.
+This would be a huge problem in a real world setting.*
 
@@
 You could have a shorter profile with more informative features.
+
+*We agree that shorter profile might have more information features, but with smaller profiles a certain level of distinctiveness is absent in the information that the chumbley algorithm in its original (chumbley et al and hadler et al) form is not able to pickup.
+Ability to distinguish such informative yet smaller sub-class characterisitic and therefore find relevant highly correlated windows becomes increasingly hard for the algoritm.The longer and wider toolmarks have more pronounced markings which the original chumbley and hadler algorithms are able to pick up in their different steps.*
+
 - Eq 1 uses a variable i, which is defined above as having the values of 1 or 2.
@@
 Possibly this is a sloppy reuse of variables?
+
+*thank you for pointing this out - we accidentally changed notations. this is fixed now. Index k=1,2 refers to the two different signatures, index i in Z refers to window locations.*
+
 - How are the original profiles (eg.
@@
 This needs a lot more description.
+
+*The methods of extracting the profiles are mentioned and described in detail in the Hare et al paper. We have used the same extracted profiles as used in the Hare et al. paper.*
+
 - Section 2.4 is not clearly written.
@@
 Note that my description is not clear either as this is a tricky thing to put into words, but is something the authors need to do.
+
+*The alternating choice of windows with respect to the location of the Optimization windows is a key concept of the Hadler and Morris Paper.
+The choice of integer windows comes because we are dealing with pixels and classify lengths in pixels.
+It primarily ensures repeating or overlapping windows are avoided.
+This means that in order to avoid overlapping windows the striations will have a certain integer number of windows which depends on the length of the marking in pixels and the size of the window in pixels.Overlapping windows are to be avoided because as described in the hadler and morris algorithm it is used to solve the problem of lack of independence which was a key incremental contribution of their paper over the original chumbley algorithm*
+
+*For the different shift step, firstly we aim to compare (as in the Hadler and Morris algorithm) windows that have a very small chance of having higher correlations.
+When starting from Right to left on one striae and going from left to right on the 2nd striae, we end up finding correlations between windows that are the middle too.
+That creates a problem as for same-source markings we will have higher correlations between such windows.*
+
+*Secondly, since the Optimization window locations has determined the possible lag in the two signatures, by going in opposite direction of the Optimization window we are able to certify that the windows being compared to calculate different-dhift correlations, will have higher correlation only by fluke or chance of nature, and not because of the two striae being from the same or different sources.
+*
+
+*Thirdly, the aim of this step is to calculate the window correlations that are the worst.
+These set of correlations can then serve as a base line for comparison with the set of higher correlations from the same-shift step.
+This enables to designate the two striae as matching (same-source) or non-matching (different source)*
+
+*Fourthly, if as described, we skip the same-shift windows it means we skip the Optimization window pairs too, as the Optimization window pairs are also a same-shift or lag infused pair of windows albeit of a larger size.
+This means comparing from left to right on one marking with right to left on the other one by skipping the optimization window.
+This is something similar to what we are doing right now.*
+
+*One the same note, we would like to add that, skipping the same-shift windows is not required, that is, if by skipping of windows, the windows being "used" in the same shift step is being meant.
+The windows designated in the validation i.e same shift and different shift substeps remain the same,  as they are non overlapping segements which have been designated.
+The way in which they pair up is what has changed.
+This means the pair of same-shift windows is what that needs to be avoided from being compared.
+And this is the issue being addressed in the current setup.*
+
 - The results presented on page 14 list the correct identification of "same-source and difference-source toolmarks of 93.5 to 94.1%".
@@
 To be clear, the true positive rate, true negative rate, false positive rate, and false negative rate should all be reported for the first dataset, perhaps added to Table 3 or another table like Table 3.
+
+*The 93.5% to 94.1% are the True Negative rates which are being mentioned.
+It is computed as 1 - Observed FPR.
+The observed False positive rates can be seen in figure 10.
+The True positive rates would then be 1 - Observed FNR or about 64% to 72%.*
+
 - The true positive rate, true negative rate, false positive rate, and false negative rate should also be presented for the second dataset (Hamby 44 scans, Sec 3.5).

### 2018-08-13 doc/Response.Rmd
reviewer 2 comments answers

@@
 
-*Secondly, since the Optimization window locations has determined the possible lag in the two signatures, by going in opposite direction of the Optimization window we are able to certify that the windows being compared to calculate different-dhift correlations, will have higher correlation only by fluke or chance of nature, and not because of the two striae being from the same or different sources.
-*
+*Secondly, since the Optimization window locations has determined the possible lag in the two signatures, by going in opposite direction of the Optimization window we are able to certify that the windows being compared to calculate different-dhift correlations, will have higher correlation only by fluke or chance of nature, and not because of the two striae being from the same or different sources.*
 
@@
 
+*Thank you for your comments, we appreciate your insight into this.
+This paper is a first measure in the adaption process of toolmark based methods to Bullets, and we therefore plan on making a comparison of this sort by including phase (as described by \cite{chu2010}) at a later stage by making such a comparison.
+In this paper we wanted to focus on the feasibility and possibility of making a successful adaption of the Chumbley method to bullet matching*
+
+*We agree on the point that the toolmarks have large differences in width and the parameters of the chumbley algorithm are optimized for tools, which is why we first do a parameter study to identify the best set of parameters that would work for the LEAs.
+Before identifying problems in doing so and trying to address those problems*
+
 Another very important point;  in none of the cited texts should there be an impression that if a ground truth match is not indicated by sufficient match score, that it falls into a miss-identification, or Type 2 error "a failed test".
@@
 This miss-characterization is a major weakness in the manuscript.
+
+*By failed test, what is being meant is that the method fails to conduct a statistical test because of incorrect identification of Maximum correlations windows required for comparison of the two striation marks.
+This means that there is no resolution as to the two striations being a match or a non-match.
+Which makes it impossible to classify it as either Type 1 (False Positive) or Type 2 (False Negative) rate.*
+
+*Thank you for pointing this out. For the terminology error rates, we do mean FPR = FP/ (TN + FP) and FNR = FN/ (TP + FN), which holds with the definition of error rates in literature.*
+
+*Thank you for pointing out the confusing terminology as seen in the "Missed identifications" Page 17. missed same-source marking. *
 
@@
 
+*By coarseness we mean the parameter used to identify smoothing span as used in the LOWESS filter. *
+
 Additionally, consider including a firearm examiner who is acquainted with objective methods of comparisons for 3D toolmark profiles to review your updated manuscript draft.
+
+*We agree that having a Forensic examiner review the draft would bring value to the validation of the method.
+We would very politely like to bring to light that the work in this paper is incremental to the work done by \citet{hadler} and \citet{chumbley}.
+The objectivity of the methods were established in these works, and we only change the pairing of certain windows in the different-shift substep which does not change the format of the algorithm to compromise objectivity in the analysis*

### 2018-08-13 doc/Response.Rmd
small change

@@
 
-*By failed test, what is being meant is that the method fails to conduct a statistical test because of incorrect identification of Maximum correlations windows required for comparison of the two striation marks.
-This means that there is no resolution as to the two striations being a match or a non-match.
-Which makes it impossible to classify it as either Type 1 (False Positive) or Type 2 (False Negative) rate.*
+*By failed test, what is being meant is that the method fails to conduct a statistical test although there is a resolution available in the ground truth as whether the lands match or dont match.
+This is because of incorrect identification of Maximum correlations windows required for comparison of the two striation marks.
+This means that there is no resolution as to the two striations being a match or a non-match, when there should be one *
 
@@
 
-*By coarseness we mean the parameter used to identify smoothing span as used in the LOWESS filter. *
+*By coarseness we mean the parameter used to identify smoothing span as used in the LOWESS filter.*
 

### 2018-08-17 doc/Response.Rmd
some info on missed classification
inconclusive line
rmds rev1 and paper

@@
 
-*Since different methods are being employed in literature and from what is being done in this paper (to the best of our knowledge, the method in this paper has never been used on bullet striation marks before), it is difficult to assess and quantify performance purely on the basis of different datasets.
-The method being employed to make the comparisons would seem to also have a say in the performance and FP and FN rates being reported in literature*
+*Thank you for pointing this out.
+You are right, we have added more information about the data, wherever available.
+We are also citing xxx to emphasize the difficulty level and resulting differences in error rates of different brands of firearms and ammunition*
 
@@
 
-*The data set being used by Hare et al.
+*Thank you for spotting the typo.
+With respect to the sample size, the data set being used by \cite{aoas}.
 has about 10384 land-to-land comparisons and the error rates being reported here are computed from Receiver operating characteristic (ROC) curves.
-This means that the error rates for a particular chosen FP rate will have a corresponding FN rate.
-The FP rates reported, therefore could have been different, and any difference in the rate would mean a corresponding change in the FN rate that is governed and controlled by the ROC curve.*
+We decided to report False Negative error rates for a fixed value of False Positives (FP) = 6.25% to make them comparable to the observed values in Figure 10 (bottom right)*
 
@@
 
-*One reason for not looking at this is because any slight change in the Maximum Correlation window (second best window) locations would still lie within the vicinity of the top priority window, which would change the lag but not with any drastic difference.
-Now for the case of failed tests where the Optimal locations on the striation marks sometimes get positioned at the ends, the maximum correlation was high because of almost completely smooth (lacking sub-class characteristics) ends.
-Any other second or third best correlation windows computed would be positioned with a wiggle room of just a few pixels (say 1 or 2) pixels.
-This makes it increasingly difficult to identify a threshold in the heirarchy of the maximal correlation optimal position calculations that would be able satisfy and correctly identify the lag.*
-
-*For the performance hit the Failed tests seem to be a very big cause because 20 to 30% of the comparison do not yield a result at all, it means that for almost a quarter of the dataset we cannot make any decision.
-This would be a huge problem in a real world setting.*
+*You are right, a failure might occur for a match if the initial optimization fails.
+The new method is still susceptible to that.
+It would be interesting to further investigate thus problem in the furture.
+We have made a note of that in the conclusions*
 
@@
 
-*We agree that shorter profile might have more information features, but with smaller profiles a certain level of distinctiveness is absent in the information that the chumbley algorithm in its original (chumbley et al and hadler et al) form is not able to pickup.
-Ability to distinguish such informative yet smaller sub-class characterisitic and therefore find relevant highly correlated windows becomes increasingly hard for the algoritm.The longer and wider toolmarks have more pronounced markings which the original chumbley and hadler algorithms are able to pick up in their different steps.*
+*We were indeed referring to the shorter length of the marks, which given the same amount of information decreases the power of the test, a purely technical consequence.
+Obviously you are right that this power calculation changes with a change in the amount of information*
 
@@
 
-*The methods of extracting the profiles are mentioned and described in detail in the Hare et al paper. We have used the same extracted profiles as used in the Hare et al. paper.*
+*We followed the \cite{aoas} paper in extracting the profiles from scans. We have expanded on the discussion at the begining of section titled "Scans for land engraved areas" in paper*
 
@@
 
-*The alternating choice of windows with respect to the location of the Optimization windows is a key concept of the Hadler and Morris Paper.
-The choice of integer windows comes because we are dealing with pixels and classify lengths in pixels.
-It primarily ensures repeating or overlapping windows are avoided.
-This means that in order to avoid overlapping windows the striations will have a certain integer number of windows which depends on the length of the marking in pixels and the size of the window in pixels.Overlapping windows are to be avoided because as described in the hadler and morris algorithm it is used to solve the problem of lack of independence which was a key incremental contribution of their paper over the original chumbley algorithm*
-
-*For the different shift step, firstly we aim to compare (as in the Hadler and Morris algorithm) windows that have a very small chance of having higher correlations.
-When starting from Right to left on one striae and going from left to right on the 2nd striae, we end up finding correlations between windows that are the middle too.
-That creates a problem as for same-source markings we will have higher correlations between such windows.*
-
-*Secondly, since the Optimization window locations has determined the possible lag in the two signatures, by going in opposite direction of the Optimization window we are able to certify that the windows being compared to calculate different-dhift correlations, will have higher correlation only by fluke or chance of nature, and not because of the two striae being from the same or different sources.*
-
-*Thirdly, the aim of this step is to calculate the window correlations that are the worst.
-These set of correlations can then serve as a base line for comparison with the set of higher correlations from the same-shift step.
-This enables to designate the two striae as matching (same-source) or non-matching (different source)*
-
-*Fourthly, if as described, we skip the same-shift windows it means we skip the Optimization window pairs too, as the Optimization window pairs are also a same-shift or lag infused pair of windows albeit of a larger size.
-This means comparing from left to right on one marking with right to left on the other one by skipping the optimization window.
-This is something similar to what we are doing right now.*
-
-*One the same note, we would like to add that, skipping the same-shift windows is not required, that is, if by skipping of windows, the windows being "used" in the same shift step is being meant.
-The windows designated in the validation i.e same shift and different shift substeps remain the same,  as they are non overlapping segements which have been designated.
-The way in which they pair up is what has changed.
-This means the pair of same-shift windows is what that needs to be avoided from being compared.
-And this is the issue being addressed in the current setup.*
+*Thank you for poining this out, you are right in your understanding of the method.
+We have included a figure for this and expanded out discussion, which hopefully makes the it easier to understand the method.*
 
@@
 
-*The 93.5% to 94.1% are the True Negative rates which are being mentioned.
+*Thank you for bringing this to our notice.
+We have included a table of error rates.
+The 93.5% to 94.1% are the True Negative rates which are being mentioned.
 It is computed as 1 - Observed FPR.
@@
 If the paper is allowed to state that the described work is a large study it will set a precedent that will be detrimental to the field.
+
+*We agree the data available to us is by no means large. Our comment was referring to the availability of 35 bullets rater than just 2 0r 3 as might be seen in actual case work*
 
@@
 
+*We are aware of the current standards of classification as having a section as inconclusive for examination.
+But for the data sets we are working on, each and every land has been identified and none of the data being used, cannot be classified as from the same-source and different source*
+
 Coarseness of the surface is not a direct indicator of sub-class influence.

### 2018-08-17 doc/paper.Rmd
rmds rev1 and paper

@@
 Potential subject bias in the assessment as well as the lack of specified error rates are the main points of criticisms  first raised by the National Research Council in 2009 \cite{NAS:2009} and later emphasized further by the President's Council of Advisors on Science and Technology \cite{pcast2016}.
-
-Technological advances, such as profilometers and confocal microscopy allow to measure 3D surfaces in a high-resolution digitized form.
+\gk{need a sentence on striated marks}
+
+Technological advances, such as profilometers and confocal microscopy allow \gk{one} to measure 3D surfaces in a high-resolution digitized form.
 This technology has become more accessible over the last decade, and has made its way into topological images of ballistics evidence, such as bullet lands and breech faces \cite{DeKinder1, DeKinder2, Bachrach1, vorburger2016}.
-Digitized images of 3D surfaces of form the data basis of statistical analysis of toolmarks.
+\gk{small change}Digitized images of 3D surfaces, form the basis of statistical analysis of toolmarks.
 A statistical approach based on data  removes both subjectivity from the assessment and allows a quantification of error rates for both false positive and false negative identifications.
@@
 \hh{changed citation style}
-Methods for matching marks for a variety of tools have been studied in the literature (see \autoref{tab:toolmarks-ER} for an overview):  in \cite{manytoolmarks1} and \cite{chumbley} digitized screwdriver marks have been analyzed using a profilometer;  in \cite{manytoolmarks2} 3D marks from screwdriver, tongue and groove pliers captured using a confocal microscope have been investigated; digitized marks from slip-joint pliers generated by a surface profilometer  have been investigated in \cite{afte-chumbley}.
-
-Analysis of these digitized markings require the use of statistical methods which can quantify the scientific mechanism of comparing markings and serve as basis for an error rate calculation.
+Methods for matching \gk{striated} marks for a variety of tools have been studied in the literature (see \autoref{tab:toolmarks-ER} for an overview):  in \cite{manytoolmarks1} and \cite{chumbley} digitized screwdriver marks have been analyzed using a profilometer;  in \cite{manytoolmarks2} 3D marks from screwdriver, tongue and groove pliers captured using a confocal microscope have been investigated; digitized marks from slip-joint pliers generated by a surface profilometer  have been investigated in \cite{afte-chumbley}.
+
+\gk{changed} These digitized markings allow the use of statistical methods which can quantify the scientific mechanism of comparing markings and serve as basis for an error rate calculation.
 In \cite{manytoolmarks2} a relative distance metric is defined and used as similarity measure between two toolmarks.
@@
 \end{tabular}}
-\caption{\label{tab:toolmarks-ER} Error Rates in same-source toolmark analysis reported in the literature. The top four reported papers use some variation of the Chumbley-score method.}
+\caption{\label{tab:toolmarks-ER} Error Rates in same-source \gk{striated} toolmark analysis reported in the literature. The top four reported papers use some variation of the Chumbley-score method.}
 \end{table}
@@
   \textbf{\citet{aoas}} & LEAs \\
-Consecutive Matching Bullets Striae (CMS) &  &  6.25\% &  33.85\% \\ 
-Consecutive Non-matching Striae (CNMS)  & &  6.26\% & 35.42\%  \\ 
-Average Distance (D) &   &   6.25\% & 45.83\% \\ 
+Consecutive Matching Bullets Striae (CMS) &  Ruger &  6.25\% &  33.85\% \\ 
+Consecutive Non-matching Striae (CNMS)  & Hamby 252 &  6.26\% & 35.42\%  \\ 
+Average Distance (D) & Winchester  &   6.25\% & 45.83\% \\ 
 Cross-correlation Function (CCF) & &  6.25\% & 17.71\%  \\ 
@@
    \textbf{\citet{chu2013}} & LEAs \\
-Consecutive Matching Bullets Striae (CMS) &  &  0\% &  52\% \\ 
+Consecutive Matching Bullets Striae (CMS) & Ruger &  0\% &  52\% \\ 
    \hline
    \textbf{\citet{chu2010}} \& \textbf{\citet{ma2004}} & LEAs \\
-Cross-Correlation Function (CCF) &  &  - &  - \\ 
+Cross-Correlation Function (CCF) & Ruger  &  - &  - \\ 
     \hline
 \end{tabular}}
-\caption{\label{tab:bullets-ER} Error Rates in same-source single feature land-to-land analysis reported in the literature.}
+\caption{\label{tab:bullets-ER} Error Rates in same-source single feature land-to-land analysis reported in the literature.
+\gk{Note that differences in error rates do not only come from different methods but also differently marking firearms and types of ammunition}}
 \end{table}
@@
 
+\gk{Bullet striae are most pronounced at the bottom of the bullet heel (because the heel has the most contact with the insides of the barrel).
+Howver, these areas are also affected by the break off due to friction effects between barrel and the bullet.
+An optimal crosscut is chosen orthogonally to the striae, close to the heel while avoiding break off, see also \citep{aoas}}.
 Comparisons of striae from bullets are usually based on comparisons of striae in land engraved areas, which are extracted in form of cross sections, called *profiles* \citep{aoas,ma2004}.

### 2018-08-17 doc/Response.Rmd
reviewer #2

@@
 
-*Thank you for your comments, we appreciate your insight into this.
-This paper is a first measure in the adaption process of toolmark based methods to Bullets, and we therefore plan on making a comparison of this sort by including phase (as described by \cite{chu2010}) at a later stage by making such a comparison.
-In this paper we wanted to focus on the feasibility and possibility of making a successful adaption of the Chumbley method to bullet matching*
+*Thank you for your comments, we appreciate your insight.
+You are right, for the purpose of matching bullet to bullet, all lands of each of the bullets and their relative positions on the bullet should be considered.
+This is not the problem we tried to address in this manuscript.
+To get from land-to-land comparisons to bullet-to-bullet comparisons, an additional step is required.
+In \cite{chu2010}, the average correlation value is taken.
+In this manuscript, the land-to-land comparisons result in a p-value.
+There are different options to aggregate these values at this stage to quantify a bullet-bullet comparison.
+*
 
-*We agree on the point that the toolmarks have large differences in width and the parameters of the chumbley algorithm are optimized for tools, which is why we first do a parameter study to identify the best set of parameters that would work for the LEAs.
-Before identifying problems in doing so and trying to address those problems*
+*We agree on the point that the toolmarks have large differences in width and the parameters of the chumbley algorithm are optimized for tools, which is why we first do a parameter study to identify the best set of parameters that would work for the LEAs.*
 

### 2018-08-17 doc/paper.Rmd
Update paper.Rmd

@@
 
-\gk{changed} These digitized markings allow the use of statistical methods which can quantify the scientific mechanism of comparing markings and serve as basis for an error rate calculation.
+\gk{changed} These digitized markings allow the use of statistical methods to quantify the scientific mechanism of comparing markings and serve as basis for an error rate calculation.
 In \cite{manytoolmarks2} a relative distance metric is defined and used as similarity measure between two toolmarks.

### 2018-08-17 doc/paper.Rmd
source of data in table

@@
   \hline
-  \textbf{\citet{aoas}} & LEAs \\
-Consecutive Matching Bullets Striae (CMS) &  Ruger &  6.25\% &  33.85\% \\ 
-Consecutive Non-matching Striae (CNMS)  & Hamby 252 &  6.26\% & 35.42\%  \\ 
+  \textbf{\citet{aoas}} & Hamby 252 (NIST scans) \\
+Consecutive Matching Bullets Striae (CMS) &   &  6.25\% &  33.85\% \\ 
+Consecutive Non-matching Striae (CNMS)  &  &  6.26\% & 35.42\%  \\ 
 Average Distance (D) & Winchester  &   6.25\% & 45.83\% \\ 
@@
    \hline
-   \textbf{\citet{chu2013}} & LEAs \\
-Consecutive Matching Bullets Striae (CMS) & Ruger &  0\% &  52\% \\ 
+   \textbf{\citet{chu2013}} & Hamby 252 (NIST scans) \\
+Consecutive Matching Bullets Striae (CMS) &  &  0\% &  52\% \\ 
    \hline
-   \textbf{\citet{chu2010}} \& \textbf{\citet{ma2004}} & LEAs \\
-Cross-Correlation Function (CCF) & Ruger  &  - &  - \\ 
+   \textbf{\citet{chu2010}} & \small{6 types of firearms, 2 types of ammunition} \\
+   \textbf{\citet{ma2004}} & \small{NIST standard bullet SRM 8240} \\
+Cross-Correlation Function (CCF) &   &  - &  - \\ 
     \hline

### 2018-08-17 doc/paper.tex
reference adher(?) to style guide
adding modified bst file
update reference to aoas paper
source of data in table

@@
 \usepackage{enumerate}
+%\usepackage[numbers]{natbib}
 \usepackage{natbib}
 \setcitestyle{numbers} 
+\makeatletter % Reference list option change
+\renewcommand\@biblabel[1]{#1. } % from [1] to 1
+\makeatother %
 
@@
 \usepackage{subfig}
-\usepackage[table]{xcolor}
+\usepackage[table,xcdraw]{xcolor}
 \usepackage{wrapfig}
 \usepackage{float}
-\usepackage{colortbl}
-\usepackage{hyperref}
+%\usepackage{colortbl}
+\usepackage[colorlinks]{hyperref}
+\hypersetup{
+  colorlinks=true,
+  citecolor=black,
+  linkcolor=black,
+  urlcolor=blue}
+
 \usepackage{pdflscape}
@@
 \usepackage{threeparttable}
+\usepackage{url} % not crucial - just used below for the URL
 
-\usepackage{url} % not crucial - just used below for the URL
+\usepackage{etoolbox}% http://ctan.org/pkg/etoolbox
+\makeatletter
+\patchcmd{\subsection}{\bfseries}{\relax}{}{}% Non-bold \subsection
+\patchcmd{\subsubsection}{\bfseries}{\relax}{}{}% Non-bold \subsection
+\makeatother
 
@@
 
-\tableofcontents
 \newpage
-
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
 
@@
 Council of Advisors on Science and Technology \cite{pcast2016}.
+{\textcolor{blue}{need a sentence on striated marks}}
 
 Technological advances, such as profilometers and confocal microscopy
-allow to measure 3D surfaces in a high-resolution digitized form. This
-technology has become more accessible over the last decade, and has made
-its way into topological images of ballistics evidence, such as bullet
-lands and breech faces
-\cite{DeKinder1, DeKinder2, Bachrach1, vorburger2016}. Digitized images
-of 3D surfaces of form the data basis of statistical analysis of
-toolmarks. A statistical approach based on data removes both
-subjectivity from the assessment and allows a quantification of error
-rates for both false positive and false negative identifications.
+allow {\textcolor{blue}{one}} to measure 3D surfaces in a
+high-resolution digitized form. This technology has become more
+accessible over the last decade, and has made its way into topological
+images of ballistics evidence, such as bullet lands and breech faces
+\cite{DeKinder1, DeKinder2, Bachrach1, vorburger2016}.
+{\textcolor{blue}{small change}}Digitized images of 3D surfaces, form
+the basis of statistical analysis of toolmarks. A statistical approach
+based on data removes both subjectivity from the assessment and allows a
+quantification of error rates for both false positive and false negative
+identifications.
 
@@
 
-Methods for matching marks for a variety of tools have been studied in
-the literature (see \autoref{tab:toolmarks-ER} for an overview): in
-\cite{manytoolmarks1} and \cite{chumbley} digitized screwdriver marks
-have been analyzed using a profilometer; in \cite{manytoolmarks2} 3D
-marks from screwdriver, tongue and groove pliers captured using a
-confocal microscope have been investigated; digitized marks from
-slip-joint pliers generated by a surface profilometer have been
-investigated in \cite{afte-chumbley}.
+Methods for matching {\textcolor{blue}{striated}} marks for a variety of
+tools have been studied in the literature (see
+\autoref{tab:toolmarks-ER} for an overview): in \cite{manytoolmarks1}
+and \cite{chumbley} digitized screwdriver marks have been analyzed using
+a profilometer; in \cite{manytoolmarks2} 3D marks from screwdriver,
+tongue and groove pliers captured using a confocal microscope have been
+investigated; digitized marks from slip-joint pliers generated by a
+surface profilometer have been investigated in \cite{afte-chumbley}.
 
-Analysis of these digitized markings require the use of statistical
-methods which can quantify the scientific mechanism of comparing
-markings and serve as basis for an error rate calculation. In
+{\textcolor{blue}{changed}} These digitized markings allow the use of
+statistical methods which can quantify the scientific mechanism of
+comparing markings and serve as basis for an error rate calculation. In
 \cite{manytoolmarks2} a relative distance metric is defined and used as
@@
 \end{tabular}}
-\caption{\label{tab:toolmarks-ER} Error Rates in same-source toolmark analysis reported in the literature. The top four reported papers use some variation of the Chumbley-score method.}
+\caption{\label{tab:toolmarks-ER} Error Rates in same-source {\textcolor{blue}{striated}} toolmark analysis reported in the literature.
+The top four reported papers use some variation of the Chumbley-score method.}
 \end{table}
@@
   \hline
-  \textbf{\citet{aoas}} & LEAs \\
+  \textbf{\citet{aoas}} & Hamby 252 (NIST scans) \\
 Consecutive Matching Bullets Striae (CMS) &   &  6.25\% &  33.85\% \\ 
 Consecutive Non-matching Striae (CNMS)  &  &  6.26\% & 35.42\%  \\ 
-Average Distance (D) &   &   6.25\% & 45.83\% \\ 
+Average Distance (D) & Winchester  &   6.25\% & 45.83\% \\ 
 Cross-correlation Function (CCF) & &  6.25\% & 17.71\%  \\ 
@@
    \hline
-   \textbf{\citet{chu2013}} & LEAs \\
+   \textbf{\citet{chu2013}} & Hamby 252 (NIST scans) \\
 Consecutive Matching Bullets Striae (CMS) &  &  0\% &  52\% \\ 
    \hline
-   \textbf{\citet{chu2010}} \& \textbf{\citet{ma2004}} & LEAs \\
+   \textbf{\citet{chu2010}} & \small{6 types of firearms, 2 types of ammunition} \\
+   \textbf{\citet{ma2004}} & \small{NIST standard bullet SRM 8240} \\
 Cross-Correlation Function (CCF) &   &  - &  - \\ 
@@
 \end{tabular}}
-\caption{\label{tab:bullets-ER} Error Rates in same-source single feature land-to-land analysis reported in the literature.}
+\caption{\label{tab:bullets-ER} Error Rates in same-source single feature land-to-land analysis reported in the literature.
+{\textcolor{blue}{Note that differences in error rates do not only come from different methods but also differently marking firearms and types of ammunition}}}
 \end{table}
@@
 \hypertarget{methods}{%
-\section{Methods}\label{methods}}
+\section*{Methods}\label{methods}}
+\addcontentsline{toc}{section}{Methods}
 
 \hypertarget{scans-for-land-engraved-areas}{%
-\subsection{Scans for land engraved
-areas}\label{scans-for-land-engraved-areas}}
+\subsection*{\texorpdfstring{\emph{Scans for land engraved
+areas}}{Scans for land engraved areas}}\label{scans-for-land-engraved-areas}}
+\addcontentsline{toc}{subsection}{\emph{Scans for land engraved areas}}
 
+{\textcolor{blue}{Bullet striae are most pronounced at the bottom of the bullet heel (because the heel has the most contact with the insides of the barrel).
+Howver, these areas are also affected by the break off due to friction effects between barrel and the bullet.
+An optimal crosscut is chosen orthogonally to the striae, close to the heel while avoiding break off, see also \citep{aoas}}}.
 Comparisons of striae from bullets are usually based on comparisons of
@@
 \hypertarget{the-chumbley-score-test}{%
-\subsection{The Chumbley Score Test}\label{the-chumbley-score-test}}
+\subsection*{\texorpdfstring{\emph{The Chumbley Score
+Test}}{The Chumbley Score Test}}\label{the-chumbley-score-test}}
+\addcontentsline{toc}{subsection}{\emph{The Chumbley Score Test}}
 
@@
 \hypertarget{a-problem-with-failed-tests}{%
-\subsection{A problem with failed
-tests}\label{a-problem-with-failed-tests}}
+\subsection*{\texorpdfstring{\emph{A problem with failed
+tests}}{A problem with failed tests}}\label{a-problem-with-failed-tests}}
+\addcontentsline{toc}{subsection}{\emph{A problem with failed tests}}
 
@@
 \hypertarget{a-modified-approach}{%
-\subsection{A modified approach}\label{a-modified-approach}}
+\subsection*{\texorpdfstring{\emph{A modified
+approach}}{A modified approach}}\label{a-modified-approach}}
+\addcontentsline{toc}{subsection}{\emph{A modified approach}}
 
@@
 \hypertarget{testing-setup}{%
-\subsection{Testing setup}\label{testing-setup}}
+\subsection*{\texorpdfstring{\emph{Testing
+setup}}{Testing setup}}\label{testing-setup}}
+\addcontentsline{toc}{subsection}{\emph{Testing setup}}
 
 \hypertarget{the-data}{%
-\subsubsection{The Data}\label{the-data}}
+\subsubsection*{The Data}\label{the-data}}
+\addcontentsline{toc}{subsubsection}{The Data}
 
@@
 \hypertarget{setup}{%
-\subsubsection{Setup}\label{setup}}
+\subsubsection*{Setup}\label{setup}}
+\addcontentsline{toc}{subsubsection}{Setup}
 
@@
 \hypertarget{results}{%
-\section{Results}\label{results}}
+\section*{Results}\label{results}}
+\addcontentsline{toc}{section}{Results}
 
 \hypertarget{failed-tests}{%
-\subsection{Failed Tests}\label{failed-tests}}
+\subsection*{\texorpdfstring{\emph{Failed
+Tests}}{Failed Tests}}\label{failed-tests}}
+\addcontentsline{toc}{subsection}{\emph{Failed Tests}}
 
@@
 \hypertarget{coarseness}{%
-\subsection{Coarseness}\label{coarseness}}
+\subsection*{\texorpdfstring{\emph{Coarseness}}{Coarseness}}\label{coarseness}}
+\addcontentsline{toc}{subsection}{\emph{Coarseness}}
 
@@
 \hypertarget{error-rate-assessment}{%
-\subsection{Error rate assessment}\label{error-rate-assessment}}
+\subsection*{\texorpdfstring{\emph{Error rate
+assessment}}{Error rate assessment}}\label{error-rate-assessment}}
+\addcontentsline{toc}{subsection}{\emph{Error rate assessment}}
 
@@
 \hypertarget{observed-versus-nominal-type-i-error-rates}{%
-\subsection{Observed versus Nominal Type I error
-rates}\label{observed-versus-nominal-type-i-error-rates}}
+\subsection*{\texorpdfstring{\emph{Observed versus Nominal Type I error
+rates}}{Observed versus Nominal Type I error rates}}\label{observed-versus-nominal-type-i-error-rates}}
+\addcontentsline{toc}{subsection}{\emph{Observed versus Nominal Type I
+error rates}}
 
@@
 \hypertarget{high-resolution-hamby-44-scans}{%
-\subsection{High resolution Hamby 44
-scans}\label{high-resolution-hamby-44-scans}}
+\subsection*{\texorpdfstring{\emph{High resolution Hamby 44
+scans}}{High resolution Hamby 44 scans}}\label{high-resolution-hamby-44-scans}}
+\addcontentsline{toc}{subsection}{\emph{High resolution Hamby 44 scans}}
 
@@
 \hypertarget{conclusions}{%
-\section{Conclusions}\label{conclusions}}
+\section*{Conclusions}\label{conclusions}}
+\addcontentsline{toc}{section}{Conclusions}
 
@@
 \hypertarget{appendix}{%
-\section{Appendix}\label{appendix}}
+\section*{Appendix}\label{appendix}}
+\addcontentsline{toc}{section}{Appendix}
 
@@
 
-\bibliographystyle{vancouver}
+\bibliographystyle{jfs-authoryear}
 \bibliography{bibliography}

### 2018-08-17 doc/paper.Rmd
added h44 table fpfnr same/ auc diff ?

@@
 
+```{r, fig.cap="h44 Overview of all of the results."}
+colnames(h44)[11]<- "coarse"
+roc44.summary <- h44 %>% group_by(wo, wv, method, coarse) %>%
+  summarize(
+    eer = EER(1-p_value, match),
+    auc = list(pROC::ci.auc(match, 1-p_value))
+  )
+
+roc44.summary <- roc44.summary %>% 
+  mutate(
+    lower = auc %>% purrr::map_dbl(.f = function(x) x[1]),
+    upper = auc %>% purrr::map_dbl(.f = function(x) x[3]),
+    AUC = auc %>% purrr::map_dbl(.f = function(x) x[2])
+  )
+
+errorsCV.h44 <- rbind(errorrate(h44, 0.001), 
+                   errorrate(h44, 0.005), 
+                   errorrate(h44, 0.01), 
+                   errorrate(h44, 0.05))
+
+roc44.summary <- roc44.summary %>% left_join(
+  errorsCV.h44 %>% dplyr::select(wv, wo, alpha, actual, beta) %>% 
+    dplyr::filter(alpha == 0.05)
+)
+
+rocs44 <- roc44.summary %>% dplyr::select(-auc, -coarse, -alpha)
+rocs44 <- rocs44 %>% mutate(
+  aucplus = sprintf("%.3f (%.3f, %.3f)", AUC, lower, upper)
+)
+rocs44 <- rocs44 %>% dplyr::filter(wv==75)
+rocs44 <- rocs44 %>% dplyr::select(-AUC, -lower, -upper, -wv)
+rocs1.44 <- rocs44 %>% dplyr::filter(method=="cs1")
+rocs2.44 <- rocs44 %>% dplyr::filter(method=="cs2")
+rocs2.44<- rocs2.44[-3,]
+rocs44_tab <- data.frame(rocs1.44 %>% ungroup() %>% dplyr::select(-wv, -method, -eer), 
+                       rocs2.44 %>% ungroup() %>% dplyr::select(actual, beta, aucplus))
+dt44 <- xtable(rocs44_tab, 
+             digits=c(0,0,3, 3,0,3,3,0))
+```
+
+\begin{table}[hbtp]
+\centering
+\begin{adjustbox}{max width=\textwidth}
+\begin{tabular}{rrrlrrl}
+  \hline
+  & \multicolumn{3}{l}{\bf CS1} & \multicolumn{3}{l}{\bf CS2} \\
+$w_o$ & \bf FPR & \bf FNR & \bf AUC (95\% C.I.) & \bf FPR & \bf FNR & \bf AUC (95\% C.I.) \\  
+  \hline
+90 & 0.066 & 0.512 & 0.680 (0.633, 0.727) & 0.066 & 0.512 & 0.729 (0.684, 0.774) \\ 
+  120 & 0.069 & 0.468 & 0.736 (0.693, 0.779) & 0.069 & 0.468 & 0.763 (0.720, 0.806) \\ 
+  140 & 0.070 & 0.473 & 0.741 (0.698, 0.784) & 0.070 & 0.473 & 0.767 (0.727, 0.808) \\ 
+  150 & 0.067 & 0.457 & 0.730 (0.685, 0.775) & 0.067 & 0.457 & 0.774 (0.733, 0.816) \\ 
+  180 & 0.071 & 0.446 & 0.773 (0.734, 0.813) & 0.071 & 0.446 & 0.773 (0.730, 0.816) \\ 
+  210 & 0.070 & 0.400 & 0.780 (0.740, 0.821) & 0.070 & 0.400 & 0.803 (0.763, 0.843) \\ 
+  240 & 0.070 & 0.396 & 0.784 (0.744, 0.824) & 0.070 & 0.396 & 0.801 (0.760, 0.841) \\ 
+  270 & 0.065 & 0.394 & 0.782 (0.741, 0.823) & 0.065 & 0.394 & 0.808 (0.769, 0.848) \\ 
+  300 & 0.066 & 0.394 & 0.791 (0.751, 0.832) & 0.066 & 0.394 & 0.814 (0.776, 0.852) \\ 
+  330 & 0.066 & 0.405 & 0.791 (0.753, 0.830) & 0.066 & 0.405 & 0.804 (0.765, 0.844) \\ 
+  360 & 0.066 & 0.413 & 0.788 (0.747, 0.828) & 0.066 & 0.413 & 0.803 (0.765, 0.842) \\ 
+  390 & 0.067 & 0.391 & 0.789 (0.748, 0.830) & 0.067 & 0.391 & 0.810 (0.771, 0.849) \\ 
+  480 & 0.063 & 0.419 & 0.755 (0.711, 0.800) & 0.063 & 0.419 & 0.801 (0.761, 0.841) \\ 
+   \hline
+\end{tabular}
+\end{adjustbox}
+\caption{\label{tab:aucsh44} Overview of high resolution scan results as shown in Figures \ref{fig:h44}.
+FPR is the observed rate of false positives (for a fixed level $\alpha = 0.05$), FNR is the false negative rate.
+Area under the curve (AUC) is shown with confidence intervals.
+}
+\end{table}
+
 # Conclusions {-}

### 2018-08-18 doc/Response.Rmd
couple more comments

@@
 
-*By failed test, what is being meant is that the method fails to conduct a statistical test although there is a resolution available in the ground truth as whether the lands match or dont match.
-This is because of incorrect identification of Maximum correlations windows required for comparison of the two striation marks.
-This means that there is no resolution as to the two striations being a match or a non-match, when there should be one *
-
-*Thank you for pointing this out. For the terminology error rates, we do mean FPR = FP/ (TN + FP) and FNR = FN/ (TP + FN), which holds with the definition of error rates in literature.*
-
-*Thank you for pointing out the confusing terminology as seen in the "Missed identifications" Page 17. missed same-source marking. *
-
-*We are aware of the current standards of classification as having a section as inconclusive for examination.
-But for the data sets we are working on, each and every land has been identified and none of the data being used, cannot be classified as from the same-source and different source*
+*You are right - the term 'error rate' is highly charged emotionally for firearms examiners.
+The Chumbley score test is peculiar in the sense, that the null hypothesis is stated as 'two markings coming from a different source'.
+For clarification, we have added a paragraph at the end of the section on the the Chumbley score that details the test hypotheses with the corresponding (statistical) errors.
+We have also added a sentence to the effect that all of the comparisons run on the Hamby 252 set only included lands that were deemed suitable for a comparison (e.g.
+not affected by tank rash), similar to Hare et al (2016).
+Lands leading to inconclusive results by FTEs are therefore not included in the comparisons and would not lead to a 'missed identifaction'.
+*
 

### 2018-08-18 doc/Response.Rmd
hypotheses  and error rates

@@
 The Chumbley score test is peculiar in the sense, that the null hypothesis is stated as 'two markings coming from a different source'.
-For clarification, we have added a paragraph at the end of the section on the the Chumbley score that details the test hypotheses with the corresponding (statistical) errors.
+For clarification, we have added a paragraph at the end of the section on the the Chumbley score that details the test hypotheses and a paragraph at the end of the following section with the corresponding (statistical) errors.
 We have also added a sentence to the effect that all of the comparisons run on the Hamby 252 set only included lands that were deemed suitable for a comparison (e.g.
@@
 
-*By coarseness we mean the parameter used to identify smoothing span as used in the LOWESS filter.*
+*A coarse Gaussian filter was indeed used to remove some class characteristics in Bachrach et al (2010) for toolmark striae.
+In this manuscript, the filter is used to remove bullet curvature.
+However, subclass characteristics expressed in form of 'waviness' in the barrel (e.g.
+introduced by specifics in the firing process) can be addressed in the same way as for toolmarks.*
 
@@
 
-*We agree that having a Forensic examiner review the draft would bring value to the validation of the method.
-We would very politely like to bring to light that the work in this paper is incremental to the work done by \citet{hadler} and \citet{chumbley}.
-The objectivity of the methods were established in these works, and we only change the pairing of certain windows in the different-shift substep which does not change the format of the algorithm to compromise objectivity in the analysis*
+*Our research center is working closely with forensic examiners, including firearms examiners. Thank you for the advice.*

### 2018-08-18 doc/paper.Rmd
hypotheses  and error rates

@@
 title: "Adapting the Chumbley Score to match striae on Land Engraved Areas of bullets"
+affiliation: Department of Statistics, Iowa State University and CSAFE (Center for
+  Statistics and Applications in Forensic Evidence)
 authors:
-# - affiliation: Department of Statistics, Iowa State University 
-- name: Ganesh Krishnan,
-  degree: Master of Science (M.Sc.)
-  # thank: The authors gratefully acknowledge funding from CSAFE through cooperative agreement \# 70NANB15H176 between NIST and Iowa State University.
-We would also like to thank the efforts of CSAFE and Roy J carver Microscopy lab in scanning the Hamby-44 set and providing the scans to us.
-# - affiliation: Department of Statistics and CSAFE, Iowa State University 
-- name: Prof. Heike Hofmann,
-  degree: Doctor of Philosophy (Ph.D.)
-affiliation: Department of Statistics, Iowa State University and CSAFE (Center for Statistics and Applications in Forensic Evidence)
-funding: The authors gratefully acknowledge funding from CSAFE through cooperative agreement \# 70NANB15H176 between NIST and Iowa State University. 
-thank: We would like to thank the efforts of CSAFE and Roy J Carver Microscopy lab in scanning the Hamby-44 set and providing the scans to us.
-We would also like to thank NIST for providing the Hamby-44 bullets to us for scanning.
-presentedat:  This work was presented in the form of a poster in the AFTE 2018 49th Annual Training Seminar, held from June 3rd to June 8th, 2018 in Charleston, West Virginia
-biblio-style: jfs
+- degree: Master of Science (M.Sc.)
+  name: Ganesh Krishnan,
+- degree: Doctor of Philosophy (Ph.D.)
+  name: Prof. Heike Hofmann,
 output:
@@
   html_document: default
+bibliography: bibliography
 blinded: 0
+funding: The authors gratefully acknowledge funding from CSAFE through cooperative
+  agreement \# 70NANB15H176 between NIST and Iowa State University.
 keywords:
@@
 - same-source problem
-bibliography: bibliography
-abstract: The same-source problem remains a major challenge in forensic toolmark and firearm examination.
-Technological advances in surface imaging allow measurements of 3D surfaces at previously unforeseen resolutions and enable digitized imaging.
-Here, we investigate the applicability of the Chumbley scoring method \cite{hadler}, developed for screwdriver markings, for same-source identification of bullet striae.
-We provide methods to identify parameters that minimize the error rates for matching of LEAs using Hamby datasets 44 and 252 measured by NIST and CSAFE.
-We suggest a remedial algorithm to alleviate the problem of failed tests in the method, increasing both the power of the test and reducing error rates.
-Type II error rates of the proposed method improve by more than one third (Type I error of 0.05) and are on average about 0.22.
-This puts the proposed method on similar footing as other single feature matching approaches in the literature.
+biblio-style: jfs
+presentedat: This work was presented in the form of a poster in the AFTE 2018 49th
+  Annual Training Seminar, held from June 3rd to June 8th, 2018 in Charleston, West
+  Virginia
+thank: We would like to thank the efforts of CSAFE and Roy J Carver Microscopy lab
+  in scanning the Hamby-44 set and providing the scans to us. We would also like to
+  thank NIST for providing the Hamby-44 bullets to us for scanning.
+abstract: The same-source problem remains a major challenge in forensic toolmark and
+  firearm examination. Technological advances in surface imaging allow measurements
+  of 3D surfaces at previously unforeseen resolutions and enable digitized imaging.
+  Here, we investigate the applicability of the Chumbley scoring method \cite{hadler},
+  developed for screwdriver markings, for same-source identification of bullet striae.
+  We provide methods to identify parameters that minimize the error rates for matching
+  of LEAs using Hamby datasets 44 and 252 measured by NIST and CSAFE. We suggest a
+  remedial algorithm to alleviate the problem of failed tests in the method, increasing
+  both the power of the test and reducing error rates. Type II error rates of the
+  proposed method improve by more than one third (Type I error of 0.05) and are on
+  average about 0.22. This puts the proposed method on similar footing as other single
+  feature matching approaches in the literature.
 ---
@@
 Consecutive Non-matching Striae (CNMS)  &  &  6.26\% & 35.42\%  \\ 
-Average Distance (D) & Winchester  &   6.25\% & 45.83\% \\ 
+Average Distance (D) &   &   6.25\% & 45.83\% \\ 
 Cross-correlation Function (CCF) & &  6.25\% & 17.71\%  \\ 
@@
 \caption{\label{tab:bullets-ER} Error Rates in same-source single feature land-to-land analysis reported in the literature.
-\gk{Note that differences in error rates do not only come from different methods but also differently marking firearms and types of ammunition}}
+\hh{Note that only error rates from the same data source can be compared directly between methods.
+Error rates from different sources are influenced by type of firearm and ammunition used.}}
 \end{table}
@@
 \centering
-\includegraphics[width=\textwidth]{images/B6-B2-L6-rescaled.png}
+\includegraphics[width=\textwidth]{images/B6-B2-L6-rescaled.pdf}
 ```{r, eval=FALSE, echo=FALSE, fig.width=8, fig.height=3, out.width='\\textwidth'}
@@
 
-\gk{Bullet striae are most pronounced at the bottom of the bullet heel (because the heel has the most contact with the insides of the barrel).
+Comparisons of striae from bullets are usually based on comparisons of striae in land engraved areas, which are extracted in form of cross sections, called *profiles* \citep{aoas,ma2004}.
+Bullet striae are most pronounced at the bottom of the bullet heel (because the heel has the most contact with the insides of the barrel).
 Howver, these areas are also affected by the break off due to friction effects between barrel and the bullet.
-An optimal crosscut is chosen orthogonally to the striae, close to the heel while avoiding break off, see also \citep{aoas}}.
-Comparisons of striae from bullets are usually based on comparisons of striae in land engraved areas, which are extracted in form of cross sections, called *profiles* \citep{aoas,ma2004}.
+An optimal cross section is chosen orthogonally to the striae, close to the heel while avoiding break off as shown in \autoref{fig:rgl}, see also \citep{aoas} for mathematical details of the extraction.
+
  From profiles bullet *signatures* \citep{chu2013,aoas} are extracted as residuals of a loess fit or Gaussian filter.
@@
 The intuition here is that for two markings from the same-source the correlation for the in-phase sample should be high, while the correlations of the out-of-phase sample  provide a measure for the base-level correlation for non-matching marks of a given length $w_v$.
-The Chumbley score is then computed as a Mann Whitney U statistic to compare between in-phase sample and out-of-phase sample.
+\hh{More specifically, the null hypothesis of the Chumbley score test is stated as **$H_o$: the markings come from two different sources**} with the alternative given as **$H_a$: the markings come from the same source**.
+A $p$-value for the Chumbley score test is } then computed as a Mann Whitney U statistic to compare between in-phase sample and out-of-phase sample.
+A low $p$-value is then interpreted as a rejection of the null hypothesis in favor of the alternative hypothesis of `same-source`.}
+
 In the original method proposed in \citet{chumbley} both in-phase and out-of-phase sample are extracted randomly, whereas \citet{hadler} proposed the above specified deterministic rules for both samples to make the resulting score deterministic while simultaneously avoiding overlaps within selected marks to ensure independence.
@@
 Looking closer at \autoref{eq.start}, we see that by definition, some number of tests will fail to produce a result.
-The problem of failed tests is first mentioned in \citet{afte-chumbley}.
+\hh{Note that this problem is different from erroneous test results.} The problem of failed tests is first mentioned in \citet{afte-chumbley}.
 Unfortunately, the authors do not provide any percentage of how many tests failed for their data.
@@
 In the remainder of the paper, we will refer to the algorithm defined by \citet{hadler} as \textbf{(CS1)} and the suggested modified algorithm as \textbf{(CS2)} and compare their performance on the available scans of the Hamby study.
+
+\hh{Performance of tests is measured with respect to the errors they make in situations where ground truth is known.
+We distinguish between two error rates: (1) false positive and (2) false negative rate.
+*False positives* (or *false identifications*) are situations where the test indicates a match (i.e.
+the test falsely rejects) but the markings come from different sources.
+This is also known as a *Type I* error.
+*False negatives* (or *missed identifications*) are situations where the test fails to reject, i.e.
+the test indicates that the markings come from different sources, but in fact the markings come from the same source.
+This is a *Type II* error.}
+
+\hh{Note that in all of the following land-to-land comparisons only lands are compared that are suitable for a comparison, i.e.
+a signature can be extracted from the scan.
+In particular, lands which exhibited tank rash  were removed from comparison \citep{aoas}.
+}
+
 The section \ref{appendix:appxfailed} of the Appendix discusses about the various scenarious of failed tests.

### 2018-08-18 doc/paper.tex
hypotheses  and error rates

@@
 {\textcolor{blue}{changed}} These digitized markings allow the use of
-statistical methods which can quantify the scientific mechanism of
-comparing markings and serve as basis for an error rate calculation. In
+statistical methods to quantify the scientific mechanism of comparing
+markings and serve as basis for an error rate calculation. In
 \cite{manytoolmarks2} a relative distance metric is defined and used as
@@
 \caption{\label{tab:bullets-ER} Error Rates in same-source single feature land-to-land analysis reported in the literature.
-{\textcolor{blue}{Note that differences in error rates do not only come from different methods but also differently marking firearms and types of ammunition}}}
+{\textcolor{orange}{Note that only error rates from the same data source can be compared directly between methods.
+Error rates from different sources are influenced by type of firearm and ammunition used.}}}
 \end{table}
@@
 \centering
-\includegraphics[width=\textwidth]{images/B6-B2-L6-rescaled.png}
+\includegraphics[width=\textwidth]{images/B6-B2-L6-rescaled.pdf}
 
@@
 
-{\textcolor{blue}{Bullet striae are most pronounced at the bottom of the bullet heel (because the heel has the most contact with the insides of the barrel).
-Howver, these areas are also affected by the break off due to friction effects between barrel and the bullet.
-An optimal crosscut is chosen orthogonally to the striae, close to the heel while avoiding break off, see also \citep{aoas}}}.
 Comparisons of striae from bullets are usually based on comparisons of
 striae in land engraved areas, which are extracted in form of cross
-sections, called \emph{profiles} \citep{aoas,ma2004}. From profiles
-bullet \emph{signatures} \citep{chu2013,aoas} are extracted as residuals
-of a loess fit or Gaussian filter. This effectively removes topographic
-structure from the data in the attempt to increase the signal to noise
-ratio. \autoref{fig:rgl} shows how the signature from a bullet land
-(bottom) lines up with the image of the land (top) from which it was
-extracted. We can see in the figure how the depth and relative position
-of the striation markings seen in the image are interpreted as peaks and
-valleys in the signature.
+sections, called \emph{profiles} \citep{aoas,ma2004}. Bullet striae are
+most pronounced at the bottom of the bullet heel (because the heel has
+the most contact with the insides of the barrel). Howver, these areas
+are also affected by the break off due to friction effects between
+barrel and the bullet. An optimal cross section is chosen orthogonally
+to the striae, close to the heel while avoiding break off as shown in
+\autoref{fig:rgl}, see also \citep{aoas} for mathematical details of the
+extraction.
+
+From profiles bullet \emph{signatures} \citep{chu2013,aoas} are
+extracted as residuals of a loess fit or Gaussian filter. This
+effectively removes topographic structure from the data in the attempt
+to increase the signal to noise ratio. \autoref{fig:rgl} shows how the
+signature from a bullet land (bottom) lines up with the image of the
+land (top) from which it was extracted. We can see in the figure how the
+depth and relative position of the striation markings seen in the image
+are interpreted as peaks and valleys in the signature.
 
@@
 measure for the base-level correlation for non-matching marks of a given
-length \(w_v\). The Chumbley score is then computed as a Mann Whitney U
-statistic to compare between in-phase sample and out-of-phase sample. In
-the original method proposed in \citet{chumbley} both in-phase and
+length \(w_v\).
+{\textcolor{orange}{More specifically, the null hypothesis of the Chumbley score test is stated as **$H_o$: the markings come from two different sources**}}
+with the alternative given as \textbf{\(H_a\): the markings come from
+the same source}. A \(p\)-value for the Chumbley score test is \} then
+computed as a Mann Whitney U statistic to compare between in-phase
+sample and out-of-phase sample. A low \(p\)-value is then interpreted as
+a rejection of the null hypothesis in favor of the alternative
+hypothesis of \texttt{same-source}.\}
+
+In the original method proposed in \citet{chumbley} both in-phase and
 out-of-phase sample are extracted randomly, whereas \citet{hadler}
@@
 Looking closer at \autoref{eq.start}, we see that by definition, some
-number of tests will fail to produce a result. The problem of failed
-tests is first mentioned in \citet{afte-chumbley}. Unfortunately, the
-authors do not provide any percentage of how many tests failed for their
-data. The algorithm fails to produce for two reasons: either the number
-of eligible same-shift pairs is zero, or the number of different-shift
-pairs is zero.
+number of tests will fail to produce a result.
+{\textcolor{orange}{Note that this problem is different from erroneous test results.}}
+The problem of failed tests is first mentioned in \citet{afte-chumbley}.
+Unfortunately, the authors do not provide any percentage of how many
+tests failed for their data. The algorithm fails to produce for two
+reasons: either the number of eligible same-shift pairs is zero, or the
+number of different-shift pairs is zero.
 
@@
 and the suggested modified algorithm as \textbf{(CS2)} and compare their
-performance on the available scans of the Hamby study. The section
-\ref{appendix:appxfailed} of the Appendix discusses about the various
-scenarious of failed tests.
+performance on the available scans of the Hamby study.
+
+{\textcolor{orange}{Performance of tests is measured with respect to the errors they make in situations where ground truth is known.
+We distinguish between two error rates: (1) false positive and (2) false negative rate.
+*False positives* (or *false identifications*) are situations where the test indicates a match (i.e.
+the test falsely rejects) but the markings come from different sources.
+This is also known as a *Type I* error.
+*False negatives* (or *missed identifications*) are situations where the test fails to reject, i.e.
+the test indicates that the markings come from different sources, but in fact the markings come from the same source.
+This is a *Type II* error.}}
+
+{\textcolor{orange}{Note that in all of the following land-to-land comparisons only lands are compared that are suitable for a comparison, i.e.
+a signature can be extracted from the scan.
+In particular, lands which exhibited tank rash  were removed from comparison \citep{aoas}.
+}}
+
+The section \ref{appendix:appxfailed} of the Appendix discusses about
+the various scenarious of failed tests.
 
@@
 
+\begin{table}[hbtp]
+\centering
+\begin{adjustbox}{max width=\textwidth}
+\begin{tabular}{rrrlrrl}
+  \hline
+  & \multicolumn{3}{l}{\bf CS1} & \multicolumn{3}{l}{\bf CS2} \\
+$w_o$ & \bf FPR & \bf FNR & \bf AUC (95\% C.I.) & \bf FPR & \bf FNR & \bf AUC (95\% C.I.) \\  
+  \hline
+90 & 0.066 & 0.512 & 0.680 (0.633, 0.727) & 0.066 & 0.512 & 0.729 (0.684, 0.774) \\ 
+  120 & 0.069 & 0.468 & 0.736 (0.693, 0.779) & 0.069 & 0.468 & 0.763 (0.720, 0.806) \\ 
+  140 & 0.070 & 0.473 & 0.741 (0.698, 0.784) & 0.070 & 0.473 & 0.767 (0.727, 0.808) \\ 
+  150 & 0.067 & 0.457 & 0.730 (0.685, 0.775) & 0.067 & 0.457 & 0.774 (0.733, 0.816) \\ 
+  180 & 0.071 & 0.446 & 0.773 (0.734, 0.813) & 0.071 & 0.446 & 0.773 (0.730, 0.816) \\ 
+  210 & 0.070 & 0.400 & 0.780 (0.740, 0.821) & 0.070 & 0.400 & 0.803 (0.763, 0.843) \\ 
+  240 & 0.070 & 0.396 & 0.784 (0.744, 0.824) & 0.070 & 0.396 & 0.801 (0.760, 0.841) \\ 
+  270 & 0.065 & 0.394 & 0.782 (0.741, 0.823) & 0.065 & 0.394 & 0.808 (0.769, 0.848) \\ 
+  300 & 0.066 & 0.394 & 0.791 (0.751, 0.832) & 0.066 & 0.394 & 0.814 (0.776, 0.852) \\ 
+  330 & 0.066 & 0.405 & 0.791 (0.753, 0.830) & 0.066 & 0.405 & 0.804 (0.765, 0.844) \\ 
+  360 & 0.066 & 0.413 & 0.788 (0.747, 0.828) & 0.066 & 0.413 & 0.803 (0.765, 0.842) \\ 
+  390 & 0.067 & 0.391 & 0.789 (0.748, 0.830) & 0.067 & 0.391 & 0.810 (0.771, 0.849) \\ 
+  480 & 0.063 & 0.419 & 0.755 (0.711, 0.800) & 0.063 & 0.419 & 0.801 (0.761, 0.841) \\ 
+   \hline
+\end{tabular}
+\end{adjustbox}
+\caption{\label{tab:aucsh44} Overview of high resolution scan results as shown in Figures \ref{fig:h44}.
+FPR is the observed rate of false positives (for a fixed level $\alpha = 0.05$), FNR is the false negative rate.
+Area under the curve (AUC) is shown with confidence intervals.
+}
+\end{table}
+
 \hypertarget{conclusions}{%

### 2018-08-18 doc/paper.Rmd
fixed table - error rates need to be calculated separately for the two methods

@@
 
-The section \ref{appendix:appxfailed} of the Appendix discusses about the various scenarious of failed tests.
+Section \ref{appendix:appxfailed} in the Appendix discusses  various scenarios of failed tests.
 \begin{figure}[hbtp]
@@
 
-errorsCV.h44 <- rbind(errorrate(h44, 0.001), 
-                   errorrate(h44, 0.005), 
-                   errorrate(h44, 0.01), 
-                   errorrate(h44, 0.05))
+# need to distinguish between methods
+h44cs1 <- h44 %>% filter(method=="cs1")
+errorsCV.h44.cs1 <- errorrate(h44cs1, 0.05)
+errorsCV.h44.cs1$method <- "cs1"
+h44cs2 <- h44 %>% filter(method=="cs2")
+errorsCV.h44.cs2 <- errorrate(h44cs2, 0.05)
+errorsCV.h44.cs2$method <- "cs2"
+
+errorsCV.h44 <- rbind(
+  errorsCV.h44.cs1,
+  errorsCV.h44.cs2
+)
 
 roc44.summary <- roc44.summary %>% left_join(
-  errorsCV.h44 %>% dplyr::select(wv, wo, alpha, actual, beta) %>% 
-    dplyr::filter(alpha == 0.05)
+  errorsCV.h44 %>% dplyr::select(-total, method) 
 )
@@
 )
-rocs44 <- rocs44 %>% dplyr::filter(wv==75)
+rocs44 <- rocs44 %>% dplyr::filter(wv==75, wo %% 30 == 0)
 rocs44 <- rocs44 %>% dplyr::select(-AUC, -lower, -upper, -wv)
@@
 rocs2.44 <- rocs44 %>% dplyr::filter(method=="cs2")
-rocs2.44<- rocs2.44[-3,]
+#rocs2.44<- rocs2.44[-3,]
 rocs44_tab <- data.frame(rocs1.44 %>% ungroup() %>% dplyr::select(-wv, -method, -eer), 
@@
              digits=c(0,0,3, 3,0,3,3,0))
+print(dt44, include.rownames = FALSE)
 ```
@@
   \hline
-90 & 0.066 & 0.512 & 0.680 (0.633, 0.727) & 0.066 & 0.512 & 0.729 (0.684, 0.774) \\ 
-  120 & 0.069 & 0.468 & 0.736 (0.693, 0.779) & 0.069 & 0.468 & 0.763 (0.720, 0.806) \\ 
-  140 & 0.070 & 0.473 & 0.741 (0.698, 0.784) & 0.070 & 0.473 & 0.767 (0.727, 0.808) \\ 
-  150 & 0.067 & 0.457 & 0.730 (0.685, 0.775) & 0.067 & 0.457 & 0.774 (0.733, 0.816) \\ 
-  180 & 0.071 & 0.446 & 0.773 (0.734, 0.813) & 0.071 & 0.446 & 0.773 (0.730, 0.816) \\ 
-  210 & 0.070 & 0.400 & 0.780 (0.740, 0.821) & 0.070 & 0.400 & 0.803 (0.763, 0.843) \\ 
-  240 & 0.070 & 0.396 & 0.784 (0.744, 0.824) & 0.070 & 0.396 & 0.801 (0.760, 0.841) \\ 
-  270 & 0.065 & 0.394 & 0.782 (0.741, 0.823) & 0.065 & 0.394 & 0.808 (0.769, 0.848) \\ 
-  300 & 0.066 & 0.394 & 0.791 (0.751, 0.832) & 0.066 & 0.394 & 0.814 (0.776, 0.852) \\ 
-  330 & 0.066 & 0.405 & 0.791 (0.753, 0.830) & 0.066 & 0.405 & 0.804 (0.765, 0.844) \\ 
-  360 & 0.066 & 0.413 & 0.788 (0.747, 0.828) & 0.066 & 0.413 & 0.803 (0.765, 0.842) \\ 
-  390 & 0.067 & 0.391 & 0.789 (0.748, 0.830) & 0.067 & 0.391 & 0.810 (0.771, 0.849) \\ 
-  480 & 0.063 & 0.419 & 0.755 (0.711, 0.800) & 0.063 & 0.419 & 0.801 (0.761, 0.841) \\ 
+90 & 0.077 & 0.546 & 0.680 (0.633, 0.727) & 0.056 & 0.478 & 0.729 (0.684, 0.774) \\ 
+  120 & 0.082 & 0.485 & 0.736 (0.693, 0.779) & 0.056 & 0.452 & 0.763 (0.720, 0.806) \\ 
+  150 & 0.080 & 0.493 & 0.730 (0.685, 0.775) & 0.055 & 0.422 & 0.774 (0.733, 0.816) \\ 
+  180 & 0.087 & 0.458 & 0.773 (0.734, 0.813) & 0.055 & 0.435 & 0.773 (0.730, 0.816) \\ 
+  210 & 0.086 & 0.413 & 0.780 (0.740, 0.821) & 0.055 & 0.387 & 0.803 (0.763, 0.843) \\ 
+  240 & 0.085 & 0.427 & 0.784 (0.744, 0.824) & 0.055 & 0.365 & 0.801 (0.760, 0.841) \\ 
+  270 & 0.082 & 0.415 & 0.782 (0.741, 0.823) & 0.048 & 0.374 & 0.808 (0.769, 0.848) \\ 
+  300 & 0.082 & 0.405 & 0.791 (0.751, 0.832) & 0.051 & 0.383 & 0.814 (0.776, 0.852) \\ 
+  330 & 0.082 & 0.428 & 0.791 (0.753, 0.830) & 0.050 & 0.383 & 0.804 (0.765, 0.844) \\ 
+  360 & 0.083 & 0.427 & 0.788 (0.747, 0.828) & 0.049 & 0.400 & 0.803 (0.765, 0.842) \\ 
+  390 & 0.085 & 0.400 & 0.789 (0.748, 0.830) & 0.050 & 0.383 & 0.810 (0.771, 0.849) \\ 
+  480 & 0.081 & 0.443 & 0.755 (0.711, 0.800) & 0.046 & 0.396 & 0.801 (0.761, 0.841) \\ 
    \hline

### 2018-08-19 doc/paper.Rmd
table reduced

@@
              digits=c(0,0,3, 3,0,3,3,0))
-print(dt44, include.rownames = FALSE)
+#print(dt44, include.rownames = FALSE)
 ```
@@
   \hline
-90 & 0.077 & 0.546 & 0.680 (0.633, 0.727) & 0.056 & 0.478 & 0.729 (0.684, 0.774) \\ 
-  120 & 0.082 & 0.485 & 0.736 (0.693, 0.779) & 0.056 & 0.452 & 0.763 (0.720, 0.806) \\ 
-  150 & 0.080 & 0.493 & 0.730 (0.685, 0.775) & 0.055 & 0.422 & 0.774 (0.733, 0.816) \\ 
-  180 & 0.087 & 0.458 & 0.773 (0.734, 0.813) & 0.055 & 0.435 & 0.773 (0.730, 0.816) \\ 
+%90 & 0.077 & 0.546 & 0.680 (0.633, 0.727) & 0.056 & 0.478 & 0.729 (0.684, 0.774) \\
+%  120 & 0.082 & 0.485 & 0.736 (0.693, 0.779) & 0.056 & 0.452 & 0.763 (0.720, 0.806) \\
+%  150 & 0.080 & 0.493 & 0.730 (0.685, 0.775) & 0.055 & 0.422 & 0.774 (0.733, 0.816) \\
+%  180 & 0.087 & 0.458 & 0.773 (0.734, 0.813) & 0.055 & 0.435 & 0.773 (0.730, 0.816) \\
   210 & 0.086 & 0.413 & 0.780 (0.740, 0.821) & 0.055 & 0.387 & 0.803 (0.763, 0.843) \\ 
@@
   390 & 0.085 & 0.400 & 0.789 (0.748, 0.830) & 0.050 & 0.383 & 0.810 (0.771, 0.849) \\ 
-  480 & 0.081 & 0.443 & 0.755 (0.711, 0.800) & 0.046 & 0.396 & 0.801 (0.761, 0.841) \\ 
    \hline
@@
 FPR is the observed rate of false positives (for a fixed level $\alpha = 0.05$), FNR is the false negative rate.
-Area under the curve (AUC) is shown with confidence intervals.
+Area under the curve (AUC) is shown with 95\% confidence intervals.
 }

### 2018-08-19 doc/paper.tex
fixed table - error rates need to be calculated separately for the two methods
table reduced

@@
 
-The section \ref{appendix:appxfailed} of the Appendix discusses about
-the various scenarious of failed tests.
+Section \ref{appendix:appxfailed} in the Appendix discusses various
+scenarios of failed tests.
 
@@
 
+\begin{verbatim}
+% latex table generated in R 3.5.1 by xtable 1.8-2 package
+% Sat Aug 18 16:58:28 2018
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrrlrrl}
+  \hline
+wo & actual & beta & aucplus & actual.1 & beta.1 & aucplus.1 \\ 
+  \hline
+90 & 0.077 & 0.546 & 0.680 (0.633, 0.727) & 0.056 & 0.478 & 0.729 (0.684, 0.774) \\ 
+  120 & 0.082 & 0.485 & 0.736 (0.693, 0.779) & 0.056 & 0.452 & 0.763 (0.720, 0.806) \\ 
+  150 & 0.080 & 0.493 & 0.730 (0.685, 0.775) & 0.055 & 0.422 & 0.774 (0.733, 0.816) \\ 
+  180 & 0.087 & 0.458 & 0.773 (0.734, 0.813) & 0.055 & 0.435 & 0.773 (0.730, 0.816) \\ 
+  210 & 0.086 & 0.413 & 0.780 (0.740, 0.821) & 0.055 & 0.387 & 0.803 (0.763, 0.843) \\ 
+  240 & 0.085 & 0.427 & 0.784 (0.744, 0.824) & 0.055 & 0.365 & 0.801 (0.760, 0.841) \\ 
+  270 & 0.082 & 0.415 & 0.782 (0.741, 0.823) & 0.048 & 0.374 & 0.808 (0.769, 0.848) \\ 
+  300 & 0.082 & 0.405 & 0.791 (0.751, 0.832) & 0.051 & 0.383 & 0.814 (0.776, 0.852) \\ 
+  330 & 0.082 & 0.428 & 0.791 (0.753, 0.830) & 0.050 & 0.383 & 0.804 (0.765, 0.844) \\ 
+  360 & 0.083 & 0.427 & 0.788 (0.747, 0.828) & 0.049 & 0.400 & 0.803 (0.765, 0.842) \\ 
+  390 & 0.085 & 0.400 & 0.789 (0.748, 0.830) & 0.050 & 0.383 & 0.810 (0.771, 0.849) \\ 
+  480 & 0.081 & 0.443 & 0.755 (0.711, 0.800) & 0.046 & 0.396 & 0.801 (0.761, 0.841) \\ 
+   \hline
+\end{tabular}
+\end{table}
+\end{verbatim}
+
 \begin{table}[hbtp]
@@
   \hline
-90 & 0.066 & 0.512 & 0.680 (0.633, 0.727) & 0.066 & 0.512 & 0.729 (0.684, 0.774) \\ 
-  120 & 0.069 & 0.468 & 0.736 (0.693, 0.779) & 0.069 & 0.468 & 0.763 (0.720, 0.806) \\ 
-  140 & 0.070 & 0.473 & 0.741 (0.698, 0.784) & 0.070 & 0.473 & 0.767 (0.727, 0.808) \\ 
-  150 & 0.067 & 0.457 & 0.730 (0.685, 0.775) & 0.067 & 0.457 & 0.774 (0.733, 0.816) \\ 
-  180 & 0.071 & 0.446 & 0.773 (0.734, 0.813) & 0.071 & 0.446 & 0.773 (0.730, 0.816) \\ 
-  210 & 0.070 & 0.400 & 0.780 (0.740, 0.821) & 0.070 & 0.400 & 0.803 (0.763, 0.843) \\ 
-  240 & 0.070 & 0.396 & 0.784 (0.744, 0.824) & 0.070 & 0.396 & 0.801 (0.760, 0.841) \\ 
-  270 & 0.065 & 0.394 & 0.782 (0.741, 0.823) & 0.065 & 0.394 & 0.808 (0.769, 0.848) \\ 
-  300 & 0.066 & 0.394 & 0.791 (0.751, 0.832) & 0.066 & 0.394 & 0.814 (0.776, 0.852) \\ 
-  330 & 0.066 & 0.405 & 0.791 (0.753, 0.830) & 0.066 & 0.405 & 0.804 (0.765, 0.844) \\ 
-  360 & 0.066 & 0.413 & 0.788 (0.747, 0.828) & 0.066 & 0.413 & 0.803 (0.765, 0.842) \\ 
-  390 & 0.067 & 0.391 & 0.789 (0.748, 0.830) & 0.067 & 0.391 & 0.810 (0.771, 0.849) \\ 
-  480 & 0.063 & 0.419 & 0.755 (0.711, 0.800) & 0.063 & 0.419 & 0.801 (0.761, 0.841) \\ 
+<!-- 90 & 0.077 & 0.546 & 0.680 (0.633, 0.727) & 0.056 & 0.478 & 0.729 (0.684, 0.774) \\  -->
+<!--   120 & 0.082 & 0.485 & 0.736 (0.693, 0.779) & 0.056 & 0.452 & 0.763 (0.720, 0.806) \\  -->
+<!--   150 & 0.080 & 0.493 & 0.730 (0.685, 0.775) & 0.055 & 0.422 & 0.774 (0.733, 0.816) \\  -->
+<!--   180 & 0.087 & 0.458 & 0.773 (0.734, 0.813) & 0.055 & 0.435 & 0.773 (0.730, 0.816) \\  -->
+  210 & 0.086 & 0.413 & 0.780 (0.740, 0.821) & 0.055 & 0.387 & 0.803 (0.763, 0.843) \\ 
+  240 & 0.085 & 0.427 & 0.784 (0.744, 0.824) & 0.055 & 0.365 & 0.801 (0.760, 0.841) \\ 
+  270 & 0.082 & 0.415 & 0.782 (0.741, 0.823) & 0.048 & 0.374 & 0.808 (0.769, 0.848) \\ 
+  300 & 0.082 & 0.405 & 0.791 (0.751, 0.832) & 0.051 & 0.383 & 0.814 (0.776, 0.852) \\ 
+  330 & 0.082 & 0.428 & 0.791 (0.753, 0.830) & 0.050 & 0.383 & 0.804 (0.765, 0.844) \\ 
+  360 & 0.083 & 0.427 & 0.788 (0.747, 0.828) & 0.049 & 0.400 & 0.803 (0.765, 0.842) \\ 
+  390 & 0.085 & 0.400 & 0.789 (0.748, 0.830) & 0.050 & 0.383 & 0.810 (0.771, 0.849) \\ 
    \hline

### 2018-08-19 doc/Response.Rmd
couple more comments

@@
 The new method is still susceptible to that.
-It would be interesting to further investigate thus problem in the furture.
-We have made a note of that in the conclusions*
+It would be interesting to further investigate this problem in the future.
+We have made a note of that in the conclusions.*
 
@@
 
+*Thank you for pointing out these issues. We have fixed those accordingly.*
+
 Minor edits should be made to accuracy and clarity:
@@
 I would also propose that the authors make this distinction earlier in the introduction.
+
+*We have added emphasis on striae as opposed to impressed toolmarks. *
+
 - p3 "Analysis of these digitized markings require the use of statistical methods".
@@
 - p14 how are the false negative and false positive rates computed?
+
+*We have added two paragraphs on the Chumbley Score test for clarification: the first paragraph expicitly states the hypotheses involved, the second defines error rates.*
 
@@
 
-*We followed the \cite{aoas} paper in extracting the profiles from scans. We have expanded on the discussion at the begining of section titled "Scans for land engraved areas" in paper*
+*We followed the \cite{aoas} paper in extracting the profiles from scans. We have expanded on the discussion at the begining of section titled "Scans for land engraved areas" in the paper.*
 
@@
 *Thank you for poining this out, you are right in your understanding of the method.
-We have included a figure for this and expanded out discussion, which hopefully makes the it easier to understand the method.*
+We have changed figure 5 for clarification and expanded our discussion, which hopefully makes it easier to understand the method.*
 
@@
 
-*Thank you for bringing this to our notice.
-We have included a table of error rates.
-The 93.5% to 94.1% are the True Negative rates which are being mentioned.
-It is computed as 1 - Observed FPR.
-The observed False positive rates can be seen in figure 10.
-The True positive rates would then be 1 - Observed FNR or about 64% to 72%.*
+*Thank you for bringing this to our notice. We have indluded a paragraph with a definition of error rates, and switched to a table to show the results as suggested below.*
 
@@
 - The results should all be listed as either fractions or all as percent (which is preferred) and not mix them.
+
+*We have changed the reporting to error percentages.*
+
 - p21 the authors state "this kind of assessment is only feasible in the setting of a large study, such as the one we presented".

### 2018-08-19 doc/paper.Rmd
couple more comments

@@
 
-Technological advances, such as profilometers and confocal microscopy allow \gk{one} to measure 3D surfaces in a high-resolution digitized form.
+Technological advances, such as profilometers and confocal microscopy, have made it possible to capture 3D surfaces in a high-resolution digitized form.
 This technology has become more accessible over the last decade, and has made its way into topological images of ballistics evidence, such as bullet lands and breech faces \cite{DeKinder1, DeKinder2, Bachrach1, vorburger2016}.
-\gk{small change}Digitized images of 3D surfaces, form the basis of statistical analysis of toolmarks.
-A statistical approach based on data  removes both subjectivity from the assessment and allows a quantification of error rates for both false positive and false negative identifications.
-
-\hh{changed citation style}
-Methods for matching \gk{striated} marks for a variety of tools have been studied in the literature (see \autoref{tab:toolmarks-ER} for an overview):  in \cite{manytoolmarks1} and \cite{chumbley} digitized screwdriver marks have been analyzed using a profilometer;  in \cite{manytoolmarks2} 3D marks from screwdriver, tongue and groove pliers captured using a confocal microscope have been investigated; digitized marks from slip-joint pliers generated by a surface profilometer  have been investigated in \cite{afte-chumbley}.
-
-\gk{changed} These digitized markings allow the use of statistical methods to quantify the scientific mechanism of comparing markings and serve as basis for an error rate calculation.
+Digitized images of 3D surfaces form the basis of statistical analysis of toolmarks.
+Statistical approaches based on this data remove both subjectivity from the assessment and allow a quantification of error rates for  false positive and false negative identifications.
+
+Methods for matching striated marks for a variety of tools have been studied in the literature (see \autoref{tab:toolmarks-ER} for an overview):  in \cite{manytoolmarks1} and \cite{chumbley} digitized screwdriver marks have been analyzed using a profilometer;  in \cite{manytoolmarks2} 3D marks from screwdriver, tongue and groove pliers captured using a confocal microscope have been investigated; digitized marks from slip-joint pliers generated by a surface profilometer  have been investigated in \cite{afte-chumbley}.
+
+These digitized markings allow the use of statistical methods to quantify the scientific mechanism of comparing markings and serve as basis for an error rate calculation.
 In \cite{manytoolmarks2} a relative distance metric is defined and used as similarity measure between two toolmarks.
@@
 \end{tabular}}
-\caption{\label{tab:toolmarks-ER} Error Rates in same-source \gk{striated} toolmark analysis reported in the literature. The top four reported papers use some variation of the Chumbley-score method.}
+\caption{\label{tab:toolmarks-ER} Error Rates in same-source analyses for  striated toolmarks reported in the literature. The top four reported papers use some variation of the Chumbley score method.}
 \end{table}
@@
 \caption{\label{tab:bullets-ER} Error Rates in same-source single feature land-to-land analysis reported in the literature.
-\hh{Note that only error rates from the same data source can be compared directly between methods.
-Error rates from different sources are influenced by type of firearm and ammunition used.}}
+Note that only error rates from the same data source can be compared directly between methods.
+Error rates from different sources are influenced by type of firearm and ammunition used.}
 \end{table}
@@
 The intuition here is that for two markings from the same-source the correlation for the in-phase sample should be high, while the correlations of the out-of-phase sample  provide a measure for the base-level correlation for non-matching marks of a given length $w_v$.
-\hh{More specifically, the null hypothesis of the Chumbley score test is stated as **$H_o$: the markings come from two different sources**} with the alternative given as **$H_a$: the markings come from the same source**.
-A $p$-value for the Chumbley score test is } then computed as a Mann Whitney U statistic to compare between in-phase sample and out-of-phase sample.
-A low $p$-value is then interpreted as a rejection of the null hypothesis in favor of the alternative hypothesis of `same-source`.}
+More specifically, the null hypothesis of the Chumbley score test is stated as **$H_o$: the markings come from two different sources** with the alternative given as **$H_a$: the markings come from the same source**.
+A $p$-value for the Chumbley score test is  then computed as a Mann Whitney U statistic to compare between in-phase sample and out-of-phase sample.
+A low $p$-value is  interpreted as a rejection of the null hypothesis in favor of the alternative hypothesis of `same-source`.
 
@@
 The algorithm fails to produce for two reasons: either the number of eligible same-shift pairs is zero, or the number of different-shift pairs is zero.
+Section \ref{appendix:appxfailed} in the Appendix discusses scenarios of failed tests in more detail.
 
@@
 
-\hh{Performance of tests is measured with respect to the errors they make in situations where ground truth is known.
+Performance of tests is measured with respect to the errors a test makes in situations where ground truth is known.
 We distinguish between two error rates: (1) false positive and (2) false negative rate.
@@
 the test indicates that the markings come from different sources, but in fact the markings come from the same source.
-This is a *Type II* error.}
-
-\hh{Note that in all of the following land-to-land comparisons only lands are compared that are suitable for a comparison, i.e.
+This is a *Type II* error.
+In both cases error rates are calculated as the ratio of the number of errors observed and the number of tests executed.
+
+Note that in all of the following land-to-land comparisons only lands are compared that are suitable for a comparison, i.e.
 a signature can be extracted from the scan.
 In particular, lands which exhibited tank rash  were removed from comparison \citep{aoas}.
-}
-
-Section \ref{appendix:appxfailed} in the Appendix discusses  various scenarios of failed tests.
+
 \begin{figure}[hbtp]
@@
 Figures \ref{same-shift-failure} and \ref{diff-shift-failure} show scenarios in which the deterministic Chumbley score can fail. 
+\autoref{diff-shift-failure} shows two situations, in which CS1 fails to identify any different-shift pairings.
+This happens, when both relative locations are close to either end of the signature.
+CS2 can still be computed in this situation as long as there are at least two same-shift pairs identified.
+
+\begin{figure}[hbtp]
+\centering
+\includegraphics[width=0.7\textwidth]{images/diff-shift-failure.png}
+
+\caption{\label{diff-shift-failure}Sketch of different-shift pairings. For CS1 no different-shift pairs can be identified, resulting in a failed test.}
+\end{figure}
+
 In \autoref{same-shift-failure} both CS1 and CS2 fail, because the lag between optimal locations is so large that no same-shift pairs can be found once the signatures are aligned.
@@
 
-\autoref{diff-shift-failure} shows two situations, in which CS1 fails to identify any different-shift pairings.
-This happens, when both relative locations are close to either end of the signature.
-CS2 can still be computed in this situation as long as there are at least two same-shift pairs identified.
-
-\begin{figure}[hbtp]
-\centering
-\includegraphics[width=0.7\textwidth]{images/diff-shift-failure.png}
-
-\caption{\label{diff-shift-failure}Sketch of different-shift pairings. For CS1 no different-shift pairs can be identified, resulting in a failed test.}
-\end{figure}
-
 \newpage

### 2018-08-19 doc/paper.tex
couple more comments

--- 2018-08-17 doc/paper.tex
@@
 
-Technological advances, such as profilometers and confocal microscopy
-allow {\textcolor{blue}{one}} to measure 3D surfaces in a
-high-resolution digitized form. This technology has become more
-accessible over the last decade, and has made its way into topological
-images of ballistics evidence, such as bullet lands and breech faces
-\cite{DeKinder1, DeKinder2, Bachrach1, vorburger2016}.
-{\textcolor{blue}{small change}}Digitized images of 3D surfaces, form
-the basis of statistical analysis of toolmarks. A statistical approach
-based on data removes both subjectivity from the assessment and allows a
-quantification of error rates for both false positive and false negative
-identifications.
-
-{\textcolor{orange}{changed citation style}}
+Technological advances, such as profilometers and confocal microscopy,
+have made it possible to capture 3D surfaces in a high-resolution
+digitized form. This technology has become more accessible over the last
+decade, and has made its way into topological images of ballistics
+evidence, such as bullet lands and breech faces
+\cite{DeKinder1, DeKinder2, Bachrach1, vorburger2016}. Digitized images
+of 3D surfaces form the basis of statistical analysis of toolmarks.
+Statistical approaches based on this data remove both subjectivity from
+the assessment and allow a quantification of error rates for false
+positive and false negative identifications.
 
@@
 \caption{\label{tab:bullets-ER} Error Rates in same-source single feature land-to-land analysis reported in the literature.
-{\textcolor{blue}{Note that differences in error rates do not only come from different methods but also differently marking firearms and types of ammunition}}}
+Note that only error rates from the same data source can be compared directly between methods.
+Error rates from different sources are influenced by type of firearm and ammunition used.}
 \end{table}
@@
 measure for the base-level correlation for non-matching marks of a given
-length \(w_v\). The Chumbley score is then computed as a Mann Whitney U
-statistic to compare between in-phase sample and out-of-phase sample. In
-the original method proposed in \citet{chumbley} both in-phase and
+length \(w_v\). More specifically, the null hypothesis of the Chumbley
+score test is stated as \textbf{\(H_o\): the markings come from two
+different sources} with the alternative given as \textbf{\(H_a\): the
+markings come from the same source}. A \(p\)-value for the Chumbley
+score test is then computed as a Mann Whitney U statistic to compare
+between in-phase sample and out-of-phase sample. A low \(p\)-value is
+interpreted as a rejection of the null hypothesis in favor of the
+alternative hypothesis of \texttt{same-source}.
+
+In the original method proposed in \citet{chumbley} both in-phase and
 out-of-phase sample are extracted randomly, whereas \citet{hadler}
@@
 Looking closer at \autoref{eq.start}, we see that by definition, some
-number of tests will fail to produce a result. The problem of failed
-tests is first mentioned in \citet{afte-chumbley}. Unfortunately, the
-authors do not provide any percentage of how many tests failed for their
-data. The algorithm fails to produce for two reasons: either the number
-of eligible same-shift pairs is zero, or the number of different-shift
-pairs is zero.
+number of tests will fail to produce a result.
+{\textcolor{orange}{Note that this problem is different from erroneous test results.}}
+The problem of failed tests is first mentioned in \citet{afte-chumbley}.
+Unfortunately, the authors do not provide any percentage of how many
+tests failed for their data. The algorithm fails to produce for two
+reasons: either the number of eligible same-shift pairs is zero, or the
+number of different-shift pairs is zero. Section
+\ref{appendix:appxfailed} in the Appendix discusses scenarios of failed
+tests in more detail.
 
@@
 and the suggested modified algorithm as \textbf{(CS2)} and compare their
-performance on the available scans of the Hamby study. The section
-\ref{appendix:appxfailed} of the Appendix discusses about the various
-scenarious of failed tests.
+performance on the available scans of the Hamby study.
+
+Performance of tests is measured with respect to the errors a test makes
+in situations where ground truth is known. We distinguish between two
+error rates: (1) false positive and (2) false negative rate. \emph{False
+positives} (or \emph{false identifications}) are situations where the
+test indicates a match (i.e.~the test falsely rejects) but the markings
+come from different sources. This is also known as a \emph{Type I}
+error. \emph{False negatives} (or \emph{missed identifications}) are
+situations where the test fails to reject, i.e.~the test indicates that
+the markings come from different sources, but in fact the markings come
+from the same source. This is a \emph{Type II} error. In both cases
+error rates are calculated as the ratio of the number of errors observed
+and the number of tests executed.
+
+Note that in all of the following land-to-land comparisons only lands
+are compared that are suitable for a comparison, i.e.~a signature can be
+extracted from the scan. In particular, lands which exhibited tank rash
+were removed from comparison \citep{aoas}.
 
@@
 
+\begin{table}[hbtp]
+\centering
+\begin{adjustbox}{max width=\textwidth}
+\begin{tabular}{rrrlrrl}
+  \hline
+  & \multicolumn{3}{l}{\bf CS1} & \multicolumn{3}{l}{\bf CS2} \\
+$w_o$ & \bf FPR & \bf FNR & \bf AUC (95\% C.I.) & \bf FPR & \bf FNR & \bf AUC (95\% C.I.) \\  
+  \hline
+%90 & 0.077 & 0.546 & 0.680 (0.633, 0.727) & 0.056 & 0.478 & 0.729 (0.684, 0.774) \\
+%  120 & 0.082 & 0.485 & 0.736 (0.693, 0.779) & 0.056 & 0.452 & 0.763 (0.720, 0.806) \\
+%  150 & 0.080 & 0.493 & 0.730 (0.685, 0.775) & 0.055 & 0.422 & 0.774 (0.733, 0.816) \\
+%  180 & 0.087 & 0.458 & 0.773 (0.734, 0.813) & 0.055 & 0.435 & 0.773 (0.730, 0.816) \\
+  210 & 0.086 & 0.413 & 0.780 (0.740, 0.821) & 0.055 & 0.387 & 0.803 (0.763, 0.843) \\ 
+  240 & 0.085 & 0.427 & 0.784 (0.744, 0.824) & 0.055 & 0.365 & 0.801 (0.760, 0.841) \\ 
+  270 & 0.082 & 0.415 & 0.782 (0.741, 0.823) & 0.048 & 0.374 & 0.808 (0.769, 0.848) \\ 
+  300 & 0.082 & 0.405 & 0.791 (0.751, 0.832) & 0.051 & 0.383 & 0.814 (0.776, 0.852) \\ 
+  330 & 0.082 & 0.428 & 0.791 (0.753, 0.830) & 0.050 & 0.383 & 0.804 (0.765, 0.844) \\ 
+  360 & 0.083 & 0.427 & 0.788 (0.747, 0.828) & 0.049 & 0.400 & 0.803 (0.765, 0.842) \\ 
+  390 & 0.085 & 0.400 & 0.789 (0.748, 0.830) & 0.050 & 0.383 & 0.810 (0.771, 0.849) \\ 
+   \hline
+\end{tabular}
+\end{adjustbox}
+\caption{\label{tab:aucsh44} Overview of high resolution scan results as shown in Figures \ref{fig:h44}.
+FPR is the observed rate of false positives (for a fixed level $\alpha = 0.05$), FNR is the false negative rate.
+Area under the curve (AUC) is shown with 95\% confidence intervals.
+}
+\end{table}
+
 \hypertarget{conclusions}{%
@@
 Figures \ref{same-shift-failure} and \ref{diff-shift-failure} show scenarios in which the deterministic Chumbley score can fail. 
-In \autoref{same-shift-failure} both CS1 and CS2 fail, because the lag between optimal locations is so large that no same-shift pairs can be found once the signatures are aligned.
-These failures are inherent to the Chumbley Score and cannot be prevented.
-\begin{figure}[hbtp]
-\centering
-\includegraphics[width=0.62\textwidth]{images/same-shift-failure.png}
-
-\caption{\label{same-shift-failure}Sketch of same-shift pairings.
-When the lag between optimal reference points is too large to accommodate a validation window in both signatures, both CS1 and CS2 fail.}
-\end{figure}
-
 \autoref{diff-shift-failure} shows two situations, in which CS1 fails to identify any different-shift pairings.
@@
 \caption{\label{diff-shift-failure}Sketch of different-shift pairings. For CS1 no different-shift pairs can be identified, resulting in a failed test.}
+\end{figure}
+
+In \autoref{same-shift-failure} both CS1 and CS2 fail, because the lag between optimal locations is so large that no same-shift pairs can be found once the signatures are aligned.
+These failures are inherent to the Chumbley Score and cannot be prevented.
+\begin{figure}[hbtp]
+\centering
+\includegraphics[width=0.62\textwidth]{images/same-shift-failure.png}
+
+\caption{\label{same-shift-failure}Sketch of same-shift pairings.
+When the lag between optimal reference points is too large to accommodate a validation window in both signatures, both CS1 and CS2 fail.}
 \end{figure}

### 2018-08-19 doc/Response.Rmd
resolution of images

@@
 You are right, we have added more information about the data, wherever available.
-We are also citing xxx to emphasize the difficulty level and resulting differences in error rates of different brands of firearms and ammunition*
+We have also added a sentence to emphasize different difficulty levels and resulting differences in error rates of different brands of firearms and ammunition*
 

### 2018-08-19 doc/paper.Rmd
resolution of images

@@
   thank NIST for providing the Hamby-44 bullets to us for scanning.
+  We would also like to thank the two anonymous reviewers for their helpful suggestions in improving the paper.
 abstract: The same-source problem remains a major challenge in forensic toolmark and
@@
                   rocs2 %>% ungroup() %>% dplyr::select(actual, beta, aucplus))
+rocs_tab[,c(2,3,5,6)] <- rocs_tab[,c(2,3,5,6)]*100
 dt <- xtable(rocs_tab, 
-             digits=c(0,0,3, 3,0,3,3,0))
+             digits=c(0,0,1, 1,0,1,1,0))
 #print(dt, include.rownames = FALSE)
@@
   & \multicolumn{3}{l}{\bf CS1} & \multicolumn{3}{l}{\bf CS2} \\
-$w_o$ & \bf FPR & \bf FNR & \bf AUC (95\% C.I.) & \bf FPR & \bf FNR & \bf AUC (95\% C.I.) \\ 
+$w_o$ & \bf \%FP & \bf \%FN & \bf AUC (95\% C.I.) & \bf \%FP & \bf \%FN & \bf AUC (95\% C.I.) \\ 
   \hline
-90 & 0.068 & 0.330 & 0.850 (0.835, 0.865) & 0.075 & 0.242 & 0.877 (0.863, 0.890) \\ 
-  120 & 0.065 & 0.309 & 0.864 (0.850, 0.878) & 0.072 & 0.236 & 0.890 (0.877, 0.903) \\ 
-  130 & 0.065 & 0.317 & 0.863 (0.850, 0.877) & 0.072 & 0.224 & 0.898 (0.886, 0.910) \\ 
-  140 & 0.065 & 0.304 & 0.873 (0.859, 0.886) & 0.071 & 0.229 & 0.902 (0.891, 0.914) \\ 
-  150 & 0.067 & 0.326 & 0.863 (0.850, 0.877) & 0.072 & 0.225 & 0.904 (0.892, 0.916) \\ 
-  180 & 0.061 & 0.326 & 0.877 (0.864, 0.890) & 0.070 & 0.229 & 0.907 (0.896, 0.919) \\ 
-  210 & 0.064 & 0.342 & 0.865 (0.851, 0.878) & 0.066 & 0.231 & 0.906 (0.895, 0.918) \\ 
-   \hline
+90 & 6.8 & 33.0 & 0.850 (0.835, 0.865) & 7.5 & 24.2 & 0.877 (0.863, 0.890) \\ 
+  120 & 6.5 & 30.9 & 0.864 (0.850, 0.878) & 7.2 & 23.6 & 0.890 (0.877, 0.903) \\ 
+  130 & 6.5 & 31.7 & 0.863 (0.850, 0.877) & 7.2 & 22.4 & 0.898 (0.886, 0.910) \\ 
+  140 & 6.5 & 30.4 & 0.873 (0.859, 0.886) & 7.1 & 22.9 & 0.902 (0.891, 0.914) \\ 
+  150 & 6.7 & 32.6 & 0.863 (0.850, 0.877) & 7.2 & 22.5 & 0.904 (0.892, 0.916) \\ 
+  180 & 6.1 & 32.6 & 0.877 (0.864, 0.890) & 7.0 & 22.9 & 0.907 (0.896, 0.919) \\ 
+  210 & 6.4 & 34.2 & 0.865 (0.851, 0.878) & 6.6 & 23.1 & 0.906 (0.895, 0.918) \\    \hline
 \end{tabular}
@@
 \caption{\label{tab:aucs} Overview of results as shown in Figures \ref{fig:roc} and \ref{fig:eer}.
-FPR is the observed rate of false positives (for a fixed level $\alpha = 0.05$), FNR is the false negative rate.
+\%FP is the observed percentage of false positives (for a fixed level $\alpha = 0.05$), \%FN is the percentage of false negatives.
 Area under the curve (AUC) is shown with confidence intervals.
@@
 
-\autoref{fig:type1} shows observed type I error rates (FPR) across a range of optimization windows $w_o$.
+\autoref{fig:type1} shows the percentages of observed type I errors (\%FP) across a range of optimization windows $w_o$.
 Generally, observed type I errors are higher than expected.
-Method CS1 shows in this instance slightly better performance than method CS2, but for both an increase in the size of the optimization window leads to a decrease in the observed type I error rates.
+Method CS1 shows in this instance slightly better performance than method CS2, but for both an increase in the size of the optimization window leads to a decrease in the observed type I errors.
 
@@
 Both methods have a failed test rate of less than 0.006, indicating, again, that the larger number of pixels alleviates the problem of test failures.
-\autoref{fig:h44} shows the resulting EER and AUC for methods CS1 and CS2 based on two sizes of validation windows ($w_v \in 75, 125$) and optimization window sizes around 300 pixels (10 percent of the average length).
+\autoref{fig:h44} shows the resulting EER and AUC for methods CS1 and CS2 based on two sizes of validation windows ($w_v \in 75, 125$) and optimization window sizes around 300 pixels (10 percent of the average length), with corresponding errors shown numerically in \autoref{tab:aucsh44}.
 Both methods show an increase in performance around $w_o = 300$ pixels.
@@
 Partly, this might be due to the particular choice of the parameters, partly the higher-resolution scans might be picking up on real differences between the lands that the lower-resolution scans fail to detect.
+Re-assuringly, the observed error rate of false positives for CS2 is closer to the nominal rate of 5\% than for the lower-resolution scans.
 
@@
 
-```{r, fig.cap="h44 Overview of all of the results."}
+```{r, fig.cap="Overview of all of the results."}
 colnames(h44)[11]<- "coarse"
@@
                        rocs2.44 %>% ungroup() %>% dplyr::select(actual, beta, aucplus))
+rocs44_tab[, c(2,3,5,6)] <- rocs44_tab[, c(2,3,5,6)]*100
 dt44 <- xtable(rocs44_tab, 
-             digits=c(0,0,3, 3,0,3,3,0))
+             digits=c(0,0,1,1,0,1,1,0))
 #print(dt44, include.rownames = FALSE)
@@
   & \multicolumn{3}{l}{\bf CS1} & \multicolumn{3}{l}{\bf CS2} \\
-$w_o$ & \bf FPR & \bf FNR & \bf AUC (95\% C.I.) & \bf FPR & \bf FNR & \bf AUC (95\% C.I.) \\  
+$w_o$ & \bf \%FP & \bf \%FN & \bf AUC (95\% C.I.) & \bf \%FP & \bf \%FN & \bf AUC (95\% C.I.) \\  
   \hline
-%90 & 0.077 & 0.546 & 0.680 (0.633, 0.727) & 0.056 & 0.478 & 0.729 (0.684, 0.774) \\
-%  120 & 0.082 & 0.485 & 0.736 (0.693, 0.779) & 0.056 & 0.452 & 0.763 (0.720, 0.806) \\
-%  150 & 0.080 & 0.493 & 0.730 (0.685, 0.775) & 0.055 & 0.422 & 0.774 (0.733, 0.816) \\
-%  180 & 0.087 & 0.458 & 0.773 (0.734, 0.813) & 0.055 & 0.435 & 0.773 (0.730, 0.816) \\
-  210 & 0.086 & 0.413 & 0.780 (0.740, 0.821) & 0.055 & 0.387 & 0.803 (0.763, 0.843) \\ 
-  240 & 0.085 & 0.427 & 0.784 (0.744, 0.824) & 0.055 & 0.365 & 0.801 (0.760, 0.841) \\ 
-  270 & 0.082 & 0.415 & 0.782 (0.741, 0.823) & 0.048 & 0.374 & 0.808 (0.769, 0.848) \\ 
-  300 & 0.082 & 0.405 & 0.791 (0.751, 0.832) & 0.051 & 0.383 & 0.814 (0.776, 0.852) \\ 
-  330 & 0.082 & 0.428 & 0.791 (0.753, 0.830) & 0.050 & 0.383 & 0.804 (0.765, 0.844) \\ 
-  360 & 0.083 & 0.427 & 0.788 (0.747, 0.828) & 0.049 & 0.400 & 0.803 (0.765, 0.842) \\ 
-  390 & 0.085 & 0.400 & 0.789 (0.748, 0.830) & 0.050 & 0.383 & 0.810 (0.771, 0.849) \\ 
+  210 & 8.6 & 41.3 & 0.780 (0.740, 0.821) & 5.5 & 38.7 & 0.803 (0.763, 0.843) \\ 
+  240 & 8.5 & 42.7 & 0.784 (0.744, 0.824) & 5.5 & 36.5 & 0.801 (0.760, 0.841) \\ 
+  270 & 8.2 & 41.5 & 0.782 (0.741, 0.823) & 4.8 & 37.4 & 0.808 (0.769, 0.848) \\ 
+  300 & 8.2 & 40.5 & 0.791 (0.751, 0.832) & 5.1 & 38.3 & 0.814 (0.776, 0.852) \\ 
+  330 & 8.2 & 42.8 & 0.791 (0.753, 0.830) & 5.0 & 38.3 & 0.804 (0.765, 0.844) \\ 
+  360 & 8.3 & 42.7 & 0.788 (0.747, 0.828) & 4.9 & 40.0 & 0.803 (0.765, 0.842) \\
+  390 & 8.5 & 40.0 & 0.789 (0.748, 0.830) & 5.0 & 38.3 & 0.810 (0.771, 0.849) \\ 
    \hline
@@
 \caption{\label{tab:aucsh44} Overview of high resolution scan results as shown in Figures \ref{fig:h44}.
-FPR is the observed rate of false positives (for a fixed level $\alpha = 0.05$), FNR is the false negative rate.
+\%FP is the percentage of observed  false positives (for a fixed level $\alpha = 0.05$), \%FN is the percentage of false negatives.
 Area under the curve (AUC) is shown with 95\% confidence intervals.
@@
 Sizes for optimization windows $w_o$ were based on cross-validation to minimize overall type 2 error rates. 
-Unfortunately, this kind of assessment is only feasible in the setting of a large study, such as the one we presented, and does not transfer immediately to case work, where a forensic examiner would only deal with a few identifications.
-Parameter settings for different studies should be investigated and reported to allow for a further refinement.
+Ideally, the exact values for parameters will be determined in a large study incorporating different types of firearms and brands of ammunition. 
+%Unfortunately, this kind of assessment is only feasible in the setting of a large study, such as the one we presented, 
+Results therefore do not transfer immediately to case work, where a forensic examiner would only deal with a few identifications.
+%Parameter settings for different studies should be investigated and reported to allow for a further refinement.
 
@@
  Algorithm CS2 is introduced here as a remedy for failed tests by introducing an alternate version of choosing different-shift pairs. 
-The algorithm CS2 introduced here is constructed in a way that achieves on average a ten-fold reduction in the number of failures.
+Algorithm CS2 is constructed in a way that achieves on average a ten-fold reduction in the number of failures.
 While reducing the failure rate, the algorithm also shows an increase in the power of the test.
@@
 
-However, bullets usually have multiple lands -- in the case of Ruger P85s as used in the Hamby study, there are six lands for each bullet.
-We might be able to get more power out of the test by adapting CS2 to work in a bullet-to-bullet comparison.
+Different avenues for improving the performance of the Chumbley-score are still open: 
+(i) When making the original Chumbley Score deterministic, \citet{hadler} rely on an optimal shift based on the maximum cross-correlation between two markings.
+An erroneous decision in  identifying the optimial shift between markings leads almost always to a missed identification.
+CS2 is still susceptible to this source of errors.
+(ii) Bullets usually have multiple lands -- in the case of Ruger P85s as used in the Hamby study, there are six lands for each bullet.
+As shown by \citet{chu2010} in the example of the cross-correlation between lands, we will be able to get more power out of the test by adapting the algorithm to make use of the relative position of a land on the bullet to combine multiple land-to-land comparisons into a single bullet-to-bullet comparison.
 

### 2018-08-19 doc/paper.tex
resolution of images

@@
   also like to thank NIST for providing the Hamby-44 bullets to us for
-  scanning.\\
+  scanning. We would also like to thank the two anonymous reviewers for
+  their helpful suggestions in improving the paper.\\
   \bigskip
@@
 
-Methods for matching {\textcolor{blue}{striated}} marks for a variety of
-tools have been studied in the literature (see
-\autoref{tab:toolmarks-ER} for an overview): in \cite{manytoolmarks1}
-and \cite{chumbley} digitized screwdriver marks have been analyzed using
-a profilometer; in \cite{manytoolmarks2} 3D marks from screwdriver,
-tongue and groove pliers captured using a confocal microscope have been
-investigated; digitized marks from slip-joint pliers generated by a
-surface profilometer have been investigated in \cite{afte-chumbley}.
-
-{\textcolor{blue}{changed}} These digitized markings allow the use of
-statistical methods to quantify the scientific mechanism of comparing
-markings and serve as basis for an error rate calculation. In
-\cite{manytoolmarks2} a relative distance metric is defined and used as
-similarity measure between two toolmarks. This approach is expanded in
-\citet{manytoolmarks1}: a set of small segments in the markings of two
-toolmarks are extracted and similarity is compared using a maximum
-Pearson correlation coefficient. The Chumbley scoring method, first
-introduced in \cite{chumbley}, uses a similar but more extensive
-framework based on a Mann-Whitney U test of the resulting correlation
-coefficients. This approach is non-deterministic, because segments are
-chosen randomly. In \cite{hadler} the score is fixed to be deterministic
-for each pair of toolmarks by choosing segments for comparison
-systematically. This approach also ensures independence between segments
-of striae.
+Methods for matching striated marks for a variety of tools have been
+studied in the literature (see \autoref{tab:toolmarks-ER} for an
+overview): in \cite{manytoolmarks1} and \cite{chumbley} digitized
+screwdriver marks have been analyzed using a profilometer; in
+\cite{manytoolmarks2} 3D marks from screwdriver, tongue and groove
+pliers captured using a confocal microscope have been investigated;
+digitized marks from slip-joint pliers generated by a surface
+profilometer have been investigated in \cite{afte-chumbley}.
+
+These digitized markings allow the use of statistical methods to
+quantify the scientific mechanism of comparing markings and serve as
+basis for an error rate calculation. In \cite{manytoolmarks2} a relative
+distance metric is defined and used as similarity measure between two
+toolmarks. This approach is expanded in \citet{manytoolmarks1}: a set of
+small segments in the markings of two toolmarks are extracted and
+similarity is compared using a maximum Pearson correlation coefficient.
+The Chumbley scoring method, first introduced in \cite{chumbley}, uses a
+similar but more extensive framework based on a Mann-Whitney U test of
+the resulting correlation coefficients. This approach is
+non-deterministic, because segments are chosen randomly. In
+\cite{hadler} the score is fixed to be deterministic for each pair of
+toolmarks by choosing segments for comparison systematically. This
+approach also ensures independence between segments of striae.
 
@@
 \end{tabular}}
-\caption{\label{tab:toolmarks-ER} Error Rates in same-source {\textcolor{blue}{striated}} toolmark analysis reported in the literature.
-The top four reported papers use some variation of the Chumbley-score method.}
+\caption{\label{tab:toolmarks-ER} Error Rates in same-source analyses for  striated toolmarks reported in the literature. The top four reported papers use some variation of the Chumbley score method.}
 \end{table}
@@
   & \multicolumn{3}{l}{\bf CS1} & \multicolumn{3}{l}{\bf CS2} \\
-$w_o$ & \bf FPR & \bf FNR & \bf AUC (95\% C.I.) & \bf FPR & \bf FNR & \bf AUC (95\% C.I.) \\ 
-  \hline
-90 & 0.068 & 0.330 & 0.850 (0.835, 0.865) & 0.075 & 0.242 & 0.877 (0.863, 0.890) \\ 
-  120 & 0.065 & 0.309 & 0.864 (0.850, 0.878) & 0.072 & 0.236 & 0.890 (0.877, 0.903) \\ 
-  130 & 0.065 & 0.317 & 0.863 (0.850, 0.877) & 0.072 & 0.224 & 0.898 (0.886, 0.910) \\ 
-  140 & 0.065 & 0.304 & 0.873 (0.859, 0.886) & 0.071 & 0.229 & 0.902 (0.891, 0.914) \\ 
-  150 & 0.067 & 0.326 & 0.863 (0.850, 0.877) & 0.072 & 0.225 & 0.904 (0.892, 0.916) \\ 
-  180 & 0.061 & 0.326 & 0.877 (0.864, 0.890) & 0.070 & 0.229 & 0.907 (0.896, 0.919) \\ 
-  210 & 0.064 & 0.342 & 0.865 (0.851, 0.878) & 0.066 & 0.231 & 0.906 (0.895, 0.918) \\ 
-   \hline
+$w_o$ & \bf \%FP & \bf \%FN & \bf AUC (95\% C.I.) & \bf \%FP & \bf \%FN & \bf AUC (95\% C.I.) \\ 
+  \hline
+90 & 6.8 & 33.0 & 0.850 (0.835, 0.865) & 7.5 & 24.2 & 0.877 (0.863, 0.890) \\ 
+  120 & 6.5 & 30.9 & 0.864 (0.850, 0.878) & 7.2 & 23.6 & 0.890 (0.877, 0.903) \\ 
+  130 & 6.5 & 31.7 & 0.863 (0.850, 0.877) & 7.2 & 22.4 & 0.898 (0.886, 0.910) \\ 
+  140 & 6.5 & 30.4 & 0.873 (0.859, 0.886) & 7.1 & 22.9 & 0.902 (0.891, 0.914) \\ 
+  150 & 6.7 & 32.6 & 0.863 (0.850, 0.877) & 7.2 & 22.5 & 0.904 (0.892, 0.916) \\ 
+  180 & 6.1 & 32.6 & 0.877 (0.864, 0.890) & 7.0 & 22.9 & 0.907 (0.896, 0.919) \\ 
+  210 & 6.4 & 34.2 & 0.865 (0.851, 0.878) & 6.6 & 23.1 & 0.906 (0.895, 0.918) \\    \hline
 \end{tabular}
@@
 
-\autoref{fig:type1} shows observed type I error rates (FPR) across a
-range of optimization windows \(w_o\). Generally, observed type I errors
-are higher than expected. Method CS1 shows in this instance slightly
-better performance than method CS2, but for both an increase in the size
-of the optimization window leads to a decrease in the observed type I
-error rates.
+\autoref{fig:type1} shows the percentages of observed type I errors
+(\%FP) across a range of optimization windows \(w_o\). Generally,
+observed type I errors are higher than expected. Method CS1 shows in
+this instance slightly better performance than method CS2, but for both
+an increase in the size of the optimization window leads to a decrease
+in the observed type I errors.
 
@@
 sizes of validation windows (\(w_v \in 75, 125\)) and optimization
-window sizes around 300 pixels (10 percent of the average length). Both
+window sizes around 300 pixels (10 percent of the average length), with
+corresponding errors shown numerically in \autoref{tab:aucsh44}. Both
 methods show an increase in performance around \(w_o = 300\) pixels. CS2
@@
 might be picking up on real differences between the lands that the
-lower-resolution scans fail to detect.
+lower-resolution scans fail to detect. Re-assuringly, the observed error
+rate of false positives for CS2 is closer to the nominal rate of 5\%
+than for the lower-resolution scans.
 
@@
   & \multicolumn{3}{l}{\bf CS1} & \multicolumn{3}{l}{\bf CS2} \\
-$w_o$ & \bf FPR & \bf FNR & \bf AUC (95\% C.I.) & \bf FPR & \bf FNR & \bf AUC (95\% C.I.) \\  
-  \hline
-%90 & 0.077 & 0.546 & 0.680 (0.633, 0.727) & 0.056 & 0.478 & 0.729 (0.684, 0.774) \\
-%  120 & 0.082 & 0.485 & 0.736 (0.693, 0.779) & 0.056 & 0.452 & 0.763 (0.720, 0.806) \\
-%  150 & 0.080 & 0.493 & 0.730 (0.685, 0.775) & 0.055 & 0.422 & 0.774 (0.733, 0.816) \\
-%  180 & 0.087 & 0.458 & 0.773 (0.734, 0.813) & 0.055 & 0.435 & 0.773 (0.730, 0.816) \\
-  210 & 0.086 & 0.413 & 0.780 (0.740, 0.821) & 0.055 & 0.387 & 0.803 (0.763, 0.843) \\ 
-  240 & 0.085 & 0.427 & 0.784 (0.744, 0.824) & 0.055 & 0.365 & 0.801 (0.760, 0.841) \\ 
-  270 & 0.082 & 0.415 & 0.782 (0.741, 0.823) & 0.048 & 0.374 & 0.808 (0.769, 0.848) \\ 
-  300 & 0.082 & 0.405 & 0.791 (0.751, 0.832) & 0.051 & 0.383 & 0.814 (0.776, 0.852) \\ 
-  330 & 0.082 & 0.428 & 0.791 (0.753, 0.830) & 0.050 & 0.383 & 0.804 (0.765, 0.844) \\ 
-  360 & 0.083 & 0.427 & 0.788 (0.747, 0.828) & 0.049 & 0.400 & 0.803 (0.765, 0.842) \\ 
-  390 & 0.085 & 0.400 & 0.789 (0.748, 0.830) & 0.050 & 0.383 & 0.810 (0.771, 0.849) \\ 
+$w_o$ & \bf \%FP & \bf \%FN & \bf AUC (95\% C.I.) & \bf \%FP & \bf \%FN & \bf AUC (95\% C.I.) \\  
+  \hline
+  210 & 8.6 & 41.3 & 0.780 (0.740, 0.821) & 5.5 & 38.7 & 0.803 (0.763, 0.843) \\ 
+  240 & 8.5 & 42.7 & 0.784 (0.744, 0.824) & 5.5 & 36.5 & 0.801 (0.760, 0.841) \\ 
+  270 & 8.2 & 41.5 & 0.782 (0.741, 0.823) & 4.8 & 37.4 & 0.808 (0.769, 0.848) \\ 
+  300 & 8.2 & 40.5 & 0.791 (0.751, 0.832) & 5.1 & 38.3 & 0.814 (0.776, 0.852) \\ 
+  330 & 8.2 & 42.8 & 0.791 (0.753, 0.830) & 5.0 & 38.3 & 0.804 (0.765, 0.844) \\ 
+  360 & 8.3 & 42.7 & 0.788 (0.747, 0.828) & 4.9 & 40.0 & 0.803 (0.765, 0.842) \\
+  390 & 8.5 & 40.0 & 0.789 (0.748, 0.830) & 5.0 & 38.3 & 0.810 (0.771, 0.849) \\ 
    \hline
@@
 CSAFE. Sizes for optimization windows \(w_o\) were based on
-cross-validation to minimize overall type 2 error rates. Unfortunately,
-this kind of assessment is only feasible in the setting of a large
-study, such as the one we presented, and does not transfer immediately
-to case work, where a forensic examiner would only deal with a few
-identifications. Parameter settings for different studies should be
-investigated and reported to allow for a further refinement.
+cross-validation to minimize overall type 2 error rates. Ideally, the
+exact values for parameters will be determined in a large study
+incorporating different types of firearms and brands of ammunition.
+\%Unfortunately, this kind of assessment is only feasible in the setting
+of a large study, such as the one we presented, Results therefore do not
+transfer immediately to case work, where a forensic examiner would only
+deal with a few identifications. \%Parameter settings for different
+studies should be investigated and reported to allow for a further
+refinement.
 
@@
 introduced here as a remedy for failed tests by introducing an alternate
-version of choosing different-shift pairs. The algorithm CS2 introduced
-here is constructed in a way that achieves on average a ten-fold
-reduction in the number of failures. While reducing the failure rate,
-the algorithm also shows an increase in the power of the test. Type II
-error rates of CS2 reach a minimum of 0.217 for an optimized window size
-of 130 pixels. This increase in power of CS2 over CS1 should also apply
-to previous studies on toolmarks. It would be interesting to see these
-results using the adjusted algorithm. Unfortunately, none of the studies
-have made the data publicly accessible.
+version of choosing different-shift pairs. Algorithm CS2 is constructed
+in a way that achieves on average a ten-fold reduction in the number of
+failures. While reducing the failure rate, the algorithm also shows an
+increase in the power of the test. Type II error rates of CS2 reach a
+minimum of 0.217 for an optimized window size of 130 pixels. This
+increase in power of CS2 over CS1 should also apply to previous studies
+on toolmarks. It would be interesting to see these results using the
+adjusted algorithm. Unfortunately, none of the studies have made the
+data publicly accessible.
 
@@
 
-However, bullets usually have multiple lands -- in the case of Ruger
-P85s as used in the Hamby study, there are six lands for each bullet. We
-might be able to get more power out of the test by adapting CS2 to work
-in a bullet-to-bullet comparison.
+Different avenues for improving the performance of the Chumbley-score
+are still open: (i) When making the original Chumbley Score
+deterministic, \citet{hadler} rely on an optimal shift based on the
+maximum cross-correlation between two markings. An erroneous decision in
+identifying the optimial shift between markings leads almost always to a
+missed identification. CS2 is still susceptible to this source of
+errors. (ii) Bullets usually have multiple lands -- in the case of Ruger
+P85s as used in the Hamby study, there are six lands for each bullet. As
+shown by \citet{chu2010} in the example of the cross-correlation between
+lands, we will be able to get more power out of the test by adapting the
+algorithm to make use of the relative position of a land on the bullet
+to combine multiple land-to-land comparisons into a single
+bullet-to-bullet comparison.
 

### 2018-08-19 doc/Response.Rmd
last couple of comments addressed

@@
 The authors do present AUC and EER curves but these are less informative and can often be used to hide weaker performing models.
+
+*We have added a table with error rates for the Hamby 44 scans.*
+
 - The results should all be listed as either fractions or all as percent (which is preferred) and not mix them.
@@
 
-*We agree the data available to us is by no means large. Our comment was referring to the availability of 35 bullets rater than just 2 0r 3 as might be seen in actual case work*
+*We agree the data available to us is by no means large.
+Our comment was referring to the availability of 35 bullets rater than just 2 0r 3 as might be seen in actual case work.
+We have changed the sentence accordingly.*
 

### 2018-08-19 doc/paper.Rmd
last couple of comments addressed

@@
 \centering
-\includegraphics[width=\textwidth]{images/B6-B2-L6-rescaled.pdf}
+\includegraphics[width=\textwidth]{images/B6-B2-L6-rescaled.png}
 ```{r, eval=FALSE, echo=FALSE, fig.width=8, fig.height=3, out.width='\\textwidth'}
@@
 A $p$-value for the Chumbley score test is  then computed as a Mann Whitney U statistic to compare between in-phase sample and out-of-phase sample.
-A low $p$-value is  interpreted as a rejection of the null hypothesis in favor of the alternative hypothesis of `same-source`.
+A low $p$-value is  interpreted as a rejection of the null hypothesis in favor of the alternative hypothesis of *same-source*.
 
@@
 
-We then define the pairs for different-shifts as
+\begin{figure}[hbtp]
+\centering
+\includegraphics[width=.7\textwidth]{images/sketch-diff-2.png}
+\caption{\label{sketch-diff-2}Sketch of adjusted different-shift pairings.
+Pairings are side from opposite ends of the two markings.
+For an odd number of pairings the middle pair is left out.
+At most one of the same-shift pairings can not be matched with a different-shift pair under algorithm (CS2).
+}
+\end{figure}
+
+We then define the pairs for different-shifts by matching up windows from opposite ends of the markings, i.e.
+the first pair consists of a matchup of the first window on the first marking and the last window on the second marking, the second pair consists of the second window on the first marking and the second to last pair on the second marking, and so on.
+In case of an odd number of pairs we need to be careful to exclude the middle pair from this assignment:  the middle pair is already part of the same-shift pair, therefore we cannot re-use the same pair as part of the different-shift pairs.
+
+Mathematically,  this assignment of pairs is written as: 
 \begin{equation}\label{eq.diff2}
@@
 1, ..., (I-1)/2, (I-1)/2 + 2, ..., I & \text{ for odd } I
-\end{cases},
+\end{cases}.
 \end{equation}
-i.e.\ for an odd number of same-shift correlations, we skip the middle pair for the different-shift correlations (see \autoref{sketch-diff-2}).
+Note that for an odd number of same-shift correlations, we skip the middle pair for the different-shift correlations (see also \autoref{sketch-diff-2}).
 This pairing ensures that the number of different-shift pairings is the same or at most one less than the number of same-shift pairings in all tests.
@@
 In particular, lands which exhibited tank rash  were removed from comparison \citep{aoas}.
-
-\begin{figure}[hbtp]
-\centering
-\includegraphics[width=.7\textwidth]{images/sketch-diff-2.png}
-\caption{\label{sketch-diff-2}Sketch of adjusted different-shift pairings. At most one of the same-shift pairings can not be matched with a different-shift pair under algorithm (CS2). }
-\end{figure}
 
@@
 
-\newpage
 \subsection{Type 2 errors: CS1 vs CS2}

### 2018-08-19 doc/paper.tex
last couple of comments addressed

@@
 %\usepackage[numbers]{natbib}
+\usepackage[nomarkers]{endfloat}
 \usepackage{natbib}
@@
 \centering
-\includegraphics[width=\textwidth]{images/B6-B2-L6-rescaled.pdf}
+\includegraphics[width=\textwidth]{images/B6-B2-L6-rescaled.png}
 
@@
 interpreted as a rejection of the null hypothesis in favor of the
-alternative hypothesis of \texttt{same-source}.
+alternative hypothesis of \emph{same-source}.
 
@@
 
-We then define the pairs for different-shifts as
+\begin{figure}[hbtp]
+\centering
+\includegraphics[width=.7\textwidth]{images/sketch-diff-2.png}
+\caption{\label{sketch-diff-2}Sketch of adjusted different-shift pairings.
+Pairings are side from opposite ends of the two markings.
+For an odd number of pairings the middle pair is left out.
+At most one of the same-shift pairings can not be matched with a different-shift pair under algorithm (CS2).
+}
+\end{figure}
+
+We then define the pairs for different-shifts by matching up windows
+from opposite ends of the markings, i.e.~the first pair consists of a
+matchup of the first window on the first marking and the last window on
+the second marking, the second pair consists of the second window on the
+first marking and the second to last pair on the second marking, and so
+on. In case of an odd number of pairs we need to be careful to exclude
+the middle pair from this assignment: the middle pair is already part of
+the same-shift pair, therefore we cannot re-use the same pair as part of
+the different-shift pairs.
+
+Mathematically, this assignment of pairs is written as:
 \begin{equation}\label{eq.diff2}
@@
 1, ..., (I-1)/2, (I-1)/2 + 2, ..., I & \text{ for odd } I
-\end{cases},
-\end{equation} i.e.~for an odd number of same-shift correlations, we
-skip the middle pair for the different-shift correlations (see
+\end{cases}.
+\end{equation} Note that for an odd number of same-shift correlations,
+we skip the middle pair for the different-shift correlations (see also
 \autoref{sketch-diff-2}). This pairing ensures that the number of
@@
 were removed from comparison \citep{aoas}.
-
-\begin{figure}[hbtp]
-\centering
-\includegraphics[width=.7\textwidth]{images/sketch-diff-2.png}
-\caption{\label{sketch-diff-2}Sketch of adjusted different-shift pairings. At most one of the same-shift pairings can not be matched with a different-shift pair under algorithm (CS2). }
-\end{figure}
 

### 2018-08-22 doc/abstract.Rmd
abstract separated out

---
title: "Abstract"
author: "Heike Hofmann"
date: "8/22/2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background

The same-source problem remains a major challenge in forensic toolmark and
  firearm examination. Technological advances in surface imaging allow measurements
  of 3D surfaces at previously unforeseen resolutions and enable digitized imaging.
  Here, we investigate the applicability of the Chumbley scoring method (Chumbley et al 2010, Hadler et al 2018),
  developed for screwdriver markings, for same-source identification of  striation marks on bullet land engraved areas (LEAs).

## Brief description of methods and results

3D scans of bullets from Hamby sets 44 and 252 measured by the National Institute of Standards and Technology (NIST) and the Center for Statistical Application and Forensic Evidence (CSAFE) were used as a basis for the evaluation.  
In the adaption of the Chumbley Scoring method, we suggest methods to identify parameters that minimize the error rates. We also suggest a remedial algorithm to alleviate the problem of failed tests in the method, increasing both the power of the test and reducing error rates.

Out of a total of 85,491 land-to-land comparisons (comprised of 84,235 known non-matches and 1256 known matches), the adapted test does not provide a result in 176 situations (down from more than 500). The Type I error rate is 7.2% (6105 out of 84235 known non-matches), the Type II error rate is 21.4% (271 out of 1256 known matches).

## Conclusions

This puts the proposed method on similar footing as other single
  feature matching approaches in the literature.

### 2018-09-17 jfs_submission/paper-submission-original/paper-title.tex
re-arranging the files

--- 2018-08-19 doc/paper.tex
@@
 \usepackage{enumerate}
-%\usepackage[numbers]{natbib}
 \usepackage{natbib}
-\setcitestyle{numbers} 
-\makeatletter % Reference list option change
-\renewcommand\@biblabel[1]{#1. } % from [1] to 1
-\makeatother %
-
 \usepackage{booktabs}
@@
 \usepackage{subfig}
-\usepackage[table,xcdraw]{xcolor}
+\usepackage[table]{xcolor}
 \usepackage{wrapfig}
 \usepackage{float}
-%\usepackage{colortbl}
-\usepackage[colorlinks]{hyperref}
-\hypersetup{
-  colorlinks=true,
-  citecolor=black,
-  linkcolor=black,
-  urlcolor=blue}
-
+\usepackage{colortbl}
+\usepackage{hyperref}
 \usepackage{pdflscape}
@@
 \usepackage{threeparttable}
+
 \usepackage{url} % not crucial - just used below for the URL
-
-\usepackage{etoolbox}% http://ctan.org/pkg/etoolbox
-\makeatletter
-\patchcmd{\subsection}{\bfseries}{\relax}{}{}% Non-bold \subsection
-\patchcmd{\subsubsection}{\bfseries}{\relax}{}{}% Non-bold \subsection
-\makeatother
 
@@
 {
-  \title{\bf Adapting the Chumbley Score to match striae on Land Engraved Areas of
-bullets}
+  \title{\bf Adapting the Chumbley Score to match striae on Land Engraved Areas (LEAs) of bullets}
 
@@
   \begin{center}
-  {
-  \textbf{Affiliation(s)}\\
+  {\textbf{Affiliation(s)}\\
   Department of Statistics, Iowa State University and CSAFE (Center for
@@
   Annual Training Seminar, held from June 3rd to June 8th, 2018 in
-  Charleston, West Virginia
-  }
+  Charleston, West Virginia}
   \end{center}
@@
 
-  % \bigskip
-
-  % \bigskip
-\newpage
-
-\if0\blind
-{
-  \bigskip
-  \bigskip
-  \bigskip
-  \begin{center}
-  \title{\LARGE\bf Adapting the Chumbley Score to match striae on Land Engraved Areas of
-bullets}
-  \end{center}
-  \bigskip
-  \maketitle
-} \fi
-
-\bigskip
-\begin{abstract}
-The same-source problem remains a major challenge in forensic toolmark
-and firearm examination. Technological advances in surface imaging allow
-measurements of 3D surfaces at previously unforeseen resolutions and
-enable digitized imaging. Here, we investigate the applicability of the
-Chumbley scoring method \cite{hadler}, developed for screwdriver
-markings, for same-source identification of bullet striae. We provide
-methods to identify parameters that minimize the error rates for
-matching of LEAs using Hamby datasets 44 and 252 measured by NIST and
-CSAFE. We suggest a remedial algorithm to alleviate the problem of
-failed tests in the method, increasing both the power of the test and
-reducing error rates. Type II error rates of the proposed method improve
-by more than one third (Type I error of 0.05) and are on average about
-0.22. This puts the proposed method on similar footing as other single
-feature matching approaches in the literature.
-\end{abstract}
-
-\noindent%
-{\it Keywords:} forensic science, toolmark, cross-correlation, Mann-Whitney U statistic, land engraved areas (LEAs), algorithm, signatures, same-source problem
-\vfill
-
-\newpage
-\spacingset{1.45} % DON'T change the spacing!
-
-\newcommand{\hh}[1]{{\textcolor{orange}{#1}}}
-\newcommand{\gk}[1]{{\textcolor{blue}{#1}}}
-\newcommand{\cited}[1]{{\textcolor{red}{#1}}}
-\setlength\parindent{0pt}
-
-\newpage
-
-Same-source analyses are a major part of a Forensic Toolmark Examiner's
-job. In current practice, examiners make these comparisons by visual
-inspection under a comparison microscope and come to one of the
-following four conclusions: identification, inconclusive, elimination or
-unsuitable for examination \cite{afte-toolmarks1998}. These conclusions
-are made on the basis of ``unique surface contours'' of the two
-toolmarks being in ``sufficient agreement'' \cite{afte-toolmarks1998}.
-AFTE describes the term ``sufficient agreement'' as the possibility of
-another tool producing the markings under comparison, as practically
-impossible \cite{afte-toolmarks1998}. Potential subject bias in the
-assessment as well as the lack of specified error rates are the main
-points of criticisms first raised by the National Research Council in
-2009 \cite{NAS:2009} and later emphasized further by the President's
-Council of Advisors on Science and Technology \cite{pcast2016}.
-{\textcolor{blue}{need a sentence on striated marks}}
-
-Technological advances, such as profilometers and confocal microscopy,
-have made it possible to capture 3D surfaces in a high-resolution
-digitized form. This technology has become more accessible over the last
-decade, and has made its way into topological images of ballistics
-evidence, such as bullet lands and breech faces
-\cite{DeKinder1, DeKinder2, Bachrach1, vorburger2016}. Digitized images
-of 3D surfaces form the basis of statistical analysis of toolmarks.
-Statistical approaches based on this data remove both subjectivity from
-the assessment and allow a quantification of error rates for false
-positive and false negative identifications.
-
-Methods for matching {\textcolor{blue}{striated}} marks for a variety of
-tools have been studied in the literature (see
-\autoref{tab:toolmarks-ER} for an overview): in \cite{manytoolmarks1}
-and \cite{chumbley} digitized screwdriver marks have been analyzed using
-a profilometer; in \cite{manytoolmarks2} 3D marks from screwdriver,
-tongue and groove pliers captured using a confocal microscope have been
-investigated; digitized marks from slip-joint pliers generated by a
-surface profilometer have been investigated in \cite{afte-chumbley}.
-
-{\textcolor{blue}{changed}} These digitized markings allow the use of
-statistical methods to quantify the scientific mechanism of comparing
-markings and serve as basis for an error rate calculation. In
-\cite{manytoolmarks2} a relative distance metric is defined and used as
-similarity measure between two toolmarks. This approach is expanded in
-\citet{manytoolmarks1}: a set of small segments in the markings of two
-toolmarks are extracted and similarity is compared using a maximum
-Pearson correlation coefficient. The Chumbley scoring method, first
-introduced in \cite{chumbley}, uses a similar but more extensive
-framework based on a Mann-Whitney U test of the resulting correlation
-coefficients. This approach is non-deterministic, because segments are
-chosen randomly. In \cite{hadler} the score is fixed to be deterministic
-for each pair of toolmarks by choosing segments for comparison
-systematically. This approach also ensures independence between segments
-of striae.
-
-\begin{table}[ht]
-\centering
-\resizebox{\textwidth}{!}{\begin{tabular}{rlrrr}
-  \hline
-Research paper & Method & Data Source & False Positives & False Negatives\\
-  \hline
- \textbf{Faden et al. (2007)} \cite{manytoolmarks1} & Maximum Pearson & Screwdrivers & & \\ & Correlation &  & - & - \\ \hline
- \textbf{Chumbley et al. (2010)} \cite{chumbley} & Randomized & Screwdrivers & & \\ (Same-Surface Same-Angle) & Chumbley Score & & 2.3\% & 8.9\%  \\ \hline
-  \textbf{Grieve et al. (2014)}
- \cite{afte-chumbley} & Randomized & Slip-joint  &  & \\ & Chumbley Score &  & - & - \\ \hline 
-  \textbf{Hadler \& Morris (2017)} \cite{hadler} & Deterministic & Screwdrivers & & \\ (Same-Surface Same-Angle) & Chumbley Score & & 0\% & 6\%  \\ \hline
-  \textbf{Bachrach et al.
-(2010)} \cite{manytoolmarks2} & Similarity Measure & Screwdrivers & & \\ (Different Surfaces-same angle) & Relative Distance Metric & & 5.9\% &  9.4\% \\ (Same Surfaces-same angle)&  & & 0.22\% &  0\% \\
-   \hline
-\end{tabular}}
-\caption{\label{tab:toolmarks-ER} Error Rates in same-source {\textcolor{blue}{striated}} toolmark analysis reported in the literature.
-The top four reported papers use some variation of the Chumbley-score method.}
-\end{table}
-
-In this paper, we are investigating the applicability of the Chumbley
-scoring method by \citet{hadler} to assess striation marks on bullet
-lands for same-source identification. Striation marks on bullets are
-made by impurities in the barrel. As the bullet travels through the
-barrel, these imperfections leave ``scratches'' on the bullet surface
-(see top of \autoref{fig:rgl}). Typically, only striation marks in the
-land engraved areas (LEAs) are considered \citep{afte-article1992}.
-Bullet lands are depressed areas between the grooves made by the rifling
-action of the barrel. Compared to toolmarks made by screwdrivers
-striation marks on bullets are typically much smaller, both in length
-and in width. Bullets also have a curved cross-sectional topography.
-
-In same-source comparisons this curvature is usually removed using some
-form of Gaussian filter \citep{ma2004} or non-parametric smoothing
-\citep{aoas}. An overview of some of the error rates reported in the
-literature on bullet matching is given in \autoref{tab:bullets-ER}.
-\citet{chu2013} use an automatic method for counting consecutive
-matching striae (CMS). The authors report an error rate of 52\% for
-known same-source land comparisons to be (incorrectly) identified as
-different-source (false negative) and zero false positives for known
-different-source lands. \citet{ma2004} and \citet{vorburger2011} discuss
-CCF (cross-correlation function) and its discriminating power and
-applicability for same-source analyses of bullets, but do not provide
-any error rates in their discussion. \citet{aoas} use multiple features,
-such as CCF, CMS, D (distance measure), etc. in a random forest based
-method and compare every land against every other land of digitized
-versions of Hamby 252 and Hamby 44 \citep{hamby} published on the NIST
-Ballistics Database \citep{nist}. The authors report an out-of-bag
-overall error rate of 0.46\%, comprised of an error rate of 30.05\% of
-same-source pairs that were not identified and an error rate of 0.026\%
-of different-source pairs that were incorrectly identified as
-same-source.
-
-\begin{table}[ht]
-\centering
-\resizebox{\textwidth}{!}{\begin{tabular}{rlrrr}
-  \hline
-Method & Data Source & False Positives & False Negatives\\
-  \hline
-  \textbf{\citet{aoas}} & Hamby 252 (NIST scans) \\
-Consecutive Matching Bullets Striae (CMS) &   &  6.25\% &  33.85\% \\ 
-Consecutive Non-matching Striae (CNMS)  &  &  6.26\% & 35.42\%  \\ 
-Average Distance (D) &   &   6.25\% & 45.83\% \\ 
-Cross-correlation Function (CCF) & &  6.25\% & 17.71\%  \\ 
-Sum of Peaks (S) &  &  6.25\% &  18.23\% \\
-   \hline
-   \textbf{\citet{chu2013}} & Hamby 252 (NIST scans) \\
-Consecutive Matching Bullets Striae (CMS) &  &  0\% &  52\% \\ 
-   \hline
-   \textbf{\citet{chu2010}} & \small{6 types of firearms, 2 types of ammunition} \\
-   \textbf{\citet{ma2004}} & \small{NIST standard bullet SRM 8240} \\
-Cross-Correlation Function (CCF) &   &  - &  - \\ 
-    \hline
-\end{tabular}}
-\caption{\label{tab:bullets-ER} Error Rates in same-source single feature land-to-land analysis reported in the literature.
-Note that only error rates from the same data source can be compared directly between methods.
-Error rates from different sources are influenced by type of firearm and ammunition used.}
-\end{table}
-
-The Chumbley score provides us with another approach in the same-source
-assessment of bullet striation marks. \citet{chumbley} compare two
-toolmarks for same-source. The data for this study was obtained from 50
-sequentially manufactured screwdriver tips. \citet{chumbley} report
-error rates for markings made by the tips at different angles. For
-markings made at 30 degree the authors report an average false negative
-error rate of 0.089 and an average false positive error rate of 0.023.
-For marks made under angles of 60 and 85 degrees, respectively, the
-false negatives error rate is 0.09 while the rate of false positives
-decreases to 0.01. The paper by \citet{hadler} is based on the same data
-but the authors focus on markings made under the same angle. The error
-rates associated with the deterministic version of the score are
-reported as 0.06 for false negatives and 0 for false positives.
-
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{images/B6-B2-L6-rescaled.pdf}
-
-\caption{\label{fig:rgl} Image of a bullet land from a confocal light microscope at 20 fold magnification (top) and a chart of the corresponding signature of the same land (bottom).
-The dotted lines connect some peaks visible in both visualizations.}
-
-\end{figure}
-
-In this paper we evaluate the adaptability of the Chumbley score as a
-measure to quantify similarity in land engraved areas (LEAs) on bullets.
-For that we briefly introduce the deterministic method suggested by
-\citet{hadler} in section 2. In the process we provide methods to
-identify parameters that minimize the error rates. We then investigate
-persistent scenarios in which the method proposed by \citet{hadler}
-fails to come to a result. We go on to provide a solution to the failed
-tests problem, consequently increasing the power of the test and
-reducing error rates in the process. We set up a testing framework to
-compare the performance of the two algorithms in section 3 and finally
-discuss results in section 4.
-
-\hypertarget{methods}{%
-\section*{Methods}\label{methods}}
-\addcontentsline{toc}{section}{Methods}
-
-\hypertarget{scans-for-land-engraved-areas}{%
-\subsection*{\texorpdfstring{\emph{Scans for land engraved
-areas}}{Scans for land engraved areas}}\label{scans-for-land-engraved-areas}}
-\addcontentsline{toc}{subsection}{\emph{Scans for land engraved areas}}
-
-Comparisons of striae from bullets are usually based on comparisons of
-striae in land engraved areas, which are extracted in form of cross
-sections, called \emph{profiles} \citep{aoas,ma2004}. Bullet striae are
-most pronounced at the bottom of the bullet heel (because the heel has
-the most contact with the insides of the barrel). Howver, these areas
-are also affected by the break off due to friction effects between
-barrel and the bullet. An optimal cross section is chosen orthogonally
-to the striae, close to the heel while avoiding break off as shown in
-\autoref{fig:rgl}, see also \citep{aoas} for mathematical details of the
-extraction.
-
-From profiles bullet \emph{signatures} \citep{chu2013,aoas} are
-extracted as residuals of a loess fit or Gaussian filter. This
-effectively removes topographic structure from the data in the attempt
-to increase the signal to noise ratio. \autoref{fig:rgl} shows how the
-signature from a bullet land (bottom) lines up with the image of the
-land (top) from which it was extracted. We can see in the figure how the
-depth and relative position of the striation markings seen in the image
-are interpreted as peaks and valleys in the signature.
-
-There are two sources of scans for sets from the Hamby study available
-to us: scans of Hamby 44 and Hamby 252 are available from the NIST
-database \citep{nist}. The physical Hamby 44 set has also been made
-available to us and has been scanned locally for CSAFE at the Roy
-J.~Carver High Resolution Microscopy Facility using a Sensofar confocal
-light microscope. Scans in the NIST database are made with a NanoFocus
-at 20x magnification. The resolutions of the two instruments are
-different: the NIST scans are taken at a resolution of 1.5625 \(\mu m\)
-per pixel, while the CSAFE scans are available at a resolution of 0.645
-\(\mu m\) per pixel. The length of an average bullet land from Hamby (9
-mm Ruger P85) is about 2 millimeter, resulting in signatures of about
-1200 pixels for NIST scans, and about 3000 pixels for CSAFE scans.
-
-In comparison, scans from the profilometer used by
-\citet{chumbley, hadler} were taken at a resolution of about 0.73
-\(\mu m\) per pixel. The screw driver toolmarks are about 7 mm in length
-\citep{manytoolmarks1}, for a total of over 9000 pixels for the width of
-these scans.
-
-This severe limitation in the amount of available data poses the main
-challenge in adapting the Chumbley score to matching bullet lands,
-because of the resulting loss in power.
-
-\hypertarget{the-chumbley-score-test}{%
-\subsection*{\texorpdfstring{\emph{The Chumbley Score
-Test}}{The Chumbley Score Test}}\label{the-chumbley-score-test}}
-\addcontentsline{toc}{subsection}{\emph{The Chumbley Score Test}}
-
-A digitized toolmark forms a spatial process \(z(t)\) with location
-indexed by \(t\). \(t\) here denotes equally spaced pixel locations for
-the striation marks under consideration. For a toolmark consisting of
-\(t\) pixels, \(t = 1, ..., T\). Let further \(z^s(t)\) denote a vector
-of markings of length \(s\) starting in location \(t\).
-
-The Chumbley score algorithm takes input in form of two digitized
-toolmarks:
-
-Let \(x(t_1)\), \(t_1 = 1,2,...T_1\) and \(y(t_2)\), \(t_2 = 1,2...T_2\)
-be two digitized toolmarks (where \(T_1\) and \(T_2\), the lengths of
-the two marks, are not necessarily equal). The toolmarks under
-consideration are potentially from two different-sources or the
-same-source.
-
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figures/sigs-profiles-1}
-
-}
-
-\caption{Bullet land profile (left) and the corresponding signature (right) for one of the lands of Hamby-44.}\label{fig:sigs-profiles}
-\end{figure}
-
-In a pre-processing step the two markings are smoothed using a LOWESS
-\citep{lowess} with coarseness parameter \(c\). Originally, this
-smoothing is intended to remove drift and (sub)class characteristics
-from individual markings, however, in the setting of matching bullet
-striae, we can also make use of this mechanism to separate bullet
-curvature in profiles from signatures before matching signatures.
-\autoref{fig:sigs-profiles} shows an example of a bullet land profile
-(left) and the corresponding signature (right).
-
-After removing sub-class structure, the Chumbley score is calculated in
-two steps: an optimization step and a validation step. In the
-optimization step, the two markings are aligned horizontally such that
-within a pre-defined window of length \(w_o\) the correlation between
-\(x(t_1)\) and \(y(t_2)\) is maximized: \[
-\left(t_1^o, t_2^o\right) = \mathop{\arg \max}\limits_{1 \le t_1 \le T_1-w_o, 1 \le t_2 \le T_2-w_o} \text{cor} \left(x^{w_o} (t_1), y^{w_o}(t_2) \right)
-\] This results in an optimal vertical (in-phase) shift of
-\(t_1^o - t_2^o\) for aligning the two markings. We will denote the
-relative optimal locations as \(t_1^*\) and \(t_2^*\), where
-\(t_k^* = t_k^o/(T_k-w_o)\) for \(k=1,2\), such that
-\(t_1^*, t_2^* \in [0,1]\). Once (sub-)class characteristics are
-removed, the relative optimal locations should be distributed according
-to a uniform distribution in \([0,1]\).
-
-In the validation step, two sets of windows of size \(w_v\) are chosen
-from both markings, see \autoref{fig:win-comparison}. In the first set,
-pairs of windows are extracted from the two markings using the optimal
-vertical shift as determined in the first step, whereas for the second
-set the windows are extracted using a different (out-of-phase) shift.
-
-More precisely, let us define starting points \(s_i^{(k)}\) for each
-signature \(k = 1, 2\) as \begin{eqnarray}\label{eq.start}
-s^{(k)}_i = 
-\begin{cases}
-t_k^* + i w_v & \text{ for } i < 0 \\
-t_k^* + w_ o + i w_v & \text{ for } i \ge 0,
-\end{cases}
-\end{eqnarray} for integer values of \(i\) with
-\(0 < s^{(k)}_i \le T_k - w_v\).
-
-Same-shift pairs of length \(w_v\) are defined in \citet{hadler} as all
-pairs \((s_i^{(1)}, s_i^{(2)})\) with integer values \(i\) for which
-both \(s_i^{(1)}\) and \(s_i^{(2)}\) are defined. Similarly,
-different-shift pairs are defined as \((s_i^{(1)}, s_{-i-1}^{(2)})\) for
-all \(i\) where both \(s_i^{(1)}\) and \(s_{-i-1}^{(2)}\) are defined
-(see \autoref{sketch-same-diff}).
-
-\begin{figure}[hbtp]
-\centering
-\includegraphics[width=.7\textwidth]{images/sketch-same.png}
-
-\includegraphics[width=.7\textwidth]{images/sketch-diff.png}
-\caption{\label{sketch-same-diff}Sketch of same-shift pairings  (top) and different-shift pairings (bottom).
-Filled in rectangles show pairings resulting in correlations, unfilled rectangles are segments without a match.}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figures/win-comparison-1}
-
-}
-
-\caption{Two same-source markings.
-For convenience, the markings are moved into phase on the left and out-of phase on the right.
-In-phase (left) and out-of-phase (right) samples are shown by the light grey background.
-The Chumbley-score is based on a Mann-Whitney U test of the correlations derived from these two sets of samples.}\label{fig:win-comparison}
-\end{figure}
-
-For both same- and different-shift pairs correlations between the
-markings are calculated. The intuition here is that for two markings
-from the same-source the correlation for the in-phase sample should be
-high, while the correlations of the out-of-phase sample provide a
-measure for the base-level correlation for non-matching marks of a given
-length \(w_v\). More specifically, the null hypothesis of the Chumbley
-score test is stated as \textbf{\(H_o\): the markings come from two
-different sources} with the alternative given as \textbf{\(H_a\): the
-markings come from the same source}. A \(p\)-value for the Chumbley
-score test is then computed as a Mann Whitney U statistic to compare
-between in-phase sample and out-of-phase sample. A low \(p\)-value is
-interpreted as a rejection of the null hypothesis in favor of the
-alternative hypothesis of \texttt{same-source}.
-
-In the original method proposed in \citet{chumbley} both in-phase and
-out-of-phase sample are extracted randomly, whereas \citet{hadler}
-proposed the above specified deterministic rules for both samples to
-make the resulting score deterministic while simultaneously avoiding
-overlaps within selected marks to ensure independence.
-
-\hypertarget{a-problem-with-failed-tests}{%
-\subsection*{\texorpdfstring{\emph{A problem with failed
-tests}}{A problem with failed tests}}\label{a-problem-with-failed-tests}}
-\addcontentsline{toc}{subsection}{\emph{A problem with failed tests}}
-
-Looking closer at \autoref{eq.start}, we see that by definition, some
-number of tests will fail to produce a result.
-{\textcolor{orange}{Note that this problem is different from erroneous test results.}}
-The problem of failed tests is first mentioned in \citet{afte-chumbley}.
-Unfortunately, the authors do not provide any percentage of how many
-tests failed for their data. The algorithm fails to produce for two
-reasons: either the number of eligible same-shift pairs is zero, or the
-number of different-shift pairs is zero. Section
-\ref{appendix:appxfailed} in the Appendix discusses scenarios of failed
-tests in more detail.
-
-The number of same-shift pairs will be zero, if the optimal locations
-\(t_1^{o}\) and \(t_2^{o}\) are so far apart, that no segments of size
-\(w_v\) are left on the same sides of the optimal locations, i.e.
-\(t_1^{o} < w_v\) and \(t_2^{o} > T_2-w_o-w_v\) or
-\(t_1^{o} < T_1- w_o - w_v\) and \(t_2^{o} < w_v\), i.e.~we have a
-failure rate of \[
-P\left( t_1^{o} < w_v \ \cap \ t_2^{o} > T_2-w_o-w_v\right) + P\left( t_1^{o} < T_1- w_o - w_v \ \cap \ t_2^{o} < w_v\right).
-\] While we can assume that once (sub-)class characteristics are
-removed, optimal locations \(t_i^{o}\) are uniformly distributed across
-the length of the profile, we cannot assume that \(t_1^o\) and \(t_2^o\)
-are independent of each other. In particular, for same-source profiles,
-we would expect a strong dependency between these locations, in which
-case a large difference between locations is unlikely. However, for
-different-source matches, we can assume that locations are independent.
-In that case, we expect a test to fail with a probability of
-\(\frac{2 w_v^2}{(T_1-w_o)(T_2-w_o)}\). For an average length of \(T_i\)
-of 1200 pixels, \(w_o = 120\) pixels and \(w_v = 30\) pixels this
-probability is about 0.0015.
-
-The number of possible different-shift pairs also depends on the
-location of the optimal locations \(t_1^o\) and \(t_2^o\). Whenever the
-optimal locations are close to the boundaries, the number of possible
-pairings decreases and reaches zero, if \(t_i^{(o)} < w_v\) or
-\(t_i^{(o)} > T_i-w_o- w_v\). Assuming a correlation between optimal
-locations \(t_1^o\) and \(t_2^o\) of close to one for same-source
-profiles, this results in an expected rate of failure of
-\(2 w_v / (T_i-w_o)\), or about 5.6\% for an average length of \(T_i\)
-of 1200 pixels, \(w_o = 120\) pixels and \(w_v = 30\) pixels. Assuming
-independence in the optimal locations for different-source profiles the
-expected probability for a failed test is, again,
-\(\frac{2 w_v^2}{(T_1-w_o)(T_2-w_o)}\).
-
-\hypertarget{a-modified-approach}{%
-\subsection*{\texorpdfstring{\emph{A modified
-approach}}{A modified approach}}\label{a-modified-approach}}
-\addcontentsline{toc}{subsection}{\emph{A modified approach}}
-
-While failures due to missing correlations from same-shift pairs are
-unavoidable by definition of the Chumbley score, failures due to missing
-correlations from different-shift pairs can be prevented by using a
-different strategy in assigning pairs.
-
-Using the same notation as in \autoref{eq.start}, we define same-shift
-pairs identical to \citet{hadler} as pairs \((s_i^{(1)}, s_i^{(2)})\)
-for all \(i\) where the boundary conditions of both sequences are met
-simultaneously. Let us assume that this results in \(I\) pairs. Define
-\(s_{(j)}^{(k)}\) to be the \(j\)th starting location in sequence
-\(k = 1, 2\),
-i.e.~\(s_{(1)}^{(k)} < s_{(2)}^{(k)} < ... < s_{(I)}^{(k)}\).
-
-We then define the pairs for different-shifts as
-\begin{equation}\label{eq.diff2}
-\left(s_{(j)}^{(1)}, s_{(I-j+1)}^{(1)} \right) \text{ for } j = 
-\begin{cases}
-1, ..., I & \text{ for even } I \\
-1, ..., (I-1)/2, (I-1)/2 + 2, ..., I & \text{ for odd } I
-\end{cases},
-\end{equation} i.e.~for an odd number of same-shift correlations, we
-skip the middle pair for the different-shift correlations (see
-\autoref{sketch-diff-2}). This pairing ensures that the number of
-different-shift pairings is the same or at most one less than the number
-of same-shift pairings in all tests. In the remainder of the paper, we
-will refer to the algorithm defined by \citet{hadler} as \textbf{(CS1)}
-and the suggested modified algorithm as \textbf{(CS2)} and compare their
-performance on the available scans of the Hamby study.
-
-Performance of tests is measured with respect to the errors a test makes
-in situations where ground truth is known. We distinguish between two
-error rates: (1) false positive and (2) false negative rate. \emph{False
-positives} (or \emph{false identifications}) are situations where the
-test indicates a match (i.e.~the test falsely rejects) but the markings
-come from different sources. This is also known as a \emph{Type I}
-error. \emph{False negatives} (or \emph{missed identifications}) are
-situations where the test fails to reject, i.e.~the test indicates that
-the markings come from different sources, but in fact the markings come
-from the same source. This is a \emph{Type II} error. In both cases
-error rates are calculated as the ratio of the number of errors observed
-and the number of tests executed.
-
-Note that in all of the following land-to-land comparisons only lands
-are compared that are suitable for a comparison, i.e.~a signature can be
-extracted from the scan. In particular, lands which exhibited tank rash
-were removed from comparison \citep{aoas}.
-
-\begin{figure}[hbtp]
-\centering
-\includegraphics[width=.7\textwidth]{images/sketch-diff-2.png}
-\caption{\label{sketch-diff-2}Sketch of adjusted different-shift pairings. At most one of the same-shift pairings can not be matched with a different-shift pair under algorithm (CS2). }
-\end{figure}
-
-\hypertarget{testing-setup}{%
-\subsection*{\texorpdfstring{\emph{Testing
-setup}}{Testing setup}}\label{testing-setup}}
-\addcontentsline{toc}{subsection}{\emph{Testing setup}}
-
-\hypertarget{the-data}{%
-\subsubsection*{The Data}\label{the-data}}
-\addcontentsline{toc}{subsubsection}{The Data}
-
-Lands for all Hamby-44 and Hamby-252 scans are made available through
-the NIST ballistics database \citep{nist} and are considered, here. Both
-of these sets of scans are part of the larger Hamby study \citep{hamby}.
-Each set consists of twenty known bullets (two each from ten
-consecutively rifled Ruger P85 barrels) and fifteen questioned bullets
-(each matching one of the ten barrels). Ground truth for both of these
-Hamby sets is known and was used to assess correctness of the tests
-results.
-
-Profiles for each bullet land were extracted from scans close to the
-heel of the bullet while avoiding break-off as described in
-\citet{aoas}.
-
-\hypertarget{setup}{%
-\subsubsection*{Setup}\label{setup}}
-\addcontentsline{toc}{subsubsection}{Setup}
-
-Both algorithms (CS1) and (CS2) are implemented in R \citep{R}. (CS1) is
-available from package \texttt{toolmaRk} \citep{toolmark}, (CS2) is
-available from a modified version of the \texttt{toolmaRk} package
-available from GitHub (\url{https://github.com/heike/toolmaRk}). We
-applied both methods to all pairwise land-to-land comparisons of the
-Hamby scans provided by NIST for a total of 85,491 land-to-land
-comparisons.
-
-\hypertarget{results}{%
-\section*{Results}\label{results}}
-\addcontentsline{toc}{section}{Results}
-
-\hypertarget{failed-tests}{%
-\subsection*{\texorpdfstring{\emph{Failed
-Tests}}{Failed Tests}}\label{failed-tests}}
-\addcontentsline{toc}{subsection}{\emph{Failed Tests}}
-
-As described above, the Chumbley-score is based on three parameters:
-coarseness \(c\) and the sizes of the optimization window \(w_o\) and
-validation window \(w_v\). In a first run of results, we applied default
-settings for the parameters, as suggested in \citet{hadler}:
-\(w_o = 120\) pixels or about 190 \(\mu m\) (ten percent of the average
-length of profiles) and coarseness \(c = 0.25\), and varied the size of
-the validation window \(w_v\) in steps of 10 from 10 pixels to 60
-pixels. Based on a significance level \(\alpha\) of 0.05 for the test,
-this results in a correct identification of same-source and
-different-source toolmarks of 93.5\% to 94.1\%, corresponding to a rate
-of false negatives between 0.28 and 0.36 and a rate of false positives
-between 0.05 and 0.06. However, the most prominent result we
-encountered, are the high number of failed tests, i.e.~the number of
-instances, in which CS1 did not return any result. \autoref{fig:fails}
-shows the percentage of failed tests among the 85,491 land-to-land
-comparisons of the NIST data for different values of the validation
-window size \(w_v\). For same-source lands up to 12.5 percent of the
-tests fail using CS1. The highest percentage of failed tests under CS2
-is 1.3\% for different-source tests using a validation window size
-\(w_v\) of 60 pixels. Rates of expected failures are based on simulation
-runs using covariances between locations of same-source profiles of
-0.854, and 0.120 for locations from different-source profiles, matching
-observed covariances for the Hamby scans. Observed failure rates are
-higher than expected. This might be due to remaining sub-class structure
-at a coarseness of 0.25 resulting in a distribution of optimal locations
-different from the assumed uniform.
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.7\textwidth]{figures/fails-1}
-
-}
-
-\caption{Percent of failed land-to-land comparisons using an optimization window $w_o = 120$ and a coarseness of $c = 0.25$.
-With an increase in the size of the validation window  a higher percentage of tests fails under both methods (CS1) and (CS2), but the percentage of failed tests is much smaller under (CS2).
-Observed failure rates of (CS1) are higher than expected rates.}\label{fig:fails}
-\end{figure}
-
-\hypertarget{coarseness}{%
-\subsection*{\texorpdfstring{\emph{Coarseness}}{Coarseness}}\label{coarseness}}
-\addcontentsline{toc}{subsection}{\emph{Coarseness}}
-
-The purpose of the coarseness parameter is to remove (sub-)class
-characteristics from profiles before comparisons for matching.
-\citet{hadler} suggest a coarseness parameter of 0.25 in the setting of
-toolmark comparisons. For bullet lands, coarseness might need to be
-adjusted because of the strong effect bullet curvature has on profiles.
-
-\autoref{fig:profile-sketch} gives an overview of the effect of
-different coarseness parameters: from left to right, coarseness levels
-\(c\) are varied in steps of 0.05 from 0.1 to 0.3. The top row shows
-resulting signatures after smoothing the profile shown in
-\autoref{fig:sigs-profiles} with different levels of coarseness. The
-histograms in the bottom row show the relative optimal location \(t^*\).
-Optimal locations are distributed uniformly once (sub-)class
-characteristics are removed. However, for coarseness values of
-\(c > 0.20\) we see quite distinct boundary effects: optimal locations
-\(t^*\) are found at the very extreme ends of a profile more often than
-one would expect based on a uniform distribution.
-
-The key effect of the optimal locations and thereby the coarseness is
-seen in the number of failed tests. Irrespective of whether CS1 or CS2
-is being used, if the relative optimal locations are at the boundaries
-we will see an increase in the number of failed tests. A balance is
-therefore needed in the selection of the coarseness parameter which
-reduces the boundary effect but does not remove important individual
-characteristics. Based on \autoref{fig:profile-sketch} a coarseness
-value of \(c = 0.15\) seems to be best suited to strike this balance for
-this example. For the remainder of the analysis, we will use this value
-for \(c\).
-
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figures/profile-sketch-1}
-
-}
-
-\caption{Overview of the effect of different coarseness parameters $c$ on the profile shown in Figure \ref{fig:sigs-profiles} (top).
-The bottom row shows histograms of  the (relative) optimal locations $t^o$ identified in the optimization step for different values of the coarseness parameter $c$.
-}\label{fig:profile-sketch}
-\end{figure}
-
-\hypertarget{error-rate-assessment}{%
-\subsection*{\texorpdfstring{\emph{Error rate
-assessment}}{Error rate assessment}}\label{error-rate-assessment}}
-\addcontentsline{toc}{subsection}{\emph{Error rate assessment}}
-
-\autoref{fig:roc} gives an overview of ROC (Receiver operating
-characteristic) curves for methods CS1 and CS2 over a range of different
-optimization window sizes \(w_o\) and two sizes for the validation
-window \(w_v\) (shape). The different color hues represent the two
-methods CS1 (red) and CS2 (blue). The ROC curves show the superior
-performance of CS2 over CS1. Generally, an optimization window \(w_o\)
-of 150 pixels or more leads to the best performance with respect to ROC
-curves. Results based on a validation window of size \(w_v = 30\) are
-generally better than results for \(w_v = 50\).
-
-\autoref{fig:eer} shows a comparison of the performance of the two
-methods CS1 and CS2 with respect to EER (equal error rate) and AUC (area
-under the curve) corresponding to the ROC curves shown in
-\autoref{fig:roc}. Equal error rates are reduced using method CS2, while
-area under the curve significantly increases (at a significance level
-\(\alpha\) of 0.05) compared to method CS1.
-
-The results from Figures \ref{fig:roc} and \ref{fig:eer} are summarized
-in numbers in \autoref{tab:aucs}. Equal error rates (EER), rates for
-false positives (FPR) and false negatives (FNR) are shown side by side
-with the area under the curve (AUR) for both methods for a set of
-different optimization windows \(w_o\) and a validation window \(w_v\)
-of 30 pixels. The rate of false positive same-source identifications is
-equal to the statistical type I error rate, which is set to
-\(\alpha = 0.05\) for this example. The rate of false negatives are
-missed same-source markings. This rate is also known as the type II
-error rate. A detailed plot on the type II error rates for CS1 and CS2
-can be found in the section \ref{appendix:appxtype2} of the Appendix.
-Area under the curve (AUC) is shown with confidence intervals as given
-by \citet{delong}. CS2 significantly outperforms CS1 with respect to its
-predictive power in most situations.
-
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figures/roc-1}
-
-}
-
-\caption{ROC curves of methods CS1 and CS2 for different sizes of optimization window $w_o$.
-Best performances with respect to  ROC curves are reached for optimization windows of sizes 150 and higher.
-Points of equal error rates (EERs) can be found at the intersection of the dotted line and the ROC curves.}\label{fig:roc}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figures/eer-1}
-
-}
-
-\caption{Comparison of results for CS1 and CS2 using area under the curve (AUC) and equal error rates (EER). }\label{fig:eer}
-\end{figure}
-
-\begin{table}[hbtp]
-\centering
-\begin{adjustbox}{max width=\textwidth}
-\begin{tabular}{rrrlrrl}
-  \hline
-  & \multicolumn{3}{l}{\bf CS1} & \multicolumn{3}{l}{\bf CS2} \\
-$w_o$ & \bf FPR & \bf FNR & \bf AUC (95\% C.I.) & \bf FPR & \bf FNR & \bf AUC (95\% C.I.) \\ 
-  \hline
-90 & 0.068 & 0.330 & 0.850 (0.835, 0.865) & 0.075 & 0.242 & 0.877 (0.863, 0.890) \\ 
-  120 & 0.065 & 0.309 & 0.864 (0.850, 0.878) & 0.072 & 0.236 & 0.890 (0.877, 0.903) \\ 
-  130 & 0.065 & 0.317 & 0.863 (0.850, 0.877) & 0.072 & 0.224 & 0.898 (0.886, 0.910) \\ 
-  140 & 0.065 & 0.304 & 0.873 (0.859, 0.886) & 0.071 & 0.229 & 0.902 (0.891, 0.914) \\ 
-  150 & 0.067 & 0.326 & 0.863 (0.850, 0.877) & 0.072 & 0.225 & 0.904 (0.892, 0.916) \\ 
-  180 & 0.061 & 0.326 & 0.877 (0.864, 0.890) & 0.070 & 0.229 & 0.907 (0.896, 0.919) \\ 
-  210 & 0.064 & 0.342 & 0.865 (0.851, 0.878) & 0.066 & 0.231 & 0.906 (0.895, 0.918) \\ 
-   \hline
-\end{tabular}
-\end{adjustbox}
-\caption{\label{tab:aucs} Overview of results as shown in Figures \ref{fig:roc} and \ref{fig:eer}.
-FPR is the observed rate of false positives (for a fixed level $\alpha = 0.05$), FNR is the false negative rate.
-Area under the curve (AUC) is shown with confidence intervals.
-}
-\end{table}
-
-\hypertarget{observed-versus-nominal-type-i-error-rates}{%
-\subsection*{\texorpdfstring{\emph{Observed versus Nominal Type I error
-rates}}{Observed versus Nominal Type I error rates}}\label{observed-versus-nominal-type-i-error-rates}}
-\addcontentsline{toc}{subsection}{\emph{Observed versus Nominal Type I
-error rates}}
-
-\autoref{fig:type1} shows observed type I error rates (FPR) across a
-range of optimization windows \(w_o\). Generally, observed type I errors
-are higher than expected. Method CS1 shows in this instance slightly
-better performance than method CS2, but for both an increase in the size
-of the optimization window leads to a decrease in the observed type I
-error rates.
-
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figures/type1-1}
-
-}
-
-\caption{Comparison of observed and nominal type I error rates  across a range of window sizes for optimization $wo$.
-The horizontal line in each facet indicates the nominal type I error rate.}\label{fig:type1}
-\end{figure}
-
-\hypertarget{high-resolution-hamby-44-scans}{%
-\subsection*{\texorpdfstring{\emph{High resolution Hamby 44
-scans}}{High resolution Hamby 44 scans}}\label{high-resolution-hamby-44-scans}}
-\addcontentsline{toc}{subsection}{\emph{High resolution Hamby 44 scans}}
-
-The high-resolution scans of Hamby set 44 are capturing images at a
-resolution of \(0.645 \mu m\) per pixel. On average, land engraved areas
-are 3000 pixels in length. A coarseness of \(c = 0.125\) seemed to be
-sufficient in removing any bullet curvature. Both methods have a failed
-test rate of less than 0.006, indicating, again, that the larger number
-of pixels alleviates the problem of test failures. \autoref{fig:h44}
-shows the resulting EER and AUC for methods CS1 and CS2 based on two
-sizes of validation windows (\(w_v \in 75, 125\)) and optimization
-window sizes around 300 pixels (10 percent of the average length). Both
-methods show an increase in performance around \(w_o = 300\) pixels. CS2
-out-performs CS1 in all scenarios, but the difference is not significant
-(using DeLong's confidence intervals). Interestingly, the overall
-performance of both CS1 and CS2 is a lot lower for the high resolution
-version of Hamby-44 than for the lower resolution scans. The area under
-the curve overall is significantly lower for the high resolution scans
-than for the previous set of scans. Partly, this might be due to the
-particular choice of the parameters, partly the higher-resolution scans
-might be picking up on real differences between the lands that the
-lower-resolution scans fail to detect.
-
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figures/h44-1}
-
-}
-
-\caption{AUC and EER for methods CS1 and CS2 based on high resolution scans of Hamby 44. }\label{fig:h44}
-\end{figure}
-
-\begin{table}[hbtp]
-\centering
-\begin{adjustbox}{max width=\textwidth}
-\begin{tabular}{rrrlrrl}
-  \hline
-  & \multicolumn{3}{l}{\bf CS1} & \multicolumn{3}{l}{\bf CS2} \\
-$w_o$ & \bf FPR & \bf FNR & \bf AUC (95\% C.I.) & \bf FPR & \bf FNR & \bf AUC (95\% C.I.) \\  
-  \hline
-%90 & 0.077 & 0.546 & 0.680 (0.633, 0.727) & 0.056 & 0.478 & 0.729 (0.684, 0.774) \\
-%  120 & 0.082 & 0.485 & 0.736 (0.693, 0.779) & 0.056 & 0.452 & 0.763 (0.720, 0.806) \\
-%  150 & 0.080 & 0.493 & 0.730 (0.685, 0.775) & 0.055 & 0.422 & 0.774 (0.733, 0.816) \\
-%  180 & 0.087 & 0.458 & 0.773 (0.734, 0.813) & 0.055 & 0.435 & 0.773 (0.730, 0.816) \\
-  210 & 0.086 & 0.413 & 0.780 (0.740, 0.821) & 0.055 & 0.387 & 0.803 (0.763, 0.843) \\ 
-  240 & 0.085 & 0.427 & 0.784 (0.744, 0.824) & 0.055 & 0.365 & 0.801 (0.760, 0.841) \\ 
-  270 & 0.082 & 0.415 & 0.782 (0.741, 0.823) & 0.048 & 0.374 & 0.808 (0.769, 0.848) \\ 
-  300 & 0.082 & 0.405 & 0.791 (0.751, 0.832) & 0.051 & 0.383 & 0.814 (0.776, 0.852) \\ 
-  330 & 0.082 & 0.428 & 0.791 (0.753, 0.830) & 0.050 & 0.383 & 0.804 (0.765, 0.844) \\ 
-  360 & 0.083 & 0.427 & 0.788 (0.747, 0.828) & 0.049 & 0.400 & 0.803 (0.765, 0.842) \\ 
-  390 & 0.085 & 0.400 & 0.789 (0.748, 0.830) & 0.050 & 0.383 & 0.810 (0.771, 0.849) \\ 
-   \hline
-\end{tabular}
-\end{adjustbox}
-\caption{\label{tab:aucsh44} Overview of high resolution scan results as shown in Figures \ref{fig:h44}.
-FPR is the observed rate of false positives (for a fixed level $\alpha = 0.05$), FNR is the false negative rate.
-Area under the curve (AUC) is shown with 95\% confidence intervals.
-}
-\end{table}
-
-\hypertarget{conclusions}{%
-\section*{Conclusions}\label{conclusions}}
-\addcontentsline{toc}{section}{Conclusions}
-
-In assessing the suitability of the (deterministic) Chumbley Score for
-matching striae on bullet lands we have gained valuable insights into
-the process: method CS1 as proposed by \citet{hadler} has a strong
-dependency on the specific choice of parameters; the defaults suggested
-by \citet{hadler} for screw drivers are not directly applicable for the
-smaller bullet lands. The coarseness parameter in particular has a
-strong impact on the performance of the test. However, we were able to
-suggest some heuristics based on the assumption that once sub-class
-characteristics are removed, optimal locations are distributed uniformly
-across the profile. For bullet lands we found a coarseness value of
-\(c = 0.15\) to be suitable for the low-resolution scans from NIST and a
-value of \(c=0.125\) suitable for the higher-resolution scans from
-CSAFE. Sizes for optimization windows \(w_o\) were based on
-cross-validation to minimize overall type 2 error rates. Unfortunately,
-this kind of assessment is only feasible in the setting of a large
-study, such as the one we presented, and does not transfer immediately
-to case work, where a forensic examiner would only deal with a few
-identifications. Parameter settings for different studies should be
-investigated and reported to allow for a further refinement.
-
-Method CS1 proposed by \citet{hadler} has a minimal type 2 error rate of
-0.272 for an optimized window size of 140 pixels -- which is
-considerably higher than the error rates achieved on matching toolmarks,
-but is similar to other single-feature methods proposed for bullet
-matching. Unfortunately, method CS1 also has a high rate of failed tests
--- situations, in which the algorithm does not provide a result, due to
-the way different-shift pairs are constructed. Algorithm CS2 is
-introduced here as a remedy for failed tests by introducing an alternate
-version of choosing different-shift pairs. The algorithm CS2 introduced
-here is constructed in a way that achieves on average a ten-fold
-reduction in the number of failures. While reducing the failure rate,
-the algorithm also shows an increase in the power of the test. Type II
-error rates of CS2 reach a minimum of 0.217 for an optimized window size
-of 130 pixels. This increase in power of CS2 over CS1 should also apply
-to previous studies on toolmarks. It would be interesting to see these
-results using the adjusted algorithm. Unfortunately, none of the studies
-have made the data publicly accessible.
-
-While significantly reduced over CS1, CS2 still has type 2 error rates
-on bullet lands that are higher than the error rates achieved on the
---much larger-- toolmarks. Applying these methods to the high-resolution
-scans provided by CSAFE shows that better scanning methodology does not
-guarantee a better matching performance.
-
-However, bullets usually have multiple lands -- in the case of Ruger
-P85s as used in the Hamby study, there are six lands for each bullet. We
-might be able to get more power out of the test by adapting CS2 to work
-in a bullet-to-bullet comparison.
-
-\hypertarget{appendix}{%
-\section*{Appendix}\label{appendix}}
-\addcontentsline{toc}{section}{Appendix}
-
-\begin{appendix}
-
-\subsection{Scenarios of failed tests}
-\label{appendix:appxfailed}
-Figures \ref{same-shift-failure} and \ref{diff-shift-failure} show scenarios in which the deterministic Chumbley score can fail. 
-\autoref{diff-shift-failure} shows two situations, in which CS1 fails to identify any different-shift pairings.
-This happens, when both relative locations are close to either end of the signature.
-CS2 can still be computed in this situation as long as there are at least two same-shift pairs identified.
-
-\begin{figure}[hbtp]
-\centering
-\includegraphics[width=0.7\textwidth]{images/diff-shift-failure.png}
-
-\caption{\label{diff-shift-failure}Sketch of different-shift pairings. For CS1 no different-shift pairs can be identified, resulting in a failed test.}
-\end{figure}
-
-In \autoref{same-shift-failure} both CS1 and CS2 fail, because the lag between optimal locations is so large that no same-shift pairs can be found once the signatures are aligned.
-These failures are inherent to the Chumbley Score and cannot be prevented.
-\begin{figure}[hbtp]
-\centering
-\includegraphics[width=0.62\textwidth]{images/same-shift-failure.png}
-
-\caption{\label{same-shift-failure}Sketch of same-shift pairings.
-When the lag between optimal reference points is too large to accommodate a validation window in both signatures, both CS1 and CS2 fail.}
-\end{figure}
-
-\newpage
-\subsection{Type 2 errors: CS1 vs CS2}
-\label{appendix:appxtype2}
-\autoref{fig:type2} gives an overview of type 2 error rates of methods CS1 and CS2 for different significance levels $\alpha$. Method CS2 is outperforming CS1 significantly in every instance.
-
-\begin{figure}[h]
-
-{\centering \includegraphics[width=0.9\textwidth]{figures/type2-1}
-
-}
-
-\caption{Type II error rates observed across a range of window sizes for optimization $w_o$.
-For a window size of $w_o = 130$ we see a minimum in type II error rate across all type I rates considered.
-Smaller validation sizes $w_v$ are typically associated with a smaller type II error.}\label{fig:type2}
-\end{figure}
-
-\end{appendix}
-
-\bibliographystyle{jfs-authoryear}
-\bibliography{bibliography}
-
 \end{document}

### 2018-09-19 doc/paper.Rmd
commented out sentences
update to percentages

@@
 ggerrors <- errorsCV %>%
-  ggplot(aes(x = wo, y = beta, colour=factor(alpha), linetype=factor(alpha))) +
+  ggplot(aes(x = wo, y = 100*beta, colour=factor(alpha*100), linetype=factor(100*alpha))) +
   geom_point(aes(shape=factor(wv)), size=2.5) +
@@
   theme_bw() +
-  scale_colour_brewer(expression("Nominal\ntype I error"~alpha),
+  scale_colour_brewer(expression("Nominal Percent \nFalse Positives "~alpha),
                       palette="Set2", guide=FALSE) +
@@
   xlab(expression("Window size for optimization "~w[o])) +
-  ylab("Type II error rate") +
+  ylab("Percent False Negatives") +
   facet_wrap(~method, labeller="label_both")
@@
   theme_bw() +
-  xlab("False Positives Rate (1 - Specificity)") +
-  ylab("True Positives Rate (Sensitivity)") +
+  xlab("Percent False Positives (100 - Specificity)") +
+  ylab("Percent True Positives (Sensitivity)") +
   guides(
@@
   geom_abline(slope=-1, intercept = 1, linetype = 3, 
-              colour="grey30", size = 0.25)
+              colour="grey30", size = 0.25) +
+  scale_x_continuous(labels=c("0", "25", "50", "75", "100")) +
+  scale_y_continuous(labels=c("0", "25", "50", "75", "100"))
 
@@
 p1 <- roc.summary %>% filter(!(wo %in% c(130, 140))) %>%
-  ggplot(aes( x = wo, y=eer, colour=method, shape=method)) +
+  ggplot(aes( x = wo, y=eer*100, colour=method, shape=method)) +
   geom_point(size=2) +
@@
   theme_bw() +
-  ylab("Equal Error rate (EER)") +
+  ylab("Equal Error (EER) in Percent") +
   xlab("Size of optimization window in pixels") +
@@
 errorsCV %>% filter(wv %in% c(30,50)) %>%
-  ggplot(aes(x = wo, y = actual, colour=factor(alpha), 
+  ggplot(aes(x = wo, y = 100*actual, colour=factor(100*alpha), 
              linetype=method)) + 
-  geom_hline(aes(yintercept=alpha), colour="grey30", size=.5) +
+  geom_hline(aes(yintercept=100*alpha), colour="grey30", size=.5) +
   geom_point(aes(shape=factor(wv.method)), size=2.5) +
@@
   theme_bw() +
-  ylab(expression("Observed type I error rate "~hat(alpha))) +
+  ylab(expression("Observed Percent False Positives "~hat(alpha))) +
   xlab(expression("Window size for optimization "~w[o])) +
   scale_colour_brewer(
-    expression("Nominal type I error "~alpha), palette="Set2") +
-  facet_wrap(~alpha, labeller=label_bquote(alpha == .(alpha)), scales="free") +
+    expression("Nominal Percent False Positives "~alpha), palette="Set2") +
+  facet_wrap(~alpha, labeller=label_bquote(alpha == .(alpha*100)), scales="free") +
   theme(legend.position="bottom") +
@@
 
-h44.summ %>% gather(type, value, AUC, EER) %>% 
+h44.summ$EER <- 100*h44.summ$EER
+h44.long <- h44.summ %>% gather(type, value, AUC, EER) %>% 
   filter(wv %in% c(75, 125), coarseness==0.125,
-         between(wo, 210, 390)) %>% 
+         between(wo, 210, 390)) 
+h44.long$type <- factor(h44.long$type)
+
+levels(h44.long$type)[2] <- "Percent Equal Error"
+
+h44.long %>% 
   ggplot(aes(x = wo, y = value, colour=Method, shape=factor(wv)))  +

### 2018-09-20 doc/paper.Rmd
update figures to percentages

@@
   scale_shape("Method") +
-  scale_y_continuous(limits=c(0, 0.25))
+  scale_y_continuous(limits=c(0, 25))
 
@@
     expression("Nominal Percent False Positives "~alpha), palette="Set2") +
-  facet_wrap(~alpha, labeller=label_bquote(alpha == .(alpha*100)), scales="free") +
+  facet_wrap(~alpha, labeller=label_bquote(alpha == .(alpha*100)~'%'), scales="free") +
   theme(legend.position="bottom") +
@@
 Smaller validation sizes $w_v$ are typically associated with a smaller type II error."}
-ggerrors
+ggerrors # write our new?
 ```

### 2018-09-24 jfs_submission/paper-pre-submission-revision-2/Responses-round2.Rmd
responses document

---
title: "Revision 2: Responses to Reviewers Round 2"
author: "Ganesh Krishnan, Heike Hofmann"
date: "September 24, 2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\newcommand{\hh}[1]{{\textcolor{orange}{#1}}}
\newcommand{\gk}[1]{{\textcolor{blue}{#1}}}

We would like to thank the reviewers for their in-depth reviews and the suggestions regarding the manuscript. Please find our detailed response addressing each of the points below.

## Editorial changes

1.  Headers are included in the abstract and they do not need to be.

2.  Several references include the DOI and year, volume page citations.  Currently we only need the latter if that is available.  Please correct.

3.  Reference 4 is a URL and when checked it indicated that it was no linger updated.  It also did not take an interested reader to the PCST report.  Please use a different reference

4.  Reference 9 is another URL and my check indicated that the page could not be found.  Please correct.

5.  Please follow the formats included in the Information for Authors and abbreviate the journal titles according to Index Medicus

6.  Reference 8 was incomplete.  Please complete per the Information for Authors.

## Reviewer #1

Comments to the Author

This revision is a major improvement from the initial submission, congratulations.

1. Pg.4, ln 14:  Change "impurities" to "surface imperfections"

2. Pg. 4, ln. 45; change to "(because the heel typically has the most contact with the rifling in the barrel)"

3. Pg.4, last sentence.  Please check for grammar or incomplete information.
Also capitalize "Lowess". (?)
Perhaps: "From these profiles, bullet signatures (16; 15) are extracted as residuals ..."

4. Pg.5, ln. 36: Change to "because of the resulting loss in discriminating power."

5. Pg.7 and on;  Please add some text to clarify your definition of "sub-class".  You are designating this term as a general form or structure that has to be removed to get at the signature.  However, the term is used quite differently by firearm examiners to describe features that may reproduce in a number of tool surfaces, and its influence if detected must be considered prior to a comparison conclusion.

6. Pg. 8; Thanks for pointing out what "failed test" is, and what it is not.

7. Pg. 10, ln. 32:  "tank rash" is a is a colloquialism that should be in quote marks since it would be unknown to many scientists outside of the forensic discipline.  Or longer: "random tool marks on the fired bullet surface caused by the impact with the interior surfaces of the bullet capture tank."

8. Pg. 10, ln. 7:  fix error "toolmaRkpack-age"
  ln. 48 space is needed between "profiles of0.854"

## Reviewer #2

Comments to the Author
Thank you for making the requested revisions. I appreciate that this takes time. A few issues remain but I believe we are getting close.

1.  p3 “These digitized markings allow the use of statistical methods to quantify the scientific mechanism of comparing markings and serve as basis for an error rate calculation.” Still a little awkward.

2.  The authors claim that they changed all numerical results to percentages but there are still a number of inconsistencies. The authors should use consistent terminology. Sometimes they say FPR, sometimes %FP, sometimes Type I, sometimes Type 1 error rate. It is ok to list all these as being the same term in the definition, but then they should pick one.  For example, in the conclusions CS1 is reported with a type 2 error rate of 0.272 which I assume is 27.2%. In the Results section the expected failures are listed as 0.854 and 0.120, should these be 85.5% and 12.0%? Figure 14 y-axes are not percentages and must be updated. Figure 8 both axes need to be updated. Figure 9 (left, EER) also likely needs to be updated.

3.  Original comment: p8 “This severe limitation in the amount of available data poses the main challenge in”. It’s unclear what ‘this’ refers to. Do the authors mean the number of bullets and LEAs in the dataset or do they mean the number of sample points per striation profile (due to the fact that LEAs are shorter than screwdriver profiles)? Both would seem to be limitations but just because the profile is shorter doesn’t inherently mean that there are fewer or less informative striaes. You could have a shorter profile with more informative features. The authors replied: “We were indeed referring to the shorter length of the marks, which given the same amount of information decreases the power of the test, a purely technical consequence. Obviously you are right that this power calculation changes with a change in the amount of information.

No change was apparently made to the manuscript. I feel an edit is required. The text is still ambiguous and the reader will not have the benefit of the author’s response letter. In addition I believe the response is incorrect. We do not have enough information to make this statement. As I mentioned, it’s possible to have a very informative short profile. Perhaps bullet lands are inherently more structured than longer screwdriver marks. For example, a small but focused photo a person is more informative than a larger blurry image.

4.  I originally wrote: The authors state “this kind of assessment is only feasible in the setting of a large study, such as the one we presented”. I strongly disagree with the use of the word “large”. The presented work is not a large study. Ten firearms were used in each of two studies (it’s not clear if both datasets used the same 10 firearms). A large study would include at least 100 different firearms and even 100 is not likely to provide statistical confidence. This is a very important point. If the paper is allowed to state that the described work is a large study it will set a precedent that will be detrimental to the field.” The authors answered: We agree the data available to us is by no means large. Our comment was referring to the availability of 35 bullets rater than just 2 0r 3 as might be seen in actual case work. We have changed the sentence accordingly.

The updated text does not appear to be changed here. The text still states: “Unfortunately, this kind of assessment is only feasible in the setting of a large study, such as the one we presented.” This needs to be changed prior to acceptance. This is not a large study. Please see my original comment.

### 2018-09-24 jfs_submission/paper-pre-submission-revision-2/Responses-round2.Rmd
inserted replies for Reviewer 1 and editor in the responses document.

@@
 
-1.  Headers are included in the abstract and they do not need to be.
+1.  Headers are included in the abstract and they do not need to be. *Done*
 
-2.  Several references include the DOI and year, volume page citations.  Currently we only need the latter if that is available.  Please correct.
+2.  Several references include the DOI and year, volume page citations.  Currently we only need the latter if that is available.  Please correct. *Done*
 
-3.  Reference 4 is a URL and when checked it indicated that it was no linger updated.  It also did not take an interested reader to the PCST report.  Please use a different reference
+3.  Reference 4 is a URL and when checked it indicated that it was no linger updated.  It also did not take an interested reader to the PCST report.  Please use a different reference. *Done*
 
-4.  Reference 9 is another URL and my check indicated that the page could not be found.  Please correct.
+4.  Reference 9 is another URL and my check indicated that the page could not be found.  Please correct. *Done*
 
-5.  Please follow the formats included in the Information for Authors and abbreviate the journal titles according to Index Medicus
+5.  Please follow the formats included in the Information for Authors and abbreviate the journal titles according to Index Medicus *Done*
 
@@
 
+*Thank you, we have incorporated this change*
+
 2. Pg. 4, ln. 45; change to "(because the heel typically has the most contact with the rifling in the barrel)"
+
+*Thank you, we have incorporated this change*
 
@@
 
+*Thank you, we have changed the grammar and incorporated this change*
+
 4. Pg.5, ln. 36: Change to "because of the resulting loss in discriminating power."
+
+*Thank you, we have incorporated this change*
 
@@
 
+*We are happy that we could clarify this*
+
 7.
@@
 
+*Thanks a lot for pointing this out, we have made the changes in the paper reflecting this*
+
 8. Pg. 10, ln. 7:  fix error "toolmaRkpack-age"
   ln. 48 space is needed between "profiles of0.854"
+
+  *Thank you, we have incorporated this change*
 

### 2018-10-03 jfs_submission/paper-pre-submission-revision-2/Responses-round2.Rmd
last couple of comments?

@@
 
-6.  Reference 8 was incomplete.  Please complete per the Information for Authors.
+6.  Reference 8 was incomplete.  Please complete per the Information for Authors. *Done*
 
@@
 
+*Thank you for this comment.
+You are right, the term sub-class characteristics is used by FTEs to describe markings that are shared across multiple tools.
+Some of these sub-class characteristics will be expressed in markings that can be removed by using the normalization methods suggested.
+However, normalization can del with more general structure, but at the same time will not address all sub-class structures.
+We have tried to clarify this point by introducing the following paragraph on page 6:
+
+The implementation of the Chumbley score in Hadler and Morris (1) uses a normalization step before going into the optimization and validation step described below.
+Normalization is done by using a s LOWESS smooth to reduce extraneous structure in the markings, such as a drift, or spatial trends introduced during the barreling.
+To a degree, this normalization can also be used to address problems stemming from sub-class characteristics, i.e.
+markings in a pattern that are not unique to a single barrel but shared across a group of barrels introduced by specifics in the manufacturing process.
+*
+
 6. Pg. 8; Thanks for pointing out what "failed test" is, and what it is not.
@@
 
+*We have changed the sentence to: This data forms the basis of a statistical analysis that allow us to quantify similarity of markings and serves as a basis for error rate calculations*
+
 2.
@@
 Figure 9 (left, EER) also likely needs to be updated.
+
+*We now have changed all reported numbers regarding errors to percentages, including all figures.*
 
@@
 
+*We have softened the sentence at the end of the paragraph.
+We also did not mean to make this sentence an absolute statement but were rather speculating ourselves.
+We have made the change to the manuscript to reflect that better.*
+
 4.
@@
 Please see my original comment.
+
+*We are very sorry - we are completely in agreement. The original sentence was left in because of an unfortunate technical hiccup on our side. It is removed now.*

### 2018-10-09 jfs_submission/paper-pre-submission-revision-2/Responses-round2.Rmd
after submission revision 2- note: the manuscript file has been rewritted, its a different file from the previous commit

@@
 
-We would like to thank the reviewers for their in-depth reviews and the suggestions regarding the manuscript. Please find our detailed response addressing each of the points below.
+We would again like to thank the reviewers for their in-depth reviews and the suggestions regarding the manuscript. Please find our detailed response addressing each of the points below.
 
@@
 
-*We have changed the sentence to: This data forms the basis of a statistical analysis that allow us to quantify similarity of markings and serves as a basis for error rate calculations*
+*Thank you for the comment.
+We have changed the sentence to: This data forms the basis of a statistical analysis that allow us to quantify similarity of markings and serves as a basis for error rate calculations*
 
@@
 
-*We now have changed all reported numbers regarding errors to percentages, including all figures.*
+*Thank you for poining this out. We now have changed all reported numbers regarding errors to percentages, including all figures.*
 
@@
 
-*We have softened the sentence at the end of the paragraph.
+*Thank you for the comment.
+We have softened the sentence at the end of the paragraph.
 We also did not mean to make this sentence an absolute statement but were rather speculating ourselves.
